### 2022-09-03 Writing/figures.Rmd
initial commit

<!-- Geographic focus -->

```{r results = "hide"}
#=== HPA boundary ===#
hpa <- st_read(here("Data/RawData/hp_bound2010.shp")) %>% 
  st_simplify()

us_states_borders <- states(cb = TRUE) %>%
  st_as_sf() %>%
  st_transform(st_crs(hpa)) %>% 
  rename(state_code = STATEFP, state_abb = STUSPS) %>%
  dplyr::select(state_code, state_abb)

```

```{r gfocus, fig.cap = "Geographic Focus of the Study by Crop Type", fig.height = 4, fig.width = 6.5}

all_counties <- readRDS(here("Data/ProcessedData/base_counties.rds"))

# results$states

gfocus_data <- results %>% 
  dplyr::select(crop_type, states) %>% 
  mutate(plot_data = list(
    all_counties[state %in% states, ] 
  )) %>% 
  dplyr::select(-states) %>% 
  unnest() %>% 
  dplyr::select(crop_type, geometry, state_name) %>% 
  st_as_sf()

gfocus_state_borders <- results %>% 
  dplyr::select(crop_type, states) %>% 
  mutate(state_data = list(
    filter(us_states_borders, state_abb %in% states)
  )) %>% 
  dplyr::select(-states) %>% 
  unnest() %>% 
  st_as_sf()

t_gfocus <- tm_shape(hpa) +
  tm_polygons(col = "blue", alpha = 0.3) +
tm_shape(gfocus_data) +
  tm_borders() +
tm_shape(gfocus_state_borders) +
  tm_borders(lwd = 1.5, col = "red") +
  tm_facets(by = "crop_type", nrow = 1)

# tmap_save(
#   t_gfocus,
#   here("Writing/Figures", "geogprahic_focus.pdf"),
#   height = 3.5,
#   width = 6
# )

t_gfocus

```

`r run_pagebreak()`

<!-- The impact of saturated thickness on yield (Intensive Margin) -->

```{r yield-intensive, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_intensive$data_for_plot[[1]] %>% 
  .[days_ab_35 == median(days_ab_35), ]

g_intensive$data_for_plot[[1]] %>% 
  .[days_ab_35 == median(days_ab_35), ]

g_intensive$data_for_plot[[1]]

str(g_intensive$data_for_plot[[1]])

g_intensive <- reg_results %>% 
  dplyr::select(crop_type, data_for_plot) %>% 
  mutate(data_for_plot_balance = list(
    data_for_plot[days_ab_35 == median(days_ab_35), ]
  )) 
  unnest(cols = "data_for_plot") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = balance
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = balance)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()

# ggsave(
#   here("Writing/Figures", "yield_intensive.pdf"),
#   g_intensive, 
#   height = 3.5,
#   width = 6
# )

g_intensive

```

`r run_pagebreak()`

<!-- The impact of saturated thickness on Irrigation Share -->

```{r ir-share, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_ir_share <- results_all %>% 
  dplyr::select(crop_type, ir_share_eval_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = ir_share_hat, 
      x = sat, 
      color = factor(balance_avg)
    )
  ) +
  facet_grid(. ~ crop_type) +
  theme_bw()

# ggsave(
#   here("Writing/Figures", "ir_share.pdf"),
#   g_ir_share, 
#   height = 3.5,
#   width = 6
# )

g_ir_share
```

`r run_pagebreak()`

<!-- The impact of saturated thickness on yield (Combined) -->

```{r yield-combined, fig.cap = "The Impact of Saturated Thickness on Yield", fig.height = 6, fig.width = 6.5}

g_sat_impact <- results_all %>% 
  dplyr::select(crop_type, sat_sim_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = mean_yield, 
      x = balance, 
      color = factor(sat)
    )
  ) +
  facet_grid(crop_type ~ ., scales = "free_y") +
  scale_color_viridis_d(name = "Saturated Thickness") +
  theme_bw() +
  theme(
    legend.position = "bottom"
  )

# ggsave(
#   here("Writing/Figures", "yield_combined.pdf"),
#   g_sat_impact, 
#   height = 3.5,
#   width = 6
# )

g_sat_impact
```

```{r }
data_sim_yield <- readRDS(here("Results/simulated_yield_data.rds"))

cl_data_current <- readRDS(here("Data/ProcessedData", "summrized_gridMET.rds"))

cl_data_future <- 
  expand.grid(
    model = model_ls,
    rcp = rcp_ls
  ) %>% 
  data.table() %>% 
  rowwise() %>% 
  #=== define maca data file names ===#
  mutate(file_name = list(
    paste0(
      "Data/ProcessedData/MACA/sum_", 
      model,
      "_",
      rcp,
      "_county.rds"
    )  
  )) %>% 
  mutate(maca_data = list(
    #=== read maca data ===#
    readRDS(file_name) %>% 
    #=== calculate 10-year balance avg ===#
    .[, 
      `:=`(
        balance_avg = mean(balance), 
        days_ab_35_avg = mean(days_ab_35)
      ), 
      by = .(floor((year - 2020) / 10), sc_code)
    ] 
  )) %>% 
  dplyr::select(model, rcp, maca_data) %>% 
  unnest() %>% 
  data.table()

cl_data <- 
  rbind(
    cl_data_current[, .(sc_code, year, balance, days_ab_35, model, rcp)],
    cl_data_future[, .(sc_code, year, balance, days_ab_35, model, rcp)]
  ) %>% 
  .[order(sc_code, year), ]

ggplot(data = cl_data_future) +
  geom_boxplot(aes(y = balance, x = factor(year))) +
  facet_grid(model ~ rcp)

felm(balance ~ year , data = cl_data_future) %>% summary()

```

### 2022-09-03 Writing/manuscript_DR.Rmd
initial commit

---
title: The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather
author:
  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota]
  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
abstract: |
  Abstract goes here. It should be a single paragraph of around 200 words.
acknowledgments: |
  We wish to thank...
bibliography: DRA.bib
csl: iopart.csl
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
---

```{r packages}
#| include: false
#| cache: false

#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(here)
library(tmap)
library(lfe)
library(modelsummary)
library(patchwork)
library(kableExtra)
```

```{r}
#| echo: false

options(knitr.duplicate.label = "allow")
```

```{r run-analys, cache = TRUE, include = FALSE}
knitr::purl(
  here::here("GitControlled/Codes/1_analysis.Rmd"),
  here::here("GitControlled/Codes/1_analysis.R"),
)
source(here::here("GitControlled/Codes/1_analysis.R"))

knitr::purl(
  here::here("GitControlled/Codes/2_figures_tables.Rmd"),
  here::here("GitControlled/Codes/2_figures_tables.R"),
)
source(here::here("GitControlled/Codes/2_figures_tables.R"))

knitr::purl(
  here::here("GitControlled/Codes/3_appendix.Rmd"),
  here::here("GitControlled/Codes/3_appendix.R"),
)
source(here::here("GitControlled/Codes/3_appendix.R"))
```

```{r setup-main}
#| include: false
#| cache: false

library(knitr)

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  fig.width = 5,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

# Introduction

Groundwater-irrigated agriculture supports X% of crop production worldwide and Y% in U.S. (citations) It is a vital natural resource that brings tremendous economic and social benefits to the society. However, groundwater has been depleting rapidly all over the world due to intensive groundwater extraction that far exceeds the natural replenishment rate (citation). The importance of irrigation has been consistently growing over the years (Irmak).

+ some mentionin of food security issues

The vast majority of studies on the climate change impacts on agricultural productivity focus on dryland production as it is far vulnerable to climate variations than irrigated production is. Another strand of literature seeks to understand the impact of climate change on water demand. Only a small collection of studies estimated the climate change impacts on irrigated production. It first seems that we do not need to understand the impact of climate change on irrigated production because irrigation provide a protection against climate variation. This would be true under the condition that farmer can apply desired amount of water without any physical or economic constraints, in which case, farmers can always take advantage of the benefit of irrigation to its full potential. Unfortunately, that is not the case for groundwater irrigated agriculture. Specifically, the stock of groundwater in the aquifer, measured as saturated thickness, affects the physical and economic constraints under which farmers operate. A decline in saturated thickness place stronger constraints on farmers and hurt the profitability of irrigated production.

There are mainly two reasons that saturated thickness affects the ability of aquifer to protect farmers from undesirable production conditions. First, as saturated thickness goes down due to groundwater depletion, the economic cost of pumping water increases. This affects overall agricultural productivity through two channels: extensive and intensive margins. When the cost of groundwater extraction becomes too high, irrigated production is no longer more profitable than dryland production. In this case, farmers will switch to irrigated production to dryland production even though groundwater still exists in the aquifer, which leads to significant reduction in crop yield (extensive margin). For those who have decided to irrigate, they would use less water as the cost of groundwater pumping is higher, thus lowering crop yield (intensive margin). Second and more important, well yield (a measure of how much water can be pumped per unit of time) declines as saturated thickness goes down (). This physical constraint has been shown to have far greater impacts on irrigation decisions than pumping cost [@mieno2021]. Similarly to the rising pumping cost case, declining well yield affects agricultural productivity through two channels. Too low a well yield makes producers shift to dryland production. For those irrigating, a lower well yield would lead to lower crop yield.

@sampson2019 estimated the impact of saturated thickness on land value.

The vast majority of existing body of studies focus on estimating the impact of climate on dryland production. A small amount of studies that look at the impact of climate on irrigated production does not take into account saturated thickness. Our study fills this important gap and sheds light on the value of groundwater by recognizing the importance of saturated thickness and quantifying its ability to protect agricultural production from undesirable weather outcomes.

<!-- Changes in saturated thickness alter the irrigation environment in two major ways. First, as saturated thickness declines, pumping cost increases as the lyft length becomes larger. Consequently, farmers are incetivized to apply less water, which then could result in lower yield. If the decline in saturated thickness is so significant and pumping costs become high enough, farmers may stop irrigating and revert back to dryland production. Another and arguably more important consequence of a decline in saturated thickness is lower well-yield. As well-yield declines, farmers may not be able to catch up with daily crop water demand, damging crop yield potential. Low well-yield is particularly problematic in extreme drought. When well-yield becomes too low, farmers revert back to dryland production. -->

[@rad2020effects]
[@foster2014modeling]
[@foster2015analysis]

# Materials and Method

In order to understand the impact of saturated thickness on agricultural production, this study examines the two channels through which saturated thickness affects average corn and soybean yields using econometric analysis. First, we quantify the impact of water deficit ($WB$) during the production season (May through September) on irrigated and dryland yields ($Y$) differentiated by the level of saturated thickness ($ST$). Mathematically, we estimate the expected value of yield conditional on water deficit, saturated thickness, and whether irrigated or not ($I$).

$$
E[Y_{i,t}|WB_{i,t}, ST_{i,t}, I_{i,t}]= f(WB_{i,t}, ST_{i,t}, I_{i,t})
$$

where $i$ indicates county and $t$ indicates year.

\noindent Second, we estimate the expected value of the share of irrigated acres as a function of saturated thickness and average water deficit (averaged over years). Historical average water deficit is used in place of concurrent water deficit $WB_{i,t}$ because the decision of whether to irrigate or not is made based on what the history of crop water demand rather than the concurrent weather that is yet to unfold at the time of decision making.

$$
E[S_{i,t}|WB_{i,t}, ST_{i,t}]= g(\overline{WB}_i, ST_{i,t}) 
$$

\noindent Once they are estimated, we can estimate the average crop yield as the share-weighted average of dryland and irrigated yields as follows:

$$
\hat{h}(WB_t, St_t)= (1 - \hat{g}(\overline{WB}, ST) )\cdot \hat{f}(WB_t, ST_t, I_t = 0)+ \hat{g}(\overline{WB}, ST) \cdot \hat{f}(WB_t, ST_t, I_t = 1)
$$

\noindent where $\hat{f}(\cdot)$ and $\hat{g}(\cdot)$ are estimated version of $f(\cdot)$ and $g(\cdot)$, respectively. For example, the impact of a change in saturate thickness level, say from $ST_{before}$ to $ST_{after}$, on crop yield is then $\hat{h}(WB_t, St_{before})$ - $\hat{h}(WB_t, St_{before})$.

## Econometric Models

We use the same model specification for corn and soybean yield regression analysis. The estimating equation is written as follows:

$$
y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t}
$$

Individual county and year are indicated by $i$ and $t$, respectively. In the model, $y_{i,t}$ is crop yield for county $i$ in year$t$. $C_j$ is a dummy variable that takes 1 if observation $i$ is under category $C_j$, 0 otherwise. The observations are grouped into four categories: dryland production, observations with saturated thickness falling under the $[0-33.3\%)$j, $[33.3-66.7\%)$, and $[66.7\% - 100\%]$ quantile ranges of saturated thickness. The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category. $\sigma_j$ represents the category specific intercept. $WB_{i,t}$ is water deficit defined as total reference evapotranspiration less precipitation between May through September. Category-specific impact of water deficit on yield is represented by $\sum_{k}^K \beta_k\phi_{j,k}(WB_{i,t})$, where $\phi_{j,k}$ is $k$th spline basis function and $\beta_k$ is its coefficient. Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner. The model can be interpreted as hierarchical generalized additive model, where each of saturated thickness level category has its own yield response function to drought (cite). To control for time-invariant  heterogeneity across counties and yearly shocks, county fixed effects ($\alpha_i$) and year fixed effects ($\phi_t$) are included. Standard error estimation is clustered by county.

The impact of saturated thickness on the share of irrigated acres is estimated using a fractional logit model under the generalize additive model framework in which the impact of saturated thickness, average water deficit, and their interactions are estimated semi-parametrically, allowing for flexible representation of their impacts. The estimating equation is as follows:

$$
log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + h(ST_{i,t}, \overline{WB}_i) + \mu_t + v_{i,t}
$$

, where $s_{i,t}$ is the share of irrigated production. $h(ST_{i,t}, \overline{WB}_i)$ consists of basis functions and their coefficients to repsent the the impact of $ST_{i,t}$, $\overline{WB}_i$, and their interactions (see Appendix ?? for further details of how $h(ST_{i,t}, \overline{WB}_i)$ are mathematically defined).

## Data

Figure \@ref(fig:geo-focus) shows the geographical coverage of the data for regression analysis for corn and soybean.

```{r, fig.dim = c(6, 5)}
#| label: geo-focus
#| fig-cap: "Geographic focus of the study for corn and soybean"

tm_map
```

First, county-level irrigated and dryland yield data for corn and soybean for 1985 through 2016 were obtained from the USDA NASS QuickStat. Both dryland and irrigated yield observations are used for counties that shares more than 75% of their area with the Ogallala aquifer. Only dryland yield observations are used for counties that do not satisfy the condition. We use dryland yield data outside of the Ogallala aquifer to supplement dryland yield data for more accurate estimation of the impact of water deficit for dryland production. Daily precipitation and reference ET were obtained from gridMET [@Abatzoglou2013]. Water deficit ($WB_{i,t}$) was then calculated by subtracting total precipitation from total reference evapotranspiration from May through September. Saturated thickness data were obtained from Dr. Erin Haacker [@haacker2016water]. Observations that have saturated thickness value lower than 12 meter were eliminated as saturated thickness below that is considered not suitable for irrigation (Tim citations here).

Figure \@ref(fig:deficit-hist) presents annual water deficit by year for the corn and soybean data. Water deficit is highly variable year to year. Year 2012 in particular stands out as an extremely dry year in the region. We use the year-to-year variations in water deficit to identify its impact on crop yield. Figure \@ref(fig:yield-hist) compares historical corn and soybean yields for dryland and irrigated production. The power of irrigation is clearly evident in this figure. First, irrigation provides protection against severe drought years like 2000, 2002, 2012. While the dryland yield dives down in those years, irrigated yields is almost unaffected by the severe droughts. First key objective of our study is to see if and how much the ability of irrigation is diminished when saturated thickness goes down. Figure \@ref(fig:sat-dist) presents the distribution of saturated thickness. Saturated thickness vary substantially, but most of the sites have saturated thickness of less than 50 meters. We exploit the observed variations in saturated thickness to identify its influence on the ability of irrigation to drought. Second, irrigated yield is always about 50% or higher than dryalnd yield irrespective of the degree of water deficit. We will investigate the negative impact of a decline in saturated thickness on the percentage of irrigated fields, thus reducing crop yields.

```{r }
#| fig-cap: "Distribution of water deficit for corn and soybean datasets"
#| label: deficit-hist
#| fig-width: 6
#| fig-height: 3.5

g_wb_box
```

```{r}
#| fig-cap: "Historical dryland and irrigated yields for corn and soybean"
#| label: yield-hist
#| fig-width: 6
#| fig-height: 3.5

g_yield_hist
```

```{r}
#| fig-cap: "Distribution of saturated thickness"
#| label: sat-dist
#| fig-width: 6
#| fig-height: 3.5

g_sat_dist
```

# Results and Discussions

## The impact of drought on dryland and irrigated yield

Regression results are presented graphically rather than in tables because individual coefficients estimates on spline basis functions cannot be readily interpreted and also because visualized results are much easier to interpret. Figure \@ref(fig:yield-response) shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).

```{r}
#| label: yield-response
#| fig-cap: "The impact of water deficit on dryland and irrigated yield by crop"
#| fig-width: 6
#| fig-height: 4

g_yield
```

In a fairly wet year without much water stress, dryland and irrigated production result in very similar crop yields for both soybean and corn yields. Irrigated yields are only slightly higher than dryland yields at a negative value of water deficit (precipitation is greater than reference evapotranspiration). However, dryland yield declines sharply and steadily as water deficit increases. On the contrary, irrigated corn yield actually increase as water deficit increases up to about water deficit of 400mm at all the saturated thickness levels for both corn and soybean. This is because irrigation can turn unfavorable high-heat conditions into favorable growing conditions up to a point (Tim, citations?). However, yield response to water deficit departs after deficit of about 400mm depending on the level of saturated thickness. While higher saturated thickness categories can withstand water deficit and crop yield stays about the same for higher water deficit values, the low saturated thickness category experience non-trivial declines in crop yields as water deficit increases.

While there are some gaps in crop yield between the lowest and the higher saturated thickness categories, the lowest category still performs quite well for both corn and soybean. This is because the samples in irrigation in the regression analysis are those who opted to stay irrigated even when the saturated thickness level is low because irrigation is still powerful enough to provide protections for farmers. So, fields who would have experienced significant declines in crop yield in severe droughts due to low saturated thickness have opted out of irrigation and thus not observable in the data for irrigated production. Therefore, In order to capture the fuller impact of a decline in saturated thickness, it is important to understand how saturated thickness affects the decision of whether to irrigate or not.

## The impact of saturated thickness on the share of irrigated production

```{r ir-share-plot}
#| label: ir-share
#| fig-cap: "The impact of saturated thickness on the share of irriagate production by crop "
#| fig-width: 6
#| fig-height: 3.5

g_share
```

As shown in \@ref(fig:ir-share), saturated thickness has significant impacts on the share of irrigated production for corn. As saturated thickness declines, the share of irrigated acres for corn decreases. As saturated thickness declines, more land will be under dryland production, which is more vulnerable to drought, damaging average corn yield. Unlike corn, a decline in saturated thickness does not have as large negative impact on the share of irrigated production for soybean as corn. This is likely because soybean is less susceptible to drought compared to corn.

## Total impact of decline in saturated thickness

Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated. For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness. The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher. In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively. These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness. In summary, our studies confirms that a decline in saturated thickness makes yield production much more vulnerable to severe drought, and its negative impact is non-negligible.

It is also noteworthy that the impact of saturated thickness on crop yield in drought years is non-linear. Specifically, according to our estimates, the decline of saturated thickness from 186 meters to 75 meters would cause about one tonne/ha of damage in corn yield at water deficit of 1100mm. However, the decline of saturated thickness from 75 meters to 30 meters would cause about 1.5 tonne/ha of damage in corn yield. This means that the negative impact of a meter of decline in saturated thickness becomes larger as the aquifer depletion progresses. This has an important policy implication. Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwater-irrigated production. Then, we need to act early before saturated thickness becomes too low so that we do not need to product at a less productive steady state. The longer you wait, the harsher damage is going to be when you finally decide to place a policy that limits groundwater pumping.

```{r total-impact}
#| fig-cap: "The impact of water deficit on average crop yield by sturated thickness"
#| label: tot-impact
#| fig-width: 6
#| fig-height: 3.5
g_total
```

# Conclusion

This article quantified the impact of saturated thickness on average crop yield...

\clearpage

# (APPENDIX) Appendix {-}

# Use SPEI in place of water deficit

Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \@ref(tab:fit-reg)).

```{r}
#| label: fit-reg
#| tbl-cap: Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean

data.table(
  Crop = c("Corn", "Corn", "Soybean", "Soybean"),
  Variable = c("Water Deficit", "SPEI", "Water Deficit", "SPEI"),
  `R-squared` = c(r2_corn_wb, r2_corn_spei, r2_soy_wb, r2_soy_spei)
) %>%
  .[, `R-squared` := round(`R-squared`, digits = 3)] %>%
  dcast(Crop ~ Variable, value.var = "R-squared") %>%
  kbl(caption = "Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean", booktabs = T) %>%
  kable_styling(latex_options = c("hold_position"))
```

Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \@ref(fig:spei).

```{r}
#| label: spei
#| fig-cap: The impact of SPEI on crop yield for corn and soybean

g_yield
```

\clearpage

### 2022-09-03 Writing/manuscript_DR.qmd
initial commit

--- 2022-09-03 Writing/manuscript_DR.Rmd
@@
 title: The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather
-author:
-  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
-  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
-  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota]
-  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
+authors:
+- name: Taro Mieno
+  address: Department of Agricultural Economics, University of Nebraska Lincoln
+  email: tmieno2@unl.edu
+  footnote: Something here
+- name: Timothy Foster
+  address: Department of WWW, University of Manchester
+  email: author@xxx.edu
+- name: Shunkei Kakimoto
+  address: Department of Agricultural Economics, University of Nebraska Lincoln
+  email: author@zzz.edu
+- name: Nicholas Brozovic
+  address: Department of Agricultural Economics, University of Nebraska Lincoln
+  email: author@zzz.edu
 abstract: |
@@
 csl: iopart.csl
-output:
-  bookdown::pdf_document2:
-    citation_package: natbib
-    keep_tex: true
-    toc: false
+format:
+  pdf:
+    cite-method: natbib
+    keep-tex: true
 ---
@@
 library(patchwork)
-library(kableExtra)
 ```
@@
 
-```{r run-analys, cache = TRUE, include = FALSE}
+```{r run-analys}
+#| include: false
+#| cache: false
 knitr::purl(
-  here::here("GitControlled/Codes/1_analysis.Rmd"),
-  here::here("GitControlled/Codes/1_analysis.R"),
+  here::here("Codes/1_analysis.Rmd"),
+  here::here("Codes/1_analysis.R"),
 )
-source(here::here("GitControlled/Codes/1_analysis.R"))
+source(here::here("Codes/1_analysis.R"))
 
 knitr::purl(
-  here::here("GitControlled/Codes/2_figures_tables.Rmd"),
-  here::here("GitControlled/Codes/2_figures_tables.R"),
+  here::here("Codes/2_figures_tables.Rmd"),
+  here::here("Codes/2_figures_tables.R"),
 )
-source(here::here("GitControlled/Codes/2_figures_tables.R"))
+source(here::here("Codes/2_figures_tables.R"))
 
 knitr::purl(
-  here::here("GitControlled/Codes/3_appendix.Rmd"),
-  here::here("GitControlled/Codes/3_appendix.R"),
+  here::here("Codes/3_appendix.Rmd"),
+  here::here("Codes/3_appendix.R"),
 )
-source(here::here("GitControlled/Codes/3_appendix.R"))
+source(here::here("Codes/3_appendix.R"))
 ```
@@
 In the model, $y_{i,t}$ is crop yield for county $i$ in year$t$.
-$C_j$ is a dummy variable that takes 1 if observation $i$ is under category $C_j$, 0 otherwise.
+$C_j$ is a dummy variable that takes 1 if observation $i$ is under catecotry $C_j$, 0 otherwise.
 The observations are grouped into four categories: dryland production, observations with saturated thickness falling under the $[0-33.3\%)$j, $[33.3-66.7\%)$, and $[66.7\% - 100\%]$ quantile ranges of saturated thickness.
@@
 $\sigma_j$ represents the category specific intercept.
-$WB_{i,t}$ is water deficit defined as total reference evapotranspiration less precipitation between May through September.
+$WB_{i,t}$ is water deficit defined as total reference evapotranspitration less precipitation between May through September.
 Category-specific impact of water deficit on yield is represented by $\sum_{k}^K \beta_k\phi_{j,k}(WB_{i,t})$, where $\phi_{j,k}$ is $k$th spline basis function and $\beta_k$ is its coefficient.
@@
 
-The impact of saturated thickness on the share of irrigated acres is estimated using a fractional logit model under the generalize additive model framework in which the impact of saturated thickness, average water deficit, and their interactions are estimated semi-parametrically, allowing for flexible representation of their impacts.
+The impact of saturated thickness on the share of irrigated acres is estiamted using a fractional logit model under the generalize additive model framwork in which the impact of saturated thickness, average water deficit, and their interactions are estimated semi-parametrically, allowing for flexible representation of their impacts.
 The estimating equation is as follows:
@@
 
-Figure \@ref(fig:geo-focus) shows the geographical coverage of the data for regression analysis for corn and soybean.
+@fig-geo-focus shows the geographical coverage of the data for regression analysis for corn and soybean.
 
-```{r, fig.dim = c(6, 5)}
-#| label: geo-focus
+```{r}
 #| fig-cap: "Geographic focus of the study for corn and soybean"
-
+#| label: fig-geo-focus
+#| fig-width: 5
 tm_map
@@
 Erin Haacker [@haacker2016water].
-Observations that have saturated thickness value lower than 12 meter were eliminated as saturated thickness below that is considered not suitable for irrigation (Tim citations here).
+Observations that have saturated thickness value lower than 12 meter were eliminated as saturated thicknees below that is considered not suitable for irrigation (Tim citations here).
 
-Figure \@ref(fig:deficit-hist) presents annual water deficit by year for the corn and soybean data.
+@fig-deficit-hist presents annual water deficit by year for the corn and soybean data.
 Water deficit is highly variable year to year.
 Year 2012 in particular stands out as an extremely dry year in the region.
-We use the year-to-year variations in water deficit to identify its impact on crop yield.
-Figure \@ref(fig:yield-hist) compares historical corn and soybean yields for dryland and irrigated production.
+We use the year-to-year variations in water deficit to idenfity its impact on crop yield.
+@fig-yield-hist compares historical corn and soybean yields for dryland and irrigated production.
 The power of irrigation is clearly evident in this figure.
@@
 While the dryland yield dives down in those years, irrigated yields is almost unaffected by the severe droughts.
-First key objective of our study is to see if and how much the ability of irrigation is diminished when saturated thickness goes down.
-Figure \@ref(fig:sat-dist) presents the distribution of saturated thickness.
-Saturated thickness vary substantially, but most of the sites have saturated thickness of less than 50 meters.
+First key objective of our study is to see if and how much the ability of irrigation is dminisihed when saturated thickness goes down.
+@fig-sat-dist presents the distribution of saturated thickness.
+Saturated thickness vary substantially, but most of the sites have saturated thickness of less than 50 meteres.
 We exploit the observed variations in saturated thickness to identify its influence on the ability of irrigation to drought.
@@
 
-```{r }
+```{r}
 #| fig-cap: "Distribution of water deficit for corn and soybean datasets"
-#| label: deficit-hist
-#| fig-width: 6
-#| fig-height: 3.5
+#| label: fig-deficit-hist
+#| fig-width: 5
 
@@
 #| fig-cap: "Historical dryland and irrigated yields for corn and soybean"
-#| label: yield-hist
-#| fig-width: 6
-#| fig-height: 3.5
+#| label: fig-yield-hist
+#| fig-width: 5
 
@@
 #| fig-cap: "Distribution of saturated thickness"
-#| label: sat-dist
-#| fig-width: 6
-#| fig-height: 3.5
+#| label: fig-sat-dist
+#| fig-width: 5
 
@@
 Regression results are presented graphically rather than in tables because individual coefficients estimates on spline basis functions cannot be readily interpreted and also because visualized results are much easier to interpret.
-Figure \@ref(fig:yield-response) shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).
+Figure @fig-yield shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).
 
 ```{r}
-#| label: yield-response
+#| label: fig-yield
 #| fig-cap: "The impact of water deficit on dryland and irrigated yield by crop"
-#| fig-width: 6
-#| fig-height: 4
+#| fig-width: 5
 
@@
 In a fairly wet year without much water stress, dryland and irrigated production result in very similar crop yields for both soybean and corn yields.
-Irrigated yields are only slightly higher than dryland yields at a negative value of water deficit (precipitation is greater than reference evapotranspiration).
+Irrigated yields are only slighly higher than dryland yields at a negative value of water deficit (precipitation is greater than reference evapotraspiration).
 However, dryland yield declines sharply and steadily as water deficit increases.
@@
 However, yield response to water deficit departs after deficit of about 400mm depending on the level of saturated thickness.
-While higher saturated thickness categories can withstand water deficit and crop yield stays about the same for higher water deficit values, the low saturated thickness category experience non-trivial declines in crop yields as water deficit increases.
+While higher saturated thickness categories can withstand water deficit and crop yield stays about the same for higer watger deficit values, the low saturated thickness category experience non-trivial declines in crop yields as water deficit increases.
 
@@
 ```{r ir-share-plot}
-#| label: ir-share
+#| label: fig-ir-share
 #| fig-cap: "The impact of saturated thickness on the share of irriagate production by crop "
-#| fig-width: 6
-#| fig-height: 3.5
 
@@
 
-As shown in \@ref(fig:ir-share), saturated thickness has significant impacts on the share of irrigated production for corn.
+As shown in @fig-ir-share, saturated thickness has significant impacts on the share of irrigated production for corn.
 As saturated thickness declines, the share of irrigated acres for corn decreases.
@@
 
-Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated.
+Taking into account both impacts of saturated thickness, figure @fig-tot-impact shows how water deficit affects yield at different saturated thickness levels evaluated.
 For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness.
@@
 This has an important policy implication.
-Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwater-irrigated production.
+Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwawter-irrigated production.
 Then, we need to act early before saturated thickness becomes too low so that we do not need to product at a less productive steady state.
@@
 #| fig-cap: "The impact of water deficit on average crop yield by sturated thickness"
-#| label: tot-impact
-#| fig-width: 6
-#| fig-height: 3.5
+#| label: fig-tot-impact
 g_total
@@
 
-\clearpage
+\newpage
 
-# (APPENDIX) Appendix {-}
+# Appendix {.appendix}
 
-# Use SPEI in place of water deficit
+## Use SPEI in place of water deficit
 
-Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \@ref(tab:fit-reg)).
+Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see @tbl-fit).
 
 ```{r}
-#| label: fit-reg
+#| label: tbl-fit
 #| tbl-cap: Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean
@@
   Variable = c("Water Deficit", "SPEI", "Water Deficit", "SPEI"),
-  `R-squared` = c(r2_corn_wb, r2_corn_spei, r2_soy_wb, r2_soy_spei)
+  `R-squared` = c(r2_corn_wb, r2_corn_spei, r2_corn_wb, r2_soy_spei)
 ) %>%
-  .[, `R-squared` := round(`R-squared`, digits = 3)] %>%
-  dcast(Crop ~ Variable, value.var = "R-squared") %>%
-  kbl(caption = "Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean", booktabs = T) %>%
-  kable_styling(latex_options = c("hold_position"))
+modelsummary::datasummary(`R-squared` ~ Crop * Variable * (Mean), data = .)
 ```
 
-Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \@ref(fig:spei).
+Qualitatively, the results with SPEI are similar to the regression reulsts with water deficit as can be seen in @fig-spei.
 
 ```{r}
-#| label: spei
+#| label: fig-spei
 #| fig-cap: The impact of SPEI on crop yield for corn and soybean
@@
 
-\clearpage
+<!-- ## Continuous saturated thickness -->
+
+\newpage
+
+### References
+
+::: {#refs}
+:::

### 2022-09-03 Writing/manuscript_DR.tex
initial commit

--- 2022-09-03 Writing/manuscript_DR.Rmd
@@
----
-title: The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather
-author:
-  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
-  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
-  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota]
-  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
-abstract: |
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather},
+  pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\title{The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather}
+\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
+\date{}
+
+\begin{document}
+\maketitle
+\begin{abstract}
 Abstract goes here. It should be a single paragraph of around 200 words.
-acknowledgments: |
-  We wish to thank...
-bibliography: DRA.bib
-csl: iopart.csl
-output:
-  bookdown::pdf_document2:
-    citation_package: natbib
-    keep_tex: true
-    toc: false
----
+\end{abstract}
 
-```{r packages}
-#| include: false
-#| cache: false
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-#--- packages ---#
-library(data.table)
-library(tidyverse)
-library(flextable)
-library(stringr)
-library(sf)
-library(here)
-library(tmap)
-library(lfe)
-library(modelsummary)
-library(patchwork)
-library(kableExtra)
-```
-
-```{r}
-#| echo: false
-
-options(knitr.duplicate.label = "allow")
-```
-
-```{r run-analys, cache = TRUE, include = FALSE}
-knitr::purl(
-  here::here("GitControlled/Codes/1_analysis.Rmd"),
-  here::here("GitControlled/Codes/1_analysis.R"),
-)
-source(here::here("GitControlled/Codes/1_analysis.R"))
-
-knitr::purl(
-  here::here("GitControlled/Codes/2_figures_tables.Rmd"),
-  here::here("GitControlled/Codes/2_figures_tables.R"),
-)
-source(here::here("GitControlled/Codes/2_figures_tables.R"))
-
-knitr::purl(
-  here::here("GitControlled/Codes/3_appendix.Rmd"),
-  here::here("GitControlled/Codes/3_appendix.R"),
-)
-source(here::here("GitControlled/Codes/3_appendix.R"))
-```
-
-```{r setup-main}
-#| include: false
-#| cache: false
-
-library(knitr)
-
-opts_chunk$set(
-  fig.align = "center",
-  fig.retina = 5,
-  fig.width = 5,
-  fig.height = 4.5,
-  warning = FALSE,
-  message = FALSE,
-  echo = FALSE
-)
-```
-
-# Introduction
-
-Groundwater-irrigated agriculture supports X% of crop production worldwide and Y% in U.S.
+Groundwater-irrigated agriculture supports X\% of crop production worldwide and Y\% in U.S.
 (citations) It is a vital natural resource that brings tremendous economic and social benefits to the society.
@@
 
-+ some mentionin of food security issues
+\begin{itemize}
+\tightlist
+\item
+  some mentionin of food security issues
+\end{itemize}
 
@@
 Second and more important, well yield (a measure of how much water can be pumped per unit of time) declines as saturated thickness goes down ().
-This physical constraint has been shown to have far greater impacts on irrigation decisions than pumping cost [@mieno2021].
+This physical constraint has been shown to have far greater impacts on irrigation decisions than pumping cost \citep{mieno2021}.
 Similarly to the rising pumping cost case, declining well yield affects agricultural productivity through two channels.
@@
 
-@sampson2019 estimated the impact of saturated thickness on land value.
+\citet{sampson2019} estimated the impact of saturated thickness on land value.
 
@@
 
-<!-- Changes in saturated thickness alter the irrigation environment in two major ways.
-First, as saturated thickness declines, pumping cost increases as the lyft length becomes larger.
-Consequently, farmers are incetivized to apply less water, which then could result in lower yield.
-If the decline in saturated thickness is so significant and pumping costs become high enough, farmers may stop irrigating and revert back to dryland production.
-Another and arguably more important consequence of a decline in saturated thickness is lower well-yield.
-As well-yield declines, farmers may not be able to catch up with daily crop water demand, damging crop yield potential.
-Low well-yield is particularly problematic in extreme drought.
-When well-yield becomes too low, farmers revert back to dryland production.
--->
+\citep{rad2020effects}
+\citep{foster2014modeling}
+\citep{foster2015analysis}
 
-[@rad2020effects]
-[@foster2014modeling]
-[@foster2015analysis]
-
-# Materials and Method
+\hypertarget{materials-and-method}{%
+\section{Materials and Method}\label{materials-and-method}}
 
 In order to understand the impact of saturated thickness on agricultural production, this study examines the two channels through which saturated thickness affects average corn and soybean yields using econometric analysis.
-First, we quantify the impact of water deficit ($WB$) during the production season (May through September) on irrigated and dryland yields ($Y$) differentiated by the level of saturated thickness ($ST$).
-Mathematically, we estimate the expected value of yield conditional on water deficit, saturated thickness, and whether irrigated or not ($I$).
+First, we quantify the impact of water deficit (\(WB\)) during the production season (May through September) on irrigated and dryland yields (\(Y\)) differentiated by the level of saturated thickness (\(ST\)).
+Mathematically, we estimate the expected value of yield conditional on water deficit, saturated thickness, and whether irrigated or not (\(I\)).
 
-$$
+\[
 E[Y_{i,t}|WB_{i,t}, ST_{i,t}, I_{i,t}]= f(WB_{i,t}, ST_{i,t}, I_{i,t})
-$$
+\]
 
-where $i$ indicates county and $t$ indicates year.
+where \(i\) indicates county and \(t\) indicates year.
 
 \noindent Second, we estimate the expected value of the share of irrigated acres as a function of saturated thickness and average water deficit (averaged over years).
-Historical average water deficit is used in place of concurrent water deficit $WB_{i,t}$ because the decision of whether to irrigate or not is made based on what the history of crop water demand rather than the concurrent weather that is yet to unfold at the time of decision making.
+Historical average water deficit is used in place of concurrent water deficit \(WB_{i,t}\) because the decision of whether to irrigate or not is made based on what the history of crop water demand rather than the concurrent weather that is yet to unfold at the time of decision making.
 
-$$
+\[
 E[S_{i,t}|WB_{i,t}, ST_{i,t}]= g(\overline{WB}_i, ST_{i,t}) 
-$$
+\]
 
@@
 
-$$
+\[
 \hat{h}(WB_t, St_t)= (1 - \hat{g}(\overline{WB}, ST) )\cdot \hat{f}(WB_t, ST_t, I_t = 0)+ \hat{g}(\overline{WB}, ST) \cdot \hat{f}(WB_t, ST_t, I_t = 1)
-$$
+\]
 
-\noindent where $\hat{f}(\cdot)$ and $\hat{g}(\cdot)$ are estimated version of $f(\cdot)$ and $g(\cdot)$, respectively.
-For example, the impact of a change in saturate thickness level, say from $ST_{before}$ to $ST_{after}$, on crop yield is then $\hat{h}(WB_t, St_{before})$ - $\hat{h}(WB_t, St_{before})$.
+\noindent where \(\hat{f}(\cdot)\) and \(\hat{g}(\cdot)\) are estimated version of \(f(\cdot)\) and \(g(\cdot)\), respectively.
+For example, the impact of a change in saturate thickness level, say from \(ST_{before}\) to \(ST_{after}\), on crop yield is then \(\hat{h}(WB_t, St_{before})\) - \(\hat{h}(WB_t, St_{before})\).
 
-## Econometric Models
+\hypertarget{econometric-models}{%
+\subsection{Econometric Models}\label{econometric-models}}
 
@@
 
-$$
+\[
 y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t}
-$$
+\]
 
-Individual county and year are indicated by $i$ and $t$, respectively.
-In the model, $y_{i,t}$ is crop yield for county $i$ in year$t$.
-$C_j$ is a dummy variable that takes 1 if observation $i$ is under category $C_j$, 0 otherwise.
-The observations are grouped into four categories: dryland production, observations with saturated thickness falling under the $[0-33.3\%)$j, $[33.3-66.7\%)$, and $[66.7\% - 100\%]$ quantile ranges of saturated thickness.
+Individual county and year are indicated by \(i\) and \(t\), respectively.
+In the model, \(y_{i,t}\) is crop yield for county \(i\) in year\(t\).
+\(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
+The observations are grouped into four categories: dryland production, observations with saturated thickness falling under the \([0-33.3\%)\)j, \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of saturated thickness.
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-$\sigma_j$ represents the category specific intercept.
-$WB_{i,t}$ is water deficit defined as total reference evapotranspiration less precipitation between May through September.
-Category-specific impact of water deficit on yield is represented by $\sum_{k}^K \beta_k\phi_{j,k}(WB_{i,t})$, where $\phi_{j,k}$ is $k$th spline basis function and $\beta_k$ is its coefficient.
+\(\sigma_j\) represents the category specific intercept.
+\(WB_{i,t}\) is water deficit defined as total reference evapotranspiration less precipitation between May through September.
+Category-specific impact of water deficit on yield is represented by \(\sum_{k}^K \beta_k\phi_{j,k}(WB_{i,t})\), where \(\phi_{j,k}\) is \(k\)th spline basis function and \(\beta_k\) is its coefficient.
 Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner.
 The model can be interpreted as hierarchical generalized additive model, where each of saturated thickness level category has its own yield response function to drought (cite).
-To control for time-invariant  heterogeneity across counties and yearly shocks, county fixed effects ($\alpha_i$) and year fixed effects ($\phi_t$) are included.
+To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are included.
 Standard error estimation is clustered by county.
@@
 
-$$
+\[
 log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + h(ST_{i,t}, \overline{WB}_i) + \mu_t + v_{i,t}
-$$
+\]
 
-, where $s_{i,t}$ is the share of irrigated production.
-$h(ST_{i,t}, \overline{WB}_i)$ consists of basis functions and their coefficients to repsent the the impact of $ST_{i,t}$, $\overline{WB}_i$, and their interactions (see Appendix ??
-for further details of how $h(ST_{i,t}, \overline{WB}_i)$ are mathematically defined).
+, where \(s_{i,t}\) is the share of irrigated production.
+\(h(ST_{i,t}, \overline{WB}_i)\) consists of basis functions and their coefficients to repsent the the impact of \(ST_{i,t}\), \(\overline{WB}_i\), and their interactions (see Appendix ??
+for further details of how \(h(ST_{i,t}, \overline{WB}_i)\) are mathematically defined).
 
-## Data
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
 
-Figure \@ref(fig:geo-focus) shows the geographical coverage of the data for regression analysis for corn and soybean.
+Figure \ref{fig:geo-focus} shows the geographical coverage of the data for regression analysis for corn and soybean.
 
-```{r, fig.dim = c(6, 5)}
-#| label: geo-focus
-#| fig-cap: "Geographic focus of the study for corn and soybean"
+\begin{figure}
 
-tm_map
-```
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/geo-focus-1}
+
+}
+
+\caption{Geographic focus of the study for corn and soybean}\label{fig:geo-focus}
+\end{figure}
 
 First, county-level irrigated and dryland yield data for corn and soybean for 1985 through 2016 were obtained from the USDA NASS QuickStat.
-Both dryland and irrigated yield observations are used for counties that shares more than 75% of their area with the Ogallala aquifer.
+Both dryland and irrigated yield observations are used for counties that shares more than 75\% of their area with the Ogallala aquifer.
 Only dryland yield observations are used for counties that do not satisfy the condition.
 We use dryland yield data outside of the Ogallala aquifer to supplement dryland yield data for more accurate estimation of the impact of water deficit for dryland production.
-Daily precipitation and reference ET were obtained from gridMET [@Abatzoglou2013].
-Water deficit ($WB_{i,t}$) was then calculated by subtracting total precipitation from total reference evapotranspiration from May through September.
-Saturated thickness data were obtained from Dr.
-Erin Haacker [@haacker2016water].
+Daily precipitation and reference ET were obtained from gridMET \citep{Abatzoglou2013}.
+Water deficit (\(WB_{i,t}\)) was then calculated by subtracting total precipitation from total reference evapotranspiration from May through September.
+Saturated thickness data were obtained from Dr.~Erin Haacker \citep{haacker2016water}.
 Observations that have saturated thickness value lower than 12 meter were eliminated as saturated thickness below that is considered not suitable for irrigation (Tim citations here).
 
-Figure \@ref(fig:deficit-hist) presents annual water deficit by year for the corn and soybean data.
+Figure \ref{fig:deficit-hist} presents annual water deficit by year for the corn and soybean data.
 Water deficit is highly variable year to year.
@@
 We use the year-to-year variations in water deficit to identify its impact on crop yield.
-Figure \@ref(fig:yield-hist) compares historical corn and soybean yields for dryland and irrigated production.
+Figure \ref{fig:yield-hist} compares historical corn and soybean yields for dryland and irrigated production.
 The power of irrigation is clearly evident in this figure.
@@
 First key objective of our study is to see if and how much the ability of irrigation is diminished when saturated thickness goes down.
-Figure \@ref(fig:sat-dist) presents the distribution of saturated thickness.
+Figure \ref{fig:sat-dist} presents the distribution of saturated thickness.
 Saturated thickness vary substantially, but most of the sites have saturated thickness of less than 50 meters.
 We exploit the observed variations in saturated thickness to identify its influence on the ability of irrigation to drought.
-Second, irrigated yield is always about 50% or higher than dryalnd yield irrespective of the degree of water deficit.
+Second, irrigated yield is always about 50\% or higher than dryalnd yield irrespective of the degree of water deficit.
 We will investigate the negative impact of a decline in saturated thickness on the percentage of irrigated fields, thus reducing crop yields.
 
-```{r }
-#| fig-cap: "Distribution of water deficit for corn and soybean datasets"
-#| label: deficit-hist
-#| fig-width: 6
-#| fig-height: 3.5
+\begin{figure}
 
-g_wb_box
-```
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/deficit-hist-1}
 
-```{r}
-#| fig-cap: "Historical dryland and irrigated yields for corn and soybean"
-#| label: yield-hist
-#| fig-width: 6
-#| fig-height: 3.5
+}
 
-g_yield_hist
-```
+\caption{Distribution of water deficit for corn and soybean datasets}\label{fig:deficit-hist}
+\end{figure}
 
-```{r}
-#| fig-cap: "Distribution of saturated thickness"
-#| label: sat-dist
-#| fig-width: 6
-#| fig-height: 3.5
+\begin{figure}
 
-g_sat_dist
-```
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/yield-hist-1}
 
-# Results and Discussions
+}
 
-## The impact of drought on dryland and irrigated yield
+\caption{Historical dryland and irrigated yields for corn and soybean}\label{fig:yield-hist}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/sat-dist-1}
+
+}
+
+\caption{Distribution of saturated thickness}\label{fig:sat-dist}
+\end{figure}
+
+\hypertarget{results-and-discussions}{%
+\section{Results and Discussions}\label{results-and-discussions}}
+
+\hypertarget{the-impact-of-drought-on-dryland-and-irrigated-yield}{%
+\subsection{The impact of drought on dryland and irrigated yield}\label{the-impact-of-drought-on-dryland-and-irrigated-yield}}
 
 Regression results are presented graphically rather than in tables because individual coefficients estimates on spline basis functions cannot be readily interpreted and also because visualized results are much easier to interpret.
-Figure \@ref(fig:yield-response) shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).
+Figure \ref{fig:yield-response} shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).
 
-```{r}
-#| label: yield-response
-#| fig-cap: "The impact of water deficit on dryland and irrigated yield by crop"
-#| fig-width: 6
-#| fig-height: 4
+\begin{figure}
 
-g_yield
-```
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/yield-response-1}
+
+}
+
+\caption{The impact of water deficit on dryland and irrigated yield by crop}\label{fig:yield-response}
+\end{figure}
 
@@
 
-## The impact of saturated thickness on the share of irrigated production
+\hypertarget{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}{%
+\subsection{The impact of saturated thickness on the share of irrigated production}\label{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}}
 
-```{r ir-share-plot}
-#| label: ir-share
-#| fig-cap: "The impact of saturated thickness on the share of irriagate production by crop "
-#| fig-width: 6
-#| fig-height: 3.5
+\begin{figure}
 
-g_share
-```
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/ir-share-1}
 
-As shown in \@ref(fig:ir-share), saturated thickness has significant impacts on the share of irrigated production for corn.
+}
+
+\caption{The impact of saturated thickness on the share of irriagate production by crop }\label{fig:ir-share}
+\end{figure}
+
+As shown in \ref{fig:ir-share}, saturated thickness has significant impacts on the share of irrigated production for corn.
 As saturated thickness declines, the share of irrigated acres for corn decreases.
@@
 
-## Total impact of decline in saturated thickness
+\hypertarget{total-impact-of-decline-in-saturated-thickness}{%
+\subsection{Total impact of decline in saturated thickness}\label{total-impact-of-decline-in-saturated-thickness}}
 
-Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated.
+Taking into account both impacts of saturated thickness, figure \ref{fig:tot-impact} shows how water deficit affects yield at different saturated thickness levels evaluated.
 For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness.
 The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher.
-In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively.
+In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15\% and 25\% higher than at 75.7 meters and 29.6 meters, respectively.
 These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness.
@@
 
-```{r total-impact}
-#| fig-cap: "The impact of water deficit on average crop yield by sturated thickness"
-#| label: tot-impact
-#| fig-width: 6
-#| fig-height: 3.5
-g_total
-```
+\begin{figure}
 
-# Conclusion
+{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/tot-impact-1}
 
-This article quantified the impact of saturated thickness on average crop yield...
+}
+
+\caption{The impact of water deficit on average crop yield by sturated thickness}\label{fig:tot-impact}
+\end{figure}
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+This article quantified the impact of saturated thickness on average crop yield\ldots{}
 
@@
 
-# (APPENDIX) Appendix {-}
+\hypertarget{appendix-appendix}{%
+\appendix}
 
-# Use SPEI in place of water deficit
+\hypertarget{use-spei-in-place-of-water-deficit}{%
+\section{Use SPEI in place of water deficit}\label{use-spei-in-place-of-water-deficit}}
 
-Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \@ref(tab:fit-reg)).
+Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \ref{tab:fit-reg}).
 
-```{r}
-#| label: fit-reg
-#| tbl-cap: Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean
+\begin{table}[!h]
 
-data.table(
-  Crop = c("Corn", "Corn", "Soybean", "Soybean"),
-  Variable = c("Water Deficit", "SPEI", "Water Deficit", "SPEI"),
-  `R-squared` = c(r2_corn_wb, r2_corn_spei, r2_soy_wb, r2_soy_spei)
-) %>%
-  .[, `R-squared` := round(`R-squared`, digits = 3)] %>%
-  dcast(Crop ~ Variable, value.var = "R-squared") %>%
-  kbl(caption = "Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean", booktabs = T) %>%
-  kable_styling(latex_options = c("hold_position"))
-```
+\caption{\label{tab:fit-reg}Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean}
+\centering
+\begin{tabular}[t]{lrr}
+\toprule
+Crop & SPEI & Water Deficit\\
+\midrule
+Corn & 0.899 & 0.903\\
+Soybean & 0.840 & 0.846\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \@ref(fig:spei).
+Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \ref{fig:spei}.
 
-```{r}
-#| label: spei
-#| fig-cap: The impact of SPEI on crop yield for corn and soybean
+\begin{figure}
 
-g_yield
-```
+{\centering \includegraphics[width=360px]{manuscript_DR_files/figure-latex/spei-1}
+
+}
+
+\caption{The impact of SPEI on crop yield for corn and soybean}\label{fig:spei}
+\end{figure}
 
 \clearpage
+
+  \bibliography{DRA.bib}
+
+\end{document}

### 2022-09-03 Writing/usda_report.Rmd
initial commit

---
title: "The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Droughts"
author:
  - Taro Mieno^[University of Nebraska Lincoln, tmieno2@unl.edu], Timothy Foster, Shunkei Kakimoto, Ahmed Chennak, Nicholas Brozovic
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: true
    reference_docx: "word_template.docx"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ": "
bibliography: DRA.bib
abstract: "Abstract: "
---

```{r echo = F}
library(knitr)
library(here)
library(officedown)
library(officer)

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = T,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(here)
library(tmap)
library(lfe)
library(modelsummary)
library(patchwork)
```

# List of activities done since April

## Groundwater and Saturated Thickness
+ Conduct a literature review
+ Collect gridMET weather data and assign weather variable values to each of the counties
+ Collect daily weather data up to 2099 under various climate change scenarios
+ Improve the modeling of corn and soybean yield response to weather variables: water deficit (reference evapotranspiration less precipitation)
+ Simulate the impact of saturated thickness on corn and soybean yield under various climate change scenarios
+ Start writing a manuscript for journal submission

## Land value

+ Continually negotiated with ERS to get access to the CoreLogic land value data
+ We are at the final stage of agreement. CoreLogic will send us a document, which all the permitted users will sign.

`r run_pagebreak()`

# Main findings with graphics/tables illustrating key points/findings.

# Key Finging 1:

Figure below show the impact of annual water deficit (the sum of reference evapotranspiration less the sum of precipitation from April through September). The greater the value of water deficit, the more severe drought is. However, it also captures the favorable growing conditions like warm weather at the same time. The figure below shows that groundwater provides strong protection against drought. For dryland production, yield starts to decline precipitously once water deficit goes beyond 0 mm (evapotranspiration and precipitation are equated). However, for irrigated production, corn and soy yields increase up to water deficit of 500 mm to 700 mm depending the level of saturated thickness, capturing the benefit of warm weather while negating damage from water deficit.

```{r}
include_graphics(here("Shared/Results/y_intensive.pdf"))
```

`r run_pagebreak()`

# Key Finging 2:

Saturated thickness has a large impact of the share of irrigated acres for both corn and soybean. As saturated thickness goes down, well yield (the amount of water that can be pumped per a certain period of time, often measured in gallons per minute) also goes down, which then results in a reduction in irrigated acres. Given the dryland production is so much more vulnerable to drought as shown above, saturated thickness has a large impact of "average" corn and soybean yield.

```{r}
include_graphics(here("Shared/Results/y_ir_share.pdf"))
```

`r run_pagebreak()`

# Key Finding 3:

The total impact of saturated thickness on corn is presented in the figure below, combining the two channels of how saturated thickness affects corn and soybean yield. As expected, the value of saturated thickness is greater in drier years. For example, for a year with water deficit of 1000 mm, the average yield at saturated thickness level of 60 and 450 are about 40 and 220, respectively. This clearly shows that maintaining or increasing saturated thickness levels is crucial from both the economic perspective and food security.

```{r }
include_graphics(here("Shared/Results/y_sat_impact_total.pdf"))
```

### 2022-09-03 Codes/0_1_hpa_counties.R

@@ -1,0 +1,175 @@
+#' ---
+#' title: "Prepare base-county data"
+#' author: "Taro Mieno"
+#' output:
+#'   html_document:
+#'     number_sections: yes
+#'     theme: flatly
+#'     highlight: zenburn
+#'     toc_float: yes
+#'     toc: yes
+#'     toc_depth: 3
+#' geometry: margin=1in
+#' ---
+
+#+ setup, include=FALSE
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = TRUE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi = 400,
+  fig.width = 7.5,
+  fig.height = 5,
+  out.width = "750px",
+  out.height = "500px"
+)
+
+# /*=================================================*/
+#' # Identify counties the overlaps with the HPA boundary
+# /*=================================================*/
+
+# /*----------------------------------*/
+#' ## Read the HPA border shapefile
+# /*----------------------------------*/
+#+ Read-HPA, cache = TRUE
+hpa <- st_read(here("Shared/Data/RawData/hp_bound2010.shp"))
+
+#' Simplify the HPA boundary for faster mapping
+hpa_simplified <-
+  here("Shared/Data/RawData/hp_bound2010.shp") %>%
+  st_read() %>%
+  st_simplify(dTolerance = 1000)
+
+saveRDS(hpa_simplified, here("Shared/Data/ProcessedData/hp_simplified.shp"))
+
+#' Map-of-HPA
+tm_shape(hpa_simplified) +
+  tm_polygons()
+
+# /*----------------------------------*/
+#' ## Get all the US counties as an sf
+# /*----------------------------------*/
+#+ Get-US-counties (entire US)
+(
+  US_county <-
+    tigris::counties(cb = TRUE) %>%
+    st_as_sf() %>%
+    st_transform(st_crs(hpa)) %>%
+    rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
+    dplyr::select(state_code, county_code)
+)
+
+# /*----------------------------------*/
+#' ## Identify counties that overlaps with `hpa`
+# /*----------------------------------*/
+#+ Identify-overlap, cache = TRUE
+hpa_counties <- US_county[hpa, ] %>%
[…]

### 2022-09-03 Codes/0_3_get_sat.R

@@ -1,0 +1,94 @@
+#' ---
+#' title: "Assign saturated thickness value "
+#' author: "Taro Mieno"
+#' output:
+#'   html_document:
+#'     number_sections: yes
+#'     theme: flatly
+#'     highlight: zenburn
+#'     toc_float: yes
+#'     toc: yes
+#'     toc_depth: 3
+#' geometry: margin=1in
+#' ---
+
+#+ setup, include=FALSE
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = FALSE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi=400,
+  fig.width=7.5,
+  fig.height=5,
+  out.width="750px",
+  out.height="500px"
+)
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+
+#' Import saturated thickness data
+sat_temp <- raster("/Volumes/Elements/Database/SaturatedThickness/SatThick_19851.tif")
+
+#' HPA counties
+hpa_counties <- 
+  here("Shared/Data/ProcessedData/base_counties.rds") %>% 
+  readRDS() %>% 
+  .[in_hpa == 1, ] %>% 
+  st_as_sf() %>% 
+  st_transform(projection(sat_temp)) 
+
+#' Define function to extract saturated thickness for each county
+get_sat <- function(y) {
+  # y <- 2000
+  sat_temp <- 
+    paste0("/Volumes/Elements/Database/SaturatedThickness/SatThick_", y, "1.tif") %>% 
+    raster()
+
+  sat_mean <- 
+    exact_extract(sat_temp, hpa_counties, "mean")
+
+  data_return <- 
+    data.table(
+      sat = sat_mean,
+      sc_code = hpa_counties$sc_code,
+      year = y
+    )
+
+  return(data_return)
+}
+
+#' Extract saturated thickness values (parallelized)
+#+ extract-sat, eval = F
+sat_all <- mclapply(1985:2016, get_sat, mc.cores = 12) %>%
+  rbindlist() %>%
+  .[sat < 0, sat := 0] %>% 
+  .[, sat := sat * 3.28084] # meter to feet
+
+
+#/*=================================================*/
+#' # Save the data
+#/*=================================================*/
+#+ save, eval = F
+saveRDS(sat_all, here("Shared/Data/ProcessedData/sat_all.rds"))
+
[…]

### 2022-09-03 Codes/0_4_ndmc_di.R

@@ -1,0 +1,46 @@
+######################################
+# Drought Index
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+#' According to Kuwayama (2018), the six key physical indicators consider percentiles
+#' from a soil moisture model, daily streamflow percentiles, the percent of normal precipitation,
+#' a standardized precipitation index, and a remotely-sensed vegetation
+#' health index (Svoboda et al. 2002). The authors also rely on supplementary indicators
+#' such as humidity and temperature departure from normal, reservoir, and lake levels,
+#' surface water supply indices, snowpack, and groundwater levels. The drought Index data is saved
+#' below:
+
+DI_1 <- fread(here("Data/RawData/NDMC_CO_IL_IN_DI_IA_KS_MS.csv"))
+DI_2 <- fread(here("Data/RawData/NDMC_DI_NM_OK_SD_TX_WY_NE.csv"))
+
+DI <- rbind(DI_1, DI_2)
+
+#' The default drought index is weekly persentage data. We changed them to monthly data by using the codes below.
+#' However, we are examing for different "seasons", from April to September, from May to October,
+#' from April to October, and from May to September. So we need to filter each season's drought index.
+
+drought_county <- DI %>%
+  .[, start_date := as.Date(ValidStart)] %>%
+  .[, end_date := as.Date(ValidEnd)] %>%
+  .[, FIPS := as.character(FIPS)] %>%
+  .[str_length(FIPS) == 4, FIPS := paste0("0", FIPS)] %>%
+  .[, state_code := str_sub(FIPS, 1, 2)] %>%
+  .[, county_code := str_sub(FIPS, 3, 5)] %>%
+  .[month(end_date) > 4 & month(start_date) < 10, ] %>%
+  #=== state-county code ===#
+  .[, sc_code := paste0(state_code, county_code)] %>% 
+  .[, .(
+    d0_5_9 = sum(D0) / 100,
+    d1_5_9 = sum(D1) / 100,
+    d2_5_9 = sum(D2) / 100,
+    d3_5_9 = sum(D3) / 100,
+    d4_5_9 = sum(D4) / 100
+  ), by = .(sc_code, year(end_date))] %>%
+  data.table()
+
+saveRDS(drought_county, here("Data/ProcessedData/DI_NDMC.rds"))
+
+

### 2022-09-03 Codes/0_6_merge_datasets.R

@@ -1,0 +1,82 @@
+library(tidyverse)
+library(ggplot2)
+library(sf)
+library(raster)
+library(data.table)
+library(here)
+
+# /*=================================================*/
+#' # Combine all the datasets
+# /*=================================================*/
+base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
+
+#' NDMC Drought index
+drought_county <- readRDS(here("Shared/Data/ProcessedData/DI_NDMC.rds"))
+
+#' SPEI
+# spei_data <- readRDS(here("Shared/Data/ProcessedData/spei_data.rds"))
+
+#' Weather
+weather_data <- readRDS(here("Shared/Data/ProcessedData", "summrized_gridMET.rds"))
+
+#' Saturated thickness
+sat_all <- readRDS(here("Shared/Data/ProcessedData/sat_all.rds"))
+
+#' Yield, acreage data
+nass_data <- readRDS(here("Shared/Data/ProcessedData/nass_data.rds"))
+
+# /*----------------------------------*/
+#' ## combine all
+# /*----------------------------------*/
+
+merged_data <- base_counties[nass_data, on = c("sc_code")] %>%
+  weather_data[., on = c("sc_code", "year")] %>%
+  sat_all[., on = c("sc_code", "year")] %>%
+  drought_county[., on = c("sc_code", "year")] %>%
+  .[order(sc_code, year), ]
+
+# merged_data[ir == "nir", state] %>% unique()
+# merged_data[ir == "ir", state] %>% unique()
+
+
+# /*=================================================*/
+#' # Create and redefine new variables, select observations
+# /*=================================================*/
+
+final_data <-
+  merged_data %>%
+  # === if not in_hpa and irrigated, drop them ===#
+  # irrigated, but saturated thickness not observed (not even sure if
+  # it is supported by surface water or groundwater). These
+  # observations cannot be used.
+  .[!(in_hpa == 0 & ir == "ir"), ] %>%
+  # === if not irrigated, set sat = 0 ===#
+  .[ir == "nir", sat := 0] %>%
+  # === if sat == 0, but indicated irrigated ===#
+  # add 0.1 to distinguish dryland production, where
+  # sat is set at 0 above
+  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
+  # === drop if yield or acres are not observed ===#
+  .[!is.na(yield) & !is.na(acres), ] %>%
+  # === if sat is missing, drop ===#
[…]

### 2022-09-03 Codes/0_data_prep.R

@@ -1,0 +1,129 @@
+#' ---
+#' title: "Data preparation"
+#' author: "Taro Mieno"
+#' output:
+#'   html_document:
+#'     number_sections: yes
+#'     theme: flatly
+#'     highlight: zenburn
+#'     toc_float: yes
+#'     toc: yes
+#'     toc_depth: 3
+#' geometry: margin=1in
+#' ---
+
+#+ setup, include=FALSE
+
+library(knitr)
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = TRUE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi = 400,
+  fig.width = 7.5,
+  fig.height = 5,
+  out.width = "750px",
+  out.height = "500px"
+)
+
+library(here)
+
+opts_knit$set(
+  root.dir = here()
+)
+
+# /*=================================================*/
+#' # Objectives
+# /*=================================================*/
+#' Prepare datasets for econometric analysis of the impacts of saturated thickness on the
+#' impacts of drought on crop yield
+#'
+
+#' + Irrigated corn and soybean yields at the county level from 2000 to 2018 within the High Plains Aquifer measured in bushels per acre, downloaded from the National Agricultural Statistic Service (NASS) website.
+#' + County-level depths of saturated thickness of High Plains Aquifer form the 2000 to 2018, measured in feet.
+#' + For different degrees of drought, we use the Drought Index used by the U.S. Drought Monitor (Kuwayama 2018). Along D0 to D4, the severity of drought increases, which represents abnormally dry, drought moderate, drought sever, drought extreme, drought exceptional, respectively.
+#'
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(ggplot2)
+library(sf)
+library(raster)
+library(rnassqs)
+library(data.table)
+library(lfe)
+library(stargazer)
+library(here)
+library(keyring)
+library(tidyUSDA)
+library(tigris)
+library(tidycensus)
+library(exactextractr)
+library(parallel)
+
+# /*=================================================*/
+#' # Identify counties the overlaps with the HPA boundary
+# /*=================================================*/
+#' in: "Data/RawData/hp_bound2010.shp"
[…]

### 2022-09-03 Codes/1_A3_feols_smooth_exp.R

@@ -1,0 +1,205 @@
+
+N <- 300
+g_N <- (N + 1) / 5
+T <- 10
+
+data <- CJ(id = 1:N, t = 1:T) %>% 
+.[, ind_fe := runif(1), by = id] %>% 
+.[, time_fe := runif(1), by = t] %>% 
+.[, x := runif(1) + ind_fe, by = .(id, t)] %>% 
+.[, group := factor(floor((id) / g_N))] %>% 
+.[, beta_group := rnorm(1), by = group] %>% 
+.[, phi_group := rnorm(1), by = group] %>% 
+.[, alpha_group := rnorm(1) / 2, by = group] %>% 
+.[, mu := rnorm(nrow(.))] %>% 
+.[, e_y := 2 + beta_group * x + phi_group * (x - alpha_group)^2] %>% 
+.[, y := e_y + mu]  
+
+
+gam_f <- formula(y ~ s(x, by = group, k = 5, m = 2))
+
+reg_data <- gen_reg_data(data, gam_formula = gam_f)
+
+# cor(reg_data$model_X)
+
+formula_feols <- 
+  names(reg_data$model_X) %>%
+  paste0(., collapse = "+") %>% 
+  paste0("y ~ ", ., " | group + t") %>% 
+  formula()
+
+fe_res <- 
+  feols(
+    formula_feols, 
+    cluster = ~ group, 
+    data = reg_data$full_data
+  )
+
+pred_data <- 
+  reg_data$full_data %>% 
+  .[, y_hat := predict(fe_res, newdata = .)]
+
+ggplot(data) +
+  geom_point(aes(y = y, x = x), size = 0.3) +
+  geom_point(data = pred_data, aes(y = y_hat, x = x), col = "red", size = 0.4) +
+  geom_line(aes(y = e_y, x = x), size = 0.5, color = "blue") +
+  facet_grid(. ~ group)
+
+gen_reg_data <- function(data, gam_formula) {
+  
+  temp_gam_reg <- 
+    gam(
+      gam_formula, 
+      data = data, 
+      method = "REML",
+      fit = FALSE
+    )
+  
+  model_X <- temp_gam_reg$X %>% 
+    data.table() %>% 
+    setnames(names(.), temp_gam_reg$term.names) %>% 
+    .[, `(Intercept)` := NULL] %>% 
+    setnames(names(.), gsub("\\(", "_", names(.))) %>% 
[…]

### 2022-09-03 Codes/1_A_2_effective_rainfall.R

@@ -1,0 +1,80 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(here)
+library(ggplot2)
+library(sf)
+library(broom)
+library(raster)
+library(RpackageTM)
+library(mgcv)
+library(fixest)
+library(gratia)
+library(data.table)
+library(lfe)
+library(fixest)
+library(stargazer)
+
+source(here("GitControlled/Codes/1_0_fcn_reg.R"))
+
+#/*=================================================*/
+#' # Prepare data
+#/*=================================================*/
+gm_data <- readRDS(here("Shared/Data/ProcessedData", "gridMET.rds"))
+
+d_data <- gm_data %>% 
+  .[, d_balance := et0 - prcp] %>% 
+  .[, too_low := fifelse(d_balance < -50, 1, 0)] %>% 
+  .[, too_low_num := sum(too_low), by = .(sc_code, year(date))] %>% 
+  .[too_low_num == 0 , ]
+
+d_smooth_data <- gen_smooth_data(d_data, formula(sc_code ~ s(d_balance, k = 7)))
+
+saveRDS(d_smooth_data, "Shared/Data/ProcessedData/d_smooth_weather.rds")
+
+s_vars <- names(d_smooth_data$model_X)
+
+ds_data <- d_smooth_data$full_data[, lapply(.SD, sum), by = .(sc_code, year(date)), .SDcols = s_vars]
+
+
+temp_data <- 
+  copy(reg_data)[crop_type == "corn", ] %>% 
+  .$reg_data_y %>% 
+  .[[1]] %>% 
+  .[ir == "nir", ] %>% 
+  ds_data[., on = c("sc_code", "year")]
+
+ds_formula <- 
+  paste0(
+    "yield ~ ",
+    paste0(s_vars, collapse = " + "),
+    "| sc_code + year"
+  ) %>% 
+  formula()
+
+ds_fe <- feols(ds_formula, data = temp_data)   
+
+pred_data <- 
+  d_smooth_data$full_data %>% 
+  .[order(d_balance), ] %>% 
+  .[seq(1, nrow(.), length = 1000), ] %>% 
+  .[, year := 2009] %>% 
[…]

### 2022-09-03 Codes/1_A_power_of_irrigation.R

@@ -1,0 +1,133 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(here)
+library(ggplot2)
+library(sf)
+library(broom)
+library(raster)
+library(RpackageTM)
+library(mgcv)
+library(fixest)
+library(gratia)
+library(data.table)
+library(lfe)
+library(fixest)
+library(scales)
+library(stargazer)
+
+source(here("GitControlled/Codes/Functions/1_0_functions.R"))
+
+#/*=================================================*/
+#' # Prepare data
+#/*=================================================*/
+
+#+ read data
+final_data <-
+  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  readRDS()
+
+# reg_data$reg_data_y[[1]][, .(balance, et0, gdd, edd)] %>% cor()
+
+#' Data used for regression analysis
+reg_data <-
+  prepare_reg_data(
+    data = final_data,
+    sat_thld = 9, # saturated thickness has to be at least 12 meters
+    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
+    balance_thld = c(-1200, 200),
+    state_ls =
+      list(
+        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
+        soy = c("KS", "NE", "TX")
+      )
+  ) %>%
+  data.table()
+
+data_y <- reg_data[crop_type == "corn", reg_data_y][[1]]
+
+within_data <- data_y[, lapply(.SD, function(x) x - mean(x)), by = .(sc_code, sat_cat == "dryland"), .SDcols = c("prcp", "et0", "gdd", "edd", "balance", "yield")]
+
+
+ir_data_y <- data_y[sat_cat != "dryland", ]
+
+semi_formula <-
+  formula(
+    yield
+    ~ s(balance, k = 4, m = 2)
+    + s(gdd, k = 4, m = 2)
+    + s(edd, k = 4, m = 2)
+  )
+
+semi_res <-
+  semi_ols(
+    semi_formula = semi_formula,
[…]

### 2022-09-03 Codes/1_A_regression_analysis.R

@@ -1,0 +1,495 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(here)
+library(ggplot2)
+library(sf)
+library(broom)
+library(raster)
+# library(RpackageTM)
+library(mgcv)
+library(fixest)
+library(gratia)
+library(data.table)
+library(scales)
+
+source(here("GitControlled/Codes/Functions/1_0_functions.R"))
+
+# /*=================================================*/
+#' # Prepare data
+# /*=================================================*/
+
+#+ read data
+final_data <-
+  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  readRDS()
+
+# reg_data$reg_data_y[[1]][, .(balance, et0, gdd, edd)] %>% cor()
+
+#' Data used for regression analysis
+reg_data <-
+  prepare_reg_data(
+    data = final_data,
+    sat_thld = 9, # saturated thickness has to be at least 12 meters
+    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
+    balance_thld = c(-1200, 200),
+    state_ls =
+      list(
+        # corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
+        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
+        soy = c("KS", "NE", "TX")
+      )
+  ) %>%
+  data.table()
+
+# ggplot(data = data_y) +
+#   geom_point(aes(y = yield, x = balance, color = sat_cat_i)) +
+#   geom_smooth(aes(y = yield, x = balance, color = sat_cat_i))
+
+# ggplot(data = data_y) +
+#   geom_point(aes(y = yield, x = spei, color = sat_cat_i)) +
+#   geom_smooth(aes(y = yield, x = spei, color = sat_cat_i))
+
+# /*=================================================*/
[…]

### 2022-09-03 Codes/1_regression_analysis.R

@@ -1,0 +1,89 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(here)
+library(ggplot2)
+library(sf)
+library(broom)
+library(raster)
+library(RpackageTM)
+library(mgcv)
+library(fixest)
+library(gratia)
+library(data.table)
+library(lfe)
+library(fixest)
+library(scales)
+
+source(here("GitControlled/Codes/Functions/1_0_functions.R"))
+
+# /*=================================================*/
+#' # Prepare data
+# /*=================================================*/
+
+#+ read data
+final_data <-
+  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  readRDS()
+
+# reg_data$reg_data_y[[1]][, .(balance, et0, gdd, edd)] %>% cor()
+
+#' Data used for regression analysis
+reg_data <-
+  prepare_reg_data(
+    data = final_data,
+    sat_thld = 9, # saturated thickness has to be at least 12 meters
+    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
+    balance_thld = c(-1200, 200),
+    state_ls =
+      list(
+        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
+        soy = c("KS", "NE", "TX")
+      )
+  ) %>%
+  data.table()
+
+# ggplot(data = data_y) +
+#   geom_point(aes(y = yield, x = balance, color = sat_cat_i)) +
+#   geom_smooth(aes(y = yield, x = balance, color = sat_cat_i))
+
+# ggplot(data = data_y) +
+#   geom_point(aes(y = yield, x = spei, color = sat_cat_i)) +
+#   geom_smooth(aes(y = yield, x = spei, color = sat_cat_i))
+
+# /*=================================================*/
+#' # Regression
+# /*=================================================*/
+
+reg_par <-
+  CJ(
+    crop = c("corn", "soy"),
+    voi = c("balance", "spei"),
+    model = c("semi", "quad")
+  )
+
+all_res <-
+  lapply(
+    1:nrow(reg_par),
+    function(x) {
+      run_analysis(
+        crop = reg_par[x, crop],
+        voi = reg_par[x, voi],
[…]

### 2022-09-03 Codes/2_A1_simulation_analysis.R

@@ -1,0 +1,440 @@
+######################################
+# Simulation based on the statistical analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+#=== load packages ===#
+library(tidyverse)
+library(RpackageTM)
+library(parallel)
+library(here)
+library(ggplot2)
+library(sf)
+library(raster)
+library(mgcv)
+library(gratia)
+library(data.table)
+library(lfe)
+library(stargazer)
+
+#=== Regression results ===#
+# created in 1_regression_analysis.R
+reg_results <- readRDS("Shared/Results/all_res.rds")
+
+#=== source functions ===#
+source(here("GitControlled/Codes/functions.R"))
+
+#/*----------------------------------*/
+#' ## Set up parameters
+#/*----------------------------------*/
+cl_model_ls <-
+  c(
+    "GFDL-ESM2M",
+    "HadGEM2-ES365",
+    "IPSL-CM5A-LR",
+    "MIROC5"
+  )
+
+rcp_ls <- c(45, 85)
+
+par_data_base <-
+  CJ(
+    crop = reg_results$crop %>% unique(),
+    voi = reg_results$voi %>% unique(),
+    model = reg_results$model %>% unique(),
+    cl_model = cl_model_ls,
+    rcp = rcp_ls
+  )
+
+#/*=================================================*/
+#' # Simulation Case 1: Multiple hypothetical counties
+#/*=================================================*/
+#' Analysis of what happens if the current farmers suddenly face the climate that they
+#' are expected to see in the future now on average.
+#'
+#' + observation unit: hypothetical farmers representative of each county
+#' + share of irrigated acres: fixed at the latest observed value for each county
+#' + saturated thickness: fixed
+#' 
+ 
+cl_sim_indiv_counties <- function(i) {
+
+  print(i)
+
+  # /*+++++++++++++++++++++++++++++++++++
+  #' # Prepare data 
+  # /*+++++++++++++++++++++++++++++++++++
+  par_data <-
+    par_data_base[i, ] %>%
[…]

### 2022-09-03 Codes/2_figures_tables.R

@@ -1,0 +1,298 @@
+## ----r setup, include=FALSE---------------------------------------------------
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = TRUE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi = 400,
+  fig.width = 7.5,
+  fig.height = 5,
+  out.width = "750px",
+  out.height = "500px"
+)
+
+
+## ----r------------------------------------------------------------------------
+#--- load pacakges ---#
+library(tidyverse)
+library(here)
+library(ggplot2)
+library(sf)
+library(broom)
+library(raster)
+library(mgcv)
+library(fixest)
+library(gratia)
+library(modelsummary)
+library(data.table)
+library(scales)
+
+#--- source functions ---#
+source(here("Codes/Functions/1_0_functions.R"))
+
+
+## ----r------------------------------------------------------------------------
+raw_data_both <-
+  rbind(data_corn, data_soy) %>%
+  .[, dry_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")]
+
+
+## ----r------------------------------------------------------------------------
+(
+  sum_table <-
+    modelsummary::datasummary(
+      type * (yield + balance) ~ dry_or_not * (Mean + SD),
+      data = raw_data_both,
+      output = "kableExtra"
+      # output = "latex"
+    )
+)
+
+
+## ----r------------------------------------------------------------------------
+(
+  g_wb_box <-
+    rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
+    ggplot(data = .) +
+    geom_boxplot(aes(y = -balance, x = factor(year), fill = type)) +
+    xlab("Year") +
+    ylab("Water Deficit (mm)") +
+    scale_fill_discrete(name = "") +
+    theme_bw() +
+    theme(
+      legend.position = "bottom",
+      axis.text.x = element_text(angle = 90)
+    )
+)
+
+
+## ----r------------------------------------------------------------------------
+g_yield_hist <-
[…]

### 2022-09-03 Codes/3_2_make_figure_misc.R

@@ -1,0 +1,454 @@
+#' ---
+#' title: "Generate figures (climate, saturated thickness)"
+#' author: "Taro Mieno"
+#' output:
+#'   html_document:
+#'     number_sections: yes
+#'     theme: flatly
+#'     highlight: zenburn
+#'     toc_float: yes
+#'     toc: yes
+#'     toc_depth: 3
+#' geometry: margin=1in
+#' ---
+
+#+ setup, include = FALSE
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = TRUE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi = 400,
+  fig.width = 7.5,
+  fig.height = 5,
+  out.width = "750px",
+  out.height = "500px"
+)
+
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+# === packages ===#
+library(downloader)
+library(terra)
+library(lubridate)
+library(tidyverse)
+library(tidycensus)
+library(data.table)
+library(ncdf4)
+library(sf)
+library(raster)
+library(parallel)
+library(tmap)
+library(here)
+library(geosphere)
+library(exactextractr)
+
+# === source functions ===#
+source(here("GitControlled/Codes/functions.R"))
+
+# === Regression results ===#
+# created in 1_regression_analysis.R
+reg_results <-
+  here("Shared/Results", "reg_results.rds") %>%
+  readRDS() %>%
+  dplyr::select(crop_type, states, data, share_data, model_name) %>%
+  unique(by = "crop_type")
+
+# /*=================================================*/
+#' # Prepare Data
+# /*=================================================*/
+#+ echo = F
+
+# === sf for yield regression ===#
+sc_y_sf <-
+  reg_results %>%
+  dplyr::select(crop_type, data) %>%
+  unique(by = "crop_type") %>%
+  unnest() %>%
+  data.table() %>%
+  .[, .(sc_code, state, geometry, crop_type)] %>%
+  unique(by = c("sc_code", "crop_type"))
+
+# === sf for ir_share regression ===#
+sc_is_sf <-
+  reg_results %>%
+  dplyr::select(crop_type, share_data) %>%
+  unnest() %>%
+  data.table() %>%
[…]

### 2022-09-03 Codes/3_appendix.R

@@ -1,0 +1,210 @@
+## ----r------------------------------------------------------------------------
+#--- define the semi-parametric portion of the model ---#
+semi_formula <- formula(yield ~ s(spei, by = sat_cat, k = 3, m = 2))
+
+#--- semi-parametric FE wih fixest ---#
+semi_res <-
+  semi_ols(
+    semi_formula = semi_formula,
+    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
+    # fe = "",
+    cluster = "sc_dry",
+    data = data_corn
+  )
+
+#--- main regression results ---#
+y_res_int <- semi_res$fe_res
+
+r2_corn_spei <- r2(y_res_int, type = "r2")
+
+
+## ----r, eval = FALSE----------------------------------------------------------
+modelsummary::msummary(y_res_int, stars = TRUE)
+
+
+## ----r------------------------------------------------------------------------
+gam_y_res <- semi_res$gam_res
+
+#--- raw data for prediction ---#
+data_for_pred_corn <-
+  CJ(
+    spei = seq(min(data_corn$spei), max(data_corn$spei), length = 50),
+    sat_cat = unique(data_corn$sat_cat)
+  ) %>%
+  .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
+  .[is.na(sat), sat := 0] %>%
+  # === can be any sc_code (need to shift yield for "average" county) ===#
+  .[, sc_code := data_corn$sc_code[20]] %>%
+  .[, sc_dry := data_corn$sc_dry[20]] %>%
+  # === can be any year (need to shift yield for "average" year) ===#
+  .[, year := 2009] %>%
+  # === fake yield data (necessary to take advantage of predict.gam) ===#
+  .[, yield := 0] %>%
+  .[, type := "pred"] %>%
+  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]
+
+#--- create bases  ---#
+yhat_data_corn <-
+  gen_smooth_data(
+    data = data_for_pred_corn,
+    gam_res = gam_y_res
+  ) %>%
+  .$data
+
+y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)
+
+yhat_data_corn[, `:=`(
+  y_hat = y_hat_with_se$fit,
+  y_hat_se = y_hat_with_se$se.fit
[…]

### 2022-09-03 Codes/ExploratoryAnalysis/1_exploratory_analysis.R

@@ -1,0 +1,332 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(ggplot2)
+library(sf)
+library(raster)
+library(mgcv)
+library(data.table)
+library(lfe)
+library(stargazer)
+
+#+ read data
+
+final_data <- readRDS("Data/ProcessedData/final_data.rds") %>% 
+	.[sat == 0 & ir == "ir", sat := sat + 0.1] %>% 
+	.[, year_from_base := year - min(year)] %>% 
+	.[!is.na(sat), ] %>% 
+	.[, balance := balance / 1000]
+
+#+ define saturated thickness break
+
+sat_breaks <- 
+quantile(
+	final_data[ir == "ir", sat], 
+	prob = c(25, 50, 75, 100) / 100, 
+	na.rm = TRUE
+)
+
+#+ Create saturated thickness dummies
+
+reg_data <- 
+final_data[, 
+	sat_cat := cut(
+		sat, 
+	  breaks = c(0, 0.01, sat_breaks),
+	  include.lowest = TRUE
+	)  
+] %>% 
+.[, sat_cat_txt := paste0("sat_", as.numeric(sat_cat))] %>% 
+cbind(
+	., 
+	model.matrix(~ sat_cat_txt, data = .) %>% 
+	.[, -1] %>% 
+	data.table() %>% 
+	setnames(names(.), paste0("sat_d_", ((1:ncol(.)) + 1)))
+) %>% 
+.[, `:=`(
+		w_balance = balance - mean(balance),
+		w_yield = yield - mean(yield)
+	), 
+	by = .(sc_code, crop)
+] %>%  
+.[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)]
+
+# corn_data <- reg_data[crop == "corn", ]
+# soy_data <- reg_data[crop == "soy", ]
+
+# corn_data[state == "TX" & ir == "ir", ]
+# soy_data[state == "TX" & ir == "ir", ]
+
+
+#/*=================================================*/
+#' # Exploratory analysis
[…]

### 2022-09-03 Codes/ExploratoryAnalysis/analysis_explore.R

@@ -1,0 +1,194 @@
+######################################
+# Regression Analysis
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+
+#+ load packages
+
+library(tidyverse)
+library(ggplot2)
+library(sf)
+library(raster)
+library(mgcv)
+library(data.table)
+library(lfe)
+library(stargazer)
+
+#+ read data
+
+final_data <- readRDS("Data/ProcessedData/final_data.rds") %>% 
+	.[sat == 0 & ir == "ir", sat := sat + 0.1] %>% 
+	.[, year_from_base := year - min(year)] %>% 
+	.[!is.na(sat), ] %>% 
+	.[, balance := balance / 1000] %>% 
+	.[, num_obs := .N, by = sc_code] %>% 
+	.[num_obs >= 10, ]
+
+#+ define saturated thickness break
+
+sat_breaks <- 
+quantile(
+	final_data[ir == "ir", sat], 
+	prob = c(25, 50, 75, 100) / 100, 
+	na.rm = TRUE
+)
+
+#+ Create saturated thickness dummies
+
+reg_data <- 
+final_data[, 
+	sat_cat := cut(
+		sat, 
+	  breaks = c(0, 0.01, sat_breaks),
+	  include.lowest = TRUE
+	)  
+] %>% 
+.[, sat_cat_txt := paste0("sat_", as.numeric(sat_cat))] %>% 
+cbind(
+	., 
+	model.matrix(~ sat_cat_txt, data = .) %>% 
+	.[, -1] %>% 
+	data.table() %>% 
+	setnames(names(.), paste0("sat_d_", ((1:ncol(.)) + 1)))
+) %>% 
+.[, 
+	`:=`(
+	w_balance = balance - mean(balance),
+		w_yield = yield - mean(yield)
+	), 
+	by = .(sc_code, crop)
+] %>%  
+.[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)]
+
+corn_states <- c("IL", "IA", "CO", "NE", "IN", "KS", "MS")
+corn_data <- reg_data[crop == "corn" & state %in% corn_states, ]
+soy_data <- reg_data[crop == "soy", ]
+
+#/*=================================================*/
+#' # Regression Analysis (GAM)
[…]

### 2022-09-03 Codes/Functions/temp_1_0_functions_test.R

@@ -1,0 +1,550 @@
+
+prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, state_ls){
+
+  #=== saturated thickness threshold in meter ===#
+  sat_thld <- measurements::conv_unit(sat_thld_m, "m", "ft")
+
+  reg_data <-
+    data.table(
+      crop_type = c("corn", "soy"),
+      states = list(state_ls[["corn"]], state_ls[["soy"]])
+    ) %>%
+    rowwise() %>%
+    mutate(base_data = list(
+      data %>%
+      #=== lower bound ===#
+      .[balance >= balance_thld[1], ] %>%
+      #=== upper bound ===#
+      .[balance <= balance_thld[2], ]
+    )) %>%
+    #=== select observations for each crop type ===#
+    mutate(data = list(
+      base_data[crop == crop_type & state %in% states, ]
+    )) %>%
+    mutate(data = list(
+      #=== share of irrigated/dryland acres ===#
+      data[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)] %>%
+      #=== filter out counties with too low saturated thickness ===#
+      # 0: dryland
+      # If irrigated, sat is set to be at least 0.01.
+      # See 0_6_merge_datasets.R
+      .[sat >= sat_thld | sat == 0, ]
+    )) %>%
+    #=== sat breaks for grouping ===#
+    #' dryland production will have NA
+    mutate(sat_breaks = list(
+      quantile(
+        base_data[ir == "ir" & sat >= sat_thld, sat],
+        # prob = c(0, 25, 50, 75, 100) / 100,
+        prob = c(0, 20, 40, 60, 80, 100) / 100,
+        na.rm = TRUE
+      )
+    )) %>%
+    #=== prepare data for regression by crop type ===#
+    mutate(reg_data_y = list(
+      data %>%
+      #=== define saturated thickness category variable ===#
+      .[,
+        sat_cat := cut(
+          sat,
+          breaks = sat_breaks,
+          include.lowest = TRUE
+        )
+      ] %>%
+      .[is.na(sat_cat), sat_cat := "dryland"] %>%
+      .[, sat_cat_i := factor(paste0("sc_", as.numeric(factor(sat_cat))))]
+    )) %>%
+    #=== prepare irrigation share data ===#
+    mutate(base_data_is = list(
[…]

### 2022-09-03 Codes/Outdated_0_2_quickstat_acres.R

@@ -1,0 +1,261 @@
+######################################
+# Get crop yield and acreage data: Corn and Soybeans
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+
+library(tidyverse)
+library(data.table)
+library(terra)
+library(raster)
+library(stars)
+library(here)
+library(sf)
+library(tidyUSDA)
+
+
+#/*=================================================*/
+#' # Prep
+#/*=================================================*/
+base_counties <- readRDS("./Data/ProcessedData/base_counties.rds")
+
+#' get the list of states
+all_states <- base_counties$state %>%
+  unique() %>% 
+  toupper() 
+
+#' get a list of data items
+tidyUSDA::allDataItem %>%
+  grep(pattern = "WHEAT", ., value = TRUE) %>%
+  grep(pattern = "YIELD", ., value = TRUE) %>%
+  grep(pattern = "IRRIGATED", ., value = TRUE)
+
+tidyUSDA::allDataItem %>%
+  grep(pattern = "COTTON", ., value = TRUE) %>%
+  grep(pattern = "YIELD", ., value = TRUE) %>%
+  grep(pattern = "IRRIGATED", ., value = TRUE)
+
+#' define a function to get QuickStat data
+get_quickstat_data <- function(year, data_item) {
+  ir_corn_yield <- getQuickstat(
+    key = "8FBF8552-57BD-38CB-9C16-E7008C5B8D7D",
+    program = "SURVEY",
+    data_item = data_item,
+    geographic_level = "COUNTY",
+    state = all_states,
+    year = year,
+    geometry = TRUE
+  ) %>%
+    #--- keep only some of the variables ---#
+    dplyr::select(year, NAME, county_code, state_fips_code, short_desc, Value) %>%
+    rename("state_code" = "state_fips_code")
+
+  return(ir_corn_yield)
+}
+
+#' Step 1: Major Crops (acre harvested) in the base states
+ item_list0 <- c(
[…]

### 2022-09-03 Codes/Outdated_0_5_2_get_daymet_et0.R

@@ -1,0 +1,175 @@
+######################################
+# Get Daymet weather data and calculate ET0
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+library(daymetr)
+
+#=== centroid-county data ===#
+centroid_to_county <- readRDS("./Data/ProcessedData/centroid_to_county.rds")
+
+#=== load the elevation data ===#
+elevation_data <- readRDS(here("Data/DEM/elevation.rds")) %>% 
+  .[, .(centroid_id, elev)]
+
+#=== merge ===#
+centroids <- left_join(
+  centroid_to_county, 
+  elevation_data,
+  by = "centroid_id"
+)
+
+# === ASCE_PenmanMonteith parameters ===#
+# define constant
+albedo <- 0.23 # Albedo for grass reference crop (Allen et al. 1998)
+G <- 0 # Assume soil heat flux is negligible
+sbc <- 4.903 * 10^(-9) # Stefan-Boltzmann constant (MJ/K^4/m2/day)
+Gs <- 0.0820 # Solar constant (MJ/m2/min)
+Cn <- 900 # Default for daily calculations for short-reference grass crop
+Cd <- 0.34 # Default for daily calculations for short-reference grass crop
+Wspd <- 3 # wind speed
+
+# /*----------------------------------*/
+#' ## define a function that download daymet and calculate et0
+# /*----------------------------------*/
+
+# i <- 1
+# get_weather(1)
+
+get_weather <- function(i) {
+
+  print(i)
+
+  temp_lat <- centroids[i, ] %>% pull(Y)
+  temp_lon <- centroids[i, ] %>% pull(X)
+  temp_site <- centroids[i, ] %>% pull(centroid_id)
+  temp_elev <- centroids[i, ] %>% pull(elev)
+
+  temp_daymet <-  
+  tryCatch(
+    {
+      download_daymet(
+        lat = temp_lat,
+        lon = temp_lon,
+        start = 1990,
+        end = 2018,
+        internal = TRUE
+      )  
+    }, 
+    error = function(cond){
+      return(NULL)
+    }
+  )
+
+  if (!is.null(temp_daymet)) {
+
[…]

### 2022-09-03 Codes/Outdated_0_5_3_summarize_weather.R

@@ -1,0 +1,133 @@
+######################################
+# Summarize weather data
+######################################
+# written by Name on Date
+# objectives:
+# 1.
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+# === uploading the packages ===#
+library(SPEI)
+library(tidyverse)
+library(data.table)
+library(here)
+library(lubridate)
+
+#=== centroid-county ===#
+centroid_to_county <- readRDS(here("Data/ProcessedData/centroid_to_county.rds")) %>% 
+  data.table() %>% 
+  .[, .(centroid_id, sc_code)]
+
+#=== daily weather data ===#
+daily_weather <- readRDS(here("Data/ProcessedData/weather.rds")) %>% 
+  .[, year := year(date)]
+
+#/*=================================================*/
+#' # weekly balance data
+#/*=================================================*/
+
+#=== weekly balance ===#
+balance_data <- daily_weather %>% 
+  .[, .SD[!any(is.na(et0)), ], by = centroid_id] %>% 
+  .[, .SD[!any(is.na(prcp)), ], by = centroid_id] %>% 
+  .[, year := year(date)] %>% 
+  .[, balance := et0 - prcp] %>% 
+  .[, base_date := min(date), by = .(centroid_id, year)] %>% 
+  .[, date_dif := date - base_date] %>% 
+  .[, week := as.numeric(date_dif) %/% 7] %>% 
+  .[, .(balance = sum(balance), num_obs = .N), by = .(centroid_id, year, week)] %>% 
+  centroid_to_county[., on = "centroid_id"] %>% 
+  .[num_obs == 7, ] %>% 
+  .[, .(balance = mean(balance)), by = .(sc_code, year, week)]
+
+#=== break balance values into 10 groups ===#
+balance_data[, balance_cat := cut(
+  balance,
+  breaks = quantile(balance, prob = 0:10/10),
+  include.lowest = TRUE 
+)] %>% 
+.[, bal_txt := paste0("b_", as.numeric(balance_cat))] 
+
+balance_data[, .(bal_txt, balance_cat)] %>% 
+  unique(by = "bal_txt") %>% 
+  saveRDS(here("Data/ProcessedData/balance_txt_meaning.rds"))
+
+#=== weekly count of balance bins ===#
+balance_w <- balance_data[, .(num_week = .N), by = .(bal_txt, sc_code, year)] %>% 
[…]

### 2022-09-03 Codes/Outdated_0_5_get_weather.R

@@ -1,0 +1,75 @@
+library(tidyverse)
+library(daymetr)
+library(lubridate)
+library(data.table)
+library(tmap)
+library(parallel)
+library(here)
+library(lwgeom)
+library(tigris)
+library(sf)
+
+
+# /*=================================================*/
+#' # Preparation
+# /*=================================================*/
+#=== hpa counties ===#
+base_counties_sf <- readRDS(here("Data/ProcessedData/base_counties.rds")) %>%
+  st_as_sf()
+
+# === Create grids and their centroids ===#
+grids <- st_make_grid(base_counties_sf, n = c(150, 150)) %>%
+  st_centroid() %>%
+  st_as_sf() %>%
+  mutate(centroid_id = 1:n())
+
+# tm_shape(base_counties_sf) +
+#   tm_borders() +
+# tm_shape(centroid_to_county) +
+#   tm_dots()
+
+#=== US counties ===#
+# US_county <- counties(cb = TRUE) %>%
+#   st_as_sf() %>%
+#   st_transform(4326) 
+
+# === find centroid to county mapping ===#
+centroid_to_county <- st_intersection(grids, base_counties_sf) %>%
+  dplyr::select(centroid_id, sc_code) %>%
+  st_transform(4326) %>%
+  cbind(., st_coordinates(.))
+
+saveRDS(centroid_to_county, "./Data/ProcessedData/centroid_to_county.rds")
+
+#/*=================================================*/
+#' # Get elevation data
+#/*=================================================*/
+#' in: "./Data/ProcessedData/centroid_to_county.rds"
+#' out: "Data/DEM/elevation.rds"
+
+if (!file.exists(here("Data/DEM/elevation.rds"))) {
+  source(here("Codes", "0_5_1_elevation.R"))
+}
+
+#/*=================================================*/
+#' # Get Daymet weather data and calculate ET0
+#/*=================================================*/
+#' in: "Data/DEM/elevation.rds"
+#' out: "Data/ProcessedData/weather.rds"
+
+if (!file.exists(here("Data/ProcessedData/weather.rds"))) {
+  source(here("Codes", "0_5_2_get_daymet_et0.R"))
+} 
+
+#/*=================================================*/
+#' # Summarize the weather data
+#/*=================================================*/
+#' in: "Data/ProcessedData/weather.rds"
+#' in: "Data/ProcessedData/centroid_to_county.rds"
[…]

### 2022-09-03 Codes/functions.R

@@ -1,0 +1,602 @@
+#/*=================================================*/
+#' # Download and process gridMET data
+#/*=================================================*/
+#' Used in 0_5_get_weather.R
+#' Objective: download gridMET data
+
+get_grid_MET <- function(var_name, year) {
+
+  target_url <-
+    paste0(
+      "http://www.northwestknowledge.net/metdata/data/",
+      var_name, "_", year,
+      ".nc"
+    )
+
+  file_name <-
+    paste0(
+      "/Volumes/Elements/Database/gridMET-historical/",
+      var_name, "_", year,
+      ".nc"
+    )
+
+  if (!file.exists(file_name)) {
+    downloader::download(
+      url = target_url,
+      destfile = file_name,
+      mode = 'wb'
+    )
+  }
+
+}
+
+#/*=================================================*/
+#' # Calculate ET0 from gridMET data
+#/*=================================================*/
+#' Used in 0_5_get_weather.R
+
+get_et0_county_year <- function(county, year) {
+
+  file_name <-
+    paste0(
+      "Shared/Data/ProcessedData/gMET-county/gMET-",
+      county$sc_code, "-", year,
+      ".rds"
+    )
+
+  if (!file.exists(file_name)) {
+    #--------------------------
+    # collect weather variables in a single data.table
+    #--------------------------
+    temp <-
+      lapply(
+        var_ls,
+        function(x) proess_gm_county(x, year, county)
+      ) %>%
+      rbindlist() %>%
+      dcast(
+        rowid + coverage_fraction + cell + x + y + date ~ var,
+        value.var = "value"
+      ) %>%
+      .[, Jday := as.numeric(date - min(date) + 1)] %>%
+      .[month(date) >= 4, ] %>%
+      .[month(date) <= 9, ] %>%
+      #=== day length in second ===#
+      .[, dayl := daylength(y, date) * 3600] %>%
+      setnames(
+        c("tmmx", "tmmn"),
+        c("tmax", "tmin"),
+      ) %>%
+      #=== Kelvin to Celsius ===#
+      .[, `:=`(
+        tmin = tmin - 273.15,
+        tmax = tmax - 273.15
+      )]
+
+    #--------------------------
+    # get elevation data
[…]

### 2022-09-03 Codes/get_spei.R

@@ -1,0 +1,48 @@
+#/*=================================================*/
+#' # Calculate SPEI
+#/*=================================================*/
+ET_data <- readRDS(here("Data", "weather.rds"))
+
+
+
+#/*----------------------------------*/
+#' ## Calculate SPEI
+#/*----------------------------------*/
+monthly_balance <- ET_data %>%
+.[, 
+  .(m_et = sum(ET0), m_prcp = sum(prcp)), 
+  by = .(centroid_id, state_code, county_code, year, month)
+] %>% 
+.[, m_balance := m_prcp - m_et] %>% 
+.[order(centroid_id, year, month), ] %>% 
+nest_by(centroid_id) %>% 
+mutate(data = list(
+  mutate(data, spei_6 = spei(data[, "m_balance"], 6)$fitted %>% as.vector())  
+)) %>% 
+mutate(data = list(
+  mutate(data, spei_3 = spei(data[, "m_balance"], 3)$fitted %>% as.vector())  
+)) %>% 
+unnest() %>% 
+data.table()
+
+#/*----------------------------------*/
+#' ## Combine SPEI 3 and 6
+#/*----------------------------------*/
+#=== Sep covers 4 ~ 9 ===#
+spei_6_Sep <- monthly_balance[month == 9, .(centroid_id, state_code, county_code, year, month, spei_6)] %>% 
+  .[, month := NULL]
+
+#=== June and Sep cover 4 ~ 6 and 7 ~ 9, respectively ===#
+spei_3_June_Sep <- monthly_balance[month %in% c(6, 9), .(centroid_id, state_code, county_code, year, month, spei_3)] %>% 
+  dcast(centroid_id + state_code + county_code + year ~ paste0("spei_3_", month), value.var = "spei_3")
+
+#=== Summarize by state and county codes ===#
+spei_data <- spei_3_June_Sep[spei_6_Sep, on = c("centroid_id", "year")] %>% 
+  .[, 
+    .(
+      spei_3_6 = mean(spei_3_6, na.rm = TRUE), 
+      spei_3_9 = mean(spei_3_9, na.rm = TRUE), 
+      spei_6 = mean(spei_6, na.rm = TRUE)
+    ), 
+    by = .(state_code, county_code, year)
+  ]  

### 2022-09-03 Codes/import_modules.py

@@ -1,0 +1,24 @@
+import econml
+
+# Main imports
+from econml.orf import DMLOrthoForest, DROrthoForest
+from econml.dml import CausalForestDML
+from econml.sklearn_extensions.linear_model import (
+    WeightedLassoCVWrapper,
+    WeightedLasso,
+    WeightedLassoCV,
+)
+
+from sklearn.ensemble import GradientBoostingRegressor
+from sklearn.multioutput import MultiOutputRegressor
+
+# Helper imports
+import numpy as np
+from itertools import product
+from sklearn.linear_model import (
+    Lasso,
+    LassoCV,
+    LogisticRegression,
+    LogisticRegressionCV,
+)
+import matplotlib.pyplot as plt

### 2022-09-03 Codes/make_figure_misc_explore.R

@@ -1,0 +1,270 @@
+#' ---
+#' title: "Generate figures (climate, saturated thickness)"
+#' author: "Taro Mieno"
+#' output:
+#'   html_document:
+#'     number_sections: yes
+#'     theme: flatly
+#'     highlight: zenburn
+#'     toc_float: yes
+#'     toc: yes
+#'     toc_depth: 3
+#' geometry: margin=1in
+#' ---
+
+#+ setup, include = FALSE
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = TRUE,
+  comment = NA,
+  message = FALSE,
+  warning = FALSE,
+  tidy = FALSE,
+  cache.lazy = FALSE,
+  #--- figure ---#
+  dpi = 400,
+  fig.width = 7.5,
+  fig.height = 5,
+  out.width = "750px",
+  out.height = "500px"
+)
+
+
+#/*=================================================*/
+#' # Preparation
+#/*=================================================*/
+#=== packages ===#
+library(downloader)
+library(terra)
+library(lubridate)
+library(tidyverse)
+library(tidycensus)
+library(data.table)
+library(ncdf4)
+library(sf)
+library(raster)
+library(parallel)
+library(tmap)
+library(here)
+library(geosphere)
+library(exactextractr)
+
+#=== source functions ===#
+source(here("GitControlled/Codes/functions.R"))
+
+#=== Regression results ===#
+# created in 1_regression_analysis.R
+reg_results <- 
+  here("Shared/Results/all_res.rds") %>% 
+  readRDS() %>% 
+  dplyr::select(crop, data_int, ir_share_data) %>% 
+  unique(by = "crop")
+
+#/*=================================================*/
+#' # Prepare Data
+#/*=================================================*/
+#+ echo = F
+
+#/*----------------------------------*/
+#' ## Regression data
+#/*----------------------------------*/
+#=== sf for yield regression ===#
+sc_y_sf <-  
+  reg_results %>% 
+  dplyr::select(crop, data_int) %>% 
+  unique(by = "crop") %>% 
+  unnest() %>% 
+  data.table() %>% 
+  .[, .(sc_code, state, year, geometry, crop, sat_cat)] %>% 
+  unique(by = c("sc_code", "crop"))
+
+#=== sf for ir_share regression ===#
+sc_is_sf <-  
+  reg_results %>% 
[…]

### 2022-09-03 Codes/spin.R

@@ -1,0 +1,24 @@
+library(knitr)
+library(tidyverse)
+library(here)
+library(prettydoc)
+
+here("GitControlled/Codes/0_1_hpa_counties.R") %>% 
+  knitr::spin(. , knit = FALSE) %>% 
+  rmarkdown::render()
+
+here("GitControlled/Codes/0_2_quickstat.R") %>% 
+  knitr::spin(. , knit = FALSE) %>% 
+  rmarkdown::render()
+
+here("GitControlled/Codes/0_3_get_sat.R") %>% 
+  knitr::spin(. , knit = FALSE) %>% 
+  rmarkdown::render()
+
+here("GitControlled/Codes/3_1_make_figures_results.R") %>% 
+  knitr::spin(. , knit = FALSE) %>% 
+  rmarkdown::render()
+
+here("GitControlled/Codes/3_2_make_figure_misc.R") %>% 
+  knitr::spin(. , knit = FALSE) %>% 
+  rmarkdown::render()

### 2023-05-03 Writing/manuscript_DR.Rmd
minor edits

@@
   - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
-  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota]
+  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota, kakim002@umn.edu]
   - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
@@
 
+```{r}
+deficit_2012 <- data_corn[year == 2012, mean(balance)]
+
+#--- corn ---#
+balace_closest <- data_final_corn[which.min(abs(balance - deficit_2012)), balance]
+
+corn_avg_yield_low <- data_final_corn[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
+corn_avg_yield_high <- data_final_corn[balance == balace_closest & sat_cat == "(194,708]", avg_yield]
+
+corn_pct_decline <- (corn_avg_yield_high - corn_avg_yield_low) / corn_avg_yield_high * 100
+
+#--- soy ---#
+balace_closest <- data_final_soy[which.min(abs(balance - deficit_2012)), balance]
+
+soy_avg_yield_low <- data_final_soy[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
+soy_avg_yield_high <- data_final_soy[balance == balace_closest & sat_cat == "(194,708]", avg_yield]
+
+soy_pct_decline <- (soy_avg_yield_high - soy_avg_yield_low) / soy_avg_yield_high * 100
+```
+
 Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated.
@@
 The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher.
-In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively.
+In the year with water deficit of 950mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively.
 These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness.

### 2023-05-03 Codes/Functions/1_0_functions.R

@@ -1,8 +1,11 @@
 
-prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, state_ls) {
+prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
 
   # === saturated thickness threshold in meter ===#
   sat_thld <- measurements::conv_unit(sat_thld_m, "m", "ft")
+
+  # data <- reg_data$data[[1]]
+  # sat_breaks <- reg_data$sat_breaks[[1]]
 
   reg_data <-
     data.table(
@@ -35,9 +38,7 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, stat
     mutate(sat_breaks = list(
       quantile(
         base_data[ir == "ir" & sat >= sat_thld, sat],
-        prob = c(0, 33.3, 66.6, 100) / 100,
-        # prob = c(0, 25, 50, 75, 100) / 100,
-        # prob = c(0, 20, 40, 60, 80, 100) / 100,
+        prob = seq(0, 1, length = sat_cat_num + 1),
         na.rm = TRUE
       )
     )) %>%
@@ -54,17 +55,18 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, stat
           )
         ] %>%
         .[is.na(sat_cat), sat_cat := "dryland"] %>%
-        .[, sat_cat_i := factor(paste0("sc_", as.numeric(factor(sat_cat))))]
+        .[, sat_cat_i := factor(paste0("sc_", as.numeric(factor(sat_cat))))] %>%
+        .[sat_cat == "dryland" | (ir == "ir" & ir_area_ratio >= ir_share_thld), ]
     )) %>%
     # === prepare irrigation share data ===#
-    mutate(base_data_is = list(
+    mutate(base_data_ir_share = list(
       data[ir == "ir" & ir_area_ratio >= ir_share_thld, ]
     )) %>%
     mutate(cl_vars = list(
       c("balance", "days_ab_30", "gdd")
     )) %>%
     mutate(avg_cl_data = list(
-      base_data_is[,
+      base_data_ir_share[,
         lapply(.SD, mean, na.rm = TRUE),
         .SDcols = cl_vars,
         by = sc_code
@@ -72,12 +74,12 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, stat
[…]

### 2023-07-18 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.rmd
revision started

---
output:
  pdf_document:
    keep_tex: true
---
Reviewer #1 (Comments for the Author):

This manuscript assesses the impact of aquifer thickness on agricultural production during drought in the area overlying the High Plains aquifer in the central United States. The authors use crop, meteorological, and aquifer thickness data in regressions to develop relationships among these quantities. Their most important finding is the impact of aquifer thickness on productivity during drought and the ramifications for aquifer conservation efforts. I have read this manuscript over a number of times and found it of considerable interest and worthy of publication in Nature Water. However, I do have a few concerns that I wanted to bring to the attention of the authors. My most important concerns are as follows (not necessarily in order of importance):

\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}

1. Aquifer thickness versus saturated thickness – The authors speak of "aquifer saturated thickness", which is a redundant term. Remember, the definition of an aquifer is a saturated unit – there is no unsaturated aquifer. Please use "aquifer thickness" to avoid propagating this confusing usage into the future.

\textcolor{blue}{All the instances of "saturated thickness" and "aquifer saturated thickness" have been replaced by "aquifer thickness".}

2. Dependence on aquifer thickness – I had a number of questions about the relationships displayed in Figure 1. First, the regression appears to have been done year by year. Are the quantiles recalculated each year to reflect that there may be changes between categories as time goes on? Second, when the authors state the lowest quantile goes from 0 to 33% are they assuming that the minimum thickness of 9 m is the zero point? Third, I assume that the county and year fixed effects are calculated as part of the regression. Are there any insights that can be gained from the interpretation of those quantities? Fourth, the authors need to point out that the same aquifer thickness can result in dramatically different transmissive characteristics. Thus, a relatively thin aquifer interval composed of coarse sands and gravels may be more resilient to drought than a much thicker aquifer dominated by fine sands and silts. Could that be a major player in the error bars in the plots? Please mention this at some point in the manuscript.

- First point: \textcolor{blue}{Quantiles are not recalculated each each. If we do so, then the same quantile would represent different aquifer thickness range every year. When modeling the impact of aquifer thickness, this is not desirable because we are implicitly making assumptions that different ranges of aquifer thickness would have the same impact.}
-  Second point: \textcolor{blue}{Actually, 0% quntile is 12 m. "9m" was repalced with "12m" in the current manuscript.}
-  Third point: \textcolor{blue}{No interesting insights will be gained from them because they are merely shifters without any intractions with aquifer thickness. They tell us how much higher yield in a single county relative to another county on average. But, we do not beleive that is useful information at least in this artcile.}
-  Fourth point: \textcolor{blue}{This is a very important point. We have added texts describing the implications of this fact in page xx.}

3. Conditioned on drought risk exposure – I wasn’t sure what the authors meant by this in the second paragraph of Section 2.2. This must be related to the water deficit but the lack of explanation in the main text or the methods section left me wondering. Please clarify at some point in the manuscript.

\textcolor{blue}{What we meant by "conditiond on drought risk exposure" is "after taking into account the impact of climate variables." In order to avoid confusion, we simply removed it from the sentence in the current manuscript.}

4. Equation 2 – Please define all terms in the equation and provide an explanation of what is meant by "estimated semi-parametrically".

\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained. We eliminated the word semi-parametrically as it is not ncecessary. Generalized additive model that estiamtes the impact of a variable using splines without functional form assumption are sometimes referred to semi-parametric regression. That's what we meant by "estimate semi-parametrically" in our previous manuscript.}

5. Kansas and California efforts – I’ve seen a number of papers describing efforts in Kansas and California to address aquifer depletion via LEMAs (Kansas) and SGMA (California). The 1985-2016 period examined by the authors was before most of the efforts in Kansas got started. My understanding is that the first LEMA was established in 2013 but then no more were set up until after 2016. This manuscript would be more impactful if the authors could speculate on how their relationships might change with greater attention to groundwater conservation. For example, I would think the decrease in yield the authors observe with negative water deficits would disappear when the irrigators are using soil moisture sensors and are focused on water-use efficiency. I would expect changes as well for tipping points.

\textcolor{blue}{Tim?}

## Minor points

+ Section 1, 3rd Paragraph – "Two aquifer pathways exist through aquifer depletion will.." – confusing sentence – please replace "through" with "by which".

\textcolor{blue}{Corrected.}

+ Section 1, 4th Paragraph – "we generate empirical evidence" seems awkward to me. How about replacing with "we present empirical evidence"? Next sentence – "Our analysis advances on previous…" reads better as "Our analysis extends previous…" There are many issues such as these and the previous comment. Please address such issues so that the manuscript is an easier read.

\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}

+ Section 2.1 title – I would delete "marginally" as it is not needed and some will find it confusing.

\textcolor{blue}{Tim, any thoughts on this. I feel like it is still important to say the increase is not much.}

+ Figure 2 - Why do the plots stop at a thickness of about 12 m when you had earlier defined the minimum thickness as 9 m and mention 10 m in the text when discussing the relationships?

\textcolor{blue}{Thank you for catching this. 9m and 10m should have been 12m.}

+ Section 2.3, First Paragraph – please clarify that "average per-area production" is a combination of irrigated and rainfed yields. That would make the manuscript easier to follow.

\textcolor{blue}{We added "(weighted-average of irrigated and rainfed yields, where weights are production area shares)" so that it is clear what average per-area production means.}

+ Figure 5 – I assume that the left plot is corn and the right soybean, but please clarify.

\textcolor{blue}{Thanks for noticing this. Fixed.}

+ Data and Code Availability – Are the data from Dr. Haacker on the Github repository? Please place them there prior to publication so future investigators can access them rather than having to track down Dr. Haacker. The data set is over seven years old so it should be made available according to the principles of open science.

\textcolor{blue}{Yes, we will make the data available online.}

+ Line before Equation (2) – "REF" should be replaced with the actual reference.

\textcolor{blue}{Tim?}

+ Discussion, First Paragraph – "reduction in aquifer depletion" should be "reduction in aquifer thickness".

\textcolor{blue}{Thank you for catching this. Fixed.}

+ Section 4.2, 3rd Paragraph – "the the impact" – remember to spell and grammar check prior to submission.

\textcolor{blue}{Thank you for catching this. Fixed.}

+ Reference in Section 3 – I doubt that "Butler Jr" is the surname, please replace with "Butler". Suffixes apparently aren’t handled well by some of the bibliography generation programs.

\textcolor{blue}{}

+ Please remember to close all parentheses and don't mispair brackets and parentheses. Such stuff can raise questions about the care used in the analysis.

### 2023-07-18 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.tex
revision started

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

Reviewer \#1 (Comments for the Author):

This manuscript assesses the impact of aquifer thickness on agricultural
production during drought in the area overlying the High Plains aquifer
in the central United States. The authors use crop, meteorological, and
aquifer thickness data in regressions to develop relationships among
these quantities. Their most important finding is the impact of aquifer
thickness on productivity during drought and the ramifications for
aquifer conservation efforts. I have read this manuscript over a number
of times and found it of considerable interest and worthy of publication
in Nature Water. However, I do have a few concerns that I wanted to
bring to the attention of the authors. My most important concerns are as
follows (not necessarily in order of importance):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Aquifer thickness versus saturated thickness -- The authors speak of
  ``aquifer saturated thickness'', which is a redundant term. Remember,
  the definition of an aquifer is a saturated unit -- there is no
  unsaturated aquifer. Please use ``aquifer thickness'' to avoid
  propagating this confusing usage into the future.
\end{enumerate}

\textcolor{blue}{ferkjnf}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
  Dependence on aquifer thickness -- I had a number of questions about
  the relationships displayed in Figure 1. First, the regression appears
  to have been done year by year. Are the quantiles recalculated each
  year to reflect that there may be changes between categories as time
  goes on? Second, when the authors state the lowest quantile goes from
  0 to 33\% are they assuming that the minimum thickness of 9 m is the
  zero point? Third, I assume that the county and year fixed effects are
  calculated as part of the regression. Are there any insights that can
  be gained from the interpretation of those quantities? Fourth, the
  authors need to point out that the same aquifer thickness can result
  in dramatically different transmissive characteristics. Thus, a
  relatively thin aquifer interval composed of coarse sands and gravels
  may be more resilient to drought than a much thicker aquifer dominated
  by fine sands and silts. Could that be a major player in the error
  bars in the plots? Please mention this at some point in the
  manuscript.
\item
  Conditioned on drought risk exposure -- I wasn't sure what the authors
  meant by this in the second paragraph of Section 2.2. This must be
  related to the water deficit but the lack of explanation in the main
  text or the methods section left me wondering. Please clarify at some
  point in the manuscript.
\item
  Equation 2 -- Please define all terms in the equation and provide an
  explanation of what is meant by ``estimated semi-parametrically''.
\item
  Kansas and California efforts -- I've seen a number of papers
  describing efforts in Kansas and California to address aquifer
  depletion via LEMAs (Kansas) and SGMA (California). The 1985-2016
  period examined by the authors was before most of the efforts in
  Kansas got started. My understanding is that the first LEMA was
  established in 2013 but then no more were set up until after 2016.
  This manuscript would be more impactful if the authors could speculate
  on how their relationships might change with greater attention to
  groundwater conservation. For example, I would think the decrease in
  yield the authors observe with negative water deficits would disappear
  when the irrigators are using soil moisture sensors and are focused on
  water-use efficiency. I would expect changes as well for tipping
  points.
\end{enumerate}

\hypertarget{minor-points}{%
\subsection{Minor points}\label{minor-points}}

Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through aquifer
depletion will..'' -- confusing sentence -- please replace ``through''
with ``by which''.

Section 1, 4th Paragraph -- ``we generate empirical evidence'' seems
awkward to me. How about replacing with ``we present empirical
evidence''? Next sentence -- ``Our analysis advances on
previous\ldots{}'' reads better as ``Our analysis extends
previous\ldots{}'' There are many issues such as these and the previous
comment. Please address such issues so that the manuscript is an easier
read.

Section 2.1 title -- I would delete ``marginally'' as it is not needed
and some will find it confusing.

Figure 2 - Why do the plots stop at a thickness of about 12 m when you
had earlier defined the minimum thickness as 9 m and mention 10 m in the
text when discussing the relationships?

Section 2.3, First Paragraph -- please clarify that ``average per-area
production'' is a combination of irrigated and rainfed yields. That
would make the manuscript easier to follow.

Figure 5 -- I assume that the left plot is corn and the right soybean,
but please clarify.

Data and Code Availability -- Are the data from Dr.~Haacker on the
Github repository? Please place them there prior to publication so
future investigators can access them rather than having to track down
Dr.~Haacker. The data set is over seven years old so it should be made
available according to the principles of open science.

Line before Equation (2) -- ``REF'' should be replaced with the actual
reference.

Discussion, First Paragraph -- ``reduction in aquifer depletion'' should
be ``reduction in aquifer thickness''.

Section 4.2, 3rd Paragraph -- ``the the impact'' -- remember to spell
and grammar check prior to submission.

Reference in Section 3 -- I doubt that ``Butler Jr'' is the surname,
please replace with ``Butler''. Suffixes apparently aren't handled well
by some of the bibliography generation programs.

Please remember to close all parentheses and don't mispair brackets and
parentheses. Such stuff can raise questions about the care used in the
analysis.

\end{document}

### 2023-07-18 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.rmd
revision started

\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}

This is an interesting paper that asks an important yet answered question in the field of agricultural sustainability - how much is groundwater depletion reducing agricultural productivity in the High Plains region in the United States? While the paper is an important contribution to the literature, there are several issues that need to be addressed prior to acceptance at any journal.

+ There needs to be more description about why aquifer saturation thickness is an appropriate proxy for farmers' ability to irrigate from wells. This is particularly true since this is a broad interest journal, and the mechanism for the main independent variable should be very clear to all readers.

\textcolor{blue}{Tim?}

+ Though I am not trained as a hydrologist, my understanding is that aquifer saturation thickness is a good measure of general water availability in an aquifer but not necessarily how much water is available at a specific location or point. If this is the case, the authors should make clear why this is an appropriate measure to use in their study and potential limitations of using this approach.

\textcolor{blue}{Tim? this is realted to one of the reviewer1's comments}

+ It is unclear to me why county and time fixed effects were not used in every regression. In particular, it does not look like they were included in the regression that modeled irrigated area as a function of saturated thickness and water deficit. Yet, accounting for time-invariant differences across counties is important, especially because this is the regression that showed the largest effect of saturated thickness on a measure of agricultural production (irrigated area). Is there not enough variation within a county through time to pick up the effect of saturated thickness on irrigated area? This should all be made clearer, and a model that better captures the causal relationship between saturated thickness and area irrigated should be used if possible. If it's not possible, the limitations of the current approach should be made very clear and potential issues with inference.

\textcolor{blue}{}

+ There are minor typos (e.g., missing reference listed as REF) that should be corrected prior to resubmission.

\textcolor{blue}{}

### 2023-07-18 Writing/NatureWater/1stRoundRevision/revised_manuscript/cover_letter.tex
revision started

--- 2022-09-03 Writing/manuscript_DR.tex
@@
 
-\title{The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather}
-\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
+\title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
+\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of Mechanical, Aerospace and Civil Engineering, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
 \date{}
@@
 \maketitle
-\begin{abstract}
-Abstract goes here. It should be a single paragraph of around 200 words.
-\end{abstract}
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+Dear Editor,\\
 
-Groundwater-irrigated agriculture supports X\% of crop production worldwide and Y\% in U.S.
-(citations) It is a vital natural resource that brings tremendous economic and social benefits to the society.
-However, groundwater has been depleting rapidly all over the world due to intensive groundwater extraction that far exceeds the natural replenishment rate (citation).
-The importance of irrigation has been consistently growing over the years (Irmak).
+Thank you for considering this manuscript for publication at Nature Water.
+In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events.
+Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer saturated thickness significantly reduces the overall productivity of both corn and soybean during drought events.
+Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion.
+Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.\\
+
+Our work is significant because we quantify the impact of a decline in saturated thickness of groundwater on the resiliency of agricultural production, which has rarely been quantified in the literature before.
+We confirm quantitatively the importance of keeping relatively high saturated thickness for sustainable high-productivity agricultural production.
+While our results are for the US High Plains, the insights we generated here are likely to hold elsewhere.\\
+
+This article does not have any realated manuscripts that are under consideration.
+However, its pre-print has been published at EarthArXiv (\url{https://eartharxiv.org/repository/view/5126/}).
+We have not had any prior discussions with a Nature Water editor about this work.
+\\
+
+Here is the list of potential reviewers that we may recommend.
 
 \begin{itemize}
-\tightlist
-\item
-  some mentionin of food security issues
+    \item James Butler (\url{https://geo.ku.edu/people/james-k-butler-jr})
+    \item Marc Muller (\url{https://engineering.nd.edu/faculty/marc-francois-muller/})
+    \item Sam Zipper (\url{https://www.samzipper.com/authors/samuel-c.-zipper/})
+    \item Christian Klassert (\url{https://www.ufz.de/index.php?en=38767})
+    \item Jen Burney (\url{http://www.jaburney.net/})
 \end{itemize}
 
-The vast majority of studies on the climate change impacts on agricultural productivity focus on dryland production as it is far vulnerable to climate variations than irrigated production is.
-Another strand of literature seeks to understand the impact of climate change on water demand.
-Only a small collection of studies estimated the climate change impacts on irrigated production.
-It first seems that we do not need to understand the impact of climate change on irrigated production because irrigation provide a protection against climate variation.
-This would be true under the condition that farmer can apply desired amount of water without any physical or economic constraints, in which case, farmers can always take advantage of the benefit of irrigation to its full potential.
-Unfortunately, that is not the case for groundwater irrigated agriculture.
-Specifically, the stock of groundwater in the aquifer, measured as saturated thickness, affects the physical and economic constraints under which farmers operate.
-A decline in saturated thickness place stronger constraints on farmers and hurt the profitability of irrigated production.
-
-There are mainly two reasons that saturated thickness affects the ability of aquifer to protect farmers from undesirable production conditions.
-First, as saturated thickness goes down due to groundwater depletion, the economic cost of pumping water increases.
-This affects overall agricultural productivity through two channels: extensive and intensive margins.
-When the cost of groundwater extraction becomes too high, irrigated production is no longer more profitable than dryland production.
-In this case, farmers will switch to irrigated production to dryland production even though groundwater still exists in the aquifer, which leads to significant reduction in crop yield (extensive margin).
-For those who have decided to irrigate, they would use less water as the cost of groundwater pumping is higher, thus lowering crop yield (intensive margin).
-Second and more important, well yield (a measure of how much water can be pumped per unit of time) declines as saturated thickness goes down ().
-This physical constraint has been shown to have far greater impacts on irrigation decisions than pumping cost \citep{mieno2021}.
-Similarly to the rising pumping cost case, declining well yield affects agricultural productivity through two channels.
-Too low a well yield makes producers shift to dryland production.
-For those irrigating, a lower well yield would lead to lower crop yield.
-
-\citet{sampson2019} estimated the impact of saturated thickness on land value.
-
-The vast majority of existing body of studies focus on estimating the impact of climate on dryland production.
-A small amount of studies that look at the impact of climate on irrigated production does not take into account saturated thickness.
-Our study fills this important gap and sheds light on the value of groundwater by recognizing the importance of saturated thickness and quantifying its ability to protect agricultural production from undesirable weather outcomes.
-
-\citep{rad2020effects}
-\citep{foster2014modeling}
-\citep{foster2015analysis}
-
-\hypertarget{materials-and-method}{%
-\section{Materials and Method}\label{materials-and-method}}
-
-In order to understand the impact of saturated thickness on agricultural production, this study examines the two channels through which saturated thickness affects average corn and soybean yields using econometric analysis.
-First, we quantify the impact of water deficit (\(WB\)) during the production season (May through September) on irrigated and dryland yields (\(Y\)) differentiated by the level of saturated thickness (\(ST\)).
-Mathematically, we estimate the expected value of yield conditional on water deficit, saturated thickness, and whether irrigated or not (\(I\)).
-
-\[
-E[Y_{i,t}|WB_{i,t}, ST_{i,t}, I_{i,t}]= f(WB_{i,t}, ST_{i,t}, I_{i,t})
-\]
-
-where \(i\) indicates county and \(t\) indicates year.
-
-\noindent Second, we estimate the expected value of the share of irrigated acres as a function of saturated thickness and average water deficit (averaged over years).
-Historical average water deficit is used in place of concurrent water deficit \(WB_{i,t}\) because the decision of whether to irrigate or not is made based on what the history of crop water demand rather than the concurrent weather that is yet to unfold at the time of decision making.
-
-\[
-E[S_{i,t}|WB_{i,t}, ST_{i,t}]= g(\overline{WB}_i, ST_{i,t}) 
-\]
-
-\noindent Once they are estimated, we can estimate the average crop yield as the share-weighted average of dryland and irrigated yields as follows:
-
-\[
-\hat{h}(WB_t, St_t)= (1 - \hat{g}(\overline{WB}, ST) )\cdot \hat{f}(WB_t, ST_t, I_t = 0)+ \hat{g}(\overline{WB}, ST) \cdot \hat{f}(WB_t, ST_t, I_t = 1)
-\]
-
-\noindent where \(\hat{f}(\cdot)\) and \(\hat{g}(\cdot)\) are estimated version of \(f(\cdot)\) and \(g(\cdot)\), respectively.
-For example, the impact of a change in saturate thickness level, say from \(ST_{before}\) to \(ST_{after}\), on crop yield is then \(\hat{h}(WB_t, St_{before})\) - \(\hat{h}(WB_t, St_{before})\).
-
-\hypertarget{econometric-models}{%
-\subsection{Econometric Models}\label{econometric-models}}
-
-We use the same model specification for corn and soybean yield regression analysis. The estimating equation is written as follows:
-
-\[
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t}
-\]
-
-Individual county and year are indicated by \(i\) and \(t\), respectively.
-In the model, \(y_{i,t}\) is crop yield for county \(i\) in year\(t\).
-\(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: dryland production, observations with saturated thickness falling under the \([0-33.3\%)\)j, \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of saturated thickness.
-The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-\(\sigma_j\) represents the category specific intercept.
-\(WB_{i,t}\) is water deficit defined as total reference evapotranspiration less precipitation between May through September.
-Category-specific impact of water deficit on yield is represented by \(\sum_{k}^K \beta_k\phi_{j,k}(WB_{i,t})\), where \(\phi_{j,k}\) is \(k\)th spline basis function and \(\beta_k\) is its coefficient.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner.
-The model can be interpreted as hierarchical generalized additive model, where each of saturated thickness level category has its own yield response function to drought (cite).
-To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are included.
-Standard error estimation is clustered by county.
-
-The impact of saturated thickness on the share of irrigated acres is estimated using a fractional logit model under the generalize additive model framework in which the impact of saturated thickness, average water deficit, and their interactions are estimated semi-parametrically, allowing for flexible representation of their impacts.
-The estimating equation is as follows:
-
-\[
-log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + h(ST_{i,t}, \overline{WB}_i) + \mu_t + v_{i,t}
-\]
-
-, where \(s_{i,t}\) is the share of irrigated production.
-\(h(ST_{i,t}, \overline{WB}_i)\) consists of basis functions and their coefficients to repsent the the impact of \(ST_{i,t}\), \(\overline{WB}_i\), and their interactions (see Appendix ??
-for further details of how \(h(ST_{i,t}, \overline{WB}_i)\) are mathematically defined).
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
-Figure \ref{fig:geo-focus} shows the geographical coverage of the data for regression analysis for corn and soybean.
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/geo-focus-1}
-
-}
-
-\caption{Geographic focus of the study for corn and soybean}\label{fig:geo-focus}
-\end{figure}
-
-First, county-level irrigated and dryland yield data for corn and soybean for 1985 through 2016 were obtained from the USDA NASS QuickStat.
-Both dryland and irrigated yield observations are used for counties that shares more than 75\% of their area with the Ogallala aquifer.
-Only dryland yield observations are used for counties that do not satisfy the condition.
-We use dryland yield data outside of the Ogallala aquifer to supplement dryland yield data for more accurate estimation of the impact of water deficit for dryland production.
-Daily precipitation and reference ET were obtained from gridMET \citep{Abatzoglou2013}.
-Water deficit (\(WB_{i,t}\)) was then calculated by subtracting total precipitation from total reference evapotranspiration from May through September.
-Saturated thickness data were obtained from Dr.~Erin Haacker \citep{haacker2016water}.
-Observations that have saturated thickness value lower than 12 meter were eliminated as saturated thickness below that is considered not suitable for irrigation (Tim citations here).
-
-Figure \ref{fig:deficit-hist} presents annual water deficit by year for the corn and soybean data.
-Water deficit is highly variable year to year.
-Year 2012 in particular stands out as an extremely dry year in the region.
-We use the year-to-year variations in water deficit to identify its impact on crop yield.
-Figure \ref{fig:yield-hist} compares historical corn and soybean yields for dryland and irrigated production.
-The power of irrigation is clearly evident in this figure.
-First, irrigation provides protection against severe drought years like 2000, 2002, 2012.
-While the dryland yield dives down in those years, irrigated yields is almost unaffected by the severe droughts.
-First key objective of our study is to see if and how much the ability of irrigation is diminished when saturated thickness goes down.
-Figure \ref{fig:sat-dist} presents the distribution of saturated thickness.
-Saturated thickness vary substantially, but most of the sites have saturated thickness of less than 50 meters.
-We exploit the observed variations in saturated thickness to identify its influence on the ability of irrigation to drought.
-Second, irrigated yield is always about 50\% or higher than dryalnd yield irrespective of the degree of water deficit.
-We will investigate the negative impact of a decline in saturated thickness on the percentage of irrigated fields, thus reducing crop yields.
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/deficit-hist-1}
-
-}
-
-\caption{Distribution of water deficit for corn and soybean datasets}\label{fig:deficit-hist}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/yield-hist-1}
-
-}
-
-\caption{Historical dryland and irrigated yields for corn and soybean}\label{fig:yield-hist}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/sat-dist-1}
-
-}
-
-\caption{Distribution of saturated thickness}\label{fig:sat-dist}
-\end{figure}
-
-\hypertarget{results-and-discussions}{%
-\section{Results and Discussions}\label{results-and-discussions}}
-
-\hypertarget{the-impact-of-drought-on-dryland-and-irrigated-yield}{%
-\subsection{The impact of drought on dryland and irrigated yield}\label{the-impact-of-drought-on-dryland-and-irrigated-yield}}
-
-Regression results are presented graphically rather than in tables because individual coefficients estimates on spline basis functions cannot be readily interpreted and also because visualized results are much easier to interpret.
-Figure \ref{fig:yield-response} shows yield response to water deficit for dryland production and groundwater-irrigated production at different levels of saturated thickness categories for corn (1st row) and soybean (2nd row).
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/yield-response-1}
-
-}
-
-\caption{The impact of water deficit on dryland and irrigated yield by crop}\label{fig:yield-response}
-\end{figure}
-
-In a fairly wet year without much water stress, dryland and irrigated production result in very similar crop yields for both soybean and corn yields.
-Irrigated yields are only slightly higher than dryland yields at a negative value of water deficit (precipitation is greater than reference evapotranspiration).
-However, dryland yield declines sharply and steadily as water deficit increases.
-On the contrary, irrigated corn yield actually increase as water deficit increases up to about water deficit of 400mm at all the saturated thickness levels for both corn and soybean.
-This is because irrigation can turn unfavorable high-heat conditions into favorable growing conditions up to a point (Tim, citations?).
-However, yield response to water deficit departs after deficit of about 400mm depending on the level of saturated thickness.
-While higher saturated thickness categories can withstand water deficit and crop yield stays about the same for higher water deficit values, the low saturated thickness category experience non-trivial declines in crop yields as water deficit increases.
-
-While there are some gaps in crop yield between the lowest and the higher saturated thickness categories, the lowest category still performs quite well for both corn and soybean.
-This is because the samples in irrigation in the regression analysis are those who opted to stay irrigated even when the saturated thickness level is low because irrigation is still powerful enough to provide protections for farmers.
-So, fields who would have experienced significant declines in crop yield in severe droughts due to low saturated thickness have opted out of irrigation and thus not observable in the data for irrigated production.
-Therefore, In order to capture the fuller impact of a decline in saturated thickness, it is important to understand how saturated thickness affects the decision of whether to irrigate or not.
-
-\hypertarget{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}{%
-\subsection{The impact of saturated thickness on the share of irrigated production}\label{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/ir-share-1}
-
-}
-
-\caption{The impact of saturated thickness on the share of irriagate production by crop }\label{fig:ir-share}
-\end{figure}
-
-As shown in \ref{fig:ir-share}, saturated thickness has significant impacts on the share of irrigated production for corn.
-As saturated thickness declines, the share of irrigated acres for corn decreases.
-As saturated thickness declines, more land will be under dryland production, which is more vulnerable to drought, damaging average corn yield.
-Unlike corn, a decline in saturated thickness does not have as large negative impact on the share of irrigated production for soybean as corn.
-This is likely because soybean is less susceptible to drought compared to corn.
-
-\hypertarget{total-impact-of-decline-in-saturated-thickness}{%
-\subsection{Total impact of decline in saturated thickness}\label{total-impact-of-decline-in-saturated-thickness}}
-
-Taking into account both impacts of saturated thickness, figure \ref{fig:tot-impact} shows how water deficit affects yield at different saturated thickness levels evaluated.
-For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness.
-The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher.
-In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15\% and 25\% higher than at 75.7 meters and 29.6 meters, respectively.
-These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness.
-In summary, our studies confirms that a decline in saturated thickness makes yield production much more vulnerable to severe drought, and its negative impact is non-negligible.
-
-It is also noteworthy that the impact of saturated thickness on crop yield in drought years is non-linear.
-Specifically, according to our estimates, the decline of saturated thickness from 186 meters to 75 meters would cause about one tonne/ha of damage in corn yield at water deficit of 1100mm.
-However, the decline of saturated thickness from 75 meters to 30 meters would cause about 1.5 tonne/ha of damage in corn yield.
-This means that the negative impact of a meter of decline in saturated thickness becomes larger as the aquifer depletion progresses.
-This has an important policy implication.
-Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwater-irrigated production.
-Then, we need to act early before saturated thickness becomes too low so that we do not need to product at a less productive steady state.
-The longer you wait, the harsher damage is going to be when you finally decide to place a policy that limits groundwater pumping.
-
-\begin{figure}
-
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/tot-impact-1}
-
-}
-
-\caption{The impact of water deficit on average crop yield by sturated thickness}\label{fig:tot-impact}
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This article quantified the impact of saturated thickness on average crop yield\ldots{}
-
-\clearpage
-
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\hypertarget{use-spei-in-place-of-water-deficit}{%
-\section{Use SPEI in place of water deficit}\label{use-spei-in-place-of-water-deficit}}
-
-Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \ref{tab:fit-reg}).
-
-\begin{table}[!h]
-
-\caption{\label{tab:fit-reg}Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean}
-\centering
-\begin{tabular}[t]{lrr}
-\toprule
-Crop & SPEI & Water Deficit\\
-\midrule
-Corn & 0.899 & 0.903\\
-Soybean & 0.840 & 0.846\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \ref{fig:spei}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=360px]{manuscript_DR_files/figure-latex/spei-1}
-
-}
-
-\caption{The impact of SPEI on crop yield for corn and soybean}\label{fig:spei}
-\end{figure}
-
-\clearpage
-
-  \bibliography{DRA.bib}
+We do not have any reviewers that we would like to avoid.
 

### 2023-07-18 Writing/NatureWater/1stRoundRevision/revised_manuscript/manuscript_DR.tex
revision started

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather},
  pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
% \author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of Mechanical, Aerospace and Civil Engineering, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
\date{}

\begin{document}
\maketitle

% \centering{\textbf{This paper is a non-peer reviewed preprint submitted to EarthArXiv.}}

% \vspace{3cm}

\begin{abstract}
Aquifer depletion poses a major threat to the ability of farmers, food supply chains, and rural economies globally to use groundwater as a means of adapting to climate variability and change. Empirical research has demonstrated the large differences in drought risk exposure that exist between rainfed and irrigated croplands, but previous work commonly assumes water supply for the latter is unconstrained. In this paper, we evaluate how aquifer depletion affects the resilience of irrigated crop production to drought risk using over 30 years of data on historical corn and soybean yields, production areas, and aquifer conditions for the High Plains region in the United States. We show that aquifer depletion reduces the ability of farmers to sustain irrigated crop yields and production areas in drought years. Our findings demonstrate that drought-related production losses on irrigated croplands increase non-linearly with aquifer depletion, highlighting the need for proactive aquifer conservation interventions to support adaptation and resilience to future increases in rainfall variability under climate change. 
\end{abstract}

% \vspace{3cm}

% \textbf{Acknowledgement}: Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.

% \newpage

\hypertarget{Main Text}{%
\section{Main Text}\label{Main Text}}

Groundwater is an essential input for agricultural production, providing a critical buffer against limited or variable surface water supplies for farmers in many parts of the world \citep{scanlon2023global}. Globally, agriculture’s dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change \citep{zhou2010impact, wada2013multimodel, wada2014sustainability, kreins2015quantification, florke2018water}. However, many of the world’s most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}, which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.

Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers capacity to adapt successfully to drought and rainfall variability.  
Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear,lobell2014greater,schlenker2010robust,zhou2020connections,borgomeo2020impact}. Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought \citep{kuwayama2019estimating,zipper2016drought,zhu2022untangling,zhu2022warming,lu2020mapping,davis2019sensitivity,li2018changes}. In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling \citep{foster2015well,cotterman2018groundwater,kahil2015modeling,yoon2021coupled,rad2020mod}, with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage \citep{jain2021groundwater,suter2021depletion}.

Two pathways exist by which aquifer depletion will impact farmers’ ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}. First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}. Second, depletion also reduces the pumping capacity of wells used to extract groundwater, which may create a physical constraint to the volume of irrigation that can be applied to a field \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}. The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress. Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production \citep{foster2014modeling, rad2020effects}. Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust,chiang2021evidence,cook2020twenty}

In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought risk using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States. Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater,lu2018crop}. We also provide new empirical insights about the relative roles of extensive (i.e., irrigated area) and intensive (i.e., per-area irrigation rates) margin water use adjustments in determining drought-related production losses, which previously had only been demonstrated through theoretical modeling \citep{foster2014modeling,foster2017effects,rad2020effects}. Our findings provide important evidence of the value of preserving aquifer storage as a buffer against drought events. In doing so, our analysis adds weight to the need to improve groundwater resource conservation and sustainability to enable adaptation to climate change in groundwater-dependent agricultural systems globally \citep{jain2021groundwater,scanlon2023global}.

%Add troy2015impact and li2018changes somewhere.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{the-impact-of-drought-on-rainfed-and-irrigated-yield}{%
\subsection{Aquifer depletion increases sensitivity of crop yields to drought}\label{the-impact-of-drought-on-rainfed-and-irrigated-yield}}

Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (1st row) and soybean (2nd row). In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean. As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases. This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling,li2020quantifying}.

Differentiating irrigated yields by the level of aquifer thickness alters these responses in two ways. First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}). This reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous}. Second, for negative water deficits, counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response} experience marginal declines in yields resulting in productivity close to rainfed yields. This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing,gibson2017case,gibson2019benchmarking}. However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e. high deficits), in particular for corn.

\begin{figure}

{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_yield_intensive.pdf}

}

\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:yield-response}
\end{figure}

\hypertarget{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}{%
\subsection{Aquifer depletion substantially reduces the share of agricultural land under irrigation}\label{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}}

While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in moderate to extreme drought conditions, the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production. One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation. Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous,deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.

Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean. Figure \ref{fig:ir-share} demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn. As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m. Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn \citep{zipper2016drought,ruess2022irrigation}.

\begin{figure}
{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_share.pdf} 
}
\caption{The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:ir-share}
\end{figure}

\hypertarget{total-impact-of-decline-in-saturated-thickness}{%
\subsection{Aquifer depletion increases overall drought-related production losses by up to 25\%}\label{total-impact-of-decline-in-saturated-thickness}}

The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \ref{the-impact-of-drought-on-rainfed-and-irrigated-yield} and \ref{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}. Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean.

For corn, aquifer thickness does not have a significant impact on  productivity when precipitation deficits are small (less than around 200mm). However, beyond this threshold, higher levels of deficit result in noticeable and significant non-linear reductions to average corn production. This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower. Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e. greater than around 600mm) and are smaller in overall magnitude. For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.

%It is also noteworthy that the impact of aquifer thickness on crop yield in drought years is non-linear. Specifically, according to our estimates, the decline of aquifer thickness from 186 meters to 75 meters would cause about one tonne/ha of damage in corn yield at water deficit of 1100mm. However, the decline of aquifer thickness from 75 meters to 30 meters would cause about 1.5 tonne/ha of damage in corn yield. This means that the negative impact of a meter of decline in aquifer thickness becomes larger as the aquifer depletion progresses. This has an important policy implication. Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwater-irrigated production. Then, we need to act early before aquifer thickness becomes too low so that we do not need to product at a less productive steady state. The longer you wait, the harsher damage is going to be when you finally decide to place a policy that limits groundwater pumping.

\begin{figure}
{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_total_impact.pdf} 
}

\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness}\label{fig:tot-impact}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

We demonstrate the important role of aquifer thickness in determining farmers ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States. Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear,lobell2014greater,schlenker2010robust,zhou2020connections,borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating,zipper2016drought,zhu2022untangling,zhu2022warming,lu2020mapping,davis2019sensitivity,li2018changes,luan2021combined}. Our analysis adds new empirical insights to this literature which show explicitly that reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events. Critically, this suggests that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.

An important implication of our findings is that depletion of groundwater resources is likely to increase significantly the vulnerability of agricultural production to current and future drought risks. Climate change is projected to increase the frequency and severity of extreme drought events in many parts of the world \citep{ukkola2020robust,chiang2021evidence,cook2020twenty}, including in parts of Kansas and Texas \citep{bradford2020robust,cook2022projected,mullens2019quantitative} where the High Plains Aquifer continues to experience prolonged and significant declines in aquifer thickness \citep{scanlon2012groundwater, haacker2016water, cotterman2018groundwater}. Similar trends of increasing drought risks coupled with declining aquifer storage have been reported in other major groundwater irrigated farming systems worldwide \citep{wada2010global, doll2014global, famiglietti2014global, feng2018groundwater, bierkens2019non}. Our findings suggest that these simultaneous changes in aquifer conditions will amplify impacts of future climate change on crop production, with associated ramifications for resilience of supply chains, commodity markets, rural economies, and food security.

Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally. Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and drought impacts is highly non-linear when accounting for farmers decisions about irrigated production areas. This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon,butler2020charting,elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought. Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production \citep{foster2017effects,deines2020transitions}. For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure  50\% of current aquifer storage remains in 50 years time (so called 50/50 rule) \citep{closas2018chronicle}. The effectiveness of such targets could likely be improved substantially by setting conservation targets based on empirical evidence about the level of aquifer thickness that is required to ensure adequate levels of drought protection, both now and with future climate change.

Several opportunities exist for future research to build upon the methods and analyses presented here. First, due to the limited resolution of our observational data, we were only able to estimate the impacts of different aquifer thickness categories on drought risk. Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining,deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed. This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.

Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types. Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact,deines2019quantifying,manning2017producer}. Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.

Finally, we focus on the impacts of aquifer thickness on agricultural resilience to seasonal water deficits. However, where water deficits are distributed unevenly during the season, negative impacts of aquifer depletion may be even larger. This is because reductions in well yields caused by aquifer depletion will constrain farmers ability to meet peak crop water demands, increasing yield losses compared with if water deficits had been distributed evenly throughout the season \citep{ortiz2019unpacking}. Furthermore, irrigation also has broader benefits for crops such as reducing heat-related stress through cooling of land surface temperatures \citep{adegoke2003impact, bonfils2007empirical, lobell2008effect, zhu2022untangling}. Reductions in aquifer thickness therefore may have other negative impacts on irrigated crop productivity, if constraints to water supply due to aquifer depletion are significant enough to limit cooling benefits of irrigation during the growing season.

% \hypertarget{conclusion}{%
% \section{Conclusion}\label{conclusion}}

% In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events. Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer thickness significantly reduces the overall productivity of both corn and soybean during drought events. Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion. Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{methods-data}{%
\subsection{Data-sets}\label{methods-data}}

Our study focuses on the High Plains Aquifer (HPA) in the Central United States. Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{fenichel2016measuring,smidt2016complex}. To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.

County-level irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S. Department of Agriculture's National Agricultural Statistics Service (USDA-NASS). USDA-NASS does not report irrigated yields disaggregated by the source of water used for irrigation. However, it is known that groundwater is the main source of water for irrigation for counties that overlie the HPA due to the limited availability of surface water resources in this area. Therefore, we retain both rainfed and irrigated yield observations for counties where greater than 75\% of the county area overlies the boundary of the HPA. For counties that do not meet this condition, only rainfed yield observations are obtained to avoid the possibility that irrigated production may be only partly dependent on use of groundwater from the HPA. We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production.

Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset \citep{Abatzoglou2013}. The gridMET data is a gridded daily weather dataset with a spatial resolution of ~4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}. Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample. Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET, and were then used to calculate daily reference crop evapotranspiration for each grid cell. From these data, we computed the seasonal water deficit (\(WB_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region. Seasonal water balance estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell. Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012. The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \ref{fig:deficit-yield-hist})

\begin{figure}
{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_y_wd_all.png} }
\caption{Historical dryland yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
\end{figure}

As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA reported in \citet{haacker2016water}. These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935, derived based on extensive water level observations from monitoring wells and detailed maps of both land surface elevation and the bedrock elevation of the HPA. aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data. aquifer thickness values were only estimated for counties with irrigated yield observations. Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring,haacker2016water,deines2020transitions}. Figure \ref{fig:sat-dist} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses. aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer influences farmers' exposure and responses to drought.

These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield (8773 observations), (2) per-acre soybean yield (5977 observations), (3) share of irrigated corn (2686 observations), and (4) share of irrigated soybean (2014 observations).

\begin{figure}
{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_map.png}}
\caption{Map showing aquifer thickness in 2016 for each county included in our corn and soybean regression analyses. Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-dist}
\end{figure}

\hypertarget{methods-models}{%
\subsection{Regression Models}\label{methods-models}}

Aquifer depletion may impact resilience of irrigated crop production to drought through two main channels: (1) a reduction in per-area yields due to the economic or physical inability to meet crop water requirements; and (2) a reduction in the area of land irrigated, resulting in reduced production due to lower yields obtained through rainfed agriculture.

To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield to drought is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year. Our regression model is given in Equation \ref{equation1}, with the same model formulation used for corn and soybean:

\begin{equation}\label{equation1}
y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t}
\end{equation}

where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit. \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise. The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3\%)\), \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness. The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category. Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient. To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.

To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalize additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms. This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF). The model is given in Equation \ref{equation2} below:

\begin{equation}\label{equation2}
    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \sum_{h}^H \gamma_{h}\rho_{h}(WB_{i}, AT_{i,t}) + \mu_t + v_{i,t}
\end{equation}

where \(s_{i,t}\) is the share of irrigated production. The impact of average water balance is captured by $\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})$, where $\phi_{k}(\cdot)$ is the $k$th spline basis function and $\alpha_{k}$ is its coefficient. The impact of aquifer thickness is captured in a similar way with $\tau_{l}(\cdot)$ and $\beta_{l}$ being the $l$th spline basis function and its coefficient, respectively. The interactive impact of average water balance and aquifer thickness is modeled by $\sum_{h}^H \gamma_{h}\rho_{h}(WB_{i}, AT_{i,t})$. Here, $\rho_{h}(\cdot)$ is a bi-variate spline basis function and $\gamma_{h}$ is its coefficient. Year fixed effects and error term are represented by $\mu_t$ and $v_{i,t}$, respectively. Unlike our regression model of the impact of water balance on yield, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable. The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time. Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.

In the final step of our analysis, we then estimate the average crop yield for each county as a function of aquifer thickness and seasonal water deficit. The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \ref{equation1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \ref{equation2}. Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area yields and the area of land irrigated.

\hypertarget{data-code-availability}{%
\subsection{Data and Code Availability}\label{data-availability}}
All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}). All the raw data-sets used in this study are publicly accessible. The boundary shape file of the High Plains aquifer was obtained from the U.S. Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}. This data is used for Figure \ref{fig:sat-dist}. Corn and soybean yield data were obtained from the U.S. Department of Agriculture’s National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}. The R tidyUSDA package was used to automatically download the relevant data. This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. Weather data (precipitation, temperature, and evapotranspiration) was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}. This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. The aquifer thickness data were obtained directly from Dr. Erin Haacker (\href{mailto:ehaacker2@unl.edu}, who can be reached at {\nolinkurl{ehaacker2@unl.edu}}). This data is used for Figure \ref{fig:sat-dist} and the regression analysis. The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.

\clearpage

%\hypertarget{appendix-appendix}{
%\appendix}

%\hypertarget{use-spei-in-place-of-water-deficit}{%
%\section{Use SPEI in place of water deficit}\label{use-spei-in-place-of-water-deficit}}

%Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \ref{tab:fit-reg}).

%\begin{table}[!h]

%\caption{\label{tab:fit-reg}Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean}
%\centering
%\begin{tabular}[t]{lrr}
%\toprule
%Crop & SPEI & Water Deficit\\
%\midrule
%Corn & 0.899 & 0.903\\
%Soybean & 0.840 & 0.846\\
%\bottomrule
%\end{tabular}
%\end{table}

%Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \ref{fig:spei}.

%\begin{figure}

%{\centering \includegraphics[width=360px]{manuscript_DR_files/figure-latex/spei-1}

%}

%\caption{The impact of SPEI on crop yield for corn and soybean}\label{fig:spei}
%\end{figure}

\clearpage

  \bibliography{DRA.bib}

\end{document}

### 2023-07-18 Codes/1_analysis.R

@@ -33,24 +33,25 @@ library(scales)
 # here::i_am("GitControlled/Codes/Functions/1_0_functions.R")
 
 #--- source functions ---#
-source(here::here("Codes/Functions/1_0_functions.R"))
-# source("GitControlled/Codes/Functions/1_0_functions.R")
+# source(here::here("Codes/Functions/1_0_functions.R"))
+source("GitControlled/Codes/Functions/1_0_functions.R")
 
 
 ## ----r------------------------------------------------------------------------
 #--- read data ---#
 final_data <-
-  # here("Shared/Data/ProcessedData/final_data.rds") %>%
-  here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
+  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
   readRDS()
 
 #--- Data used for regression analysis ---#
 reg_data <-
   prepare_reg_data(
     data = final_data,
-    sat_thld = 9, # saturated thickness has to be at least 12 meters
+    sat_thld_m = 12, # saturated thickness has to be at least 12 meters
     ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
     balance_thld = c(-1200, 200),
+    sat_cat_num = 3,
     state_ls =
       list(
         corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
@@ -61,6 +62,7 @@ reg_data <-
 
 
 ## ----r------------------------------------------------------------------------
+
 data_corn <-
   reg_data[crop_type == "corn", reg_data_y][[1]] %>%
   copy() %>%
@@ -121,9 +123,9 @@ data_for_pred_corn <-
   ) %>%
   # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
   .[, sat := case_when(
-    sat_cat == "[29.6,75.7]" ~ 30,
-    sat_cat == "(75.7,186]" ~ 100,
-    sat_cat == "(186,708]" ~ 200,
+    sat_cat == "[39.4,82]" ~ 40,
+    sat_cat == "(82,194]" ~ 100,
+    sat_cat == "(194,708]" ~ 200,
     sat_cat == "dryland" ~ 0
   )] %>%
   .[is.na(sat), sat := 0] %>%
[…]

### 2023-07-18 Writing/NatureWater/1stRoundRevision/knit.R

@@ -1,0 +1,3 @@
+setwd(here::here("GitControlled/Writing/NatureWater/1stRoundRevision"))
+
+

### 2023-07-21 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.rmd
Response to reviewers added

@@
 
-\textcolor{blue}{Tim? this is realted to one of the reviewer1's comments}
+\textcolor{blue}{Tim? this is realted to one of the reviewer 1's comments}
 
@@
 
-\textcolor{blue}{}
+\textcolor{blue}{Thank you for this comment, which we believe led to a slighly better model and acknowledgin important limitations of our methods.
+County and year fixed effects were included in equation (1), but only year fixed effects were included in equation (2).
+As you gueesed, the problem of including county FEs lies in the massive elimination of variation in aquifer thickness.
+Please see Figure 6 in Appendix A for the histogram of (1) aquifer thickness demeaned by the county average, (2) aquifer thickness demeaned by the state average, and (3) aquifer thickness without demeaning.
+As you can see in the figure, inclusion of county FEs (mathematically equivalent to regression using county-demeaned aquifer thickness) dramatically reduces the total variation in aquifer thickness and not enough variations are left for accurately estimating its impact of the share of irrigated acres.
+The inclusion of State FEs does not diminish the variation in aquifer thickness.
+Therefore, while the inclusion of State FEs certainly does not provide as tight of a control as county FEs, we have decided to include State FEs for both corn and soybean irrigated share regrssions as it is better than not including them.
+Figure 7 compares the estiamted impacts of aquifer thickness with or without State FEs.
+The negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included.
+Therefore, it may be that aquifer thickness is not very highly correlated with important unobservable variables that are affecting irrigated share significantly.
+Given the direction of potential bias, providing tighter controls with county fixed effects may result in even more significant drop in the share of irrigated acres (we may be underestimating the impact of aquifer thickness.).
+All the discussion we are providing here is now seen in Appenddix A.
+}
 

### 2023-07-21 Codes/Retired/3_1_make_figures_results.Rmd
Response to reviewers added

---
title: "Make Figures"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi=400,
  fig.width=7.5,
  fig.height=5,
  out.width="750px",
  out.height="500px"
)
```

# Prep

```{r package, include = F}
library(tmap)
library(tidyverse)
library(data.table)
library(tidycensus)

#=== regression results ===#
reg_results <- readRDS(here("Shared/Results/reg_results.rds"))

#=== hypothetical sim for regression results illustration ===#
sim_results <- readRDS(here("Shared/Results", "hypothetical_sim_results.rds"))

#=== simulation results ===#
data_sim_yield <- readRDS(here("Shared/Results", "simulated_yield_data.rds"))

#=== county sf ===#
data(fips_codes)

all_counties <-
  tigris::counties() %>% 
  rename(
    state_code = STATEFP, 
    county_code = COUNTYFP,
    county_name = NAME
  ) %>% 
  left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
  dplyr::select(state_code, state, county_code, county_name) %>% 
  mutate(sc_code = paste0(state_code, county_code)) %>% 
  data.table()
```

# Yield response to drought and heat
## Water Deficit (Balance) on Yield by Saturated Thickness Level

```{r yield-intensive-balance, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}
(
g_intensive_balance <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  mutate(days_ab_35_med = 
    yield_hat_eval_data[which.min(abs(days_ab_35-median(days_ab_35))), days_ab_35]
  ) %>% 
  mutate(yield_hat_eval_data = list(
    yield_hat_eval_data[days_ab_35 == days_ab_35_med, ]
  )) %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  unnest(cols = "yield_hat_eval_data") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = balance
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = balance)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()
)

# (
# g_intensive_balance <- 
#   sim_resulits %>% 
#   filter(model_name == "et0_prcp") %>% 
#   dplyr::select(crop_type, yield_hat_eval_data) %>% 
#   mutate(prcp_med = 
#     yield_hat_eval_data[which.min(abs(prcp-median(prcp))), prcp]
#   ) %>% 
#   mutate(yield_hat_eval_data = list(
#     yield_hat_eval_data[prcp == prcp_med, ]
#   )) %>% 
#   dplyr::select(crop_type, yield_hat_eval_data) %>% 
#   unnest(cols = "yield_hat_eval_data") %>% 
#   ggplot(data = .) +
#     geom_ribbon(
#       aes(
#         ymin = yhat_d, 
#         ymax = yhat_u, 
#         x = et0
#       ),
#       alpha = 0.4
#     ) +
#     geom_line(aes(y = y_hat, x = et0)) +
#     facet_grid(crop_type ~ sat, scales = "free_y") +
#     theme_bw()
# )
```

## Days above 35 on Yield by Saturated Thickness Level

```{r yield-intensive-d35, fig.cap = "The Impact of the Number of Days above 35 on Irrigated Yield at Various Saturated Thickness Levels", fig.height = 4, fig.width = 6.5}
(
g_intensive_d35 <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  mutate(balance_med = 
    yield_hat_eval_data[which.min(abs(balance-median(balance))), balance]
  ) %>% 
  mutate(yield_hat_eval_data = list(
    yield_hat_eval_data[balance == balance_med, ] %>% 
      .[, type := "days_ab_35"]
  )) %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  unnest(cols = "yield_hat_eval_data") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = days_ab_35
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = days_ab_35)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()
)
```

# The impact of saturated thickness on Irrigation Share

```{r ir-share, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_ir_share <- 
  sim_results %>% 
  dplyr::select(crop_type, ir_share_eval_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = ir_share_hat, 
      x = sat, 
      color = factor(balance_avg)
    )
  ) +
  facet_grid(. ~ crop_type) +
  theme_bw()

# ggsave(
#   here("Writing/Figures", "ir_share.pdf"),
#   g_ir_share, 
#   height = 3.5,
#   width = 6
# )

g_ir_share
```

# The Impact of Water Deficit on Yield by Saturated thickness (intensive and extensive margins combined)

```{r yield-combined, fig.cap = "The Impact of Saturated Thickness on Yield", fig.height = 6, fig.width = 6.5}

g_sat_impact <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, avg_yield_hat_eval_data) %>% 
  mutate(days_ab_35_med = 
    avg_yield_hat_eval_data[which.min(abs(days_ab_35-median(days_ab_35))), days_ab_35]
  ) %>% 
  mutate(balance_avg_med = 
    avg_yield_hat_eval_data[which.min(abs(balance_avg-median(balance_avg))), balance_avg]
  ) %>% 
  mutate(days_ab_35_avg_med = 
    avg_yield_hat_eval_data[which.min(abs(days_ab_35_avg-median(days_ab_35_avg))), days_ab_35_avg]
  ) %>% 
  mutate(plot_data = list(
    avg_yield_hat_eval_data %>% 
    .[days_ab_35 == days_ab_35_med, ] %>% 
    .[balance_avg == balance_avg_med, ] %>% 
    .[days_ab_35_avg == days_ab_35_avg_med, ] 
  )) %>% 
  dplyr::select(crop_type, plot_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = mean_yield, 
      x = balance, 
      color = factor(sat)
    )
  ) +
  facet_grid(crop_type ~ ., scales = "free_y") +
  scale_color_viridis_d(name = "Saturated Thickness") +
  theme_bw() +
  theme(
    legend.position = "bottom"
  )

g_sat_impact
```

# Climate change impact

```{r echo = F, eval = F}
data_sim_yield <- 
  here("Shared/Results/simulated_yield_data.rds") %>% 
  readRDS() %>% 
  data.table() %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  .[, yield := y_hat_dr * (1 - ir_share) + y_hat * ir_share]
```

## Overall trend (Corn)

```{r echo = F, eval = F}
plot_data_corn_ir <- 
  data_sim_yield %>% 
  .[crop_type == "corn", ] %>% 
  .[sat_type == "hypothetical", ] %>% 
  .[, .(yield, y_hat_dr, cl_period, sc_code, rcp, real_year)]

plot_data_corn_dr <-
  copy(plot_data_corn_ir) %>% 
  .[, yield := NULL] %>% 
  setnames("y_hat_dr", "yield") %>% 
  unique(by = c("cl_period", "sc_code", "rcp", "real_year"))

ggplot(plot_data_corn) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)

ggplot(data_overall[crop_type == "corn", ]) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)

data_overall <- 
  data_sim_yield %>% 
  .[sat_type == "hypothetical", ] %>% 
  .[, .(yield = mean(yield)), by = .(rcp, sc_code, cl_period, crop_type, cl_type, sat)]

ggplot(data_overall[crop_type == "corn", ]) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)
```

# Climate change simulation analysis

Variables that can be varied:

+ annual weather
  * balance (water deficit): irrigated and dryland yield 
  * days_ab_35: irrigated and dryland yield 
+ climate variable
  * balance average: share of irrigated acres 
  * days_ab_35 average: share of irrigated acres

Key limitations:

Water use is not modeled. As climate change progresses, more water is necessary.
However, its effect on saturated thickness is not modeled. Any simulations will not 
be realistic. Suppose we assume the historical trend of saturated thickness decline and use it
to project future saturated thickness level. This is not realistic since saturated thickness decline
should be faster because of greater water demand in future years than the past years.

Simulation variants:

+ **Irrigation share (fixed) and saturated thickness (fixed)**: 
This option answers the question of what would happen to average yield if farmers in the current 
production condition (irrigation share and saturated thickness) suddenly face weather like you will see
in the future.  
+ Irrigation share (varied based on climate) and saturated thickness (fixed): 
This option allows climate to influence irrigation share as climate change progresses. 
Under this option, irrigation share increases as the demand for irrigation increases. 
However, this ignores that fact that an increase in irrigation share has to reduce saturated 
thickness, which is assumed fixed.
+ Irrigation share (fixed) and saturated thickness (varied): inconsistent 
+ **Irrigation share (varied based on sat and climate) and saturated thickness (varied)**:

## Saturated thickness and ir_share fixed

```{r }
sim_ob <- 
  readRDS(here("Shared/Results/cl_simulationi_obs.rds")) %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"] %>% 
  nest_by(sc_code, crop_type) %>% 
  mutate(data_plot = list(
    data.table(data)[, .(y_ir_hat, y_dr_hat, avg_y_hat, cl_period, rcp)] %>% 
    melt(id.var = c("cl_period", "rcp")) %>% 
    .[, type := case_when(
      variable == "y_ir_hat" ~ "Irrigated Yield",
      variable == "y_dr_hat" ~ "Dryland Yield",
      variable == "avg_y_hat" ~ "Average Yield"
    )]
  )) %>% 
  mutate(g_cl_impact = list(
    ggplot(data_plot) +
      geom_boxplot(aes(y = value, x = cl_period, fill = factor(rcp))) +
      facet_grid(type ~ .) +
      ggtitle(
        paste0(
          "State = ",
          unique(data$state),
          ", County = ",
          unique(data$county_name),
          "\nSaturated Thickness = ", 
          round(unique(data$sat), digits = 2), 
          ", Share of Irrigated Acres = ", 
          round(unique(data$ir_share), digits = 3)
        )
      )
  ))
```

**Contributing Factors**

+ the impact of weather on irrigated yield (vary by saturate thickness)
+ the impact of weather on dryland yield
+ irrigated acreage share (fixed and not affected by saturated thickness)

By far the biggest contributor of the climate change impact of yield
is the second and third items, especially irrigated acreage share.
Irrigated yield can increase on average. Yield increases as balance increases
for a certain range (see the figures above). Depending on the change in the distribution of balance, 
climate change can increase yield on average.

### Corn

```{r }
filter(sim_ob, crop_type == "corn")$g_cl_impact
```

### Soy

```{r }
filter(sim_ob, crop_type == "soy")$g_cl_impact
```

## Average representative farm experiencing different kind of climate

Share of Irrigated Acres modeled for a few set of saturated thickness
levels, which are fixed for the entire period. Observed irrigated
share and saturated thickness level not respected.

Observations:
Share of irrigated acres goes up significantly because of warming climate without
any punishment on saturated thickness. Consequently, irrigated acres go up almost 
up to 1, leading to extreme resilience of average yield to climate change.

```{r }
sim_h <- 
  readRDS(here("Shared/Results/cl_simulationi_h.rds")) %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"] %>% 
  nest_by(sc_code, crop_type) %>% 
  mutate(data_plot = list(
    data.table(data)[, .(y_ir_hat, y_dr_hat, avg_y_hat, cl_period, rcp, sat, ir_share)] %>% 
    melt(id.var = c("cl_period", "rcp", "sat")) %>% 
    .[, type := case_when(
      variable == "y_ir_hat" ~ "Irrigated Yield",
      variable == "y_dr_hat" ~ "Dryland Yield",
      variable == "avg_y_hat" ~ "Average Yield",
      variable == "ir_share" ~ "Share of Irrigated Acres"
    )]
  )) %>% 
  mutate(g_cl_impact = list(
    ggplot(data_plot) +
      geom_boxplot(aes(y = value, x = cl_period, fill = factor(sat))) +
      facet_grid(type ~ rcp, scales = "free_y") +
      ggtitle(
        paste0(
          "State = ",
          unique(data$state),
          ", County = ",
          unique(data$county_name)
        )
      )
  ))

```
```{r eval = F, echo = F}
sim_h$g_cl_impact[[1]]
sim_h$data_plot[[1]]
data.table(sim_h$data[[1]])[, .(real_year, sat, ir_share)]
```

### Corn

```{r }
filter(sim_h, crop_type == "corn")$g_cl_impact
```

### Soy

```{r }
filter(sim_h, crop_type == "soy")$g_cl_impact
```

### 2023-07-21 Codes/Retired/3_2_make_figure_misc.Rmd
Response to reviewers added

---
title: "Generate figures (climate, saturated thickness)"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Preparation

```{r }
#=== packages ===#
library(downloader)
library(terra)
library(lubridate)
library(tidyverse)
library(tidycensus)
library(data.table)
library(ncdf4)
library(sf)
library(raster)
library(parallel)
library(tmap)
library(here)
library(geosphere)
library(exactextractr)

#=== source functions ===#
source(here("GitControlled/Codes/functions.R"))

#=== Regression results ===#
# created in 1_regression_analysis.R
reg_results <- 
  here("Shared/Results", "reg_results.rds") %>% 
  readRDS() %>% 
  dplyr::select(crop_type, states, data, share_data, model_name) %>% 
  unique(by = "crop_type")
```

# Prepare Data

```{r echo = F}

#=== sf for yield regression ===#
sc_y_sf <-  
  reg_results %>% 
  dplyr::select(crop_type, data) %>% 
  unique(by = "crop_type") %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, state, geometry, crop_type)] %>% 
  unique(by = c("sc_code", "crop_type"))

#=== sf for ir_share regression ===#
sc_is_sf <-  
  reg_results %>% 
  dplyr::select(crop_type, share_data) %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, state, geometry, crop_type)] %>% 
  unique(by = c("sc_code", "crop_type"))

#=== hpa ===#
hpa_simplified <- 
  here("Shared/Data/ProcessedData/hp_simplified.shp") %>% 
  readRDS()

#=== all counties ===#
data(fips_codes)

all_counties <-
  tigris::counties() %>% 
  rename(
    state_code = STATEFP, 
    county_code = COUNTYFP,
    county_name = NAME
  ) %>% 
  left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
  dplyr::select(state_code, state, county_code, county_name) %>% 
  mutate(sc_code = paste0(state_code, county_code)) %>% 
  data.table()

all_states <- 
  tigris::states() %>% 
  rename(state_code = STATEFP) %>% 
  left_join(
    ., 
    unique(data.table(fips_codes), by = "state"), 
    by = "state_code"
  )

counties_states <- 
  reg_results %>%  
  mutate(all_counties = list(
    tigris::counties() %>% 
    rename(state_code = STATEFP, county_code = COUNTYFP) %>% 
    left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
    filter(state %in% unique(share_data$state)) %>% 
    dplyr::select(state)  
  )) %>% 
  mutate(all_states = list(
    tigris::states() %>% 
    rename(state_code = STATEFP) %>% 
    left_join(
      ., 
      unique(data.table(fips_codes), by = "state"), 
      by = "state_code"
    ) %>% 
    filter(state %in% unique(share_data$state)) %>% 
    dplyr::select(state) 
  ))
```

# Geographic Focus
## Yield regression
Map of counties used for yield regression

+ Irrigated yield observations outside of HPA are dropped because saturated thickness is not observed
+ Dryland yield observations are kept even if outside of HPA
+ Both dryland and irrigated yield observations are use if inside of HPA

```{r }
states_y <- 
  sc_y_sf %>% 
  nest_by(crop_type) %>% 
  mutate(state_ls = list(
    unique(data$state)
  )) %>% 
  mutate(state_sf = list(
    filter(all_states, state %in% state_ls)
  )) %>% 
  dplyr::select(crop_type, state_sf) %>% 
  unnest()

ggplot() + 
  geom_sf(data = st_as_sf(sc_y_sf), fill = "green", alpha = 0.4) +
  geom_sf(data = st_as_sf(states_y), color = "red", fill = NA) +
  geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.2) +
  facet_grid(. ~ crop_type)
```

## Irrigation Share (and Simulation)
Map of counties used for irrigation share regression and simulations

+ Only the counties with 95% of its area over HPA are kept

```{r }
states_is <- 
  sc_is_sf %>% 
  nest_by(crop_type) %>% 
  mutate(state_ls = list(
    unique(data$state)
  )) %>% 
  mutate(state_sf = list(
    filter(all_states, state %in% state_ls)
  )) %>% 
  dplyr::select(crop_type, state_sf) %>% 
  unnest()

ggplot() + 
  geom_sf(data = st_as_sf(sc_is_sf), fill = "green", alpha = 0.4) +
  geom_sf(data = st_as_sf(states_is), color = "red", fill = NA) +
  geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.2) +
  facet_grid(. ~ crop_type)
```

# Saturated Thickness

```{r }
sat_sf <- 
  reg_results %>% 
  dplyr::select(crop_type, share_data) %>% 
  unique(by = "crop_type") %>% 
  unnest() %>% 
  nest_by(sc_code, crop_type) %>% 
  #=== sat trend ===#
  mutate(beta_sat = 
    lm(sat ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== average sat level ===#
  mutate(sat_mean = 
    mean(data$sat, na.rm = TRUE)
  ) %>% 
  mutate(geometry = 
    unique(data$geometry)
  ) %>% 
  st_as_sf() 
```

## Saturated thickness trend by sc_code
`beta_sat` measures the annual rate of change in saturated thickness
observed 1990 - 2016

```{r hist-sat-trend, fig.cap = "Saturated thickness trend (1990-2016)"}
ggplot() +
  geom_histogram(
    data = sat_sf, 
    aes(x = beta_sat), 
    color = "blue",
    fill = NA
  ) +
  facet_grid(crop_type ~ .)
```

## Histogram of saturated thickness level by sc_code

```{r hist-sat-level, fig.cap = "Mean saturated thickness (1990-2016)"}
ggplot() +
  geom_histogram(
    data = sat_sf, 
    aes(x = sat_mean), 
    color = "blue",
    fill = NA,
    bins = 50
  ) +
  facet_grid(crop_type ~ .)
```

## Histogram of saturated thickness level (below 100)

```{r hist-sat-level-b100, fig.cap = "Mean saturated thickness (less than 100) (1990-2018)"}
ggplot() +
  geom_histogram(
    data = filter(sat_sf, sat_mean <= 100), 
    aes(x = sat_mean), 
    color = "blue",
    fill = NA,
    bins = 50
  ) +
  facet_grid(crop_type ~ .)
```

## Map of saturated thickness level 
Counties with fill colors are used for simulation and ir-share regression. 
Corn: NE, KS, CO, SD 
Soy: NE, KS

```{r }
sc_is_sf <-
  left_join(sc_is_sf, dplyr::select(sat_sf, sc_code, beta_sat, sat_mean), by = "sc_code")

```
```{r hist-sat-map, fig.cap = "Map of mean saturated thickness (1990-2018)"}
ggplot() +
geom_sf(data = st_as_sf(states_is), fill = NA, col = "red") + 
geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.4) + 
geom_sf(data = st_as_sf(sc_is_sf), aes(fill = sat_mean)) + 
facet_grid(. ~ crop_type) +
scale_fill_viridis_c()
```

# Basic data visualization

```{r }
data_y <- reg_results %>% 
  dplyr::select(crop_type, data) %>% 
  unique(by = "crop_type") %>% 
  unnest()
```

## Yield histogram by saturated thickness category

```{r yield-hist}
ggplot(data_y) +
  geom_boxplot(aes(y = yield, x = sat_cat)) +
  facet_grid(crop_type ~ ., scales = "free_y")
```

## Yield against water deficit (balance)

```{r yield-balance}
ggplot(data_y) +
  geom_point(aes(y = yield, x = balance, color = sat_cat), size = 0.4) +
  geom_smooth(aes(y = yield, x = balance, color = sat_cat)) +
  facet_grid(crop_type ~ ., scales = "free_y")
```

# Climate data 
## Prepare data

```{r }
model_ls <- 
  c(
    "GFDL-ESM2M", 
    "HadGEM2-ES365", 
    "IPSL-CM5A-LR", 
    "MIROC5"
  )

rcp_ls <- c(45, 85)

model_rcp <- 
  expand.grid(
    cl_model = model_ls,
    rcp = rcp_ls
  ) %>% 
  data.table()

#=== weather data in the data for regression ===#
# cl_data_current <- 
#   reg_data[, .(sc_code, year, balance, et0, prcp, days_ab_35)] %>% 
#   expand_grid_df(., model_rcp)

#=== future climate data ===#
cl_data_future <- 
  model_rcp %>% 
  rowwise() %>% 
  #=== define maca data file names ===#
  mutate(file_name = list(
    paste0(
      "Shared/Data/ProcessedData/MACA/sum_", 
      cl_model,
      "_",
      rcp,
      "_county.rds"
    ) %>% 
    here()
  )) %>% 
  mutate(maca_data = list(
    #=== read maca data ===#
    readRDS(file_name) %>% 
    #=== calculate 10-year balance avg ===#
    .[, 
      `:=`(
        balance_avg = mean(balance), 
        days_ab_35_avg = mean(days_ab_35)
      ), 
      by = .(floor((year - 2020) / 10), sc_code)
    ] 
  )) %>% 
  dplyr::select(cl_model, rcp, maca_data) %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, year, balance, et0, prcp, days_ab_35, cl_model, rcp)]  %>% 
  .[, cl_period := case_when(
    year <= 2020 ~ "past",
    year > 2020 & year <= 2040 ~ "2021-2040",
    year > 2040 & year <= 2060 ~ "2041-2060",
    year > 2060 & year <= 2080 ~ "2061-2080",
    year > 2080 & year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"]  
```

## Weather variables by period (model mixed)

```{r cl-trend-future, fig.cap = "Water deficit, ET0, and precipitation by period (model mixed)"}
cl_data_future[, .(balance, et0, prcp, days_ab_35, rcp, sc_code, year, cl_period)] %>% 
melt(id.var = c("sc_code", "year", "rcp", "cl_period")) %>% 
ggplot(data = ) +
  geom_density(aes(x = value, fill = cl_period), alpha = 0.4) +
  facet_grid(rcp ~ variable, scales = "free") +
  theme(
    legend.position = "bottom"
  )
```

## Rate of increase in water deficit by county and model

```{r }
cl_trend_county <- 
  cl_data_future %>% 
  nest_by(sc_code, cl_model, rcp) %>% 
  #=== rate of balance increase ===#
  mutate(beta_balance = 
    lm(balance ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of et0 increase ===#
  mutate(beta_et0 = 
    lm(et0 ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of precip increase ===#
  mutate(beta_prcp = 
    lm(prcp ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of precip increase ===#
  mutate(beta_d35 = 
    lm(days_ab_35 ~ year, data = data)$coef[["year"]]
  ) %>% 
  mutate(geometry = 
    unique(data$geometry)
  ) %>% 
  st_as_sf()
```

### Balance

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_balance)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### Precipitation

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_prcp)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### ET0

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_et0)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### days_above_35

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_d35)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### 2023-07-21 Writing/NatureWater/1stRoundRevision/revised_manuscript/manuscript_DR (LAGAE-6N57C7R9D's conflicted copy 2023-07-19).tex
Response to reviewers added

--- 2023-07-18 Writing/NatureWater/1stRoundRevision/revised_manuscript/manuscript_DR.tex
@@
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous}.
-Second, for negative water deficits, counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response} experience marginal declines in yields resulting in productivity close to rainfed yields.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous}.
+Second, at negative water deficits, counties with the second highest and highest levels of aquifer thickness (i.e., 2nd and 3rd quantile in Figure \ref{fig:yield-response}) experience lower yields than counties with the lowest level of aquifer thickness (1st quantile), resulting in productivity close to rainfed yields.
 This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing,gibson2017case,gibson2019benchmarking}.
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalize additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
 This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).

### 2023-07-21 Writing/manuscript.Rmd
Response to reviewers added

---
title: Aquifer depletion exacerbates agricultural drought losses in the US High Plains
author:
  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota, kakim002@umn.edu]
  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
abstract: |
  Aquifer depletion poses a major threat to the ability of farmers, food supply chains, and rural economies globally to use groundwater as a means of adapting to climate variability and change. Empirical research has demonstrated the large differences in drought risk exposure that exist between rainfed and irrigated croplands, but previous work commonly assumes water supply for the latter is unconstrained. In this paper, we evaluate how aquifer depletion affects the resilience of irrigated crop production to drought risk using over 30 years of data on historical corn and soybean yields, production areas, and aquifer conditions for the High Plains region in the United States. We show that aquifer depletion reduces the ability of farmers to sustain irrigated crop yields and production areas in drought years. Our findings demonstrate that drought-related production losses on irrigated croplands increase non-linearly with aquifer depletion, highlighting the need for proactive aquifer conservation interventions to support adaptation and resilience to future increases in rainfall variability under climate change.
acknowledgments: |
  Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.
bibliography: DRA.bib
csl: iopart.csl
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
    includes:
      in_header: "preamble.tex"
---

```{r packages}
#| include: false
#| cache: false

#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(here)
library(tmap)
library(lfe)
library(modelsummary)
library(patchwork)
library(kableExtra)
```

```{r}
#| echo: false

options(knitr.duplicate.label = "allow")
```

```{r run-analys, cache = TRUE, include = FALSE}
# knitr::purl(
#   here::here("GitControlled/Codes/1_1_main_analysis.Rmd"),
#   here::here("GitControlled/Codes/1_1_main_analysis.R"),
# )
# source(here::here("GitControlled/Codes/1_1_main_analysis.R"))

# knitr::purl(
#   here::here("GitControlled/Codes/1_2_sub_analysis.Rmd"),
#   here::here("GitControlled/Codes/1_2_sub_analysis.R"),
# )
# source(here::here("GitControlled/Codes/1_2_sub_analysis.R"))

# knitr::purl(
#   here::here("GitControlled/Codes/2_figures_tables.Rmd"),
#   here::here("GitControlled/Codes/2_figures_tables.R"),
# )
# source(here::here("GitControlled/Codes/2_figures_tables.R"))

# knitr::purl(
#   here::here("GitControlled/Codes/3_appendix.Rmd"),
#   here::here("GitControlled/Codes/3_appendix.R"),
# )

# source(here::here("GitControlled/Codes/3_appendix.R"))
```

```{r setup-main}
#| include: false
#| cache: false

library(knitr)

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  fig.width = 5,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

# Main

Groundwater is an essential input for agricultural production, providing a critical buffer against limited or variable surface water supplies for farmers in many parts of the world [@scanlon2023global]. Globally, agriculture's dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change [@zhou2010impact; @wada2013multimodel; @wada2014sustainability; @kreins2015quantification; @florke2018water]. However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades [@wada2010global; @famiglietti2011satellites; @scanlon2012groundwater; @konikow2015long; @bierkens2019non], which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.

Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers capacity to adapt successfully to drought and rainfall variability. Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks [@schlenker2009nonlinear; @lobell2014greater; @schlenker2010robust; @zhou2020connections; @borgomeo2020impact]. Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought [@kuwayama2019estimating; @zipper2016drought; @zhu2022untangling; @zhu2022warming; @lu2020mapping; @davis2019sensitivity; @li2018changes]. In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling [@foster2015well; @cotterman2018groundwater; @kahil2015modeling; @yoon2021coupled; @rad2020mod], with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage [@jain2021groundwater; @suter2021depletion].

Two pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability [@foster2015analysis]. First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs [@mieno2017price; @bhattarai2021impact]. Second, depletion also reduces the pumping capacity of wells used to extract groundwater, which may create a physical constraint to the volume of irrigation that can be applied to a field [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous]. The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress. Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production [@foster2014modeling; @rad2020effects]. Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change [@ukkola2020robust; @chiang2021evidence; @cook2020twenty]

In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought risk using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States. Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production [@schlenker2009nonlinear; @lobell2014greater; @lu2018crop]. We also provide new empirical insights about the relative roles of extensive (i.e., irrigated area) and intensive (i.e., per-area irrigation rates) margin water use adjustments in determining drought-related production losses, which previously had only been demonstrated through theoretical modeling [@foster2014modeling; @foster2017effects; @rad2020effects]. Our findings provide important evidence of the value of preserving aquifer storage as a buffer against drought events. In doing so, our analysis adds weight to the need to improve groundwater resource conservation and sustainability to enable adaptation to climate change in groundwater-dependent agricultural systems globally [@jain2021groundwater; @scanlon2023global].

# Results

## Aquifer depletion increases sensitivity of crop yields to drought {#impact-intensive}

Figure \@ref(fig:yield-response) shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (1st row) and soybean (2nd row). In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean. As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases. This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors [@zhu2022untangling; @li2020quantifying].

Differentiating irrigated yields by the level of aquifer thickness alters these responses in two ways. First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \@ref(fig:yield-response)). This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits [@rad2020effects, @foster2014modeling; @hrozencik2017heterogeneous]. Second, at negative water deficits, counties with the second highest and highest levels of aquifer thickness (i.e., 2nd and 3rd quantile in Figure \@ref(fig:yield-response)) experience lower yields than counties with the lowest level of aquifer thickness (1st quantile), resulting in productivity close to rainfed yields. This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities [@foster2019assessing; @gibson2017case; @gibson2019benchmarking]. However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e. high deficits), in particular for corn.

```{r yield-response, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains", out.height = "5in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_intensive.pdf"))
```

## Aquifer depletion substantially reduces the share of agricultural land under irrigation {#impact-share}

While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in moderate to extreme drought conditions, the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production. One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation. Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically [@rad2020effects; @foster2014modeling; @hrozencik2017heterogeneous; @deines2020transitions], but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.

Figure \@ref(fig:ir-share) shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean. Figure \@ref(fig:ir-share) demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn. As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m. Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn [@zipper2016drought; @ruess2022irrigation].

```{r ir-share, fig.cap = "The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)", out.height = "4in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share.pdf"))
```

## Aquifer depletion increases overall drought-related production losses by up to $25\%$

The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share). Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean.

For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm). However, beyond this threshold, higher levels of water deficit result in noticeable and significant non-linear reductions to average corn production. This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower. Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e. greater than around 600mm) and are smaller in overall magnitude. For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.

```{r tot-impact, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness", out.height = "5in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_total_impact.pdf"))
```

```{r dif-tot-impact, fig.cap = "Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean", cache = FALSE}
knitr::include_graphics(here::here("Shared/Results/g_avg_yield_dif.pdf"))
```

\clearpage
# Discussion

We demonstrate the important role of aquifer thickness in determining farmers ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States. Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production [@schlenker2009nonlinear; @lobell2014greater; @schlenker2010robust; @zhou2020connections; @borgomeo2020impact] or for irrigated lands where water supply constraints are assumed to be minimal [@kuwayama2019estimating; @zipper2016drought; @zhu2022untangling; @zhu2022warming; @lu2020mapping; @davis2019sensitivity; @li2018changes; @luan2021combined]. Our analysis adds new empirical insights to this literature which show explicitly that reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events. Critically, this suggests that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.

An important implication of our findings is that depletion of groundwater resources is likely to increase significantly the vulnerability of agricultural production to current and future drought risks. Climate change is projected to increase the frequency and severity of extreme drought events in many parts of the world [@ukkola2020robust; @chiang2021evidence; @cook2020twenty], including in parts of Kansas and Texas [@bradford2020robust; @cook2022projected; @mullens2019quantitative] where the High Plains Aquifer continues to experience prolonged and significant declines in aquifer thickness [@scanlon2012groundwater; @haacker2016water; @cotterman2018groundwater]. Similar trends of increasing drought risks coupled with declining aquifer storage have been reported in other major groundwater irrigated farming systems worldwide [@wada2010global; @doll2014global; @famiglietti2014global; @feng2018groundwater; @bierkens2019non]. Our findings suggest that these simultaneous changes in aquifer conditions will amplify impacts of future climate change on crop production, with associated ramifications for resilience of supply chains, commodity markets, rural economies, and food security.

Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally. Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and drought impacts is highly non-linear when accounting for farmers decisions about irrigated production areas. This indicates that sustainable groundwater conservation efforts [@macewan2017hydroecon; @butler2020charting; @elshall2020groundwater] should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought. Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production [@foster2017effects; @deines2020transitions]. For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure  50\% of current aquifer storage remains in 50 years time (so called 50/50 rule) [@closas2018chronicle]. The effectiveness of such targets could likely be improved substantially by setting conservation targets based on empirical evidence about the level of aquifer thickness that is required to ensure adequate levels of drought protection, both now and with future climate change.

Several opportunities exist for future research to build upon the methods and analyses presented here. First, due to the limited resolution of our observational data, we were only able to estimate the impacts of different aquifer thickness categories on drought risk. Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring [@edreira2020combining; @deines2021million], would allow models of the continuous effect of aquifer thickness on drought risk to be developed. This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.

Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types. Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies [@bhattarai2021impact; @deines2019quantifying; @manning2017producer]. Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.

Finally, we focus on the impacts of aquifer thickness on agricultural resilience to seasonal water deficits. However, where water deficits are distributed unevenly during the season, negative impacts of aquifer depletion may be even larger. This is because reductions in well yields caused by aquifer depletion will constrain farmers ability to meet peak crop water demands, increasing yield losses compared with if water deficits had been distributed evenly throughout the season [@ortiz2019unpacking]. Furthermore, irrigation also has broader benefits for crops such as reducing heat-related stress through cooling of land surface temperatures [@adegoke2003impact; @bonfils2007empirical; @lobell2008effect; @zhu2022untangling]. Reductions in aquifer thickness therefore may have other negative impacts on irrigated crop productivity, if constraints to water supply due to aquifer depletion are significant enough to limit cooling benefits of irrigation during the growing season.

# Methods

## Data-sets

Our study focuses on the High Plains Aquifer (HPA) in the Central United States. Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually [@fenichel2016measuring @smidt2016complex]. To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.

County-level irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S. Department of Agriculture's National Agricultural Statistics Service (USDA-NASS). USDA-NASS does not report irrigated yields disaggregated by the source of water used for irrigation. However, it is known that groundwater is the main source of water for irrigation for counties that overlie the HPA due to the limited availability of surface water resources in this area. Therefore, we retain both rainfed and irrigated yield observations for counties where greater than 75\% of the county area overlies the boundary of the HPA. For counties that do not meet this condition, only rainfed yield observations are obtained to avoid the possibility that irrigated production may be only partly dependent on use of groundwater from the HPA. We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production.

Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset [@Abatzoglou2013]. The gridMET data is a gridded daily weather dataset with a spatial resolution of ~4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States [@abatzoglou2016impact; @pereira2015crop; @crane2018machine; @venkatappa2021impacts; @zhu2019dissecting]. Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample. Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET, and were then used to calculate daily reference crop evapotranspiration for each grid cell. From these data, we computed the seasonal water deficit (\(WB_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region. Seasonal water balance estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell. Figure \@ref(fig:deficit-yield-hist) presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012. The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \@ref(fig:deficit-yield-hist)).

```{r deficit-yield-hist, fig.cap = "Historical dryland yield, irrigated yield, and water deficit", out.height = "6in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_y_wd_all.pdf"))
```

As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA available at Hydroshare [@haacker2023]. These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935, derived based on extensive water level observations from monitoring wells and detailed maps of both land surface elevation and the bedrock elevation of the HPA. aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data. aquifer thickness values were only estimated for counties with irrigated yield observations. Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [@fenichel2016measuring; @haacker2016water; @deines2020transitions]. Figure \@ref(fig:sat-map) presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses. aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer influences farmers' exposure and responses to drought.

These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield (8773 observations), (2) per-acre soybean yield (5977 observations), (3) share of irrigated corn (2686 observations), and (4) share of irrigated soybean (2014 observations).

```{r sat-map, fig.cap = "Map showing aquifer thickness in 2016 for each county included in our corn and soybean regression analyses. Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system", out.width = "6in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_map.pdf"))
```

## Regression Models

Aquifer depletion may impact resilience of irrigated crop production to drought through two main channels: (1) a reduction in per-area yields due to the economic or physical inability to meet crop water requirements; and (2) a reduction in the area of land irrigated, resulting in reduced production due to lower yields obtained through rainfed agriculture.

To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield to drought is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year. Our regression model is given in Equation \@ref(eq:eqn1), with the same model formulation used for corn and soybean:

```{=Tex}
\begin{equation}
y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t} (\#eq:eqn1)
\end{equation}
```

where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit. \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise. The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3\%)\), \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness. The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category. Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient. To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.

To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalizez additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms. This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF). The model is given in Equation \@ref(eq:eqn2) below:

```{=Tex}
\begin{equation}
    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \theta_s + \mu_t + v_{i,t} (\#eq:eqn2)
\end{equation}
```

where \(s_{i,t}\) is the share of irrigated production. The impact of average water balance is captured by $\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})$, where $\phi_{k}(\cdot)$ is the $k$th spline basis function and $\alpha_{k}$ is its coefficient. The impact of aquifer thickness is captured in a similar way with $\tau_{l}(\cdot)$ and $\beta_{l}$ being the $l$th spline basis function and its coefficient, respectively. State fixed effects, year fixed effects and error term are represented by $\theta_s$, $\mu_t$, and $v_{i,t}$, respectively. Unlike our regression model of the impact of water balance on yield, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable. The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time. Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.

county FE

In the final step of our analysis, we then estimate the average crop yield for each county as a function of aquifer thickness and seasonal water deficit. The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \@ref(eq:eqn1), weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \@ref(eq:eqn2). Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area yields and the area of land irrigated.

## Data and Code Availability
All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}). All the raw data-sets used in this study are publicly accessible. The boundary shape file of the High Plains aquifer was obtained from the U.S. Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}. This data is used for Figure \@ref(fig:sat-map). Corn and soybean yield data were obtained from the U.S. Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}. The R tidyUSDA package was used to automatically download the relevant data. This data is used for Figure \@ref(fig:deficit-yield-hist) and the regression analysis. Weather data (precipitation, temperature, and evapotranspiration) was obtained from the gridMET dataset [@Abatzoglou2013], which is available at \url{https://www.climatologylab.org/gridmet.html}. This data is used for Figure \@ref(fig:deficit-yield-hist) and the regression analysis. The aquifer thickness data were obtained from Hydroshare [@haacker2023]. This data is used for Figure \@ref(fig:sat-map) and the regression analysis. The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.

\clearpage

# (APPENDIX) Appendix {-}

# More on the impact of water deficit on irrigated and average yield

In the main article, the confidence intervals of the estiamted yield responses to water deficit were not presented as they would crowd the figures excessively. This section provides visualization of yield response with the 95 $\%$ confidence interval to convery the degree of uncertainty in our estimation. Figures \@ref(fig:irrigated-yield-ind-corn) and \@ref(fig:irrigated-yield-ind-soy) present the impact of water deficit on crop yield by aquifer thickness category with the 95$\%$ confidence interval for irrigated corn and soybean, respectively. Figure \@ref(fig:avg-yield-ind) presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95$\%$ confidence interval.

```{r irrigated-yield-ind-corn, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
knitr::include_graphics(here::here("Shared/Results/g_yield_with_conf_corn.pdf"))
```

```{r irrigated-yield-ind-soy, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of soybean in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
knitr::include_graphics(here::here("Shared/Results/g_yield_with_conf_soy.pdf"))
```

```{r avg-yield-ind, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (The shaded areas represent the 95 percent confidence interval)", out.width = "6.5in", cache = FALSE}
knitr::include_graphics(here::here("Shared/Results/g_avg_yield.rds"))
```

\clearpage

# County Fixed Effects in Irrigation Share Regression

```{=Tex}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A.\arabic{figure}}
```

Figure \@ref(fig:variation-left) presents the histogram of aquifer thickness after demeaning them by its overall average (a), state average (b), and county average (c) to gain insights into how much variation for identification is left under no county or state fixed effects (a), state fixed effects (b), and county fixed effects (c). As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation. Consequently, we found that including county fixed effects results in extremely inaccurate estimates (high standard error). For this reason, we just included state fixed effects.

```{r variation-left, fig.cap = "Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.", out.height = "6in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_variation.pdf"))
```

Figure \@ref(fig:state-fe-with-without) compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean. For corn, the negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included. Therefore, it may be that there are important unobservables that are spatially correlated with saturated thickness, but their correlation may not be very strong. Given the direction of potential bias, providing tighter controls with county fixed effects may result in even more significant drop in the share of irrigated acres in response to a reduction in aquifer thickness. That is we may be underestimating the true negative impact of a reduction in aquifer thickness.

```{r state-fe-with-without, fig.cap = "Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects", out.height = "6in"}
knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share_comp.pdf"))
```

# Spline basis functions and GAM estimation

```{=Tex}
\setcounter{figure}{0}
\renewcommand{\thefigure}{B.\arabic{figure}}
```

We used generalized additive models (GAM) for our regression analysis. In GAM, the impact of a variable on the dependent variable can be expressed as a series of spline basis functions and their coefficients. Suppose you are interested in modeling the following model.

```{=Tex}
\begin{equation}
y = f(x) + \mu
\end{equation}
```

Where $f(x)$ represents the impact of variable $x$ on $y$. The shape of $f(x)$ is not known to the researcher. GAM attemps to represent $f(x)$ with a series of spline basis functions and their coefficients like below.

```{=Tex}
\begin{equation}
y = \sum_{k=1}^K \beta_k \phi_k(x) + \mu
\end{equation}
```

Where $\phi_k(\cdot)$ is the $k$th spline basis function and $\beta_k$ is its coefficient. Figure \@ref(fig:sp-basis) provides a visual illustration of a series of thin-plate spline basis functions.

```{r sp-basis, fig.cap = "Example of a series of thin-plate spline basis functions", out.height = "4in", out.width = "6in"}
library(mgcv)

set.seed(83944)

# === generate data ===#
N <- 400 # number of observations
x <- seq(1, 250, length = N)
y_det <- 240 * (1 - 0.4 * exp(-0.03 * x))
e <- 25 * rnorm(N) # error
data <- data.table(x = x, y = y_det + e, y_det = y_det)

basis_data <-
  gam(y_det ~ s(x, k = 5, bs = "cr"), data = data) %>%
  predict(., type = "lpmatrix") %>%
  data.table() %>%
  .[, x := data[, x]] %>%
  melt(id.var = "x") %>%
  .[, variable := case_when(
    variable == "(Intercept)" ~ "Intercept",
    variable == "s(x).1" ~ "1st basis",
    variable == "s(x).2" ~ "2nd basis",
    variable == "s(x).3" ~ "3rd basis",
    variable == "s(x).4" ~ "4th basis"
  )] %>%
  .[, variable := factor(
    variable,
    levels= c("Intercept", "1st basis", "2nd basis", "3rd basis","4th basis")
  )]

ggplot(data = basis_data) +
  geom_line(aes(y = value, x = x)) +
  facet_grid(variable ~ .) +
  theme_bw() +
  ylab("")

```

Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of $x$ on $y$. Figure \@ref(fig:ex-fx) presents two examples of two different sets of coefficients and their resulting represention of $f(x)$. In the first case, $\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}$. In the second case, $\{\beta_1, \dots, \beta_5\} = \{50, 20, 40, 60, 80\}$. Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best. Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate $\beta_k$ to fit the data best.

```{r ex-fx, fig.cap = "Example of a function represented by different coefficient values", out.height = "4in", out.width = "6in"}
library(patchwork)

g_il_1 <-
  data.table(
    variable = unique(basis_data$variable),
    coef = rep(60, 5)
  ) %>%
  .[basis_data, on = "variable"] %>%
  .[, sum(coef * value), by = x] %>%
  ggplot(data = .) +
  geom_line(aes(y = V1, x = x)) +
  theme_bw() +
  ylab("y") +
  ggtitle("Case 1")

g_il_2 <-
  data.table(
    variable = unique(basis_data$variable),
    coef = c(50, 20, 40, 60, 80)
  ) %>%
  .[basis_data, on = "variable"] %>%
  .[, sum(coef * value), by = x] %>%
  ggplot(data = .) +
  geom_line(aes(y = V1, x = x)) +
  theme_bw() +
  ylab("y")  +
  ggtitle("Case 2")

g_il_1 / g_il_2
```

\clearpage

### 2023-07-21 Writing/manuscript.tex
Response to reviewers added

--- 2023-07-21 Writing/NatureWater/1stRoundRevision/revised_manuscript/manuscript_DR (LAGAE-6N57C7R9D's conflicted copy 2023-07-19).tex
@@
 \setcounter{secnumdepth}{5}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+\usepackage{lineno}
+\linenumbers
 \ifLuaTeX
@@
 \hypersetup{
-  pdftitle={The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather},
+  pdftitle={Aquifer depletion exacerbates agricultural drought losses in the US High Plains},
   pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
@@
 \title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
-% \author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of Mechanical, Aerospace and Civil Engineering, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
+\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
 \date{}
@@
 \maketitle
-
-% \centering{\textbf{This paper is a non-peer reviewed preprint submitted to EarthArXiv.}}
-
-% \vspace{3cm}
-
 \begin{abstract}
@@
 
-% \vspace{3cm}
-
-% \textbf{Acknowledgement}: Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.
-
-% \newpage
-
-\hypertarget{Main Text}{%
-\section{Main Text}\label{Main Text}}
+\hypertarget{main}{%
+\section{Main}\label{main}}
 
 Groundwater is an essential input for agricultural production, providing a critical buffer against limited or variable surface water supplies for farmers in many parts of the world \citep{scanlon2023global}.
-Globally, agriculture’s dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change \citep{zhou2010impact, wada2013multimodel, wada2014sustainability, kreins2015quantification, florke2018water}.
-However, many of the world’s most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}, which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
+Globally, agriculture's dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change \citep{zhou2010impact, wada2013multimodel, wada2014sustainability, kreins2015quantification, florke2018water}.
+However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}, which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
 
 Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers capacity to adapt successfully to drought and rainfall variability.
-Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear,lobell2014greater,schlenker2010robust,zhou2020connections,borgomeo2020impact}.
-Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought \citep{kuwayama2019estimating,zipper2016drought,zhu2022untangling,zhu2022warming,lu2020mapping,davis2019sensitivity,li2018changes}.
-In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling \citep{foster2015well,cotterman2018groundwater,kahil2015modeling,yoon2021coupled,rad2020mod}, with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage \citep{jain2021groundwater,suter2021depletion}.
+Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact}.
+Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes}.
+In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling \citep{foster2015well, cotterman2018groundwater, kahil2015modeling, yoon2021coupled, rad2020mod}, with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage \citep{jain2021groundwater, suter2021depletion}.
 
-Two pathways exist by which aquifer depletion will impact farmers’ ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
+Two pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
@@
 Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production \citep{foster2014modeling, rad2020effects}.
-Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust,chiang2021evidence,cook2020twenty}
+Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}
 
 In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought risk using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
-Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater,lu2018crop}.
-We also provide new empirical insights about the relative roles of extensive (i.e., irrigated area) and intensive (i.e., per-area irrigation rates) margin water use adjustments in determining drought-related production losses, which previously had only been demonstrated through theoretical modeling \citep{foster2014modeling,foster2017effects,rad2020effects}.
+Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater, lu2018crop}.
+We also provide new empirical insights about the relative roles of extensive (i.e., irrigated area) and intensive (i.e., per-area irrigation rates) margin water use adjustments in determining drought-related production losses, which previously had only been demonstrated through theoretical modeling \citep{foster2014modeling, foster2017effects, rad2020effects}.
 Our findings provide important evidence of the value of preserving aquifer storage as a buffer against drought events.
-In doing so, our analysis adds weight to the need to improve groundwater resource conservation and sustainability to enable adaptation to climate change in groundwater-dependent agricultural systems globally \citep{jain2021groundwater,scanlon2023global}.
-
-%Add troy2015impact and li2018changes somewhere.
+In doing so, our analysis adds weight to the need to improve groundwater resource conservation and sustainability to enable adaptation to climate change in groundwater-dependent agricultural systems globally \citep{jain2021groundwater, scanlon2023global}.
 
@@
 
-\hypertarget{the-impact-of-drought-on-rainfed-and-irrigated-yield}{%
-\subsection{Aquifer depletion increases sensitivity of crop yields to drought}\label{the-impact-of-drought-on-rainfed-and-irrigated-yield}}
+\hypertarget{impact-intensive}{%
+\subsection{Aquifer depletion increases sensitivity of crop yields to drought}\label{impact-intensive}}
 
@@
 As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases.
-This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling,li2020quantifying}.
+This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling, li2020quantifying}.
 
@@
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous}.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
 Second, at negative water deficits, counties with the second highest and highest levels of aquifer thickness (i.e., 2nd and 3rd quantile in Figure \ref{fig:yield-response}) experience lower yields than counties with the lowest level of aquifer thickness (1st quantile), resulting in productivity close to rainfed yields.
-This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing,gibson2017case,gibson2019benchmarking}.
-However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e.
-high deficits), in particular for corn.
+This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}.
+However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e.~high deficits), in particular for corn.
 
@@
 
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_yield_intensive.pdf}
+{\centering \includegraphics[height=5in]{Figures/g_yield_intensive}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:yield-response}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains}\label{fig:yield-response}
 \end{figure}
 
-\hypertarget{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}{%
-\subsection{Aquifer depletion substantially reduces the share of agricultural land under irrigation}\label{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}}
+\hypertarget{impact-share}{%
+\subsection{Aquifer depletion substantially reduces the share of agricultural land under irrigation}\label{impact-share}}
 
@@
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous,deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.
+Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.
 
@@
 As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m.
-Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn \citep{zipper2016drought,ruess2022irrigation}.
+Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn \citep{zipper2016drought, ruess2022irrigation}.
 
 \begin{figure}
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_share.pdf} 
+
+{\centering \includegraphics[height=4in]{Figures/g_share}
+
 }
-\caption{The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:ir-share}
+
+\caption{The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)}\label{fig:ir-share}
 \end{figure}
 
-\hypertarget{total-impact-of-decline-in-saturated-thickness}{%
-\subsection{Aquifer depletion increases overall drought-related production losses by up to 25\%}\label{total-impact-of-decline-in-saturated-thickness}}
+\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}{%
+\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
 
-The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \ref{the-impact-of-drought-on-rainfed-and-irrigated-yield} and \ref{the-impact-of-saturated-thickness-on-the-share-of-irrigated-production}.
+The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
 Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean.
@@
 For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm).
-However, beyond this threshold, higher levels of deficit result in noticeable and significant non-linear reductions to average corn production.
+However, beyond this threshold, higher levels of water deficit result in noticeable and significant non-linear reductions to average corn production.
 This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.
-greater than around 600mm) and are smaller in overall magnitude.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.~greater than around 600mm) and are smaller in overall magnitude.
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
 
-%It is also noteworthy that the impact of aquifer thickness on crop yield in drought years is non-linear.
-Specifically, according to our estimates, the decline of aquifer thickness from 186 meters to 75 meters would cause about one tonne/ha of damage in corn yield at water deficit of 1100mm.
-However, the decline of aquifer thickness from 75 meters to 30 meters would cause about 1.5 tonne/ha of damage in corn yield.
-This means that the negative impact of a meter of decline in aquifer thickness becomes larger as the aquifer depletion progresses.
-This has an important policy implication.
-Suppose we intend to limit groundwater pumping to maintain the stable level of groundwater stock by matching the natural replenishment rate for the purpose of sustainable groundwater-irrigated production.
-Then, we need to act early before aquifer thickness becomes too low so that we do not need to product at a less productive steady state.
-The longer you wait, the harsher damage is going to be when you finally decide to place a policy that limits groundwater pumping.
+\begin{figure}
 
-\begin{figure}
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_total_impact.pdf} 
+{\centering \includegraphics[height=5in]{Figures/g_total_impact}
+
 }
@@
 
+\begin{figure}
+
+{\centering \includegraphics{../../Shared/Results/g_avg_yield_dif}
+
+}
+
+\caption{Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean}\label{fig:dif-tot-impact}
+\end{figure}
+
+\clearpage
+
 \hypertarget{discussion}{%
@@
 We demonstrate the important role of aquifer thickness in determining farmers ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States.
-Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear,lobell2014greater,schlenker2010robust,zhou2020connections,borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating,zipper2016drought,zhu2022untangling,zhu2022warming,lu2020mapping,davis2019sensitivity,li2018changes,luan2021combined}.
+Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes, luan2021combined}.
 Our analysis adds new empirical insights to this literature which show explicitly that reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
@@
 An important implication of our findings is that depletion of groundwater resources is likely to increase significantly the vulnerability of agricultural production to current and future drought risks.
-Climate change is projected to increase the frequency and severity of extreme drought events in many parts of the world \citep{ukkola2020robust,chiang2021evidence,cook2020twenty}, including in parts of Kansas and Texas \citep{bradford2020robust,cook2022projected,mullens2019quantitative} where the High Plains Aquifer continues to experience prolonged and significant declines in aquifer thickness \citep{scanlon2012groundwater, haacker2016water, cotterman2018groundwater}.
+Climate change is projected to increase the frequency and severity of extreme drought events in many parts of the world \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}, including in parts of Kansas and Texas \citep{bradford2020robust, cook2022projected, mullens2019quantitative} where the High Plains Aquifer continues to experience prolonged and significant declines in aquifer thickness \citep{scanlon2012groundwater, haacker2016water, cotterman2018groundwater}.
 Similar trends of increasing drought risks coupled with declining aquifer storage have been reported in other major groundwater irrigated farming systems worldwide \citep{wada2010global, doll2014global, famiglietti2014global, feng2018groundwater, bierkens2019non}.
@@
 Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and drought impacts is highly non-linear when accounting for farmers decisions about irrigated production areas.
-This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon,butler2020charting,elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
-Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production \citep{foster2017effects,deines2020transitions}.
+This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon, butler2020charting, elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
+Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production \citep{foster2017effects, deines2020transitions}.
 For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure 50\% of current aquifer storage remains in 50 years time (so called 50/50 rule) \citep{closas2018chronicle}.
@@
 First, due to the limited resolution of our observational data, we were only able to estimate the impacts of different aquifer thickness categories on drought risk.
-Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining,deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
+Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining, deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
 This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
@@
 Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types.
-Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact,deines2019quantifying,manning2017producer}.
+Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact, deines2019quantifying, manning2017producer}.
 Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
@@
 
-% \hypertarget{conclusion}{%
-% \section{Conclusion}\label{conclusion}}
-
-% In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events.
-Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer thickness significantly reduces the overall productivity of both corn and soybean during drought events.
-Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion.
-Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.
-
 \hypertarget{methods}{%
@@
 
-\hypertarget{methods-data}{%
-\subsection{Data-sets}\label{methods-data}}
+\hypertarget{data-sets}{%
+\subsection{Data-sets}\label{data-sets}}
 
 Our study focuses on the High Plains Aquifer (HPA) in the Central United States.
-Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{fenichel2016measuring,smidt2016complex}.
+Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep[ \citet{smidt2016complex}]{fenichel2016measuring}.
 To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.
@@
 Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset \citep{Abatzoglou2013}.
-The gridMET data is a gridded daily weather dataset with a spatial resolution of ~4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
+The gridMET data is a gridded daily weather dataset with a spatial resolution of \textasciitilde4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
 Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample.
@@
 Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
-The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \ref{fig:deficit-yield-hist})
+The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \ref{fig:deficit-yield-hist}).
 
 \begin{figure}
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_y_wd_all.png} }
+
+{\centering \includegraphics[height=6in]{Figures/g_y_wd_all}
+
+}
+
 \caption{Historical dryland yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
@@
 
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA reported in \citet{haacker2016water}.
+As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA available at Hydroshare \citep{haacker2023}.
 These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935, derived based on extensive water level observations from monitoring wells and detailed maps of both land surface elevation and the bedrock elevation of the HPA.
@@
 aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring,haacker2016water,deines2020transitions}.
-Figure \ref{fig:sat-dist} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
+Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
+Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
 aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer influences farmers' exposure and responses to drought.
@@
 \begin{figure}
-{\centering \includegraphics[width=432px]{manuscript_DR_files/figure-latex/g_map.png}}
+
+{\centering \includegraphics[width=6in]{Figures/g_map}
+
+}
+
 \caption{Map showing aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
-Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-dist}
+Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
 \end{figure}
 
-\hypertarget{methods-models}{%
-\subsection{Regression Models}\label{methods-models}}
+\hypertarget{regression-models}{%
+\subsection{Regression Models}\label{regression-models}}
 
@@
 To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield to drought is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
-Our regression model is given in Equation \ref{equation1}, with the same model formulation used for corn and soybean:
+Our regression model is given in Equation \eqref{eq:eqn1}, with the same model formulation used for corn and soybean:
 
-\begin{equation}\label{equation1}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t}
+\begin{equation}
+y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalizez additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
 This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).
-The model is given in Equation \ref{equation2} below:
+The model is given in Equation \eqref{eq:eqn2} below:
 
-\begin{equation}\label{equation2}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \sum_{h}^H \gamma_{h}\rho_{h}(WB_{i}, AT_{i,t}) + \mu_t + v_{i,t}
+\begin{equation}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \theta_s + \mu_t + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water balance is captured by $\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})$, where $\phi_{k}(\cdot)$ is the $k$th spline basis function and $\alpha_{k}$ is its coefficient.
-The impact of aquifer thickness is captured in a similar way with $\tau_{l}(\cdot)$ and $\beta_{l}$ being the $l$th spline basis function and its coefficient, respectively.
-The interactive impact of average water balance and aquifer thickness is modeled by $\sum_{h}^H \gamma_{h}\rho_{h}(WB_{i}, AT_{i,t})$.
-Here, $\rho_{h}(\cdot)$ is a bi-variate spline basis function and $\gamma_{h}$ is its coefficient.
-Year fixed effects and error term are represented by $\mu_t$ and $v_{i,t}$, respectively.
+The impact of average water balance is captured by \(\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\alpha_{k}\) is its coefficient.
+The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
+State fixed effects, year fixed effects and error term are represented by \(\theta_s\), \(\mu_t\), and \(v_{i,t}\), respectively.
 Unlike our regression model of the impact of water balance on yield, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
@@
 
+county FE
+
 In the final step of our analysis, we then estimate the average crop yield for each county as a function of aquifer thickness and seasonal water deficit.
-The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \ref{equation1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \ref{equation2}.
+The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area yields and the area of land irrigated.
 
-\hypertarget{data-code-availability}{%
-\subsection{Data and Code Availability}\label{data-availability}}
+\hypertarget{data-and-code-availability}{%
+\subsection{Data and Code Availability}\label{data-and-code-availability}}
+
 All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
@@
 Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}.
-This data is used for Figure \ref{fig:sat-dist}.
+This data is used for Figure \ref{fig:sat-map}.
 Corn and soybean yield data were obtained from the U.S.
-Department of Agriculture’s National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
+Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
 The R tidyUSDA package was used to automatically download the relevant data.
@@
 This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-The aquifer thickness data were obtained directly from Dr.
-Erin Haacker (\href{mailto:ehaacker2@unl.edu}, who can be reached at {\nolinkurl{ehaacker2@unl.edu}}).
-This data is used for Figure \ref{fig:sat-dist} and the regression analysis.
+The aquifer thickness data were obtained from Hydroshare \citep{haacker2023}.
+This data is used for Figure \ref{fig:sat-map} and the regression analysis.
 The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.
@@
 
-%\hypertarget{appendix-appendix}{
-%\appendix}
+\hypertarget{appendix-appendix}{%
+\appendix}
 
-%\hypertarget{use-spei-in-place-of-water-deficit}{%
-%\section{Use SPEI in place of water deficit}\label{use-spei-in-place-of-water-deficit}}
+\hypertarget{more-on-the-impact-of-water-deficit-on-irrigated-and-average-yield}{%
+\section{More on the impact of water deficit on irrigated and average yield}\label{more-on-the-impact-of-water-deficit-on-irrigated-and-average-yield}}
 
-%Some studies use SPEI to represent drought severity. We also considered using SPEI, but the regressions with water deficit provides a better fit for both corn and soybean (see \ref{tab:fit-reg}).
+In the main article, the confidence intervals of the estiamted yield responses to water deficit were not presented as they would crowd the figures excessively.
+This section provides visualization of yield response with the 95 \(\%\) confidence interval to convery the degree of uncertainty in our estimation.
+Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\(\%\) confidence interval for irrigated corn and soybean, respectively.
+Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\(\%\) confidence interval.
 
-%\begin{table}[!h]
+\begin{figure}
 
-%\caption{\label{tab:fit-reg}Comparison of R-squared of the regressions with water deficit and SPEI for corn and soybean}
-%\centering
-%\begin{tabular}[t]{lrr}
-%\toprule
-%Crop & SPEI & Water Deficit\\
-%\midrule
-%Corn & 0.899 & 0.903\\
-%Soybean & 0.840 & 0.846\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
+{\centering \includegraphics[width=6in]{../../Shared/Results/g_yield_with_conf_corn}
 
-%Qualitatively, the results with SPEI are similar to the regression results with water deficit as can be seen in \ref{fig:spei}.
+}
 
-%\begin{figure}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-corn}
+\end{figure}
 
-%{\centering \includegraphics[width=360px]{manuscript_DR_files/figure-latex/spei-1}
+\begin{figure}
 
-%}
+{\centering \includegraphics[width=6in]{../../Shared/Results/g_yield_with_conf_soy}
 
-%\caption{The impact of SPEI on crop yield for corn and soybean}\label{fig:spei}
-%\end{figure}
+}
+
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of soybean in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-soy}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=6.5in]{../../Shared/Results/g_avg_yield}
+
+}
+
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (The shaded areas represent the 95 percent confidence interval)}\label{fig:avg-yield-ind}
+\end{figure}
+
+\clearpage
+
+\hypertarget{county-fixed-effects-in-irrigation-share-regression}{%
+\section{County Fixed Effects in Irrigation Share Regression}\label{county-fixed-effects-in-irrigation-share-regression}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A.\arabic{figure}}
+
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its overall average (a), state average (b), and county average (c) to gain insights into how much variation for identification is left under no county or state fixed effects (a), state fixed effects (b), and county fixed effects (c).
+As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation.
+Consequently, we found that including county fixed effects results in extremely inaccurate estimates (high standard error).
+For this reason, we just included state fixed effects.
+
+\begin{figure}
+
+{\centering \includegraphics[height=6in]{Figures/g_variation}
+
+}
+
+\caption{Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.}\label{fig:variation-left}
+\end{figure}
+
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean.
+For corn, the negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included.
+Therefore, it may be that there are important unobservables that are spatially correlated with saturated thickness, but their correlation may not be very strong.
+Given the direction of potential bias, providing tighter controls with county fixed effects may result in even more significant drop in the share of irrigated acres in response to a reduction in aquifer thickness.
+That is we may be underestimating the true negative impact of a reduction in aquifer thickness.
+
+\begin{figure}
+
+{\centering \includegraphics[height=6in]{Figures/g_share_comp}
+
+}
+
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}\label{fig:state-fe-with-without}
+\end{figure}
+
+\hypertarget{spline-basis-functions-and-gam-estimation}{%
+\section{Spline basis functions and GAM estimation}\label{spline-basis-functions-and-gam-estimation}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{B.\arabic{figure}}
+
+We used generalized additive models (GAM) for our regression analysis.
+In GAM, the impact of a variable on the dependent variable can be expressed as a series of spline basis functions and their coefficients.
+Suppose you are interested in modeling the following model.
+
+\begin{equation}
+y = f(x) + \mu
+\end{equation}
+
+Where \(f(x)\) represents the impact of variable \(x\) on \(y\).
+The shape of \(f(x)\) is not known to the researcher.
+GAM attemps to represent \(f(x)\) with a series of spline basis functions and their coefficients like below.
+
+\begin{equation}
+y = \sum_{k=1}^K \beta_k \phi_k(x) + \mu
+\end{equation}
+
+Where \(\phi_k(\cdot)\) is the \(k\)th spline basis function and \(\beta_k\) is its coefficient.
+Figure \ref{fig:sp-basis} provides a visual illustration of a series of thin-plate spline basis functions.
+
+\begin{figure}
+
+{\centering \includegraphics[width=6in,height=4in]{manuscript_files/figure-latex/sp-basis-1}
+
+}
+
+\caption{Example of a series of thin-plate spline basis functions}\label{fig:sp-basis}
+\end{figure}
+
+Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of \(x\) on \(y\).
+Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting represention of \(f(x)\).
+In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\).
+In the second case, \(\{\beta_1, \dots, \beta_5\} = \{50, 20, 40, 60, 80\}\).
+Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
+Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate \(\beta_k\) to fit the data best.
+
+\begin{figure}
+
+{\centering \includegraphics[width=6in,height=4in]{manuscript_files/figure-latex/ex-fx-1}
+
+}
+
+\caption{Example of a function represented by different coefficient values}\label{fig:ex-fx}
+\end{figure}
 

### 2023-07-21 Codes/1_A3_feols_smooth_exp.R

@@ -1,205 +1,0 @@
-
-N <- 300
-g_N <- (N + 1) / 5
-T <- 10
-
-data <- CJ(id = 1:N, t = 1:T) %>% 
-.[, ind_fe := runif(1), by = id] %>% 
-.[, time_fe := runif(1), by = t] %>% 
-.[, x := runif(1) + ind_fe, by = .(id, t)] %>% 
-.[, group := factor(floor((id) / g_N))] %>% 
-.[, beta_group := rnorm(1), by = group] %>% 
-.[, phi_group := rnorm(1), by = group] %>% 
-.[, alpha_group := rnorm(1) / 2, by = group] %>% 
-.[, mu := rnorm(nrow(.))] %>% 
-.[, e_y := 2 + beta_group * x + phi_group * (x - alpha_group)^2] %>% 
-.[, y := e_y + mu]  
-
-
-gam_f <- formula(y ~ s(x, by = group, k = 5, m = 2))
-
-reg_data <- gen_reg_data(data, gam_formula = gam_f)
-
-# cor(reg_data$model_X)
-
-formula_feols <- 
-  names(reg_data$model_X) %>%
-  paste0(., collapse = "+") %>% 
-  paste0("y ~ ", ., " | group + t") %>% 
-  formula()
-
-fe_res <- 
-  feols(
-    formula_feols, 
-    cluster = ~ group, 
-    data = reg_data$full_data
-  )
-
-pred_data <- 
-  reg_data$full_data %>% 
-  .[, y_hat := predict(fe_res, newdata = .)]
-
-ggplot(data) +
-  geom_point(aes(y = y, x = x), size = 0.3) +
-  geom_point(data = pred_data, aes(y = y_hat, x = x), col = "red", size = 0.4) +
-  geom_line(aes(y = e_y, x = x), size = 0.5, color = "blue") +
-  facet_grid(. ~ group)
-
-gen_reg_data <- function(data, gam_formula) {
-  
-  temp_gam_reg <- 
-    gam(
-      gam_formula, 
-      data = data, 
-      method = "REML",
-      fit = FALSE
-    )
-  
-  model_X <- temp_gam_reg$X %>% 
-    data.table() %>% 
-    setnames(names(.), temp_gam_reg$term.names) %>% 
-    .[, `(Intercept)` := NULL] %>% 
-    setnames(names(.), gsub("\\(", "_", names(.))) %>% 
[…]

### 2023-07-21 Codes/1_A_2_effective_rainfall.R

@@ -1,80 +1,0 @@
-######################################
-# Regression Analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(here)
-library(ggplot2)
-library(sf)
-library(broom)
-library(raster)
-library(RpackageTM)
-library(mgcv)
-library(fixest)
-library(gratia)
-library(data.table)
-library(lfe)
-library(fixest)
-library(stargazer)
-
-source(here("GitControlled/Codes/1_0_fcn_reg.R"))
-
-#/*=================================================*/
-#' # Prepare data
-#/*=================================================*/
-gm_data <- readRDS(here("Shared/Data/ProcessedData", "gridMET.rds"))
-
-d_data <- gm_data %>% 
-  .[, d_balance := et0 - prcp] %>% 
-  .[, too_low := fifelse(d_balance < -50, 1, 0)] %>% 
-  .[, too_low_num := sum(too_low), by = .(sc_code, year(date))] %>% 
-  .[too_low_num == 0 , ]
-
-d_smooth_data <- gen_smooth_data(d_data, formula(sc_code ~ s(d_balance, k = 7)))
-
-saveRDS(d_smooth_data, "Shared/Data/ProcessedData/d_smooth_weather.rds")
-
-s_vars <- names(d_smooth_data$model_X)
-
-ds_data <- d_smooth_data$full_data[, lapply(.SD, sum), by = .(sc_code, year(date)), .SDcols = s_vars]
-
-
-temp_data <- 
-  copy(reg_data)[crop_type == "corn", ] %>% 
-  .$reg_data_y %>% 
-  .[[1]] %>% 
-  .[ir == "nir", ] %>% 
-  ds_data[., on = c("sc_code", "year")]
-
-ds_formula <- 
-  paste0(
-    "yield ~ ",
-    paste0(s_vars, collapse = " + "),
-    "| sc_code + year"
-  ) %>% 
-  formula()
-
-ds_fe <- feols(ds_formula, data = temp_data)   
-
-pred_data <- 
-  d_smooth_data$full_data %>% 
-  .[order(d_balance), ] %>% 
-  .[seq(1, nrow(.), length = 1000), ] %>% 
-  .[, year := 2009] %>% 
[…]

### 2023-07-21 Codes/1_A_DMLOF.R

@@ -1,185 +1,0 @@
-######################################
-# Regression Analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-# /*=================================================*/
-#' # Preparation
-# /*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(here)
-library(ggplot2)
-library(sf)
-library(broom)
-library(raster)
-library(RpackageTM)
-library(mgcv)
-library(fixest)
-library(gratia)
-library(data.table)
-library(lfe)
-library(fixest)
-library(scales)
-library(stargazer)
-
-source(here("GitControlled/Codes/Functions/1_0_functions.R"))
-
-# /*=================================================*/
-#' # Prepare data
-# /*=================================================*/
-
-#+ read data
-final_data <-
-  here("Shared/Data/ProcessedData/final_data.rds") %>%
-  readRDS()
-
-# reg_data$reg_data_y[[1]][, .(balance, et0, gdd, edd)] %>% cor()
-
-#' Data used for regression analysis
-reg_data <-
-  prepare_reg_data(
-    data = final_data,
-    sat_thld = 9, # saturated thickness has to be at least 12 meters
-    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
-    balance_thld = c(-1200, 200),
-    state_ls =
-      list(
-        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
-        soy = c("KS", "NE", "TX")
-      )
-  ) %>%
-  data.table()
-
-# ggplot(data = data_y) +
-#   geom_point(aes(y = yield, x = balance, color = sat_cat_i)) +
-#   geom_smooth(aes(y = yield, x = balance, color = sat_cat_i))
-
-# ggplot(data = data_y) +
-#   geom_point(aes(y = yield, x = spei, color = sat_cat_i)) +
-#   geom_smooth(aes(y = yield, x = spei, color = sat_cat_i))
-
-# /*=================================================*/
-#' # Regression (DML-OF)
-# /*=================================================*/
-# source DML-OF-c
[…]

### 2023-07-21 Codes/1_A_power_of_irrigation.R

@@ -1,133 +1,0 @@
-######################################
-# Regression Analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(here)
-library(ggplot2)
-library(sf)
-library(broom)
-library(raster)
-library(RpackageTM)
-library(mgcv)
-library(fixest)
-library(gratia)
-library(data.table)
-library(lfe)
-library(fixest)
-library(scales)
-library(stargazer)
-
-source(here("GitControlled/Codes/Functions/1_0_functions.R"))
-
-#/*=================================================*/
-#' # Prepare data
-#/*=================================================*/
-
-#+ read data
-final_data <-
-  here("Shared/Data/ProcessedData/final_data.rds") %>%
-  readRDS()
-
-# reg_data$reg_data_y[[1]][, .(balance, et0, gdd, edd)] %>% cor()
-
-#' Data used for regression analysis
-reg_data <-
-  prepare_reg_data(
-    data = final_data,
-    sat_thld = 9, # saturated thickness has to be at least 12 meters
-    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
-    balance_thld = c(-1200, 200),
-    state_ls =
-      list(
-        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
-        soy = c("KS", "NE", "TX")
-      )
-  ) %>%
-  data.table()
-
-data_y <- reg_data[crop_type == "corn", reg_data_y][[1]]
-
-within_data <- data_y[, lapply(.SD, function(x) x - mean(x)), by = .(sc_code, sat_cat == "dryland"), .SDcols = c("prcp", "et0", "gdd", "edd", "balance", "yield")]
-
-
-ir_data_y <- data_y[sat_cat != "dryland", ]
-
-semi_formula <-
-  formula(
-    yield
-    ~ s(balance, k = 4, m = 2)
-    + s(gdd, k = 4, m = 2)
-    + s(edd, k = 4, m = 2)
-  )
-
-semi_res <-
-  semi_ols(
-    semi_formula = semi_formula,
[…]

### 2023-07-21 Codes/renv.R

@@ -1,0 +1,2 @@
+install.packages("httgpd")
+renv::snapshot()

### 2023-07-21 Codes/spinned/0_4_ndmc_di.R

@@ -30,8 +30,8 @@ drought_county <- DI %>%
   .[, state_code := str_sub(FIPS, 1, 2)] %>%
   .[, county_code := str_sub(FIPS, 3, 5)] %>%
   .[month(end_date) > 4 & month(start_date) < 10, ] %>%
-  # === state-county code ===#
-  .[, sc_code := paste0(state_code, county_code)] %>%
+  #=== state-county code ===#
+  .[, sc_code := paste0(state_code, county_code)] %>% 
   .[, .(
     d0_5_9 = sum(D0) / 100,
     d1_5_9 = sum(D1) / 100,
@@ -42,3 +42,5 @@ drought_county <- DI %>%
   data.table()
 
 saveRDS(drought_county, here("Data/ProcessedData/DI_NDMC.rds"))
+
+

### 2023-07-22 Codes/0_1_hpa_counties.Rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Prepare base-county data"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Identify counties the overlaps with the HPA boundary
## Read the HPA border shapefile

```{r Read-HPA, cache = TRUE}
hpa <- st_read(here("Shared/Data/RawData/hp_bound2010.shp"))
```

Map-of-HPA

```{r }
tm_shape(st_simplify(hpa)) +
  tm_polygons()
```

## Get all the US counties as an sf

```{r Get-US-counties}
(
US_county <- tigris::counties(cb = TRUE) %>%
  st_as_sf() %>%
  st_transform(st_crs(hpa)) %>%
  rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
  dplyr::select(state_code, county_code)
)
```

## Identify counties that overlaps with `hpa`

```{r Identify-overlap, cache = TRUE}
hpa_counties <- US_county[hpa, ] %>%
  mutate(whole_area = st_area(.) %>% as.numeric())
```

Map-of-HPA and the intersecting counties

```{r }
tm_shape(st_simplify(hpa)) +
  tm_polygons() +
tm_shape(hpa_counties) +
  tm_polygons(col = "blue", alpha = 0.4)
```

## Find the area of the portion that is intersecting with hpa
Intersect HPA and counties in HPA

```{r intersect-HPA-county}
intersected_portion <- st_intersection(hpa, hpa_counties) %>%
  #=== get the intersecting area ===#
  mutate(area = st_area(.) %>% as.numeric()) %>%
  data.table() %>%
  .[, .(interesected_area = sum(area)), by = .(state_code, county_code)]
```

Merge the intersection data

```{r }
hpa_counties <- 
  left_join(
    hpa_counties,
    intersected_portion,
    by = c("state_code", "county_code")
  ) %>%
  #=== find the proportion of the county area that overlaps with HPA ===#
  mutate(ir_area_ratio = interesected_area / whole_area) %>%
  data.table() %>%
  .[, geometry := NULL]
```

Histogram of coverage ratio

```{r }
hpa_counties$ir_area_ratio %>% hist
```

Get the list of the states

```{r }
hpa_states <- hpa_counties$state_code %>% unique()
```

# Additional counties 
Additional state codes:
these supplement the dryland production data for understanding the impact of drought severity
on dryland crop production

```{r }
data(fips_codes)

additional_states <- fips_codes %>%
  data.table() %>%
  unique(by = "state_name") %>%
  .[state_name %in% c("Illinois", "Iowa", "Missouri", "Indiana"), ] %>%
  .[, state_code]

all_states <- c(hpa_states, additional_states)
```

# All counties
**Get the state abbreviation names**

```{r }
fips_codes <- fips_codes %>%
  data.table() %>%
  .[, .(state, state_code, state_name, county_code)]
```

Get all the counties in the target states

```{r }
(
all_counties <- filter(US_county, state_code %in% all_states) %>%
  data.table() %>%
  #=== merge with the data on counties in HPA ===#
  hpa_counties[., on = c("state_code", "county_code")] %>%
  #=== merge with FIPS codes data ===#
  fips_codes[., on = c("state_code", "county_code")] %>%
  #=== indicator of whether intersecting with HPA or not ===#
  .[, in_hpa := 0] %>%
  .[!is.na(ir_area_ratio), in_hpa := 1] %>%
  #=== generate state-code variable ===#
  .[, sc_code := paste0(state_code, county_code)] %>%
  #=== keep only the relevant variables ===#
  .[, .(state, state_name, sc_code, ir_area_ratio, geometry, in_hpa)]
)
```

Map of all the counties: 
Note: not all the counties are used in our eventual analysis

```{r map-all-counties, cache = TRUE}
tm_shape(st_as_sf(all_counties)) +
tm_polygons() +
tm_shape(
  tigris::states() %>% 
    filter(STATEFP %in% all_states)
) +
tm_borders(col = "red")
```

# save the county data

```{r }
saveRDS(all_counties, here("Shared/Data/ProcessedData/base_counties.rds"))

# all_counties[, state_name] %>% unique()
```

### 2023-07-22 Codes/0_2_quickstat.Rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Get crop yield and acreage data: Corn and Soybeans"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Preparation
Read the counties data

```{r }
base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
```

Get the list of states

```{r }
all_states <- base_counties$state_name %>%
  unique() %>%
  toupper()
```

Define a function to get QuickStat data

```{r }
get_quickstat_data <- function(year, data_item) {
  desc <-
    tidyUSDA::allDataItem %>%
    grep(pattern = data_item, ., value = TRUE) %>%
    .[!grepl(., pattern = "NET")]

  temp_data <- getQuickstat(
    key = key_get("usda_nass_qs_api"),
    program = "SURVEY",
    data_item = desc,
    geographic_level = "COUNTY",
    state = all_states,
    year = year,
    geometry = FALSE
  ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
    rename("state_code" = "state_fips_code")

  return(temp_data)
}
```

List of items to download

```{r }
item_list <-
  c(
    "CORN, GRAIN, IRRIGATED - YIELD",
    "CORN, GRAIN, NON-IRRIGATED - YIELD",
    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, IRRIGATED - YIELD",
    "SOYBEANS, NON-IRRIGATED - YIELD",
    "SOYBEANS, IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED"
  )
```

List of year-item combinations

```{r }
(
  par_list <-
    expand.grid(
      data_item = item_list,
      year = as.character(1985:2018)
    ) %>%
    data.table()
)
```

# Download the data

```{r download-nass-data, eval = F}

# === Download the relevant NASS data ===#
(
  nass_data <-
    lapply(
      seq_len(nrow(par_list)),
      function(x) {
        print(x)
        data <- get_quickstat_data(
          par_list[x, year],
          par_list[x, data_item]
        ) %>%
          data.table()
        return(data)
      }
    ) %>%
    rbindlist()
)

# === save the data ===#
saveRDS(nass_data, here("Shared/Data/RawData/nass_data_raw.rds"))
```

# Process the nass-data

```{r process-nass-data}
(
  nass_data_wide <-
    here("Shared/Data/RawData/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not? ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

Quick Summary

```{r }
library(modelsummary)
datasummary(
  crop * ir * (Mean + SD)
  ~ acres + yield,
  data = nass_data_wide
)
```

# Save the data

```{r eval = F}
saveRDS(nass_data_wide, here("Shared/Data/ProcessedData/nass_data.rds"))
```

### 2023-07-22 Codes/0_3_get_sat.Rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Assign saturated thickness value "
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi=400,
  fig.width=7.5,
  fig.height=5,
  out.width="750px",
  out.height="500px"
)
```

# Preparation
Import saturated thickness data

```{r }
sat_temp <- raster("/Volumes/Elements/Database/SaturatedThickness/SatThick_19851.tif")
```

HPA counties

```{r }
hpa_counties <- 
  here("Shared/Data/ProcessedData/base_counties.rds") %>% 
  readRDS() %>% 
  .[in_hpa == 1, ] %>% 
  st_as_sf() %>% 
  st_transform(projection(sat_temp)) 
```

Define function to extract saturated thickness for each county

```{r }
get_sat <- function(y) {
  # y <- 2000
  sat_temp <- 
    paste0("/Volumes/Elements/Database/SaturatedThickness/SatThick_", y, "1.tif") %>% 
    raster()

  sat_mean <- 
    exact_extract(sat_temp, hpa_counties, "mean")

  data_return <- 
    data.table(
      sat = sat_mean,
      sc_code = hpa_counties$sc_code,
      year = y
    )

  return(data_return)
}
```

Extract saturated thickness values (parallelized)

```{r extract-sat, eval = F}
sat_all <- mclapply(1985:2016, get_sat, mc.cores = 12) %>%
  rbindlist() %>%
  .[sat < 0, sat := 0] %>% 
  .[, sat := sat * 3.28084] # meter to feet
```

# Save the data

```{r save, eval = F}
saveRDS(sat_all, here("Shared/Data/ProcessedData/sat_all.rds"))
```

# Quick visualization

```{r }
sat_all <- 
  here("Shared/Data/ProcessedData/sat_all.rds") %>% 
  readRDS()
```

Saturate thickness history by county

```{r }
ggplot(sat_all) +
  geom_line(aes(y = sat, x = year, group = sc_code))
```

Histogram

```{r }
ggplot(sat_all) +
  geom_histogram(aes(x = sat))
```

### 2023-07-22 Codes/1_1_main_analysis.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Regression Analysis and Prediction"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare data

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    sat_thld_m = 12, # saturated thickness has to be at least 12 meters
    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
    balance_thld = c(-1200, 200),
    sat_cat_num = 3,
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table()

saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))
```

# Regression and prediciton: Corn

## Intensive Margin

### Regression

```{r}
data_corn <-
  reg_data[crop_type == "corn", reg_data_y][[1]] %>%
  copy() %>%
  #--- bu/acre to tonne/ha ---#
  .[, yield := yield * 12.553 / 200] %>%
  .[, type := "Corn"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
  .[, balance := -balance]

#--- define the semi-parametric portion of the model ---#
semi_formula <-
  formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res_corn <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_corn
  )

#--- main regression results ---#
y_res_int <- semi_res_corn$fe_res

r2_corn_wb <- r2(y_res_int, type = "r2")
```

### Prediction

```{r}
gam_y_res <- semi_res_corn$gam_res

sc_dry_base <- data_corn[sc_code == "31011", sc_dry][1]
sc_base <- "31011"

#--- raw data for prediction ---#
data_for_pred_corn <-
  CJ(
    balance = seq(min(data_corn$balance), max(data_corn$balance), length = 50),
    sat_cat = unique(data_corn$sat_cat)
  ) %>%
  # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
  .[, sat := case_when(
    sat_cat == "[39.4,82]" ~ 40,
    sat_cat == "(82,194]" ~ 194,
    sat_cat == "(194,708]" ~ 348,
    sat_cat == "dryland" ~ 0
  )] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  .[, sc_code := sc_base] %>%
  .[, sc_dry := sc_dry_base] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_corn <-
  gen_smooth_data(
    data = data_for_pred_corn,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)

yhat_data_corn[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_corn[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]
```

```{r}
(
  g_yield_balance_corn <-
    ggplot(yhat_data_corn) +
    geom_ribbon(
      aes(
        ymin = y_hat - 1.96 * y_hat_se,
        ymax = y_hat + 1.96 * y_hat_se,
        x = balance,
        fill = factor(sat_cat_q)
      ),
      alpha = 0.3
    ) +
    geom_line(aes(
      y = y_hat,
      x = balance,
      color = factor(sat_cat_q)
    )) +
    xlab("Water Balance") +
    ylab("Yield (tonne/ha)") +
    scale_color_discrete(name = "") +
    scale_fill_discrete(guide = "none") +
    theme(
      legend.position = "bottom"
    ) +
    theme_bw()
)
```

## Extensive Margin (acreage share)

```{r}
ir_share_data_corn <-
  data_corn %>%
  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
  .[, balance_avg := mean(balance), by = sc_code] %>%
  .[, type := "Corn"]

ir_share_res_corn <-
  gam(
    acres_ratio ~
      s(sat, k = 3, m = 2) +
      s(balance_avg, k = 3, m = 2) +
      factor(state_name) +
      factor(year),
    data = ir_share_data_corn,
    # weight = ~acres,
    family = binomial
  )
```

### Prediction

```{r}
share_pred_data_corn <-
  CJ(
    sat = seq(
      min(ir_share_data_corn$sat),
      max(ir_share_data_corn$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_corn$balance_avg),
    year = 2016,
    state_name = "Nebraska"
  )

ir_share_hat_data <-
  predict(
    ir_share_res_corn,
    newdata = share_pred_data_corn,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]

ggplot(share_pred_data_corn[sat <= 600, ]) +
  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat * 0.3048
    ),
    alpha = 0.4
  ) +
  ylim(0, NA) +
  xlab("Saturated Thickness (meter)") +
  ylab("Share of Irrigated Acres")
```

## Average (rainfed and irrigated combined)

```{r}
data_corn_total <-
  yhat_data_corn[, .(y_hat, y_hat_se, balance, sat_cat, sat, type, year)] %>%
  .[, state_name := "Nebraska"] %>%
  .[, balance_avg := mean(balance)] %>%
  .[, c("ir_share_hat", "ir_share_hat_se") := predict(ir_share_res_corn, newdata = ., type = "response", se.fit = TRUE)]

data_ir_corn <-
  data_corn_total[sat_cat != "dryland"] %>%
  setnames(c("y_hat", "y_hat_se"), c("y_hat_ir", "y_hat_se_ir"))

data_dry_corn <-
  data_corn_total[sat_cat == "dryland"] %>%
  .[, .(y_hat, y_hat_se, balance)] %>%
  setnames(c("y_hat", "y_hat_se"), c("y_hat_dry", "y_hat_se_dry"))

data_avg_yield_corn <-
  data_dry_corn[data_ir_corn, on = "balance"] %>%
  .[, avg_yield := ir_share_hat * y_hat_ir + (1 - ir_share_hat) * y_hat_dry] %>%
  .[, type := "Corn"]
```

# Regression and prediciton: Soybean

## Intensive margin

### Regression

```{r}
data_soy <-
  reg_data[crop_type == "soy", reg_data_y][[1]] %>%
  copy() %>%
  #--- bu/acre to tonne/ha ---#
  .[, yield := yield * 12.553 / 200] %>%
  .[, type := "Soybean"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
  .[, balance := -balance]

#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res_soy <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_soy
  )

#--- main regression results ---#
y_res_int <- semi_res_soy$fe_res

r2_soy_wb <- r2(y_res_int, type = "r2")
```

### Prediction

```{r}
gam_y_res <- semi_res_soy$gam_res

sc_dry_base <- data_soy[sc_code == "31011", sc_dry][1]
sc_base <- "31011"

#--- raw data for prediction ---#
data_for_pred_soy <-
  CJ(
    balance = seq(min(data_soy$balance), max(data_soy$balance), length = 50),
    sat_cat = unique(data_soy$sat_cat)
  ) %>%
  .[, sat := case_when(
    sat_cat == "[39.4,82]" ~ 40,
    sat_cat == "(82,194]" ~ 194,
    sat_cat == "(194,708]" ~ 348,
    sat_cat == "dryland" ~ 0
  )] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  .[, sc_code := data_soy$sc_code[1]] %>%
  .[, sc_dry := data_soy$sc_dry[1]] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_soy <-
  gen_smooth_data(
    data = data_for_pred_soy,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_soy, se.fit = TRUE)

yhat_data_soy[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_soy[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]
```

```{r}
(
  g_yield_balance_soy <-
    ggplot(yhat_data_soy) +
    geom_ribbon(
      aes(
        ymin = y_hat - 1.96 * y_hat_se,
        ymax = y_hat + 1.96 * y_hat_se,
        x = balance,
        fill = factor(sat_cat_q)
      ),
      alpha = 0.3
    ) +
    geom_line(aes(
      y = y_hat,
      x = balance,
      color = factor(sat_cat_q)
    )) +
    xlab("balance") +
    ylab("Yield (tonne/ha)") +
    scale_color_discrete(name = "") +
    theme(
      legend.position = "bottom"
    )
)
```

## Extensive Margin (acreage share)

```{r}
ir_share_data_soy <-
  data_soy %>%
  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
  .[, balance_avg := mean(balance), by = sc_code] %>%
  .[, type := "Soybean"]

ir_share_res_soy <-
  gam(
    acres_ratio ~
      s(sat, k = 4, m = 2) +
      s(balance_avg, k = 3, m = 2) +
      factor(state_name) +
      factor(year),
    data = ir_share_data_soy,
    # weight = ~acres,
    family = binomial
  )
```

### Prediction

```{r}
share_pred_data_soy <-
  CJ(
    sat = seq(
      min(ir_share_data_soy$sat),
      max(ir_share_data_soy$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_soy$balance_avg),
    year = 2016,
    state_name = "Nebraska"
  )

ir_share_hat_data <-
  predict(
    ir_share_res_soy,
    newdata = share_pred_data_soy,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]

ggplot(share_pred_data_soy[sat <= 600, ]) +
  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat * 0.3048
    ),
    alpha = 0.4
  ) +
  ylim(0, NA) +
  xlab("Saturated Thickness (meter)") +
  ylab("Share of Irrigated Acres")
```

## Average (rainfed and irrigated combined)

```{r}
data_soy_total <-
  yhat_data_soy[, .(y_hat, y_hat_se, balance, sat_cat, sat, type, year)] %>%
  .[, state_name := "Nebraska"] %>%
  .[, balance_avg := mean(balance)] %>%
  .[, c("ir_share_hat", "ir_share_hat_se") := predict(ir_share_res_soy, newdata = ., type = "response", se.fit = TRUE)]

data_ir_soy <-
  data_soy_total[sat_cat != "dryland"] %>%
  setnames(c("y_hat", "y_hat_se"), c("y_hat_ir", "y_hat_se_ir"))

data_dry_soy <-
  data_soy_total[sat_cat == "dryland"] %>%
  .[, .(y_hat, y_hat_se, balance)] %>%
  setnames(c("y_hat", "y_hat_se"), c("y_hat_dry", "y_hat_se_dry"))

data_avg_yield_soy <-
  data_dry_soy[data_ir_soy, on = "balance"] %>%
  .[, avg_yield := ir_share_hat * y_hat_ir + (1 - ir_share_hat) * y_hat_dry] %>%
  .[, type := "Soybean"]
```

# Save the results

```{r}
reg_results_data <-
  data.table(
    crop = c("Corn", "Soybean"),
    data = list(data_corn, data_soy),
    semi_res = list(semi_res_corn, semi_res_soy),
    yhat_data = list(yhat_data_corn, yhat_data_soy),
    share_data = list(ir_share_data_corn, ir_share_data_soy),
    share_res = list(ir_share_res_corn, ir_share_res_soy),
    share_pred_data = list(share_pred_data_corn, share_pred_data_soy),
    avg_yield_data = list(data_avg_yield_corn, data_avg_yield_soy)
  )

saveRDS(reg_results_data, here::here("Shared/Results/reg_results_data.rds"))
```

### 2023-07-22 Codes/1_2_test_dif_significance_ir_yield.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: Test of the significance of yield difference (intensive)
---

# Preparation

```{r}
results <- readRDS("Shared/Results/reg_results_data.rds")
```

# Define the function to test

```{r}
test_dif_in_yield <- function(crop, balance, base_q, comp_q, results) {
  #++++++++++++++++++++++++++++++++++++
  #+ For testing
  #++++++++++++++++++++++++++++++++++++
  # crop <- "Corn"
  # balance <- 1000
  # base_q <- 1
  # comp_q <- 3

  #++++++++++++++++++++++++++++++++++++
  #+ Main
  #++++++++++++++++++++++++++++++++++++
  q_data <-
    data.table(
      quater = c(1, 2, 3),
      coef_name = list(
        c("s_balance_sat_cat_39_4_82__1", "s_balance_sat_cat_39_4_82__2", "sat_cat::[39.4,82]"),
        c("s_balance_sat_cat_82_194__1", "s_balance_sat_cat_82_194__2", "sat_cat::(82,194]"),
        c("s_balance_sat_cat_194_708__1", "s_balance_sat_cat_194_708__2", "sat_cat::(194,708]")
      ),
      quarter_names = c("[39.4,82]", "(82,194]", "(194,708]")
    )

  crop_w <- crop
  results_w <- results[crop == crop_w, ]
  fe_res <- results_w[, semi_res][[1]]$fe_res
  vcov <- vcov(fe_res)
  gam_res <- results_w[, semi_res][[1]]$gam_res
  data_w <- results_w[, data][[1]]
  balance_w <- balance

  #--- create data for prediction ---# 
  data_for_pred <-
    CJ(
      balance = balance_w,
      sat_cat = unique(data_w$sat_cat)
    ) %>%
    # === can be any sc_code (need to shift yield for "average" county) ===#
    # .[, sc_code := data_w$sc_code[2]] %>%
    .[, sc_code := data_w[sat_cat != "dryland"]$sc_code[1]] %>%
    .[, sc_dry := paste0(data_w$sc_dry[1])] %>%
    # === can be any year (need to shift yield for "average" year) ===#
    .[, year := 2009] %>%
    # === fake yield data (necessary to take advantage of predict.gam) ===#
    .[, yield := 0] %>%
    .[, type := "pred"] %>%
    .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

  #--- create bases  ---#
  yhat_data <-
    gen_smooth_data(
      data = data_for_pred,
      gam_res = gam_res
    ) %>%
    .$data

  #++++++++++++++++++++++++++++++++++++
  #+ point estimate
  #++++++++++++++++++++++++++++++++++++
  coef_data <- tidy(fe_res) %>% data.table()

  #--- coefficients ---#
  base_coef_names <- q_data[quater == base_q, coef_name][[1]]
  comp_coef_names <- q_data[quater == comp_q, coef_name][[1]]

  base_coefs <- coef_data[term %in% base_coef_names, estimate]
  comp_coefs <- coef_data[term %in% comp_coef_names, estimate]

  coefs <- c(comp_coefs, base_coefs)

  #--- X ---#
  base_x_names <- base_coef_names[-3]
  comp_x_names <- comp_coef_names[-3]

  base_x <-
    c(
      yhat_data[balance == balance_w & sat_cat == q_data[quater == base_q, quarter_names], ..base_x_names] %>% unlist(),
      1
    )

  comp_x <-
    c(
      yhat_data[balance == balance_w & sat_cat == q_data[quater == comp_q, quarter_names], ..comp_x_names] %>% unlist(),
      1
    )

  x_vec <- c(comp_x, -base_x)

  #--- difference in yield ---#
  dif_hat <- sum(coefs * x_vec)

  #++++++++++++++++++++++++++++++++++++
  #+ SE
  #++++++++++++++++++++++++++++++++++++
  coef_names <- c(base_coef_names, comp_coef_names)
  which_col <- purrr::map_dbl(coef_names, function(x) which(x == colnames(vcov)))
  rel_vcov <- vcov[which_col, which_col]

  se <- sqrt(t(x_vec) %*% rel_vcov %*% x_vec) %>% .[1, 1]

  test_result <-
    data.table(
      crop = crop_w,
      balance = balance_w,
      base_q = base_q,
      comp_q = comp_q,
      dif_hat = dif_hat,
      se = se,
      t = dif_hat / se
    )

  return(test_result)
}
```

# Run tests for a series of values of water deficit

```{r}
data_for_test <-
  expand.grid(
    crop = c("Corn", "Soybean"),
    balance = seq(-200, 1200, by = 20),
    comp_q = c(2, 3)
  ) %>% data.table()

test_results <-
  lapply(
    1:nrow(data_for_test),
    function(x) {
      test_dif_in_yield(
        crop = data_for_test[x, crop],
        balance = data_for_test[x, balance],
        base_q = 1,
        comp_q = data_for_test[x, comp_q],
        results = results
      )
    }
  ) %>%
  rbindlist()

saveRDS(test_results, "Shared/Results/test_results_intensive.rds")
```

### 2023-07-22 Codes/1_3_bootstrap_avgyield.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

# Testing of the significance of yield difference (total)

```{r}
results_data <- readRDS("Shared/Results/reg_results_data.rds")
```

## Corn

```{r}
#--- corn ---#
data_corn <- results_data[crop == "Corn", data][[1]]

#--- soy ---#
data_soy <- results_data[crop == "Soybean", data][[1]]

#--- sc_code and sc_dry (same for corn and soybean) ---#
#* both corn and soybean data have them
sc_dry_base <- data_corn[sc_code == "31011", sc_dry][1]
sc_base <- "31011"
```

```{r}
bootstrap <- function(i, data, sc_base, sc_dry_base) {
  #++++++++++++++++++++++++++++++++++++
  #+ Debug
  #++++++++++++++++++++++++++++++++++++
  # i <- 1
  # data <- data_corn
  # sc_base <- sc_base_corn
  # sc_dry_base <- sc_dry_base_corn
  #++++++++++++++++++++++++++++++++++++
  #+ Main
  #++++++++++++++++++++++++++++++++++++
  print(i)

  sc_ls <- data[, sc_code] %>% unique()

  data_boot <-
    rbind(
      data[sc_code %in% sample(sc_ls, length(sc_ls) - 1, replace = TRUE), ],
      data[sc_code == sc_base, ],
      data[sat_cat == "dryland", ]
    )

  #--- define the semi-parametric portion of the model ---#
  semi_formula <-
    formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))

  #--- semi-parametric FE wih fixest ---#
  semi_res <-
    semi_ols(
      semi_formula = semi_formula,
      c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
      # fe = "",
      cluster = "sc_dry",
      data = data_boot
    )

  #--- main regression results ---#
  y_res_int <- semi_res$fe_res

  r2_corn_wb <- r2(y_res_int, type = "r2")

  #--- raw data for prediction ---#
  data_for_pred <-
    CJ(
      balance = seq(min(data_boot$balance), max(data_boot$balance), length = 50),
      sat_cat = unique(data_boot$sat_cat)
    ) %>%
    # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
    .[, sat := case_when(
      sat_cat == "[39.4,82]" ~ 40,
      sat_cat == "(82,194]" ~ 194,
      sat_cat == "(194,708]" ~ 348,
      sat_cat == "dryland" ~ 0
    )] %>%
    .[is.na(sat), sat := 0] %>%
    # === can be any sc_code (need to shift yield for "average" county) ===#
    .[, sc_code := sc_base] %>%
    .[, sc_dry := sc_dry_base] %>%
    # === can be any year (need to shift yield for "average" year) ===#
    .[, year := 2009] %>%
    # === fake yield data (necessary to take advantage of predict.gam) ===#
    .[, yield := 0] %>%
    .[, type := "pred"] %>%
    .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

  #--- create bases  ---#
  yhat_data <-
    gen_smooth_data(
      data = data_for_pred,
      gam_res = semi_res$gam_res
    ) %>%
    .$data

  y_hat_with_se <- predict(y_res_int, newdata = yhat_data, se.fit = TRUE)

  yhat_data[, `:=`(
    y_hat = y_hat_with_se$fit,
    y_hat_se = y_hat_with_se$se.fit
  )]

  yhat_data[, sat_cat_q :=
    as.numeric(sat_cat) %>%
    ordinal() %>%
    paste0(., " quintile")]

  ir_share_data <-
    data_boot %>%
    .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
    .[, balance_avg := mean(balance), by = sc_code] %>%
    .[, type := "Corn"]

  ir_share_res <-
    gam(
      acres_ratio ~
        s(sat, k = 3, m = 2) +
        s(balance_avg, k = 3, m = 2) +
        factor(state_name) +
        factor(year),
      data = ir_share_data,
      # weight = ~acres,
      family = binomial
    )

  share_pred_data_corn <-
    CJ(
      sat = seq(
        min(ir_share_data$sat),
        max(ir_share_data$sat),
        length = 100
      ),
      balance_avg = mean(ir_share_data$balance_avg),
      year = 2016,
      state_name = "Nebraska"
    )

  ir_share_hat_data <-
    predict(
      ir_share_res,
      newdata = share_pred_data_corn,
      type = "response",
      se.fit = TRUE
    )

  data_total <-
    yhat_data[, .(y_hat, y_hat_se, balance, sat_cat, sat, type, year)] %>%
    .[, state_name := "Nebraska"] %>%
    .[, balance_avg := mean(balance)] %>%
    .[, c("ir_share_hat", "ir_share_hat_se") := predict(ir_share_res, newdata = ., type = "response", se.fit = TRUE)]

  data_ir <-
    data_total[sat_cat != "dryland"] %>%
    setnames(c("y_hat", "y_hat_se"), c("y_hat_ir", "y_hat_se_ir"))

  data_dry <-
    data_total[sat_cat == "dryland"] %>%
    .[, .(y_hat, y_hat_se, balance)] %>%
    setnames(c("y_hat", "y_hat_se"), c("y_hat_dry", "y_hat_se_dry"))

  data_final <-
    data_dry[data_ir, on = "balance"] %>%
    .[, avg_yield := ir_share_hat * y_hat_ir + (1 - ir_share_hat) * y_hat_dry] %>%
    .[, .(sat, sat_cat, avg_yield, balance, ir_share_hat, y_hat_dry, y_hat_ir)] %>%
    .[, type := "Corn"] %>%
    .[, q1_yield := .SD[sat_cat == "[39.4,82]", avg_yield], by = balance] %>%
    .[, dif_yield := avg_yield - q1_yield]

  return(data_final)
}
```

## Bootstrap

```{r}
boot_results_data <-
  data.table(
    crop = c("Corn", "Soybean"),
    data = list(data_corn, data_soy)
  ) %>%
  mutate(boot_results = purrr::map(data, \(data) {
    parallel::mclapply(
      1:1000,
      function(i) bootstrap(i, data, sc_base, sc_dry_base),
      mc.cores = 12
    ) %>%
      rbindlist()
  })) %>%
  dplyr::mutate(se_data = purrr::map(boot_results, \(boot_results){
    boot_results[, .(avg_yield_sd = sd(avg_yield), dif_yield_sd = sd(dif_yield)), by = .(sat_cat, balance, type)]
  }))

saveRDS(boot_results_data, "Shared/Results/boot_results.rds")
```

### 2023-07-22 Codes/2_figures_tables.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Create Figures and Tables"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Preparation

## Load the data and results

```{r}
results <- readRDS("Shared/Results/reg_results_data.rds")
```

## Set themes
```{r}
theme_fig <-
  theme_bw() +
  theme(
    axis.title.x =
      element_text(
        size = 9, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
      ),
    axis.title.y =
      element_text(
        size = 9, angle = 90, hjust = .5, vjust = .9, family = "Times"
      ),
    axis.text.x =
      element_text(
        size = 8, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
      ),
    axis.text.y =
      element_text(
        size = 8, angle = 0, hjust = 1, vjust = 0, family = "Times"
      ),
    axis.ticks =
      element_line(
        linewidth = 0.3, linetype = "solid"
      ),
    axis.ticks.length = unit(.15, "cm"),
    #--- legend ---#
    legend.text =
      element_text(
        size = 9, angle = 0, hjust = 0, vjust = 0.5, family = "Times"
      ),
    legend.title =
      element_text(
        size = 9, angle = 0, hjust = 0, vjust = 0, family = "Times"
      ),
    legend.key.size = unit(0.5, "cm"),
    #--- strip (for faceting) ---#
    strip.text = element_text(size = 9, family = "Times"),
    #--- plot title ---#
    plot.title = element_text(family = "Times", face = "bold", size = 9),
    #--- margin ---#
    # plot.margin = margin(0, 0, 0, 0, "cm"),
    #--- panel ---#
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA)
  )
```

# Summary statistics

```{r, eval = FALSE}
data_corn <- results[crop == "Corn", data][[1]]
data_soy <- results[crop == "Soybean", data][[1]]

raw_data_both <-
  rbind(data_corn, data_soy) %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")]
```

## Summary data

```{r, eval = FALSE}
(
  sum_table <-
    modelsummary::datasummary(
      type * (yield + balance) ~ dry_or_not * (Mean + SD),
      data = raw_data_both,
      output = "kableExtra"
      # output = "latex"
    )
)
```

## Boxplot of water deficit by year

```{r}
data_corn <- results[crop == "Corn", data][[1]]
data_soy <- results[crop == "Soybean", data][[1]]

wd_data <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[, .(wd = mean(balance)), by = .(year, type)] %>%
  .[, ]

yield_data <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[, ir_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")] %>%
  .[, .(yield = mean(yield), wb = mean(balance)), by = .(year, type, ir_or_not)]

coeff_corn <- 70

g_yield_wd_corn <-
  ggplot() +
  geom_line(
    data = yield_data[type == "Corn" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield")
  ) +
  geom_point(
    data = yield_data[type == "Corn" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield"),
    size = 0.8
  ) +
  geom_line(
    data = yield_data[type == "Corn" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield")
  ) +
  geom_point(
    data = yield_data[type == "Corn" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield"),
    size = 0.8
  ) +
  geom_line(
    data = wd_data[type == "Corn"],
    aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
    linetype = 2
  ) +
  geom_point(
    data = wd_data[type == "Corn"],
    aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
    size = 0.8
  ) +
  scale_y_continuous(
    sec.axis = sec_axis(~ . * coeff_corn, name = "Water Deficit (mm)")
  ) +
  scale_x_continuous(
    name = "Year",
    breaks = 1985:2016
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "Dryland Yield" = "red",
      "Irrigated Yield" = "blue",
      "Water Deficit" = "black"
    )
  ) +
  ylab("Yield (bu/acre)") +
  xlab("Year") +
  theme_fig +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90)
  ) +
  ggtitle("A. Corn")

coeff_soy <- 200
g_yield_wd_soy <-
  ggplot() +
  geom_line(
    data = yield_data[type == "Soybean" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield")
  ) +
  geom_point(
    data = yield_data[type == "Soybean" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield"),
    size = 0.8
  ) +
  geom_line(
    data = yield_data[type == "Soybean" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield")
  ) +
  geom_point(
    data = yield_data[type == "Soybean" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield"),
    size = 0.8
  ) +
  geom_line(
    data = wd_data[type == "Soybean"],
    aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
    linetype = 2
  ) +
  geom_point(
    data = wd_data[type == "Soybean"],
    aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
    size = 0.8
  ) +
  scale_y_continuous(
    sec.axis = sec_axis(~ . * coeff_soy, name = "Water Deficit (mm)")
  ) +
  scale_x_continuous(
    name = "Year",
    breaks = 1985:2016
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "Dryland Yield" = "red",
      "Irrigated Yield" = "blue",
      "Water Deficit" = "black"
    )
  ) +
  ylab("Yield (bu/acre)") +
  xlab("Year") +
  theme_fig +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90)
  ) +
  ggtitle("B. Soybean")

library(patchwork)

g_y_wd <- g_yield_wd_corn / g_yield_wd_soy

ggsave(
  file = "GitControlled/Writing/Figures/g_y_wd_all.pdf",
  g_y_wd,
  height = 6,
  width = 6
)
```

## Saturated thickness

```{r, eval = FALSE}
g_sat_dist <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[sat != 0, ] %>%
  ggplot(data = .) +
  geom_histogram(aes(x = sat * 0.3048), fill = "white", color = "blue") +
  facet_grid(type ~ ., scale = "free_y") +
  ylab("Count") +
  xlab("Aquifer Thickness (meter)") +
  theme_fig
```

## Map

```{r echo = F}
# === hpa ===#
hpa_simplified <-
  # here("../Shared/Data/ProcessedData/hp_simplified.shp") %>%
  here("Shared/Data/ProcessedData/hp_simplified.shp") %>%
  readRDS()

# === all counties ===#
data(fips_codes, package = "tidycensus")

all_counties <-
  tigris::counties() %>%
  rename(
    state_code = STATEFP,
    county_code = COUNTYFP,
    county_name = NAME
  ) %>%
  left_join(., fips_codes, by = c("state_code", "county_code")) %>%
  dplyr::select(state_code, state, county_code, county_name) %>%
  mutate(sc_code = paste0(state_code, county_code)) %>%
  data.table()

all_states <-
  tigris::states() %>%
  rename(state_code = STATEFP) %>%
  left_join(
    .,
    unique(data.table(fips_codes), by = "state"),
    by = "state_code"
  )

states_of_interest <-
  dplyr::filter(all_states, state %in% data_corn[, unique(state)])

#--- corn ---#
counties_corn_all <-
  all_counties[sc_code %in% unique(data_corn$sc_code), ] %>%
  mutate(type = "Corn") %>%
  st_as_sf()

counties_corn_ir <-
  data_corn %>%
  .[sat_cat != "dryland", ] %>%
  .[, .(sat, sc_code)] %>%
  .[, .(sat = mean(sat, na.rm = TRUE)), by = sc_code] %>%
  left_join(all_counties, by = "sc_code") %>%
  .[, type := "Corn"] %>%
  st_as_sf()

#--- soy ---#
counties_soy_all <-
  all_counties[sc_code %in% unique(data_soy$sc_code), ] %>%
  mutate(type = "Soybean") %>%
  st_as_sf()

counties_soy_ir <-
  data_soy %>%
  .[sat_cat != "dryland", ] %>%
  .[, .(sat, sc_code)] %>%
  .[, .(sat = mean(sat, na.rm = TRUE)), by = sc_code] %>%
  left_join(all_counties, by = "sc_code") %>%
  .[, type := "Soybean"] %>%
  st_as_sf()

counties_both_all <- rbind(counties_corn_all, counties_soy_all)
counties_both_ir <-
  rbind(counties_corn_ir, counties_soy_ir) %>%
  mutate(sat_meter = sat * 0.3048)

g_corn <-
  ggplot() +
  geom_sf(data = counties_corn_all, fill = "gray") +
  geom_sf(data = counties_corn_ir, aes(fill = sat * 0.3048)) +
  scale_fill_viridis_c(
    name = "Aquifer Thickness (meter)",
    guide = guide_colorbar(
      title.position = "top"
    )
  ) +
  geom_sf(data = states_of_interest, fill = NA, color = "blue") +
  geom_sf(data = hpa_simplified, fill = NA, color = "red") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  ) +
  ggtitle("A. Corn")

g_soy <-
  ggplot() +
  geom_sf(data = counties_soy_all, fill = "gray") +
  geom_sf(data = counties_soy_ir, aes(fill = sat * 0.3048)) +
  scale_fill_viridis_c(
    name = "Aquifer Thickness (meter)",
    guide = guide_colorbar(
      title.position = "top"
    )
  ) +
  geom_sf(data = states_of_interest, fill = NA, color = "blue") +
  geom_sf(data = hpa_simplified, fill = NA, color = "red") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  ) +
  ggtitle("B. Soybean")

g_map <- g_corn | g_soy

ggsave(file = "GitControlled/Writing/Figures/g_map.png", g_map, height = 7, width = 6, dpi = 600)
ggsave(file = "GitControlled/Writing/Figures/g_map.pdf", g_map, height = 7, width = 6)
```

# Yield response to water deficit by saturated thickness level (intensive margin)

```{r}
g_ir_yield_both <-
  results[, .(crop, yhat_data)] %>%
  dplyr::mutate(plot_data = purrr::map2(yhat_data, crop, \(yhat_data, crop) {
    yhat_data[, sat_cat_text := case_when(
      sat_cat == "[39.4,82]" ~ "1st quantile - [12, 25]",
      sat_cat == "(82,194]" ~ "2nd quantile - (25, 59.1]",
      sat_cat == "(194,708]" ~ "3rd quantile - (59.1, 215]",
      sat_cat == "dryland" ~ "Rainfed"
    )] %>%
      .[, sat_cat_text := factor(
        sat_cat_text,
        levels = c("Rainfed", "1st quantile - [12, 25]", "2nd quantile - (25, 59.1]", "3rd quantile - (59.1, 215]")
      )] %>%
      .[, crop_type := crop] %>%
      .[, crop_type := case_when(
        crop_type == "Corn" ~ "A. Corn",
        crop_type == "Soybean" ~ "(b) Sybean"
      )]
  })) %>%
  dplyr::mutate(g_yield = purrr::map(plot_data, \(plot_data) {
    ggplot(plot_data) +
      geom_line(aes(
        y = y_hat,
        x = balance,
        color = factor(sat_cat_text)
      )) +
      scale_x_continuous(
        name = "Water Deficit (mm)",
        breaks = (-2:11) * 100
      ) +
      scale_y_continuous(
        name = "Yield (tonne/ha)",
        breaks = 0:(ceiling(max(plot_data$y_hat)) + 0.5),
        limits = c(NA, (ceiling(max(plot_data$y_hat)) + 0.5))
      ) +
      scale_color_discrete(
        name = "Aquifer Thickness",
        guide = guide_legend(
          title.position = "top",
          nrow = 2
        )
      ) +
      scale_fill_discrete(guide = "none") +
      theme_fig +
      coord_cartesian(xlim = c(-200, 1150)) +
      theme(legend.position = "bottom")
  }))

g_ir_yield_corn <- g_ir_yield_both$g_yield[[1]] + theme(legend.position = "none")
g_ir_yield_soy <- g_ir_yield_both$g_yield[[2]]

g_wds <-
  results[, .(crop, data)] %>%
  dplyr::mutate(g_fig = purrr::map(
    data,
    function(data) {
      ggplot(data) +
        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
        theme_fig +
        xlab("Water Deficit (mm)") +
        ylab("") +
        theme_void()
    }
  ))

g_wd_corn <-
  g_wds$g_fig[[1]] +
  ggtitle("A. Corn") +
  theme(plot.title = element_text(size = 9, family = "Times", face = "bold"))

g_wd_soy <-
  g_wds$g_fig[[2]] +
  ggtitle("B. Soybean") +
  theme(plot.title = element_text(size = 9, family = "Times", face = "bold"))

g_yield_response <-
  cowplot::plot_grid(
    g_wd_corn, g_ir_yield_corn, g_wd_soy, g_ir_yield_soy,
    ncol = 1,
    align = "v",
    rel_heights = c(0.15 / 2, 0.72 / 2, 0.15 / 2, 0.98 / 2)
  )

cowplot::ggsave2("GitControlled/Writing/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)

#++++++++++++++++++++++++++++++++++++
#+ With confidence interval (seprately)
#++++++++++++++++++++++++++++++++++++
g_yield_with_conf <-
  results[, .(crop, yhat_data)] %>%
  dplyr::mutate(plot_data = purrr::map2(yhat_data, crop, \(yhat_data, crop) {
    yhat_data[, sat_cat_text := case_when(
      sat_cat == "[39.4,82]" ~ "1st quantile - [12, 25]",
      sat_cat == "(82,194]" ~ "2nd quantile - (25, 59.1]",
      sat_cat == "(194,708]" ~ "3rd quantile - (59.1, 215]",
      sat_cat == "dryland" ~ "Rainfed"
    )] %>%
      .[, sat_cat_text := factor(
        sat_cat_text,
        levels = c("Rainfed", "1st quantile - [12, 25]", "2nd quantile - (25, 59.1]", "3rd quantile - (59.1, 215]")
      )] %>%
      .[, crop_type := crop] %>%
      .[, crop_type := case_when(
        crop_type == "Corn" ~ "A. Corn",
        crop_type == "Soybean" ~ "(b) Sybean"
      )]
  })) %>%
  mutate(g_fig = purrr::map(
    plot_data,
    \(plot_data) {
      ggplot(plot_data) +
        geom_ribbon(
          aes(
            ymin = y_hat - 1.96 * y_hat_se,
            ymax = y_hat + 1.96 * y_hat_se,
            x = balance
          ),
          alpha = 0.3
        ) +
        geom_line(aes(
          y = y_hat,
          x = balance
        )) +
        facet_wrap(~sat_cat_text, ncol = 2) +
        scale_x_continuous(
          name = "Water Deficit (mm)",
          breaks = (-1:5) * 200
        ) +
        scale_y_continuous(
          name = "Yield (tonne/ha)",
          breaks = 0:13
        ) +
        theme_fig +
        theme(
          legend.position = "bottom",
          strip.background = element_blank()
        )
    }
  )) %>%
  dplyr::select(crop, g_fig) %>%
  data.table()

ggsave("Shared/Results/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "Corn", g_fig][[1]], width = 6, height = 3.5)
ggsave("Shared/Results/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "Soybean", g_fig][[1]], width = 6, height = 3.5)
```

# The impact of saturated thickness on the share of irrigated production

```{r}
plot_data_ir_share <-
  results[, .(crop, share_pred_data)] %>%
  unnest(cols = share_pred_data) %>%
  data.table() %>%
  .[, crop_type := case_when(
    crop == "Corn" ~ "A. Corn",
    crop == "Soybean" ~ "B. Soybean"
  )]

g_share <-
  ggplot(plot_data_ir_share[sat <= 500, ]) +
  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat * 0.3048
    ),
    alpha = 0.4
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
  xlab("Aquifer Thickness (meter)") +
  ylab("Estimated Share of Irrigated Acres") +
  facet_wrap(crop_type ~ ., ncol = 1) +
  theme_fig +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
  )

ggsave(
  file = "GitControlled/Writing/Figures/g_share.pdf",
  g_share,
  height = 4,
  width = 5
)
```

# Total impact of a decline in saturated thickness on average yield

```{r}
g_total_both <-
  rbindlist(results$avg_yield_data) %>%
  .[, crop_type := case_when(
    type == "Corn" ~ "A. Corn",
    type == "Soybean" ~ "B. Soybean"
  )] %>%
  .[, sat_m := measurements::conv_unit(sat, "ft", "m") %>% format(digits = 3)] %>%
  nest_by(crop_type) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data) {
    ggplot(data) +
      geom_line(aes(y = avg_yield, x = balance, color = sat_m)) +
      theme_fig +
      ylim(min(data$avg_yield) - 0.5, max(data$avg_yield) + 0.5) +
      ylab("Yield (tonne/ha)") +
      scale_color_discrete(
        "Aquifer Thickness (meter)",
        guide = guide_legend(
          title.position = "top",
          nrow = 1
        )
      ) +
      scale_x_continuous(name = "Water Deficit (mm)", breaks = (-2:11) * 100, limits = c(-200, 1200)) +
      theme(legend.position = "bottom")
  }))

g_total_corn <-
  g_total_both$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  ylim(6, 13) +
  ggtitle("A. Corn")

g_total_soy <-
  g_total_both$g_fig[[2]] +
  ylim(1.5, 4) +
  ggtitle("B. Soybean")

g_total <- g_total_corn / g_total_soy

# g_total <-
#   ggplot(data = combined_data) +
#   geom_line(aes(y = avg_yield, x = balance, color = sat_cat_text)) +
#   facet_wrap(crop_type ~ ., scale = "free_y", ncol = 1) +
#   theme_fig +
#   ylim(0, NA) +
#   ylab("Yield (tonne/ha)") +
#   scale_color_discrete(
#     "Aquifer Thickness",
#     guide = guide_legend(
#       title.position = "top",
#       nrow = 1
#     )
#   ) +
#   scale_x_continuous(name = "Water Deficit (mm)", breaks = (-2:11) * 100) +
#   theme(
#     legend.position = "bottom",
#     strip.background = element_blank(),
#     legend.text = element_text(size = 8)
#   )

ggsave(
  file = "GitControlled/Writing/Figures/g_total_impact.pdf",
  g_total,
  height = 6,
  width = 6
)
```

# Test the significance of the difference in irrigated yield between the aquifer thickness categories

```{r}
test_results <- readRDS("Shared/Results/test_results_intensive.rds")

plot_data <-
  test_results %>%
  .[, upper := dif_hat + se * 1.96] %>%
  .[, lower := dif_hat - se * 1.96] %>%
  .[, crop_type := case_when(
    crop == "Corn" ~ "A. Corn",
    crop == "Soybean" ~ "B. Soybean"
  )] %>%
  .[, as_q := case_when(
    comp_q == 2 ~ "2nd Aquifer Thickness Quantile",
    comp_q == 3 ~ "3rd Aquifer Thickness Quantile"
  )] %>%
  nest_by(crop) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(
    data,
    \(data) {
      ggplot(data) +
        facet_grid(. ~ as_q) +
        geom_ribbon(
          aes(
            x = balance,
            ymin = lower,
            ymax = upper
          ),
          fill = "blue",
          alpha = 0.3
        ) +
        geom_line(
          aes(
            x = balance,
            y = dif_hat
          ),
          color = "red"
        ) +
        geom_hline(yintercept = 0, col = "black") +
        scale_color_discrete(name = "") +
        scale_x_continuous(breaks = (-1:6) * 200) +
        ylab("Difference in Yield (tonne/ha)") +
        xlab("Water Deficit (mm)") +
        theme_fig +
        theme(
          strip.background = element_blank(),
          legend.position = "bottom",
          panel.spacing = unit(1, "lines")
        )
    }
  ))

g_ir_yield_dif <-
  cowplot::plot_grid(
    plot_data$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
    plot_data$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
    ncol = 1,
    align = "v",
    rel_heights = c(0.5, 0.5)
  )

cowplot::ggsave2("GitControlled/Writing/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
```

# Testing of the significance of yield difference (average total impact)

```{r}
boot_results_data <- readRDS("Shared/Results/boot_results.rds")

g_avg_yield_dif <-
  boot_results_data[, .(crop, boot_results)] %>%
  unnest(cols = c(boot_results)) %>%
  data.table() %>%
  .[, .(
    dif_yield = mean(dif_yield),
    lower = quantile(dif_yield, prob = 0.05),
    upper = quantile(dif_yield, prob = 0.95)
  ),
  by = .(sat, balance, crop)
  ] %>%
  .[, sat_text := paste0(
    "Aquifer Thickness = ",
    measurements::conv_multiunit(sat, "ft", "m") %>% format(digits = 3),
    " meter"
  )] %>%
  .[sat != 40, ] %>%
  nest_by(crop) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data){
    ggplot(data) +
      geom_ribbon(aes(x = balance, ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
      geom_line(aes(x = balance, y = dif_yield), color = "red") +
      geom_hline(yintercept = 0) +
      facet_grid(. ~ sat_text, scales = "free_y") +
      xlab("Water Deficit (mm)") +
      ylab("Difference in Yield (tonne/ha)") +
      xlim(-200, 1200) +
      theme_fig +
      theme(
        strip.background = element_blank(),
        strip.text = element_text(size = 9, family = "Times"),
        panel.spacing.y = unit(1, "cm")
      )
  }))

g_ay_corn <-
  g_avg_yield_dif$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(breaks = -2:5, limits = c(-2, 5)) +
  ggtitle("A. Corn")

g_ay_soy <-
  g_avg_yield_dif$g_fig[[2]] +
  scale_y_continuous(breaks = -2:2, limits = c(-2, 2)) +
  ggtitle("B. Soybean")

g_ay <- g_ay_corn / g_ay_soy

# g_avg_yield_dif <-
#   ggplot(plot_data_dif) +
#   geom_ribbon(aes(x = -balance, ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
#   geom_line(aes(x = -balance, y = dif_yield), color = "red") +
#   geom_hline(yintercept = 0) +
#   facet_grid(crop ~ sat_cat_text, scales = "free_y") +
#   xlab("Water Deficit (mm)") +
#   ylab("Difference in Yield (tonne/ha)") +
#   theme_fig +
#   theme(
#     strip.background = element_blank(),
#     strip.text = element_text(size = 9, family = "Times"),
#     panel.spacing.y = unit(1, "cm")
#   )

ggsave("Shared/Results/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
```

# The impact of water deficit on average yield by aquifer thickness category with confidence interval

```{r}
g_ay_response <-
  boot_results_data[, .(crop, boot_results)] %>%
  unnest(cols = c(boot_results)) %>%
  data.table() %>%
  .[, .(
    avg_yield = mean(avg_yield),
    lower = quantile(avg_yield, prob = 0.05),
    upper = quantile(avg_yield, prob = 0.95)
  ),
  by = .(sat, balance, crop)
  ] %>%
  .[, sat_text := paste0(
    "Aquifer Thickness = ",
    measurements::conv_multiunit(sat, "ft", "m") %>% format(digits = 3),
    " meter"
  )] %>%
  nest_by(crop) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data) {
    ggplot(data) +
      geom_ribbon(aes(x = balance, ymin = lower, ymax = upper), alpha = 0.3) +
      geom_line(aes(x = balance, y = avg_yield)) +
      facet_grid(. ~ sat_text, scales = "free_y") +
      ylim(min(data$avg_yield) - 1, max(data$avg_yield) + 0.5) +
      xlab("Water Deficit (mm)") +
      ylab("Yield (tonne/ha)") +
      theme_fig +
      theme(
        strip.background.x = element_blank(),
        panel.spacing = unit(0.1, "cm")
      )
  }))

g_ay_response_corn <-
  g_ay_response$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  ggtitle("A. Corn")

g_ay_response_soy <-
  g_ay_response$g_fig[[2]] +
  ggtitle("B. Soybean")

g_ay_response <- g_ay_response_corn / g_ay_response_soy

# g_avg_yield_response <-
#   ggplot(plot_data_response) +
#   geom_ribbon(aes(x = -balance, ymin = lower, ymax = upper), alpha = 0.3) +
#   geom_line(aes(x = -balance, y = avg_yield)) +
#   facet_grid(crop ~ sat_cat_text, scales = "free_y") +
#   xlab("Water Deficit (mm)") +
#   ylab("Yield (tonne/ha)") +
#   theme_fig +
#   theme(
#     strip.background.x = element_blank(),
#     panel.spacing = unit(0.2, "cm")
#   )

ggsave("Shared/Results/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
```

### 2023-07-22 Codes/A2_state_fe.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

In the main regresion analysis in the manuscript, state FEs are included. County FEs are not included as it eliminates too much of the variation in aquifer thickness. This appendix compares the impact of aquifer thickness on irrigated share with and without State FEs.

# Regression with and without state fixed effects

## Without State FEs

### Corn

```{r}
data_corn <- readRDS(here("Shared/Data/ProcessedData/data_corn.rds"))

#--- define the semi-parametric portion of the model ---#
semi_formula <-
  formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_corn
  )

#--- main regression results ---#
y_res_int <- semi_res$fe_res

r2_corn_wb <- r2(y_res_int, type = "r2")

gam_y_res <- semi_res$gam_res

#--- raw data for prediction ---#
data_for_pred_corn <-
  CJ(
    balance = seq(min(data_corn$balance), max(data_corn$balance), length = 50),
    sat_cat = unique(data_corn$sat_cat)
  ) %>%
  .[, sat := case_when(
    sat_cat == "[39.4,82]" ~ 40,
    sat_cat == "(82,194]" ~ 194,
    sat_cat == "(194,708]" ~ 348,
    sat_cat == "dryland" ~ 0
  )] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  # .[, sc_code := data_corn$sc_code[2]] %>%
  .[, sc_code := data_corn[sat_cat != "dryland"]$sc_code[1]] %>%
  .[, sc_dry := paste0(data_corn$sc_dry[1])] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_corn <-
  gen_smooth_data(
    data = data_for_pred_corn,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)

yhat_data_corn[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_corn[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]

ir_share_data_corn <-
  data_corn %>%
  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
  .[, balance_avg := mean(balance), by = sc_code] %>%
  .[, type := "Corn"]

ir_share_res_corn <-
  gam(
    acres_ratio ~
      s(sat, k = 3, m = 2) + s(balance_avg, k = 3, m = 2) +
      ti(sat, balance_avg, k = 3) + factor(year),
    data = ir_share_data_corn,
    # weight = ~acres,
    family = binomial
  )

share_pred_data_corn <-
  CJ(
    sat = seq(
      min(ir_share_data_corn$sat),
      max(ir_share_data_corn$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_corn$balance_avg),
    year = 2016
  )

ir_share_hat_data <-
  predict(
    ir_share_res_corn,
    newdata = share_pred_data_corn,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
```

### Soybean

```{r}
data_soy <- readRDS(here("Shared/Data/ProcessedData/data_soy.rds"))

#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_soy
  )

#--- main regression results ---#
y_res_int <- semi_res$fe_res

r2_soy_wb <- r2(y_res_int, type = "r2")

gam_y_res <- semi_res$gam_res

#--- raw data for prediction ---#
data_for_pred_soy <-
  CJ(
    balance = seq(min(data_soy$balance), max(data_soy$balance), length = 50),
    sat_cat = unique(data_soy$sat_cat)
  ) %>%
  .[, sat := case_when(
    sat_cat == "[39.4,82]" ~ 40,
    sat_cat == "(82,194]" ~ 194,
    sat_cat == "(194,708]" ~ 348,
    sat_cat == "dryland" ~ 0
  )] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  .[, sc_code := data_soy$sc_code[1]] %>%
  .[, sc_dry := data_soy$sc_dry[1]] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_soy <-
  gen_smooth_data(
    data = data_for_pred_soy,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_soy, se.fit = TRUE)

yhat_data_soy[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_soy[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]

ir_share_data_soy <-
  data_soy %>%
  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
  .[, balance_avg := mean(balance), by = sc_code] %>%
  .[, type := "Soybean"]

ir_share_res_soy <-
  gam(
    acres_ratio ~
      s(sat, k = 4, m = 2) +
      s(balance_avg, k = 3, m = 2) +
      ti(sat, balance_avg, k = 3, m = 2) +
      factor(state_name) +
      factor(year),
    data = ir_share_data_soy,
    # weight = ~acres,
    family = binomial
  )

share_pred_data_soy <-
  CJ(
    sat = seq(
      min(ir_share_data_soy$sat),
      max(ir_share_data_soy$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_soy$balance_avg),
    year = 2016,
    state_name = "Nebraska"
  )

ir_share_hat_data <-
  predict(
    ir_share_res_soy,
    newdata = share_pred_data_soy,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
```

```{r}
plot_data_ir_share_no_sfe <-
  rbind(
    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
  ) %>%
  .[, sfe := "No State FE"]
```

## With State FEs (Main)

### Corn

```{r}
ir_share_res_corn <-
  gam(
    acres_ratio ~
      s(sat, k = 3, m = 2) +
      s(balance_avg, k = 3, m = 2) +
      ti(sat, balance_avg, k = 3) +
      factor(state_name) +
      factor(year),
    data = ir_share_data_corn,
    # weight = ~acres,
    family = binomial
  )

share_pred_data_corn <-
  CJ(
    sat = seq(
      min(ir_share_data_corn$sat),
      max(ir_share_data_corn$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_corn$balance_avg),
    year = 2016,
    state_name = "Nebraska"
  )

ir_share_hat_data <-
  predict(
    ir_share_res_corn,
    newdata = share_pred_data_corn,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
```

### Soybean

```{r}
ir_share_res_soy <-
  gam(
    acres_ratio ~
      s(sat, k = 4, m = 2) +
      s(balance_avg, k = 3, m = 2) +
      ti(sat, balance_avg, k = 3, m = 2) +
      factor(state_name) +
      factor(year),
    data = ir_share_data_soy,
    # weight = ~acres,
    family = binomial
  )

share_pred_data_soy <-
  CJ(
    sat = seq(
      min(ir_share_data_soy$sat),
      max(ir_share_data_soy$sat),
      length = 100
    ),
    balance_avg = mean(ir_share_data_soy$balance_avg),
    year = 2016,
    state_name = "Nebraska"
  )

ir_share_hat_data <-
  predict(
    ir_share_res_soy,
    newdata = share_pred_data_soy,
    type = "response",
    se.fit = TRUE
  )

share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
```

```{r}
plot_data_ir_share_with_sfe <-
  rbind(
    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
  ) %>%
  .[, sfe := "With State FE"]
```

```{r}
plot_data_ir_share <-
  rbind(
    plot_data_ir_share_with_sfe,
    plot_data_ir_share_no_sfe
  )

g_share_comp <-
  ggplot(plot_data_ir_share[sat <= 500, ]) +
  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat * 0.3048
    ),
    alpha = 0.4
  ) +
  ylim(0.25, 1) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(10, 160, by = 20)) +
  xlab("Saturated Thickness (meter)") +
  ylab("Share of Irrigated Acres") +
  facet_grid(type ~ sfe) +
  theme_fig +
  theme(
    strip.text = element_text(face = "bold"),
    strip.background = element_blank()
  )

ggsave(
  file = "GitControlled/Writing/Figures/g_share_comp.pdf",
  g_share_comp,
  height = 5,
  width = 6
)
```

# Loss of variation in aquifer thickness

```{r}
ir_share_data_corn[, sat_m := measurements::conv_unit(sat, "ft", "m")]

s_deam_data <-
  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m)), by = state_name] %>%
  .[, type := "B. Demeaned by State Average"] %>%
  .[, .(sat_m, type)]

c_deam_data <-
  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m)), by = sc_code] %>%
  .[, type := "C. Demeaned by County Average"] %>%
  .[, .(sat_m, type)]

non_deam_data <-
  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m))] %>%
  .[, type := "A. Demeaned by Overall Average"]

plot_data <- rbind(s_deam_data, c_deam_data, non_deam_data)

g_variation <-
  ggplot(plot_data) +
  geom_histogram(aes(x = sat_m)) +
  facet_wrap(type ~ ., ncol = 1) +
  theme_fig +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(hjust = 0, face = "bold")
  ) +
  xlab("Aquifer Thickness") +
  ylab("Count")

ggsave(
  file = "GitControlled/Writing/Figures/g_variation.pdf",
  g_variation,
  height = 5,
  width = 6
)
```

### 2023-07-22 Codes/A4_spei.rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

# Regression with SPEI

## Corn: Intensive Margin

### Regression

```{r}
#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(spei, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_corn
  )

#--- main regression results ---#
y_res_int <- semi_res$fe_res

r2_corn_spei <- r2(y_res_int, type = "r2")
```

```{r, eval = FALSE}
modelsummary::msummary(y_res_int, stars = TRUE)
```

### Prediction

```{r}
gam_y_res <- semi_res$gam_res

#--- raw data for prediction ---#
data_for_pred_corn <-
  CJ(
    spei = seq(min(data_corn$spei), max(data_corn$spei), length = 50),
    sat_cat = unique(data_corn$sat_cat)
  ) %>%
  .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  .[, sc_code := data_corn$sc_code[20]] %>%
  .[, sc_dry := data_corn$sc_dry[20]] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_corn <-
  gen_smooth_data(
    data = data_for_pred_corn,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)

yhat_data_corn[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_corn[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]
```

## Soybean: Intensive Margin

### Regression

```{r}
#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(spei, by = sat_cat, k = 3, m = 2))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_dry",
    data = data_soy
  )

#--- main regression results ---#
y_res_int <- semi_res$fe_res

r2_soy_spei <- r2(y_res_int, type = "r2")
```

```{r}
modelsummary::msummary(y_res_int, stars = TRUE)
```

### Prediction

```{r}
gam_y_res <- semi_res$gam_res

#--- raw data for prediction ---#
data_for_pred_soy <-
  CJ(
    spei = seq(min(data_soy$spei), max(data_soy$spei), length = 50),
    sat_cat = unique(data_soy$sat_cat)
  ) %>%
  .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
  .[is.na(sat), sat := 0] %>%
  # === can be any sc_code (need to shift yield for "average" county) ===#
  .[, sc_code := data_soy$sc_code[20]] %>%
  .[, sc_dry := data_soy$sc_dry[20]] %>%
  # === can be any year (need to shift yield for "average" year) ===#
  .[, year := 2009] %>%
  # === fake yield data (necessary to take advantage of predict.gam) ===#
  .[, yield := 0] %>%
  .[, type := "pred"] %>%
  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

#--- create bases  ---#
yhat_data_soy <-
  gen_smooth_data(
    data = data_for_pred_soy,
    gam_res = gam_y_res
  ) %>%
  .$data

y_hat_with_se <- predict(y_res_int, newdata = yhat_data_soy, se.fit = TRUE)

yhat_data_soy[, `:=`(
  y_hat = y_hat_with_se$fit,
  y_hat_se = y_hat_with_se$se.fit
)]

yhat_data_soy[, sat_cat_q :=
  as.numeric(sat_cat) %>%
  ordinal() %>%
  paste0(., " quintile")]
```

## Yield response to water deficit by saturated thickness level (intensive margin)

```{r}
yield_data <-
  rbind(
    yhat_data_corn[, .(y_hat, y_hat_se, spei, sat_cat, sat)][, type := "Corn"],
    yhat_data_soy[, .(y_hat, y_hat_se, spei, sat_cat, sat)][, type := "Soybean"]
  )

g_yield_spei <-
  ggplot(yield_data) +
  geom_ribbon(
    aes(
      ymin = y_hat - 1.96 * y_hat_se,
      ymax = y_hat + 1.96 * y_hat_se,
      x = spei,
      fill = factor(sat_cat)
    ),
    alpha = 0.3
  ) +
  geom_line(aes(
    y = y_hat,
    x = spei,
    color = factor(sat_cat)
  )) +
  facet_grid(type ~ ., scale = "free_y") +
  scale_x_continuous(
    name = "Water Deficit (mm)",
    breaks = (-2:12) * 100
  ) +
  scale_y_continuous(
    name = "Yield (tonne/ha)",
    breaks = 0:13
  ) +
  scale_color_discrete(name = "") +
  scale_fill_discrete(guide = "none") +
  theme(
    legend.position = "bottom"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Regression with continuous saturated thickness

## Corn

```{r}
#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(balance, k = 4, m = 2) + s(sat, k = 4, m = 2) + ti(balance, sat, k = 4))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_code",
    data = data_corn
  )

# y_res_int <- semi_res$fe_res

#--- main regression results ---#
r2_corn_sat_cont <- r2(semi_res$fe_res, type = "r2")
```

## Soybean

```{r}
#--- define the semi-parametric portion of the model ---#
semi_formula <- formula(yield ~ s(balance, k = 4, m = 2) + s(sat, k = 4, m = 2) + ti(balance, sat, k = 4))

#--- semi-parametric FE wih fixest ---#
semi_res <-
  semi_ols(
    semi_formula = semi_formula,
    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
    # fe = "",
    cluster = "sc_code",
    data = data_soy
  )

#--- main regression results ---#
r2_soy_sat_cont <- r2(semi_res$fe_res, type = "r2")
```

### 2023-07-24 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.tex
updates

@@
 \usepackage{amsmath,amssymb}
+\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math} % this also loads fontspec
+  \usepackage{unicode-math}
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else
-  % xetex/luatex font selection
 \fi
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same}
+\urlstyle{same} % disable monospaced font for URLs
 \hypersetup{
@@
 
+\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}
+
 \begin{enumerate}
@@
 
-\textcolor{blue}{ferkjnf}
+\textcolor{blue}{All the instances of "saturated thickness" and "aquifer saturated thickness" have been replaced by "aquifer thickness".}
 
@@
 \setcounter{enumi}{1}
+\tightlist
 \item
@@
   manuscript.
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  First point:
+  \textcolor{blue}{Quantiles are not recalculated each each.
+If we do so, then the same quantile would represent different aquifer thickness range every year.
+When modeling the impact of aquifer thickness, this is not desirable because we are implicitly making assumptions that different ranges of aquifer thickness would have the same impact.}
+\item
+  Second point: \textbackslash textcolor\{blue\}\{Actually, 0\% quntile
+  is 12 m. ``9m'' was repalced with ``12m'' in the current manuscript.\}
+\item
+  Third point:
+  \textcolor{blue}{No interesting insights will be gained from them because they are merely shifters without any intractions with aquifer thickness.
+They tell us how much higher yield in a single county relative to another county on average.
+But, we do not beleive that is useful information at least in this artcile.}
+\item
+  Fourth point:
+  \textcolor{blue}{This is a very important point. We have added texts describing the implications of this fact in page xx.}
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
 \item
@@
   point in the manuscript.
+\end{enumerate}
+
+\textcolor{blue}{What we meant by "conditiond on drought risk exposure" is "after taking into account the impact of climate variables." In order to avoid confusion, we simply removed it from the sentence in the current manuscript.}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
 \item
@@
   explanation of what is meant by ``estimated semi-parametrically''.
+\end{enumerate}
+
+\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained.
+We eliminated the word semi-parametrically as it is not ncecessary.
+Generalized additive model that estiamtes the impact of a variable using splines without functional form assumption are sometimes referred to semi-parametric regression.
+That's what we meant by "estimate semi-parametrically" in our previous manuscript.}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
 \item
@@
 
+\textcolor{blue}{Tim?}
+
 \hypertarget{minor-points}{%
@@
 
-Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through aquifer
-depletion will..'' -- confusing sentence -- please replace ``through''
-with ``by which''.
+\begin{itemize}
+\tightlist
+\item
+  Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through
+  aquifer depletion will..'' -- confusing sentence -- please replace
+  ``through'' with ``by which''.
+\end{itemize}
 
+\textcolor{blue}{Corrected.}
+
+\begin{itemize}
+\tightlist
+\item
   Section 1, 4th Paragraph -- ``we generate empirical evidence'' seems
@@
   previous\ldots{}'' reads better as ``Our analysis extends
-previous\ldots{}'' There are many issues such as these and the previous
-comment. Please address such issues so that the manuscript is an easier
-read.
+  previous\ldots{}'' There are many issues such as these and the
+  previous comment. Please address such issues so that the manuscript is
+  an easier read.
+\end{itemize}
 
+\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}
+
+\begin{itemize}
+\tightlist
+\item
   Section 2.1 title -- I would delete ``marginally'' as it is not needed
   and some will find it confusing.
+\end{itemize}
 
+\textcolor{blue}{Tim, any thoughts on this. I feel like it is still important to say the increase is not much.}
+
+\begin{itemize}
+\tightlist
+\item
   Figure 2 - Why do the plots stop at a thickness of about 12 m when you
-had earlier defined the minimum thickness as 9 m and mention 10 m in the
-text when discussing the relationships?
+  had earlier defined the minimum thickness as 9 m and mention 10 m in
+  the text when discussing the relationships?
+\end{itemize}
 
+\textcolor{blue}{Thank you for catching this. 9m and 10m should have been 12m.}
+
+\begin{itemize}
+\tightlist
+\item
   Section 2.3, First Paragraph -- please clarify that ``average per-area
@@
   would make the manuscript easier to follow.
+\end{itemize}
 
+\textcolor{blue}{We added "(weighted-average of irrigated and rainfed yields, where weights are production area shares)" so that it is clear what average per-area production means.}
+
+\begin{itemize}
+\tightlist
+\item
   Figure 5 -- I assume that the left plot is corn and the right soybean,
   but please clarify.
+\end{itemize}
 
+\textcolor{blue}{Thanks for noticing this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
   Data and Code Availability -- Are the data from Dr.~Haacker on the
@@
   available according to the principles of open science.
+\end{itemize}
 
+\textcolor{blue}{Yes, we will make the data available online.}
+
+\begin{itemize}
+\tightlist
+\item
   Line before Equation (2) -- ``REF'' should be replaced with the actual
   reference.
+\end{itemize}
 
-Discussion, First Paragraph -- ``reduction in aquifer depletion'' should
-be ``reduction in aquifer thickness''.
+\textcolor{blue}{Tim?}
 
+\begin{itemize}
+\tightlist
+\item
+  Discussion, First Paragraph -- ``reduction in aquifer depletion''
+  should be ``reduction in aquifer thickness''.
+\end{itemize}
+
+\textcolor{blue}{Thank you for catching this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
   Section 4.2, 3rd Paragraph -- ``the the impact'' -- remember to spell
   and grammar check prior to submission.
+\end{itemize}
 
+\textcolor{blue}{Thank you for catching this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
   Reference in Section 3 -- I doubt that ``Butler Jr'' is the surname,
-please replace with ``Butler''. Suffixes apparently aren't handled well
-by some of the bibliography generation programs.
+  please replace with ``Butler''. Suffixes apparently aren't handled
+  well by some of the bibliography generation programs.
+\end{itemize}
 
-Please remember to close all parentheses and don't mispair brackets and
-parentheses. Such stuff can raise questions about the care used in the
-analysis.
+\textcolor{blue}{}
+
+\begin{itemize}
+\tightlist
+\item
+  Please remember to close all parentheses and don't mispair brackets
+  and parentheses. Such stuff can raise questions about the care used in
+  the analysis.
+\end{itemize}
 

### 2023-07-24 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.tex
updates

--- 2023-07-21 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.rmd
@@
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\author{}
+\date{\vspace{-2.5em}}
+
+\begin{document}
+
 \textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}
 
-This is an interesting paper that asks an important yet answered question in the field of agricultural sustainability - how much is groundwater depletion reducing agricultural productivity in the High Plains region in the United States?
-While the paper is an important contribution to the literature, there are several issues that need to be addressed prior to acceptance at any journal.
+This is an interesting paper that asks an important yet answered
+question in the field of agricultural sustainability - how much is
+groundwater depletion reducing agricultural productivity in the High
+Plains region in the United States? While the paper is an important
+contribution to the literature, there are several issues that need to be
+addressed prior to acceptance at any journal.
 
-+ There needs to be more description about why aquifer saturation thickness is an appropriate proxy for farmers' ability to irrigate from wells.
-This is particularly true since this is a broad interest journal, and the mechanism for the main independent variable should be very clear to all readers.
+\begin{itemize}
+\tightlist
+\item
+  There needs to be more description about why aquifer saturation
+  thickness is an appropriate proxy for farmers' ability to irrigate
+  from wells. This is particularly true since this is a broad interest
+  journal, and the mechanism for the main independent variable should be
+  very clear to all readers.
+\end{itemize}
 
@@
 
-+ Though I am not trained as a hydrologist, my understanding is that aquifer saturation thickness is a good measure of general water availability in an aquifer but not necessarily how much water is available at a specific location or point.
-If this is the case, the authors should make clear why this is an appropriate measure to use in their study and potential limitations of using this approach.
+\begin{itemize}
+\tightlist
+\item
+  Though I am not trained as a hydrologist, my understanding is that
+  aquifer saturation thickness is a good measure of general water
+  availability in an aquifer but not necessarily how much water is
+  available at a specific location or point. If this is the case, the
+  authors should make clear why this is an appropriate measure to use in
+  their study and potential limitations of using this approach.
+\end{itemize}
 
@@
 
-+ It is unclear to me why county and time fixed effects were not used in every regression.
-In particular, it does not look like they were included in the regression that modeled irrigated area as a function of saturated thickness and water deficit.
-Yet, accounting for time-invariant differences across counties is important, especially because this is the regression that showed the largest effect of saturated thickness on a measure of agricultural production (irrigated area).
-Is there not enough variation within a county through time to pick up the effect of saturated thickness on irrigated area?
-This should all be made clearer, and a model that better captures the causal relationship between saturated thickness and area irrigated should be used if possible.
-If it's not possible, the limitations of the current approach should be made very clear and potential issues with inference.
+\begin{itemize}
+\tightlist
+\item
+  It is unclear to me why county and time fixed effects were not used in
+  every regression. In particular, it does not look like they were
+  included in the regression that modeled irrigated area as a function
+  of saturated thickness and water deficit. Yet, accounting for
+  time-invariant differences across counties is important, especially
+  because this is the regression that showed the largest effect of
+  saturated thickness on a measure of agricultural production (irrigated
+  area). Is there not enough variation within a county through time to
+  pick up the effect of saturated thickness on irrigated area? This
+  should all be made clearer, and a model that better captures the
+  causal relationship between saturated thickness and area irrigated
+  should be used if possible. If it's not possible, the limitations of
+  the current approach should be made very clear and potential issues
+  with inference.
+\end{itemize}
 
@@
 
-+ There are minor typos (e.g., missing reference listed as REF) that should be corrected prior to resubmission.
+\begin{itemize}
+\tightlist
+\item
+  There are minor typos (e.g., missing reference listed as REF) that
+  should be corrected prior to resubmission.
+\end{itemize}
 
 \textcolor{blue}{}
+
+\end{document}

### 2023-07-24 Codes/0_7_ssurgo.Rmd
updates

---
title: "Prepare SSURGO Data"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare

## Read the counties sf
```{r}
all_counties <-
  readRDS(here("Shared/Data/ProcessedData/base_counties.rds")) %>%
  st_as_sf() %>%
  dplyr::mutate(geometry = st_make_valid(geometry)) %>%
  st_transform(4326)
```

```{r}
get_ssurgo_props <- function(field, vars) {
  # Get SSURGO mukeys for polygon intersection
  download_try <-
    soilDB::SDA_spatialQuery(
      field,
      what = "geom",
      db = "SSURGO"
    )

  if ("try-error" %in% class(download_try)) {
    #* county too large fot a single download
    print("County too large. Breaking it into pieces and work on them separately.")

    grids <-
      st_bbox(field) %>%
      st_as_sfc() %>%
      st_make_grid(n = c(3, 3))

    field_parts <-
      st_intersection(grids, field) %>%
      st_make_valid() %>%
      st_as_sf()

    ssurgo_geom <-
      lapply(
        1:nrow(field_parts),
        \(x) {
          print(paste0("working on ", x))
          field_part_w <- field_parts[x, ]
          download_try <-
            soilDB::SDA_spatialQuery(
              field_part_w,
              what = "geom",
              db = "SSURGO"
            ) %>%
            st_make_valid() %>%
            st_intersection(field_part_w, .)
        }
      ) %>%
      bind_rows() %>%
      st_make_valid() %>%
      dplyr::mutate(
        area = as.numeric(st_area(.)),
        area_weight = area / sum(area)
      )

    ssurgo_data_sum <- get_summarized_ssurgo(ssurgo_geom, vars)
  } else {
    ssurgo_geom <-
      download_try %>%
      st_make_valid() %>%
      st_intersection(field, .) %>%
      mutate(
        area = as.numeric(st_area(.)),
        area_weight = area / sum(area)
      )

    ssurgo_data_sum <- get_summarized_ssurgo(ssurgo_geom, vars)
  }

  return(ssurgo_data_sum)
}

get_summarized_ssurgo <- function(ssurgo_geom, vars) {
  # Get soil properties for each mukey
  mukeydata <-
    soilDB::get_SDA_property(
      property = vars,
      method = "Weighted Average",
      mukeys = unique(ssurgo_geom$mukey),
      top_depth = 0,
      bottom_depth = 150
    )

  ssurgo_data <- left_join(ssurgo_geom, mukeydata, by = "mukey")

  ssurgo_data_sum <-
    ssurgo_data %>%
    data.table() %>%
    .[,
      lapply(.SD, function(x) weighted.mean(x, w = area_weight, na.rm = TRUE)),
      .SDcols = vars
    ] %>%
    data.table()
  return(ssurgo_data_sum)
}
```

## Download

```{r download, cache = TRUE}
vars <- c("sandtotal_r", "silttotal_r", "claytotal_r", "awc_r", "om_r", "dbovendry_r")

parallel::mclapply(
  1:nrow(all_counties),
  \(x) {
    print(x)
    file_name <- here::here(paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
    if (!file.exists(file_name)) {
      print(paste0("working on ", x))
      temp <- get_ssurgo_props(all_counties[x, ], vars)
      saveRDS(temp, here::here(paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds")))
    } else {
      print(paste0("skipping ", x))
    }
  },
  mc.cores = 6,
  mc.preschedule = FALSE
)

# lapply(
#   1:nrow(all_counties),
#   \(x) {
#     print(x)
#     file_name <- paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds")
#     if (!file.exists(file_name)) {
#       temp <- get_ssurgo_props(field = all_counties[x, ], vars)
#       saveRDS(temp, paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
#     }
#   }
# )
```

# save the SSURGO data

```{r }
ssurgo_soil <-
  fs::dir_ls("Shared/Data/RawData/SSURGO") %>%
  purrr::map(readRDS) %>%
  lapply(function(x) x[, geometry := NULL]) %>%
  rbindlist() %>%
  .[, `:=`(
    state = NULL,
    state_name = NULL,
    ir_area_ratio = NULL,
    in_hpa = NULL
  )]

saveRDS(ssurgo_soil, here("Shared/Data/ProcessedData/ssurgo_soil.rds"))
# all_counties[, state_name] %>% unique()
```

### 2023-07-24 Codes/1_1_main_analysis.rmd
updates

@@
   .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
+  .[, state_year := paste0(state_name, "_", year)] %>%
   .[, balance := -balance]
@@
     balance = seq(min(data_corn$balance), max(data_corn$balance), length = 50),
-    sat_cat = unique(data_corn$sat_cat)
+    sat = c(0, 40, 194, 348, 602)
   ) %>%
   # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
-  .[, sat := case_when(
-    sat_cat == "[39.4,82]" ~ 40,
-    sat_cat == "(82,194]" ~ 194,
-    sat_cat == "(194,708]" ~ 348,
-    sat_cat == "dryland" ~ 0
+  .[, sat_cat := case_when(
+    sat == 40 ~ "[39.4,82]",
+    sat == 194 ~ "(82,194]",
+    sat == 348 ~ "(194,708]",
+    sat == 602 ~ "(194,708]",
+    sat == 0 ~ "dryland"
   )] %>%
-  .[is.na(sat), sat := 0] %>%
   # === can be any sc_code (need to shift yield for "average" county) ===#
@@
 )]
-
-yhat_data_corn[, sat_cat_q :=
-  as.numeric(sat_cat) %>%
-  ordinal() %>%
-  paste0(., " quintile")]
 ```
 
-```{r}
+```{r, eval = FALSE}
 (
@@
         x = balance,
-        fill = factor(sat_cat_q)
+        fill = factor(sat)
       ),
@@
       x = balance,
-      color = factor(sat_cat_q)
+      color = factor(sat)
     )) +
@@
       s(balance_avg, k = 3, m = 2) +
-      factor(state_name) +
-      factor(year),
+      factor(state_year),
     data = ir_share_data_corn,
@@
   )
+```
+
+```{r}
+gam_formula_ir_share <-
+  formula(
+    acres_ratio ~
+      s(sat, k = 4, m = 2) +
+      s(balance_avg, k = 3, m = 2)
+  )
+
+reg_data_ir_share <-
+  gen_smooth_data(
+    data = ir_share_data_corn,
+    gam_formula = gam_formula_ir_share
+  )
+
+formula_feols <-
+  names(reg_data_ir_share$model_X) %>%
+  paste0(., collapse = "+") %>%
+  paste0("acres_ratio ~ ", .) %>%
+  # paste0(., "+", ) %>%
+  paste0(., " | ", "state_year") %>%
+  formula()
+
+share_res <-
+  feglm(
+    formula_feols,
+    cluster = "sc_code",
+    data = reg_data_ir_share$data,
+    family = binomial()
+  )
+
+share_pred_data_corn <-
+  CJ(
+    sat = seq(
+      min(ir_share_data_corn$sat),
+      max(ir_share_data_corn$sat),
+      length = 100
+    ),
+    balance_avg =
+      c(
+        seq(
+          min(ir_share_data_corn$balance_avg),
+          max(ir_share_data_corn$balance_avg),
+          length = 20
+        ),
+        mean(ir_share_data_corn$balance_avg)
+      ),
+    state_year = "Nebraska_2009"
+  ) %>%
+  .[, acres_ratio := 0]
+
+share_hat_data <-
+  gen_smooth_data(
+    data = share_pred_data_corn,
+    gam_res = reg_data_ir_share$gam_res
+  ) %>%
+  .$data %>%
+  .[balance_avg == mean(ir_share_data_corn$balance_avg), ]
+
+share_hat <- 
+  predict(
+    share_res, 
+    newdata = share_hat_data
+  )
+
+share_hat_data[, `:=`(
+  share_hat = share_hat
+)]
+
+ggplot(share_hat_data) +
+  geom_line(aes(y = share_hat, x = sat))
 ```
@@
     balance_avg = mean(ir_share_data_corn$balance_avg),
-    year = 2016,
-    state_name = "Nebraska"
+    state_year = "Nebraska_2009"
   )
@@
 
-ggplot(share_pred_data_corn[sat <= 600, ]) +
-  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
-  geom_ribbon(
-    aes(
-      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
-      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
-      x = sat * 0.3048
-    ),
-    alpha = 0.4
-  ) +
-  ylim(0, NA) +
-  xlab("Saturated Thickness (meter)") +
-  ylab("Share of Irrigated Acres")
+# ggplot(share_pred_data_corn[sat <= 600, ]) +
+#   geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
+#   geom_ribbon(
+#     aes(
+#       ymin = ir_share_hat - 1.96 * ir_share_hat_se,
+#       ymax = ir_share_hat + 1.96 * ir_share_hat_se,
+#       x = sat * 0.3048
+#     ),
+#     alpha = 0.4
+#   ) +
+#   ylim(0, NA) +
+#   xlab("Saturated Thickness (meter)") +
+#   ylab("Share of Irrigated Acres")
 ```
@@
   yhat_data_corn[, .(y_hat, y_hat_se, balance, sat_cat, sat, type, year)] %>%
-  .[, state_name := "Nebraska"] %>%
+  .[, state_year := "Nebraska_2009"] %>%
   .[, balance_avg := mean(balance)] %>%
@@
     balance = seq(min(data_soy$balance), max(data_soy$balance), length = 50),
-    sat_cat = unique(data_soy$sat_cat)
+    sat = c(0, 40, 194, 348, 602)
   ) %>%
-  .[, sat := case_when(
-    sat_cat == "[39.4,82]" ~ 40,
-    sat_cat == "(82,194]" ~ 194,
-    sat_cat == "(194,708]" ~ 348,
-    sat_cat == "dryland" ~ 0
+  # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
+  .[, sat_cat := case_when(
+    sat == 40 ~ "[39.4,82]",
+    sat == 194 ~ "(82,194]",
+    sat == 348 ~ "(194,708]",
+    sat == 602 ~ "(194,708]",
+    sat == 0 ~ "dryland"
   )] %>%
-  .[is.na(sat), sat := 0] %>%
   # === can be any sc_code (need to shift yield for "average" county) ===#
@@
 )]
-
-yhat_data_soy[, sat_cat_q :=
-  as.numeric(sat_cat) %>%
-  ordinal() %>%
-  paste0(., " quintile")]
 ```
@@
 ```{r}
-(
-  g_yield_balance_soy <-
-    ggplot(yhat_data_soy) +
-    geom_ribbon(
-      aes(
-        ymin = y_hat - 1.96 * y_hat_se,
-        ymax = y_hat + 1.96 * y_hat_se,
-        x = balance,
-        fill = factor(sat_cat_q)
-      ),
-      alpha = 0.3
-    ) +
-    geom_line(aes(
-      y = y_hat,
-      x = balance,
-      color = factor(sat_cat_q)
-    )) +
-    xlab("balance") +
-    ylab("Yield (tonne/ha)") +
-    scale_color_discrete(name = "") +
-    theme(
-      legend.position = "bottom"
-    )
-)
+# (
+#   g_yield_balance_soy <-
+#     ggplot(yhat_data_soy) +
+#     geom_ribbon(
+#       aes(
+#         ymin = y_hat - 1.96 * y_hat_se,
+#         ymax = y_hat + 1.96 * y_hat_se,
+#         x = balance,
+#         fill = factor(sat_cat_q)
+#       ),
+#       alpha = 0.3
+#     ) +
+#     geom_line(aes(
+#       y = y_hat,
+#       x = balance,
+#       color = factor(sat_cat_q)
+#     )) +
+#     xlab("balance") +
+#     ylab("Yield (tonne/ha)") +
+#     scale_color_discrete(name = "") +
+#     theme(
+#       legend.position = "bottom"
+#     )
+# )
 ```
@@
       s(balance_avg, k = 3, m = 2) +
-      factor(state_name) +
-      factor(year),
+      factor(state_year),
     data = ir_share_data_soy,
@@
     year = 2016,
-    state_name = "Nebraska"
+    state_year = "Nebraska_2009"
   )
@@
 
-ggplot(share_pred_data_soy[sat <= 600, ]) +
-  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
-  geom_ribbon(
-    aes(
-      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
-      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
-      x = sat * 0.3048
-    ),
-    alpha = 0.4
-  ) +
-  ylim(0, NA) +
-  xlab("Saturated Thickness (meter)") +
-  ylab("Share of Irrigated Acres")
+# ggplot(share_pred_data_soy[sat <= 600, ]) +
+#   geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
+#   geom_ribbon(
+#     aes(
+#       ymin = ir_share_hat - 1.96 * ir_share_hat_se,
+#       ymax = ir_share_hat + 1.96 * ir_share_hat_se,
+#       x = sat * 0.3048
+#     ),
+#     alpha = 0.4
+#   ) +
+#   ylim(0, NA) +
+#   xlab("Saturated Thickness (meter)") +
+#   ylab("Share of Irrigated Acres")
 ```
@@
   yhat_data_soy[, .(y_hat, y_hat_se, balance, sat_cat, sat, type, year)] %>%
-  .[, state_name := "Nebraska"] %>%
+  .[, state_year := "Nebraska_2009"] %>%
   .[, balance_avg := mean(balance)] %>%

### 2023-07-24 Codes/1_1_main_analysis_SHORT.rmd
updates

---
title: "Regression Analysis and Prediction"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare data

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    sat_thld_m = 12, # saturated thickness has to be at least 12 meters
    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
    balance_thld = c(-1200, 200),
    sat_cat_num = 3,
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table()

saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))
```

```{r}
plan(multicore, workers = parallel::detectCores() - 2)
```

# Regression and prediciton: Corn

## Yield

```{r}
set.seed(283397)

#--- global parameters ---#
sc_base <- "31011"
sc_dry_base <- "31011_0"

all_results <-
  reg_data %>%
  setnames("crop_type", "crop") %>%
  #--- further necessary transformation ---#
  dplyr::mutate(data = purrr::map(
    reg_data_y,
    \(reg_data_y) {
      copy(reg_data_y) %>%
        #--- bu/acre to tonne/ha ---#
        .[, yield := yield * 12.553 / 200] %>%
        .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
        .[, state_year := paste0(state_name, "_", year)] %>%
        .[, balance := -balance]
    }
  )) %>%
  #--- bootstrapped data ---#
  dplyr::mutate(boot_data = purrr::map(
    data,
    \(data) {
      replicate(20, boot(data), simplify = FALSE)
      # replicate(1000, boot(data), simplify = FALSE)
    }
  )) %>%
  #--- analysis on each of the bootstrapped datasets ---#
  dplyr::mutate(results = purrr::map(
    boot_data,
    \(boot_data) {
      parallel::mclapply(
        1:length(boot_data),
        \(x) {
          print(x)
          run_analysis(boot_data[[x]])
        },
        mc.cores = 5
      )
    }
  )) %>%
  #--- extract ---#
  .[, .(crop, data, results)] %>%
  dplyr::mutate(
    yield_response = purrr::map(
      results,
      \(x) {
        purrr::map(x, 1) %>% rbindlist()
      }
    ),
    share = purrr::map(
      results,
      \(x) {
        purrr::map(x, 2) %>% rbindlist()
      }
    ),
    avg_yield = purrr::map(
      results,
      \(x) {
        purrr::map(x, 3) %>% rbindlist()
      }
    )
  ) %>%
  dplyr::mutate(yield_response_sum = purrr::map(
    yield_response,
    \(yield_response) {
      yield_response[, .(
        y_hat = mean(y_hat),
        lower = quantile(y_hat, prob = 0.025),
        upper = quantile(y_hat, prob = 0.975),
        dif_y_hat = mean(dif_y_hat),
        dif_lower = quantile(dif_y_hat, prob = 0.025),
        dif_upper = quantile(dif_y_hat, prob = 0.975)
      ),
      by = .(balance, sat, sat_cat)
      ]
    }
  )) %>%
  dplyr::mutate(share_sum = purrr::map(
    share,
    \(share) {
      share[, .(
        ir_share_hat = mean(ir_share_hat),
        lower = quantile(ir_share_hat, prob = 0.025),
        upper = quantile(ir_share_hat, prob = 0.975)
      ),
      by = .(sat)
      ]
    }
  )) %>%
  dplyr::mutate(avg_yield_sum = purrr::map(
    avg_yield,
    \(avg_yield) {
      avg_yield[, .(
        avg_yield = mean(avg_yield),
        lower = quantile(avg_yield, prob = 0.025),
        upper = quantile(avg_yield, prob = 0.975),
        dif_avg_yield = mean(dif_avg_yield),
        dif_lower = quantile(dif_avg_yield, prob = 0.025),
        dif_upper = quantile(dif_avg_yield, prob = 0.975)
      ),
      by = .(balance, sat)
      ]
    }
  ))
# all_results[1, crop := "Corn"]
# all_results[2, crop := "Soybean"]
```

```{r}
saveRDS(all_results, "Shared/Results/all_results.rds")
```

### 2023-07-24 Codes/1_3_bootstrap_avgyield.rmd
updates

@@
       balance = seq(min(data_boot$balance), max(data_boot$balance), length = 50),
-      sat_cat = unique(data_boot$sat_cat)
+      sat = c(0, 40, 194, 348, 602)
     ) %>%
     # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
-    .[, sat := case_when(
-      sat_cat == "[39.4,82]" ~ 40,
-      sat_cat == "(82,194]" ~ 194,
-      sat_cat == "(194,708]" ~ 348,
-      sat_cat == "dryland" ~ 0
+    .[, sat_cat := case_when(
+      sat == 40 ~ "[39.4,82]",
+      sat == 194 ~ "(82,194]",
+      sat == 348 ~ "(194,708]",
+      sat == 602 ~ "(194,708]",
+      sat == 0 ~ "dryland"
     )] %>%
-    .[is.na(sat), sat := 0] %>%
     # === can be any sc_code (need to shift yield for "average" county) ===#
@@
         s(balance_avg, k = 3, m = 2) +
-        factor(state_name) +
-        factor(year),
+        factor(state_year),
       data = ir_share_data,
@@
 
-  share_pred_data_corn <-
+  share_pred_data <-
     CJ(
@@
       year = 2016,
-      state_name = "Nebraska"
+      state_year = "Nebraska_2009"
     )
@@
       ir_share_res,
-      newdata = share_pred_data_corn,
+      newdata = share_pred_data,
       type = "response",
@@
       function(i) bootstrap(i, data, sc_base, sc_dry_base),
-      mc.cores = 12
+      mc.cores = parallel::detectCores() - 2
     ) %>%

### 2023-07-24 Codes/2_figures_tables.rmd
updates

@@
 results <- readRDS("Shared/Results/reg_results_data.rds")
+all_results <- readRDS("Shared/Results/all_results.rds")
+
 ```
@@
     legend.position = "bottom",
-    legend.key.width = unit(1, "cm")
+    legend.key.width = unit(1, "cm"),
+    plot.title = element_text(size = 9, family = "Times"),
+    legend.title = element_text(family = "Times"),
+    legend.text = element_text(family = "Times")
   ) +
@@
 g_ir_yield_both <-
-  results[, .(crop, yhat_data)] %>%
-  dplyr::mutate(plot_data = purrr::map2(yhat_data, crop, \(yhat_data, crop) {
-    yhat_data[, sat_cat_text := case_when(
-      sat_cat == "[39.4,82]" ~ "1st quantile - [12, 25]",
-      sat_cat == "(82,194]" ~ "2nd quantile - (25, 59.1]",
-      sat_cat == "(194,708]" ~ "3rd quantile - (59.1, 215]",
+  all_results[, .(crop, yield_response_sum)] %>%
+  dplyr::mutate(plot_data = purrr::map2(yield_response_sum, crop, \(x, crop) {
+    x[, sat_cat_text := case_when(
+      sat_cat == "[39.4,82]" ~ "1st quantile: [12m, 25m]",
+      sat_cat == "(82,194]" ~ "2nd quantile: (25m, 59.1m]",
+      sat_cat == "(194,708]" ~ "3rd quantile: (59.1m, 215m]",
       sat_cat == "dryland" ~ "Rainfed"
@@
         sat_cat_text,
-        levels = c("Rainfed", "1st quantile - [12, 25]", "2nd quantile - (25, 59.1]", "3rd quantile - (59.1, 215]")
+        levels = c("Rainfed", "1st quantile: [12m, 25m]", "2nd quantile: (25m, 59.1m]", "3rd quantile: (59.1m, 215m]")
       )] %>%
@@
       .[, crop_type := case_when(
-        crop_type == "Corn" ~ "A. Corn",
-        crop_type == "Soybean" ~ "(b) Sybean"
+        crop_type == "corn" ~ "A. Corn",
+        crop_type == "soy" ~ "(b) Sybean"
       )]
@@
 g_wds <-
-  results[, .(crop, data)] %>%
+  all_results[, .(crop, data)] %>%
   dplyr::mutate(g_fig = purrr::map(
@@
 g_yield_with_conf <-
-  results[, .(crop, yhat_data)] %>%
-  dplyr::mutate(plot_data = purrr::map2(yhat_data, crop, \(yhat_data, crop) {
-    yhat_data[, sat_cat_text := case_when(
-      sat_cat == "[39.4,82]" ~ "1st quantile - [12, 25]",
-      sat_cat == "(82,194]" ~ "2nd quantile - (25, 59.1]",
-      sat_cat == "(194,708]" ~ "3rd quantile - (59.1, 215]",
+  all_results[, .(crop, yield_response_sum)] %>%
+  dplyr::mutate(plot_data = purrr::map2(yield_response_sum, crop, \(x, crop) {
+    x[, sat_cat_text := case_when(
+      sat_cat == "[39.4,82]" ~ "1st quantile: [12m, 25m]",
+      sat_cat == "(82,194]" ~ "2nd quantile: (25m, 59.1m]",
+      sat_cat == "(194,708]" ~ "3rd quantile: (59.1m, 215m]",
       sat_cat == "dryland" ~ "Rainfed"
@@
           aes(
-            ymin = y_hat - 1.96 * y_hat_se,
-            ymax = y_hat + 1.96 * y_hat_se,
+            ymin = lower,
+            ymax = upper,
             x = balance
@@
 
-ggsave("Shared/Results/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "Corn", g_fig][[1]], width = 6, height = 3.5)
-ggsave("Shared/Results/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "Soybean", g_fig][[1]], width = 6, height = 3.5)
+ggsave("Shared/Results/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
+ggsave("Shared/Results/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
 ```
@@
 plot_data_ir_share <-
-  results[, .(crop, share_pred_data)] %>%
-  unnest(cols = share_pred_data) %>%
+  all_results[, .(crop, share_sum)] %>%
+  unnest(cols = share_sum) %>%
   data.table() %>%
   .[, crop_type := case_when(
-    crop == "Corn" ~ "A. Corn",
-    crop == "Soybean" ~ "B. Soybean"
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
   )]
@@
     aes(
-      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
-      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
+      ymin = lower,
+      ymax = upper,
       x = sat * 0.3048
@@
 g_total_both <-
-  rbindlist(results$avg_yield_data) %>%
+  all_results[, .(crop, avg_yield_sum)] %>%
+  unnest() %>%
+  data.table() %>%
   .[, crop_type := case_when(
-    type == "Corn" ~ "A. Corn",
-    type == "Soybean" ~ "B. Soybean"
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
   )] %>%
@@
     measurements::conv_multiunit(sat, "ft", "m") %>% format(digits = 3),
-    " meter"
+    "m"
   )] %>%

### 2023-07-24 Codes/A2_state_fe.rmd
updates

@@
 
+```{r}
+reg_results_data <- readRDS(here::here("Shared/Results/reg_results_data.rds"))
+```
+
 ## Without State FEs
@@
 ```{r}
-data_corn <- readRDS(here("Shared/Data/ProcessedData/data_corn.rds"))
-
-#--- define the semi-parametric portion of the model ---#
-semi_formula <-
-  formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))
-
-#--- semi-parametric FE wih fixest ---#
-semi_res <-
-  semi_ols(
-    semi_formula = semi_formula,
-    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
-    # fe = "",
-    cluster = "sc_dry",
-    data = data_corn
-  )
-
-#--- main regression results ---#
-y_res_int <- semi_res$fe_res
-
-r2_corn_wb <- r2(y_res_int, type = "r2")
-
-gam_y_res <- semi_res$gam_res
-
-#--- raw data for prediction ---#
-data_for_pred_corn <-
-  CJ(
-    balance = seq(min(data_corn$balance), max(data_corn$balance), length = 50),
-    sat_cat = unique(data_corn$sat_cat)
-  ) %>%
-  .[, sat := case_when(
-    sat_cat == "[39.4,82]" ~ 40,
-    sat_cat == "(82,194]" ~ 194,
-    sat_cat == "(194,708]" ~ 348,
-    sat_cat == "dryland" ~ 0
-  )] %>%
-  .[is.na(sat), sat := 0] %>%
-  # === can be any sc_code (need to shift yield for "average" county) ===#
-  # .[, sc_code := data_corn$sc_code[2]] %>%
-  .[, sc_code := data_corn[sat_cat != "dryland"]$sc_code[1]] %>%
-  .[, sc_dry := paste0(data_corn$sc_dry[1])] %>%
-  # === can be any year (need to shift yield for "average" year) ===#
-  .[, year := 2009] %>%
-  # === fake yield data (necessary to take advantage of predict.gam) ===#
-  .[, yield := 0] %>%
-  .[, type := "pred"] %>%
-  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]
-
-#--- create bases  ---#
-yhat_data_corn <-
-  gen_smooth_data(
-    data = data_for_pred_corn,
-    gam_res = gam_y_res
-  ) %>%
-  .$data
-
-y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)
-
-yhat_data_corn[, `:=`(
-  y_hat = y_hat_with_se$fit,
-  y_hat_se = y_hat_with_se$se.fit
-)]
-
-yhat_data_corn[, sat_cat_q :=
-  as.numeric(sat_cat) %>%
-  ordinal() %>%
-  paste0(., " quintile")]
-
-ir_share_data_corn <-
-  data_corn %>%
-  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-  .[, balance_avg := mean(balance), by = sc_code] %>%
-  .[, type := "Corn"]
+ir_share_data_corn <- reg_results_data[crop == "Corn", share_data][[1]]
 
@@
 ```{r}
-data_soy <- readRDS(here("Shared/Data/ProcessedData/data_soy.rds"))
-
-#--- define the semi-parametric portion of the model ---#
-semi_formula <- formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))
-
-#--- semi-parametric FE wih fixest ---#
-semi_res <-
-  semi_ols(
-    semi_formula = semi_formula,
-    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
-    # fe = "",
-    cluster = "sc_dry",
-    data = data_soy
-  )
-
-#--- main regression results ---#
-y_res_int <- semi_res$fe_res
-
-r2_soy_wb <- r2(y_res_int, type = "r2")
-
-gam_y_res <- semi_res$gam_res
-
-#--- raw data for prediction ---#
-data_for_pred_soy <-
-  CJ(
-    balance = seq(min(data_soy$balance), max(data_soy$balance), length = 50),
-    sat_cat = unique(data_soy$sat_cat)
-  ) %>%
-  .[, sat := case_when(
-    sat_cat == "[39.4,82]" ~ 40,
-    sat_cat == "(82,194]" ~ 194,
-    sat_cat == "(194,708]" ~ 348,
-    sat_cat == "dryland" ~ 0
-  )] %>%
-  .[is.na(sat), sat := 0] %>%
-  # === can be any sc_code (need to shift yield for "average" county) ===#
-  .[, sc_code := data_soy$sc_code[1]] %>%
-  .[, sc_dry := data_soy$sc_dry[1]] %>%
-  # === can be any year (need to shift yield for "average" year) ===#
-  .[, year := 2009] %>%
-  # === fake yield data (necessary to take advantage of predict.gam) ===#
-  .[, yield := 0] %>%
-  .[, type := "pred"] %>%
-  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]
-
-#--- create bases  ---#
-yhat_data_soy <-
-  gen_smooth_data(
-    data = data_for_pred_soy,
-    gam_res = gam_y_res
-  ) %>%
-  .$data
-
-y_hat_with_se <- predict(y_res_int, newdata = yhat_data_soy, se.fit = TRUE)
-
-yhat_data_soy[, `:=`(
-  y_hat = y_hat_with_se$fit,
-  y_hat_se = y_hat_with_se$se.fit
-)]
-
-yhat_data_soy[, sat_cat_q :=
-  as.numeric(sat_cat) %>%
-  ordinal() %>%
-  paste0(., " quintile")]
-
-ir_share_data_soy <-
-  data_soy %>%
-  .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-  .[, balance_avg := mean(balance), by = sc_code] %>%
-  .[, type := "Soybean"]
+ir_share_data_soy <- reg_results_data[crop == "Soybean", share_data][[1]]
 
@@
 
+## With State-year FEs (Main)
+
+### Corn
+
+```{r}
+ir_share_data_corn[, state_year := paste0(state_name, year)]
+
+ir_share_res_corn <-
+  gam(
+    acres_ratio ~
+      s(sat, k = 3, m = 2) +
+      s(balance_avg, k = 3, m = 2) +
+      ti(sat, balance_avg, k = 3) +
+      factor(state_year),
+    data = ir_share_data_corn,
+    # weight = ~acres,
+    family = binomial
+  )
+
+share_pred_data_corn <-
+  CJ(
+    sat = seq(
+      min(ir_share_data_corn$sat),
+      max(ir_share_data_corn$sat),
+      length = 100
+    ),
+    balance_avg = mean(ir_share_data_corn$balance_avg),
+    year = 2016,
+    state_year = "Nebraska2009"
+  )
+
+ir_share_hat_data <-
+  predict(
+    ir_share_res_corn,
+    newdata = share_pred_data_corn,
+    type = "response",
+    se.fit = TRUE
+  )
+
+share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
+share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
+```
+
+### Soybean
+
+```{r}
+ir_share_data_soy[, state_year := paste0(state_name, year)]
+
+ir_share_res_soy <-
+  gam(
+    acres_ratio ~
+      s(sat, k = 4, m = 2) +
+      s(balance_avg, k = 3, m = 2) +
+      ti(sat, balance_avg, k = 3, m = 2) +
+      factor(state_year),
+    data = ir_share_data_soy,
+    # weight = ~acres,
+    family = binomial
+  )
+
+share_pred_data_soy <-
+  CJ(
+    sat = seq(
+      min(ir_share_data_soy$sat),
+      max(ir_share_data_soy$sat),
+      length = 100
+    ),
+    balance_avg = mean(ir_share_data_soy$balance_avg),
+    year = 2016,
+    state_year = "Nebraska2009"
+  )
+
+ir_share_hat_data <-
+  predict(
+    ir_share_res_soy,
+    newdata = share_pred_data_soy,
+    type = "response",
+    se.fit = TRUE
+  )
+
+share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
+share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
+```
+
+```{r}
+plot_data_ir_share_with_syfe <-
+  rbind(
+    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
+    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
+  ) %>%
+  .[, sfe := "With State-year FE"]
+```
+
 ```{r}
@@
     plot_data_ir_share_with_sfe,
+    plot_data_ir_share_with_syfe,
     plot_data_ir_share_no_sfe
@@
   facet_grid(type ~ sfe) +
-  theme_fig +
+  # theme_fig +
   theme(
@@
 
-s_deam_data <-
-  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m)), by = state_name] %>%
-  .[, type := "B. Demeaned by State Average"] %>%
+sy_demeaan_data <-
+  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = .(state_name, year)] %>%
+  .[, type := "B. Demeaned by State-Year Average"] %>%
   .[, .(sat_m, type)]
 
-c_deam_data <-
-  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m)), by = sc_code] %>%
-  .[, type := "C. Demeaned by County Average"] %>%
+c_y_demeaan_data <-
+  copy(ir_share_data_corn)[, sat_m := sat_m - mean(sat_m), by = sc_code] %>%
+  .[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  .[, type := "C. Demeaned by County and Yearly Average"] %>%
   .[, .(sat_m, type)]
 
-non_deam_data <-
-  ir_share_data_corn[, .(sat_m = sat_m - mean(sat_m))] %>%
-  .[, type := "A. Demeaned by Overall Average"]
+y_demeaan_data <-
+  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  .[, type := "A. Demeaned by Yearly Average"] %>%
+  .[, .(sat_m, type)]
 
-plot_data <- rbind(s_deam_data, c_deam_data, non_deam_data)
+plot_data <- rbind(sy_demeaan_data, c_y_demeaan_data, y_demeaan_data)
 

### 2023-07-24 Writing/manuscript.Rmd
updates

@@
   bookdown::pdf_document2:
+    extra_dependencies: ["float"]
     citation_package: natbib
@@
 #| echo: false
+#| cache: false
+knitr::opts_chunk$set(
+  echo = TRUE,
+  cache = FALSE,
+  fig.pos = "H",
+  out.extra = ""
+)
 
 options(knitr.duplicate.label = "allow")
-```
-
-```{r run-analys, cache = TRUE, include = FALSE}
-# knitr::purl(
-#   here::here("GitControlled/Codes/1_1_main_analysis.Rmd"),
-#   here::here("GitControlled/Codes/1_1_main_analysis.R"),
-# )
-# source(here::here("GitControlled/Codes/1_1_main_analysis.R"))
-
-# knitr::purl(
-#   here::here("GitControlled/Codes/1_2_sub_analysis.Rmd"),
-#   here::here("GitControlled/Codes/1_2_sub_analysis.R"),
-# )
-# source(here::here("GitControlled/Codes/1_2_sub_analysis.R"))
-
-# knitr::purl(
-#   here::here("GitControlled/Codes/2_figures_tables.Rmd"),
-#   here::here("GitControlled/Codes/2_figures_tables.R"),
-# )
-# source(here::here("GitControlled/Codes/2_figures_tables.R"))
-
-# knitr::purl(
-#   here::here("GitControlled/Codes/3_appendix.Rmd"),
-#   here::here("GitControlled/Codes/3_appendix.R"),
-# )
-
-# source(here::here("GitControlled/Codes/3_appendix.R"))
 ```
@@
 
-Figure \@ref(fig:yield-response) shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (1st row) and soybean (2nd row).
+```{r}
+test_results <- readRDS(here::here("Shared/Results/test_results_intensive.rds"))
+```
+
+Figure \@ref(fig:yield-response) shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B))^[Confidence intervals were not presented to avoid over-crowding the figures.
+Please see Appendix \@ref(reg-conf) for estimated yield response curves with 95 percent confidence interval.].
+The histogram of annual water deficit is presented at the top for each crop.
 In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean.
@@
 
-Differentiating irrigated yields by the level of aquifer thickness alters these responses in two ways.
+We now focus on comparing yield response to water deficitn among the three auifer thickness categories (all irrigated).
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \@ref(fig:yield-response)).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits [@rad2020effects, @foster2014modeling; @hrozencik2017heterogeneous].
-Second, at negative water deficits, counties with the second highest and highest levels of aquifer thickness (i.e., 2nd and 3rd quantile in Figure \@ref(fig:yield-response)) experience lower yields than counties with the lowest level of aquifer thickness (1st quantile), resulting in productivity close to rainfed yields.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits [@rad2020effects, @foster2014modeling; @hrozencik2017heterogeneous].
+However, yield gradually increases as water deficit increases for counties with the highest level of aquifer thickness.
+
+Figure \@ref(fig:yield-dif) shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quatile with the 95 percent confidence interval for corn an soybean.
+For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
+At an extrement water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is `r round(test_results[balance == 1000 & crop == "Corn" & comp_q == 3, dif_hat], digits = 2)` tonne/ha.
+At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \@ref(fig:yield-response)) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
 This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities [@foster2019assessing; @gibson2017case; @gibson2019benchmarking].
-However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e.
+Unlike corn, soybean yields are not statistically distinguishable among the three aquifer thickness groups.
+This is likey because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
+
+<!-- However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e.
 high deficits), in particular for corn.
+-->
+```{r yield-response, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn and soybean in US High Plains", out.width = "6.5in", cache = FALSE}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_intensive.pdf"))
+```
 
-```{r yield-response, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains", out.height = "5in"}
-knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_intensive.pdf"))
+```{r yield-dif, out.width = "6in", fig.cap = "The difference in yield for the counties with the second and third (highest) aquifer thickness groups relative to the first (lowest) aquifer thickness group."}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_ir_yield_dif.pdf"))
 ```
@@
 
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in moderate to extreme drought conditions, the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production.
+While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (more than 800mm water deficit), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production.
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
@@
 
-```{r ir-share, fig.cap = "The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)", out.height = "4in"}
+```{r ir-share, fig.cap = "The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)", out.width = "5in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share.pdf"))
@@
 
-The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
-Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean.
+```{r}
+results <- readRDS(here::here("Shared/Results/reg_results_data.rds"))
 
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm).
-However, beyond this threshold, higher levels of water deficit result in noticeable and significant non-linear reductions to average corn production.
+res_at_950 <-
+  rbindlist(results$avg_yield_data) %>%
+  .[(abs(balance - 950)) < 1, ]
+```
+The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
+Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, and 106.1 meters^[Confidence intervals were not presented to avoid over-crowding the figures.
+Please see Appendix \@ref(reg-conf) for estimated average yield response curves with 95 percent confidence interval.].
+While you can see similar trends as presented in Figure \@ref(fig:yield-response), the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronouced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
+
+Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 59.1m and 106.1m aquifer thickness leves relative to 12.2m with the 95 percent confidence interval for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 59.1m and 106.1m aquifer thickness, respectively).
+However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
 This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
 Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.
-greater than around 600mm) and are smaller in overall magnitude.
+greater than around 800mm) and are smaller in overall magnitude.
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
+Further, notice that the decline in average per-area production at a sesonal water deicit of 950mm is not linear.
+Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in `r round(res_at_950[sat == 348, avg_yield] - res_at_950[sat == 194, avg_yield], digits = 2)` tonne/ha of production loss for corn.
+However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in `r round(res_at_950[sat == 194, avg_yield] - res_at_950[sat == 40, avg_yield], digits = 2)` tonne/ha of production loss for corn.
+That is the same degree of decline is much more damaging when the decline is from a low aquifer thickness level thant a higher aquifer thickness level.
+This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \@ref(fig:ir-share).
+When aquifer thickness level is high enough (e.g., 100m), a small decline does not reduce the share of irrigated acres.
+However, when aquifer thickness level is low (e.g., 40m), a small decline does reduce the share of irrigated acres.
 
-```{r tot-impact, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness", out.height = "5in"}
+```{r tot-impact, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness", out.width = "6in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_total_impact.pdf"))
@@
 
-```{r dif-tot-impact, fig.cap = "Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean", cache = FALSE}
+```{r dif-tot-impact, fig.cap = "Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean", out.width = "6in", cache = FALSE}
 knitr::include_graphics(here::here("Shared/Results/g_avg_yield_dif.pdf"))
@@
 
-Aquifer depletion may impact resilience of irrigated crop production to drought through two main channels: (1) a reduction in per-area yields due to the economic or physical inability to meet crop water requirements; and (2) a reduction in the area of land irrigated, resulting in reduced production due to lower yields obtained through rainfed agriculture.
+Aquifer depletion may impact resilience of irrigated crop production to drought through two main channels: (1) a reduction in per-area irrigated yields due to the economic or physical inability to meet crop water requirements; and (2) a reduction in the area of land irrigated, resulting in reduced production due to lower yields obtained through rainfed agriculture.
 
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \@ref(spline-basis) for further explanations on how spline-based regression works).
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
@@
 State fixed effects, year fixed effects and error term are represented by $\theta_s$, $\mu_t$, and $v_{i,t}$, respectively.
-Unlike our regression model of the impact of water balance on yield, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
+One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
 The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time.
 Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.
+The trade-off here is tighter controls on potential unobservables that may be correlated with aquifer thickness (see Appendix \@ref(county-fe) for more discussions).
 
-county FE
-
-In the final step of our analysis, we then estimate the average crop yield for each county as a function of aquifer thickness and seasonal water deficit.
+In the final step of our analysis, we then estimate the average crop yield as a function of aquifer thickness and seasonal water deficit.
 The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \@ref(eq:eqn1), weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \@ref(eq:eqn2).
-Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area yields and the area of land irrigated.
+Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
+Its standard error is estimated using bootstrap clustered by county.
 
@@
 
-# More on the impact of water deficit on irrigated and average yield
+# More on the impact of water deficit on irrigated and average yield {#reg-conf}
+
+```{=Tex}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A.\arabic{figure}}
+```
 
@@
 
-```{r irrigated-yield-ind-corn, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
-knitr::include_graphics(here::here("Shared/Results/g_yield_with_conf_corn.pdf"))
+```{r irrigated-yield-ind-corn, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_with_conf_corn.pdf"))
 ```
 
-```{r irrigated-yield-ind-soy, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of soybean in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
-knitr::include_graphics(here::here("Shared/Results/g_yield_with_conf_soy.pdf"))
+```{r irrigated-yield-ind-soy, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_with_conf_soy.pdf"))
 ```
@@
 ```{r avg-yield-ind, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (The shaded areas represent the 95 percent confidence interval)", out.width = "6.5in", cache = FALSE}
-knitr::include_graphics(here::here("Shared/Results/g_avg_yield.rds"))
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_avg_yield.pdf"))
 ```
@@
 
-# County Fixed Effects in Irrigation Share Regression
+# County Fixed Effects in Irrigation Share Regression {#county-fe}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{A.\arabic{figure}}
+\renewcommand{\thefigure}{B.\arabic{figure}}
 ```
 
-Figure \@ref(fig:variation-left) presents the histogram of aquifer thickness after demeaning them by its overall average (a), state average (b), and county average (c) to gain insights into how much variation for identification is left under no county or state fixed effects (a), state fixed effects (b), and county fixed effects (c).
+Figure \@ref(fig:variation-left) presents the histogram of aquifer thickness after demeaning them by its overall average (A), state average (B), and county average (C) to gain insights into how much variation for identification is left under no county or state fixed effects (A), state fixed effects (B), and county fixed effects (C).
 As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation.
@@
 
-```{r variation-left, fig.cap = "Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.", out.height = "6in"}
+```{r variation-left, fig.cap = "Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.", out.width = "6in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_variation.pdf"))
@@
 
-```{r state-fe-with-without, fig.cap = "Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects", out.height = "6in"}
+```{r state-fe-with-without, fig.cap = "Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects", out.width = "6in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share_comp.pdf"))
@@
 
-# Spline basis functions and GAM estimation
+\clearpage
+
+# Spline basis functions and GAM estimation {#spline-basis}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
+\renewcommand{\thefigure}{C.\arabic{figure}}
 ```
@@
 
-```{r sp-basis, fig.cap = "Example of a series of thin-plate spline basis functions", out.height = "4in", out.width = "6in"}
-library(mgcv)
-
-set.seed(83944)
-
-# === generate data ===#
-N <- 400 # number of observations
-x <- seq(1, 250, length = N)
-y_det <- 240 * (1 - 0.4 * exp(-0.03 * x))
-e <- 25 * rnorm(N) # error
-data <- data.table(x = x, y = y_det + e, y_det = y_det)
-
-basis_data <-
-  gam(y_det ~ s(x, k = 5, bs = "cr"), data = data) %>%
-  predict(., type = "lpmatrix") %>%
-  data.table() %>%
-  .[, x := data[, x]] %>%
-  melt(id.var = "x") %>%
-  .[, variable := case_when(
-    variable == "(Intercept)" ~ "Intercept",
-    variable == "s(x).1" ~ "1st basis",
-    variable == "s(x).2" ~ "2nd basis",
-    variable == "s(x).3" ~ "3rd basis",
-    variable == "s(x).4" ~ "4th basis"
-  )] %>%
-  .[, variable := factor(
-    variable,
-    levels= c("Intercept", "1st basis", "2nd basis", "3rd basis","4th basis")
-  )]
-
-ggplot(data = basis_data) +
-  geom_line(aes(y = value, x = x)) +
-  facet_grid(variable ~ .) +
-  theme_bw() +
-  ylab("")
-
+```{r sp-basis, fig.cap = "Example of a series of thin-plate spline basis functions", out.width = "6in"}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_basis.pdf"))
 ```
@@
 In the first case, $\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}$.
-In the second case, $\{\beta_1, \dots, \beta_5\} = \{50, 20, 40, 60, 80\}$.
+In the second case, $\{\beta_1, \dots, \beta_5\} = \{20, -20, -40, 40, 60\}$.
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
@@
 
-```{r ex-fx, fig.cap = "Example of a function represented by different coefficient values", out.height = "4in", out.width = "6in"}
-library(patchwork)
-
-g_il_1 <-
-  data.table(
-    variable = unique(basis_data$variable),
-    coef = rep(60, 5)
-  ) %>%
-  .[basis_data, on = "variable"] %>%
-  .[, sum(coef * value), by = x] %>%
-  ggplot(data = .) +
-  geom_line(aes(y = V1, x = x)) +
-  theme_bw() +
-  ylab("y") +
-  ggtitle("Case 1")
-
-g_il_2 <-
-  data.table(
-    variable = unique(basis_data$variable),
-    coef = c(50, 20, 40, 60, 80)
-  ) %>%
-  .[basis_data, on = "variable"] %>%
-  .[, sum(coef * value), by = x] %>%
-  ggplot(data = .) +
-  geom_line(aes(y = V1, x = x)) +
-  theme_bw() +
-  ylab("y")  +
-  ggtitle("Case 2")
-
-g_il_1 / g_il_2
+```{r ex-fx, fig.cap = "Example of a function represented by different coefficient values", out.width = "6in"}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_il_y.pdf"))
 ```

### 2023-07-24 Writing/manuscript.tex
updates

@@
 \linenumbers
+\usepackage{float}
 \ifLuaTeX
@@
 
-Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (1st row) and soybean (2nd row).
+Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B))\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
+Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.}.
+The histogram of annual water deficit is presented at the top for each crop.
 In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean.
@@
 
-Differentiating irrigated yields by the level of aquifer thickness alters these responses in two ways.
+We now focus on comparing yield response to water deficitn among the three auifer thickness categories (all irrigated).
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
-Second, at negative water deficits, counties with the second highest and highest levels of aquifer thickness (i.e., 2nd and 3rd quantile in Figure \ref{fig:yield-response}) experience lower yields than counties with the lowest level of aquifer thickness (1st quantile), resulting in productivity close to rainfed yields.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
+However, yield gradually increases as water deficit increases for counties with the highest level of aquifer thickness.
+
+Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quatile with the 95 percent confidence interval for corn an soybean.
+For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
+At an extrement water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is 1.15 tonne/ha.
+At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
 This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}.
-However, it is important to highlight that impacts of aquifer thickness on yields for negative deficits is much less pronounced than that for extreme droughts (i.e.~high deficits), in particular for corn.
+Unlike corn, soybean yields are not statistically distinguishable among the three aquifer thickness groups.
+This is likey because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=5in]{Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_yield_intensive}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains}\label{fig:yield-response}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn and soybean in US High Plains}\label{fig:yield-response}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_ir_yield_dif}
+
+}
+
+\caption{The difference in yield for the counties with the second and third (highest) aquifer thickness groups relative to the first (lowest) aquifer thickness group.}\label{fig:yield-dif}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=4in]{Figures/g_share}
+{\centering \includegraphics[width=5in,height=500px,]{Figures/g_share}
 
@@
 
-The overall effects of aquifer depletion on drought-related production losses is a function of both per-area yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean.
+The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
+Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, and 106.1 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
+Please see Appendix \ref{reg-conf} for estimated average yield response curves with 95 percent confidence interval.}.
+While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronouced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
 
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm).
-However, beyond this threshold, higher levels of water deficit result in noticeable and significant non-linear reductions to average corn production.
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 59.1m and 106.1m aquifer thickness leves relative to 12.2m with the 95 percent confidence interval for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 300mm).
+However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
 This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.~greater than around 600mm) and are smaller in overall magnitude.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.~greater than around 800mm) and are smaller in overall magnitude.
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
+Further, notice that the decline in average per-area production at a sesonal water deicit of 950mm is not linear.
+Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in 1.01 tonne/ha of production loss for corn.
+However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in 2.12 tonne/ha of production loss for corn.
+That is the same degree of decline is much more damaging when the decline is from a low aquifer thickness level thant a higher aquifer thickness level.
+This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \ref{fig:ir-share}.
+When aquifer thickness level is high enough (e.g., 100m), a small decline does not reduce the share of irrigated acres.
+However, when aquifer thickness level is low (e.g., 40m), a small decline does reduce the share of irrigated acres.
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=5in]{Figures/g_total_impact}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_total_impact}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics{../../Shared/Results/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_avg_yield_dif}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=6in]{Figures/g_y_wd_all}
+{\centering \includegraphics[width=750px,height=6in,]{Figures/g_y_wd_all}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6in]{Figures/g_map}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_map}
 
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient.
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
@@
 State fixed effects, year fixed effects and error term are represented by \(\theta_s\), \(\mu_t\), and \(v_{i,t}\), respectively.
-Unlike our regression model of the impact of water balance on yield, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
+One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
 The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time.
 Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.
-
-county FE
+The trade-off here is tighter controls on potential unobservables that may be correlated with aquifer thickness (see Appendix \ref{county-fe} for more discussions).
 
@@
 The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
-Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area yields and the area of land irrigated.
+Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
 
@@
 
-\hypertarget{more-on-the-impact-of-water-deficit-on-irrigated-and-average-yield}{%
-\section{More on the impact of water deficit on irrigated and average yield}\label{more-on-the-impact-of-water-deficit-on-irrigated-and-average-yield}}
+\hypertarget{reg-conf}{%
+\section{More on the impact of water deficit on irrigated and average yield}\label{reg-conf}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A.\arabic{figure}}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6in]{../../Shared/Results/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_corn}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-corn}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-corn}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6in]{../../Shared/Results/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_soy}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of soybean in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-soy}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-soy}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6.5in]{../../Shared/Results/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Shared/Results/g_avg_yield}
 
@@
 
-\hypertarget{county-fixed-effects-in-irrigation-share-regression}{%
-\section{County Fixed Effects in Irrigation Share Regression}\label{county-fixed-effects-in-irrigation-share-regression}}
+\hypertarget{county-fe}{%
+\section{County Fixed Effects in Irrigation Share Regression}\label{county-fe}}
 
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{A.\arabic{figure}}
+\renewcommand{\thefigure}{B.\arabic{figure}}
 
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its overall average (a), state average (b), and county average (c) to gain insights into how much variation for identification is left under no county or state fixed effects (a), state fixed effects (b), and county fixed effects (c).
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its overall average (A), state average (B), and county average (C) to gain insights into how much variation for identification is left under no county or state fixed effects (A), state fixed effects (B), and county fixed effects (C).
 As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation.
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=6in]{Figures/g_variation}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_variation}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[height=6in]{Figures/g_share_comp}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_share_comp}
 
@@
 
-\hypertarget{spline-basis-functions-and-gam-estimation}{%
-\section{Spline basis functions and GAM estimation}\label{spline-basis-functions-and-gam-estimation}}
+\clearpage
+
+\hypertarget{spline-basis}{%
+\section{Spline basis functions and GAM estimation}\label{spline-basis}}
 
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
+\renewcommand{\thefigure}{C.\arabic{figure}}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=4in]{manuscript_files/figure-latex/sp-basis-1}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_basis}
 
@@
 In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\).
-In the second case, \(\{\beta_1, \dots, \beta_5\} = \{50, 20, 40, 60, 80\}\).
+In the second case, \(\{\beta_1, \dots, \beta_5\} = \{20, -20, -40, 40, 60\}\).
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=4in]{manuscript_files/figure-latex/ex-fx-1}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_il_y}
 

### 2023-07-24 Codes/Functions/functions.R

@@ -1,6 +1,4 @@
-
 prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
-
   # === saturated thickness threshold in meter ===#
   sat_thld <- measurements::conv_unit(sat_thld_m, "m", "ft")
 
@@ -149,386 +147,310 @@ semi_ols <- function(semi_formula, c_formula, cluster, data) {
   # cor(reg_data$model_X)
 }
 
+reg_yield <- function(yield_data) {
+  #++++++++++++++++++++++++++++++++++++
+  #+ For Debugging
+  #++++++++++++++++++++++++++++++++++++
+  # yield_data <- all_results[crop == "corn", data][[1]]
 
-# /*=================================================*/
-#' # Prepare a data set
-# /*=================================================*/
-#' used in 1_regression_analysis.R
+  #++++++++++++++++++++++++++++++++++++
+  #+ Main
+  #++++++++++++++++++++++++++++++++++++
 
-# vars <- c("balance", "et0")
+  #---------------------
+  #- Regression
+  #---------------------
+  #--- define the semi-parametric portion of the model ---#
+  semi_formula <-
+    formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))
 
-gen_pred_data <- function(data, vars, sat_ls) {
-  var_1 <- vars[1]
-  var_2 <- vars[2]
-
-  return_data <-
-    expand.grid(
-      var_1 =
-        seq(
-          min(data[, ..var_1], na.rm = TRUE),
-          max(data[, ..var_1], na.rm = TRUE),
-          length = 50
-        ),
-      var_2 =
-        seq(
-          min(data[, ..var_2], na.rm = TRUE),
-          max(data[, ..var_2], na.rm = TRUE),
-          length = 50
-        ),
-      sat = sat_ls
-    ) %>%
-    data.table() %>%
-    .[
-      ,
-      sat_cat := cut(
-        sat,
-        breaks = sat_breaks,
-        include.lowest = TRUE
-      )
-    ] %>%
-    .[is.na(sat_cat), sat_cat := "dryland"] %>%
-    setnames(
-      c("var_1", "var_2"),
-      vars
+  #--- semi-parametric FE wih fixest ---#
+  semi_res_corn <-
+    semi_ols(
+      semi_formula = semi_formula,
[…]

### 2023-07-29 Codes/0_7_ssurgo.Rmd
change the aquifer thickness threshold to 9

@@
 
+  ssurgo_data_sum[, sc_code := field$sc_code]
   return(ssurgo_data_sum)
@@
     } else {
+      temp <- readRDS(file_name)
+      if (ncol(temp) < 7) {
+        temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+        saveRDS(temp, paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
+      } else {
         print(paste0("skipping ", x))
+      }
     }
@@
 
-# lapply(
-#   1:nrow(all_counties),
-#   \(x) {
-#     print(x)
-#     file_name <- paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds")
-#     if (!file.exists(file_name)) {
-#       temp <- get_ssurgo_props(field = all_counties[x, ], vars)
-#       saveRDS(temp, paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
-#     }
-#   }
-# )
+lapply(
+  1:nrow(all_counties),
+  \(x) {
+    print(x)
+    file_name <- paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds")
+    if (!file.exists(file_name)) {
+      temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+      saveRDS(temp, paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
+    } else {
+      temp <- readRDS(file_name)
+      if (ncol(temp) < 7) {
+        temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+        saveRDS(temp, paste0("Shared/Data/RawData/SSURGO/ssurgo_", x, "rds"))
+      }
+    }
+  }
+)
 ```
@@
   purrr::map(readRDS) %>%
-  lapply(function(x) x[, geometry := NULL]) %>%
-  rbindlist() %>%
-  .[, `:=`(
-    state = NULL,
-    state_name = NULL,
-    ir_area_ratio = NULL,
-    in_hpa = NULL
-  )]
+  rbindlist()
 
 saveRDS(ssurgo_soil, here("Shared/Data/ProcessedData/ssurgo_soil.rds"))
-# all_counties[, state_name] %>% unique()
 ```

### 2023-07-29 Codes/1_Analysis.rmd
change the aquifer thickness threshold to 9

---
title: "Regression Analysis and Prediction"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare data

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    sat_thld_m = 9, # saturated thickness has to be at least 12 meters
    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
    balance_thld = c(-1200, 200),
    sat_cat_num = 3,
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop") %>%
  #--- further necessary transformation ---#
  dplyr::mutate(data = purrr::map(
    reg_data_y,
    \(reg_data_y) {
      copy(reg_data_y) %>%
        #--- bu/acre to tonne/ha ---#
        .[, yield := yield * 12.553 / 200] %>%
        .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
        .[, state_year := paste0(state_name, "_", year)] %>%
        .[, balance := -balance]
    }
  ))

# saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))

# ggplot(reg_data$data[[1]]) +
#   geom_point(aes(y = acres_ratio, x = sat)) +
#   geom_smooth(aes(y = acres_ratio, x = sat))
```

# Regression and prediciton

## Yield

```{r}
set.seed(283397)

#--- global parameters ---#
sc_base <- "31011"
sc_dry_base <- "31011_0"

main_analysis <-
  reg_data %>%
  rowwise() %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(balance_seq = list(
    seq(
      min(data$balance),
      max(data$balance),
      length = 100
    )
  )) %>%
  dplyr::mutate(sat_seq_eval = list(
    c(10, 50, 90, 130)
  )) %>%
  dplyr::mutate(share_reg_data = list(
    data %>%
      .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
      .[, balance_avg := mean(balance), by = sc_code]
  )) %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(sat_seq = list(
    share_reg_data[, seq(min(sat),
      max(sat),
      length = 100
    )] %>%
      c(., sat_seq_eval)
  )) %>%
  #--- values of the controls to be used in prediction ---#
  dplyr::mutate(
    sandtotal_med = median(share_reg_data$sandtotal_r),
    silttotal_med = median(share_reg_data$silttotal_r),
    awc_med = median(share_reg_data$awc_r)
  ) %>%
  #++++++++++++++++++++++++++++++++++++
  #+ Main analysis
  #++++++++++++++++++++++++++++++++++++
  dplyr::mutate(yield_pred_data = list(
    yield_analysis(data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
  )) %>%
  dplyr::mutate(share_pred_data = list(
    share_analysis(share_reg_data, sat_seq, sandtotal_med, silttotal_med, awc_med)
  )) %>%
  dplyr::mutate(avg_yield_pred_data = list(
    get_average_yield(yield_pred_data, share_pred_data)
  ))
```

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Bootstrap
#++++++++++++++++++++++++++++++++++++
boot_results_added <-
  main_analysis %>%
  #--- bootsrap data ---#
  dplyr::mutate(boot_data = list(
    replicate(1000, boot(data), simplify = FALSE)
  )) %>%
  #--- get share data ---#
  dplyr::mutate(share_boot_data = list(
    lapply(
      1:length(boot_data),
      \(x) {
        boot_data[[x]] %>%
          .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
          .[, balance_avg := mean(balance), by = sc_code]
      }
    )
  )) %>%
  #--- yield analysis ---#
  dplyr::mutate(yield_pred_data_boot = list(
    parallel::mclapply(
      1:length(boot_data),
      \(x) {
        print(x)
        yield_analysis(boot_data[[x]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  #--- share analysis ---#
  dplyr::mutate(share_pred_data_boot = list(
    parallel::mclapply(
      1:length(share_boot_data),
      \(x) {
        print(x)
        share_analysis(share_boot_data[[x]], sat_seq, sandtotal_med, silttotal_med, awc_med)
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  #--- average yield analysis ---#
  dplyr::mutate(avg_yield_pred_data_boot = list(
    parallel::mclapply(
      1:length(boot_data),
      \(x) {
        get_average_yield(yield_pred_data_boot[[x]], share_pred_data_boot[[x]])
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  dplyr::mutate(
    yield_pred_data_boot = list(rbindlist(yield_pred_data_boot)),
    share_pred_data_boot = list(rbindlist(share_pred_data_boot)),
    avg_yield_pred_data_boot = list(rbindlist(avg_yield_pred_data_boot))
  ) %>%
  dplyr::select(
    crop, data, sat_seq, sat_text_data,
    yield_pred_data, yield_pred_data_boot,
    share_pred_data, share_pred_data_boot,
    avg_yield_pred_data, avg_yield_pred_data_boot,
  )

saveRDS(boot_results_added, "Shared/Results/boot_results_added.rds")
```

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Summarize
#++++++++++++++++++++++++++++++++++++
boot_results_added <- readRDS("Shared/Results/boot_results_added.rds")

all_results <-
  boot_results_added %>%
  # rowwise() %>%
  dplyr::mutate(yield_response_sum = list(
    yield_pred_data_boot[, .(
      # y_hat_boot = median(y_hat, na.rm = TRUE),
      y_hat_se = sd(y_hat, na.rm = TRUE),
      # dif_y_hat_boot = mean(dif_y_hat, na.rm = TRUE),
      dif_y_hat_se = sd(dif_y_hat, na.rm = TRUE)
    ),
    by = .(balance, sat, sat_cat)
    ] %>%
      .[yield_pred_data, on = c("balance", "sat")] %>%
      .[, `:=`(
        lower = y_hat - 1.96 * y_hat_se,
        upper = y_hat + 1.96 * y_hat_se,
        dif_lower = dif_y_hat - 1.96 * dif_y_hat_se,
        dif_upper = dif_y_hat + 1.96 * dif_y_hat_se
      )]
  )) %>%
  dplyr::mutate(share_sum = list(
    share_pred_data_boot[, .(
      # ir_share_hat = mean(ir_share_hat, na.rm = TRUE),
      ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE)
    ),
    by = .(sat)
    ] %>%
      .[share_pred_data, on = "sat"] %>%
      .[, `:=`(
        lower = ir_share_hat - 1.96 * ir_share_hat_se,
        upper = ir_share_hat + 1.96 * ir_share_hat_se
      )]
  )) %>%
  dplyr::mutate(avg_yield_sum = list(
    avg_yield_pred_data_boot[, .(
      avg_yield_se = sd(avg_yield, na.rm = TRUE),
      dif_avg_yield_se = sd(dif_avg_yield, na.rm = TRUE)
    ),
    by = .(balance, sat)
    ] %>%
      .[avg_yield_pred_data, on = c("balance", "sat")] %>%
      .[, `:=`(
        upper = avg_yield + 1.96 * avg_yield_se,
        lower = avg_yield - 1.96 * avg_yield_se,
        dif_upper = dif_avg_yield + 1.96 * dif_avg_yield_se,
        dif_lower = dif_avg_yield - 1.96 * dif_avg_yield_se
      )]
  )) %>%
  data.table()

saveRDS(all_results, "Shared/Results/all_results.rds")
```

### 2023-07-29 Codes/2_figures_tables.rmd
change the aquifer thickness threshold to 9

@@
 ```{r}
-results <- readRDS("Shared/Results/reg_results_data.rds")
 all_results <- readRDS("Shared/Results/all_results.rds")
-
 ```
@@
 
-```{r, eval = FALSE}
-data_corn <- results[crop == "Corn", data][[1]]
-data_soy <- results[crop == "Soybean", data][[1]]
-
-raw_data_both <-
-  rbind(data_corn, data_soy) %>%
-  .[, dry_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")]
-```
-
-## Summary data
-
-```{r, eval = FALSE}
-(
-  sum_table <-
-    modelsummary::datasummary(
-      type * (yield + balance) ~ dry_or_not * (Mean + SD),
-      data = raw_data_both,
-      output = "kableExtra"
-      # output = "latex"
-    )
-)
-```
-
 ## Boxplot of water deficit by year
@@
 ```{r}
-data_corn <- results[crop == "Corn", data][[1]]
-data_soy <- results[crop == "Soybean", data][[1]]
+data_corn <- all_results[crop == "corn", data][[1]]
+data_soy <- all_results[crop == "soy", data][[1]]
 
@@
   rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
-  .[, .(wd = mean(balance)), by = .(year, type)] %>%
+  .[, .(wd = mean(balance)), by = .(year, crop)] %>%
   .[, ]
@@
   .[, ir_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")] %>%
-  .[, .(yield = mean(yield), wb = mean(balance)), by = .(year, type, ir_or_not)]
+  .[, .(yield = mean(yield), wb = mean(balance)), by = .(year, crop, ir_or_not)]
 
@@
   geom_line(
-    data = yield_data[type == "Corn" & ir_or_not == "Irrigated", ],
+    data = yield_data[crop == "corn" & ir_or_not == "Irrigated", ],
     aes(y = yield, x = year, color = "Irrigated Yield")
@@
   geom_point(
-    data = yield_data[type == "Corn" & ir_or_not == "Irrigated", ],
+    data = yield_data[crop == "corn" & ir_or_not == "Irrigated", ],
     aes(y = yield, x = year, color = "Irrigated Yield"),
@@
   geom_line(
-    data = yield_data[type == "Corn" & ir_or_not == "Dryland", ],
+    data = yield_data[crop == "corn" & ir_or_not == "Dryland", ],
     aes(y = yield, x = year, color = "Dryland Yield")
@@
   geom_point(
-    data = yield_data[type == "Corn" & ir_or_not == "Dryland", ],
+    data = yield_data[crop == "corn" & ir_or_not == "Dryland", ],
     aes(y = yield, x = year, color = "Dryland Yield"),
@@
   geom_line(
-    data = wd_data[type == "Corn"],
+    data = wd_data[crop == "corn"],
     aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
@@
   geom_point(
-    data = wd_data[type == "Corn"],
+    data = wd_data[crop == "corn"],
     aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
@@
   geom_line(
-    data = yield_data[type == "Soybean" & ir_or_not == "Irrigated", ],
+    data = yield_data[crop == "soy" & ir_or_not == "Irrigated", ],
     aes(y = yield, x = year, color = "Irrigated Yield")
@@
   geom_point(
-    data = yield_data[type == "Soybean" & ir_or_not == "Irrigated", ],
+    data = yield_data[crop == "soy" & ir_or_not == "Irrigated", ],
     aes(y = yield, x = year, color = "Irrigated Yield"),
@@
   geom_line(
-    data = yield_data[type == "Soybean" & ir_or_not == "Dryland", ],
+    data = yield_data[crop == "soy" & ir_or_not == "Dryland", ],
     aes(y = yield, x = year, color = "Dryland Yield")
@@
   geom_point(
-    data = yield_data[type == "Soybean" & ir_or_not == "Dryland", ],
+    data = yield_data[crop == "soy" & ir_or_not == "Dryland", ],
     aes(y = yield, x = year, color = "Dryland Yield"),
@@
   geom_line(
-    data = wd_data[type == "Soybean"],
+    data = wd_data[crop == "soy"],
     aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
@@
   geom_point(
-    data = wd_data[type == "Soybean"],
+    data = wd_data[crop == "soy"],
     aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
@@
   ggplot(data = .) +
-  geom_histogram(aes(x = sat * 0.3048), fill = "white", color = "blue") +
+  geom_histogram(aes(x = sat), fill = "white", color = "blue") +
   facet_grid(type ~ ., scale = "free_y") +
@@
 
-```{r echo = F}
+```{r echo = F, eval = F}
 # === hpa ===#
@@
   rbind(counties_corn_ir, counties_soy_ir) %>%
-  mutate(sat_meter = sat * 0.3048)
+  mutate(sat_meter = sat)
 
@@
   geom_sf(data = counties_corn_all, fill = "gray") +
-  geom_sf(data = counties_corn_ir, aes(fill = sat * 0.3048)) +
+  geom_sf(data = counties_corn_ir, aes(fill = sat)) +
   scale_fill_viridis_c(
@@
   geom_sf(data = counties_soy_all, fill = "gray") +
-  geom_sf(data = counties_soy_ir, aes(fill = sat * 0.3048)) +
+  geom_sf(data = counties_soy_ir, aes(fill = sat)) +
   scale_fill_viridis_c(
@@
 ```{r}
-g_ir_yield_both <-
+plot_yield_data <-
   all_results[, .(crop, yield_response_sum)] %>%
-  dplyr::mutate(plot_data = purrr::map2(yield_response_sum, crop, \(x, crop) {
-    x[, sat_cat_text := case_when(
-      sat_cat == "[39.4,82]" ~ "1st quantile: [12m, 25m]",
-      sat_cat == "(82,194]" ~ "2nd quantile: (25m, 59.1m]",
-      sat_cat == "(194,708]" ~ "3rd quantile: (59.1m, 215m]",
-      sat_cat == "dryland" ~ "Rainfed"
-    )] %>%
-      .[, sat_cat_text := factor(
-        sat_cat_text,
-        levels = c("Rainfed", "1st quantile: [12m, 25m]", "2nd quantile: (25m, 59.1m]", "3rd quantile: (59.1m, 215m]")
-      )] %>%
-      .[, crop_type := crop] %>%
+  rowwise() %>%
+  dplyr::mutate(plot_data = list(
+    yield_response_sum[, crop_type := crop] %>%
       .[, crop_type := case_when(
         crop_type == "corn" ~ "A. Corn",
-        crop_type == "soy" ~ "(b) Sybean"
+        crop_type == "soy" ~ "B. Sybean"
       )]
-  })) %>%
-  dplyr::mutate(g_yield = purrr::map(plot_data, \(plot_data) {
+  )) %>%
+  dplyr::mutate(sat_cat_text_ls = list(
+    plot_data[, .(sat_rank, sat_cat_text)] %>%
+      .[!is.na(sat_rank), ] %>%
+      unique(by = "sat_rank") %>%
+      .[order(sat_rank), sat_cat_text]
+  )) %>%
+  dplyr::mutate(plot_data = list(
+    plot_data %>%
+      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
+      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))]
+  ))
+
+g_ir_yield_both <-
+  plot_yield_data %>%
+  dplyr::mutate(g_yield = list(
     ggplot(plot_data) +
@@
       theme(legend.position = "bottom")
-  }))
+  ))
 
@@
 g_yield_with_conf <-
-  all_results[, .(crop, yield_response_sum)] %>%
-  dplyr::mutate(plot_data = purrr::map2(yield_response_sum, crop, \(x, crop) {
-    x[, sat_cat_text := case_when(
-      sat_cat == "[39.4,82]" ~ "1st quantile: [12m, 25m]",
-      sat_cat == "(82,194]" ~ "2nd quantile: (25m, 59.1m]",
-      sat_cat == "(194,708]" ~ "3rd quantile: (59.1m, 215m]",
-      sat_cat == "dryland" ~ "Rainfed"
-    )] %>%
-      .[, sat_cat_text := factor(
-        sat_cat_text,
-        levels = c("Rainfed", "1st quantile: [12m, 25m]", "2nd quantile: (25m, 59.1m]", "3rd quantile: (59.1m, 215m]")
-      )] %>%
-      .[, crop_type := crop] %>%
-      .[, crop_type := case_when(
-        crop_type == "corn" ~ "A. Corn",
-        crop_type == "soy" ~ "(b) Sybean"
-      )]
-  })) %>%
-  mutate(g_fig = purrr::map(
-    plot_data,
-    \(plot_data) {
+  plot_yield_data %>%
+  mutate(g_fig = list(
       ggplot(plot_data) +
@@
         )
-    }
   )) %>%
@@
 
-ggsave("Shared/Results/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
-ggsave("Shared/Results/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
+ggsave("GitControlled/Writing/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
+ggsave("GitControlled/Writing/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
 ```
@@
 g_share <-
-  ggplot(plot_data_ir_share[sat <= 500, ]) +
-  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
+  ggplot(plot_data_ir_share[sat <= 150, ]) +
+  geom_line(aes(y = ir_share_hat, x = sat)) +
   geom_ribbon(
@@
       ymax = upper,
-      x = sat * 0.3048
+      x = sat
     ),
@@
   )] %>%
-  .[, sat_m := measurements::conv_unit(sat, "ft", "m") %>% format(digits = 3)] %>%
   nest_by(crop_type) %>%
@@
     ggplot(data) +
-      geom_line(aes(y = avg_yield, x = balance, color = sat_m)) +
+      geom_line(aes(y = avg_yield, x = balance, color = factor(sat))) +
       theme_fig +
@@
 
-# g_total <-
-#   ggplot(data = combined_data) +
-#   geom_line(aes(y = avg_yield, x = balance, color = sat_cat_text)) +
-#   facet_wrap(crop_type ~ ., scale = "free_y", ncol = 1) +
-#   theme_fig +
-#   ylim(0, NA) +
-#   ylab("Yield (tonne/ha)") +
-#   scale_color_discrete(
-#     "Aquifer Thickness",
-#     guide = guide_legend(
-#       title.position = "top",
-#       nrow = 1
-#     )
-#   ) +
-#   scale_x_continuous(name = "Water Deficit (mm)", breaks = (-2:11) * 100) +
-#   theme(
-#     legend.position = "bottom",
-#     strip.background = element_blank(),
-#     legend.text = element_text(size = 8)
-#   )
-
 ggsave(
@@
 ```{r}
-test_results <- readRDS("Shared/Results/test_results_intensive.rds")
-
-plot_data <-
-  test_results %>%
-  .[, upper := dif_hat + se * 1.96] %>%
-  .[, lower := dif_hat - se * 1.96] %>%
+g_dif <-
+  all_results[, .(crop, yield_response_sum)] %>%
+  rowwise() %>%
+  dplyr::mutate(plot_data = list(
+    yield_response_sum[, crop_type := crop] %>%
       .[, crop_type := case_when(
-    crop == "Corn" ~ "A. Corn",
-    crop == "Soybean" ~ "B. Soybean"
-  )] %>%
-  .[, as_q := case_when(
-    comp_q == 2 ~ "2nd Aquifer Thickness Quantile",
-    comp_q == 3 ~ "3rd Aquifer Thickness Quantile"
-  )] %>%
-  nest_by(crop) %>%
-  data.table() %>%
-  dplyr::mutate(g_fig = purrr::map(
-    data,
-    \(data) {
-      ggplot(data) +
-        facet_grid(. ~ as_q) +
+        crop_type == "corn" ~ "A. Corn",
+        crop_type == "soy" ~ "B. Sybean"
+      )]
+  )) %>%
+  dplyr::mutate(sat_cat_text_ls = list(
+    plot_data[, .(sat_rank, sat_cat_text)] %>%
+      .[!is.na(sat_rank), ] %>%
+      unique(by = "sat_rank") %>%
+      .[order(sat_rank), sat_cat_text]
+  )) %>%
+  dplyr::mutate(plot_data = list(
+    plot_data %>%
+      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
+      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))] %>%
+      .[!is.na(sat_rank), ] %>%
+      .[sat_rank != 1, ] %>%
+      unique(by = c("crop_type", "balance", "sat_cat"))
+  )) %>%
+  dplyr::mutate(g_fig = list(
+    ggplot(plot_data) +
+      facet_grid(. ~ sat_cat_text) +
       geom_ribbon(
@@
           x = balance,
-            ymin = lower,
-            ymax = upper
+          ymin = dif_lower,
+          ymax = dif_upper
         ),
@@
           x = balance,
-            y = dif_hat
+          y = dif_y_hat
         ),
@@
       )
-    }
   ))
@@
   cowplot::plot_grid(
-    plot_data$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
-    plot_data$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
+    g_dif$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
+    g_dif$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
     ncol = 1,
@@
 ```{r}
-boot_results_data <- readRDS("Shared/Results/boot_results.rds")
-
 g_avg_yield_dif <-
-  boot_results_data[, .(crop, boot_results)] %>%
-  unnest(cols = c(boot_results)) %>%
+  all_results[, .(crop, avg_yield_sum)] %>%
+  unnest(cols = c(avg_yield_sum)) %>%
   data.table() %>%
-  .[, .(
-    dif_yield = mean(dif_yield),
-    lower = quantile(dif_yield, prob = 0.05),
-    upper = quantile(dif_yield, prob = 0.95)
-  ),
-  by = .(sat, balance, crop)
-  ] %>%
   .[, sat_text := paste0(
     "Aquifer Thickness = ",
-    measurements::conv_multiunit(sat, "ft", "m") %>% format(digits = 3),
+    format(sat, digits = 3),
     "m"
   )] %>%
-  .[sat != 40, ] %>%
+  .[sat != min(sat), ] %>%
   nest_by(crop) %>%
@@
     ggplot(data) +
-      geom_ribbon(aes(x = balance, ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
-      geom_line(aes(x = balance, y = dif_yield), color = "red") +
+      geom_ribbon(aes(x = balance, ymin = dif_lower, ymax = dif_upper), alpha = 0.3, fill = "blue") +
+      geom_line(aes(x = balance, y = dif_avg_yield), color = "red") +
       geom_hline(yintercept = 0) +
@@
 
-ggsave("Shared/Results/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
+ggsave("GitControlled/Writing/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
 ```
@@
 g_ay_response <-
-  boot_results_data[, .(crop, boot_results)] %>%
-  unnest(cols = c(boot_results)) %>%
+  all_results[, .(crop, avg_yield_sum)] %>%
+  unnest(cols = c(avg_yield_sum)) %>%
   data.table() %>%
-  .[, .(
-    avg_yield = mean(avg_yield),
-    lower = quantile(avg_yield, prob = 0.05),
-    upper = quantile(avg_yield, prob = 0.95)
-  ),
-  by = .(sat, balance, crop)
-  ] %>%
   .[, sat_text := paste0(
     "Aquifer Thickness = ",
-    measurements::conv_multiunit(sat, "ft", "m") %>% format(digits = 3),
+    format(sat, digits = 3),
     "m"
   )] %>%
+  .[sat != min(sat), ] %>%
   nest_by(crop) %>%
@@
       facet_grid(. ~ sat_text, scales = "free_y") +
-      ylim(min(data$avg_yield) - 1, max(data$avg_yield) + 0.5) +
+      ylim(min(data$avg_yield) * 0.7, max(data$avg_yield) + 1) +
       xlab("Water Deficit (mm)") +
@@
 
-ggsave("Shared/Results/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
+ggsave("GitControlled/Writing/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
 ```

### 2023-07-29 Codes/A2_state_fe.rmd
change the aquifer thickness threshold to 9

@@
 ```{r}
-reg_results_data <- readRDS(here::here("Shared/Results/reg_results_data.rds"))
+all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
 ```
 
-## Without State FEs
+```{r}
+#++++++++++++++++++++++++++++++++++++
+#+ Formula
+#++++++++++++++++++++++++++++++++++++
+formula_text <- "acres_ratio ~
+    s(sat, k = 5, m = 2) + s(balance_avg, k = 5, m = 2) +
+    sandtotal_r + silttotal_r + awc_r +
+    factor(year)"
 
-### Corn
+share_formula_yfe <- formula(formula_text)
 
-```{r}
-ir_share_data_corn <- reg_results_data[crop == "Corn", share_data][[1]]
+share_formula_sandyfe <-
+  formula_text %>%
+  gsub("factor\\(year\\)", "factor\\(year\\) + factor\\(state_name\\)", .) %>%
+  formula()
 
-ir_share_res_corn <-
+share_formula_syfe <-
+  formula_text %>%
+  gsub("factor\\(year\\)", "factor\\(state_year\\)", .) %>%
+  formula()
+
+formula_data <-
+  data.table(
+    formula = c(share_formula_yfe, share_formula_sandyfe, share_formula_syfe),
+    type = c("Year FE", "State and Year FE", "State-Year FE")
+  )
+
+results <-
+  all_results[, .(crop, data)] %>%
+  expand_grid_df(., formula_data) %>%
+  rowwise() %>%
+  dplyr::mutate(ir_share_data = list(
+    data[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
+      .[, balance_avg := mean(balance), by = sc_code]
+  )) %>%
+  dplyr::mutate(ir_share_res = list(
     gam(
-    acres_ratio ~
-      s(sat, k = 3, m = 2) + s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3) + factor(year),
+      formula,
       data = ir_share_data_corn,
-    # weight = ~acres,
       family = binomial
     )
-
-share_pred_data_corn <-
+  )) %>%
+  dplyr::mutate(share_pred_data = list(
     CJ(
       sat = seq(
-      min(ir_share_data_corn$sat),
-      max(ir_share_data_corn$sat),
+        min(ir_share_data$sat),
+        max(ir_share_data$sat),
         length = 100
       ),
-    balance_avg = mean(ir_share_data_corn$balance_avg),
-    year = 2016
+      balance_avg = mean(ir_share_data$balance_avg),
+      awc_r = mean(ir_share_data$awc_r),
+      sandtotal_r = mean(ir_share_data$sandtotal_r),
+      silttotal_r = mean(ir_share_data$silttotal_r),
+      state_year = "Nebraska_2009",
+      state_name = "Nebraska",
+      year = 2009
     )
-
-ir_share_hat_data <-
+  )) %>%
+  dplyr::mutate(ir_share_hat_data = list(
     predict(
-    ir_share_res_corn,
-    newdata = share_pred_data_corn,
+      ir_share_res,
+      newdata = share_pred_data,
       type = "response",
@@
     )
-
-share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
-```
-
-### Soybean
-
-```{r}
-ir_share_data_soy <- reg_results_data[crop == "Soybean", share_data][[1]]
-
-ir_share_res_soy <-
-  gam(
-    acres_ratio ~
-      s(sat, k = 4, m = 2) +
-      s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3, m = 2) +
-      factor(state_name) +
-      factor(year),
-    data = ir_share_data_soy,
-    # weight = ~acres,
-    family = binomial
+  )) %>%
+  dplyr::mutate(share_pred_data = list(
+    dplyr::mutate(
+      share_pred_data,
+      ir_share_hat = ir_share_hat_data$fit,
+      ir_share_hat_se = ir_share_hat_data$se.fit
     )
-
-share_pred_data_soy <-
-  CJ(
-    sat = seq(
-      min(ir_share_data_soy$sat),
-      max(ir_share_data_soy$sat),
-      length = 100
-    ),
-    balance_avg = mean(ir_share_data_soy$balance_avg),
-    year = 2016,
-    state_name = "Nebraska"
-  )
-
-ir_share_hat_data <-
-  predict(
-    ir_share_res_soy,
-    newdata = share_pred_data_soy,
-    type = "response",
-    se.fit = TRUE
-  )
-
-share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
+  )) %>%
+  dplyr::select(crop, type, share_pred_data) %>%
+  unnest(cols = c(share_pred_data)) %>%
+  data.table()
 ```
@@
 ```{r}
-plot_data_ir_share_no_sfe <-
-  rbind(
-    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
-    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
-  ) %>%
-  .[, sfe := "No State FE"]
-```
-
-## With State FEs (Main)
-
-### Corn
-
-```{r}
-ir_share_res_corn <-
-  gam(
-    acres_ratio ~
-      s(sat, k = 3, m = 2) +
-      s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3) +
-      factor(state_name) +
-      factor(year),
-    data = ir_share_data_corn,
-    # weight = ~acres,
-    family = binomial
-  )
-
-share_pred_data_corn <-
-  CJ(
-    sat = seq(
-      min(ir_share_data_corn$sat),
-      max(ir_share_data_corn$sat),
-      length = 100
-    ),
-    balance_avg = mean(ir_share_data_corn$balance_avg),
-    year = 2016,
-    state_name = "Nebraska"
-  )
-
-ir_share_hat_data <-
-  predict(
-    ir_share_res_corn,
-    newdata = share_pred_data_corn,
-    type = "response",
-    se.fit = TRUE
-  )
-
-share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
-```
-
-### Soybean
-
-```{r}
-ir_share_res_soy <-
-  gam(
-    acres_ratio ~
-      s(sat, k = 4, m = 2) +
-      s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3, m = 2) +
-      factor(state_name) +
-      factor(year),
-    data = ir_share_data_soy,
-    # weight = ~acres,
-    family = binomial
-  )
-
-share_pred_data_soy <-
-  CJ(
-    sat = seq(
-      min(ir_share_data_soy$sat),
-      max(ir_share_data_soy$sat),
-      length = 100
-    ),
-    balance_avg = mean(ir_share_data_soy$balance_avg),
-    year = 2016,
-    state_name = "Nebraska"
-  )
-
-ir_share_hat_data <-
-  predict(
-    ir_share_res_soy,
-    newdata = share_pred_data_soy,
-    type = "response",
-    se.fit = TRUE
-  )
-
-share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
-```
-
-```{r}
-plot_data_ir_share_with_sfe <-
-  rbind(
-    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
-    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
-  ) %>%
-  .[, sfe := "With State FE"]
-```
-
-## With State-year FEs (Main)
-
-### Corn
-
-```{r}
-ir_share_data_corn[, state_year := paste0(state_name, year)]
-
-ir_share_res_corn <-
-  gam(
-    acres_ratio ~
-      s(sat, k = 3, m = 2) +
-      s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3) +
-      factor(state_year),
-    data = ir_share_data_corn,
-    # weight = ~acres,
-    family = binomial
-  )
-
-share_pred_data_corn <-
-  CJ(
-    sat = seq(
-      min(ir_share_data_corn$sat),
-      max(ir_share_data_corn$sat),
-      length = 100
-    ),
-    balance_avg = mean(ir_share_data_corn$balance_avg),
-    year = 2016,
-    state_year = "Nebraska2009"
-  )
-
-ir_share_hat_data <-
-  predict(
-    ir_share_res_corn,
-    newdata = share_pred_data_corn,
-    type = "response",
-    se.fit = TRUE
-  )
-
-share_pred_data_corn[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_corn[, ir_share_hat_se := ir_share_hat_data$se.fit]
-```
-
-### Soybean
-
-```{r}
-ir_share_data_soy[, state_year := paste0(state_name, year)]
-
-ir_share_res_soy <-
-  gam(
-    acres_ratio ~
-      s(sat, k = 4, m = 2) +
-      s(balance_avg, k = 3, m = 2) +
-      ti(sat, balance_avg, k = 3, m = 2) +
-      factor(state_year),
-    data = ir_share_data_soy,
-    # weight = ~acres,
-    family = binomial
-  )
-
-share_pred_data_soy <-
-  CJ(
-    sat = seq(
-      min(ir_share_data_soy$sat),
-      max(ir_share_data_soy$sat),
-      length = 100
-    ),
-    balance_avg = mean(ir_share_data_soy$balance_avg),
-    year = 2016,
-    state_year = "Nebraska2009"
-  )
-
-ir_share_hat_data <-
-  predict(
-    ir_share_res_soy,
-    newdata = share_pred_data_soy,
-    type = "response",
-    se.fit = TRUE
-  )
-
-share_pred_data_soy[, ir_share_hat := ir_share_hat_data$fit]
-share_pred_data_soy[, ir_share_hat_se := ir_share_hat_data$se.fit]
-```
-
-```{r}
-plot_data_ir_share_with_syfe <-
-  rbind(
-    share_pred_data_corn[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Corn"],
-    share_pred_data_soy[, .(ir_share_hat, ir_share_hat_se, sat)][, type := "Soybean"]
-  ) %>%
-  .[, sfe := "With State-year FE"]
-```
-
-```{r}
-plot_data_ir_share <-
-  rbind(
-    plot_data_ir_share_with_sfe,
-    plot_data_ir_share_with_syfe,
-    plot_data_ir_share_no_sfe
-  )
-
 g_share_comp <-
-  ggplot(plot_data_ir_share[sat <= 500, ]) +
-  geom_line(aes(y = ir_share_hat, x = sat * 0.3048)) +
+  results[sat <= 500, ] %>%
+  .[, type := factor(type, levels = c("Year FE", "State and Year FE", "State-Year FE"))] %>%
+  ggplot() +
+  geom_line(aes(y = ir_share_hat, x = sat)) +
   geom_ribbon(
@@
       ymax = ir_share_hat + 1.96 * ir_share_hat_se,
-      x = sat * 0.3048
+      x = sat
     ),
@@
   ylab("Share of Irrigated Acres") +
-  facet_grid(type ~ sfe) +
-  # theme_fig +
+  facet_grid(type ~ crop) +
+  theme_fig +
   theme(

### 2023-07-29 Writing/manuscript.Rmd
change the aquifer thickness threshold to 9

@@
 
-Figure \@ref(fig:yield-response) shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B))^[Confidence intervals were not presented to avoid over-crowding the figures.
+Figure \@ref(fig:yield-response) shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B)^[Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \@ref(reg-conf) for estimated yield response curves with 95 percent confidence interval.].
@@
 
-We now focus on comparing yield response to water deficitn among the three auifer thickness categories (all irrigated).
+We now focus on comparing yield response to water deficit among the three aquifer thickness categories (all irrigated).
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \@ref(fig:yield-response)).
@@
 
-Figure \@ref(fig:yield-dif) shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quatile with the 95 percent confidence interval for corn an soybean.
+Figure \@ref(fig:yield-dif) shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95 percent confidence interval for corn an soybean.
 For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extrement water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is `r round(test_results[balance == 1000 & crop == "Corn" & comp_q == 3, dif_hat], digits = 2)` tonne/ha.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is `r round(test_results[balance == 1000 & crop == "Corn" & comp_q == 3, dif_hat], digits = 2)` tonne/ha.
 At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \@ref(fig:yield-response)) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
@@
 Unlike corn, soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likey because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
+This is likely because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
 
@@
 ```{r}
-results <- readRDS(here::here("Shared/Results/reg_results_data.rds"))
+all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
 
 res_at_950 <-
-  rbindlist(results$avg_yield_data) %>%
-  .[(abs(balance - 950)) < 1, ]
+  all_results[, .(crop, avg_yield_pred_data)] %>%
+  unnest(cols = c(avg_yield_pred_data)) %>%
+  data.table() %>%
+  .[(abs(balance - 950)) < 1, by = crop]
 ```
 The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
-Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, and 106.1 meters^[Confidence intervals were not presented to avoid over-crowding the figures.
+Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, 106.1, and 153.0  meters^[Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \@ref(reg-conf) for estimated average yield response curves with 95 percent confidence interval.].
-While you can see similar trends as presented in Figure \@ref(fig:yield-response), the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronouced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
+While you can see similar trends as presented in Figure \@ref(fig:yield-response), the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
 
-Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 59.1m and 106.1m aquifer thickness leves relative to 12.2m with the 95 percent confidence interval for corn and soybean.
+Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 59.1m, 106.1m, and 153.0m aquifer thickness levels relative to 12.2m with the 95 percent confidence interval for corn and soybean.
 For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 59.1m and 106.1m aquifer thickness, respectively).
@@
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
-Further, notice that the decline in average per-area production at a sesonal water deicit of 950mm is not linear.
-Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in `r round(res_at_950[sat == 348, avg_yield] - res_at_950[sat == 194, avg_yield], digits = 2)` tonne/ha of production loss for corn.
-However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in `r round(res_at_950[sat == 194, avg_yield] - res_at_950[sat == 40, avg_yield], digits = 2)` tonne/ha of production loss for corn.
-That is the same degree of decline is much more damaging when the decline is from a low aquifer thickness level thant a higher aquifer thickness level.
+Further, notice that the decline in average per-area production at a seasonal water deficit of 950mm is not linear.
+Going from 153.0m aquifer thickness to 106.1m (decline of 46.9m) would result in `r round(res_at_950[sat == 502 & crop == "corn", avg_yield] - res_at_950[sat == 348 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
+Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in `r round(res_at_950[sat == 348 & crop == "corn", avg_yield] - res_at_950[sat == 194 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in `r round(res_at_950[sat == 194 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
 This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \@ref(fig:ir-share).
@@
 ```{r dif-tot-impact, fig.cap = "Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean", out.width = "6in", cache = FALSE}
-knitr::include_graphics(here::here("Shared/Results/g_avg_yield_dif.pdf"))
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_avg_yield_dif.pdf"))
 ```
@@
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalizez additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
 This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \theta_s + \mu_t + v_{i,t} (\#eq:eqn2)
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} (\#eq:eqn2)
 \end{equation}
@@
 The impact of aquifer thickness is captured in a similar way with $\tau_{l}(\cdot)$ and $\beta_{l}$ being the $l$th spline basis function and its coefficient, respectively.
-State fixed effects, year fixed effects and error term are represented by $\theta_s$, $\mu_t$, and $v_{i,t}$, respectively.
+County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by $X_i$.
+State-year fixed effects and error term are represented by $\theta_{s,t}$ and $v_{i,t}$, respectively.
 One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
@@
 All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All the raw data-sets used in this study are publicly accessible.
+All the computations were done in R [@R] and all the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
@@
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
-The R tidyUSDA package was used to automatically download the relevant data.
+The R tidyUSDA package [@RtidyUSDA] was used to automatically download the relevant data.
 This data is used for Figure \@ref(fig:deficit-yield-hist) and the regression analysis.
@@
 This data is used for Figure \@ref(fig:sat-map) and the regression analysis.
+Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which is used for regression analysis.
+The R soilDB package [@Rsoildb] was used to download the data.
+Regression analyses are implemented using the R fixest package [@Rfixest].
+Other key R packages include ggplot2 [@Rggplot2] for creating plots, data.table [@Rdatatable] for data wrangling, and sf [@Rsf] for spatial data operations.
 The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.
@@
 
-In the main article, the confidence intervals of the estiamted yield responses to water deficit were not presented as they would crowd the figures excessively.
-This section provides visualization of yield response with the 95 $\%$ confidence interval to convery the degree of uncertainty in our estimation.
+In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented as they would crowd the figures excessively.
+This section provides visualization of yield response with the 95 $\%$ confidence interval to convey the degree of uncertainty in our estimation.
 Figures \@ref(fig:irrigated-yield-ind-corn) and \@ref(fig:irrigated-yield-ind-soy) present the impact of water deficit on crop yield by aquifer thickness category with the 95$\%$ confidence interval for irrigated corn and soybean, respectively.
@@
 
-Figure \@ref(fig:state-fe-with-without) compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean.
-For corn, the negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included.
-Therefore, it may be that there are important unobservables that are spatially correlated with saturated thickness, but their correlation may not be very strong.
-Given the direction of potential bias, providing tighter controls with county fixed effects may result in even more significant drop in the share of irrigated acres in response to a reduction in aquifer thickness.
-That is we may be underestimating the true negative impact of a reduction in aquifer thickness.
+Figure \@ref(fig:state-fe-with-without) compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
+Most important, adding state fixed effects, which control for unobservables at the state level, does not change the estimation results much.
+Therefore, even though there may be unobservables that are spatially correlated with aquifer thickness, their correlation may not be very strong to cause significant bias.
+Clearly, this does not mean that our inability to control for other county-level unobservables beyond sand percentage, silt percentage, and water holding capacity is not causing bias in our estimation of the impact of aquifer thickness on the share of irrigated acres.
 
@@
 The shape of $f(x)$ is not known to the researcher.
-GAM attemps to represent $f(x)$ with a series of spline basis functions and their coefficients like below.
+GAM attempts to represent $f(x)$ with a series of spline basis functions and their coefficients like below.
 
@@
 Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of $x$ on $y$.
-Figure \@ref(fig:ex-fx) presents two examples of two different sets of coefficients and their resulting represention of $f(x)$.
+Figure \@ref(fig:ex-fx) presents two examples of two different sets of coefficients and their resulting representation of $f(x)$.
 In the first case, $\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}$.

### 2023-07-29 Writing/manuscript_DR.Rmd
change the aquifer thickness threshold to 9

@@
 corn_avg_yield_low <- data_final_corn[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
-corn_avg_yield_high <- data_final_corn[balance == balace_closest & sat_cat == "(194,708]", avg_yield]
+corn_avg_yield_high <- data_final_corn[balance == balace_closest & sat_cat == "(186,708]", avg_yield]
 
@@
 soy_avg_yield_low <- data_final_soy[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
-soy_avg_yield_high <- data_final_soy[balance == balace_closest & sat_cat == "(194,708]", avg_yield]
+soy_avg_yield_high <- data_final_soy[balance == balace_closest & sat_cat == "(186,708]", avg_yield]
 

### 2023-07-29 Writing/manuscript_comp.tex
change the aquifer thickness threshold to 9

--- 2023-07-24 Writing/manuscript.tex
@@
 
-Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B))\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
+Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B)\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.}.
@@
 
-We now focus on comparing yield response to water deficitn among the three auifer thickness categories (all irrigated).
+We now focus on comparing yield response to water deficit among the three aquifer thickness categories (all irrigated).
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
@@
 
-Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quatile with the 95 percent confidence interval for corn an soybean.
+Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95 percent confidence interval for corn an soybean.
 For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extrement water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is 1.15 tonne/ha.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is 1.15 tonne/ha.
 At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
@@
 The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, and 106.1 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
+Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, 106.1, and 153.0 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \ref{reg-conf} for estimated average yield response curves with 95 percent confidence interval.}.
-While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronouced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
+While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
 
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 59.1m and 106.1m aquifer thickness leves relative to 12.2m with the 95 percent confidence interval for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 300mm).
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 59.1m, 106.1m, and 153.0m aquifer thickness levels relative to 12.2m with the 95 percent confidence interval for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 59.1m and 106.1m aquifer thickness, respectively).
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
@@
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
-Further, notice that the decline in average per-area production at a sesonal water deicit of 950mm is not linear.
-Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in 1.01 tonne/ha of production loss for corn.
-However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in 2.12 tonne/ha of production loss for corn.
-That is the same degree of decline is much more damaging when the decline is from a low aquifer thickness level thant a higher aquifer thickness level.
+Further, notice that the decline in average per-area production at a seasonal water deficit of 950mm is not linear.
+Going from 153.0m aquifer thickness to 106.1m (decline of 46.9m) would result in tonne/ha of production loss for corn.
+This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
+Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in tonne/ha of production loss for corn.
+However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in tonne/ha of production loss for corn.
+That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
 This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \ref{fig:ir-share}.
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_avg_yield_dif}
 
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + \theta_s + \mu_t + v_{i,t} \label{eq:eqn2}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
-State fixed effects, year fixed effects and error term are represented by \(\theta_s\), \(\mu_t\), and \(v_{i,t}\), respectively.
+County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by \(X_i\).
+State-year fixed effects and error term are represented by \(\theta_{s,t}\) and \(v_{i,t}\), respectively.
 One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
@@
 
-In the final step of our analysis, we then estimate the average crop yield for each county as a function of aquifer thickness and seasonal water deficit.
+In the final step of our analysis, we then estimate the average crop yield as a function of aquifer thickness and seasonal water deficit.
 The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
+Its standard error is estimated using bootstrap clustered by county.
 
@@
 All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All the raw data-sets used in this study are publicly accessible.
+All the computations were done in R \citep{R} and all the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
@@
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
-The R tidyUSDA package was used to automatically download the relevant data.
+The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data.
 This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
@@
 This data is used for Figure \ref{fig:sat-map} and the regression analysis.
+Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which is used for regression analysis.
+The R soilDB package \citep{Rsoildb} was used to download the data.
+Regression analyses are implemented using the R fixest package \citep{Rfixest}.
+Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
 The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.
@@
 
-In the main article, the confidence intervals of the estiamted yield responses to water deficit were not presented as they would crowd the figures excessively.
-This section provides visualization of yield response with the 95 \(\%\) confidence interval to convery the degree of uncertainty in our estimation.
+In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented as they would crowd the figures excessively.
+This section provides visualization of yield response with the 95 \(\%\) confidence interval to convey the degree of uncertainty in our estimation.
 Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\(\%\) confidence interval for irrigated corn and soybean, respectively.
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_corn}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_soy}
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Shared/Results/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_avg_yield}
 
@@
 
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean.
-For corn, the negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included.
-Therefore, it may be that there are important unobservables that are spatially correlated with saturated thickness, but their correlation may not be very strong.
-Given the direction of potential bias, providing tighter controls with county fixed effects may result in even more significant drop in the share of irrigated acres in response to a reduction in aquifer thickness.
-That is we may be underestimating the true negative impact of a reduction in aquifer thickness.
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
+Most important, adding state fixed effects, which control for unobservables at the state level, does not change the estimation results much.
+Therefore, even though there may be unobservables that are spatially correlated with aquifer thickness, their correlation may not be very strong to cause significant bias.
+Clearly, this does not mean that our inability to control for other county-level unobservables beyond sand percentage, silt percentage, and water holding capacity is not causing bias in our estimation of the impact of aquifer thickness on the share of irrigated acres.
 
@@
 The shape of \(f(x)\) is not known to the researcher.
-GAM attemps to represent \(f(x)\) with a series of spline basis functions and their coefficients like below.
+GAM attempts to represent \(f(x)\) with a series of spline basis functions and their coefficients like below.
 
@@
 Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of \(x\) on \(y\).
-Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting represention of \(f(x)\).
+Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting representation of \(f(x)\).
 In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\).

### 2023-07-29 Codes/Functions/functions.R

@@ -1,9 +1,9 @@
 prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
   # === saturated thickness threshold in meter ===#
-  sat_thld <- measurements::conv_unit(sat_thld_m, "m", "ft")
-
   # data <- reg_data$data[[1]]
   # sat_breaks <- reg_data$sat_breaks[[1]]
+
+  # reg_data_y <- reg_data$reg_data_y[[1]]
 
   reg_data <-
     data.table(
@@ -16,7 +16,9 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
         # === lower bound ===#
         .[balance >= balance_thld[1], ] %>%
         # === upper bound ===#
-        .[balance <= balance_thld[2], ]
+        .[balance <= balance_thld[2], ] %>%
+        #--- unit conversion ---#
+        .[, sat := measurements::conv_unit(sat, "ft", "m")]
     )) %>%
     # === select observations for each crop type ===#
     mutate(data = list(
@@ -29,13 +31,13 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
         # 0: dryland
         # If irrigated, sat is set to be at least 0.01.
         # See 0_6_merge_datasets.R
-        .[sat >= sat_thld | sat == 0, ]
+        .[sat >= sat_thld_m | sat == 0, ]
     )) %>%
     # === sat breaks for grouping ===#
     #' dryland production will have NA
     mutate(sat_breaks = list(
       quantile(
-        base_data[ir == "ir" & sat >= sat_thld, sat],
+        base_data[ir == "ir" & sat >= sat_thld_m, sat],
         prob = seq(0, 1, length = sat_cat_num + 1),
         na.rm = TRUE
       )
@@ -53,32 +55,30 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
           )
         ] %>%
         .[is.na(sat_cat), sat_cat := "dryland"] %>%
-        .[, sat_cat_i := factor(paste0("sc_", as.numeric(factor(sat_cat))))] %>%
+        .[, sat_q_rank := as.numeric(factor(sat_cat))] %>%
[…]

### 2023-07-29 Codes/spinned/1_analysis.R

@@ -124,8 +124,8 @@ data_for_pred_corn <-
   # .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
   .[, sat := case_when(
     sat_cat == "[39.4,82]" ~ 40,
-    sat_cat == "(82,194]" ~ 100,
-    sat_cat == "(194,708]" ~ 200,
+    sat_cat == "(75.7,186]" ~ 100,
+    sat_cat == "(186,708]" ~ 200,
     sat_cat == "dryland" ~ 0
   )] %>%
   .[is.na(sat), sat := 0] %>%

### 2023-07-29 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.rmd
analysis updated

@@
 
-\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}
+\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
+We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}
 
 1.
-Aquifer thickness versus saturated thickness – The authors speak of "aquifer saturated thickness", which is a redundant term.
+Aquifer thickness versus saturated thickness – The authors speak of ``aquifer saturated thickness'', which is a redundant term.
 Remember, the definition of an aquifer is a saturated unit – there is no unsaturated aquifer.
-Please use "aquifer thickness" to avoid propagating this confusing usage into the future.
+Please use ``aquifer thickness'' to avoid propagating this confusing usage into the future.
 
-\textcolor{blue}{All the instances of "saturated thickness" and "aquifer saturated thickness" have been replaced by "aquifer thickness".}
+\textcolor{blue}{We agree that there is some redundant use of words here.
+All the instances of ``saturated thickness'' and ``aquifer saturated thickness'' have been replaced by ``aquifer thickness'' in the revised manuscript.}
 
@@
 If we do so, then the same quantile would represent different aquifer thickness range every year.
-When modeling the impact of aquifer thickness, this is not desirable because we are implicitly making assumptions that different ranges of aquifer thickness would have the same impact.}
--  Second point: \textcolor{blue}{Actually, 0% quntile is 12 m. "9m" was repalced with "12m" in the current manuscript.}
--  Third point: \textcolor{blue}{No interesting insights will be gained from them because they are merely shifters without any intractions with aquifer thickness.
-They tell us how much higher yield in a single county relative to another county on average.
-But, we do not beleive that is useful information at least in this artcile.}
--  Fourth point: \textcolor{blue}{This is a very important point. We have added texts describing the implications of this fact in page xx.}
+When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk.}
+
+-  Second point: \textcolor{blue}{Thank you for highlighting this point.
+For the 0\% quantile, we assume a saturated thickness of 9m.
+This minimum of saturated thickness was selected based on REF, which defines 9m as the minimum aquifer thickness needed to provide a functioning water supply.
+We have modified the text on page X lines X-X of the revised manuscript to note these points.}
+
+-  Third point: \textcolor{blue}{Yes, country and year fixed effects are calculated as part of the regressions.
+However, these variables can be thought of simply as shifters that tell us how much higher average yields are in a single county relative to another county and to account for trends in yields over time.
+As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analysing these variables - and indeed it is standard to not report and analyse fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
+
+-  Fourth point: 
+  \textcolor{blue}{This is an important point, which we thank the reviewer for highlighting.
+The reviewer is correct that unobserved factors, such as aquifer heterogeneity influencing well yield response to changes in saturated thickness, are a likely underlying cause of the uncertainty ranges found in our analyses.
+Variations in aquifer conductivity would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between saturated thickness and drought risk exposure that would contribute to uncertainty in our estimates of observed relationships between drought impacts and aquifer conditions.
+Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought.
+To highlight these points, we have added the following text on page X lines X-X of our revised manuscript:}
+
+  \textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer saturated thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger saturated thickness (Butler et al., 2013; Korus and Hensen, 2020).
+While we identify statistically robust impacts of saturated thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.''}
+
+  \textcolor{blue}{New references added:}
+
+  + \textcolor{blue}{Butler, J.
+J., Stotler, R.
+L., Whittemore, D.
+O., and Reboulet, E.
+C.
+(2013).
+Interpretation of water level changes in the High Plains aquifer in western Kansas.
+Groundwater, 51(2), 180-190.}
+
+  + \textcolor{blue}{Korus, J.
+T., and Hensen, H.
+J.
+(2020).
+Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks.
+Environmental Earth Sciences, 79(16), 382.}
 
 3.
-Conditioned on drought risk exposure – I wasn’t sure what the authors meant by this in the second paragraph of Section 2.2.
+Conditioned on drought risk exposure – I wasn't sure what the authors meant by this in the second paragraph of Section 2.2.
 This must be related to the water deficit but the lack of explanation in the main text or the methods section left me wondering.
@@
 
-\textcolor{blue}{What we meant by "conditiond on drought risk exposure" is "after taking into account the impact of climate variables." In order to avoid confusion, we simply removed it from the sentence in the current manuscript.}
+\textcolor{blue}{By ``conditioned on drought risk exposure'', we were referring to the fact that Figure 4 displays the relationship between saturated thickness and the share of production area that is irrigated while also accounting for variations in climate across the aquifer.
+I.e.
+that the Figure shows the impact of saturated thickness on share of production irrigated after removing the effects of climate on share of production irrigated.
+To improve clarity, we have modified this phrase to now state ``...after accounting for spatio-temporal variations in drought risk exposure across the aquifer''.}
 
-4. Equation 2 – Please define all terms in the equation and provide an explanation of what is meant by "estimated semi-parametrically".
+4. Equation 2 – Please define all terms in the equation and provide an explanation of what is meant by ``estimated semi-parametrically''.
 
-\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained.
-We eliminated the word semi-parametrically as it is not ncecessary.
-Generalized additive model that estiamtes the impact of a variable using splines without functional form assumption are sometimes referred to semi-parametric regression.
-That's what we meant by "estimate semi-parametrically" in our previous manuscript.}
+\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained in Section 4.2 of our revised manuscript.
+The word semi-parametrically refers to the fact that generalized additive models, such as used in our analysis, estimate the impact of a variable using splines without functional form assumption.
+These are sometimes referred to as semi-parametric regression methods.
+However, we agree that the term is potentially confusing, and therefore have removed the term ``semi-parametrically'' in our revised manuscript.}
 
 5.
-Kansas and California efforts – I’ve seen a number of papers describing efforts in Kansas and California to address aquifer depletion via LEMAs (Kansas) and SGMA (California).
+Kansas and California efforts – I've seen a number of papers describing efforts in Kansas and California to address aquifer depletion via LEMAs (Kansas) and SGMA (California).
 The 1985-2016 period examined by the authors was before most of the efforts in Kansas got started.
@@
 
-\textcolor{blue}{Tim?}
+\textcolor{blue}{This is an interesting idea - thank you for raising it.
+We agree that efforts to improve aquifer management could help to mitigate some of the changes in productivity driven by climate and aquifer conditions in our analysis.
+Greater attention to groundwater conservation, and use of improved scheduling technologies such as soil moisture probes, would be expected to mitigate production losses in years with negative water deficits (i.e.
+wetter years) which we attribute to increased possibility for over-irrigation when well capacities are high (and hence when farmers do not face or observe physical scarcity).
+In years with large positive water deficits (i.e.
+drought years), improved irrigation water management would also be expected to mitigate some of the additional production risk caused by lower saturated thickness.
+However, we would still expect lower saturated thickness to result in elevated exposure to drought risk even in these circumstances, as declining well yields would create a physical limit to water supply which would stop farmers being able to fully meet crop water needs in these years.
+Formally testing these effects is beyond the scope and data of our study, but we agree that these are valuable areas for future work and have added the following text to the discussion of our revised manuscript on page X lines X-X:}
+
+\textcolor{blue}{``Improvements to irrigation scheduling and water management practices, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas (Deines et al., 2019; Glose et al., 2022) and proposed as part of California's SGMA (Sustainable Groundwater Management Act) (Berbel and Esteban, 2019; Lubell et al., 2020), could also play an important role in improving the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+These kinds of interventions would be likely to reduce - but not eliminate - the increases in drought risk caused by declining aquifer saturated thickness, and could bring additional benefits of reducing over-use of water in wetter years through greater awareness of efficiency saving potentials (Foster et al., 2019) and the importance of conservation (Marston et al., 2022) amongst growers.''}
+
+\textcolor{blue}{New references added:}
+
+  + \textcolor{blue}{Glose, T.
+J., Zipper, S., Hyndman, D.
+W., Kendall, A.
+D., Deines, J.
+M., and Butler, J.
+J.
+(2022).
+Quantifying the impact of lagged hydrological responses on the effectiveness of groundwater conservation.
+Water Resources Research, 58(7), e2022WR032295.}
+
+  + \textcolor{blue}{Berbel, J., and Esteban, E. (2019). Droughts as a catalyst for water policy change. Analysis of Spain, Australia (MDB), and California. Global Environmental Change, 58, 101969.}
+
+  + \textcolor{blue}{Lubell, M., Blomquist, W., and Beutler, L.
+(2020).
+Sustainable groundwater management in California: A grand experiment in environmental governance.
+Society and Natural Resources, 33(12), 1447-1467.}
+
+  + \textcolor{blue}{Marston, L.
+T., Zipper, S., Smith, S.
+M., Allen, J.
+J., Butler, J.
+J., Gautam, S., and David, J.
+Y.
+(2022).
+The importance of fit in groundwater self-governance.
+Environmental Research Letters, 17(11), 111001.}
 
@@
 
-+ Section 1, 3rd Paragraph – "Two aquifer pathways exist through aquifer depletion will.." – confusing sentence – please replace "through" with "by which".
++ Section 1, 3rd Paragraph – ``Two aquifer pathways exist through aquifer depletion will..'' – confusing sentence – please replace ``through'' with ``by which''.
 
@@
 
-+ Section 1, 4th Paragraph – "we generate empirical evidence" seems awkward to me.
-How about replacing with "we present empirical evidence"?
-Next sentence – "Our analysis advances on previous…" reads better as "Our analysis extends previous…" There are many issues such as these and the previous comment.
++ Section 1, 4th Paragraph – ``we generate empirical evidence'' seems awkward to me.
+How about replacing with ``we present empirical evidence''?
+Next sentence – ``Our analysis advances on previous $\dots$'' reads better as ``Our analysis extends previous $\dots$'' There are many issues such as these and the previous comment.
 Please address such issues so that the manuscript is an easier read.
@@
 
-+ Section 2.1 title – I would delete "marginally" as it is not needed and some will find it confusing.
++ Section 2.1 title – I would delete ``marginally'' as it is not needed and some will find it confusing.
 
@@
 
-+ Section 2.3, First Paragraph – please clarify that "average per-area production" is a combination of irrigated and rainfed yields. That would make the manuscript easier to follow.
++ Section 2.3, First Paragraph – please clarify that ``average per-area production'' is a combination of irrigated and rainfed yields. That would make the manuscript easier to follow.
 
-\textcolor{blue}{We added "(weighted-average of irrigated and rainfed yields, where weights are production area shares)" so that it is clear what average per-area production means.}
+\textcolor{blue}{We added ``weighted-average of irrigated and rainfed yields, where weights are production area shares'' so that it is clear what average per-area production means.}
 
@@
 
-+ Line before Equation (2) – "REF" should be replaced with the actual reference.
++ Line before Equation (2) – ``REF'' should be replaced with the actual reference.
 
@@
 
-+ Discussion, First Paragraph – "reduction in aquifer depletion" should be "reduction in aquifer thickness".
++ Discussion, First Paragraph – ``reduction in aquifer depletion'' should be ``reduction in aquifer thickness''.
 
@@
 
-+ Section 4.2, 3rd Paragraph – "the the impact" – remember to spell and grammar check prior to submission.
++ Section 4.2, 3rd Paragraph – ``the the impact'' – remember to spell and grammar check prior to submission.
 
@@
 
-+ Reference in Section 3 – I doubt that "Butler Jr" is the surname, please replace with "Butler". Suffixes apparently aren’t handled well by some of the bibliography generation programs.
++ Reference in Section 3 – I doubt that ``Butler Jr'' is the surname, please replace with ``Butler''. Suffixes apparently aren't handled well by some of the bibliography generation programs.
 

### 2023-07-29 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.tex
analysis updated

--- 2023-07-18 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.tex
@@
 
+\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
+We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}
+
 \begin{enumerate}
@@
   Aquifer thickness versus saturated thickness -- The authors speak of
-  ``aquifer saturated thickness'', which is a redundant term. Remember,
-  the definition of an aquifer is a saturated unit -- there is no
-  unsaturated aquifer. Please use ``aquifer thickness'' to avoid
-  propagating this confusing usage into the future.
+  \texttt{aquifer\ saturated\ thickness\textquotesingle{}\textquotesingle{},\ which\ is\ a\ redundant\ term.\ Remember,\ the\ definition\ of\ an\ aquifer\ is\ a\ saturated\ unit\ –\ there\ is\ no\ unsaturated\ aquifer.\ Please\ use}aquifer
+  thickness'\,' to avoid propagating this confusing usage into the
+  future.
 \end{enumerate}
 
-\textcolor{blue}{ferkjnf}
+\textcolor{blue}{We agree that there is some redundant use of words here.
+All the instances of ``saturated thickness'' and ``aquifer saturated thickness'' have been replaced by ``aquifer thickness'' in the revised manuscript.}
 
@@
   manuscript.
+\end{enumerate}
+
+\begin{itemize}
+\item
+  First point:
+  \textcolor{blue}{Quantiles are not recalculated each each.
+If we do so, then the same quantile would represent different aquifer thickness range every year.
+When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk.}
+\item
+  Second point:
+  \textcolor{blue}{Thank you for highlighting this point.
+For the 0\% quantile, we assume a saturated thickness of 9m.
+This minimum of saturated thickness was selected based on REF, which defines 9m as the minimum aquifer thickness needed to provide a functioning water supply.
+We have modified the text on page X lines X-X of the revised manuscript to note these points.}
+\item
+  Third point:
+  \textcolor{blue}{Yes, country and year fixed effects are calculated as part of the regressions.
+However, these variables can be thought of simply as shifters that tell us how much higher average yields are in a single county relative to another county and to account for trends in yields over time.
+As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analysing these variables - and indeed it is standard to not report and analyse fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
+\item
+  Fourth point:
+  \textcolor{blue}{This is an important point, which we thank the reviewer for highlighting.
+The reviewer is correct that unobserved factors, such as aquifer heterogeneity influencing well yield response to changes in saturated thickness, are a likely underlying cause of the uncertainty ranges found in our analyses.
+Variations in aquifer conductivity would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between saturated thickness and drought risk exposure that would contribute to uncertainty in our estimates of observed relationships between drought impacts and aquifer conditions.
+Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought.
+To highlight these points, we have added the following text on page X lines X-X of our revised manuscript:}
+\end{itemize}
+
+\textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer saturated thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger saturated thickness (Butler et al., 2013; Korus and Hensen, 2020).
+While we identify statistically robust impacts of saturated thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.''}
+
+\textcolor{blue}{New references added:}
+
+\begin{itemize}
+\item
+  \textcolor{blue}{Butler, J.
+J., Stotler, R.
+L., Whittemore, D.
+O., and Reboulet, E.
+C.
+(2013).
+Interpretation of water level changes in the High Plains aquifer in western Kansas.
+Groundwater, 51(2), 180-190.}
+\item
+  \textcolor{blue}{Korus, J.
+T., and Hensen, H.
+J.
+(2020).
+Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks.
+Environmental Earth Sciences, 79(16), 382.}
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
 \item
@@
   point in the manuscript.
+\end{enumerate}
+
+\textcolor{blue}{By ``conditioned on drought risk exposure'', we were referring to the fact that Figure 4 displays the relationship between saturated thickness and the share of production area that is irrigated while also accounting for variations in climate across the aquifer.
+I.e.
+that the Figure shows the impact of saturated thickness on share of production irrigated after removing the effects of climate on share of production irrigated.
+To improve clarity, we have modified this phrase to now state ``...after accounting for spatio-temporal variations in drought risk exposure across the aquifer''.}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
 \item
   Equation 2 -- Please define all terms in the equation and provide an
-  explanation of what is meant by ``estimated semi-parametrically''.
+  explanation of what is meant by ``estimated semi-parametrically'\,'.
+\end{enumerate}
+
+\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained in Section 4.2 of our revised manuscript.
+The word semi-parametrically refers to the fact that generalized additive models, such as used in our analysis, estimate the impact of a variable using splines without functional form assumption.
+These are sometimes referred to as semi-parametric regression methods.
+However, we agree that the term is potentially confusing, and therefore have removed the term ``semi-parametrically'' in our revised manuscript.}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
 \item
@@
 
+\textcolor{blue}{This is an interesting idea - thank you for raising it.
+We agree that efforts to improve aquifer management could help to mitigate some of the changes in productivity driven by climate and aquifer conditions in our analysis.
+Greater attention to groundwater conservation, and use of improved scheduling technologies such as soil moisture probes, would be expected to mitigate production losses in years with negative water deficits (i.e.
+wetter years) which we attribute to increased possibility for over-irrigation when well capacities are high (and hence when farmers do not face or observe physical scarcity).
+In years with large positive water deficits (i.e.
+drought years), improved irrigation water management would also be expected to mitigate some of the additional production risk caused by lower saturated thickness.
+However, we would still expect lower saturated thickness to result in elevated exposure to drought risk even in these circumstances, as declining well yields would create a physical limit to water supply which would stop farmers being able to fully meet crop water needs in these years.
+Formally testing these effects is beyond the scope and data of our study, but we agree that these are valuable areas for future work and have added the following text to the discussion of our revised manuscript on page X lines X-X:}
+
+\textcolor{blue}{``Improvements to irrigation scheduling and water management practices, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas (Deines et al., 2019; Glose et al., 2022) and proposed as part of California's SGMA (Sustainable Groundwater Management Act) (Berbel and Esteban, 2019; Lubell et al., 2020), could also play an important role in improving the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+These kinds of interventions would be likely to reduce - but not eliminate - the increases in drought risk caused by declining aquifer saturated thickness, and could bring additional benefits of reducing over-use of water in wetter years through greater awareness of efficiency saving potentials (Foster et al., 2019) and the importance of conservation (Marston et al., 2022) amongst growers.''}
+
+\textcolor{blue}{New references added:}
+
+\begin{itemize}
+\item
+  \textcolor{blue}{Glose, T.
+J., Zipper, S., Hyndman, D.
+W., Kendall, A.
+D., Deines, J.
+M., and Butler, J.
+J.
+(2022).
+Quantifying the impact of lagged hydrological responses on the effectiveness of groundwater conservation.
+Water Resources Research, 58(7), e2022WR032295.}
+\item
+  \textcolor{blue}{Berbel, J., and Esteban, E. (2019). Droughts as a catalyst for water policy change. Analysis of Spain, Australia (MDB), and California. Global Environmental Change, 58, 101969.}
+\item
+  \textcolor{blue}{Lubell, M., Blomquist, W., and Beutler, L.
+(2020).
+Sustainable groundwater management in California: A grand experiment in environmental governance.
+Society and Natural Resources, 33(12), 1447-1467.}
+\item
+  \textcolor{blue}{Marston, L.
+T., Zipper, S., Smith, S.
+M., Allen, J.
+J., Butler, J.
+J., Gautam, S., and David, J.
+Y.
+(2022).
+The importance of fit in groundwater self-governance.
+Environmental Research Letters, 17(11), 111001.}
+\end{itemize}
+
 \hypertarget{minor-points}{%
@@
 
-Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through aquifer
-depletion will..'' -- confusing sentence -- please replace ``through''
-with ``by which''.
+\begin{itemize}
+\tightlist
+\item
+  Section 1, 3rd Paragraph --
+  \texttt{Two\ aquifer\ pathways\ exist\ through\ aquifer\ depletion\ will..\textquotesingle{}\textquotesingle{}\ –\ confusing\ sentence\ –\ please\ replace}through'\,'
+  with ``by which'\,'.
+\end{itemize}
 
-Section 1, 4th Paragraph -- ``we generate empirical evidence'' seems
-awkward to me. How about replacing with ``we present empirical
-evidence''? Next sentence -- ``Our analysis advances on
-previous\ldots{}'' reads better as ``Our analysis extends
-previous\ldots{}'' There are many issues such as these and the previous
-comment. Please address such issues so that the manuscript is an easier
-read.
+\textcolor{blue}{Corrected.}
 
-Section 2.1 title -- I would delete ``marginally'' as it is not needed
-and some will find it confusing.
+\begin{itemize}
+\tightlist
+\item
+  Section 1, 4th Paragraph --
+  \texttt{we\ generate\ empirical\ evidence\textquotesingle{}\textquotesingle{}\ seems\ awkward\ to\ me.\ How\ about\ replacing\ with}we
+  present empirical evidence'`? Next sentence --
+  \texttt{Our\ analysis\ advances\ on\ previous\ \$\textbackslash{}dots\$\textquotesingle{}\textquotesingle{}\ reads\ better\ as}Our
+  analysis extends previous \(\dots\)'' There are many issues such as
+  these and the previous comment. Please address such issues so that the
+  manuscript is an easier read.
+\end{itemize}
 
+\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}
+
+\begin{itemize}
+\tightlist
+\item
+  Section 2.1 title -- I would delete ``marginally'\,' as it is not
+  needed and some will find it confusing.
+\end{itemize}
+
+\textcolor{blue}{Tim, any thoughts on this. I feel like it is still important to say the increase is not much.}
+
+\begin{itemize}
+\tightlist
+\item
   Figure 2 - Why do the plots stop at a thickness of about 12 m when you
-had earlier defined the minimum thickness as 9 m and mention 10 m in the
-text when discussing the relationships?
+  had earlier defined the minimum thickness as 9 m and mention 10 m in
+  the text when discussing the relationships?
+\end{itemize}
 
+\textcolor{blue}{Thank you for catching this. 9m and 10m should have been 12m.}
+
+\begin{itemize}
+\tightlist
+\item
   Section 2.3, First Paragraph -- please clarify that ``average per-area
-production'' is a combination of irrigated and rainfed yields. That
+  production'\,' is a combination of irrigated and rainfed yields. That
   would make the manuscript easier to follow.
+\end{itemize}
 
+\textcolor{blue}{We added ``weighted-average of irrigated and rainfed yields, where weights are production area shares'' so that it is clear what average per-area production means.}
+
+\begin{itemize}
+\tightlist
+\item
   Figure 5 -- I assume that the left plot is corn and the right soybean,
   but please clarify.
+\end{itemize}
 
+\textcolor{blue}{Thanks for noticing this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
   Data and Code Availability -- Are the data from Dr.~Haacker on the
@@
   available according to the principles of open science.
+\end{itemize}
 
-Line before Equation (2) -- ``REF'' should be replaced with the actual
-reference.
+\textcolor{blue}{Yes, we will make the data available online.}
 
-Discussion, First Paragraph -- ``reduction in aquifer depletion'' should
-be ``reduction in aquifer thickness''.
+\begin{itemize}
+\tightlist
+\item
+  Line before Equation (2) -- ``REF'\,' should be replaced with the
+  actual reference.
+\end{itemize}
 
-Section 4.2, 3rd Paragraph -- ``the the impact'' -- remember to spell
-and grammar check prior to submission.
+\textcolor{blue}{Tim?}
 
-Reference in Section 3 -- I doubt that ``Butler Jr'' is the surname,
-please replace with ``Butler''. Suffixes apparently aren't handled well
-by some of the bibliography generation programs.
+\begin{itemize}
+\tightlist
+\item
+  Discussion, First Paragraph --
+  \texttt{reduction\ in\ aquifer\ depletion\textquotesingle{}\textquotesingle{}\ should\ be}reduction
+  in aquifer thickness'\,'.
+\end{itemize}
 
-Please remember to close all parentheses and don't mispair brackets and
-parentheses. Such stuff can raise questions about the care used in the
-analysis.
+\textcolor{blue}{Thank you for catching this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
+  Section 4.2, 3rd Paragraph -- ``the the impact'\,' -- remember to
+  spell and grammar check prior to submission.
+\end{itemize}
+
+\textcolor{blue}{Thank you for catching this. Fixed.}
+
+\begin{itemize}
+\tightlist
+\item
+  Reference in Section 3 -- I doubt that
+  \texttt{Butler\ Jr\textquotesingle{}\textquotesingle{}\ is\ the\ surname,\ please\ replace\ with}Butler'\,'.
+  Suffixes apparently aren't handled well by some of the bibliography
+  generation programs.
+\end{itemize}
+
+\textcolor{blue}{}
+
+\begin{itemize}
+\tightlist
+\item
+  Please remember to close all parentheses and don't mispair brackets
+  and parentheses. Such stuff can raise questions about the care used in
+  the analysis.
+\end{itemize}
 

### 2023-07-29 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.rmd
updates
analysis updated

@@
-\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}
+---
+output:
+  pdf_document:
+    keep_tex: true
+---
+
+\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
+We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}
 
@@
 
-+ There needs to be more description about why aquifer saturation thickness is an appropriate proxy for farmers' ability to irrigate from wells.
+1.
+There needs to be more description about why aquifer saturation thickness is an appropriate proxy for farmers' ability to irrigate from wells.
 This is particularly true since this is a broad interest journal, and the mechanism for the main independent variable should be very clear to all readers.
 
-\textcolor{blue}{Tim?}
+\textcolor{blue}{Saturated thickness affects farmers ability to meet the water requirements of their crops due to its impacts on the pumping yields (or capacities) of groundwater irrigation wells.
+As the saturated thickness of an aquifer declines, the rate at which water can flow from the aquifer to the pumping well is reduced proportionally to the aquifer's transmissivity (equal to the hydraulic conductivity of the aquifer geology multiplied by the saturated thickness of that aquifer geology).
+For small relative changes in saturated thickness (e.g.
+for initial depletion of an aquifer with large initial saturated thickness), the change in well yield is small and unlikely to have substantial negative impacts on drought risk exposure as a farmer could simply run their pump for slightly longer to compensate for the small reduction in instantaneous well yield.
+However, for larger relative changes in saturated thickness (e.g.
+due to significant aquifer depletion and/or thinner aquifer bodies), the magnitude of well yield reductions will be much larger.
+These reductions in water supply capacity cannot be mitigated by changes to irrigation scheduling or management practices.
+The farmer will often be unable to keep up with crop water demands during the peak of the growing season in the absence of rainfall, even if running their system continually.
+This effect can also typically not be mitigated by drilling additional or deeper wells, due to a combination of both regulatory moratoriums on well drilling - which are common across the HPA - and high costs of deep well construction.
+To more clearly highlight for non-specialist readers these connections between saturated thickness, well yields, and drought resilience of agriculture, we have added the following text on page X lines X-X of our revised manuscript.
+We note that we are limited in elaborating further on background information given journal word constraints, and hence provide key supporting studies in citations for readers interested in deepening understanding of these dynamics.}
 
-+ Though I am not trained as a hydrologist, my understanding is that aquifer saturation thickness is a good measure of general water availability in an aquifer but not necessarily how much water is available at a specific location or point.
+\textcolor{blue}{``...depletion also reduces the saturated thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation (Konikow and Kendy, 2005; Foster et al., 2015; Hrozencik et al., 2017).
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer saturated thickness is depleted (Hecox et al., 2002; Korus and Hensen, 2020) - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits (Foster et al., 2015a; Rouhi-Rad et al., 2020).''
+}
+
+\textcolor{blue}{Additional references added:}
+
+  - \textcolor{blue}{Hecox, G.R., Macfarlane, P.A., and Wilson B.B.
+(2002).
+Calculation of yield for High Plains wells: relationship between saturated thickness and well yield.
+Open file report 2002–25C, Kansas Geological Survey: Lawrence, Kansas, USA.}
+
+  - \textcolor{blue}{Korus, J.
+T., and Hensen, H.
+J.
+(2020).
+Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks.
+Environmental Earth Sciences, 79(16), 382.}
+
+  - \textcolor{blue}{Rouhi-Rad, M., Araya, A., and Zambreski, Z. T. (2020). Downside risk of aquifer depletion. Irrigation Science, 38, 577-591.}
+
+2.
+Though I am not trained as a hydrologist, my understanding is that aquifer saturation thickness is a good measure of general water availability in an aquifer but not necessarily how much water is available at a specific location or point.
 If this is the case, the authors should make clear why this is an appropriate measure to use in their study and potential limitations of using this approach.
@@
 
-+ It is unclear to me why county and time fixed effects were not used in every regression.
+3.
+It is unclear to me why county and time fixed effects were not used in every regression.
 In particular, it does not look like they were included in the regression that modeled irrigated area as a function of saturated thickness and water deficit.
@@
 
-\textcolor{blue}{Thank you for this comment, which we believe led to a slighly better model and acknowledgin important limitations of our methods.
+\textcolor{blue}{Thank you for this comment, which we believe led to a slighly better model and acknowledging important limitations of our methods.
 County and year fixed effects were included in equation (1), but only year fixed effects were included in equation (2).
@@
 
-+ There are minor typos (e.g., missing reference listed as REF) that should be corrected prior to resubmission.
+4. There are minor typos (e.g., missing reference listed as REF) that should be corrected prior to resubmission.
 

### 2023-07-29 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.tex
analysis updated

@@
 
-\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}
+\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
+We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
 \tightlist
@@
   very clear to all readers.
+\end{enumerate}
+
+\textcolor{blue}{Saturated thickness affects farmers ability to meet the water requirements of their crops due to its impacts on the pumping yields (or capacities) of groundwater irrigation wells.
+As the saturated thickness of an aquifer declines, the rate at which water can flow from the aquifer to the pumping well is reduced proportionally to the aquifer's transmissivity (equal to the hydraulic conductivity of the aquifer geology multiplied by the saturated thickness of that aquifer geology).
+For small relative changes in saturated thickness (e.g.
+for initial depletion of an aquifer with large initial saturated thickness), the change in well yield is small and unlikely to have substantial negative impacts on drought risk exposure as a farmer could simply run their pump for slightly longer to compensate for the small reduction in instantaneous well yield.
+However, for larger relative changes in saturated thickness (e.g.
+due to significant aquifer depletion and/or thinner aquifer bodies), the magnitude of well yield reductions will be much larger.
+These reductions in water supply capacity cannot be mitigated by changes to irrigation scheduling or management practices.
+The farmer will often be unable to keep up with crop water demands during the peak of the growing season in the absence of rainfall, even if running their system continually.
+This effect can also typically not be mitigated by drilling additional or deeper wells, due to a combination of both regulatory moratoriums on well drilling - which are common across the HPA - and high costs of deep well construction.
+To more clearly highlight for non-specialist readers these connections between saturated thickness, well yields, and drought resilience of agriculture, we have added the following text on page X lines X-X of our revised manuscript.
+We note that we are limited in elaborating further on background information given journal word constraints, and hence provide key supporting studies in citations for readers interested in deepening understanding of these dynamics.}
+
+\textcolor{blue}{``...depletion also reduces the saturated thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation (Konikow and Kendy, 2005; Foster et al., 2015; Hrozencik et al., 2017).
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer saturated thickness is depleted (Hecox et al., 2002; Korus and Hensen, 2020) - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits (Foster et al., 2015a; Rouhi-Rad et al., 2020).''
+}
+
+\textcolor{blue}{Additional references added:}
+
+\begin{itemize}
+\item
+  \textcolor{blue}{Hecox, G.R., Macfarlane, P.A., and Wilson B.B.
+(2002).
+Calculation of yield for High Plains wells: relationship between saturated thickness and well yield.
+Open file report 2002–25C, Kansas Geological Survey: Lawrence, Kansas, USA.}
+\item
+  \textcolor{blue}{Korus, J.
+T., and Hensen, H.
+J.
+(2020).
+Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks.
+Environmental Earth Sciences, 79(16), 382.}
+\item
+  \textcolor{blue}{Rouhi-Rad, M., Araya, A., and Zambreski, Z. T. (2020). Downside risk of aquifer depletion. Irrigation Science, 38, 577-591.}
 \end{itemize}
 
-\textcolor{blue}{Tim?}
-
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
 \tightlist
@@
   their study and potential limitations of using this approach.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
 \tightlist
@@
   with inference.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
 \tightlist
@@
   should be corrected prior to resubmission.
-\end{itemize}
+\end{enumerate}
 

### 2023-07-29 Codes/1_2_test_dif_significance_ir_yield.rmd
analysis updated

@@
 
+```{r}
+main_analysis <- readRDS("Shared/Results/main_analysis.rds")
+
+semi_res <- main_analysis$yield_semi_res[[1]]
+sat_text_data <- main_analysis$sat_text_data[[1]]
+```
+
 # Define the function to test
@@
 ```{r}
-test_dif_in_yield <- function(crop, balance, base_q, comp_q, results) {
+main_analysis <-
+  main_analysis %>%
+  dplyr::mutate(yield_dif_test = list(
+    lapply(
+      2:3,
+      \(x) {
+        test_dif_in_yield(
+          balance_ls = seq(min(data$balance), max(data$balance), length = 50),
+          base_q = 1,
+          comp_q = x,
+          yield_semi_res,
+          sat_text_data,
+          sc_base
+        ) %>%
+        .[, comp_q := x]
+      }
+    ) %>%
+    rbindlist()
+  ))
+
+test_dif_in_yield <- function(balance_ls, base_q, comp_q, yield_semi_res, sat_text_data, sc_base) {
   #++++++++++++++++++++++++++++++++++++
@@
   #++++++++++++++++++++++++++++++++++++
-  # crop <- "Corn"
+  # crop_w <- "corn"
   # balance <- 1000
@@
 
+  # balance_ls <- c(200, 300, 400)
+
   #++++++++++++++++++++++++++++++++++++
@@
   #++++++++++++++++++++++++++++++++++++
+  sc_dry_0 <- paste0(sc_base, "_1")
+  sc_dry_ir <- paste0(sc_base, "_0")
+
   q_data <-
-    data.table(
-      quater = c(1, 2, 3),
-      coef_name = list(
-        c("s_balance_sat_cat_39_4_82__1", "s_balance_sat_cat_39_4_82__2", "sat_cat::[39.4,82]"),
-        c("s_balance_sat_cat_82_194__1", "s_balance_sat_cat_82_194__2", "sat_cat::(82,194]"),
-        c("s_balance_sat_cat_194_708__1", "s_balance_sat_cat_194_708__2", "sat_cat::(194,708]")
-      ),
-      quarter_names = c("[39.4,82]", "(82,194]", "(194,708]")
+    sat_text_data %>%
+    .[, coef_name := paste0("s_balance_sat_cat", sat_cat)] %>%
+    .[, coef_name := gsub("\\[", "_", coef_name)] %>%
+    .[, coef_name := gsub("\\]", "", coef_name)] %>%
+    .[, coef_name := gsub("\\(", "_", coef_name)] %>%
+    .[, coef_name := gsub("\\,", "_", coef_name)] %>%
+    .[, coef_name := gsub("\\.", "_", coef_name)] %>%
+    rowwise() %>%
+    dplyr::mutate(coef_name = list(
+      c(
+        paste0(coef_name, "__", c(1, 2)),
+        paste0("sat_cat::", sat_cat)
       )
+    )) %>%
+    data.table()
 
-  crop_w <- crop
-  results_w <- results[crop == crop_w, ]
-  fe_res <- results_w[, semi_res][[1]]$fe_res
+  fe_res <- yield_semi_res$fe_res
   vcov <- vcov(fe_res)
-  gam_res <- results_w[, semi_res][[1]]$gam_res
-  data_w <- results_w[, data][[1]]
-  balance_w <- balance
+  gam_res <- yield_semi_res$gam_res
 
@@
     CJ(
-      balance = balance_w,
-      sat_cat = unique(data_w$sat_cat)
+      balance = balance_ls,
+      sat_cat = unique(q_data$sat_cat)
     ) %>%
@@
     # .[, sc_code := data_w$sc_code[2]] %>%
-    .[, sc_code := data_w[sat_cat != "dryland"]$sc_code[1]] %>%
-    .[, sc_dry := paste0(data_w$sc_dry[1])] %>%
+    .[, sc_dry := sc_dry_ir] %>%
     # === can be any year (need to shift yield for "average" year) ===#
@@
     # === fake yield data (necessary to take advantage of predict.gam) ===#
-    .[, yield := 0] %>%
-    .[, type := "pred"] %>%
-    .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]
+    .[, yield := 0]
 
@@
   #++++++++++++++++++++++++++++++++++++
-  #+ point estimate
+  #+ Coefficients
   #++++++++++++++++++++++++++++++++++++
@@
   #--- coefficients ---#
-  base_coef_names <- q_data[quater == base_q, coef_name][[1]]
-  comp_coef_names <- q_data[quater == comp_q, coef_name][[1]]
+  base_coef_names <- q_data[sat_rank == base_q, coef_name][[1]]
+  comp_coef_names <- q_data[sat_rank == comp_q, coef_name][[1]]
 
@@
 
+  #++++++++++++++++++++++++++++++++++++
+  #+ loop over balance
+  #++++++++++++++++++++++++++++++++++++
+  lapply(
+    1:length(balance_ls),
+    \(x) {
+      balance_w <- balance_ls[x]
+      #---------------------
+      #- point estimate
+      #---------------------
       base_x <-
         c(
-      yhat_data[balance == balance_w & sat_cat == q_data[quater == base_q, quarter_names], ..base_x_names] %>% unlist(),
+          yhat_data[balance == balance_w & sat_cat == q_data[sat_rank == base_q, sat_cat], ..base_x_names] %>% unlist(),
           1
@@
         c(
-      yhat_data[balance == balance_w & sat_cat == q_data[quater == comp_q, quarter_names], ..comp_x_names] %>% unlist(),
+          yhat_data[balance == balance_w & sat_cat == q_data[sat_rank == comp_q, sat_cat], ..comp_x_names] %>% unlist(),
           1
@@
         data.table(
-      crop = crop_w,
           balance = balance_w,
@@
     }
+  ) %>%
+    rbindlist()
+}
+
 ```

### 2023-07-29 Codes/D1_Analysis.rmd
analysis updated

--- 2023-07-29 Codes/1_Analysis.rmd
@@
   data.table() %>%
-  setnames("crop_type", "crop") %>%
-  #--- further necessary transformation ---#
-  dplyr::mutate(data = purrr::map(
-    reg_data_y,
-    \(reg_data_y) {
-      copy(reg_data_y) %>%
-        #--- bu/acre to tonne/ha ---#
-        .[, yield := yield * 12.553 / 200] %>%
-        .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
-        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
-        .[, state_year := paste0(state_name, "_", year)] %>%
-        .[, balance := -balance]
-    }
-  ))
+  setnames("crop_type", "crop")
 
@@
 ```{r}
-set.seed(283397)
 
 #--- global parameters ---#
-sc_base <- "31011"
-sc_dry_base <- "31011_0"
 
@@
 ```{r}
-#++++++++++++++++++++++++++++++++++++
-#+ Bootstrap
-#++++++++++++++++++++++++++++++++++++
+set.seed(837943)
+
+boot_results_added$yield_pred_data_boot[[1]] %>% 
+  rbindlist() %>%
+  .[, .(
+      # y_hat_boot = median(y_hat, na.rm = TRUE),
+      y_hat_se = sd(y_hat, na.rm = TRUE),
+      # dif_y_hat_boot = mean(dif_y_hat, na.rm = TRUE),
+      dif_y_hat_se = sd(dif_y_hat, na.rm = TRUE)
+    ),
+    by = .(balance, sat, sat_cat)
+    ]
+
 boot_results_added <-
-  main_analysis %>%
+  main_analysis[1, ]  %>%
   #--- bootsrap data ---#
@@
         print(x)
-        yield_analysis(boot_data[[x]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
+        yield_analysis_manual(boot_data[[x]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
       },
@@
     )
+  ))
+
+```
+```{r}
+#++++++++++++++++++++++++++++++++++++
+#+ Bootstrap
+#++++++++++++++++++++++++++++++++++++
+set.seed(837943)
+
+balance_seq <- main_analysis$balance_seq[[1]]
+  balance_seq <- main_analysis$balance_seq[[1]]
+  sat_seq_eval <- main_analysis$sat_seq_eval[[1]]
+  sat_breaks <- main_analysis$sat_breaks[[1]]
+  sat_text_data <- main_analysis$sat_text_data[[1]]
+
+temp <- lapply(1:500, get_coef) %>% rbindlist
+
+main_analysis$data[[1]][sc_dry == "56031_1", ]
+
+get_coef <- function(i) {
+
+yield_data <- main_analysis$boot_data[[1]][[i]]
+semi_res_corn <-
+    semi_ols(
+      semi_formula = semi_formula,
+      c_formula = paste0("i(sat_cat, ref = 'dryland') + i(year, ref = 2009) + i(sc_dry, ref = '", sc_dry_base, "')"),
+      # fe = "",
+      cluster = "sc_dry",
+      data = yield_data
+    )
+
+y_res_int <- semi_res_corn$fe_res
+
+return_data[, mean(y_hat), by = sat_cat]
+yield_data[, mean(yield), by = sat_cat]
+
+return(
+tidy(y_res_int) %>%
+data.table() %>%
+.[term %in% c("sat_cat::[9.01,23.1]", "sat_cat::(23.1,56.8]", "sat_cat::(56.8,216]"), estimate]
+)
+
+}
+
+main_analysis <-
+  main_analysis %>%
+  #--- bootsrap data ---#
+  dplyr::mutate(boot_data = list(
+    replicate(1000, boot(data), simplify = FALSE)
   )) %>%
-  #--- share analysis ---#
-  dplyr::mutate(share_pred_data_boot = list(
-    parallel::mclapply(
-      1:length(share_boot_data),
-      \(x) {
-        print(x)
-        share_analysis(share_boot_data[[x]], sat_seq, sandtotal_med, silttotal_med, awc_med)
-      },
-      mc.cores = parallel::detectCores() - 2
-    )
-  )) %>%
-  #--- average yield analysis ---#
-  dplyr::mutate(avg_yield_pred_data_boot = list(
-    parallel::mclapply(
+  #--- get share data ---#
+  dplyr::mutate(share_boot_data = list(
+    lapply(
       1:length(boot_data),
       \(x) {
-        get_average_yield(yield_pred_data_boot[[x]], share_pred_data_boot[[x]])
-      },
-      mc.cores = parallel::detectCores() - 2
+        boot_data[[x]] %>%
+          .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
+          .[, balance_avg := mean(balance), by = sc_code]
+      }
     )
-  )) %>%
-  dplyr::mutate(
-    yield_pred_data_boot = list(rbindlist(yield_pred_data_boot)),
-    share_pred_data_boot = list(rbindlist(share_pred_data_boot)),
-    avg_yield_pred_data_boot = list(rbindlist(avg_yield_pred_data_boot))
-  ) %>%
-  dplyr::select(
-    crop, data, sat_seq, sat_text_data,
-    yield_pred_data, yield_pred_data_boot,
-    share_pred_data, share_pred_data_boot,
-    avg_yield_pred_data, avg_yield_pred_data_boot,
-  )
-
-saveRDS(boot_results_added, "Shared/Results/boot_results_added.rds")
+  ))
 ```
 
+```{r}
+boot_data <- boot_results_added$boot_data[[1]]
+
+balance_seq <- boot_results_added$balance_seq[[1]]
+sat_seq_eval <- boot_results_added$sat_seq_eval[[1]]
+sat_breaks <- boot_results_added$sat_breaks[[1]]
+sat_text_data <- boot_results_added$sat_text_data[[1]]
+
+temp <- yield_analysis(boot_data[[200]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
+
+ggplot(temp[sat == 50, ]) +
+  geom_point(aes(y = y_hat, x = balance)) +
+  ylim(0, 15)
+
+all_results$yield_pred_data_boot[[1]][sat == 50, sd(y_hat), by = balance]
+
+all_results$yield_pred_data_boot[[1]][sat == 10, ]
+
+all_results$yield_pred_data_boot[[1]][sat == 10, ] %>%
+ggplot(.) +
+  geom_point(aes(y = y_hat, x = balance))
+
+boot_corn <- 
+  boot_results_added$yield_pred_data_boot[[1]] %>%
+  rbindlist(idcol = "boot") %>%
+  .[, mean(y_hat), by = .(boot, sat)]
+
+boot_ids <- boot_corn[V1 < 11 & sat == 10, boot]
+
+boot_results_added$boot_data[[1]][[11]] %>%
+  .[, .N, by = sat_cat]
+
+ggplot(data = boot_results_added$boot_data[[1]][[12]]) +
+  geom_point(aes(y = yield, x = balance)) +
+  facet_grid(sat_cat ~ .)
+
+boot_corn[sat %in% c(10, 50), ] %>%
+  dcast(boot ~ sat, value.var = "V1") %>%
+  ggplot() + 
+  geom_point(aes(y = `50`, x = `10`))
+
+  ggplot(boot_corn) +
+    geom_histogram(aes(x = V1)) +
+    facet_grid(sat ~ .)
+
+Check if the average yield and predicted average yield are similar.
+
+```
 ```{r}
@@
 boot_results_added <- readRDS("Shared/Results/boot_results_added.rds")
+
+# all_results[, data][[1]] %>%
+#   ggplot() +  
+#   geom_point(aes(y = yield, x = balance)) +
+#   facet_grid(sat_cat ~ .)
 

### 2023-07-29 Writing/manuscript.Rmd
analysis updated

@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs [@mieno2017price; @bhattarai2021impact].
-Second, depletion also reduces the pumping capacity of wells used to extract groundwater, which may create a physical constraint to the volume of irrigation that can be applied to a field [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous].
+Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous].
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted [@hecox2002calculation; @korus2020depletion] - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
 The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-Figure \@ref(fig:ir-share) shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean.
+```{r}
+all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
+
+corn_share <- all_results[crop == "corn", share_pred_data][[1]]
+
+share_high <- 
+  corn_share[which.min(abs(sat - 150)), ir_share_hat] %>%
+  round(digits = 2)
+
+share_low <- 
+  corn_share[which.min(abs(sat - 10)), ir_share_hat] %>%
+  round(digits = 2)
+```
+
+Figure \@ref(fig:ir-share) shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in drought risk exposure across the aquifer.
 Figure \@ref(fig:ir-share) demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m.
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around `r share_high` (`r share_high * 100`\%) for an average county-level aquifer thickness of 150m to `r share_low` (`r share_low * 100`\%) for an average county-level aquifer thickness of 12m.
 Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn [@zipper2016drought; @ruess2022irrigation].
@@
 
-## Aquifer depletion increases overall drought-related production losses by up to $25\%$
+## Aquifer depletion increases overall drought-related production losses by up to $25\%$j
 
@@
   data.table() %>%
-  .[(abs(balance - 950)) < 1, by = crop]
+  .[crop == "corn", ] %>%
+  .[(abs(balance - 950)) < 5, ]
 ```
+
 The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
-Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, 106.1, and 153.0  meters^[Confidence intervals were not presented to avoid over-crowding the figures.
+Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 10, 50, 90, and 130 meters^[Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \@ref(reg-conf) for estimated average yield response curves with 95 percent confidence interval.].
@@
 
-Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 59.1m, 106.1m, and 153.0m aquifer thickness levels relative to 12.2m with the 95 percent confidence interval for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 59.1m and 106.1m aquifer thickness, respectively).
+Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 50m, 90m, and 130m aquifer thickness levels relative to 10m with the 95 percent confidence interval for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 50m and 90m aquifer thickness, respectively).
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
@@
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
-Further, notice that the decline in average per-area production at a seasonal water deficit of 950mm is not linear.
-Going from 153.0m aquifer thickness to 106.1m (decline of 46.9m) would result in `r round(res_at_950[sat == 502 & crop == "corn", avg_yield] - res_at_950[sat == 348 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+
+Further, notice that the impact of aquifer thickness on the decline in average per-area production is highly non-linear at high seasonal water deficits.
+Going from 130m aquifer thickness to 90m (decline of 40m) would result in merely `r round(res_at_950[sat == 130 & crop == "corn", avg_yield] - res_at_950[sat == 90 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
-Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in `r round(res_at_950[sat == 348 & crop == "corn", avg_yield] - res_at_950[sat == 194 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
-However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in `r round(res_at_950[sat == 194 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+Going from 90m aquifer thickness to 50m (decline of 40m) would result in `r round(res_at_950[sat == 90 & crop == "corn", avg_yield] - res_at_950[sat == 50 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+However, going from 50m aquifer thickness to 10m (decline of 40m) would result in `r round(res_at_950[sat == 50 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
 That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
@@
 Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring [@edreira2020combining; @deines2021million], would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
+Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness [@butler2013interpretation; @korus2020depletion].
+While we identify statistically robust impacts of aquifer thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
 This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
@@
 Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies [@bhattarai2021impact; @deines2019quantifying; @manning2017producer].
+Improvements to irrigation scheduling and water management practices, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas [@deines2019quantifying; @glose2022quantifying] and proposed as part of California's SGMA (Sustainable Groundwater Management Act) [@berbel2019droughts; @lubell2020sustainable], could also play an important role in improving the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+These kinds of interventions would be likely to reduce - but not eliminate - the increases in drought risk caused by declining aquifer thickness, and could bring additional benefits of reducing over-use of water in wetter years through greater awareness of efficiency saving potentials [@foster2019assessing] and the importance of conservation [@marston2022] amongst growers.
 Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
@@
 To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
-This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).
+This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work [@foster2015well; @foster2015analysis; @hrozencik2017heterogeneous].
 The model is given in Equation \@ref(eq:eqn2) below:

### 2023-07-29 Writing/manuscript.tex
analysis updated

@@
 \usepackage{float}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \ifLuaTeX
@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
-Second, depletion also reduces the pumping capacity of wells used to extract groundwater, which may create a physical constraint to the volume of irrigation that can be applied to a field \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
+Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
 The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,]{Figures/g_yield_intensive}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_ir_yield_dif}
+{\centering \includegraphics[width=6in,]{Figures/g_ir_yield_dif}
 
@@
 
-Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean.
+Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in drought risk exposure across the aquifer.
 Figure \ref{fig:ir-share} demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m.
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 150m to 0.63 (63\%) for an average county-level aquifer thickness of 12m.
 Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn \citep{zipper2016drought, ruess2022irrigation}.
@@
 
-{\centering \includegraphics[width=5in,height=500px,]{Figures/g_share}
+{\centering \includegraphics[width=5in,]{Figures/g_share}
 
@@
 
-\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}{%
-\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
+\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25j}{%
+\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)j}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%j}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25j}}
 
 The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, and 106.1 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
+Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 10, 50, 90, and 130 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
 Please see Appendix \ref{reg-conf} for estimated average yield response curves with 95 percent confidence interval.}.
-While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronouced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
+While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
 
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 59.1m and 106.1m aquifer thickness leves relative to 12.2m with the 95 percent confidence interval for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 300mm).
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 50m, 90m, and 130m aquifer thickness levels relative to 10m with the 95 percent confidence interval for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 50m and 90m aquifer thickness, respectively).
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
@@
 For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
-Further, notice that the decline in average per-area production at a sesonal water deicit of 950mm is not linear.
-Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in 1.01 tonne/ha of production loss for corn.
-However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in 2.12 tonne/ha of production loss for corn.
-That is the same degree of decline is much more damaging when the decline is from a low aquifer thickness level thant a higher aquifer thickness level.
+
+Further, notice that the impact of aquifer thickness on the decline in average per-area production is highly non-linear at high seasonal water deficits.
+Going from 130m aquifer thickness to 90m (decline of 40m) would result in merely 0.1 tonne/ha of production loss for corn.
+This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
+Going from 90m aquifer thickness to 50m (decline of 40m) would result in 0.42 tonne/ha of production loss for corn.
+However, going from 50m aquifer thickness to 10m (decline of 40m) would result in 2.18 tonne/ha of production loss for corn.
+That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
 This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \ref{fig:ir-share}.
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_total_impact}
+{\centering \includegraphics[width=6in,]{Figures/g_total_impact}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,]{Figures/g_avg_yield_dif}
 
@@
 Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining, deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
+Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness \citep{butler2013interpretation, korus2020depletion}.
+While we identify statistically robust impacts of aquifer thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
 This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
@@
 Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact, deines2019quantifying, manning2017producer}.
+Improvements to irrigation scheduling and water management practices, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas \citep{deines2019quantifying, glose2022quantifying} and proposed as part of California's SGMA (Sustainable Groundwater Management Act) \citep{berbel2019droughts, lubell2020sustainable}, could also play an important role in improving the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+These kinds of interventions would be likely to reduce - but not eliminate - the increases in drought risk caused by declining aquifer thickness, and could bring additional benefits of reducing over-use of water in wetter years through greater awareness of efficiency saving potentials \citep{foster2019assessing} and the importance of conservation \citep{marston2022} amongst growers.
 Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
@@
 
-{\centering \includegraphics[width=750px,height=6in,]{Figures/g_y_wd_all}
+{\centering \includegraphics[height=6in,]{Figures/g_y_wd_all}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_map}
+{\centering \includegraphics[width=6in,]{Figures/g_map}
 
@@
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalizez additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
-This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
+This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work \citep{foster2015well, foster2015analysis, hrozencik2017heterogeneous}.
 The model is given in Equation \eqref{eq:eqn2} below:
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_corn}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Shared/Results/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_soy}
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Shared/Results/g_avg_yield}
+{\centering \includegraphics[width=6.5in,]{Figures/g_avg_yield}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_variation}
+{\centering \includegraphics[width=6in,]{Figures/g_variation}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_share_comp}
+{\centering \includegraphics[width=6in,]{Figures/g_share_comp}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_basis}
+{\centering \includegraphics[width=6in,]{Figures/g_basis}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_il_y}
+{\centering \includegraphics[width=6in,]{Figures/g_il_y}
 

### 2023-07-29 Codes/Functions/retired_functions.R

@@ -179,7 +179,7 @@ semi_ols_fe <- function(semi_formula, c_formula, fe, cluster, data) {
   # cor(reg_data$model_X)
 }
 
-yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data, sc_base, bootstrap = FALSE) {
+yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data, sc_base) {
   #++++++++++++++++++++++++++++++++++++
   #+ For Debugging
   #++++++++++++++++++++++++++++++++++++
@@ -205,7 +205,7 @@ yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sa
     formula(yield ~ s(balance, by = sat_cat, k = 3, m = 2))
 
   #--- semi-parametric FE wih fixest ---#
-  semi_res <-
+  semi_res_corn <-
     semi_ols(
       semi_formula = semi_formula,
       c_formula = paste0("i(sat_cat, ref = 'dryland') + i(year, ref = 2009) + i(sc_dry)"),
@@ -215,12 +215,12 @@ yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sa
     )
 
   #--- main regression results ---#
-  y_res_int <- semi_res$fe_res
+  y_res_int <- semi_res_corn$fe_res
 
   #---------------------
   #- Prediction
   #---------------------
-  gam_y_res <- semi_res$gam_res
+  gam_y_res <- semi_res_corn$gam_res
 
   #--- raw data for prediction ---#
   data_for_pred_corn <-
@@ -259,7 +259,7 @@ yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sa
 
   yield_hat_data[, `:=`(
     y_hat = y_hat_with_se$fit,
-    y_hat_se_modeled = y_hat_with_se$se.fit
+    y_hat_se = y_hat_with_se$se.fit
   )]
 
   return_data <-
@@ -269,16 +269,7 @@ yield_analysis <- function(yield_data, balance_seq, sat_seq_eval, sat_breaks, sa
 
   return_data[, dif_y_hat := y_hat - q1_yield]
 
-  if (bootstrap == TRUE) {
-    return(
-     yield_pred_data = return_data[, .(balance, sat, sat_cat, sat_cat_text, sat_rank, y_hat, y_hat_se_modeled, dif_y_hat)]
-    )
[…]

### 2023-07-31 Codes/test/1_Analysis.rmd
made revisions (semi-finalized)

---
title: "Regression Analysis and Prediction"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare data

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

# reg_data[, data][[2]] %>% .[, .N, by = sat_cat]

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    sat_thld_m = 9, # saturated thickness has to be at least 12 meters
    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
    balance_thld = c(-1200, 200),
    sat_cat_num = 3,
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop")

# saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))

# ggplot(reg_data$data[[1]]) +
#   geom_point(aes(y = acres_ratio, x = sat)) +
#   geom_smooth(aes(y = acres_ratio, x = sat))
```

# Regression and prediciton

## Yield

```{r}
main_analysis <-
  reg_data %>%
  rowwise() %>%
  dplyr::mutate(sc_base = "31011") %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(balance_seq = list(
    seq(
      min(data$balance),
      max(data$balance),
      length = 100
    )
  )) %>%
  dplyr::mutate(sat_seq_eval = list(
    c(10, 50, 90, 130)
  )) %>%
  dplyr::mutate(share_reg_data = list(
    data %>%
      .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
      .[, balance_avg := mean(balance), by = sc_code]
  )) %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(sat_seq = list(
    share_reg_data[, seq(min(sat),
      max(sat),
      length = 100
    )] %>%
      c(., sat_seq_eval)
  )) %>%
  #--- values of the controls to be used in prediction ---#
  dplyr::mutate(
    sandtotal_med = median(share_reg_data$sandtotal_r),
    silttotal_med = median(share_reg_data$silttotal_r),
    awc_med = median(share_reg_data$awc_r)
  ) %>%
  #++++++++++++++++++++++++++++++++++++
  #+ Main analysis
  #++++++++++++++++++++++++++++++++++++
  dplyr::mutate(yield_analysis_results = list(
    yield_analysis(data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data, sc_base)
  )) %>%
  dplyr::mutate(
    yield_pred_data = list(yield_analysis_results$yield_pred_data),
    yield_semi_res = list(yield_analysis_results$semi_res)
  ) %>%
  dplyr::mutate(yield_dif_test = list(
    lapply(
      2:3,
      \(x) {
        test_dif_in_yield(
          balance_ls = seq(min(data$balance), max(data$balance), length = 50),
          base_q = 1,
          comp_q = x,
          yield_semi_res,
          sat_text_data,
          sc_base
        ) %>%
          .[, comp_q := x]
      }
    ) %>%
      rbindlist()
  )) %>%
  dplyr::mutate(share_pred_data = list(
    share_analysis(share_reg_data, sat_seq, sandtotal_med, silttotal_med, awc_med)
  )) %>%
  dplyr::mutate(avg_yield_pred_data = list(
    get_average_yield(yield_pred_data, share_pred_data)
  ))

saveRDS(main_analysis, "Shared/Results/main_analysis.rds")
```

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Bootstrap
#++++++++++++++++++++++++++++++++++++
set.seed(383954)

all_results <-
  main_analysis %>%
  #--- bootsrap data ---#
  dplyr::mutate(boot_results = list(
    parallel::mclapply(
      1:1000,
      \(x) {
        print(x)

        #--- bootstrap data ---#
        boot_data <- boot(data, sc_base)

        #--- yield analysis ---#
        yield_pred_data <-
          yield_analysis(
            boot_data,
            balance_seq,
            sat_seq_eval,
            sat_breaks,
            sat_text_data,
            sc_base,
            boot = TRUE
          )

        #--- share analysis ---#
        share_boot_data <-
          boot_data %>%
          .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
          .[, balance_avg := mean(balance), by = sc_code]

        share_pred_data <-
          share_analysis(
            ir_share_data = share_boot_data,
            sat_seq,
            sandtotal_med,
            silttotal_med,
            awc_med
          )
        #--- average yield ---#
        avg_yield_pred_data <- get_average_yield(yield_pred_data, share_pred_data)

        return(list(share_pred_data = share_pred_data, avg_yield_pred_data = avg_yield_pred_data))
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  dplyr::mutate(
    share_pred_data_boot = list(purrr::map(boot_results, "share_pred_data") %>% rbindlist()),
    avg_yield_pred_data_boot = list(purrr::map(boot_results, "avg_yield_pred_data") %>% rbindlist())
  ) %>%
  dplyr::mutate(share_sum = list(
    share_pred_data_boot[, .(
      ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE),
      dif_ir_share_hat_se = sd(dif_ir_share_hat, na.rm = TRUE)
    ),
    by = .(sat)
    ] %>%
      .[share_pred_data, on = "sat"] %>%
      .[, `:=`(
        lower = ir_share_hat - 1.96 * ir_share_hat_se,
        upper = ir_share_hat + 1.96 * ir_share_hat_se,
        dif_lower = dif_ir_share_hat - 1.96 * dif_ir_share_hat_se,
        dif_upper = dif_ir_share_hat + 1.96 * dif_ir_share_hat_se
      )]
  )) %>%
  dplyr::mutate(avg_yield_sum = list(
    avg_yield_pred_data_boot[, .(
      avg_yield_se = sd(avg_yield, na.rm = TRUE),
      dif_avg_yield_se = sd(dif_avg_yield, na.rm = TRUE)
    ),
    by = .(balance, sat)
    ] %>%
      .[avg_yield_pred_data, on = c("balance", "sat")] %>%
      .[, `:=`(
        upper = avg_yield + 1.96 * avg_yield_se,
        lower = avg_yield - 1.96 * avg_yield_se,
        dif_upper = dif_avg_yield + 1.96 * dif_avg_yield_se,
        dif_lower = dif_avg_yield - 1.96 * dif_avg_yield_se
      )]
  )) %>%
  data.table() %>%
  dplyr::select(
    crop, data, sat_seq, sat_text_data,
    yield_pred_data,
    yield_dif_test,
    share_pred_data,
    share_sum,
    avg_yield_pred_data,
    avg_yield_pred_data_boot,
    avg_yield_sum
  )

saveRDS(all_results, "Shared/Results/all_results.rds")
```

### 2023-07-31 Codes/test/functions.R

@@ -1,0 +1,816 @@
+prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
+  reg_data <-
+    data.table(
+      crop_type = c("corn", "soy"),
+      states = state_ls
+    ) %>%
+    rowwise() %>%
+    mutate(data = list(
+      data %>%
+        .[crop == crop_type, ] %>%
+        .[state %in% states, ] %>%
+        #--- yield ---#
+        .[, yield := yield * 12.553 / 200] %>%
+        #--- balance ---#
+        .[balance >= balance_thld[1], ] %>%
+        .[balance <= balance_thld[2], ] %>%
+        .[, balance := -balance] %>%
+        #--- unit conversion ---#
+        .[, sat := measurements::conv_unit(sat, "ft", "m")] %>%
+        #--- filter out counties with too low saturated thickness ---#
+        .[sat >= sat_thld_m | sat == 0, ] %>%
+        #--- share of irrigated/dryland acres ---#
+        .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
+        .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ] %>%
+        .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
+        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
+        .[, state_year := paste0(state_name, "_", year)] %>%
+        .[, num_obs := .N, by = sc_dry] %>%
+        .[num_obs > 5, ] %>%
+        .[, num_obs := NULL]
+    )) %>%
+    # === sat breaks for grouping ===#
+    #' dryland production will have NA
+    mutate(sat_breaks = list(
+      quantile(
+        data[ir == "ir", sat],
+        prob = seq(0, 1, length = sat_cat_num + 1),
+        na.rm = TRUE
+      )
+    ))
+
+  #--- sat-breaks almost identical so, make them the same ---#
+  reg_data[2, ]$sat_breaks <- reg_data[1, ]$sat_breaks
+
+  return_data <-
+    reg_data %>%
+    mutate(data = list(
+      data %>%
+        # === define saturated thickness category variable ===#
+        .[
+          ,
+          sat_cat := cut(
+            sat,
+            breaks = sat_breaks,
[…]

### 2023-08-01 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.rmd
finalized for second submission

@@
 \textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
-We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}
+We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}
 
@@
 
-- First point: \textcolor{blue}{Quantiles are not recalculated each each.
+- First point: \textcolor{blue}{Quantiles are not recalculated each year.
 If we do so, then the same quantile would represent different aquifer thickness range every year.
-When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk.}
+When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk irrespective of year.}
 
@@
 For the 0\% quantile, we assume a saturated thickness of 9m.
-This minimum of saturated thickness was selected based on REF, which defines 9m as the minimum aquifer thickness needed to provide a functioning water supply.
+This minimum of saturated thickness was selected based on \textcolor{red}{Tim, please add REFs}, which defines 9m as the minimum aquifer thickness needed to provide a functioning water supply.
 We have modified the text on page X lines X-X of the revised manuscript to note these points.}
@@
 However, these variables can be thought of simply as shifters that tell us how much higher average yields are in a single county relative to another county and to account for trends in yields over time.
-As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analysing these variables - and indeed it is standard to not report and analyse fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
+As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analyzing these variables - and indeed it is standard to not report and analyze fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
 
@@
 Variations in aquifer conductivity would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between saturated thickness and drought risk exposure that would contribute to uncertainty in our estimates of observed relationships between drought impacts and aquifer conditions.
-Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+\textcolor{red}{(Tim, I am adding this sentence.) This noise is referred to as measurement error in the field of statistics and econometrics, and it has a well-known consequence in regression analysis, called attenuation bias.
+Specifically, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it true is.
+So, in our context, our estimated impact of aquifer thickness on the share of irrigation is conservative.
+In reality, the impact of aquifer thickness is likely to be larger.} Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
 This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought.
-To highlight these points, we have added the following text on page X lines X-X of our revised manuscript:}
+To highlight these points, we have added the following text in the Discussion section on page 10, the 4th paragraph of our revised manuscript:}
 
-  \textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer saturated thickness, well yields, and drought risk exposure.
-For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger saturated thickness (Butler et al., 2013; Korus and Hensen, 2020).
-While we identify statistically robust impacts of saturated thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.''}
+  \textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020].
+.
+These uncertainties can be considered as measurement errors in aquifer thickness as a proxy for well yield.
+A well-known consequence of measurement errors in regression analysis is attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001].
+That is, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it true is.
+Therefore, it is likely that our estimated impacts of aquifer thickness on drought risk is rather conservative, and the true impact of groundwater depletion is likely to be greater than what we reported in this article.
+Importantly, however, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Nonetheless, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
+This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.''}
 
@@
 Environmental Earth Sciences, 79(16), 382.}
+
+  + \textcolor{blue}{John Bound and Alan B Krueger. The extent of measurement error in longitudinal earnings data: Do two wrongs make a right? Journal of labor economics, 9(1):1–24, 1991.}
+
+  + \textcolor{blue}{Dean R Hyslop and Guido W Imbens. Bias from classical and other forms of measurement error. Journal of Business \& Economic Statistics, 19(4):475–481, 2001}
 
@@
 
-+ Section 1, 3rd Paragraph – ``Two aquifer pathways exist through aquifer depletion will..'' – confusing sentence – please replace ``through'' with ``by which''.
++ Section 1, 3rd Paragraph – ``Two aquifer pathways exist through aquifer depletion will.'' – confusing sentence – please replace ``through'' with ``by which''.
 

### 2023-08-01 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.tex
finalized for second submission

@@
   First point:
-  \textcolor{blue}{Quantiles are not recalculated each each.
+  \textcolor{blue}{Quantiles are not recalculated each year.
 If we do so, then the same quantile would represent different aquifer thickness range every year.
-When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk.}
+When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk irrespective of year.}
 \item
@@
 However, these variables can be thought of simply as shifters that tell us how much higher average yields are in a single county relative to another county and to account for trends in yields over time.
-As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analysing these variables - and indeed it is standard to not report and analyse fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
+As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analyzing these variables - and indeed it is standard to not report and analyze fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
 \item
@@
 Variations in aquifer conductivity would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between saturated thickness and drought risk exposure that would contribute to uncertainty in our estimates of observed relationships between drought impacts and aquifer conditions.
-Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+\textcolor{red}{(Tim, I am adding this sentence.) This noise is referred to as measurement error in the field of statistics and econometrics, and it has a well-known consequence in regression analysis, called attenuation bias.
+Specifically, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it true is.
+So, in our context, our estimated impact of aquifer thickness on the share of irrigation is conservative.
+In reality, the impact of aquifer thickness is likely to be larger.} Importantly though, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
 This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought.
-To highlight these points, we have added the following text on page X lines X-X of our revised manuscript:}
+To highlight these points, we have added the following text in the Discussion section on page 10, the 4th paragraph of our revised manuscript:}
 \end{itemize}
 
-\textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer saturated thickness, well yields, and drought risk exposure.
-For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger saturated thickness (Butler et al., 2013; Korus and Hensen, 2020).
-While we identify statistically robust impacts of saturated thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.''}
+\textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020].
+.
+These uncertainties can be considered as measurement errors in aquifer thickness as a proxy for well yield.
+A well-known consequence of measurement errors in regression analysis is attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001].
+That is, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it true is.
+Therefore, it is likely that our estimated impacts of aquifer thickness on drought risk is rather conservative, and the true impact of groundwater depletion is likely to be greater than what we reported in this article.
+Importantly, however, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Nonetheless, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
+This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.''}
 
@@
 Environmental Earth Sciences, 79(16), 382.}
+\item
+  \textcolor{blue}{John Bound and Alan B Krueger. The extent of measurement error in longitudinal earnings data: Do two wrongs make a right? Journal of labor economics, 9(1):1–24, 1991.}
+\item
+  \textcolor{blue}{Dean R Hyslop and Guido W Imbens. Bias from classical and other forms of measurement error. Journal of Business \& Economic Statistics, 19(4):475–481, 2001}
 \end{itemize}

### 2023-08-01 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.rmd
finalized for second submission

@@
 \textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript.
-We have endeavoured to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}
+We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}
 
@@
 
-\textcolor{blue}{Tim? this is realted to one of the reviewer 1's comments}
+\textcolor{blue}{This is a great point, which is also pointed out by another reviewer.
+In response, we have added the following texts in the Discussion section on page 10, the 4th paragraph of our revised manuscript.}
+
+\textcolor{blue}{``Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020].
+.
+These uncertainties can be considered as measurement errors in aquifer thickness as a proxy for well yield.
+A well-known consequence of measurement errors in regression analysis is attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001].
+That is, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it true is.
+Therefore, it is likely that our estimated impacts of aquifer thickness on drought risk is rather conservative, and the true impact of groundwater depletion is likely to be greater than what we reported in this article.
+Importantly, however, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Nonetheless, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
+This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.''}
 
@@
 
-\textcolor{blue}{Thank you for this comment, which we believe led to a slighly better model and acknowledging important limitations of our methods.
+\textcolor{blue}{Thank you for this comment, which we believe led to a slightly better model and acknowledging important limitations of our methods.
 County and year fixed effects were included in equation (1), but only year fixed effects were included in equation (2).
-As you gueesed, the problem of including county FEs lies in the massive elimination of variation in aquifer thickness.
+As you guessed, the problem of including county FEs lies in the massive elimination of variation in aquifer thickness.
 Please see Figure 6 in Appendix A for the histogram of (1) aquifer thickness demeaned by the county average, (2) aquifer thickness demeaned by the state average, and (3) aquifer thickness without demeaning.
@@
 The inclusion of State FEs does not diminish the variation in aquifer thickness.
-Therefore, while the inclusion of State FEs certainly does not provide as tight of a control as county FEs, we have decided to include State FEs for both corn and soybean irrigated share regrssions as it is better than not including them.
-Figure 7 compares the estiamted impacts of aquifer thickness with or without State FEs.
+Therefore, while the inclusion of State FEs certainly does not provide as tight of a control as county FEs, we have decided to include State FEs for both corn and soybean irrigated share regressions as it is better than not including them.
+Figure 7 compares the estimated impacts of aquifer thickness with or without State FEs.
 The negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included.
@@
 
-\textcolor{blue}{}
+\textcolor{blue}{We believe we fixed those typos.}

### 2023-08-01 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_2.tex
finalized for second submission

@@
 
-\textcolor{blue}{Thank you for this comment, which we believe led to a slighly better model and acknowledgin important limitations of our methods.
+\textcolor{blue}{Thank you for this comment, which we believe led to a slightly better model and acknowledging important limitations of our methods.
 County and year fixed effects were included in equation (1), but only year fixed effects were included in equation (2).
-As you gueesed, the problem of including county FEs lies in the massive elimination of variation in aquifer thickness.
+As you guessed, the problem of including county FEs lies in the massive elimination of variation in aquifer thickness.
 Please see Figure 6 in Appendix A for the histogram of (1) aquifer thickness demeaned by the county average, (2) aquifer thickness demeaned by the state average, and (3) aquifer thickness without demeaning.
@@
 
-\textcolor{blue}{}
+\textcolor{blue}{We believe we fixed those typos.}
 

### 2023-08-01 Codes/1_Analysis.rmd
finalized for second submission

@@
 ```{r}
-set.seed(283397)
-
-#--- global parameters ---#
-sc_base <- "31011"
-sc_dry_base <- "31011_0"
+# ir_share_data <- main_analysis$share_reg_data[[1]]
+# sat_seq <- main_analysis$sat_seq[[1]]
+# sandtotal_e <- main_analysis$sandtotal_med[[1]]
+# silttotal_e <- main_analysis$silttotal_med[[1]]
+# awc_e <- main_analysis$awc_med[[1]]
 
@@
   rowwise() %>%
+  dplyr::mutate(sc_base = "31011") %>%
   #--- the same sequence used for all the bootstrap iterations ---#
@@
   #++++++++++++++++++++++++++++++++++++
-  dplyr::mutate(yield_pred_data = list(
-    yield_analysis(data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
+  dplyr::mutate(yield_analysis_results = list(
+    yield_analysis(data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data, sc_base)
+  )) %>%
+  dplyr::mutate(
+    yield_pred_data = list(yield_analysis_results$yield_pred_data),
+    yield_semi_res = list(yield_analysis_results$semi_res)
+  ) %>%
+  dplyr::mutate(yield_dif_test = list(
+    lapply(
+      2:3,
+      \(x) {
+        test_dif_in_yield(
+          balance_ls = seq(min(data$balance), max(data$balance), length = 50),
+          base_q = 1,
+          comp_q = x,
+          yield_semi_res,
+          sat_text_data,
+          sc_base
+        ) %>%
+          .[, comp_q := x]
+      }
+    ) %>%
+      rbindlist()
   )) %>%
@@
   ))
+
+saveRDS(main_analysis, "Shared/Results/main_analysis.rds")
 ```
@@
 #++++++++++++++++++++++++++++++++++++
-boot_results_added <-
+set.seed(383954)
+
+all_results <-
   main_analysis %>%
   #--- bootsrap data ---#
-  dplyr::mutate(boot_data = list(
-    replicate(1000, boot(data), simplify = FALSE)
-  )) %>%
-  #--- get share data ---#
-  dplyr::mutate(share_boot_data = list(
-    lapply(
-      1:length(boot_data),
+  dplyr::mutate(boot_results = list(
+    parallel::mclapply(
+      1:1000,
       \(x) {
-        boot_data[[x]] %>%
+        print(x)
+
+        #--- bootstrap data ---#
+        boot_data <- boot(data, sc_base)
+
+        #--- yield analysis ---#
+        yield_pred_data <-
+          yield_analysis(
+            boot_data,
+            balance_seq,
+            sat_seq_eval,
+            sat_breaks,
+            sat_text_data,
+            sc_base,
+            boot = TRUE
+          )
+
+        #--- share analysis ---#
+        share_boot_data <-
+          boot_data %>%
           .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
           .[, balance_avg := mean(balance), by = sc_code]
-      }
+
+        share_pred_data <-
+          share_analysis(
+            ir_share_data = share_boot_data,
+            sat_seq,
+            sandtotal_med,
+            silttotal_med,
+            awc_med
           )
-  )) %>%
-  #--- yield analysis ---#
-  dplyr::mutate(yield_pred_data_boot = list(
-    parallel::mclapply(
-      1:length(boot_data),
-      \(x) {
-        print(x)
-        yield_analysis(boot_data[[x]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
-      },
-      mc.cores = parallel::detectCores() - 2
-    )
-  )) %>%
-  #--- share analysis ---#
-  dplyr::mutate(share_pred_data_boot = list(
-    parallel::mclapply(
-      1:length(share_boot_data),
-      \(x) {
-        print(x)
-        share_analysis(share_boot_data[[x]], sat_seq, sandtotal_med, silttotal_med, awc_med)
-      },
-      mc.cores = parallel::detectCores() - 2
-    )
-  )) %>%
-  #--- average yield analysis ---#
-  dplyr::mutate(avg_yield_pred_data_boot = list(
-    parallel::mclapply(
-      1:length(boot_data),
-      \(x) {
-        get_average_yield(yield_pred_data_boot[[x]], share_pred_data_boot[[x]])
+        #--- average yield ---#
+        avg_yield_pred_data <- get_average_yield(yield_pred_data, share_pred_data)
+
+        return(list(share_pred_data = share_pred_data, avg_yield_pred_data = avg_yield_pred_data))
       },
@@
   dplyr::mutate(
-    yield_pred_data_boot = list(rbindlist(yield_pred_data_boot)),
-    share_pred_data_boot = list(rbindlist(share_pred_data_boot)),
-    avg_yield_pred_data_boot = list(rbindlist(avg_yield_pred_data_boot))
+    share_pred_data_boot = list(purrr::map(boot_results, "share_pred_data") %>% rbindlist()),
+    avg_yield_pred_data_boot = list(purrr::map(boot_results, "avg_yield_pred_data") %>% rbindlist())
   ) %>%
-  dplyr::select(
-    crop, data, sat_seq, sat_text_data,
-    yield_pred_data, yield_pred_data_boot,
-    share_pred_data, share_pred_data_boot,
-    avg_yield_pred_data, avg_yield_pred_data_boot,
-  )
-
-saveRDS(boot_results_added, "Shared/Results/boot_results_added.rds")
-```
-
-```{r}
-#++++++++++++++++++++++++++++++++++++
-#+ Summarize
-#++++++++++++++++++++++++++++++++++++
-boot_results_added <- readRDS("Shared/Results/boot_results_added.rds")
-
-all_results <-
-  boot_results_added %>%
-  # rowwise() %>%
-  dplyr::mutate(yield_response_sum = list(
-    yield_pred_data_boot[, .(
-      # y_hat_boot = median(y_hat, na.rm = TRUE),
-      y_hat_se = sd(y_hat, na.rm = TRUE),
-      # dif_y_hat_boot = mean(dif_y_hat, na.rm = TRUE),
-      dif_y_hat_se = sd(dif_y_hat, na.rm = TRUE)
-    ),
-    by = .(balance, sat, sat_cat)
-    ] %>%
-      .[yield_pred_data, on = c("balance", "sat")] %>%
-      .[, `:=`(
-        lower = y_hat - 1.96 * y_hat_se,
-        upper = y_hat + 1.96 * y_hat_se,
-        dif_lower = dif_y_hat - 1.96 * dif_y_hat_se,
-        dif_upper = dif_y_hat + 1.96 * dif_y_hat_se
-      )]
-  )) %>%
   dplyr::mutate(share_sum = list(
     share_pred_data_boot[, .(
-      # ir_share_hat = mean(ir_share_hat, na.rm = TRUE),
       ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE)
@@
   )) %>%
-  data.table()
+  data.table() %>%
+  dplyr::select(
+    crop, data, sat_seq, sat_text_data,
+    yield_pred_data,
+    yield_dif_test,
+    share_pred_data,
+    share_sum,
+    avg_yield_pred_data,
+    avg_yield_pred_data_boot,
+    avg_yield_sum
+  )
 

### 2023-08-01 Codes/2_figures_tables.rmd
finalized for second submission

@@
 plot_yield_data <-
-  all_results[, .(crop, yield_response_sum)] %>%
+  all_results[, .(crop, yield_pred_data)] %>%
   rowwise() %>%
   dplyr::mutate(plot_data = list(
-    yield_response_sum[, crop_type := crop] %>%
+    yield_pred_data[, crop_type := crop] %>%
       .[, crop_type := case_when(
@@
   plot_yield_data %>%
-  mutate(g_fig = list(
+  dplyr::mutate(plot_data = list(
+    plot_data[, `:=`(
+      lower = y_hat - 1.96 * y_hat_se_modeled,
+      upper = y_hat + 1.96 * y_hat_se_modeled
+    )]
+  )) %>%
+  dplyr::mutate(g_fig = list(
       ggplot(plot_data) +
@@
 
+# Difference in the share of irrigated production (from 130)
+
+```{r}
+g_share_dif <-
+  ggplot(plot_data_ir_share[sat < 130, ]) +
+  geom_line(aes(y = dif_ir_share_hat, x = sat)) +
+  geom_ribbon(
+    aes(
+      ymin = dif_lower,
+      ymax = dif_upper,
+      x = sat
+    ),
+    alpha = 0.4
+  ) +
+  geom_hline(yintercept = 0, color = "red") +
+  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
+  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
+  xlab("Aquifer Thickness (meter)") +
+  ylab("Estimated Share of Irrigated Acres") +
+  facet_wrap(crop_type ~ ., ncol = 1) +
+  theme_fig +
+  theme(
+    legend.position = "bottom",
+    strip.background = element_blank(),
+    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
+  )
+
+ggsave(
+  file = "GitControlled/Writing/Figures/g_share_dif.pdf",
+  g_share_dif,
+  height = 4,
+  width = 5
+)
+```
+
 # Total impact of a decline in saturated thickness on average yield
@@
 g_dif <-
-  all_results[, .(crop, yield_response_sum)] %>%
+  all_results[, .(crop, yield_dif_test, sat_text_data)] %>%
   rowwise() %>%
   dplyr::mutate(plot_data = list(
-    yield_response_sum[, crop_type := crop] %>%
+    yield_dif_test[, crop_type := crop] %>%
       .[, crop_type := case_when(
@@
         crop_type == "soy" ~ "B. Sybean"
-      )]
+      )] %>%
+      setnames("comp_q", "sat_rank") %>%
+      sat_text_data[., on = "sat_rank"] %>%
+      .[, dif_upper := dif_hat + 1.96 * se] %>%
+      .[, dif_lower := dif_hat - 1.96 * se]
   )) %>%
@@
   dplyr::mutate(plot_data = list(
-    plot_data %>%
-      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
-      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))] %>%
-      .[!is.na(sat_rank), ] %>%
-      .[sat_rank != 1, ] %>%
-      unique(by = c("crop_type", "balance", "sat_cat"))
+     plot_data[, sat_cat_text := factor(sat_cat_text, levels = sat_cat_text_ls)]
   )) %>%
@@
           x = balance,
-          y = dif_y_hat
+          y = dif_hat
         ),
@@
   ) +
-  scale_y_continuous(breaks = -2:5, limits = c(-2, 5)) +
+  scale_y_continuous(breaks = -2:5, limits = c(-2, 6)) +
   ggtitle("A. Corn")
@@
 g_ay <- g_ay_corn / g_ay_soy
-
-# g_avg_yield_dif <-
-#   ggplot(plot_data_dif) +
-#   geom_ribbon(aes(x = -balance, ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
-#   geom_line(aes(x = -balance, y = dif_yield), color = "red") +
-#   geom_hline(yintercept = 0) +
-#   facet_grid(crop ~ sat_cat_text, scales = "free_y") +
-#   xlab("Water Deficit (mm)") +
-#   ylab("Difference in Yield (tonne/ha)") +
-#   theme_fig +
-#   theme(
-#     strip.background = element_blank(),
-#     strip.text = element_text(size = 9, family = "Times"),
-#     panel.spacing.y = unit(1, "cm")
-#   )
 
@@
 
-# g_avg_yield_response <-
-#   ggplot(plot_data_response) +
-#   geom_ribbon(aes(x = -balance, ymin = lower, ymax = upper), alpha = 0.3) +
-#   geom_line(aes(x = -balance, y = avg_yield)) +
-#   facet_grid(crop ~ sat_cat_text, scales = "free_y") +
-#   xlab("Water Deficit (mm)") +
-#   ylab("Yield (tonne/ha)") +
-#   theme_fig +
-#   theme(
-#     strip.background.x = element_blank(),
-#     panel.spacing = unit(0.2, "cm")
-#   )
-
 ggsave("GitControlled/Writing/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)

### 2023-08-01 Writing/manuscript.Rmd
finalized for second submission

--- 2023-07-29 Writing/manuscript.Rmd
@@
 In this paper, we evaluate how aquifer depletion affects the resilience of irrigated crop production to drought risk using over 30 years of data on historical corn and soybean yields, production areas, and aquifer conditions for the High Plains region in the United States.
-We show that aquifer depletion reduces the ability of farmers to sustain irrigated crop yields and production areas in drought years.
+We show that aquifer depletion reduces the ability of farmers to sustain irrigated crop yields and production areas in years and locations with large growing season water deficits.
 Our findings demonstrate that drought-related production losses on irrigated croplands increase non-linearly with aquifer depletion, highlighting the need for proactive aquifer conservation interventions to support adaptation and resilience to future increases in rainfall variability under climate change.
@@
   Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.
+
 bibliography: DRA.bib
@@
 # Main
-
 Groundwater is an essential input for agricultural production, providing a critical buffer against limited or variable surface water supplies for farmers in many parts of the world [@scanlon2023global].
 Globally, agriculture's dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change [@zhou2010impact; @wada2013multimodel; @wada2014sustainability; @kreins2015quantification; @florke2018water].
-However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades [@wada2010global; @famiglietti2011satellites; @scanlon2012groundwater; @konikow2015long; @bierkens2019non], which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
+However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades [@wada2010global; @famiglietti2011satellites; @scanlon2012groundwater; @konikow2015long; @bierkens2019non].
+If not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
 
@@
 Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks [@schlenker2009nonlinear; @lobell2014greater; @schlenker2010robust; @zhou2020connections; @borgomeo2020impact].
-Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought [@kuwayama2019estimating; @zipper2016drought; @zhu2022untangling; @zhu2022warming; @lu2020mapping; @davis2019sensitivity; @li2018changes].
+Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification and expansion to mitigate negative impacts of drought [@kuwayama2019estimating; @zipper2016drought; @zhu2022untangling; @zhu2022warming; @lu2020mapping; @davis2019sensitivity; @li2018changes].
 In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling [@foster2015well; @cotterman2018groundwater; @kahil2015modeling; @yoon2021coupled; @rad2020mod], with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage [@jain2021groundwater; @suter2021depletion].
@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs [@mieno2017price; @bhattarai2021impact].
-Second, depletion also reduces the pumping capacity of wells used to extract groundwater, which may create a physical constraint to the volume of irrigation that can be applied to a field [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous].
-The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
+Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous].
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted [@hecox2002calculation; @korus2020depletion] - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits [@foster2015well;@rouhi2020downside].
+The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
 Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production [@foster2014modeling; @rad2020effects].
-Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change [@ukkola2020robust; @chiang2021evidence; @cook2020twenty]
+Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change [@ukkola2020robust; @chiang2021evidence; @cook2020twenty].
 
-In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought risk using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
+In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
 Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production [@schlenker2009nonlinear; @lobell2014greater; @lu2018crop].
@@
 
-## Aquifer depletion increases sensitivity of crop yields to drought {#impact-intensive}
+## Aquifer depletion increases sensitivity of crop yields to water deficits {#impact-intensive}
 
@@
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \@ref(fig:yield-response)).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits [@rad2020effects, @foster2014modeling; @hrozencik2017heterogeneous].
-However, yield gradually increases as water deficit increases for counties with the highest level of aquifer thickness.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended water deficits [@rad2020effects, @foster2014modeling; @hrozencik2017heterogeneous].
+In contrast, yield continues to increase gradually as water deficit increases beyond 400 mm for counties with the highest level of aquifer thickness.
 
-Figure \@ref(fig:yield-dif) shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95 percent confidence interval for corn an soybean.
-For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is `r round(test_results[balance == 1000 & crop == "Corn" & comp_q == 3, dif_hat], digits = 2)` tonne/ha.
-At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \@ref(fig:yield-response)) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
-This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities [@foster2019assessing; @gibson2017case; @gibson2019benchmarking].
-Unlike corn, soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likely because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
+Figure \@ref(fig:yield-dif) shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95% confidence interval for corn and soybean.
+For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is `r round(test_results[balance == 1000 & crop == "Corn" & comp_q == 3, dif_hat], digits = 2)` tonne/ha.
+Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
+This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower saturated thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
+
+At low water deficits (below 200mm), Figure \@ref(fig:yield-dif) illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \@ref(fig:yield-response) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
+In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
+These counter-intuitive results reflect a combination of two factors.
+First, in years with small water deficits there is less need for irrigation and hence irrigated and rainfed yield differences are reduced.
+Second, years with small or even negative water deficits are associated with wetter conditions during the growing season, which can negatively impact yield if waterlogging occurs leading to greater risks of physical crop damage (e.g., lodging of grains), delays to planting/harvesting, nutrient leaching, and restricted root growth in waterlogged soils [@li2019excessive].
+Declines in irrigated yields could be indicative of over-irrigation by farmers, which has been shown to occur in the HPA region in wet years [@foster2019assessing; @gibson2017case; @gibson2019benchmarking], exacerbating these risks.
+However, the small declines in irrigated yields at low or negative levels of water deficit may also reflect the fact that wetter years are typically cooler, reducing positive heat effects on crop growth and yield formation.
 
@@
 -->
-```{r yield-response, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn and soybean in US High Plains", out.width = "6.5in", cache = FALSE}
+```{r yield-response, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.
+Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+The histograms on the top of each panel show the distribution of county-year water deficit values included in regressions for each crop.", out.width = "6.5in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_intensive.pdf"))
@@
 
-```{r yield-dif, out.width = "6in", fig.cap = "The difference in yield for the counties with the second and third (highest) aquifer thickness groups relative to the first (lowest) aquifer thickness group."}
+```{r yield-dif, out.width = "6in", fig.cap = "The difference in estimated county-level irrigated corn (top panels) and soybean (bottom panels) yields for counties in the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile."}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_ir_yield_dif.pdf"))
@@
 
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (more than 800mm water deficit), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production.
+While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greter than 700mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically [@rad2020effects; @foster2014modeling; @hrozencik2017heterogeneous; @deines2020transitions], but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.
+Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically [@rad2020effects; @foster2014modeling; @hrozencik2017heterogeneous; @deines2020transitions], but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
 
-Figure \@ref(fig:ir-share) shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean.
-Figure \@ref(fig:ir-share) demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.8 (80\%) for an average county-level aquifer thickness of 150m to 0.5 (50\%) for an average county-level aquifer thickness of 12m.
-Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn [@zipper2016drought; @ruess2022irrigation].
+```{r}
+all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
 
-```{r ir-share, fig.cap = "The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)", out.width = "5in"}
+corn_share <- all_results[crop == "corn", share_pred_data][[1]]
+
+share_high <-
+  corn_share[which.min(abs(sat - 130)), ir_share_hat] %>%
+  round(digits = 2)
+
+share_low <-
+  corn_share[which.min(abs(sat - 10)), ir_share_hat] %>%
+  round(digits = 2)
+```
+
+Figure \@ref(fig:ir-share) shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
+Figure \@ref(fig:ir-share) demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around `r share_high` (`r share_high * 100`%) for an average county-level aquifer thickness of 130m to `r share_low` (`r share_low * 100`%)  for an average county-level aquifer thickness of 10m (See Appendix \@ref(test-dif-share) for formal statistical testings).
+Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn [@zipper2016drought; @ruess2022irrigation]
+
+```{r ir-share, fig.cap = "The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95 percent confidence interval)", out.width = "5in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share.pdf"))
@@
   data.table() %>%
-  .[(abs(balance - 950)) < 1, by = crop]
+  .[crop == "corn", ] %>%
+  .[(abs(balance - 950)) < 5, ]
 ```
-The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
-Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 12.2, 59.1, 106.1, and 153.0  meters^[Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \@ref(reg-conf) for estimated average yield response curves with 95 percent confidence interval.].
-While you can see similar trends as presented in Figure \@ref(fig:yield-response), the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
 
-Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area production of the 59.1m, 106.1m, and 153.0m aquifer thickness levels relative to 12.2m with the 95 percent confidence interval for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 59.1m and 106.1m aquifer thickness, respectively).
+The overall influence of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \@ref(impact-intensive) and \@ref(impact-share).
+Combining both responses, Figure \@ref(fig:tot-impact) shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where the weights are production area shares) for corn and soybean at four values of aquifer thickness: 10, 50, 90, and 130 meters.
+While similar trends are observed as in Figure \@ref(fig:yield-response), the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced due to lower estimated irrigated area share for smaller aquifer thickness values.
+For example, for an aquifer thickness of 10 m, reductions in average yields occur for smaller water deficits than are required to trigger reductions in per-area irrigated yields alone.
+Similarly, the overall magnitude of yield reduction is greater for average yields than for per-area irrigated yields.
+This is because low aquifer thickness values have a greater estimated share of crop area under rainfed production, which magnifies average yield declines with increasing water deficits.
+
+Figure \@ref(fig:dif-tot-impact) shows the difference in average per-area productivity for 50 m, 90 m, and 130 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 350mm.
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
 This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.
-greater than around 800mm) and are smaller in overall magnitude.
-For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
-Further, notice that the decline in average per-area production at a seasonal water deficit of 950mm is not linear.
-Going from 153.0m aquifer thickness to 106.1m (decline of 46.9m) would result in `r round(res_at_950[sat == 502 & crop == "corn", avg_yield] - res_at_950[sat == 348 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
-This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
-Going from 106.1m aquifer thickness to 59.1m (decline of 47m) would result in `r round(res_at_950[sat == 348 & crop == "corn", avg_yield] - res_at_950[sat == 194 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
-However, going from 59.1m aquifer thickness to 12.2m (decline of 46.9m) would result in `r round(res_at_950[sat == 194 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
-That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
-This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \@ref(fig:ir-share).
-When aquifer thickness level is high enough (e.g., 100m), a small decline does not reduce the share of irrigated acres.
-However, when aquifer thickness level is low (e.g., 40m), a small decline does reduce the share of irrigated acres.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600mm) and are smaller in overall magnitude.
+For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25% and 15% lower, respectively, when differentiating between areas with the 130m and 10m aquifer thicknesses.
 
-```{r tot-impact, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness", out.width = "6in"}
+Figure \@ref(fig:tot-impact) and Figure \@ref(fig:dif-tot-impact) also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
+At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only `r round(res_at_950[sat == 130 & crop == "corn", avg_yield] - res_at_950[sat == 90 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \@ref(fig:ir-share).
+Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in `r round(res_at_950[sat == 90 & crop == "corn", avg_yield] - res_at_950[sat == 50 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in `r round(res_at_950[sat == 50 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+Similar but less pronounced effects are also observed for soybean.
+This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \@ref(fig:ir-share)), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers ability to effectively and reliably buffer crops against water deficits during the growing season.
+
+```{r tot-impact, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers choices about irrigated area shares in response to different water deficit and aquifer conditions.
+See Appendix A for estimated average yield response curves with 95 percent confidence intervals displayed, which are omitted here for visual clarity.", out.width = "6in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_total_impact.pdf"))
@@
 
-```{r dif-tot-impact, fig.cap = "Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean", out.width = "6in", cache = FALSE}
+```{r dif-tot-impact, fig.cap = "Difference in average yield of corn and soybean for aquifer thickness levels of 50 m, 90 m, and 130 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95 percent confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5 percent level of significance.", out.width = "6in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_avg_yield_dif.pdf"))
@@
 Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production [@schlenker2009nonlinear; @lobell2014greater; @schlenker2010robust; @zhou2020connections; @borgomeo2020impact] or for irrigated lands where water supply constraints are assumed to be minimal [@kuwayama2019estimating; @zipper2016drought; @zhu2022untangling; @zhu2022warming; @lu2020mapping; @davis2019sensitivity; @li2018changes; @luan2021combined].
-Our analysis adds new empirical insights to this literature which show explicitly that reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
-Critically, this suggests that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.
+Our analysis adds new empirical insights to this literature by quantifying how reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
+Critically, our findings suggest that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.
 
@@
 Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally.
-Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and drought impacts is highly non-linear when accounting for farmers decisions about irrigated production areas.
+Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and climatic water deficits during the growing season is highly non-linear when accounting for farmers decisions about irrigated production areas.
 This indicates that sustainable groundwater conservation efforts [@macewan2017hydroecon; @butler2020charting; @elshall2020groundwater] should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
 Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production [@foster2017effects; @deines2020transitions].
-For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure  50\% of current aquifer storage remains in 50 years time (so called 50/50 rule) [@closas2018chronicle].
+For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure  50% of current aquifer storage and thickness remains in 50 years time (so called 50/50 rule) [@closas2018chronicle].
 The effectiveness of such targets could likely be improved substantially by setting conservation targets based on empirical evidence about the level of aquifer thickness that is required to ensure adequate levels of drought protection, both now and with future climate change.
@@
 Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring [@edreira2020combining; @deines2021million], would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
+Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness [@butler2013interpretation; @korus2020depletion].
+These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias [@bound1991extent; @hyslop2001bias], suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk.
+Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA.
 This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
 
-Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types.
+Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types or irrigation management practices.
 Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies [@bhattarai2021impact; @deines2019quantifying; @manning2017producer].
-Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
+Improvements to irrigation scheduling practices and technologies, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas [@deines2019quantifying; @glose2022quantifying] and in parts of Texas [@mrad2020peak] and which are proposed as part of California's Sustainable Groundwater Management Act (SGMA) [@berbel2019droughts; @lubell2020sustainable], could also help to improve the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials [@foster2019assessing] and the importance of conservation [@marston2022].
+However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks [@grafton2018paradox; @perez2021agricultural].
+Future research should explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
 
@@
 Our study focuses on the High Plains Aquifer (HPA) in the Central United States.
-Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually [@fenichel2016measuring @smidt2016complex].
+Irrigated agriculture is integral to the landscape and economy of the HPA, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually [@smidt2016complex; @fenichel2016measuring].
 To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.
 
-County-level irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S.
+County-level annual irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S.
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS).
@@
 However, it is known that groundwater is the main source of water for irrigation for counties that overlie the HPA due to the limited availability of surface water resources in this area.
-Therefore, we retain both rainfed and irrigated yield observations for counties where greater than 75\% of the county area overlies the boundary of the HPA.
+Therefore, we retain both rainfed and irrigated yield observations for counties where greater than 75% of the county area overlies the boundary of the HPA.
 For counties that do not meet this condition, only rainfed yield observations are obtained to avoid the possibility that irrigated production may be only partly dependent on use of groundwater from the HPA.
-We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production.
+We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production as a point of comparison to our main irrigated crop yield analyses.
 
 Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset [@Abatzoglou2013].
-The gridMET data is a gridded daily weather dataset with a spatial resolution of ~4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States [@abatzoglou2016impact; @pereira2015crop; @crane2018machine; @venkatappa2021impacts; @zhu2019dissecting].
+gridMET is a gridded daily weather dataset with a spatial resolution of ~4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States [@abatzoglou2016impact; @pereira2015crop; @crane2018machine; @venkatappa2021impacts; @zhu2019dissecting].
 Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample.
-Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET, and were then used to calculate daily reference crop evapotranspiration for each grid cell.
+Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET in order to calculate daily reference crop evapotranspiration for each grid cell.
 From these data, we computed the seasonal water deficit (\(WB_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
@@
 Figure \@ref(fig:deficit-yield-hist) presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
-The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \@ref(fig:deficit-yield-hist)).
+The importance of irrigation is clearly evident in years with large water deficits, with rainfed yields dipping substantially whereas irrigated yields are, on average, largely unaffected (Figure \@ref(fig:deficit-yield-hist)).
 
-```{r deficit-yield-hist, fig.cap = "Historical dryland yield, irrigated yield, and water deficit", out.height = "6in"}
+```{r deficit-yield-hist, fig.cap = "Historical rainfed yield, irrigated yield, and water deficit", out.height = "6in"}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_y_wd_all.pdf"))
@@
 
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA available at Hydroshare [@haacker2023].
-These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935, derived based on extensive water level observations from monitoring wells and detailed maps of both land surface elevation and the bedrock elevation of the HPA.
-aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data.
-aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [@fenichel2016measuring; @haacker2016water; @deines2020transitions].
+As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research [@haacker2016water; @haacker2023].
+These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935.
+Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
+Aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data.
+Aquifer thickness values were only estimated for counties with irrigated yield observations.
+Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [@fenichel2016measuring; @haacker2016water; @deines2020transitions].
 Figure \@ref(fig:sat-map) presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
@@
 
-```{r sat-map, fig.cap = "Map showing aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+```{r sat-map, fig.cap = "Map showing county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
 Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system", out.width = "6in"}
@@
 
-To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield to drought is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
+To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield response to seasonal water deficits is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
 Our regression model is given in Equation \@ref(eq:eqn1), with the same model formulation used for corn and soybean:
@@
 \begin{equation}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t} (\#eq:eqn1)
+y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} (\#eq:eqn1)
 \end{equation}
@@
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3\%)\), \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness.
+The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3%)\), \([33.3-66.7%)\), and \([66.7% - 100%]\) quantile ranges of aquifer thickness.
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
 Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \@ref(spline-basis) for further explanations on how spline-based regression works).
-To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
+To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
-This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work (REF).
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without the prior requirement to assume any functional forms.
+This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work [@foster2015well; @foster2015analysis; @hrozencik2017heterogeneous].
 The model is given in Equation \@ref(eq:eqn2) below:
@@
 State-year fixed effects and error term are represented by $\theta_{s,t}$ and $v_{i,t}$, respectively.
-One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
+Unlike for our regression models of per-area yields, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
 The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time.
-Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.
-The trade-off here is tighter controls on potential unobservables that may be correlated with aquifer thickness (see Appendix \@ref(county-fe) for more discussions).
+Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
+Instead, through robustness checks, we demonstrate that our results are consistent when using state fixed effects.
+A formal discussion of these results and trade-offs associated with the choice of fixed effects is provided in Appendix \@ref(county-fe).
 
@@
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
-Its standard error is estimated using bootstrap clustered by county.
+Standard errors of this analysis are estimated using bootstrap methods clustered by county.
+Note that irrigated area share and average crop yield estimates use  data only from counties that have both irrigated and rainfed production observations.
+Counties with solely rainfed production observation data (Section \@ref(data-sets)) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \@ref(eq:eqn1)).
 
@@
 All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All the computations were done in R [@R] and all the raw data-sets used in this study are publicly accessible.
+All analyses were performed in R [@R] and all the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
@@
 This data is used for Figure \@ref(fig:deficit-yield-hist) and the regression analysis.
-The aquifer thickness data were obtained from Hydroshare [@haacker2023].
+The aquifer thickness data were obtained from \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/} [@haacker2023].
 This data is used for Figure \@ref(fig:sat-map) and the regression analysis.
@@
 Other key R packages include ggplot2 [@Rggplot2] for creating plots, data.table [@Rdatatable] for data wrangling, and sf [@Rsf] for spatial data operations.
-The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.
+The data used for the regression analysis can be generated using our codes by following the instructions available at the linked Github repository.
 
@@
 
-# More on the impact of water deficit on irrigated and average yield {#reg-conf}
+# Uncertainties in estimated impacts of water deficit on irrigated and average yield {#reg-conf}
 
@@
 
-In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented as they would crowd the figures excessively.
-This section provides visualization of yield response with the 95 $\%$ confidence interval to convey the degree of uncertainty in our estimation.
-Figures \@ref(fig:irrigated-yield-ind-corn) and \@ref(fig:irrigated-yield-ind-soy) present the impact of water deficit on crop yield by aquifer thickness category with the 95$\%$ confidence interval for irrigated corn and soybean, respectively.
-Figure \@ref(fig:avg-yield-ind) presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95$\%$ confidence interval.
+In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
+This section provides visualization of yield response with the 95% confidence interval to convey the degree of uncertainty in our estimation.
+Figures \@ref(fig:irrigated-yield-ind-corn) and \@ref(fig:irrigated-yield-ind-soy) present the impact of water deficit on crop yield by aquifer thickness category with the 95% confidence interval for irrigated corn and soybean, respectively.
+Figure \@ref(fig:avg-yield-ind) presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95% confidence interval.
 
-```{r irrigated-yield-ind-corn, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
+```{r irrigated-yield-ind-corn, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)", out.width = "6in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_with_conf_corn.pdf"))
@@
 
-```{r irrigated-yield-ind-soy, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains with 95 percent confidence interval", out.width = "6in", cache = FALSE}
+```{r irrigated-yield-ind-soy, fig.cap = "The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)", out.width = "6in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_with_conf_soy.pdf"))
@@
 
-```{r avg-yield-ind, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (The shaded areas represent the 95 percent confidence interval)", out.width = "6.5in", cache = FALSE}
+```{r avg-yield-ind, fig.cap = "Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95 percent confidence interval (grey shaded area)", out.width = "6.5in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_avg_yield.pdf"))
@@
 \clearpage
+
+# Test of the Impact of Aquiefr Thickness on the Share of Irrigated Acres {#test-dif-share}
+
+```{=Tex}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{C.\arabic{figure}}
+```
+
+```{r}
+sat_ref <-
+  readRDS(here::here("Shared/Results/all_results.rds")) %>%
+  .[1, sat_seq] %>%
+  .[[1]] %>%
+  .[length(.)]
+```
+
+Figure \@ref(fig:test-sat-share) presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of `r sat_ref`m.
+
+```{r test-sat-share, fig.cap = "Decline in the share of irrigated acres as a function of aquifer thickness relative to aquifer thickness level of 130m (The shaded areas represent the 95 percent confidence interval)", out.width = "6.5in", cache = FALSE}
+knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_share_dif.pdf"))
+```
 
@@
 For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
-Most important, adding state fixed effects, which control for unobservables at the state level, does not change the estimation results much.
-Therefore, even though there may be unobservables that are spatially correlated with aquifer thickness, their correlation may not be very strong to cause significant bias.
-Clearly, this does not mean that our inability to control for other county-level unobservables beyond sand percentage, silt percentage, and water holding capacity is not causing bias in our estimation of the impact of aquifer thickness on the share of irrigated acres.
+Importantly, adding state fixed effects, which control for unobservables at the state level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
 
@@
 
-Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of $x$ on $y$.
+Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the impact of $x$ on $y$.
 Figure \@ref(fig:ex-fx) presents two examples of two different sets of coefficients and their resulting representation of $f(x)$.

### 2023-08-01 Writing/manuscript.tex
finalized for second submission

@@
 \usepackage{float}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \ifLuaTeX
@@
 Globally, agriculture's dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change \citep{zhou2010impact, wada2013multimodel, wada2014sustainability, kreins2015quantification, florke2018water}.
-However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}, which, if not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
+However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}.
+If not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
 
@@
 Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact}.
-Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification to mitigate negative impacts of drought \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes}.
+Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification and expansion to mitigate negative impacts of drought \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes}.
 In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling \citep{foster2015well, cotterman2018groundwater, kahil2015modeling, yoon2021coupled, rad2020mod}, with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage \citep{jain2021groundwater, suter2021depletion}.
@@
 This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
-If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
-The combined effect of these two factors could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
+If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well, rouhi2020downside}.
+The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
 Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production \citep{foster2014modeling, rad2020effects}.
-Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}
+Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}.
 
-In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought risk using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
+In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
 Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater, lu2018crop}.
@@
 \hypertarget{impact-intensive}{%
-\subsection{Aquifer depletion increases sensitivity of crop yields to drought}\label{impact-intensive}}
+\subsection{Aquifer depletion increases sensitivity of crop yields to water deficits}\label{impact-intensive}}
 
@@
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended precipitation deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
-However, yield gradually increases as water deficit increases for counties with the highest level of aquifer thickness.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended water deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
+In contrast, yield continues to increase gradually as water deficit increases beyond 400 mm for counties with the highest level of aquifer thickness.
 
-Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95 percent confidence interval for corn an soybean.
-For corn, in years with water deficit higher than about 800 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference is 1.15 tonne/ha.
-At low water deficits (below 200mm), counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields than counties with the lowest level of aquifer thickness (1st quantile).
-This counter-intuitive result may be due to the increased possibility of over-irrigation in regions with high aquifer thickness and well capacities \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}.
-Unlike corn, soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likely because soybean is not as water demanding as corn and even the counties the lowest aquifer thickness level can catch up with water demand in severe drought years just as well as counties with higher aquifer thickness levels.
+Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95\% confidence interval for corn and soybean.
+For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is 1.15 tonne/ha.
+Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
+This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower saturated thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
+
+At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response} experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
+In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
+These counter-intuitive results reflect a combination of two factors.
+First, in years with small water deficits there is less need for irrigation and hence irrigated and rainfed yield differences are reduced.
+Second, years with small or even negative water deficits are associated with wetter conditions during the growing season, which can negatively impact yield if waterlogging occurs leading to greater risks of physical crop damage (e.g., lodging of grains), delays to planting/harvesting, nutrient leaching, and restricted root growth in waterlogged soils \citep{li2019excessive}.
+Declines in irrigated yields could be indicative of over-irrigation by farmers, which has been shown to occur in the HPA region in wet years \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}, exacerbating these risks.
+However, the small declines in irrigated yields at low or negative levels of water deficit may also reflect the fact that wetter years are typically cooler, reducing positive heat effects on crop growth and yield formation.
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn and soybean in US High Plains}\label{fig:yield-response}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.
+Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+The histograms on the top of each panel show the distribution of county-year water deficit values included in regressions for each crop.}\label{fig:yield-response}
 \end{figure}
@@
 
-\caption{The difference in yield for the counties with the second and third (highest) aquifer thickness groups relative to the first (lowest) aquifer thickness group.}\label{fig:yield-dif}
+\caption{The difference in estimated county-level irrigated corn (top panels) and soybean (bottom panels) yields for counties in the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.}\label{fig:yield-dif}
 \end{figure}
@@
 
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (more than 800mm water deficit), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production.
+While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greter than 700mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations of drought exposure and groundwater scarcity.
+Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
 
-Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in drought risk exposure across the aquifer.
-Figure \ref{fig:ir-share} demonstrates clearly that aquifer thickness has large and significant impacts on the share of irrigated production, in particular for corn.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 150m to 0.63 (63\%) for an average county-level aquifer thickness of 12m.
-Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn potentially due to soybean's lower water requirements and susceptibility to drought than corn \citep{zipper2016drought, ruess2022irrigation}.
+Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
+Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 130m to 0.61 (61\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
+Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}
 
@@
 
-\caption{The impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean in the US High Plains (Note: shaded area represents 95 percent confidence interval)}\label{fig:ir-share}
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95 percent confidence interval)}\label{fig:ir-share}
 \end{figure}
 
-\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25j}{%
-\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)j}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%j}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25j}}
+\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}{%
+\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
 
-The overall effects of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where weights are production area shares) for corn and soybean at thee values of aquifer thickness: 10, 50, 90, and 130 meters\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \ref{reg-conf} for estimated average yield response curves with 95 percent confidence interval.}.
-While you can see similar trends as presented in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced mainly dues to lower irrigated area for lower aquifer thickness level (rainfed production has a larger share).
+The overall influence of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
+Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where the weights are production area shares) for corn and soybean at four values of aquifer thickness: 10, 50, 90, and 130 meters.
+While similar trends are observed as in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced due to lower estimated irrigated area share for smaller aquifer thickness values.
+For example, for an aquifer thickness of 10 m, reductions in average yields occur for smaller water deficits than are required to trigger reductions in per-area irrigated yields alone.
+Similarly, the overall magnitude of yield reduction is greater for average yields than for per-area irrigated yields.
+This is because low aquifer thickness values have a greater estimated share of crop area under rainfed production, which magnifies average yield declines with increasing water deficits.
 
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area production of the 50m, 90m, and 130m aquifer thickness levels relative to 10m with the 95 percent confidence interval for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when precipitation deficits are small (less than around 200mm and 300mm for 50m and 90m aquifer thickness, respectively).
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 50 m, 90 m, and 130 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 350mm.
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
 This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (i.e.~greater than around 800mm) and are smaller in overall magnitude.
-For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the highest (3rd quantile) and lowest (1st quantile) aquifer thickness.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600mm) and are smaller in overall magnitude.
+For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 130m and 10m aquifer thicknesses.
 
-Further, notice that the impact of aquifer thickness on the decline in average per-area production is highly non-linear at high seasonal water deficits.
-Going from 130m aquifer thickness to 90m (decline of 40m) would result in merely 0.1 tonne/ha of production loss for corn.
-This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to aquifer thickness when the aquifer thickness level is high.
-Going from 90m aquifer thickness to 50m (decline of 40m) would result in 0.42 tonne/ha of production loss for corn.
-However, going from 50m aquifer thickness to 10m (decline of 40m) would result in 2.18 tonne/ha of production loss for corn.
-That is, the same degree of decline is much more damaging when the decline is from a low aquifer thickness level than a higher aquifer thickness level.
-This is primarily due to the non-linear impact of aquifer thickness on the share of irrigated corn acres as shown in Figure \ref{fig:ir-share}.
-When aquifer thickness level is high enough (e.g., 100m), a small decline does not reduce the share of irrigated acres.
-However, when aquifer thickness level is low (e.g., 40m), a small decline does reduce the share of irrigated acres.
+Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
+At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only 0.11 tonne/ha of production loss for corn.
+This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
+Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in 0.77 tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in 2.09 tonne/ha of production loss for corn.
+Similar but less pronounced effects are also observed for soybean.
+This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers ability to effectively and reliably buffer crops against water deficits during the growing season.
 
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness}\label{fig:tot-impact}
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers choices about irrigated area shares in response to different water deficit and aquifer conditions.
+See Appendix A for estimated average yield response curves with 95 percent confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
 \end{figure}
@@
 
-\caption{Difference in average yield relative to the 1st aquifer thickness quantile for corn and soybean}\label{fig:dif-tot-impact}
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 50 m, 90 m, and 130 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95 percent confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5 percent level of significance.}\label{fig:dif-tot-impact}
 \end{figure}
@@
 Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes, luan2021combined}.
-Our analysis adds new empirical insights to this literature which show explicitly that reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
-Critically, this suggests that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.
+Our analysis adds new empirical insights to this literature by quantifying how reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
+Critically, our findings suggest that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.
 
@@
 Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally.
-Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and drought impacts is highly non-linear when accounting for farmers decisions about irrigated production areas.
+Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and climatic water deficits during the growing season is highly non-linear when accounting for farmers decisions about irrigated production areas.
 This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon, butler2020charting, elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
 Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production \citep{foster2017effects, deines2020transitions}.
-For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure 50\% of current aquifer storage remains in 50 years time (so called 50/50 rule) \citep{closas2018chronicle}.
+For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure 50\% of current aquifer storage and thickness remains in 50 years time (so called 50/50 rule) \citep{closas2018chronicle}.
 The effectiveness of such targets could likely be improved substantially by setting conservation targets based on empirical evidence about the level of aquifer thickness that is required to ensure adequate levels of drought protection, both now and with future climate change.
@@
 Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining, deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
-Similarly, unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk exposure.
-For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to drought despite similar or larger aquifer thickness \citep{butler2013interpretation, korus2020depletion}.
-While we identify statistically robust impacts of aquifer thickness on drought risk despite these uncertainties, improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to refine understanding of patterns of drought vulnerability across the HPA.
+Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk.
+For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness \citep{butler2013interpretation, korus2020depletion}.
+These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias \citep{bound1991extent, hyslop2001bias}, suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk.
+Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA.
 This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
 
-Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types.
+Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types or irrigation management practices.
 Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact, deines2019quantifying, manning2017producer}.
-Improvements to irrigation scheduling and water management practices, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas \citep{deines2019quantifying, glose2022quantifying} and proposed as part of California's SGMA (Sustainable Groundwater Management Act) \citep{berbel2019droughts, lubell2020sustainable}, could also play an important role in improving the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
-These kinds of interventions would be likely to reduce - but not eliminate - the increases in drought risk caused by declining aquifer thickness, and could bring additional benefits of reducing over-use of water in wetter years through greater awareness of efficiency saving potentials \citep{foster2019assessing} and the importance of conservation \citep{marston2022} amongst growers.
-Future research should therefore explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
+Improvements to irrigation scheduling practices and technologies, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas \citep{deines2019quantifying, glose2022quantifying} and in parts of Texas \citep{mrad2020peak} and which are proposed as part of California's Sustainable Groundwater Management Act (SGMA) \citep{berbel2019droughts, lubell2020sustainable}, could also help to improve the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
+Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials \citep{foster2019assessing} and the importance of conservation \citep{marston2022}.
+However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks \citep{grafton2018paradox, perez2021agricultural}.
+Future research should explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
 
@@
 Our study focuses on the High Plains Aquifer (HPA) in the Central United States.
-Irrigated agriculture is central to the landscape and economy of the High Plains Aquifer, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep[ \citet{smidt2016complex}]{fenichel2016measuring}.
+Irrigated agriculture is integral to the landscape and economy of the HPA, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{smidt2016complex, fenichel2016measuring}.
 To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.
 
-County-level irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S.
+County-level annual irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S.
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS).
@@
 For counties that do not meet this condition, only rainfed yield observations are obtained to avoid the possibility that irrigated production may be only partly dependent on use of groundwater from the HPA.
-We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production.
+We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production as a point of comparison to our main irrigated crop yield analyses.
 
 Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset \citep{Abatzoglou2013}.
-The gridMET data is a gridded daily weather dataset with a spatial resolution of \textasciitilde4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
+gridMET is a gridded daily weather dataset with a spatial resolution of \textasciitilde4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
 Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample.
-Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET, and were then used to calculate daily reference crop evapotranspiration for each grid cell.
+Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET in order to calculate daily reference crop evapotranspiration for each grid cell.
 From these data, we computed the seasonal water deficit (\(WB_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
@@
 Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
-The importance of irrigation is clearly evident in these extreme drought years, with rainfed yields dipping substantially whereas irrigated yields are mostly unaffected (Figure \ref{fig:deficit-yield-hist}).
+The importance of irrigation is clearly evident in years with large water deficits, with rainfed yields dipping substantially whereas irrigated yields are, on average, largely unaffected (Figure \ref{fig:deficit-yield-hist}).
 
@@
 
-\caption{Historical dryland yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
+\caption{Historical rainfed yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
 \end{figure}
 
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness for the HPA available at Hydroshare \citep{haacker2023}.
-These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935, derived based on extensive water level observations from monitoring wells and detailed maps of both land surface elevation and the bedrock elevation of the HPA.
-aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data.
-aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 12m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
+As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water, haacker2023}.
+These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935.
+Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
+Aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data.
+Aquifer thickness values were only estimated for counties with irrigated yield observations.
+Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
 Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
@@
 
-\caption{Map showing aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+\caption{Map showing county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
 Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
@@
 
-To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield to drought is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
+To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield response to seasonal water deficits is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
 Our regression model is given in Equation \eqref{eq:eqn1}, with the same model formulation used for corn and soybean:
@@
 \begin{equation}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \phi_t + v_{i,t} \label{eq:eqn1}
+y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
@@
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3\%)\), \([33.3-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness.
+The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3%)
+\), \([33.3-66.7%)
+\), and \([66.7% - 100%]
+\) quantile ranges of aquifer thickness.
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
 Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
-To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\phi_t\)) are also included and standard error estimation is clustered by county.
+To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
 
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without needing to assume any functional forms.
+To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without the prior requirement to assume any functional forms.
 This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work \citep{foster2015well, foster2015analysis, hrozencik2017heterogeneous}.
@@
 State-year fixed effects and error term are represented by \(\theta_{s,t}\) and \(v_{i,t}\), respectively.
-One of the limitations here is that we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable unlike our regression model of the impact of water balance on yield.
+Unlike for our regression models of per-area yields, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
 The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time.
-Including county fixed effects effectively eliminate the cross-sectional variation in aquifer thickness.
-The trade-off here is tighter controls on potential unobservables that may be correlated with aquifer thickness (see Appendix \ref{county-fe} for more discussions).
+Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
+Instead, through robustness checks, we demonstrate that our results are consistent when using state fixed effects.
+A formal discussion of these results and trade-offs associated with the choice of fixed effects is provided in Appendix \ref{county-fe}.
 
@@
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
-Its standard error is estimated using bootstrap clustered by county.
+Standard errors of this analysis are estimated using bootstrap methods clustered by county.
+Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations.
+Counties with solely rainfed production observation data (Section \ref{data-sets}) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \eqref{eq:eqn1}).
 
@@
 All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All the computations were done in R \citep{R} and all the raw data-sets used in this study are publicly accessible.
+All analyses were performed in R \citep{R} and all the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
@@
 This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-The aquifer thickness data were obtained from Hydroshare \citep{haacker2023}.
+The aquifer thickness data were obtained from \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/} \citep{haacker2023}.
 This data is used for Figure \ref{fig:sat-map} and the regression analysis.
@@
 Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
-The data used for the regression analysis can be generated using our codes by following the instruction available at the Github repository.
+The data used for the regression analysis can be generated using our codes by following the instructions available at the linked Github repository.
 
@@
 \hypertarget{reg-conf}{%
-\section{More on the impact of water deficit on irrigated and average yield}\label{reg-conf}}
+\section{Uncertainties in estimated impacts of water deficit on irrigated and average yield}\label{reg-conf}}
 
@@
 
-In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented as they would crowd the figures excessively.
-This section provides visualization of yield response with the 95 \(\%\) confidence interval to convey the degree of uncertainty in our estimation.
-Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\(\%\) confidence interval for irrigated corn and soybean, respectively.
-Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\(\%\) confidence interval.
+In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
+This section provides visualization of yield response with the 95\% confidence interval to convey the degree of uncertainty in our estimation.
+Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\% confidence interval for irrigated corn and soybean, respectively.
+Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\% confidence interval.
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-corn}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-corn}
 \end{figure}
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains with 95 percent confidence interval}\label{fig:irrigated-yield-ind-soy}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-soy}
 \end{figure}
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (The shaded areas represent the 95 percent confidence interval)}\label{fig:avg-yield-ind}
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:avg-yield-ind}
 \end{figure}
@@
 \clearpage
+
+\hypertarget{test-dif-share}{%
+\section{Test of the Impact of Aquiefr Thickness on the Share of Irrigated Acres}\label{test-dif-share}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{C.\arabic{figure}}
+
+Figure \ref{fig:test-sat-share} presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 130m.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6.5in,]{Figures/g_share_dif}
+
+}
+
+\caption{Decline in the share of irrigated acres as a function of aquifer thickness relative to aquifer thickness level of 130m (The shaded areas represent the 95 percent confidence interval)}\label{fig:test-sat-share}
+\end{figure}
 
@@
 
-Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the imapact of \(x\) on \(y\).
+Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the impact of \(x\) on \(y\).
 Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting representation of \(f(x)\).

### 2023-08-01 Codes/Functions/functions.R

@@ -1,48 +1,49 @@
 prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
-  # === saturated thickness threshold in meter ===#
-  # data <- reg_data$data[[1]]
-  # sat_breaks <- reg_data$sat_breaks[[1]]
-
-  # reg_data_y <- reg_data$reg_data_y[[1]]
-
   reg_data <-
     data.table(
       crop_type = c("corn", "soy"),
-      states = list(state_ls[["corn"]], state_ls[["soy"]])
+      states = state_ls
     ) %>%
     rowwise() %>%
-    mutate(base_data = list(
+    mutate(data = list(
       data %>%
-        # === lower bound ===#
+        .[crop == crop_type, ] %>%
+        .[state %in% states, ] %>%
+        #--- yield ---#
+        .[, yield := yield * 12.553 / 200] %>%
+        #--- balance ---#
         .[balance >= balance_thld[1], ] %>%
-        # === upper bound ===#
         .[balance <= balance_thld[2], ] %>%
+        .[, balance := -balance] %>%
         #--- unit conversion ---#
-        .[, sat := measurements::conv_unit(sat, "ft", "m")]
-    )) %>%
-    # === select observations for each crop type ===#
-    mutate(data = list(
-      base_data[crop == crop_type & state %in% states, ]
-    )) %>%
-    mutate(data = list(
-      # === share of irrigated/dryland acres ===#
-      data[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)] %>%
-        # === filter out counties with too low saturated thickness ===#
-        # 0: dryland
-        # If irrigated, sat is set to be at least 0.01.
-        # See 0_6_merge_datasets.R
-        .[sat >= sat_thld_m | sat == 0, ]
+        .[, sat := measurements::conv_unit(sat, "ft", "m")] %>%
+        #--- filter out counties with too low saturated thickness ---#
+        .[sat >= sat_thld_m | sat == 0, ] %>%
+        #--- share of irrigated/dryland acres ---#
+        .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
[…]

### 2023-08-30 Codes/0_2_quickstat_wrangle.rmd
minor updates

```{r}
yields <-
  fread("Shared/Data/RawData/USDA-NASS/yields.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "yield", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, yield_type := case_when(
    str_detect(data_item, "NET") ~ "yield_npa",
    !str_detect(data_item, "NET") ~ "yield_pa"
  )] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, .(sc_code, year, crop, ir, yield, yield_type)] %>%
  dcast(sc_code + crop + year + ir ~ yield_type, value.var = "yield")

harvested_acres <-
  fread("Shared/Data/RawData/USDA-NASS/harvested_acres.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "acres", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, harvested_acres := as.numeric(gsub(",", "", acres))] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(crop, sc_code, year)] %>%
  .[comp == 2, ] %>%
  .[, acres := as.numeric(gsub(",", "", acres))] %>%
  .[, acres_ratio_h := acres / sum(acres), by = .(crop, sc_code, year)] %>%
  .[, .(sc_code, year, crop, ir, harvested_acres, acres_ratio_h)]

planted_acres <-
  fread("Shared/Data/RawData/USDA-NASS/planted_acres.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "acres", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, planted_acres := as.numeric(gsub(",", "", acres))] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(crop, sc_code, year)] %>%
  .[comp == 2, ] %>%
  .[, acres := as.numeric(gsub(",", "", acres))] %>%
  .[, acres_ratio_p := acres / sum(acres), by = .(crop, sc_code, year)] %>%
  .[, .(sc_code, year, crop, ir, planted_acres, acres_ratio_p)]
```

```{r}
nass_data_new <-
  harvested_acres[yields, on = c("sc_code", "ir", "year", "crop")] %>%
  planted_acres[., on = c("sc_code", "ir", "year", "crop")]

nass_data_new <- saveRDS(nass_data_new, "Shared/Data/ProcessedData/nass_data_new.rds")
```

### 2023-08-30 Codes/0_6_merge_datasets.rmd
minor updates

# Combine all the datasets

```{r}
base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))

#' NDMC Drought index
drought_county <- readRDS(here("Shared/Data/ProcessedData/DI_NDMC.rds"))

#' SPEI
# spei_data <- readRDS(here("Shared/Data/ProcessedData/spei_data.rds"))

#' Weather
weather_data <- readRDS(here("Shared/Data/ProcessedData", "summrized_gridMET.rds"))

#' Saturated thickness
sat_all <- readRDS(here("Shared/Data/ProcessedData/sat_all.rds"))

#' Yield, acreage data
nass_data <- readRDS(here("Shared/Data/ProcessedData/nass_data.rds"))

#* SSURGO

ssurgo_soil <- readRDS(here("Shared/Data/ProcessedData/ssurgo_soil.rds"))
```

## combine all

```{r}
merged_data <-
  base_counties[nass_data, on = c("sc_code")] %>%
  weather_data[., on = c("sc_code", "year")] %>%
  sat_all[., on = c("sc_code", "year")] %>%
  drought_county[., on = c("sc_code", "year")] %>%
  .[order(sc_code, year), ] %>%
  ssurgo_soil[, on = "sc_code"]

# merged_data[ir == "nir", state] %>% unique()
# merged_data[ir == "ir", state] %>% unique()
```

# Create and redefine new variables, select observations

```{r}
final_data <-
  merged_data %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  # === if not irrigated, set sat = 0 ===#
  .[ir == "nir", sat := 0] %>%
  # === if sat == 0, but indicated irrigated ===#
  # add 0.1 to distinguish dryland production, where
  # sat is set at 0 above
  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
  # === drop if yield or acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  # === if sat is missing, drop ===#
  #' just 2017 and 2018 data because saturated thickness data
  #' is observed only until 2016
  .[!is.na(sat), ] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_code] %>%
  # === keep only if more than 2 observations are available ===#
  .[num_obs >= 10, ] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ]
```

# Save

```{r}
saveRDS(final_data, here("Shared/Data/ProcessedData/final_data.rds"))
```

### 2023-08-30 Codes/0_2_quickstat.Rmd
minor updates

@@
 title: "Get crop yield and acreage data: Corn and Soybeans"
-author: "Taro Mieno"
-output:
-  html_document:
-    number_sections: yes
-    theme: flatly
-    highlight: zenburn
-    toc_float: yes
-    toc: yes
-    toc_depth: 3
-geometry: margin=1in
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-```
 
@@
 ```{r }
-all_states <- base_counties$state_name %>%
+all_states <-
+  base_counties$state_name %>%
   unique() %>%
@@
     tidyUSDA::allDataItem %>%
-    grep(pattern = data_item, ., value = TRUE) %>%
-    .[!grepl(., pattern = "NET")]
+    grep(pattern = data_item, ., value = TRUE)
 
-  temp_data <- getQuickstat(
+  temp_data <-
+    tidyUSDA::getQuickstat(
       key = key_get("usda_nass_qs_api"),
@@
   c(
-    "CORN, GRAIN, IRRIGATED - YIELD",
-    "CORN, GRAIN, NON-IRRIGATED - YIELD",
+    "CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
     "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
     "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
-    "SOYBEANS, IRRIGATED - YIELD",
-    "SOYBEANS, NON-IRRIGATED - YIELD",
+    "SOYBEANS, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
     "SOYBEANS, IRRIGATED - ACRES HARVESTED",
@@
   )
+
+# tidyUSDA::allDataItem %>%
+#     grep(pattern = "PLANTED", ., value = TRUE) %>%
+#     grep(pattern = "SOYBEANS", ., value = TRUE) %>%
+#     grep(pattern = "IRRIGATED", ., value = TRUE)
+
+# for (i in 1:length(item_list)) {
+#   desc <-
+#     tidyUSDA::allDataItem %>%
+#     grep(pattern = data_item, ., value = TRUE)
+#   print(desc)
+# }
 ```
@@
 ```{r }
-(
 par_list <-
-    expand.grid(
+  data.table::CJ(
     data_item = item_list,
     year = as.character(1985:2018)
-    ) %>%
-    data.table()
   )
@@
 ```{r download-nass-data, eval = F}
-
-# === Download the relevant NASS data ===#
-(
 nass_data <-
@@
       print(x)
-        data <- get_quickstat_data(
-          par_list[x, year],
-          par_list[x, data_item]
+      data <-
+        get_quickstat_data(
+          year = par_list[x, year],
+          data_item = par_list[x, data_item]
         ) %>%
@@
     }
-    ) %>%
-    rbindlist()
   )

### 2023-08-30 Codes/1_Analysis.rmd
minor updates

@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-```
 
@@
     data = final_data,
-    sat_thld_m = 9, # saturated thickness has to be at least 12 meters
-    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
+    #--- saturated thickness has to be at least 9 meters ---#
+    sat_thld_m = 9, 
+    #--- at least 75% of total county area has to be overlapped with HPA ---#
+    ir_share_thld = 0.75, 
+    #--- drop obs with extreme balance ---#
     balance_thld = c(-1200, 200),
+    #--- number of saturated thickness categories ---#
     sat_cat_num = 3,
+    #--- select states of focus ---#
+    # no observations are dropped with the current set of states 
     state_ls =
@@
 
+# main_analysis$share_reg_data[[1]]
+# main_analysis$share_reg_data[[2]]
+
 main_analysis <-
@@
     data %>%
+      .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
+      .[comp == 2, ] %>%
+      .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
       .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
@@
   dplyr::mutate(share_pred_data = list(
-    share_analysis(share_reg_data, sat_seq, sandtotal_med, silttotal_med, awc_med)
+    share_analysis(
+      share_reg_data, 
+      sat_seq, 
+      sandtotal_med, 
+      silttotal_med, 
+      awc_med, 
+      crop
+    )
   )) %>%
@@
     parallel::mclapply(
-      1:1000,
+      1:200,
       \(x) {
@@
           boot_data %>%
+          .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
+          .[comp == 2, ] %>%
+          .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
           .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%

### 2023-08-30 Codes/2_figures_tables.rmd
minor updates

@@
 ```{r}
+# all_results <-  data.table(all_results)
+
 plot_yield_data <-
@@
       coord_cartesian(xlim = c(-200, 1150)) +
-      theme(legend.position = "bottom")
+      theme(
+        legend.position = "bottom",
+        plot.margin = margin(0, 0, 0, 0, "cm"),
+        plot.title = element_blank()
+      )
   ))
@@
         theme_fig +
-        xlab("Water Deficit (mm)") +
+        xlab("") +
         ylab("") +
-        theme_void()
+        # scale_x_continuous(breaks = (-2:11)*100) +
+        coord_cartesian(xlim = c(-200, 1150))
+        # +
+        # # xlim(NA, max_balance) +
+        # theme_void()
     }
@@
   ggtitle("A. Corn") +
-  theme(plot.title = element_text(size = 9, family = "Times", face = "bold"))
+  theme(
+    plot.title = element_text(size = 9, family = "Times", face = "bold"),
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.border = element_blank(),
+    plot.margin = margin(0, 0, 0, 0, "cm")
+  )
 
@@
   ggtitle("B. Soybean") +
-  theme(plot.title = element_text(size = 9, family = "Times", face = "bold"))
+  theme(
+    plot.title = element_text(size = 9, family = "Times", face = "bold"),
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.border = element_blank(),
+    plot.margin = margin(0, 0, 0, 0, "cm")
+    )
 
@@
   )
+
+# g_yield_response <-
+#   cowplot::plot_grid(
+#     g_wd_corn, g_ir_yield_corn,
+#     ncol = 1,
+#     align = "v",
+#     rel_heights = c(0.15, 0.72)
+#   )
+
+# g_yield_response_soy <-
+#   cowplot::plot_grid(
+#     g_wd_soy, g_ir_yield_soy,
+#     ncol = 1,
+#     align = "v",
+#     rel_heights = c(0.15, 0.72)
+#   )
 
@@
           name = "Yield (tonne/ha)",
-          breaks = 0:13
+          breaks = 0:14
         ) +
@@
   xlab("Aquifer Thickness (meter)") +
-  ylab("Estimated Share of Irrigated Acres") +
+  ylab("Decline in Estimated Share of Irrigated Acres") +
   facet_wrap(crop_type ~ ., ncol = 1) +
@@
 g_share_dif <-
-  ggplot(plot_data_ir_share[sat < 130, ]) +
+  ggplot(plot_data_ir_share[sat < 105, ]) +
   geom_line(aes(y = dif_ir_share_hat, x = sat)) +
@@
       facet_grid(. ~ sat_text, scales = "free_y") +
-      ylim(min(data$avg_yield) * 0.7, max(data$avg_yield) + 1) +
+      ylim(min(data$lower) * 0.7, max(data$upper) + 1) +
       xlab("Water Deficit (mm)") +

### 2023-08-30 Codes/test/1_Analysis.rmd
minor updates

@@
 
-# reg_data[, data][[2]] %>% .[, .N, by = sat_cat]
-
 #--- Data used for regression analysis ---#
@@
     data = final_data,
-    sat_thld_m = 9, # saturated thickness has to be at least 12 meters
+    sat_thld_m = 9, # saturated thickness has to be at least 9 meters
     ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
@@
   dplyr::mutate(sat_seq_eval = list(
-    c(10, 50, 90, 130)
+    c(15, 45, 75, 105)
   )) %>%
@@
     data %>%
+      .[, comp := sum(ir == "ir" | ir == "nir"), by = .(sc_code, year)] %>%
+      .[comp == 2, ] %>%
+      .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
       .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
       .[, balance_avg := mean(balance), by = sc_code]
+
   )) %>%
@@
 
+# data <- main_analysis$data[[1]]
+
+# temp <- 
+#   data_boot %>%
+#   .[, comp := sum(ir == "ir" | ir == "nir"), by = .(sc_code, year)] %>%
+#   .[comp == 2, ] %>%
+#   .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
+#   .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
+#   .[, balance_avg := mean(balance), by = sc_code]
+
 all_results <-
@@
           boot_data %>%
+          .[, comp := sum(ir == "ir" | ir == "nir"), by = .(sc_code, year)] %>%
+          .[comp == 2, ] %>%
+          .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
           .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
@@
       },
-      mc.cores = parallel::detectCores() - 2
+      mc.cores = parallel::detectCores() - 2,
+      mc.preschedule = FALSE
     )

### 2023-08-30 Writing/manuscript.Rmd
minor updates

@@
 However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades [@wada2010global; @famiglietti2011satellites; @scanlon2012groundwater; @konikow2015long; @bierkens2019non].
-If not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
+If not addressed, these reductions will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
 
@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs [@mieno2017price; @bhattarai2021impact].
-Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+Second, depletion also reduces the thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
 This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation [@konikow2005groundwater; @foster2014modeling; @hrozencik2017heterogeneous].
-If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted [@hecox2002calculation; @korus2020depletion] - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits [@foster2015well;@rouhi2020downside].
+If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted [@hecox2002calculation; @korus2020depletion] - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits [@foster2015well; @rouhi2020downside].
 The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
+In this paper, we present empirical evidence about how aquifer depletion affects farmers' ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
 Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production [@schlenker2009nonlinear; @lobell2014greater; @lu2018crop].
@@
 Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower saturated thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
+This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower aquifer thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
 
@@
 Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-year water deficit values included in regressions for each crop.", out.width = "6.5in", cache = FALSE}
+The histograms on the top of each panel show the distribution of county-level water deficit values included in regressions for each crop.", out.width = "6.5in", cache = FALSE}
 knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_yield_intensive.pdf"))
@@
 
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greter than 700mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
+While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greater than 700 mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
@@
 These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias [@bound1991extent; @hyslop2001bias], suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk.
-Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquifer thickness.
 Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA.
@@
 Figure \@ref(fig:sat-map) presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
-aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer influences farmers' exposure and responses to drought.
+Aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer thickness influences farmers' exposure and responses to drought.
 

### 2023-08-30 Codes/ExploratoryAnalysis/1_exploratory_analysis.R

@@ -330,3 +330,14 @@ soy_reg <- felm(soy_yield ~
 
 stargazer(soy_reg, type = "text")
 
+
+
+# 
+
+main_analysis$data[[1]]
+
+temp <- main_analysis$share_reg_data[[1]]
+temp[sat > 200, ]
+
+ggplot(data = temp) +
+  geom_point(aes(y = acres_ratio, x = sat))

### 2023-08-30 Codes/ExploratoryAnalysis/analysis_explore.R

@@ -192,3 +192,188 @@ ggplot(soy_data[ir == "ir" & ir_area_ratio >= 0.95, ]) +
 	geom_point(aes(x = sat, y = acres_ratio)) +
 	geom_smooth(aes(x = sat, y = acres_ratio)) 
 
+main_analysis$data[[1]][ir == "ir", acres] %>% hist
+main_analysis$data[[1]][ir == "ir" & sc_code == "31113", acres]
+
+
+#!===========================================================
+#! Share regression
+#!===========================================================
+  ir_share_data <- main_analysis$share_reg_data[[1]]
+  sat_seq <- main_analysis$sat_seq[[1]]
+  sandtotal_e <- main_analysis$sandtotal_med[[1]]
+  silttotal_e <- main_analysis$silttotal_med[[1]]
+  awc_e <- main_analysis$awc_med[[1]]
+
+  #++++++++++++++++++++++++++++++++++++
+  #+ Main
+  #++++++++++++++++++++++++++++++++++++
+
+  ir_share_data <- 
+    ir_share_data[!(sc_code == "31117" & year == 2003), ]
+    # %>%
+    # .[sat < 180, ]
+    #  %>%
+    # .[sc_code != "31113", ]
+
+  #---------------------
+  #- define the formula for the flexible part
+  #---------------------
+  gam_formula_ir_share <-
+    formula(
+      acres_ratio ~
+        s(sat, k = 4, m = 2) +
+        s(balance_avg, k = 4, m = 2)
+    )
+
+  #---------------------
+  #- Run preliminary gam estimation
+  #---------------------
+  #* model data
+  #* gam reg data to replicate
+  reg_data_ir_share <-
+    gen_smooth_data(
+      data = ir_share_data,
+      gam_formula = gam_formula_ir_share
+    )
+
+  #---------------------
+  #- define formula for the fe-logit
+  #---------------------
+  formula_feols <-
+    names(reg_data_ir_share$model_X) %>%
+    paste0(., collapse = "+") %>%
+    paste0("acres_ratio ~ ", .) %>%
+    paste0(., "+ sandtotal_r + silttotal_r + awc_r + I(awc_r^2)") %>%
+    #--- add state-year FE ---#
+    paste0(., " | state_year") %>%
[…]

### 2023-08-30 Codes/Functions/functions.R

@@ -20,7 +20,7 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
         #--- filter out counties with too low saturated thickness ---#
         .[sat >= sat_thld_m | sat == 0, ] %>%
         #--- share of irrigated/dryland acres ---#
-        .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
+        # .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
         .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ] %>%
         .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
         .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
@@ -285,15 +285,18 @@ share_analysis <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e, awc
   #++++++++++++++++++++++++++++++++++++
   #+ Main
   #++++++++++++++++++++++++++++++++++++
+  ir_share_data <- ir_share_data[sat < 180, ]
+
   #---------------------
   #- define the formula for the flexible part
   #---------------------
   gam_formula_ir_share <-
     formula(
       acres_ratio ~
-        s(sat, k = 5, m = 2) +
-        s(balance_avg, k = 5, m = 2)
+        s(sat, k = 4, m = 2) +
+        s(balance_avg, k = 4, m = 2)
     )
+
   #---------------------
   #- Run preliminary gam estimation
   #---------------------
@@ -373,18 +376,27 @@ share_analysis_gam <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e,
   #++++++++++++++++++++++++++++++++++++
   #+ Debug
   #++++++++++++++++++++++++++++++++++++
+  # ir_share_data <- main_analysis$share_reg_data[[1]]
+  # sat_seq <- main_analysis$sat_seq[[1]]
+  # sandtotal_e <- main_analysis$sandtotal_med[[1]]
+  # silttotal_e <- main_analysis$silttotal_med[[1]]
+  # awc_e <- main_analysis$awc_med[[1]]
 
   #++++++++++++++++++++++++++++++++++++
   #+ Main
   #++++++++++++++++++++++++++++++++++++
[…]

### 2023-08-30 Codes/test/test_cut.R

@@ -41,3 +41,16 @@ mutate(sat_breaks = list(
             include.lowest = TRUE
           )
         ] 
+
+
+gam_k3 <- gam(yield ~ s(balance, by = sat_cat, k = 3, m = 2) + factor(sat_cat) +  factor(year) + factor(sc_dry), data = yield_data)
+
+BIC(gam_k3)
+
+gam_k4 <- gam(yield ~ s(balance, by = sat_cat, k = 4, m = 2) + factor(sat_cat) +  factor(year) + factor(sc_dry), data = yield_data)
+
+BIC(gam_k4)
+
+gam_k5 <- gam(yield ~ s(balance, by = sat_cat, k = 10, m = 2) + factor(sat_cat) +  factor(year) + factor(sc_dry), data = yield_data)
+
+BIC(gam_k5)

### 2023-08-30 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewers_R1.tex
added filed from overleaf for syncing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{url}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\section{Reply to reviewer \#1}

This manuscript assesses the impact of aquifer thickness on agricultural
production during drought in the area overlying the High Plains aquifer
in the central United States. The authors use crop, meteorological, and
aquifer thickness data in regressions to develop relationships among
these quantities. Their most important finding is the impact of aquifer
thickness on productivity during drought and the ramifications for
aquifer conservation efforts. I have read this manuscript over a number
of times and found it of considerable interest and worthy of publication
in Nature Water. However, I do have a few concerns that I wanted to
bring to the attention of the authors. My most important concerns are as
follows (not necessarily in order of importance):

\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript. We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Aquifer thickness versus saturated thickness -- The authors speak of
  ``aquifer saturated thickness'', which is a redundant term. Remember, the definition of an aquifer is a saturated unit – there is no unsaturated aquifer. Please use ``aquifer
  thickness'', to avoid propagating this confusing usage into the
  future.
\end{enumerate}

\textcolor{blue}{We agree that there is some redundant use of words here. All the instances of ``saturated thickness'' and ``aquifer saturated thickness'' have been replaced by ``aquifer thickness'' in the revised manuscript.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Dependence on aquifer thickness -- I had a number of questions about
  the relationships displayed in Figure 1. First, the regression appears
  to have been done year by year. Are the quantiles recalculated each
  year to reflect that there may be changes between categories as time
  goes on? Second, when the authors state the lowest quantile goes from
  0 to 33\% are they assuming that the minimum thickness of 9 m is the
  zero point? Third, I assume that the county and year fixed effects are
  calculated as part of the regression. Are there any insights that can
  be gained from the interpretation of those quantities? Fourth, the
  authors need to point out that the same aquifer thickness can result
  in dramatically different transmissive characteristics. Thus, a
  relatively thin aquifer interval composed of coarse sands and gravels
  may be more resilient to drought than a much thicker aquifer dominated
  by fine sands and silts. Could that be a major player in the error
  bars in the plots? Please mention this at some point in the
  manuscript.
\end{enumerate}

\begin{itemize}
\item
  First point:
  \textcolor{blue}{Quantiles are not recalculated each year. If we do so, then the same quantile would represent different aquifer thickness range every year. When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk irrespective of year.}
\item
  Second point:
  \textcolor{blue}{Thank you for highlighting this point. For the 0\% quantile, we assume a aquifer thickness of 9m. This minimum of aquifer thickness was selected based on existing literature (Fenichel et al., 2016; Haacker et al., 2016; Deines et al., 2020), which typically defines 9m as the minimum aquifer thickness needed to provide a functioning water supply. We have modified the text on page 13 lines 298-302 of the revised manuscript to note these points.}
\item
  Third point:
  \textcolor{blue}{Yes, county and year fixed effects are calculated as part of the regressions. However, these variables can be interpreted simply as shifters that tell us how much higher yields in a single county are on average relative to another county and to account for trends in yields over time. As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analyzing these variables - and indeed it is standard practice to not report and directly analyze fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
\item
  Fourth point:
  \textcolor{blue}{This is an important point. You are correct that unobserved factors, such as aquifer heterogeneity influencing well yield responses to changes in aquifer thickness, are a likely underlying cause of the uncertainty ranges found in our analyses. Variations in aquifer conductivity, for example, would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between aquifer thickness and drought risk exposure. This noise is referred to as measurement error in econometric studies, and can result in attenuation bias. Specifically, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it truly is. So, in our context, our estimated impact of aquifer thickness on the share of irrigation is conservative because measurement error will be attenuating some of the true impact of aquifer thickness on drought risk exposure. Furthermore, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquifer thickness. This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought. To highlight these points, we have added the following text in the Discussion section on page 10 lines 213-227 of our revised manuscript:}
\end{itemize}

\textcolor{blue}{"Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk. For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020]. These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001], suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk. Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquiefr thickness. Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA. This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies."}

\textcolor{blue}{New references added:}

\begin{itemize}
\item
  \textcolor{blue}{Butler, J. J., Stotler, R. L., Whittemore, D. O., and Reboulet, E. C. (2013). Interpretation of water level changes in the High Plains aquifer in western Kansas. Groundwater, 51(2), 180-190.}
\item
  \textcolor{blue}{Korus, J. T., and Hensen, H. J. (2020). Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks. Environmental Earth Sciences, 79(16), 382.}
\item
  \textcolor{blue}{John Bound and Alan B Krueger. The extent of measurement error in longitudinal earnings data: Do two wrongs make a right? Journal of labor economics, 9(1):1–24, 1991.}
\item
  \textcolor{blue}{Dean R Hyslop and Guido W Imbens. Bias from classical and other forms of measurement error. Journal of Business \& Economic Statistics, 19(4):475–481, 2001}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Conditioned on drought risk exposure -- I wasn't sure what the authors
  meant by this in the second paragraph of Section 2.2. This must be
  related to the water deficit but the lack of explanation in the main
  text or the methods section left me wondering. Please clarify at some
  point in the manuscript.
\end{enumerate}

\textcolor{blue}{By "conditioned on drought risk exposure", we were referring to the fact that Figure 4 displays the relationship between aquifer thickness and the share of production area that is irrigated while also accounting for variations in climate across the aquifer. I.e. that the Figure shows the impact of aquifer thickness on share of production irrigated after removing the effects of climate on share of production irrigated. To improve clarity, we have modified this phrase to now state "...after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer".}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Equation 2 -- Please define all terms in the equation and provide an
  explanation of what is meant by ``estimated semi-parametrically'\,'.
\end{enumerate}

\textcolor{blue}{We now added a more detailed summary of the mathematical expressions of the model with all the terms explained in Section 4.2 of our revised manuscript. The word semi-parametrically refers to the fact that generalized additive models, such as used in our analysis, estimate the impact of a variable using splines without functional form assumption. These are sometimes referred to as semi-parametric regression methods. However, we agree that the term is potentially confusing, and therefore have removed the term "semi-parametrically" in our revised manuscript.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Kansas and California efforts -- I've seen a number of papers
  describing efforts in Kansas and California to address aquifer
  depletion via LEMAs (Kansas) and SGMA (California). The 1985-2016
  period examined by the authors was before most of the efforts in
  Kansas got started. My understanding is that the first LEMA was
  established in 2013 but then no more were set up until after 2016.
  This manuscript would be more impactful if the authors could speculate
  on how their relationships might change with greater attention to
  groundwater conservation. For example, I would think the decrease in
  yield the authors observe with negative water deficits would disappear
  when the irrigators are using soil moisture sensors and are focused on
  water-use efficiency. I would expect changes as well for tipping
  points.
\end{enumerate}

\textcolor{blue}{This is an interesting idea - thank you for raising it. We agree that efforts to improve aquifer management could help to mitigate some of the changes in productivity driven by climate and aquifer conditions in our analysis. Greater attention to groundwater conservation, and use of improved scheduling technologies such as soil moisture probes, would be expected to mitigate production losses in years with negative water deficits (i.e. wetter years) which we attribute to increased possibility for over-irrigation when well capacities are high (and hence when farmers do not face or observe physical scarcity). In years with large positive water deficits (i.e. drought years), improved irrigation water management would also be expected to mitigate some of the additional production risk caused by smaller aquifer thickness. However, we would still expect smaller aquifer thickness to result in elevated exposure to drought risk even in these circumstances, as declining well yields would create a physical limit to water supply which would stop farmers being able to fully meet crop water needs in these years. Formally testing these effects is beyond the scope and data of our study, but we agree that these are valuable areas for future work and have added the following text to the discussion of our revised manuscript on page 11 lines 233-246:}

\textcolor{blue}{"Improvements to irrigation scheduling practices and technologies, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas (Deines et al., 2019; Glose et al., 2022) and in parts of Texas (Mrad et al., 2020) and which are proposed as part of California's Sustainable Groundwater Management Act (SGMA) (Berbel and Esteban, 2019; Lubell et al., 2020), could also help to improve the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities. Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials (Foster et al., 2019) and the importance of conservation (Marston et al., 2022). However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks (Grafton et al., 2018; Perez-Blanco et al., 2021)."}

\textcolor{blue}{New references added:}

\begin{itemize}
\item
  \textcolor{blue}{Glose, T. J., Zipper, S., Hyndman, D. W., Kendall, A. D., Deines, J. M., and Butler, J. J. (2022). Quantifying the impact of lagged hydrological responses on the effectiveness of groundwater conservation. Water Resources Research, 58(7), e2022WR032295.}
\item
  \textcolor{blue}{Berbel, J., and Esteban, E. (2019). Droughts as a catalyst for water policy change. Analysis of Spain, Australia (MDB), and California. Global Environmental Change, 58, 101969.}
\item
  \textcolor{blue}{Lubell, M., Blomquist, W., and Beutler, L. (2020). Sustainable groundwater management in California: A grand experiment in environmental governance. Society and Natural Resources, 33(12), 1447-1467.}
\item
  \textcolor{blue}{Marston, L. T., Zipper, S., Smith, S. M., Allen, J. J., Butler, J. J., Gautam, S., and David, J. Y. (2022). The importance of fit in groundwater self-governance. Environmental Research Letters, 17(11), 111001.}
\item 
  \textcolor{blue}{Pérez-Blanco, C. D., Loch, A., Ward, F., Perry, C., and Adamson, D. (2021). Agricultural water saving through technologies: a zombie idea. Environmental Research Letters, 16(11), 114032.}
\item 
  \textcolor{blue}{Grafton, R.Q., Williams, J., Perry, C.J., Molle, F., Ringler, C., Steduto, P., Udall, B., Wheeler, S.A., Wang, Y., Garrick, D. and Allen, R.G. (2018). The paradox of irrigation efficiency. Science, 361(6404), 748-750.}
\item 
  \textcolor{blue}{Mrad, A., Katul, G.G., Levia, D.F., Guswa, A.J., Boyer, E.W., Bruen, M., Carlyle-Moses, D.E., Coyte, R., Creed, I.F., Van De Giesen, N. and Grasso, D. (2020). Peak grain forecasts for the US High Plains amid withering waters. Proceedings of the National Academy of Sciences, 117(42), 26145-26150.}
\end{itemize}

\hypertarget{minor-points}{%
\subsection{Minor points}\label{minor-points}}

\begin{itemize}
\tightlist
\item
  Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through aquifer depletion will.'' – confusing sentence – please replace ``through'' with ``by which''
\end{itemize}

\textcolor{blue}{We have corrected the text in line with the reviewer's comment.}

\begin{itemize}
\tightlist
\item
  Section 1, 4th Paragraph --``we generate empirical evidence'' seems awkward to me. How about replacing with ``we present empirical evidence''? Next sentence – ``Our analysis advances on previous $\dots$'' reads better as ``Our analysis extends previous $\dots$'' There are many issues such as these and the previous comment. Please address such issues so that the manuscript is an easier read.
\end{itemize}

\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}

\begin{itemize}
\tightlist
\item
  Section 2.1 title -- I would delete ``marginally'\,' as it is not
  needed and some will find it confusing.
\end{itemize}

\textcolor{blue}{We agree and have removed this term for clarity.}

\begin{itemize}
\tightlist
\item
  Figure 2 - Why do the plots stop at a thickness of about 12 m when you
  had earlier defined the minimum thickness as 9 m and mention 10 m in
  the text when discussing the relationships?
\end{itemize}

\textcolor{blue}{Thank you for catching this typo. The correct value is 9 m and the text has been modified to reflect this and ensure consistency. Plots have also been updated to finish at an aquifer thickness of 9 m where relevant.}

\begin{itemize}
\tightlist
\item
  Section 2.3, First Paragraph -- please clarify that ``average per-area
  production'\,' is a combination of irrigated and rainfed yields. That
  would make the manuscript easier to follow.
\end{itemize}

\textcolor{blue}{We have modified the text in Section 2.3, first paragraph to now more clearly state: "weighted-average of irrigated and rainfed yields, where weights are production area shares".}

\begin{itemize}
\tightlist
\item
  Figure 5 -- I assume that the left plot is corn and the right soybean,
  but please clarify.
\end{itemize}

\textcolor{blue}{Thanks for noticing this. We have added this detail to the figure and its caption.}

\begin{itemize}
\tightlist
\item
  Data and Code Availability -- Are the data from Dr.~Haacker on the
  Github repository? Please place them there prior to publication so
  future investigators can access them rather than having to track down
  Dr.~Haacker. The data set is over seven years old so it should be made
  available according to the principles of open science.
\end{itemize}

\textcolor{blue}{The aquifer thickness data from Dr. Haacker are in the public and can be accessed from the following link: \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}. We have added this information to our data and code availability section of our revised manuscript.}

\begin{itemize}
\tightlist
\item
  Line before Equation (2) -- ``REF'\,' should be replaced with the
  actual reference.
\end{itemize}

\textcolor{blue}{The missing references have been added as follows: "[Foster et al., 2015a,b, Hrozencik et al., 2017]".}

\begin{itemize}
\tightlist
\item
  Discussion, First Paragraph -- ``reduction in aquifer depletion'' should be ``reduction in aquifer thickness''.
\end{itemize}

\textcolor{blue}{Thank you for catching this. This typo has been corrected in the revised manuscript.}

\begin{itemize}
\tightlist
\item
  Section 4.2, 3rd Paragraph -- ``the the impact'\,' -- remember to
  spell and grammar check prior to submission.
\end{itemize}

\textcolor{blue}{Thank you for highlighting this error. We have amended the text and also done a further proofread of our manuscript to capture and address and further spelling and grammatical errors.}

\begin{itemize}
\tightlist
\item
  Reference in Section 3 -- I doubt that ``Butler Jr'' is the surname, please replace with ``Butler''. Suffixes apparently aren't handled well by some of the bibliography generation programs.
\end{itemize}

\textcolor{blue}{The reference has been updated in the revised manuscript}

\begin{itemize}
\tightlist
\item
  Please remember to close all parentheses and don't mispair brackets
  and parentheses. Such stuff can raise questions about the care used in
  the analysis.
\end{itemize}

\textcolor{blue}{All parentheses and brackets have been checked and corrected where necessary in the revised manuscript}

\section{Reply to reviewer \#2}

\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript. We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}

This is an interesting paper that asks an important yet answered
question in the field of agricultural sustainability - how much is
groundwater depletion reducing agricultural productivity in the High
Plains region in the United States? While the paper is an important
contribution to the literature, there are several issues that need to be
addressed prior to acceptance at any journal.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  There needs to be more description about why aquifer saturation
  thickness is an appropriate proxy for farmers' ability to irrigate
  from wells. This is particularly true since this is a broad interest
  journal, and the mechanism for the main independent variable should be
  very clear to all readers.
\end{enumerate}

\textcolor{blue}{Aquifer thickness affects farmers ability to meet the water requirements of their crops due to its impacts on the pumping yields (or capacities) of groundwater irrigation wells. As the thickness of an aquifer declines, the rate at which water can flow from the aquifer to the pumping well is reduced proportionally to the aquifer's transmissivity (equal to the hydraulic conductivity of the aquifer geology multiplied by the saturated thickness of that aquifer geology). For small relative changes in aquifer thickness (e.g. for initial depletion of an aquifer with large initial thickness), the change in well yield is small and unlikely to have substantial negative impacts on drought risk exposure as a farmer could simply run their pump for slightly longer to compensate for the small reduction in instantaneous well yield. However, for larger relative changes in aquifer thickness (e.g. due to significant aquifer depletion and/or thinner initial aquifer bodies), the magnitude of well yield reductions will be much larger. These reductions in water supply capacity cannot be mitigated by changes to irrigation scheduling or management practices. The farmer will often be unable to keep up with crop water demands during the peak of the growing season in the absence of rainfall, even if running their system continually. This effect can also typically not be mitigated by drilling additional or deeper wells, due to a combination of both regulatory moratoriums on well drilling - which are common across the HPA - and high costs of deep well construction. To more clearly highlight for non-specialist readers these connections between saturated thickness, well yields, and drought resilience of agriculture, we have added the following text on page 1 lines 40-47 of our revised manuscript. We note that we are limited in elaborating further on background information given journal word constraints, and hence provide key supporting studies in citations for readers interested in deepening understanding of these dynamics.}

\textcolor{blue}{"...depletion also reduces the thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells). This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation (Konikow and Kendy, 2005; Foster et al., 2015; Hrozencik et al., 2017). If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted (Hecox et al., 2002; Korus and Hensen, 2020) - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits (Foster et al., 2015a; Rouhi-Rad et al., 2020)."
}

\textcolor{blue}{Additional references added:}

\begin{itemize}
\item
  \textcolor{blue}{Hecox, G.R., Macfarlane, P.A., and Wilson B.B. (2002). Calculation of yield for High Plains wells: relationship between saturated thickness and well yield. Open file report 2002–25C, Kansas Geological Survey: Lawrence, Kansas, USA.}
\item
  \textcolor{blue}{Korus, J. T., and Hensen, H. J. (2020). Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks. Environmental Earth Sciences, 79(16), 382.}
\item
  \textcolor{blue}{Rouhi-Rad, M., Araya, A., and Zambreski, Z. T. (2020). Downside risk of aquifer depletion. Irrigation Science, 38, 577-591.}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Though I am not trained as a hydrologist, my understanding is that
  aquifer saturation thickness is a good measure of general water
  availability in an aquifer but not necessarily how much water is
  available at a specific location or point. If this is the case, the
  authors should make clear why this is an appropriate measure to use in
  their study and potential limitations of using this approach.
\end{enumerate}

\textcolor{blue}{Aquifer thickness is a spatially and temporally variable property. It  represents the variation in water storage at specific places in an aquifer and at specific points in time, not the overall aggregated storage within an aquifer. Aquifer thickness also influences the transmissivity of an aquifer, which in turn is proportional to well yields. Aquifer thickness therefore influences the potential rate of water extraction at different points within an aquifer and how this varies over time (e.g., due to depletion reducing aquifer thickness). As noted in our reply to another reviewer, it is important to also note that the relationship between aquifer thickness and well yields will also be influenced by spatial variations in aquifer geology (specifically hydraulic conductivity which together with saturated thickness defines aquifer transmissivity). Data on county-level spatial variations in aquifer hydraulic conductivity is not available for HPA (or indeed for other aquifers in general), meaning we cannot control for these variations thus adding noise or uncertainy to our analyses. This noise is referred to as measurement error in econometric studies, and can result in attenuation bias. Specifically, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it truly is. So, in our context, our estimated impact of aquifer thickness on the share of irrigation is conservative because measurement error in the relationship between aquifer thickness and well yields will be attenuating some of the true impact of aquifer thickness on drought risk. To highlight these points, we have added the following texts in the Discussion section on page 10 lines 210-224 of our revised manuscript.}

\textcolor{blue}{"Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk. For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020]. These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001], suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk. Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness. Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA. This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies."}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  It is unclear to me why county and time fixed effects were not used in
  every regression. In particular, it does not look like they were
  included in the regression that modeled irrigated area as a function
  of saturated thickness and water deficit. Yet, accounting for
  time-invariant differences across counties is important, especially
  because this is the regression that showed the largest effect of
  saturated thickness on a measure of agricultural production (irrigated
  area). Is there not enough variation within a county through time to
  pick up the effect of saturated thickness on irrigated area? This
  should all be made clearer, and a model that better captures the
  causal relationship between saturated thickness and area irrigated
  should be used if possible. If it's not possible, the limitations of
  the current approach should be made very clear and potential issues
  with inference.
\end{enumerate}

\textcolor{blue}{Thank you for this comment, which we believe is an important point to clarify in terms of our methods and the assumptions made in our analysis. County and year fixed effects were included in equation (1), but only year fixed effects were included in equation (2). As you guessed, the problem of including county FEs lies in the fact that this would eliminate a large proportion of the variation in aquifer thickness, which is predominantly cross-sectional rather than longitudinal. To illustrate this, Figure 6 in Appendix A provides histograms of (1) aquifer thickness demeaned by the county average, (2) aquifer thickness demeaned by the state average, and (3) aquifer thickness without demeaning. As shown in Figure 6, inclusion of county FEs (mathematically equivalent to regression using county-demeaned aquifer thickness) dramatically reduces the total variation in aquifer thickness leaving little variations to accurately estimate the impact of aquifer thickness on the share of irrigated acres. The inclusion of State FEs does not diminish the variation in aquifer thickness. Therefore, while the inclusion of State FEs does not provide as tight of a control as county FEs, we have decided to include State FEs for both corn and soybean irrigated share regressions to provide a balance between preserving spatial variations in aquifer thickness and controlling for other spatially dependent effects. Figure 7 compares the estimated impacts of aquifer thickness with or without State FEs. The negative impact of reduction in aquifer thickness on the share of irrigated acres is slightly greater when state fixed effects are included. This suggests that aquifer thickness is unlikely to be very highly correlated with important unobservable variables that are affecting irrigated share significantly. Given the direction of potential bias, providing tighter controls through inclusion of county fixed effects therefore may result in even more significant drop in the share of irrigated acres with smaller aquifer thickness suggesting our results are possibly a conservative estimate of the true impact of aquifer thickness on drought risk. A detailed discussion of these analyses and tradeoffs is now provided in Appendix A.
}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  There are minor typos (e.g., missing reference listed as REF) that
  should be corrected prior to resubmission.
\end{enumerate}

\textcolor{blue}{The missing in-text citation has been corrected, and the revised manuscript has been thoroughly proofread to catch any remaining spelling, grammatical, and formatting errors.}

\section{Reply to reviewer \#3}

The paper aims to show the impacts of aquifer depletion on irrigated
crop yields and production areas in the United States High Plains. To do
so, the paper applies a panel data analysis approach to a detailed
county-level dataset of saturated thickness, water deficit, yields, and
land-use, using a spline model. The authors find that aquifer depletion
reduces farmers' ability to cope with water deficits.

The significance and novelty of the paper stems from the clear and
encompassing empirical approach to the subject matter. The depletion of
the Ogallala aquifer and implications for irrigated agriculture are not
new per se. A compelling quality of the paper is the clear statistical
evaluation of a large panel dataset, constructed to show the negative
impacts of aquifer depletion on farmer's ability to cope with water
deficits, over a long period of time. To fully harness the potential of
this quality for readers, the authors should mainly consider improving
the clarity of the figures presented and of their description in the
text, and addressing some other points that remained unclear (see
details below).

\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript. We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}

\hypertarget{main-points}{%
\section{MAIN POINTS:}\label{main-points}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Abstract:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  ``We show that aquifer depletion reduces the ability of farmers to
  sustain irrigated crop yields and production areas in drought years.''
  - If stated this way, please clarify in the results and/or discussion,
  how causality (direction of causality, no omitted factors) is
  established.
\end{itemize}

\textcolor{blue}{Thank you for raising this important point - we agree that clarifying the extent to which we can infer causality from our analysis is important. We discuss the underlying direction of causality in detail in the main section of our analysis, highlighting the biophysical mechanisms through saturated thickness of an aquifer influences well yields and how well yields will in turn affect drought risk exposure for a farmer. This discussion has been expanded on page 1 lines 37-54 to provide more detail on the physical mechanisms, and supporting empirical and theoretical evidence, connecting saturated thickness, well yields, and water-related crop production risks. For these reasons, we believe that our results and interpretation of our findings are strongly supported by and consistent with hydrologic and economic theories, and provide useful insights for groundwater management even if causality cannot be formally determined.}

\textcolor{blue}{We acknowledge that a further important factor when seeking to establish causality is the issue of possible bias in the estimation of the impact of a variable interest - in our case the effects of saturated thickness on agricultural drought risk. We do not intend to claim that our estimation is free of bias. Indeed, in all empirical studies, it is almost impossible to have no unobserved (omitted) factors unless the data is from a randomized control trial. Other sources of bias (e.g., model misspecification error) will also almost always exist. In our analysis, one of the main limitations is not including county fixed effects in our regressions, which would have controlled all the county-specific time-invariant characteristics. County-level fixed effects were not included because the majority of the variation in saturated thickness is cross-sectional rather than time-dependent, and incorporating county fixed effects would therefore have removed this variation. We acknowledge this as a limitation, and have added a detailed discussion in both the Methods (page 15 lines 339-344) and Appendix C of our revised manuscript.}

\textcolor{blue}{Furthermore, our analysis will also be influenced by measurement errors in aquifer thickness and its relationship with  well yield (e.g., due to spatially variable hydraulic conductivity of aquifer layers). This limitation is highlighted (with expanded text compared to our original manuscript) in the Discussion section (page 10 lines 210-220). Importantly, these measurement errors would be expected to result in attenuation bias, making our estimates of the impact of aquifer thickness smaller in magnitude than the likely true effect.}

\begin{itemize}
\tightlist
\item
  ``\ldots{} in drought years'' - Please define in the in the main text
  (or methods) how drought and drought years are defined. E.g. are any
  effects linked to increasing water deficits considered drought-related
  or only those above a certain level?
\end{itemize}

\textcolor{blue}{In this paper, we do not seek to formally define a threshold level of water deficit beyond which a year is considered to be a `drought'. Instead, we define meteorological conditions during the growing season in terms of a common water deficit metric - which provides a continuous measure of how `drought' a given year was in a specific county considering both precipitation and crop evapotranspiration demands. A water deficit of 1000 mm, for example, could be considered to be a "severe" drought as precipitation is substantially less than rainfall with the result that corn and soybean yields are estimated to be almost zero under rainfed production. In contrast, a water deficit of -200 mm would be considered not to be a drought because precipitation exceeds crop evapotranspiration, and as a result irrigated and rainfed yields are almost identical. We do agree that the wording "in drought years" may be somewhat confusing in this context given the lack of a formal threshold beyond which a year is considered to be drought. Therefore, we have amended this phrase throughout the manuscript to instead say "in years with large water deficits".}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Figure 1: This figure does not convey the effects of saturated
  thickness very clearly, either because the underlying analysis does
  not produce substantial effects, or because they are not presented in
  a compelling manner. It might also be helpful to use a visualization
  that more effectively illustrates the statements in the text. See
  detailed points below:
\end{enumerate}

\textcolor{blue}{We appreciate this comment. In response, we have added an additional figure - Figure 2 in the revised manuscript - which shows in more detail the difference in yield between the 2nd (medium level of saturated thickness) and 3rd (highest level of saturated thickness) quantiles relative to the 1st quantile (lowest level of saturated thickness). This illustrates the differences in irrigated corn and soybean yields across different aquifer thickness levels. Please see our detailed responses to your comments below for more explanation of the interpretation of this figure with regard to distinguishing significant and non-significant results from our analysis.}

\begin{itemize}
\tightlist
\item
  Please distinguish more clearly between significant and
  non-significant results. P. 2 states that ``counties with the highest
  level of saturated thickness \ldots{} experience \ldots{} productivity
  close to rainfed yields''. From the figure it is difficult to tell if
  the 2nd and 3rd quantile are significantly different for soybean. For
  corn this does not seem to be the case. Similarly, the text at the
  bottom of p.~2 states that ``saturated thickness has a statistically
  significant impact on per-area irrigated crop yields in moderate to
  extreme drought conditions''. Please clarify how the significance is
  evaluated here. Which range of deficits is meant by ``moderate to
  extreme''? Are the soybean yields considered significantly different
  between quantiles? Are only the 1st and 3rd quantile considered
  significantly different for corn? - It could even be of interest to
  readers which results are not significant here, in contrast to the
  greater significance of the effects analyzed in Fig. 2 and Fig. 3.
\end{itemize}

\textcolor{blue}{We have broken the above comment into separate points, which we answer  individually below.}

Your comment: Please distinguish more clearly between significant and
non-significant results.

\textcolor{blue}{
Figure 2 in our revised manuscript shows the difference in irrigated crop yields between the 2nd (medium level of saturated thickness) and 3rd (highest level of saturated thickness) quantiles relative to the 1st quantile (lowest level of saturated thickness). The shaded area denotes the 95$\%$ confidence interval around these response curves - indicating whether the difference in irrigated yield between the different levels of saturated thickness is significant at at least the 5$\%$ level. From Figure 2, it is clearly seen that large water deficits - equivalent to a severe drought event as noted above - lead to irrigated corn and soybean yields that are significantly lower for both corn and soybean when saturated thickness is low compared to when saturated thickness is high. The size of the effect is greatest for corn compared to soybean, and when comparing the third quantile (highest saturated thickness) to the first quantile (lowest saturated thickness). As discussed in our original manuscript and expanded upon in the revised manuscript (page 2 lines 86-108), this reflects the larger water requirements and drought sensitivity of corn along with the non-linearities in the response of well yields to saturated thickness.
}

Your comment: P. 2 states that ``counties with the highest level of
saturated thickness \ldots{} experience \ldots{} productivity close to
rainfed yields''. From the figure it is difficult to tell if the 2nd and
3rd quantile are significantly different for soybean. For corn this does
not seem to be the case.

\textcolor{blue}{
You are right that there are only small differences between yields for both corn and soybean when comparing the second and third quantiles of saturated thickness. As noted in the manuscript, this reflects the non-linearities in the relationships between saturated thickness and well yield, which mean that impacts of aquifer depletion on drought risk are not typically felt until saturated thickness has been lowered sufficiently to reduce well yields to levels that impair ability to meet crop water demands. However, in this part of the text, our intention was not to distinguish the 2nd and 3rd quantiles. Instead, we were seeking to point out that - in years with low levels of water deficit (i.e., wetter or non-drought years) - counties with higher aquifer thickness (2nd and 3rd) have lower yields than counties that have the lowest aquifer thickness (1st quantile). Furthermore, yields of counties with higher levels of saturated thickness are comparable to that of rainfed production - in particular for soybean production. To more clearly highlight these points, we have modified the phrasing of the text (page 2 lines 95-108 in the revised manuscript) to now state:}

\textcolor{blue}{"At low water deficits (below 200mm), Figure 2 illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure 1) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile). In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production."}

Your comment: Similarly, the text at the bottom of p.~2 states that
``saturated thickness has a statistically significant impact on per-area
irrigated crop yields in moderate to extreme drought conditions''.
Please clarify how the significance is evaluated here. Which range of
deficits is meant by ``moderate to extreme''?

\textcolor{blue}{We agree that our manuscript would benefit from greater clarity around wording on significance of aquifer depletion impacts on productivity at different levels of drought. In our revised manuscript, we have therefore replace words like "moderate to extreme" with more specific water deficit thresholds/ranges. For the phrase quoted by the reviewer, the text in the revised manuscript (page 2 lines 87-88) has been modified to now state:``in years with water deficit higher than about 750 mm.'' We have also explicitly added the term ``statistically'' when we are talking about statistical significance, as opposed to economic or agronomic significance (i.e. implications) of the results.
}

Your comment: Are the soybean yields considered significantly different
between quantiles? Are only the 1st and 3rd quantile considered
significantly different for corn? - It could even be of interest to
readers which results are not significant here, in contrast to the
greater significance of the effects analyzed in Fig. 2 and Fig. 3.

\textcolor{blue}{As discussed in our responses to earlier comments, Figure 2 and the revised supporting text added to our revised manuscript (page 2 lines 86-94) now provide an in-depth analysis of the statistical significance of yield impacts between different saturated thickness categories. For corn, Figure 2 shows that there are statistically significant differences in yields between the 3rd and 1st quantiles of saturated thickness, and also between the 2nd and 1st quantiles of saturated thickness in the case of extreme water deficits. For soybeans, yields are not significantly different across the three saturated thickness categories - likely due to soybeans having lower water requirements and therefore lower vulnerability to reduced well yield linked to smaller saturated thickness. We have added the following short statement about the insignificance of the impact of saturated thickness for per-area soybean yields on page 2 lines 91-94 of the revised manuscript:}

\textcolor{blue}{"Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups. This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower saturated thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions."}

\begin{itemize}
\tightlist
\item
  In light of point (1.a), please also consider whether additional
  panels showing tests for specific hypotheses or other visual aids
  could make the Fig. 1 more accessible to readers with regards to the
  statements made. The figure is quite crowded, with many confidence
  bands overlapping. Currently, it takes time for readers to parse where
  the confidence bands for some of the plot lines end. The figure could
  be of greater benefit to readers if it was more accessible.
\end{itemize}

\textcolor{blue}{Thank you for this valuable suggestion. As we have discussed above, we have added a new Figure 2, which provides a more detailed comparison of yield responses for the higher levels (2nd and 3rd quantiles) of saturated thickness compared to the lowest levels (1st quantile) of saturated thickness. We believe this change has made it significantly easier for the readers to understand the impact of saturated thickness on farmers ability to buffer crops against water deficits and drought. For Figure 1, it is important to have all the four curves in the same panel as that would allow the readers to compare across both rainfed and the three irrigated yield categories. Confidence intervals around the curves were removed for visual interpretation, taking on board your comments made about difficulty interpreting the original Figure 1. Instead, we have displayed the estimated curves with confidence intervals individually in separate panels in Figure A1 for corn and A2 for soybean of Appendix A.}

\begin{itemize}
\tightlist
\item
  How were confidence bands determined? Were additional tests performed
  to evaluate their validity? Please consider showing individual data
  points.
\end{itemize}

\textcolor{blue}{Confidence bands are derived from the standard error of the coefficients in the model represented by Equation 1. This is a standard approach for determining confidence intervals from a statistically estimated model, and evaluating alternate measures of defining confidence intervals is beyond the scope of our study. Instead, the role of these confidence intervals is to quantify the uncertainty of the point estimates using standard statistical approaches.}

\begin{itemize}
\tightlist
\item
  P. 2 states that ``for water deficits above around 400mm, yield
  changes become negative for counties with the lowest levels of
  saturated thickness (i.e., 1st quantile in Figure 1). This reflects
  the restrictions that lower saturated thickness and associated
  reductions in well yields place on farmers ability to fully meet crop
  water requirements during periods of more severe or extended
  precipitation deficits''. How valid is this interpretation in view of
  the fact that these counties see a similarly large yield decrease
  below 400mm and that counties with the highest levels of saturated
  thickness even see a much stronger increase in yields with increasing
  deficits?
\end{itemize}

\textcolor{blue}{We believe our interpretation is valid, but accept that this is a hypothesis and could require further testing to substantiate the effect that is beyond the scope of our data. We attribute yield reductions with declining saturated thickness at low levels of water deficit (i.e. wetter conditions) to possible over-irrigation by farmer. Over-irrigation in wetter years has been documented empirically by previous research in the High Plains Aquifer (e.g., Foster et al., 2019; Gibson et al., 2017; Gibson et al., 2019), and excess water is known to negatively affect grain yields through a variety of mechanisms including physical damage to crops, nutrient leaching, delays to planting and harvesting, and other factors (e.g., Li et al., 2019). Reductions in yields in this scenario could also reflect the lower availability of positive heat in wetter years, which would diminish the positive benefits of irrigation and possibly explain some of the convergence to rainfed yields. This alone would not explain the fact that yield declines for low water deficits are greatest for counties with the largest aquifer thickness (and by implication greatest potential to over-irrigate). To reflect these different possible interpretations and their uncertainties, we have modified the text on page 2 lines 95-108 of our revised manuscript to now state:}

\textcolor{blue}{"At low water deficits (below 200mm), Figure 2 illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure 1) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile). In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production. These counter-intuitive results reflect a combination of two factors. First, in years with small water deficits there is less need for irrigation and hence irrigated and rainfed yield differences are reduced. Second, years with small or even negative water deficits are associated with wetter conditions during the growing season, which can negatively impact yield if waterlogging occurs leading to greater risks of physical crop damage (e.g., lodging of grains), delays to planting/harvesting, nutrient leaching, and restricted root growth in waterlogged soils (Li et al., 2019). Declines in irrigated yields could be indicative of over-irrigation by farmers, which has been shown to occur in the HPA region in wet years (Foster et al., 2019; Gibson et al., 2017; Gibson et al., 2019), exacerbating these risks. However, the small declines in irrigated yields at low or negative levels of water deficit may also reflect the fact that wetter years are typically cooler, reducing positive heat effects on crop growth and yield formation."}

\textcolor{blue}{New reference added:}

\textcolor{blue}{- Li, Y., Guan, K., Schnitkey, G. D., DeLucia, E., and Peng, B. (2019). Excessive rainfall leads to maize yield loss of a comparable magnitude to extreme drought in the United States. Global Change Biology, 25(7), 2325-2337.}

\begin{itemize}
\tightlist
\item
  Please clarify in the legend or caption that the quantiles indicate
  levels of saturated thickness.
\end{itemize}

\textcolor{blue}{Thank you for this comment. We made changes to all the relevant figures and their captions according to this suggestion.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Figure 2: Please clarify in the text how the significance of
  reductions in the share of irrigated acres is evaluated (e.g.~relative
  to which aquifer depletion level(s)?), and at which level of aquifer
  depletion the reduced irrigation share becomes significant for each
  crop.
\end{enumerate}

\textcolor{blue}{We have added a new figure in Appendix B - Figure B.1 - that presents the estimated difference in the share of irrigated acres at different values of aquifer thickness levels relative to an aquifer thickness of 130 m (equivalent to a very thick aquifer in our third quantile group) including 95\% confidence intervals. This figure shows that the reduction in irrigated acres is statistically significant at the 5\% level for aquifer thickness of below approximately 35 m for corn and 30 m for soybean. Given space constraints imposed for the main body of our manuscript, we decided to place this figure in the appendix as it is supplementary to the main discussion.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Figure 3:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Please clarify why this figure has no confidence bands. The text
  describes it as a combination of the effects shown in Fig. 1 and Fig.
  2, which both have confidence bands. Please clarify how significance
  is evaluated.
\end{itemize}

\textcolor{blue}{
Thank you for highlighting this. To maintain consistency with our response to your earlier comments and valuable points about ensuring clarity of our figures, we have decided to not present confidence intervals for this figure as in the case Figure 1 to avoid overcluttering the plot. Instead, we now have added a new Figure 5, which is similar to Figure 2. It compares average (irrigated and rainfed combined) yield for counties with aquifer thickness of 50 m, 90 m, and 130 m against those with 10m. As in Figure 2, the shaded area denotes the 95\% confidence interval and can be used to interpret significance at the 5\% level. Since these numbers are equidistant, they allow us to illustrate non-linearity in the impact of aquifer thickness. Figure 5 shows that for for 50 m, 90 m, and 130 m aquifer thickness levels, yields are significantly higher than for counties with 10 m of aquifer thickness for water deficits above 400 mm for corn and above around 600 mm for soybean. The magnitude of yield gains is greatest for 90 m and 130 m aquifer thickness, but somewhat lower for aquifer thickness of 50 m. These responses indicate the non-linearities in the response of yields to both water deficits and aquifer conditions, which we have also expanded discussion of in the text in the final paragraph of Section 2.3 on page 7 lines 154-164 of our revised manuscript
}

\begin{itemize}
\tightlist
\item
  Relatedly, the text states that deficits beyond 200mm yield
  significant reductions in corn production. 200mm is where the curves
  intersect. Please clarify if this actually is also where significance
  starts, or if significance starts at a higher deficit.
\end{itemize}

\textcolor{blue}{You are correct that the text in the original manuscript referred to the point at which the curves intersect. The threshold level of water deficit at which saturated thickness has statistically significant impact on the response of per-area corn production is approximately 350 mm. We believe this is now more clearly demonstrated in Figure 5, and have also clarifies this point in the revised text linked to this figure (paragraph 2 of section 2.3 on page 6).}

\begin{itemize}
\tightlist
\item
  Please double-check if the y-axis label is correct. The label (``yield
  (tonne/ha)'') is the same as for Fig. 1, which makes it difficult to
  distinguish the different results presented in Fig. 1 and 3.
\end{itemize}

\textcolor{blue}{Yield (tonne/ha) is correct. Figures 4 (previously 3) and 5 (new figure) both report yield. Figures 1 and 2 report "irrigated" and "rainfed" yields separately, while Figures 4 and 5 report "average" yields reflecting the combination of irrigated and rainfed yields conditional on estimated irrigated production shares.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Discussion:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  P. 6: ``Previous research focused on impacts\ldots{}'' this sentence
  largely repeats a sentence from the introduction, not in the same
  words but in content (and using the same references). Please consider
  removing or modifying the sentence to add new insights for readers.
\end{itemize}

\textcolor{blue}{While we agree that this sentence is similar to one provided in the introduction, we believe it is important to restate this point here in order to provide context when discussing the contribution of our findings to existing literature assessing impacts of drought on agriculture. We therefore have not modified this sentence in our revised manuscript, as the purpose is to reinforce how our analysis differs from the existing state of methods and knowledge in the literature.}

\begin{itemize}
\tightlist
\item
  P. 6: ``\ldots{} demonstrate that the relationship between aquifer
  saturated thickness and drought impacts is highly nonlinear when
  accounting for farmers decisions about irrigated production areas.
  This indicates that sustainable groundwater conservation efforts
  {[}\ldots{]} should be targeted in space and time to avoid depletion
  exceeding critical thresholds \ldots{}'' Is the assumption of
  non-linearity tested? It is difficult to assess the validity of this
  statement based on the figures provided. Please clarify (e.g.~in the
  results) what supports this statement.
\end{itemize}

\textcolor{blue}{We now have statements that clarify this point in the last paragraph of section 2.3 on page 7 of the revised manuscript. Non-linearity in the relationship is visually inferred from examining the shape of the yield and production responses to combinations of weather (represented by water deficit) and aquifer (represented by saturated thickness) conditions. These relationships clearly demonstrate that responses are non-linear - for example increasing curvature of yield responses with higher water deficits, or the variable differences in yields and production when comparing different saturated thickness quantiles. Further formal statistical testing of non-linearity is beyond the scope of our paper given page and word limits, and given that non-linearities can be visually inferred as described above.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Methods - data sets:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  P. 7: ``For counties that do not meet this condition, only rainfed
  yield observations are obtained to avoid the possibility that
  irrigated production may be only partly dependent on use of
  groundwater from the HPA.'' - Please consider whether this assumption
  could bias results and how potential biases can be ruled out.
\end{itemize}

\textcolor{blue}{
This will not bias the main conclusions from our the regression analysis, specifically the finding that lower saturated thickness reduces returns to irrigated crop production. Rainfed yields are included as part of the modeling merely to provide a comparison to the magnitude of yield impacts across saturated thickness categories for irrigated production. However, these rainfed yield data do not have any influence on the identification of irrigated yield response to water deficit for the three saturated thickness categories. Moreover, counties with only rainfed production observations are not included in regressions to predict irrigated area production shares, which focus solely on counties with both irrigated and rainfed production data in the same county. We have also added the following text on page 16 lines 352-355 of our revised manuscript to clarify this point:}

\textcolor{blue}{"Note that irrigated area share and average crop yield estimates use  data only from counties that have both irrigated and rainfed production observations. Counties with solely rainfed production observation data (Section 4.1) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation 1)."}

\begin{itemize}
\tightlist
\item
  P. 9: Saturated thickness data - the number of monitoring wells would
  be of interest for readers to better judge the data, if available.
\end{itemize}

\textcolor{blue}{Aquifer thickness data were derived from interpolation of spatially dense water level observations from the United States Geological Survey's (USGS) national monitoring network. For the years of our analysis, aquifer thickness maps were generated based on between 14,000 and 20,000 annual point water level observations per year, distributed approximately evenly between counties conditional on variations in county area. While these data are a key input to our analysis, a detailed description of how saturated thickness data were generated is beyond the scope of our paper given the word limits imposed by articles in Nature Water. We refer the reader to  Haacker et al., for further information on the generation of saturated thickness data, which includes details about the number of well observations per year and quality control measures applied in production of saturated thickness maps. We have also added the following text to our revised manuscript (page 13 lines 291-293) to provide basic information on the number and distribution of water level observations used to construct aquifer thickness data:}

\textcolor{blue}{"Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA."}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Methods - regression models:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  P. 9: If the model includes fixed effects, how are the single plot
  lines for each category shown in Fig. 1 derived?
\end{itemize}

\textcolor{blue}{
That is because a single county-year was selected in predicting yields. We could pick any county-year without changing any of the conclusions we are drawing. This is because fixed effects are merely a shifter. The relative difference in yield across all the categories is the same no matter what county-year we pick, and the curvature will also be the same. Therefore, presenting lines for all the counties would unnecessarily complicate and crowd the figure without producing any more insights than a single estimated curve for a particular county.
}

\begin{itemize}
\tightlist
\item
  P. 9: Developments over time and independent variables: Please address
  early in the main text which overall trends shape developments in the
  aquifer studied. Please include key references about the case study
  region to inform this, such as Mrad et al.~(2020),
  doi.org/10.1073/pnas.2008383117. In light of this, please discuss the
  appropriateness of the approach. The model accounts for year-specific
  ``shocks'', but not other developments over time (e.g.~lagged effects,
  gradual depletion trends, the introduction of regulation, etc.).
  Relatedly, the analysis only saturated thickness and water deficits to
  explain yields, rather than adding more explanatory variables. The
  advantage of the approach taken here might be the simplicity and
  clarity of assumptions. Readers would benefit from a short explicit
  discussion.
\end{itemize}

\textcolor{blue}{Our analysis considers how aquifer depletion influences the resilience of irrigated crop production to drought risk. As part of this analysis, we explicitly consider trends in aquifer depletion over time through use of time-varying saturated thickness data for counties across the HPA. Consideration of other factors that might cause time variations in the exposure to drought risk are accounted for through the use of year and state fixed effects terms in our analysis. This allows us to isolate the effects of differences in saturated thickness on crop production responses to drought effects, which are the main focus of our paper. Explicit analysis of factors such as introduction of regulations is not feasible with available data, but some of these effects are embedded implicitly in our analysis through use of fixed effects terms. Moreover, our discussion includes detailed text (paragraph 5, section 3) outlining the need for future work to examine how adaptive responses by farmers and regulatory changes may alter relationships between aquifer thickness and drought risks, to which we have also added the suggested reference to support this discussion. Given word constraints and the focus of our paper, we are not in a position to provide a detailed background to the history of groundwater development in the HPA, for which we refer readers to the range of relevant references and supporting texts provided.}

\textcolor{blue}{One important omission from analysis, however, was variations in soil characteristics which might influence relationships between saturated thickness and drought risk at sub-state spatial scales. For example, if an area has sandier soils with lower water holding capacity then the impacts of lower saturated thickness (and therefore lower well yields) may be greater than a county dominated by soils with greater water holding capacities. As previously noted, we cannot include county fixed effects - an alternative way of accounting for soil type variability. This is because the majority of saturated thickness variation in our sample is cross-sectional rather than time-dependent, and hence using county-level fixed effects would remove the majority of saturated thickness variability from our analysis. Therefore, we re-ran our regressions this time including sand percentage, silt percentage, and water holding capacity from SSURGO alongside water deficit and saturated thickness predictors. Inclusion of soil properties did not change the conclusions or significance of our previously reported results.}

\textcolor{blue}{New references added:}

\textcolor{blue}{}

\begin{itemize}
\tightlist
\item
  P. 10: Please provide details on the basis functions in the methods
  section or the supplement.
\end{itemize}

\textcolor{blue}{We have added Appendix D in our revised manuscript, which provides a full overview of the basis functions used in our analysis.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  Supplementary information: Please also consider providing additional
  information to give readers more clarity on the points outlined above
  in the supplementary information.
\end{enumerate}

\textcolor{blue}{Thank you for this suggestion. We have added a set of Appendices, which provide details about (A) the impact of water deficit on irrigated and average yield with confidence intervals, (B) statistical test of the impact of aquifer depletion on the share of irrigated acres relative to 130m aquifer thickness, (C) explanation of the reason for not including county fixed effects and tests of its potential implications for our results, and (D) how spline basis functions are used to estimate the relationship between two variables in a flexible manner without assuming any functional form. We believe that inclusion of these details have made our paper much stronger, and have referenced the appendices above in response to specific comments.}

\hypertarget{minor-points}{%
\subsection{MINOR POINTS:}\label{minor-points}}

\textcolor{blue}{Thank you for noticing there errors. All of these points have been addressed in the revised manuscript.}

\begin{itemize}
\item[$\boxtimes$]
  M1. Please add line numbers.
\item[$\boxtimes$]
  M2. P. 2: Missing closing parenthesis after ``(i.e., 3rd quantile
  \ldots{}''
\item[$\boxtimes$]
  M3. Fig. 1, 2, 3, and 5: Please add panel tags (``a'', ``b'' \ldots)
  to make it easier for readers to understand the captions.
\item[$\boxtimes$]
  M4. Fig. 3: Closing parenthesis of y-axis title missing
\item[$\boxtimes$]
  M5. P. 9: ``WB'' and ``ST'' should not be in italics, unless ``W'',
  ``B'', ``S'', and ``T'' are separate variables
\item[$\boxtimes$]
  M6. P. 9: ``{[}0 - 33.3\(\%\))j'' - the ``j'' does not seem to belong
  here
\item[$\boxtimes$]
  M7. P. 9: ``with sigma\_j represents the category specific intercept''
  should be ``with sigma\_j representing the category-specific
  intercept'' or something similar
\item[$\boxtimes$]
  M8. P. 9: ``under the generalize additive model framework'' should be
  ``generalized'' and ``modeling'' or something similar
\item[$\boxtimes$]
  M9. P. 9: What is ``mu\_t'' in equation 2? Should ``theta\_t'' in
  equation 1 be ``mu\_t''? Because ``theta\_jk'' is already the spline
  basis function.
\item[$\boxtimes$]
  M10. P. 9: One reference seems to be missing (the text includes
  ``(REF)'').
\end{itemize}

\end{document}

### 2023-08-30 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
added filed from overleaf for syncing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{url}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\section{Reply to reviewer \#1}

\textcolor{blue}{Thank you for your feedback on our manuscript. We have addressed all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}

I have read the revised manuscript and the authors’ responses to my initial review. I think they have done a nice job in responding to the points I raised. Thus, I recommend acceptance after the authors address the minor points that I noted while reading the revised manuscript. Those points are as follows:

\begin{itemize}
    \item Line 58 – ``farmers'' should be ``farmers’''.

\textcolor{blue}{Thank you for catching this. Fixed}

\item Figure 1 caption – I found the term ``county-year water deficit values'' confusing. I recommend use of something like ``county-level average water deficit values''. Might also consider modifying line 290.

\textcolor{blue}{We agree. Instead of county-level "average" water deficit, we say county-level annual water deficit because they are not averaged values.}

\item Line 305 – replace ``aquifer'' with ``aquifer thickness'' or ``they''.

\textcolor{blue}{Thank you for catching this. Fixed}

\item Equations 1 and 2 – shouldn’t ``k'' be ``k=1'' – if not, please explain how k is determined. Also, explain ``K'' and how it was determined.

\textcolor{blue}{Please see line 327 for the explanation of what K represents and how K is selected. We added similar explanations for equation 2 at line 341.}

\item Lines 385-387 – Figure A.3 is not presenting results by aquifer thickness category. It is presenting results for three aquifer thicknesses, not categories. Please rewrite.

\textcolor{blue}{Figure A.3 indeed presents what we intended to present. Please note that we made similar changes for average productivity analysis in the main text as shown in Figures 4 and 5. While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable. In our average productivity analysis, we use four aquifer thickness values: 10, 50, 90, and 130 meter. For example, for the analysis of 10 meter, estimated yield response function for the first category is used as it falls within the [9m, 30.1 m] range. Estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres. In our initial submission, we were doing the same kind of analysis (using one number from each aquifer thickness category), but reported by category. In the first round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter). This is in response to another reviewer's point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.}

\item Section B title – revise ``the the''.

\textcolor{blue}{Thank you for catching this. Fixed}

\item Figure B.1 – y-axis label is not correct – this should be ``Decline in Estimated….''.

\textcolor{blue}{Thank you for catching this. Fixed}

\item Lines 398-407 – this paragraph needs to be rewritten. Use of ``demeaning'' or ``demeaned'' is incorrect. My guess is that the authors mean something like ``detrending'' or ``detrended'' – please revise. Also, fix noun-verb disagreement, etc. in this paragraph.

\textcolor{blue}{We understand that demeaning sounds wrong, however, it is a term that is used specifically for the types of operations in Economics.}

\textcolor{red}{Tim, can you work on making the paragraph free of grammatical errors?}

\item Figure C.2 – ``soy'' should be ``soybean'' to be consistent with the rest of the manuscript.

\item Line 428 – is ``a linear in parameter model'' correct? Wouldn’t ``a linear parameter model'' be better?

\textcolor{blue}{We rewrote it to "linear-in-parameter" model. Linear-in-parameter model is a more explicit and correct way of saying what we people call a linear model.}

\end{itemize}

\section{Reply to reviewer \#3}

The manuscript has been considerably improved by the substantial edits in response to the first round reviewer comments. The highly relevant insights of the paper can now be more readily understood by readers. The inclusion of new Figures 2 and 5 and Appendices A and B make it easier to parse the results. Main text additions and Appendices C and D clarify the methods and result interpretation.

Most of the previous comments have been addressed. Some open points, however, still remain and a few new ones have arisen:

\subsection{MAIN COMMENTS}

\begin{itemize}
\item New comment 1 – Clarification about moderate result changes in Figures 1, 3, and 4: Could you explain in more detail how and why the results in these figures have changed? The response to my previous comment (7.b) mentions that the regressions were re-run with sand percentage, silt percentage, and water holding capacity. The variable ``X'' corresponding to these soil characteristics, however, is only included in Equation (2) (irrigated area) and not in Equation (1) (yields).

  \textcolor{blue}{As clearly stated in the manuscript, county fixed effects were included in equation 1 (reperesented by $\alpha_i$). This means that any county-specific time-invariant variables like those in $X$ will be dropped. This does not mean $X$ are not accounted for. They are accounted for as county fixed effects will capture them.}

  \begin{itemize}
  \item (1.a) Does the addition of ``X'' to Equation (2) explain the changed irrigated area results in Figure 3?

  \textcolor{blue}{Yes. Moreover, the results for the share of irrigated acre in this manuscript are different from the previous round due to the changes in the choice of the number of spline basis functions ($K$ and $L$). Previously, we were using 5 for both. However, in responding to reviewer 1's request to explain how the number of spline basis functions are determined, we realized that we were not using the optimal numbers. We tested different numbers earlier in the project and found that the results are not sensitive to the choice of the number of spline basis functions. So, we were simply using reasonable numbers back then. Even though the choice of $K$ and $L$ does not affect the results much, we now pick the numbers based on BIC.}

  \item (1.b) What motivates the other changes made to Equation (2)?

  \textcolor{blue}{The change made other than the ones explained above and including state-year FE is whether we allow the interactions of $WB$ and $AT$. Since we found that there is not significant interactions between them, we have decided to make the model more parsimonious and simply model the impact of $WB$ and $AT$ separately. The results are virtually the same whether we used the old and current models.}

  \item (1.c) What explains the change in the yield results in Figures 1 and 4?

  \textcolor{blue}{We should have mentioned this change in our previous replies. We found that we were actually not grouping observations into three groups in a way that their number of observations are the same as we claimed due to a coding error. We fixed this problem in the previous round. The changes observed in Figure 1 is due solely to this change. For Figure 3, the changes are combination of this change and also the change in the results of the share of irrigated acres as mentioned above. The previous round of Figure 3 of course is different from the one in the initial version as we added another aquifer level to make our claim of the non-linear impact of aquifer much clearer. Figure 3 in the current version is different from the one in the previous round as we changed the number of spline basis function for equation 2 as mentioned above. Overall, the observed changes in results are really minor.}

  \item (1.d) Please add a short reference to the soil characteristics dataset used in Section 4.1 or 4.2 to allow readers can assess the assumptions made without checking the data availability statement.

  \textcolor{blue}{We added the following sentence at line 292.}

  \textcolor{blue}{Soil characteristics data (sand percentage, silt percentage, and water holding capacity) were obtained from the Soil Survey Geographic Dataset (SSURGO). Observation units of the SSURGO data are polygons. For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables.}

  \end{itemize}

\item New comment 2 – Figure 2 results below 200mm: Lines 99 states that quantiles 3 has ``statistically lower yields''. Please clarify this. From visual inspection of Figure 2, the quantile 3 yields do not seem to be significantly below the quantile 1 values (at least for corn). Could you please clarify this? If the result is not statistically significant, does it make sense to discuss it in lines 98-111? Since the paper otherwise applies statistical significance as a criterion for meaningful results, I would recommend being consistent.

    \textcolor{blue}{Thank you for noticing this. We did not realize the difference is no longer statistically significant. Following your suggestion, we simply dropped the pararaph describing the results.}

\item New comment 3 - Figure 4 category changes: Why were the aquifer thickness categories changed in Figure 4? Please also clarify how function estimates for point values of aquifer thickness (rather than value ranges) are derived from the data and explain what motivates using different categories in Figures 1 and 4.

    \textcolor{blue}{We are not using the different categories between Figures 1 and 4. While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable. In our average productivity analysis which combines the two kinds of analysis, we use four aquifer thickness values: 10, 50, 90, and 130 meter. For example, for the analysis of 10 meter, the estimated yield response function to water deficit for the ``first'' category is used as it falls within the [9m, 30.1 m] range. The estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres (equation 2). In our initial submission, we were doing the same of analysis (using one number from each aquifer thickness category), but reported by category instead of reporting aquifer levels we were using for each category. In the first round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter). 50 meter falls in the second category. This is in response to your point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.}

\item New comment 4 – causality and fixed effects: The manuscript has been strengthened by the expansion of the section explaining the reasons for assuming declining aquifer thickness causally reduces the ability to use irrigation to maintain crop yields (lines 40-57). Please clarify the following:

  \begin{itemize}
  \item (4.a) Please clarify for readers if other major causal pathways might be relevant. For example: Are there relevant regulations that limit abstractions in counties where aquifers are too depleted? If so, the causality would be similar, but the implications somewhat different. If not, then stating this would be helpful so readers can rule out potentially strong alternative causal relationships.

  \textcolor{red}{Tim, I don't think this guys understood what we wrote.}

  \item (4.b) Relatedly, the responses to my previous comments 1.a and 7.b indicated that county-level fixed effects were not included in any regressions (only state-level). The response to reviewer \#2 and the Methods section suggest that county-level fixed effects were used in Equation 1. Please clarify the discrepancy and make sure this is clear in the paper.

  \textcolor{blue}{From the initial submission, it has been clearly stated that county fixed effects $\alpha_i$ are included in equation 1. However, we understand your confusion because we stated the following in our response.}

  \textcolor{blue}{``Other sources of bias (e.g., model misspecification error) will also almost always
exist. In our analysis, one of the main limitations is not including county fixed effects in our regression\textcolor{red}{s}, which would have controlled all the county-specific time-invariant characteristics.''}

  \textcolor{blue}{It is our mistake that we added \textcolor{red}{s} to ``regression''. As we stated above, we do include county fixed effect in equation 1. }

  \item (4.c) The following conclusion in Appendix C, lines 410-414 does not seem fully convincing in its present form: The fact that state-level fixed effects are not influential does not immediately imply that county-level fixed effects are ``unlikely to cause significant bias''. There could e.g. easily be relevant county-level effects that average out at the state level.

  \textcolor{blue}{We agree. It is too strong of a claim. We simply tone down our confidence.}

  \textcolor{blue}{We also changed our sentence in line .}

  \end{itemize}

\item New comment 5 – Figure 1:

  \begin{itemize}
   \item (5.a) The histograms are a helpful addition to Figure 1. Please consider adding the total number of observations to the caption to help readers interpret the values. I also recommend adding such histograms to Figures 3 and 4 for consistency. Please clarify why the histogram reaches further than the curves in Figure 1.B and correct this if necessary.

  \item (5.b) Lines 77-78 state ``rainfed yields of both crops decline significantly'' with reference to Figure 1. Please quantify, visualize, or otherwise provide evidence of the significance to readers, or rephrase.
  \end{itemize}

\end{itemize}

\subsection{MINOR COMMENTS}

\begin{itemize}

\item (M1) The response letter indicates that ``all [minor] points have been addressed''. Most minor comments were addressed correctly. The following minor comments from the previous round, however, have not yet been addressed:

\item (M1.a) P. 15: ``WB'' and ``ST'' [now ``AT''] should not be in italics, unless ``W'', ``B'', ``S'', and ``T'' are separate variables

\item (M1.b) Lines 325-326: ``with $\sigma_j$ represents the category specific intercept'' should be ``with $\sigma_j$ representing the category-specific intercept'' or something similar

\item (M1.c) Equation 2: What is ``$\mu_t$'' in equation 2?

\item (M2) Some additional minor points have arisen:

\item (M2.a) Line 315: ``Y'' is capitalized here, but not in equation (1)

\item (M2.b) Equation 2: Please replace the re-used symbol ``alpha'' with a different symbol to avoid confusion.

\item (M2.c) Lines 427-428: Please clarify the grammatical structure of the following sentence: ``Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate $\beta_k$ to fit the data best.''

\end{itemize}

\end{document}

### 2023-08-30 Writing/NatureWater/1stRoundRevision/replies/cover_letter.tex
added filed from overleaf for syncing

--- 2023-07-18 Writing/NatureWater/1stRoundRevision/revised_manuscript/cover_letter.tex
@@
 
-Dear Editor,\\
+Dear Dr. Sjostrand,\\
 
-Thank you for considering this manuscript for publication at Nature Water.
-In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events.
-Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer saturated thickness significantly reduces the overall productivity of both corn and soybean during drought events.
-Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion.
-Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.\\
+Thank you for the opportunity to revise and resubmit our manuscript.
+We have addressed all the comments provided by the reviewers in our new manuscript, and have included detailed responses to each point alongside our revised manuscript.
 
-Our work is significant because we quantify the impact of a decline in saturated thickness of groundwater on the resiliency of agricultural production, which has rarely been quantified in the literature before.
-We confirm quantitatively the importance of keeping relatively high saturated thickness for sustainable high-productivity agricultural production.
-While our results are for the US High Plains, the insights we generated here are likely to hold elsewhere.\\
-
-This article does not have any realated manuscripts that are under consideration.
-However, its pre-print has been published at EarthArXiv (\url{https://eartharxiv.org/repository/view/5126/}).
-We have not had any prior discussions with a Nature Water editor about this work.
-\\
-
-Here is the list of potential reviewers that we may recommend.
+Specifically, in our revised manuscript, we have made the following major changes based on the reviewers' comments, alongside several minor adjustments to wording and to correct spelling and grammatical errors:
 
 \begin{itemize}
-    \item James Butler (\url{https://geo.ku.edu/people/james-k-butler-jr})
-    \item Marc Muller (\url{https://engineering.nd.edu/faculty/marc-francois-muller/})
-    \item Sam Zipper (\url{https://www.samzipper.com/authors/samuel-c.-zipper/})
-    \item Christian Klassert (\url{https://www.ufz.de/index.php?en=38767})
-    \item Jen Burney (\url{http://www.jaburney.net/})
+    \item Updated the main results section to include additional variables (e.g., soil characteristics, state fixed effects) to strengthen how we account for unobservable factors in our regression analyses.
+    \item Added several new figures (Figure 2 and Figure 5) to more clearly highlight the non-linearity and statistical significance of declining saturated thickness on per-area crop yields (Figure 2) and overall crop productivity (Figure 5), which, combined with revised discussion of our results, we believe enable readers to gain deeper understanding of the significance of our findings.
+    \item Conduct rigorous statistical testing of our main findings, in particular to highlight the specific climate and aquifer conditions at which aquifer depletion has statistically robust and significant impacts on drought impacts and risks, including how these thresholds vary between corn and soybean.
+    \item Strengthened the descriptions on the underlying theoretical and empirical evidence for causal links between crop yield, share of irrigated acres, and aquifer thickness, which are one of the main empirical findings from our paper.
+    \item Added a series of four appendices to discuss in more details our methods, robustness checks, limitations imposed by underlying data. 
 \end{itemize}
 
-We do not have any reviewers that we would like to avoid.
+We believe that these changes have substantially strengthen both the methodological approaches in our manuscript, as well the implications of our findings for groundwater conservation and management.
 
+Thank you once for again for consideration of our manuscript, and we would be happy to provide any further information that you may require.
+
+Sincerely,
+
+Taro Mieno (on behalf of the authors)
 \end{document}

### 2023-08-30 Writing/NatureWater/2ndRoundRevision/manuscript.tex
added filed from overleaf for syncing

--- 2023-08-01 Writing/manuscript.tex
@@
 \title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
-\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
+\author{}
+% \author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
 \date{}
@@
 However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}.
-If not addressed, will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
+If not addressed, these reductions will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
 
@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
-Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+Second, depletion also reduces the thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
 This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
-If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well, rouhi2020downside}.
+If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
 The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
+In this paper, we present empirical evidence about how aquifer depletion affects farmers' ability to effectively and reliably buffer crops against drought  using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
 Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater, lu2018crop}.
@@
 Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower saturated thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
+This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower aquifer thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
 
-At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response} experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
+% At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
 In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_yield_intensive}
 
@@
 Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-year water deficit values included in regressions for each crop.}\label{fig:yield-response}
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}\label{fig:yield-response}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_ir_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_ir_yield_dif}
 
@@
 
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greter than 700mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
+While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greater than 700 mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
+Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
 
@@
 As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 130m to 0.61 (61\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
-Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}
+Reductions in the share of production area under irrigation is also evident for soybean although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}.
 
@@
 
-{\centering \includegraphics[width=5in,]{Figures/g_share}
+{\centering \includegraphics[width=5in,height=500px,]{Figures/g_share}
 
@@
 
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95 percent confidence interval)}\label{fig:ir-share}
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:ir-share}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_total_impact}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_total_impact}
 
@@
 \caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix A for estimated average yield response curves with 95 percent confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
+See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_avg_yield_dif}
 
@@
 \caption{Difference in average yield of corn and soybean for aquifer thickness levels of 50 m, 90 m, and 130 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95 percent confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5 percent level of significance.}\label{fig:dif-tot-impact}
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}\label{fig:dif-tot-impact}
 \end{figure}
@@
 These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias \citep{bound1991extent, hyslop2001bias}, suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk.
-Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining saturated thickness.
+Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquifer thickness.
 Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA.
@@
 Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials \citep{foster2019assessing} and the importance of conservation \citep{marston2022}.
-However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks \citep{grafton2018paradox, perez2021agricultural}.
+However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks \citep{grafton2018paradox,perez2021agricultural}.
 Future research should explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
@@
 Our study focuses on the High Plains Aquifer (HPA) in the Central United States.
-Irrigated agriculture is integral to the landscape and economy of the HPA, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{smidt2016complex, fenichel2016measuring}.
+Irrigated agriculture is integral to the landscape and economy of the HPA, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{smidt2016complex,fenichel2016measuring}.
 To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.
@@
 
+Soil characteristics data (sand percentage, silt percentage, and water holding capacity) were obtained from the Soil Survey Geographic
+Dataset (SSURGO). Observation units of the SSURGO data are polygons. For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables.
+
 \begin{figure}[H]
 
-{\centering \includegraphics[height=6in,]{Figures/g_y_wd_all}
+{\centering \includegraphics[width=750px,height=6in,]{Figures/g_y_wd_all}
 
@@
 
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water, haacker2023}.
+As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
 These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935.
@@
 Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
-aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer influences farmers' exposure and responses to drought.
+Aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer thickness influences farmers' exposure and responses to drought.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_map}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_map}
 
@@
 \begin{equation}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
+y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
@@
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3%)
-\), \([33.3-66.7%)
-\), and \([66.7% - 100%]
-\) quantile ranges of aquifer thickness.
+The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0\%-33.3\%)\), \([33.3\%-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness.
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
 Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+$K$ is the total number of spline basis functions.
+$K = 3$ was chosen for this regression as it minimizes BIC.
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} \label{eq:eqn2}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
+$K$ and $L$ are the numbers of basis functions for $WB$ and $AT$, respectively.
+$K = 4$ and $L = 4$ were chosen as they minimize BIC.
 County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by \(X_i\).
@@
 Note that irrigated area share and average crop yield estimates use  data only from counties that have both irrigated and rainfed production observations.
-Counties with solely rainfed production observation data (Section \ref{data-sets}) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \eqref{eq:eqn1}).
+Counties with solely rainfed production observation data (Section \ref{data-sets}) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \ref{eq:eqn1}).
 
@@
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
+Aquifer thickness data were obtained from \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/} \citep{haacker2023}.
 The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data.
@@
 This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-The aquifer thickness data were obtained from \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/} \citep{haacker2023}.
+The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}.
 This data is used for Figure \ref{fig:sat-map} and the regression analysis.
@@
 In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
-This section provides visualization of yield response with the 95\% confidence interval to convey the degree of uncertainty in our estimation.
-Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\% confidence interval for irrigated corn and soybean, respectively.
-Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\% confidence interval.
+This section provides visualization of yield response with the 95 \(\%\) confidence interval to convey the degree of uncertainty in our estimation.
+Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\(\%\) confidence interval for irrigated corn and soybean, respectively.
+Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\(\%\) confidence interval.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_corn}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-corn}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-corn}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_soy}
 
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-soy}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-soy}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_avg_yield}
 
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95 percent confidence interval (grey shaded area)}\label{fig:avg-yield-ind}
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:avg-yield-ind}
 \end{figure}
@@
 \hypertarget{test-dif-share}{%
-\section{Test of the Impact of Aquiefr Thickness on the Share of Irrigated Acres}\label{test-dif-share}}
+\section{Comparison of the impacts of different aquifer thickness levels on irrigated area share}\label{test-dif-share}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{B.\arabic{figure}}
+
+Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 130m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
+Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
+Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 130 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_share_dif}
+
+}
+
+\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 130 m (black line).
+The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
+\end{figure}
+
+\clearpage
+
+\hypertarget{county-fe}{%
+\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
 
@@
 
-Figure \ref{fig:test-sat-share} presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 130m.
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its yearly average (A), state-year average (B), and county average (C).
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left under only year fixed effects (A), state-year fixed effects (B), and county and year fixed effects (C).
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects do not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces its total variation in aquifer thickness resulting in extremely inaccurate estimates (high standard error).
+For this reason, we included state-year fixed effects in irrigated production share regression analysis as a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
 
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_share_dif}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_variation}
 
@@
 
-\caption{Decline in the share of irrigated acres as a function of aquifer thickness relative to aquifer thickness level of 130m (The shaded areas represent the 95 percent confidence interval)}\label{fig:test-sat-share}
-\end{figure}
-
-\hypertarget{county-fe}{%
-\section{County Fixed Effects in Irrigation Share Regression}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its overall average (A), state average (B), and county average (C) to gain insights into how much variation for identification is left under no county or state fixed effects (A), state fixed effects (B), and county fixed effects (C).
-As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation.
-Consequently, we found that including county fixed effects results in extremely inaccurate estimates (high standard error).
-For this reason, we just included state fixed effects.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,]{Figures/g_variation}
-
-}
-
-\caption{Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.}\label{fig:variation-left}
+\caption{Comparison of variations in aquifer thickness after demeaning by yearly average, state-year average, and county average.}\label{fig:variation-left}
 \end{figure}
@@
 For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
-Importantly, adding state fixed effects, which control for unobservables at the state level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
 This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_share_comp}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_share_comp}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
+\renewcommand{\thefigure}{D.\arabic{figure}}
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_basis}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_basis}
 
@@
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
-Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate \(\beta_k\) to fit the data best.
+Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate \(\beta_k\) to fit the data best.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_il_y}
+{\centering \includegraphics[width=6in,height=500px,]{Figures/g_il_y}
 

### 2023-08-30 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
updates on the manuscript

@@
 
+\section{Changes unrelated to reviewers' comments}
+
+\textcolor{blue}{We made a small correction to our codes.
+In the previous round, we included counties with only irrigated acres for the irrigation share regression.
+For such counties, the share of irrigated acres was calculated to be 1.
+Therefore, we dropped them.
+While it resulted in some minor changes in our results from those provided in the previous round, all the qualitative results and conclusions are the same as the previous rounds.}
+
 \section{Reply to reviewer \#1}
-
-\textcolor{blue}{Thank you for your feedback on our manuscript.
-We have addressed all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black).}
 
@@
 
+\textcolor{blue}{Thank you for your feedback on our manuscript.
+We have addressed all your comments and suggestions, and our replies to your comments can be found in blue below (original comments in black).
+Please take a look at our comments about a small change in our codes at the top as well.}
+
 \begin{itemize}
@@
 
-\textcolor{blue}{Thank you for catching this. Fixed}
+\textcolor{blue}{Thank you for catching this. Fixed.}
 
@@
 While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable.
-In our average productivity analysis, we use four aquifer thickness values: 10, 50, 90, and 130 meter.
+In our average productivity analysis, we use four aquifer thickness values: 10, 40, 70, and 100 meter.
 For example, for the analysis of 10 meter, estimated yield response function for the first category is used as it falls within the [9m, 30.1 m] range.
@@
 In our initial submission, we were doing the same kind of analysis (using one number from each aquifer thickness category), but reported by category.
-In the first round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter).
-This is in response to another reviewer's point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.}
+In the previous round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter).
+This is in response to another reviewer's point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.
+Given the chosen series of aquifer thickness numbers are equi-distance, it has become much easier to observe the non-linearity in the impact of saturated thickness on overall productivity.
+Please note that we now present results for 10, 40, 70, and 100 meters, which are in the first, second, third, and third saturated thickness categories.
+}
 
@@
 
-\textcolor{blue}{We understand that demeaning sounds wrong, however, it is a term that is used specifically for the types of operations in Economics.}
+\textcolor{blue}{We understand that demeaning sounds wrong, however, it is a term that is used specifically for the types of operations we are conducting in that analysis in Economics.}
 
@@
 Most of the previous comments have been addressed. Some open points, however, still remain and a few new ones have arisen:
+
+\textcolor{blue}{Thank you for your feedback on our manuscript.
+We have addressed all your comments and suggestions, and our replies to your comments can be found in blue below (original comments in black).
+Please take a look at our comments about a small change in our codes at the top as well.}
 

### 2023-08-30 Codes/1_Analysis.rmd
updates on the manuscript

@@
   dplyr::mutate(sat_seq_eval = list(
-    c(10, 50, 90, 130)
+    c(10, 40, 70, 100)
   )) %>%
@@
   dplyr::mutate(share_pred_data = list(
-    share_analysis(
+    share_analysis_gam(
       share_reg_data,
@@
       silttotal_med,
-      awc_med, 
-      crop
+      awc_med
     )
@@
     parallel::mclapply(
-      1:200,
+      1:1000,
       \(x) {
@@
     share_pred_data_boot[, .(
-      ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE)
+      ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE),
+      dif_ir_share_hat_se = sd(dif_ir_share_hat, na.rm = TRUE)
     ),
@@
         lower = ir_share_hat - 1.96 * ir_share_hat_se,
-        upper = ir_share_hat + 1.96 * ir_share_hat_se
+        upper = ir_share_hat + 1.96 * ir_share_hat_se,
+        dif_lower = dif_ir_share_hat - 1.96 * dif_ir_share_hat_se,
+        dif_upper = dif_ir_share_hat + 1.96 * dif_ir_share_hat_se
       )]

### 2023-08-30 Codes/2_figures_tables.rmd
updates on the manuscript

@@
 coeff_soy <- 200
+
 g_yield_wd_soy <-
@@
 
-## Saturated thickness
-
-```{r, eval = FALSE}
-g_sat_dist <-
-  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
-  .[sat != 0, ] %>%
-  ggplot(data = .) +
-  geom_histogram(aes(x = sat), fill = "white", color = "blue") +
-  facet_grid(type ~ ., scale = "free_y") +
-  ylab("Count") +
-  xlab("Aquifer Thickness (meter)") +
-  theme_fig
-```
-
-## Map
+## Saturated Thickness Map
 
@@
   tigris::counties() %>%
+  st_simplify(dTolerance = 100) %>%
   rename(
@@
 counties_corn_ir <-
-  data_corn %>%
+  copy(data_corn) %>%
   .[sat_cat != "dryland", ] %>%
@@
   left_join(all_counties, by = "sc_code") %>%
+  data.table() %>%
   .[, type := "Corn"] %>%
@@
 counties_soy_ir <-
-  data_soy %>%
+  copy(data_soy) %>%
   .[sat_cat != "dryland", ] %>%
@@
   left_join(all_counties, by = "sc_code") %>%
+  data.table() %>%
   .[, type := "Soybean"] %>%
@@
 
-ggsave(file = "GitControlled/Writing/Figures/g_map.png", g_map, height = 7, width = 6, dpi = 600)
 ggsave(file = "GitControlled/Writing/Figures/g_map.pdf", g_map, height = 7, width = 6)
@@
     panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
     panel.border = element_blank(),
-    plot.margin = margin(0, 0, 0, 0, "cm")
+    plot.margin = margin(0, 0, -0.5, 0, "cm")
   )
@@
     panel.border = element_blank(),
-    plot.margin = margin(0, 0, 0, 0, "cm")
+    plot.margin = margin(0, 0, -0.5, 0, "cm")
   )
@@
     align = "v",
-    rel_heights = c(0.15 / 2, 0.72 / 2, 0.15 / 2, 0.98 / 2)
+    rel_heights = c(0.18 / 2, 0.69 / 2, 0.18 / 2, 0.95 / 2)
   )
@@
 ```{r}
+#++++++++++++++++++++++++++++++++++++
+#+ Share response
+#++++++++++++++++++++++++++++++++++++
+plot_data_ir_share <-
+  all_results[, .(crop, share_pred_data)] %>%
+  unnest(cols = share_pred_data) %>%
+  data.table() %>%
+  .[, crop_type := case_when(
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
+  )]
+
+g_share <-
+  plot_data_ir_share %>%
+  nest_by(crop) %>%
+  dplyr::mutate(g_fig = list(
+    ggplot(dplyr::filter(data, sat <= 150)) +
+      geom_line(aes(y = ir_share_hat, x = sat)) +
+      geom_ribbon(
+        aes(
+          ymin = ir_share_hat - 1.96 * ir_share_hat_se,
+          ymax = ir_share_hat + 1.96 * ir_share_hat_se,
+          x = sat
+        ),
+        alpha = 0.4
+      ) +
+      scale_y_continuous(breaks = seq(0, 1, by = 0.1), limits = c(0.4, 1)) +
+      scale_x_continuous(breaks = seq(10, 150, by = 10)) +
+      coord_cartesian(xlim = c(9, 150)) +
+      xlab("Aquifer Thickness (meter)") +
+      ylab("Share of Irrigated Acres") +
+      theme_fig +
+      theme(
+        legend.position = "bottom"
+      )
+  ))
+
+#++++++++++++++++++++++++++++++++++++
+#+ Saturated thickness histogram
+#++++++++++++++++++++++++++++++++++++
+g_sat_hist <-
+  all_results[, .(crop, data)] %>%
+  dplyr::mutate(g_fig = purrr::map(
+    data,
+    function(data) {
+      ggplot(data[sat < 150 & ir == "ir", ]) +
+        geom_histogram(aes(x = sat), bins = 30, col = "darkgreen", fill = NA) +
+        theme_fig +
+        xlab("") +
+        ylab("") +
+        coord_cartesian(xlim = c(9, 150)) +
+        theme(
+          plot.title = element_text(size = 9, family = "Times", face = "bold"),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.ticks.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank(),
+          plot.margin = margin(0, 0, -0.5, 0, "cm")
+        )
+    }
+  ))
+
+g_ir_share <-
+  cowplot::plot_grid(
+    g_sat_hist$g_fig[[1]] + ggtitle("A. Corn"),
+    g_share$g_fig[[1]] + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
+    g_sat_hist$g_fig[[2]] + ggtitle("B. Soybean"),
+    g_share$g_fig[[2]],
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.18 / 2, 0.64 / 2, 0.18 / 2, 0.64 / 2)
+  )
+
+ggsave(
+  file = "GitControlled/Writing/Figures/g_share.pdf",
+  g_ir_share,
+  height = 6,
+  width = 6
+)
+```
+
+# Difference in the share of irrigated production (from 100)
+
+```{r}
 plot_data_ir_share <-
@@
 
-g_share <-
-  ggplot(plot_data_ir_share[sat <= 150, ]) +
-  geom_line(aes(y = ir_share_hat, x = sat)) +
-  geom_ribbon(
-    aes(
-      ymin = lower,
-      ymax = upper,
-      x = sat
-    ),
-    alpha = 0.4
-  ) +
-  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
-  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
-  xlab("Aquifer Thickness (meter)") +
-  ylab("Decline in Estimated Share of Irrigated Acres") +
-  facet_wrap(crop_type ~ ., ncol = 1) +
-  theme_fig +
-  theme(
-    legend.position = "bottom",
-    strip.background = element_blank(),
-    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
-  )
-
-ggsave(
-  file = "GitControlled/Writing/Figures/g_share.pdf",
-  g_share,
-  height = 4,
-  width = 5
-)
-```
-
-# Difference in the share of irrigated production (from 130)
-
-```{r}
 g_share_dif <-
-  ggplot(plot_data_ir_share[sat < 105, ]) +
+  ggplot(plot_data_ir_share[sat < 100, ]) +
   geom_line(aes(y = dif_ir_share_hat, x = sat)) +
@@
       ) +
-      scale_x_continuous(name = "Water Deficit (mm)", breaks = (-2:11) * 100, limits = c(-200, 1200)) +
+      scale_x_continuous(
+        name = "Water Deficit (mm)",
+        breaks = (-2:12) * 100
+      ) +
+      coord_cartesian(xlim = c(-200, 1150)) +
       theme(legend.position = "bottom")
@@
   g_total_both$g_fig[[1]] +
-  theme(
-    legend.position = "none",
-    axis.title.x = element_blank()
-  ) +
-  ylim(6, 13) +
-  ggtitle("A. Corn")
+  theme(legend.position = "none") +
+  ylim(6, 13)
 
@@
   g_total_both$g_fig[[2]] +
-  ylim(1.5, 4) +
-  ggtitle("B. Soybean")
+  ylim(1.5, 4)
 
-g_total <- g_total_corn / g_total_soy
+g_wds <-
+  all_results[, .(crop, data)] %>%
+  dplyr::mutate(g_fig = purrr::map(
+    data,
+    function(data) {
+      ggplot(data) +
+        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
+        theme_fig +
+        xlab("") +
+        ylab("") +
+        coord_cartesian(xlim = c(-200, 1150)) +
+        theme(
+          plot.title = element_text(size = 9, family = "Times", face = "bold"),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.ticks.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank(),
+          plot.margin = margin(0, 0, -0.5, 0, "cm")
+        )
+    }
+  ))
+
+g_total_impact <-
+  cowplot::plot_grid(
+    g_wds$g_fig[[1]] + ggtitle("A. Corn"),
+    g_total_corn + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
+    g_wds$g_fig[[2]] + ggtitle("B. Soybean"),
+    g_total_soy,
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.18 / 2, 0.65 / 2, 0.18 / 2, 0.8 / 2)
+  )
 
@@
   file = "GitControlled/Writing/Figures/g_total_impact.pdf",
-  g_total,
+  g_total_impact,
   height = 6,

### 2023-08-30 Codes/A2_state_fe.rmd
updates on the manuscript

@@
 formula_text <- "acres_ratio ~
-    s(sat, k = 5, m = 2) + s(balance_avg, k = 5, m = 2) +
+    s(sat, k = 4, m = 2) + s(balance_avg, k = 4, m = 2) +
     sandtotal_r + silttotal_r + awc_r +
@@
 
+# ggplot(results$ir_share_data[[1]]) +
+#   geom_point(aes(y = acres_ratio, x = sat))
+
 results <-
@@
   dplyr::mutate(ir_share_data = list(
-    data[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-      .[, balance_avg := mean(balance), by = sc_code]
+    data %>%
+      .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
+      .[comp == 2, ] %>%
+      .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
+      .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
+      .[, balance_avg := mean(balance), by = sc_code] %>%
+      .[sat < 180, ]
   )) %>%
@@
       formula,
-      data = ir_share_data_corn,
+      data = ir_share_data,
       family = binomial
@@
 g_share_comp <-
-  results[sat <= 500, ] %>%
+  results[sat <= 150, ] %>%
+  .[, crop := case_when(
+    crop == "corn" ~ "Corn",
+    crop == "soy" ~ "Soybean"
+  )] %>%
   .[, type := factor(type, levels = c("Year FE", "State and Year FE", "State-Year FE"))] %>%
@@
   ) +
-  ylim(0.25, 1) +
-  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
-  scale_x_continuous(breaks = seq(10, 160, by = 20)) +
+  scale_y_continuous(breaks = seq(0, 1, by = 0.1), limits = c(0.4, 1)) +
+  scale_x_continuous(breaks = seq(10, 150, by = 20)) +
   xlab("Saturated Thickness (meter)") +

### 2023-08-30 Writing/NatureWater/2ndRoundRevision/manuscript.tex
updates on the manuscript

@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_yield_intensive}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_ir_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_ir_yield_dif}
 
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 130m to 0.61 (61\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100m to 0.72 (72\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}.
@@
 
-{\centering \includegraphics[width=5in,height=500px,]{Figures/g_share}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_total_impact}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_total_impact}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_avg_yield_dif}
 
@@
 
-{\centering \includegraphics[width=750px,height=6in,]{Figures/g_y_wd_all}
+{\centering \includegraphics[width=750px,height=6in,]{../../Figures/g_y_wd_all}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_map}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_map}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_corn}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_soy}
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_avg_yield}
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_share_dif}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_share_dif}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_variation}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_variation}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_share_comp}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share_comp}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_basis}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_basis}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{Figures/g_il_y}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_il_y}
 

### 2023-08-30 Codes/test/functions.R

@@ -286,8 +286,8 @@ share_analysis <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e, awc
   gam_formula_ir_share <-
     formula(
       acres_ratio ~
-        s(sat, k = 5, m = 2) +
-        s(balance_avg, k = 5, m = 2)
+        s(sat, k = 4, m = 2) +
+        s(balance_avg, k = 4, m = 2)
     )
   #---------------------
   #- Run preliminary gam estimation

### 2023-08-31 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
manuscripts updated and replies to reviewers created to be reviewed

@@
 
-\textcolor{blue}{We made a small correction to our codes.
+\textcolor{blue}{We made a small correction to our codes and data.
 In the previous round, we included counties with only irrigated acres for the irrigation share regression.
@@
 
-I have read the revised manuscript and the authors’ responses to my initial review.
+I have read the revised manuscript and the authors' responses to my initial review.
 I think they have done a nice job in responding to the points I raised.
@@
 \begin{itemize}
-    \item Line 58 – ``farmers'' should be ``farmers’''.
+    \item Line 58 – ``farmers'' should be ``farmers'''.
 
@@
 
-\item Equations 1 and 2 – shouldn’t ``k'' be ``k=1'' – if not, please explain how k is determined. Also, explain ``K'' and how it was determined.
+\item Equations 1 and 2 – shouldn't ``k'' be ``k=1'' – if not, please explain how k is determined. Also, explain ``K'' and how it was determined.
 
@@
 While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable.
-In our average productivity analysis, we use four aquifer thickness values: 10, 40, 70, and 100 meter.
+In our average productivity analysis that combines both analysis results, we use four aquifer thickness values: 10, 40, 70, and 100 meter.
 For example, for the analysis of 10 meter, estimated yield response function for the first category is used as it falls within the [9m, 30.1 m] range.
 Estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres.
+Please note that it is ``necessary'' to use point values for this analysis.
+The estimated model of the share of irrigated acres do not accept an aquifer thickness category or range.
+It accepts a single value of aquifer thickness and returns an estimate of the share of irrigated acres by design.}
+
+\textcolor{blue}{
 In our initial submission, we were doing the same kind of analysis (using one number from each aquifer thickness category), but reported by category.
@@
 Given the chosen series of aquifer thickness numbers are equi-distance, it has become much easier to observe the non-linearity in the impact of saturated thickness on overall productivity.
-Please note that we now present results for 10, 40, 70, and 100 meters, which are in the first, second, third, and third saturated thickness categories.
-}
+Please note that we now present results for 10, 40, 70, and 100 meters, which are in the first, second, third, and third saturated thickness categories, respectively.
+We made this change because 130 meters is really high and rarely observed, and we wanted the range of analysis to be more representative of the region.}
 
@@
 
-\item Line 428 – is ``a linear in parameter model'' correct? Wouldn’t ``a linear parameter model'' be better?
+\textcolor{blue}{Thank you for catching this. Fixed.}
 
-\textcolor{blue}{We rewrote it to "linear-in-parameter" model. Linear-in-parameter model is a more explicit and correct way of saying what we people call a linear model.}
+\item Line 428 – is ``a linear in parameter model'' correct? Wouldn't ``a linear parameter model'' be better?
+
+\textcolor{blue}{We rewrote it to the following:}
+
+\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate $\beta_k$ to fit the data best.''}
+
+\textcolor{blue}{In statistics, ``linear'' model means that the model is linear in parameter, but not necessarily in independent variable.
+For example, the model below is linear in parameter, but not linear in $x$.}
+
+\begin{align}
+  y = \beta x^2 + v
+\end{align}
+
+\textcolor{blue}{ GAM is linear in parameter, but not linear in independent variable. We wanted to emphasize the distinction and used the term linear-in-parameter model instead of linear model.}
 
@@
 
-  \textcolor{blue}{As clearly stated in the manuscript, county fixed effects were included in equation 1 (reperesented by $\alpha_i$).
+  \textcolor{blue}{As stated in the manuscript, county fixed effects were included in equation 1 (represented by $\alpha_i$).
 This means that any county-specific time-invariant variables like those in $X$ will be dropped.
@@
   \textcolor{blue}{Yes.
-Moreover, the results for the share of irrigated acre in this manuscript are different from the previous round due to the changes in the choice of the number of spline basis functions ($K$ and $L$).
-Previously, we were using 5 for both.
-However, in responding to reviewer 1's request to explain how the number of spline basis functions are determined, we realized that we were not using the optimal numbers.
-We tested different numbers earlier in the project and found that the results are not sensitive to the choice of the number of spline basis functions.
-So, we were simply using reasonable numbers back then.
-Even though the choice of $K$ and $L$ does not affect the results much, we now pick the numbers based on BIC.}
+Please note that the results for the share of irrigated acre in this manuscript are different from the previous round as well due to the changes we mentioned at the top of this document.}
 
@@
 
-  \textcolor{blue}{The change made other than the ones explained above and including state-year FE is whether we allow the interactions of $WB$ and $AT$.
-Since we found that there is not significant interactions between them, we have decided to make the model more parsimonious and simply model the impact of $WB$ and $AT$ separately.
+  \textcolor{blue}{The change made other than the ones explained above and including state-year FE (which is provides stronger controls than just including year FE) is whether we allow the interactions of $WB$ and $AT$.
+Since we found that there is no significant interactions between them, we have decided to make the model more parsimonious and simply model the impact of $WB$ and $AT$ individually.
 The results are virtually the same whether we used the old and current models.}
@@
   \textcolor{blue}{We should have mentioned this change in our previous replies.
-We found that we were actually not grouping observations into three groups in a way that their number of observations are the same as we claimed due to a coding error.
-We fixed this problem in the previous round.
-The changes observed in Figure 1 is due solely to this change.
-For Figure 3, the changes are combination of this change and also the change in the results of the share of irrigated acres as mentioned above.
-The previous round of Figure 3 of course is different from the one in the initial version as we added another aquifer level to make our claim of the non-linear impact of aquifer much clearer.
-Figure 3 in the current version is different from the one in the previous round as we changed the number of spline basis function for equation 2 as mentioned above.
-Overall, the observed changes in results are really minor.}
+We found that we were actually not grouping observations into three groups in a way that their number of observations are the same as we claimed due to a small coding error.
+This problem was fixed in the previous round.
+The changes observed in Figure 1 in the previous round is due solely to this change.
+For Figure 4 of the previous version (1st round revision), the changes are combination of this change and also the change in the results of the share of irrigated acres as mentioned above.
+This is because Figure 4 presents the results of the analysis that combines what are presented in Figures 1 and Figure 3 of the previous version (1st round revision).
+Figure 3 in the 1st round and Figure 4 in the previous round are different because as we added another aquifer level to make our claim of the non-linear impact of aquifer much clearer in response to one of your comments in the previous round.
+Further, Figure 4 in the current version is different from the one in the previous round as we changed the aquifer levels for the analysis.
+Instead of 10, 50, 90, and 130 meters, we use 10, 40, 70, and 100 meters.
+This is because 130 meters is really high and rarely observed, and we wanted the range of analysis to be more representative of the region.
+Please see our reply to New  comment 3 for further details on this.}
 
@@
 
-    \textcolor{blue}{We are not using the different categories between Figures 1 and 4.
+    \textcolor{blue}{(Please note that the description in this paragraph is about the \textcolor{red}{previous} rounds, but not the current round.) In the previous round, we were not using the different categories between Figures 1 and 4.
 While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable.
-In our average productivity analysis which combines the two kinds of analysis, we use four aquifer thickness values: 10, 50, 90, and 130 meter.
-For example, for the analysis of 10 meter, the estimated yield response function to water deficit for the ``first'' category is used as it falls within the [9m, 30.1 m] range.
-The estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres (equation 2).
-In our initial submission, we were doing the same of analysis (using one number from each aquifer thickness category), but reported by category instead of reporting aquifer levels we were using for each category.
-In the first round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter).
-50 meter falls in the second category.
+In our average productivity analysis which combines the two kinds of analysis, we used four aquifer thickness values: 10, 50, 90, and 130 meters in the previous round.
+For example, for the analysis of 10 meters, the estimated yield response function to water deficit for the ``first'' category is used as it falls within the [9m, 30.1 m] range.
+The estimated irrigation share for 10 meters can be obtained simply by plugging 10 into the trained model of the share of irrigated acres (equation 2).
+Please note that it is ``necessary'' to use point values for this analysis.
+The estimated model of the share of irrigated acres does not accept an aquifer thickness category or range.
+It accepts a single value of aquifer thickness and returns an estimate of the share of irrigated acres by design.
+Indeed, in our initial submission, we were doing the same of analysis (using one number from each aquifer thickness category), but reported by category instead of reporting aquifer levels we were using for each of the point values.
+Specifically, we were using 9.144, 30.48, and 60.96 meters as the point values for the first, second, and third aquifer thickness categories, respectively.
+In the first round of revision, we decided to have two aquifer thickness values in the 3rd aquifer thickness category (90 and 130 meter).
 This is in response to your point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.}
+
+    \textcolor{blue}{(Please note that the description in this paragraph is about the \textcolor{red}{current and previous} round.) Instead of 10, 50, 90, and 130 seen in the previous round, we now use 10 (1st category), 40 (2nd category), 70 (3rd category), and 100 (3rd category).
+This is because 130 meters is really high and rarely observed, and we wanted the range of analysis to be more representative of the region.
+Please note that key qualitative message is not any different whether we use either set of values.}
 
@@
   \textcolor{blue}{From the initial submission, it has been clearly stated that county fixed effects $\alpha_i$ are included in equation 1.
-However, we understand your confusion because we stated the following in our response.}
+Here is the screenshot of the relevant part from the initial version.
+The last sentence of the paragraph below states that county fixed effects are included in equation 1.}
 
-  \textcolor{blue}{``Other sources of bias (e.g., model misspecification error) will also almost always
-exist.
+  \includegraphics{fixedeffects}
+
+  \textcolor{blue}{However, we may have caused your confusion because we stated the following in our response.}
+
+  \textcolor{blue}{``Other sources of bias (e.g., model misspecification error) will also almost always exist.
 In our analysis, one of the main limitations is not including county fixed effects in our regression\textcolor{red}{s}, which would have controlled all the county-specific time-invariant characteristics.''}
@@
 
-  \textcolor{blue}{We agree. It is too strong of a claim. We simply tone down our confidence.}
+  \textcolor{blue}{We agree that it is too strong of a claim. Please see line 404 for our revision of the statement.}
 
-  \textcolor{blue}{We also changed our sentence in line .}
+  \textcolor{blue}{"This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+However, it is not possible to understand the true consequences of not being able to include county-fixed effects cannot be fully investigated with our datasets, and we defer this matter to future studies with less spatially disaggregated data with enough temporal variations in aquifer thickness levels."}
+
+  \textcolor{blue}{We also removed the sentence below in line 345 - 346 in the previous version to the following:}
+
+  \textcolor{blue}{"Instead, through robustness checks, we demonstrate that our results are consistent when using state fixed effects."}
 
@@
 
+   \textcolor{blue}{We have added total number of observations to the captions of Figures 1 and 3.
+Figure 4 is the results of simulation based on the estimated model represented by equations 1 and 2.
+So, there is no appropriate number of observations for the figure.}
+
+   \textcolor{blue}{We have added histograms for Figures 3 and 4.
+The inconsistency in the range of the histograms and the main figures were caused by a small coding error that did not align two figures vertically in the right way.
+The problems are fixed now.
+Please note that because of the width (range) of the histrogram bins, it looks like the bins at the left and right ends are beyond the range of the main figure.
+But they are not.}
+
   \item (5.b) Lines 77-78 state ``rainfed yields of both crops decline significantly'' with reference to Figure 1.
 Please quantify, visualize, or otherwise provide evidence of the significance to readers, or rephrase.
+
+  \textcolor{blue}{Figure A.1 provides the estimated response curve for rainfed production.
+As you can see, the confidence intervals are quite narrow at all the water deficit values.
+We do not think we need additional figures or explanations are necessarily for both statistical and real-world significance of the impacts of water deficit on rainfed production.}
+
   \end{itemize}
@@
 
+\textcolor{red}{Tim, where is this rule coming from? It is not clear he is referring to them in equations or in-line.}
+
 \item (M1.b) Lines 325-326: ``with $\sigma_j$ represents the category specific intercept'' should be ``with $\sigma_j$ representing the category-specific intercept'' or something similar
 
+\textcolor{blue}{You are correct. It is fixed now.}
+
 \item (M1.c) Equation 2: What is ``$\mu_t$'' in equation 2?
+
+\textcolor{blue}{It used to represent year fixed effects. But, now that we included state-year FE represented $\theta_{s,t}$, we should have removed it. Equation 2 no longer has $mu_t$.}
 
@@
 
+\textcolor{blue}{We changed $y_{i,t}$ in equation (1) to $y_{i,t}$.}
+
 \item (M2.b) Equation 2: Please replace the re-used symbol ``alpha'' with a different symbol to avoid confusion.
 
+\textcolor{blue}{We changed all the occurrences of $\alpha_{k}$ to $\eta_{k}$.}
+
 \item (M2.c) Lines 427-428: Please clarify the grammatical structure of the following sentence: ``Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate $\beta_k$ to fit the data best.''
+
+\textcolor{blue}{We rewrote it to the following:}
+
+\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate $\beta_k$ to fit the data best.''}
+
+\textcolor{blue}{In statistics, ``linear'' model means that the model is linear in parameter, but not necessarily in independent variable.
+For example, the model below is linear in parameter, but not linear in $x$.}
+
+\begin{align}
+  y = \beta x^2 + v
+\end{align}
+
+\textcolor{blue}{ GAM is linear in parameter, but not linear in independent variable. We wanted to emphasize the distinction and used the term linear-in-parameter model instead of linear model.}
 

### 2023-08-31 Codes/2_figures_tables.rmd
manuscripts updated and replies to reviewers created to be reviewed

@@
 g_share_dif <-
-  ggplot(plot_data_ir_share[sat < 100, ]) +
+  ggplot(plot_data_ir_share[sat <= 100, ]) +
   geom_line(aes(y = dif_ir_share_hat, x = sat)) +
@@
   scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
-  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
+  scale_x_continuous(breaks = seq(10, 150, by = 10)) +
   xlab("Aquifer Thickness (meter)") +
@@
     strip.background = element_blank(),
-    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
+    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = 9)
   )
@@
   height = 4,
-  width = 5
+  width = 6
 )

### 2023-08-31 Writing/NatureWater/2ndRoundRevision/manuscript.tex
manuscripts updated and replies to reviewers created to be reviewed

@@
 Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}\label{fig:yield-response}
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
+Total number of observations are $8,773$ and $5,977$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
 \end{figure}
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100m to 0.72 (72\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
-Reductions in the share of production area under irrigation is also evident for soybean although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}.
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100 m to 0.72 (72\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+Reductions in the share of production area under irrigation is also evident for soybean.
 
@@
 
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval)}\label{fig:ir-share}
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval).
+The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
+Total number of observations are $2,523$ and $1,558$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
 \end{figure}
@@
 The overall influence of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where the weights are production area shares) for corn and soybean at four values of aquifer thickness: 10, 50, 90, and 130 meters.
+Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where the weights are production area shares) for corn and soybean at four values of aquifer thickness: 10, 40, 70, and 100 meters.
 While similar trends are observed as in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced due to lower estimated irrigated area share for smaller aquifer thickness values.
@@
 
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 50 m, 90 m, and 130 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 350mm.
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 40 m, 70 m, and 100 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 400 mm.
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
-This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600mm) and are smaller in overall magnitude.
-For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 130m and 10m aquifer thicknesses.
+This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas where aquifer thickness is lower.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600 mm) and are smaller in overall magnitude.
+For example, for a seasonal water deficit of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 100 m and 10 m aquifer thicknesses.
 
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only 0.11 tonne/ha of production loss for corn.
+At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.09 tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in 0.77 tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in 2.09 tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading in 1.57 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 
-\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 50 m, 90 m, and 130 m, relative to a baseline aquifer thickness of 10 m.
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
 Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}\label{fig:dif-tot-impact}
@@
 
-These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield (8773 observations), (2) per-acre soybean yield (5977 observations), (3) share of irrigated corn (2686 observations), and (4) share of irrigated soybean (2014 observations).
+These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($8,773$ observations), (2) per-acre soybean yield ($5,977$ observations), (3) share of irrigated corn ($2,523$ observations), and (4) share of irrigated soybean ($1,558$ observations).
 
@@
 \begin{equation}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
+Y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
 $K$ is the total number of spline basis functions.
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} \label{eq:eqn2}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \eta_{k}\phi_{k}(WB_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water balance is captured by \(\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\alpha_{k}\) is its coefficient.
+The impact of average water balance is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(WB_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
@@
 Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
-Instead, through robustness checks, we demonstrate that our results are consistent when using state fixed effects.
-A formal discussion of these results and trade-offs associated with the choice of fixed effects is provided in Appendix \ref{county-fe}.
+A formal discussion of the trade-offs associated with the choice of fixed effects is provided in Appendix \ref{county-fe}.
 
@@
 
-Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 130m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
+Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100 m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
 Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
-Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 130 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 100 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
 
@@
 
-\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 130 m (black line).
+\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 100 m (black line).
 The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
@@
 Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
+This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+However, it is not possible to understand the true consequences of not being able to include county-fixed effects cannot be fully investigated with our datasets, and we defer this matter to future studies with less spatially disaggregated data with enough temporal variations in aquifer thickness levels.
 

### 2023-08-31 Writing/manuscript.Rmd
manuscripts updated and replies to reviewers created to be reviewed

@@
 Figure \@ref(fig:tot-impact) and Figure \@ref(fig:dif-tot-impact) also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only `r round(res_at_950[sat == 130 & crop == "corn", avg_yield] - res_at_950[sat == 90 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only `r round(res_at_950[sat == 100 & crop == "corn", avg_yield] - res_at_950[sat == 70 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \@ref(fig:ir-share).
-Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in `r round(res_at_950[sat == 90 & crop == "corn", avg_yield] - res_at_950[sat == 50 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in `r round(res_at_950[sat == 50 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
+Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in `r round(res_at_950[sat == 70 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in `r round(res_at_950[sat == 40 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)` tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.

### 2023-08-31 Codes/Functions/functions.R

@@ -436,6 +436,10 @@ share_analysis_gam <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e,
     ir_share_hat_se = share_hat$se.fit
   )]
 
+  base_ir_share <- share_hat_data[sat == sat_seq[length(sat_seq)], ir_share_hat]
+
+  share_hat_data[, dif_ir_share_hat := base_ir_share - ir_share_hat]
+
   # ggplot(share_hat_data) +
   #   geom_line(aes(y = ir_share_hat, x = sat)) +
   #   geom_ribbon(aes(
@@ -445,7 +449,7 @@ share_analysis_gam <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e,
   #   ), alpha = 0.4)
 
   return(
-    share_hat_data = share_hat_data[, .(sat, ir_share_hat, ir_share_hat_se)]
+    share_hat_data = share_hat_data[, .(sat, ir_share_hat, ir_share_hat_se, dif_ir_share_hat)]
   )
 }
 

### 2023-08-31 Writing/NatureWater/2ndRoundRevision/replies/cover_letter.tex
Edits to cover letter.

--- 2023-08-30 Writing/NatureWater/1stRoundRevision/replies/cover_letter.tex
@@
 
-Specifically, in our revised manuscript, we have made the following major changes based on the reviewers' comments, alongside several minor adjustments to wording and to correct spelling and grammatical errors:
+The vast majority of the changes we made are minor. Some notable changes are the following:
 
 \begin{itemize}
-    \item Updated the main results section to include additional variables (e.g., soil characteristics, state fixed effects) to strengthen how we account for unobservable factors in our regression analyses.
-    \item Added several new figures (Figure 2 and Figure 5) to more clearly highlight the non-linearity and statistical significance of declining saturated thickness on per-area crop yields (Figure 2) and overall crop productivity (Figure 5), which, combined with revised discussion of our results, we believe enable readers to gain deeper understanding of the significance of our findings.
-    \item Conduct rigorous statistical testing of our main findings, in particular to highlight the specific climate and aquifer conditions at which aquifer depletion has statistically robust and significant impacts on drought impacts and risks, including how these thresholds vary between corn and soybean.
-    \item Strengthened the descriptions on the underlying theoretical and empirical evidence for causal links between crop yield, share of irrigated acres, and aquifer thickness, which are one of the main empirical findings from our paper.
-    \item Added a series of four appendices to discuss in more details our methods, robustness checks, limitations imposed by underlying data. 
+    \item Fix a coding error that was creating invalid observations for our regressions analysis of the share of irrigated acres.
+Consequently, our results changed slightly.
+Our main conclusions and qualitative results are robust and did not change.
+    \item We added histograms of aquifer thickness in Figure 3.
 \end{itemize}

### 2023-09-02 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
minor text change

@@
 
-\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate $\beta_k$ to fit the data best.''}
+\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear (in-parameter) model and estimate $\beta_k$ to fit the data best.''}
 
@@
 
-\textcolor{blue}{ GAM is linear in parameter, but not linear in independent variable. We wanted to emphasize the distinction and used the term linear-in-parameter model instead of linear model.}
+\textcolor{blue}{ GAM is linear in parameter, but not linear in independent variable. We wanted to emphasize the distinction instead of simply saying linear model.}
 
@@
 
-\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate $\beta_k$ to fit the data best.''}
+\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear (in parameter) model and estimate $\beta_k$ to fit the data best.''}
 

### 2023-09-02 Writing/NatureWater/2ndRoundRevision/manuscript.tex
Edits to cover letter.
minor text change

@@
 $K$ is the total number of spline basis functions.
-$K = 3$ was chosen for this regression as it minimizes BIC.
+$K = 3$ was chosen for this regression as it minimizes Bayesian Information Criterion (BIC).
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
@@
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
-Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate \(\beta_k\) to fit the data best.
+Notice that unlike non-parametric regression, GAM is actually a linear (in parameter) model and estimate \(\beta_k\) to fit the data best.
 

### 2023-09-02 Codes/2_figures_tables.rmd
minor fixes in yield intensive figures

@@
         legend.position = "bottom",
-        plot.margin = margin(0, 0, 0, 0, "cm"),
+        plot.margin = margin(0, 0, 0.0, 0, "cm"),
         plot.title = element_blank()
@@
   cowplot::plot_grid(
-    g_wd_corn, g_ir_yield_corn, g_wd_soy, g_ir_yield_soy,
+    g_wd_corn, 
+    g_ir_yield_corn + theme(plot.margin = margin(0, 0, 0.2, 0, "cm")), g_wd_soy, 
+    g_ir_yield_soy,
     ncol = 1,

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
comments and edits added by Tim

--- 2023-08-30 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
@@
 \begin{itemize}
-    \item Line 58 – ``farmers'' should be ``farmers’''.
+    \item Line 58 – ``farmers'' should be ``farmers'''.
 
-\textcolor{blue}{Thank you for catching this. Fixed.}
+\textcolor{blue}{Thank you for catching this. We have fixed all instances of this typo.}
 
@@
 
-\textcolor{blue}{We agree. Instead of county-level "average" water deficit, we say county-level annual water deficit because they are not averaged values.}
+\textcolor{blue}{We agree. Instead of county-level "average" water deficit, in our revised manuscript we now say county-level annual water deficit because they are not averaged values.}
 
@@
 
-\textcolor{blue}{Thank you for catching this. Fixed.}
+\textcolor{blue}{Thank you for catching this. This typo has been fixed.}
 
-\item Equations 1 and 2 – shouldn’t ``k'' be ``k=1'' – if not, please explain how k is determined. Also, explain ``K'' and how it was determined.
+\item Equations 1 and 2 – shouldn't ``k'' be ``k=1'' – if not, please explain how k is determined. Also, explain ``K'' and how it was determined.
 
-\textcolor{blue}{Please see line 327 for the explanation of what K represents and how K is selected. We added similar explanations for equation 2 at line 341.}
+\textcolor{blue}{K represents the total number of spline basis functions, which influences the complexity of the spline shape in a GAM.
+A summary of this detail can be found on line 316 of our revised manuscript.
+We have also added similar explanations for equation 2 on line 330.}
 
@@
 
-\textcolor{blue}{Figure A.3 indeed presents what we intended to present.
-Please note that we made similar changes for average productivity analysis in the main text as shown in Figures 4 and 5.
-While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable.
-In our average productivity analysis, we use four aquifer thickness values: 10, 40, 70, and 100 meter.
-For example, for the analysis of 10 meter, estimated yield response function for the first category is used as it falls within the [9m, 30.1 m] range.
-Estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres.
-In our initial submission, we were doing the same kind of analysis (using one number from each aquifer thickness category), but reported by category.
-In the previous round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter).
-This is in response to another reviewer's point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.
-Given the chosen series of aquifer thickness numbers are equi-distance, it has become much easier to observe the non-linearity in the impact of saturated thickness on overall productivity.
-Please note that we now present results for 10, 40, 70, and 100 meters, which are in the first, second, third, and third saturated thickness categories.
-}
+\textcolor{blue}{While the irrigated yield regression (equation 1) uses aquifer thickness categories, aquifer thickness is specified as a continuous variable for our irrigated area share regression.
+In our average productivity analysis, the results of yield and area regression results are combined by assigning a specific aquifer thickness value (e.g.
+10m) to the relevant category for the yield component and using the specific thickness value for the area component of the regression.
+This is a necessary step given the differences in the format (i.e.
+categorical vs continuous) of the aquifer thickness variable in each of the regression components of our combined average productivity model.
+In Figure A.3.
+and also Figure 5, the set of thickness values (10, 40, 70, and 100m - covering the first, second, third, and third saturated thickness categories, respectively) were selected to be equi-distance, so as to enable more direct exploration of whether the impact of aquifer thickness on average productivity is non-linear in response to a comment raised by one of the other reviewers.
+To clarify this feature of our average productivity model we have added the following sentence to line 341 of the methods of our revised manuscript:}
+
+\textcolor{blue}{``In this combined model, aquifer thickness is a continuous variable for the irrigated area share regression component and a categorical variable for the irrigated yield regression following the format of Equations 2 and 1, respectively.''}
 
@@
 
-\textcolor{blue}{Thank you for catching this. Fixed.}
+\textcolor{blue}{Thank you for catching this. This typo has been corrected.}
 
@@
 
-\textcolor{blue}{We understand that demeaning sounds wrong, however, it is a term that is used specifically for the types of operations we are conducting in that analysis in Economics.}
-
-\textcolor{red}{Tim, can you work on making the paragraph free of grammatical errors?}
+\textcolor{blue}{While we understand the confusion around the term demeaning, we feel it is important to retain this term in the manuscript as it an important and established terms in statistics and economics that denotes the centering of data by removing the mean.
+This is subtly different to detrending, and hence we retain demeaning as a term in our manuscript.
+Noun-verb disagreements in this paragraph have been corrected in this section of the manuscript.}
 
@@
 
-\item Line 428 – is ``a linear in parameter model'' correct? Wouldn’t ``a linear parameter model'' be better?
+\textcolor{blue}{Thank you for catching this. The text has been modified for consistency as suggested.}
 
-\textcolor{blue}{We rewrote it to "linear-in-parameter" model. Linear-in-parameter model is a more explicit and correct way of saying what we people call a linear model.}
+\item Line 428 – is ``a linear in parameter model'' correct? Wouldn't ``a linear parameter model'' be better?
+
+\textcolor{blue}{From a statistical perspective, a ``linear'' model means that the model is linear in parameter, but not necessarily with regard to the independent variable.
+For example, the hypothetical model below is linear in parameter, but not linear in $x$.}
+
+\begin{align}
+  y = \beta x^2 + v
+\end{align}
+
+\textcolor{blue}{GAM is an example of a type of model that is linear in parameter, but non-linear with respect to their independent variable(s).
+We have therefore retained the term linear in parameter to ensure a clear distinction with non-linear regression models.
+To more clearly clarify this point, we have also rewritten the sentence flagged by the reviewer as follows:}
+
+\textcolor{blue}{``Unlike non-parametric regression, a GAM is a linear (in-parameter) model and estimates $\beta_k$ to best fit the observation data.''}
 
@@
 We have addressed all your comments and suggestions, and our replies to your comments can be found in blue below (original comments in black).
-Please take a look at our comments about a small change in our codes at the top as well.}
+Please also note our comments about a small change in our codes at the top as well.}
 
@@
 
-  \textcolor{blue}{As clearly stated in the manuscript, county fixed effects were included in equation 1 (reperesented by $\alpha_i$).
-This means that any county-specific time-invariant variables like those in $X$ will be dropped.
-This does not mean $X$ are not accounted for.
-They are accounted for as county fixed effects will capture them.}
+  \textcolor{blue}{As stated in the manuscript, county fixed effects were included in equation 1 (represented by $\alpha_i$).
+This means that any county-specific time-invariant variables - such as those related to soil properties as represented in $X$ - will already be accounted for implicitly in the county-level fixed effects term.
+In contrast, soil property variables are included explicitly in the irrigated area share regression as this regression only includes state-year fixed effects for reasons discussed in Appendix C}
 
@@
   \textcolor{blue}{Yes.
-Moreover, the results for the share of irrigated acre in this manuscript are different from the previous round due to the changes in the choice of the number of spline basis functions ($K$ and $L$).
-Previously, we were using 5 for both.
-However, in responding to reviewer 1's request to explain how the number of spline basis functions are determined, we realized that we were not using the optimal numbers.
-We tested different numbers earlier in the project and found that the results are not sensitive to the choice of the number of spline basis functions.
-So, we were simply using reasonable numbers back then.
-Even though the choice of $K$ and $L$ does not affect the results much, we now pick the numbers based on BIC.}
+Please note that the results for the share of irrigated acre in this manuscript are also different from the previous version of our manuscript  due to the code updates that we have summarized at the top of this document.
+All the qualitative results and conclusions regarding the impacts of aquifer thickness are the same as the previous versions of our manuscript.}
 
@@
 
-  \textcolor{blue}{The change made other than the ones explained above and including state-year FE is whether we allow the interactions of $WB$ and $AT$.
-Since we found that there is not significant interactions between them, we have decided to make the model more parsimonious and simply model the impact of $WB$ and $AT$ separately.
-The results are virtually the same whether we used the old and current models.}
+  \textcolor{blue}{The other changes made to Equation 2 include the addition state-year FE, in order to provide stronger controls than just including year FE.
+In addition, we choose not to allow the interactions of $WB$ and $AT$.
+This decision was made in the interests of developing a more parsimonious model that considers the impacts of $WB$ and $AT$ individually.
+Our results remain virtually identical though whether this the interaction between $WB$ and $AT$ is included or not.
+}
 
@@
 
-  \textcolor{blue}{We should have mentioned this change in our previous replies.
-We found that we were actually not grouping observations into three groups in a way that their number of observations are the same as we claimed due to a coding error.
-We fixed this problem in the previous round.
-The changes observed in Figure 1 is due solely to this change.
-For Figure 3, the changes are combination of this change and also the change in the results of the share of irrigated acres as mentioned above.
-The previous round of Figure 3 of course is different from the one in the initial version as we added another aquifer level to make our claim of the non-linear impact of aquifer much clearer.
-Figure 3 in the current version is different from the one in the previous round as we changed the number of spline basis function for equation 2 as mentioned above.
-Overall, the observed changes in results are really minor.}
+  \textcolor{blue}{Thank you for highlighting this important omission from our previous response in the last review round.
+In our earlier revision of the manuscript, we found a small coding error that meant we were actually not grouping observations equally into three groups as stated in our methods.
+This coding error was corrected in our previous revised submission, and the changes observed in Figure 1 are due solely to this change.
+For Figure 4 of the previous version (1st round revision), the changes are combination of this change and also the change in the results of the regression for share of irrigated acres as mentioned above.
+This is because Figure 4 presents the results of the analysis that combines what are presented in Figures 1 and Figure 3 of the previous version (1st round revision).
+Figure 3 in the 1st round and Figure 4 in the previous round are different because we added another aquifer level to make our claim of the non-linear impact of aquifer much clearer in response to one of your comments in the previous round.
+Further, Figure 4 in the current version is different from the one in the previous round as we changed the aquifer levels for this analysis.
+Instead of 10, 50, 90, and 130 meters, we use 10, 40, 70, and 100 meters.
+We made this change to illustrate that significant non-linear changes in total productivity can happen within a narrower range of evenly spaced aquifer thickness compared to the previous sets of aquifer thickness values used in this figure.}
 
@@
 
-  \textcolor{blue}{We added the following sentence at line 292.}
+  \textcolor{blue}{We added the following sentence at line 277.}
 
@@
 Observation units of the SSURGO data are polygons.
-For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables.}
+For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables were calculated.}
 
@@
     \textcolor{blue}{Thank you for noticing this.
-We did not realize the difference is no longer statistically significant.
-Following your suggestion, we simply dropped the pararaph describing the results.}
+You are correct that this difference is no longer statistically significant.
+Following your suggestion, we have removed this paragraph from the results to ensure consistency in what is discussed as significant in the context of our analysis.}
 
@@
 
-    \textcolor{blue}{We are not using the different categories between Figures 1 and 4.
-While the irrigated yield regression (equation 1) uses aquifer thickness category, the irrigated share regression uses aquifer thickness as it is as a continuous variable.
-In our average productivity analysis which combines the two kinds of analysis, we use four aquifer thickness values: 10, 50, 90, and 130 meter.
-For example, for the analysis of 10 meter, the estimated yield response function to water deficit for the ``first'' category is used as it falls within the [9m, 30.1 m] range.
-The estimated irrigation share for 10 meter can be obtained simply by plugging 10 into the trained model of the share of irrigated acres (equation 2).
-In our initial submission, we were doing the same of analysis (using one number from each aquifer thickness category), but reported by category instead of reporting aquifer levels we were using for each category.
-In the first round of revision, we decided to have two aquifer thickness values in the 3rd quantile category (90 and 130 meter).
-50 meter falls in the second category.
-This is in response to your point that it is not clear if indeed the impact of aquifer thickness on average productivity is non-linear as we claimed in our initial submission.}
+    \textcolor{blue}{Thank you for this helpful point, which highlgihts a need to clarify some details about how aquifer thickness enters different parts of our analysis.
+While the irrigated yield regression (equation 1, Figure 1) uses aquifer thickness category, the irrigated share regression (Figure 4) uses aquifer thickness as it is as a continuous variable.
+In our average productivity analysis (Figure 4) which combines these two regressions, aquifer thickness is specified as a specific point value for the component related to irrigated area share and as a categorical value for component related to irrigated per-area yields.
+This step has been the same throughout all versions of our manuscript, and is not a change that has been introduced in current or prior revisions.
+We have added the following sentence on line X of our methods to clarify this point:}
+
+    \textcolor{blue}{``In this combined model, aquifer thickness is a continuous variable for the irrigated area share regression component and a categorical variable for the irrigated yield regression following the format of Equations 2 and 1, respectively.''}
+
+    \textcolor{blue}{In the previous version of our average productivity analysis, we assess the impacts of four aquifer thickness values: 10, 50, 90, and 130 meters in the initial version of our manuscript, with equal spacing chosen to illustrate non-linearity in impacts of aquifer thickness.
+In our last round of revisions, we changed these values to instead be 10, 40, 70, and 100m.
+These values ensure that we have a range of values spanning different aquifer thickness categories.
+Please note that key qualitative message and findings is identical whichever set of values we use.
+As we mentioned earlier, we made this change to illustrate that significant non-linear changes in total productivity can happen within a narrower range of evenly spaced aquifer thickness compared to the previous sets of aquifer thickness values used in this figure.
+This change also provide results that are more realistically centred on the range of aquifer thickness values observed within our study area and period.}
 
@@
 
-  \textcolor{red}{Tim, I don't think this guys understood what we wrote.}
+  \textcolor{blue}{Regulatory pumping restrictions are not a binding constraint to water use decisions in our study area, and therefore are unlikely to impact the ability to use groundwater as a buffer against drought.
+In contrast, the impacts of aquifer depletion on farmers ability to meet crop water needs through changes in well yields are linked directly to hydrologic and agronomic theory, and have demonstrated extensively through prior work cited in our manuscript.
+While we note that restrictions in pumping are imposed in some parts of the High Plains Aquifer, there are several reasons why we can confidently state these would not be a confounding source of causality in our analysis:
+  \begin{enumerate}
+      \item The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher (Rimsaite and Brozovic, 2023).
+This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers water use decisions are not constrained by such regulations even in extreme drought years such as 2012 (Foster et al., 2019).
+      \item Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas (Marston et al., 2022; Rimsaite and Brozoic, 2023), these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
+The earliest of these groundwater management in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
+While this overlaps with the final 4 years of our study period, the LEMA area covers only parts of two counties and hence represents a very small proportion of our overall study sample while the years 2013-2016 were all characterised by average or above average growing season rainfall (i.e.
+not drought years).
+Moreover, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions (Drysdale and Hendricks, 2018; Deines et al., 2021).
+These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
+      \item Areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction due to laws that typically provide farmers with absolute rights to use groundwater under their land coupled with low levels of metering that constrain potential for implementing extensive and binding pumping restrictions (Wheeler et al., 2016; Closas and Molle, 2018).
+  \end{enumerate}
+  }
+
+  \textcolor{blue}{References referred to above are as follows:
+  \begin{itemize}
+      \item Rimsaite, R., and Brozovic, N.  (2023). Groundwater Development Paths in the U.S. High Plains. Oxford Research Encyclopedia of Environmental Science. 
+      \item Foster, T., Goncalves, I.
+Z., Campos, I., Neale, C.
+M.
+U., and Brozovic, N.
+(2019).
+Assessing landscape scale heterogeneity in irrigation water use with remote sensing and in situ monitoring.
+Environmental Research Letters, 14(2), 024004.
+      \item Marston, L.
+T., Zipper, S., Smith, S.
+M., Allen, J.
+J., Butler, J.
+J., Gautam, S., and David, J.
+Y.
+(2022).
+The importance of fit in groundwater self-governance.
+Environmental Research Letters, 17(11), 111001.
+      \item Closas, A., and Molle, F. (2018). Chronicle of a Demise Foretold: State vs. Local Groundwater Management in Texas and the High Plains Aquifer System. Water Alternatives, 11(3), 511-532.
+      \item Wheeler, S.
+A., Schoengold, K., and Bjornlund, H.
+(2016).
+Lessons to be learned from groundwater trading in Australia and the United States.
+Integrated Groundwater Management: Concepts, Approaches and Challenges, 493-517.
+      \item Deines, J.
+M., Kendall, A.
+D., Butler, J.
+J., Basso, B., and Hyndman, D.
+W.
+(2021).
+Combining remote sensing and crop models to assess the sustainability of stakeholder‐driven groundwater management in the US high plains aquifer.
+Water Resources Research, 57(3), e2020WR027756.
+      \item Drysdale, K.
+M., and Hendricks, N.
+P.
+(2018).
+Adaptation to an irrigation water restriction imposed through local governance.
+Journal of Environmental Economics and Management, 91, 150-165.
+  \end{itemize}
+  }
 
@@
 
-  \textcolor{blue}{From the initial submission, it has been clearly stated that county fixed effects $\alpha_i$ are included in equation 1.
-However, we understand your confusion because we stated the following in our response.}
-
-  \textcolor{blue}{``Other sources of bias (e.g., model misspecification error) will also almost always
-exist.
-In our analysis, one of the main limitations is not including county fixed effects in our regression\textcolor{red}{s}, which would have controlled all the county-specific time-invariant characteristics.''}
-
-  \textcolor{blue}{It is our mistake that we added \textcolor{red}{s} to ``regression''. As we stated above, we do include county fixed effect in equation 1. }
+  \textcolor{blue}{In our initial submission, we noted in our methods that county fixed effects $\alpha_i$ were included in equation 1 but not in equation 2.
+The reason for this difference is outlined in Appendix C, which provides a detailed discussion of why including county fixed effects for our irrigated area share regression is problematic.}
 
@@
 
-  \textcolor{blue}{We agree. It is too strong of a claim. We simply tone down our confidence.}
+  \textcolor{blue}{We recognize your point that this may be potentially too strong of a claim.
+To more carefully summarise the implications of our choice of fixed effects, we have modified the text on lines XXX-XXX for our revision of the statement.}
 
-  \textcolor{blue}{We also changed our sentence in line .}
+  \textcolor{blue}{"This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects."}
+
+  \textcolor{blue}{We also removed the sentence below that was state on lines 345 - 346 in the previous version of our manuscript to reflect that we can only hypothesize about the likely differences between using state and county fixed effects:}
+
+  \textcolor{blue}{"Instead, through robustness checks, we demonstrate that our results are consistent when using state fixed effects."}
 
@@
 
+   \textcolor{blue}{We have added total number of observations to the captions of Figures 1 and 3.
+Figure 4 is the results of simulation based on the estimated model represented by equations 1 and 2, and therefore there is no appropriate number of observations for the figure so the value is omitted.}
+
+   \textcolor{blue}{We have added histograms for Figures 3 and 4.
+The inconsistency in the range of the histograms and the main figures were caused by a small coding error that did not correctly align two figures vertically.
+These problems are now fixed.
+Please note that because of the width (range) of the histrogram bins, it looks like the bins at the left and right ends are beyond the range of the main figure, but this is not the case and is simply a plotting feature.}
+
   \item (5.b) Lines 77-78 state ``rainfed yields of both crops decline significantly'' with reference to Figure 1.
 Please quantify, visualize, or otherwise provide evidence of the significance to readers, or rephrase.
+
+  \textcolor{blue}{Figure A.1 provides the estimated response curve for rainfed production.
+As is visible, the confidence intervals are narrow at all the water deficit values.
+We do not think we need additional figures or explanations are necessarily for both statistical and real-world significance of the impacts of water deficit on rainfed production.}
+
   \end{itemize}
@@
 
+\textcolor{blue}{``WB'' and ``ST'' terms have been changed to italics throughout}
+
 \item (M1.b) Lines 325-326: ``with $\sigma_j$ represents the category specific intercept'' should be ``with $\sigma_j$ representing the category-specific intercept'' or something similar
 
+\textcolor{blue}{You are correct. This has been corrected in the revised manuscript.}
+
 \item (M1.c) Equation 2: What is ``$\mu_t$'' in equation 2?
+
+\textcolor{blue}{``$\mu_t$'' represents year fixed effects.
+However, as we now use state-year FE represented by $\theta_{s,t}$, the ``$\mu_t$' should have been removed.
+Equation 2 no longer includes $mu_t$.}
 
@@
 
+\textcolor{blue}{We changed $y_{i,t}$ in equation (1) to $y_{i,t}$.}
+
 \item (M2.b) Equation 2: Please replace the re-used symbol ``alpha'' with a different symbol to avoid confusion.
 
+\textcolor{blue}{We changed all the occurrences of $\alpha_{k}$ to $\eta_{k}$.}
+
 \item (M2.c) Lines 427-428: Please clarify the grammatical structure of the following sentence: ``Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate $\beta_k$ to fit the data best.''
+
+\textcolor{blue}{We have rewritten the sentence as follows:}
+
+\textcolor{blue}{``Unlike non-parametric regression, a GAM is a linear (in-parameter) model and estimates $\beta_k$ to best fit the observation data.''}
 

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/manuscript.tex
comments and edits added by Tim

@@
 Soil characteristics data (sand percentage, silt percentage, and water holding capacity) were obtained from the Soil Survey Geographic
-Dataset (SSURGO). Observation units of the SSURGO data are polygons. For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables.
+Dataset (SSURGO).
+Observation units of the SSURGO data are polygons.
+For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables were calculated.
 
@@
 The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
+In this combined model, aquifer thickness is a continuous variable for the irrigated area share regression component and a categorical variable for the irrigated yield regression following the format of Equations \eqref{eq:eqn2} and \eqref{eq:eqn1}, respectively.
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
@@
 
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its yearly average (A), state-year average (B), and county average (C).
-This analysis is used to gain insights into how much variation in aquifer thickness for identification is left under only year fixed effects (A), state-year fixed effects (B), and county and year fixed effects (C).
-As shown in Figure \ref{fig:variation-left}, including state-year fixed effects do not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
-However, including county fixed effects dramatically reduces its total variation in aquifer thickness resulting in extremely inaccurate estimates (high standard error).
-For this reason, we included state-year fixed effects in irrigated production share regression analysis as a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
+For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
 
@@
 Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same across when using different possible sets of fixed effects.
 Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
-However, it is not possible to understand the true consequences of not being able to include county-fixed effects cannot be fully investigated with our datasets, and we defer this matter to future studies with less spatially disaggregated data with enough temporal variations in aquifer thickness levels.
+This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
 
@@
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
-Notice that unlike non-parametric regression, GAM is actually a linear (in parameter) model and estimate \(\beta_k\) to fit the data best.
+Unlike non-parametric regression, a GAM is a linear (in-parameter) model and estimates $\beta_k$ to best fit the observation data.
 

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
manuscript modified

@@
 As we mentioned earlier, we made this change to illustrate that significant non-linear changes in total productivity can happen within a narrower range of evenly spaced aquifer thickness compared to the previous sets of aquifer thickness values used in this figure.
-This change also provide results that are more realistically centred on the range of aquifer thickness values observed within our study area and period.}
+This change also provide results that are more realistically centered on the range of aquifer thickness values observed within our study area and period.}
 
@@
 
-  \textcolor{blue}{Regulatory pumping restrictions are not a binding constraint to water use decisions in our study area, and therefore are unlikely to impact the ability to use groundwater as a buffer against drought.
-In contrast, the impacts of aquifer depletion on farmers ability to meet crop water needs through changes in well yields are linked directly to hydrologic and agronomic theory, and have demonstrated extensively through prior work cited in our manuscript.
-While we note that restrictions in pumping are imposed in some parts of the High Plains Aquifer, there are several reasons why we can confidently state these would not be a confounding source of causality in our analysis:
+  \textcolor{blue}{Regulatory pumping restrictions have limited impacts on water use decisions in our study area, and therefore are unlikely to impact the ability to use groundwater as a buffer against drought.
+In contrast, the impacts of aquifer depletion on farmers' ability to meet crop water needs through changes in well yields are linked directly to hydrologic and agronomic theory, and have demonstrated extensively through prior work cited in our manuscript.
+While we note that restrictions in pumping are imposed in some parts of the High Plains Aquifer, there are several reasons why we can confidently state these would not be a confounding source of causality in our analysis.
+Please see the list below.
+We added the following paragraphs as Appendix A in the latest manuscript.
+
   \begin{enumerate}
       \item The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher (Rimsaite and Brozovic, 2023).
-This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers water use decisions are not constrained by such regulations even in extreme drought years such as 2012 (Foster et al., 2019).
+This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers' water use decisions are not constrained by such regulations even in extreme drought years such as 2012 (Foster et al., 2019).
       \item Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas (Marston et al., 2022; Rimsaite and Brozoic, 2023), these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
-The earliest of these groundwater management in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
-While this overlaps with the final 4 years of our study period, the LEMA area covers only parts of two counties and hence represents a very small proportion of our overall study sample while the years 2013-2016 were all characterised by average or above average growing season rainfall (i.e.
+The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
+While this overlaps with the final 4 years of our study period, the LEMA area covers only parts of two counties and hence represents a very small proportion of our overall study sample while the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
 not drought years).
@@
 
-  \textcolor{blue}{"This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
+  \textcolor{blue}{"This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
 We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects."}

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/manuscript.tex
manuscript modified

@@
 
-Two pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
+Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}\footnote{Another pathway is through groundwater use regulation in response to low or declining saturated thickness.
+However, it is unlikely that this effects is captured in our analysis.
+Please see Appendix \ref{pathway-reg}}.
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
@@
 Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET in order to calculate daily reference crop evapotranspiration for each grid cell.
-From these data, we computed the seasonal water deficit (\(WB_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
+From these data, we computed the seasonal water deficit (\(D_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
 Seasonal water balance estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell.
@@
 
-To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(WB\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield response to seasonal water deficits is conditioned on the level of aquifer thickness (\(AT\)) in the county in a given year.
+To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(D\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield response to seasonal water deficits is conditioned on the level of aquifer thickness (\(A\)) in the county in a given year.
 Our regression model is given in Equation \eqref{eq:eqn1}, with the same model formulation used for corn and soybean:
@@
 \begin{equation}
-Y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
+Y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(D_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
 
-where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit.
+where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(D\) is seasonal water deficit.
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \eta_{k}\phi_{k}(WB_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + v_{i,t} \label{eq:eqn2}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \eta_{k}\phi_{k}(D_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(A_{i,t}) + X_i + \theta_{s,t} + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water balance is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(WB_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
+The impact of average water balance is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(D_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
-$K$ and $L$ are the numbers of basis functions for $WB$ and $AT$, respectively.
+$K$ and $L$ are the numbers of basis functions for $D$ and $A$, respectively.
 $K = 4$ and $L = 4$ were chosen as they minimize BIC.
@@
 \appendix}
+
+\hypertarget{pathway-reg}{%
+\section{Groundwater management policies as a potential causal pathway}\label{pathway-reg}}
+
+The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}.
+This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers' water use decisions are not constrained by such regulations even in extreme drought years such as 2012 \citep{foster2019assessing}.
+
+Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022importance, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
+The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
+While this overlaps with the final 4 years of our study period, the LEMA area covers only parts of two counties and hence represents a very small proportion of our overall study sample while the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
+not drought years).
+Moreover, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining}.
+These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
+
+Areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction due to laws that typically provide farmers with absolute rights to use groundwater under their land coupled with low levels of metering that constrain potential for implementing extensive and binding pumping restrictions \citep{wheeler2016lessons, closas2018chronicle}.
 
@@
 Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
+This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
 We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
edits

@@
 
-  \textcolor{blue}{The other changes made to Equation 2 include the addition state-year FE, in order to provide stronger controls than just including year FE.
+  \textcolor{blue}{The other changes made to Equation 2 include the addition of state-year FE, in order to provide stronger controls than just including year FE.
 In addition, we choose not to allow the interactions of $WB$ and $AT$.
@@
 Please see the list below.
-We added the following paragraphs as Appendix A in the latest manuscript.
+We added the following paragraphs as Appendix A in this round of manuscript.}
 
   \begin{enumerate}
-      \item The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher (Rimsaite and Brozovic, 2023).
-This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers' water use decisions are not constrained by such regulations even in extreme drought years such as 2012 (Foster et al., 2019).
-      \item Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas (Marston et al., 2022; Rimsaite and Brozoic, 2023), these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
+      \item \textcolor{blue}{The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher (Rimsaite and Brozovic, 2023).
+This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers' water use decisions are not constrained by such regulations even in extreme drought years such as 2012 (Foster et al., 2019).}
+      \item \textcolor{blue}{Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas (Marston et al., 2022; Rimsaite and Brozoic, 2023), these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
 The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
@@
 Moreover, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions (Drysdale and Hendricks, 2018; Deines et al., 2021).
-These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
-      \item Areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction due to laws that typically provide farmers with absolute rights to use groundwater under their land coupled with low levels of metering that constrain potential for implementing extensive and binding pumping restrictions (Wheeler et al., 2016; Closas and Molle, 2018).
+These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.}
+      \item \textcolor{blue}{Areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction due to laws that typically provide farmers with absolute rights to use groundwater under their land coupled with low levels of metering that constrain potential for implementing extensive and binding pumping restrictions (Wheeler et al., 2016; Closas and Molle, 2018).}
   \end{enumerate}
-  }
 
@@
   \textcolor{blue}{We recognize your point that this may be potentially too strong of a claim.
-To more carefully summarise the implications of our choice of fixed effects, we have modified the text on lines XXX-XXX for our revision of the statement.}
+To more carefully summarise the implications of our choice of fixed effects, we have modified the text on lines 430-435 for our revision of the statement.}
 
@@
 These problems are now fixed.
-Please note that because of the width (range) of the histrogram bins, it looks like the bins at the left and right ends are beyond the range of the main figure, but this is not the case and is simply a plotting feature.}
+Please note that because of the width (range) of the histogram bins, it looks like the bins at the left and right ends are beyond the range of the main figure, but this is not the case and is simply a plotting feature.}
 
@@
 
-\textcolor{blue}{``WB'' and ``ST'' terms have been changed to italics throughout}
+\textcolor{blue}{Instead of making $WB$ and $AT$ un-italicized, we simply replaced them with $D$ and $A$, respectively for all of their instances.}
 
@@
 However, as we now use state-year FE represented by $\theta_{s,t}$, the ``$\mu_t$' should have been removed.
-Equation 2 no longer includes $mu_t$.}
+Equation 2 no longer includes $\mu_t$.}
 

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/manuscript.tex
edits

@@
 
-Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers capacity to adapt successfully to drought and rainfall variability.
+Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers' capacity to adapt successfully to drought and rainfall variability.
 Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact}.
@@
 First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers ability to fully meet crop water requirements during periods of more severe or extended water deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
+This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended water deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
 In contrast, yield continues to increase gradually as water deficit increases beyond 400 mm for counties with the highest level of aquifer thickness.
@@
 One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
+Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
 
@@
 Similar but less pronounced effects are also observed for soybean.
-This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers ability to effectively and reliably buffer crops against water deficits during the growing season.
+This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers' ability to effectively and reliably buffer crops against water deficits during the growing season.
 
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers choices about irrigated area shares in response to different water deficit and aquifer conditions.
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
 See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
@@
 
-We demonstrate the important role of aquifer thickness in determining farmers ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States.
+We demonstrate the important role of aquifer thickness in determining farmers' ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States.
 Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes, luan2021combined}.
@@
 Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally.
-Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and climatic water deficits during the growing season is highly non-linear when accounting for farmers decisions about irrigated production areas.
+Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and climatic water deficits during the growing season is highly non-linear when accounting for farmers' decisions about irrigated production areas.
 This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon, butler2020charting, elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
@@
 
-Second, our analysis considered the effects of aquifer thickness on farmers decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types or irrigation management practices.
+Second, our analysis considered the effects of aquifer thickness on farmers' decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types or irrigation management practices.
 Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact, deines2019quantifying, manning2017producer}.
@@
 However, where water deficits are distributed unevenly during the season, negative impacts of aquifer depletion may be even larger.
-This is because reductions in well yields caused by aquifer depletion will constrain farmers ability to meet peak crop water demands, increasing yield losses compared with if water deficits had been distributed evenly throughout the season \citep{ortiz2019unpacking}.
+This is because reductions in well yields caused by aquifer depletion will constrain farmers' ability to meet peak crop water demands, increasing yield losses compared with if water deficits had been distributed evenly throughout the season \citep{ortiz2019unpacking}.
 Furthermore, irrigation also has broader benefits for crops such as reducing heat-related stress through cooling of land surface temperatures \citep{adegoke2003impact, bonfils2007empirical, lobell2008effect, zhu2022untangling}.
@@
 To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without the prior requirement to assume any functional forms.
-This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers irrigated area choices under water supply constraints that have been noted in prior work \citep{foster2015well, foster2015analysis, hrozencik2017heterogeneous}.
+This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers' irrigated area choices under water supply constraints that have been noted in prior work \citep{foster2015well, foster2015analysis, hrozencik2017heterogeneous}.
 The model is given in Equation \eqref{eq:eqn2} below:
@@
 
+Some parts of the HPA region face regulatory pumping restrictions in response to declines in aquifer thickness.
+Resulting groundwater use reductions can in turn negatively affect crop yield.
+This causal pathway of aquifer thickness affecting crop yield is distinct from the pathways via well yield discussed in the \ref{main} section.
+However, it is unlikely that our regression analysis capture this effect.
+
 The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}.
@@
 
-Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022importance, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
+Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
 The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.

### 2023-09-12 Writing/NatureWater/2ndRoundRevision/replies/cover_letter.tex
edits

@@
 Consequently, our results changed slightly.
-Our main conclusions and qualitative results are robust and did not change.
-    \item We added histograms of aquifer thickness in Figure 3.
+But, our main conclusions and qualitative results are robust and did not change.
+    \item We added histograms of aquifer thickness in Figure 3 as suggested by reviewer 3.
+    \item We added another section in Appendix (Appendix A) that discusses groundwater regulations in response to declines in aquifer thickness as a causal pathway through which aquifer thickness affects yield.
+This is in response to a comment from reviewer 3.
 \end{itemize}

### 2023-09-13 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex
second round submission (finalized)

@@
 \textcolor{blue}{We agree. Instead of county-level "average" water deficit, in our revised manuscript we now say county-level annual water deficit because they are not averaged values.}
+
+\textcolor{blue}{We also changed "evident in county-years with large water deficits" in line 290 in the previous manuscript to "evident in years with large water deficits," which is in line 275 in the current manuscript.}
 
@@
 This step has been the same throughout all versions of our manuscript, and is not a change that has been introduced in current or prior revisions.
-We have added the following sentence on line X of our methods to clarify this point:}
+We have added the following sentence on line 341 of our methods to clarify this point:}
 
@@
 Please see the list below.
-We added the following paragraphs as Appendix A in this round of manuscript.}
+We added the following paragraphs to a new Appendix A in this round of manuscript, as we feel these are important points to include but are constrained by word limits in the main manuscript.}
 
@@
   \textcolor{blue}{We recognize your point that this may be potentially too strong of a claim.
-To more carefully summarise the implications of our choice of fixed effects, we have modified the text on lines 430-435 for our revision of the statement.}
+To more carefully summarise the implications of our choice of fixed effects, we have modified the text on lines 448-453 for our revision of the statement.}
 
@@
 
-  \textcolor{blue}{We also removed the sentence below that was state on lines 345 - 346 in the previous version of our manuscript to reflect that we can only hypothesize about the likely differences between using state and county fixed effects:}
+  \textcolor{blue}{We also removed the sentence below that was on lines 345 - 346 in the previous version of our manuscript to reflect that we can only hypothesize about the likely differences between using state and county fixed effects:}
 
@@
 
+\textcolor{blue}{We apologize for this mistake. This time, all the minor comments are addressed.}
+
 \item (M1.a) P. 15: ``WB'' and ``ST'' [now ``AT''] should not be in italics, unless ``W'', ``B'', ``S'', and ``T'' are separate variables
@@
 
-\textcolor{blue}{We changed $y_{i,t}$ in equation (1) to $y_{i,t}$.}
+\textcolor{blue}{We changed $y_{i,t}$ in equation (1) to $Y_{i,t}$.}
 

### 2023-09-13 Writing/NatureWater/2ndRoundRevision/manuscript.tex
second round submission (finalized)

@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
 $K$ is the total number of spline basis functions.
@@
 
-Some parts of the HPA region face regulatory pumping restrictions in response to declines in aquifer thickness.
-Resulting groundwater use reductions can in turn negatively affect crop yield.
-This causal pathway of aquifer thickness affecting crop yield is distinct from the pathways via well yield discussed in the \ref{main} section.
-However, it is unlikely that our regression analysis capture this effect.
+Our analysis evaluates the links between aquifer thickness and drought resilience, proposing a causal link between declining aquifer thickness and irrigated production during droughts.
+In the first section of our manuscript, we propose that this connection is driven by impacts of aquifer thickness on well yields and, hence, farmers ability to use groundwater to fully buffer crops against water deficits during drought.
 
-The majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period, and farmers have also been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}.
-This enables farmers to increase water withdrawals substantially in drought years when demand is higher, and prior research has shown that farmers' water use decisions are not constrained by such regulations even in extreme drought years such as 2012 \citep{foster2019assessing}.
+It is important to assess the extent to which this causal relationship may be affected or confounded by other factors.
+One key alternative hypothesis is that regulations limiting groundwater withdrawals could have been introduced as a response to declining aquifer thickness, with resulting reductions in irrigated areas and irrigated crop yields driven by regulatory rather than hydrologic (i.e.
+well yield) constraints.
+Below, we discuss the reasons that regulatory constraints are highly unlikely to be the cause of the relationships between aquifer thickness and drought impacts observed in our analysis for the HPA.
 
-Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to 2016.
-The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in NW Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
-While this overlaps with the final 4 years of our study period, the LEMA area covers only parts of two counties and hence represents a very small proportion of our overall study sample while the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
+First, it is important the majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period.
+It is also common for farmers to have been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}, further increasing their regulatory quotas above defined caps.
+Together, these factors mean that farmers in parts of the HPA where groundwater withdrawals are regulated or restricted are still able to increase water withdrawals substantially in drought years when demand is higher.
+Prior research has shown that as a result farmers water use decisions in HPA are not constrained by such regulations even in extreme drought years such as 2012 \citep{foster2019assessing}.
+
+Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to only 2016.
+The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in Northwest Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
+While this overlaps with the final 4 years of our study period, the Sheridan 6 LEMA area covers only parts of two counties in Kansas and hence represents a very small proportion of our overall study sample.
+Additionally, the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
 not drought years).
-Moreover, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining}.
+Furthermore, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining}.
 These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
 
-Areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction due to laws that typically provide farmers with absolute rights to use groundwater under their land coupled with low levels of metering that constrain potential for implementing extensive and binding pumping restrictions \citep{wheeler2016lessons, closas2018chronicle}.
+Finally, we also highlight that areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction.
+This is due to laws that typically provide farmers with absolute rights to use groundwater under their land, coupled with low levels of metering.
+Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigators in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
+In contrast, as discussed in the \ref{main} section of our manuscript, there is extensive evidence highlighting the reductions in well yields that have occurred in these parts of the HPA along with anecdotal, theoretical and simulation based evidence for why these restrictions would impose meaningful restrictions on farmers ability to adequately meet crop water needs during droughts.
 

### 2023-10-06 Writing/manuscript.tex
minor edits

@@
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
-Second, depletion also reduces the aquifer thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
+Second, depletion also reduces the thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
 This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
-If the reduction in well yields is substantial - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well, rouhi2020downside}.
+If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well, rouhi2020downside}.
 The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-In this paper, we present empirical evidence about how aquifer depletion affects farmers ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
+In this paper, we present empirical evidence about how aquifer depletion affects farmers' ability to effectively and reliably buffer crops against drought using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
 Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater, lu2018crop}.
@@
 Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-year water deficit values included in regressions for each crop.}\label{fig:yield-response}
+The histograms on the top of each panel show the distribution of county-level water deficit values included in regressions for each crop.}\label{fig:yield-response}
 \end{figure}
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 130m to 0.61 (61\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.89 (89\%) for an average county-level aquifer thickness of 130m to 0.72 (72\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}
@@
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only 0.11 tonne/ha of production loss for corn.
+At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only 0.14 tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in 0.77 tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in 2.09 tonne/ha of production loss for corn.
+Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in 0.94 tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in 1.44 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 
-Figure \ref{fig:test-sat-share} presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 130m.
+Figure \ref{fig:test-sat-share} presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100m.
 

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
edits made

---
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---

# Reply to Reviewer 1

I have read the second revision of this manuscript and the authors' responses to my last review. I think the authors have done an excellent job in responding to the points I raised in that review. Thus, I recommend acceptance after the authors address the very minor points that I noted while reading this version of the manuscript. Those points are as follows:

\textcolor{blue}{Thank you for your feedback on our manuscript. We have addressed all your comments and suggestions, and our replies to your comments can be found in blue below (original comments in black).}

+ Footnote 1 – change "saturated thickness" to "aquifer thickness" to be consistent with usage elsewhere. Same for x-axis label on Figure C.2.

\textcolor{blue}{Thank you for catching this. In the footnote, "saturated thickness" was replaced with "aquifer thickness".}

+ Line 110 – can you briefly describe how this "accounting" is done?

\textcolor{blue}{The figure is designed to showcase the effect of aquifer thickness on the percentage of irrigated acres, without the influence of spatio-temporal variations in weather conditions and drought risk exposure across the aquifer. In deriving the relationship between aquifer thickness and the proportion of irrigated acres using our model, these factors are kept constant, allowing us to focus solely on the impact of aquifer thickness. We have opted to omit the latter part of the sentence to avoid potential confusion. While Figure 1 was constructed using a similar procedure—adjusting only the water deficit while keeping other factors unchanged — we did not explicitly state this either. We believe this omission avoids unnecessary confusion.}

+ Line 147 – "leading in" should be "leading to".

\textcolor{blue}{Thank you for catching this. In the footnote, "leading in" was replaced with "leading to".}

+ Lines 282-283 – delete "of the aquifer" - redundant.

\textcolor{blue}{Thank you for catching this. In the sentence, "of the aquifer" was removed to avoid redundancy.}

+ Line 289 – "thickness" should be "thicknesses".

\textcolor{blue}{Thank you for catching this. In the sentence, "thickness" was replaced with "thicknesses".}

+ Figure 7 – need a 200 m label for scale bar on soybean map. Are the blank areas in the aquifer in Texas and Oklahoma an indication that the aquifer is thin or that there is a lack of data? Please clarify.

+ Line 308 – "y" should be "Y".

\textcolor{blue}{Thank you for catching this. The suggested changes has been made.}

+ Lines 357-359 – delete sentence as it is repeated on lines 363-364 where it fits better.

\textcolor{blue}{Thank you for catching this. The sentenced was removed.}

+ Lines 401-402 – a recent paper in Agricultural Water Management addresses this issue in more detail 
(https://www.sciencedirect.com/science/article/pii/S0378377423002731).

\textcolor{blue}{Thank you for suggesting this article. We now cite this article at the end of the sentence.}

+ Lines 410-413 – very awkward sentence – please rewrite.

\textcolor{blue}{We have rewritten the sentence to the following:}

\textcolor{blue}{"As discussed in section 1 of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA. This evidence ranges from anecdotal and theoretical insights to simulation-based studies. Collectively, they suggest that these restrictions significantly constrain farmers' ability to meet crop water demands during droughts."}

+ Line 446 – delete "across".

\textcolor{blue}{Thank you for catching this. The suggested changes has been made.}

# Reply to Reviewer 3

The second round comments have been addressed convincingly. The discussion of the role of groundwater management policies in Appendix A is particularly useful addition, as it clarifies how the results presented here relate to the findings of analyses which focused more on policy factors. Ruling out regulations as a substantial causal factor based on the limited temporal overlap with the study period is coherent. The other edits implemented in response to the other comments have further improved the strength and accuracy of the presentation.

Two relatively minor points remain to be addressed, based on one of the responses and the data and code correction you mentioned:

\textcolor{blue}{Thank you for your feedback on our manuscript. We have addressed all your comments and suggestions, and our replies to your comments can be found in blue below (original comments in black).}

+ (1) I would recommend a minor addition regarding the response to my second round comment 5.b. You have pointed out that the claim of a significant yield decline in lines 77-78 is supported by Figure A.1 (and also A.2 for soybeans). Please add a reference to both figures in this sentence to clarify for readers where they can find evidence for the statistical significance.

\textcolor{blue}{Thank you for this suggestion. We modify the sentence as follows:}

\textcolor{blue}{As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases (see Figures A.1 and A.2 for the 95\% confidence intervals)}

+ (2) The beginning of the response letter describes a code and data correction unrelated to the previous reviewer comments. "[Counties] with only irrigated acres" were removed from the irrigation share regression. The immediate acknowledgement of this correction is appreciated. I trust that it was based on sensible considerations, though the rationale does not immediately become clear from the short statement, especially given that this does not seem to have been relevant previously. As you point out, the result changes due to this correction are quantitatively relatively minor (most visible in Figs. 3 and 4), and do not affect the qualitative findings or conclusions.

\textcolor{blue}{We apologize for the confusion we caused. Our sentnce on this issue was not clear in the last round of revision. We meants that we excluded counties where only the irrigated acres are "observed" because only irrigated acres were reported by USDA, but not rainfed acres. So, it is very much possible that those counties had rainfed production, but rainfed acres were simply not reported by USDA for the counties we dropped. So, we were not trying to drop the fully irrigated counties. Of course, they may be fully irrigated and actually had zero rainfed production though it is very unlikely. If we keep them, they would all have irrigated share of 1 (100\%) just because drayland acres were not repoted by USDA, which can be a significant over-estimation of their actual share of irrigated acres. Therefore, we should drop them to avoid the problem lowering the quality of the data and consequently analysis. This is more of a data limitation than anything.}

I recommend, however, making sure that the correction, its rationale, and potential implications for the result interpretation are made clear to readers. Lines 347-348 state that "irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations". This statement covers the removal of fully irrigated counties, but could be clearer: Lines 348-350 mention the removal of "solely rainfed" counties explicitly - mentioning this but not the removal of fully irrigated counties could be confusing to readers. Please include an explicit statement about the removal of fully irrigated counties to avoid confusion. Please also make clear whether this removal pertains to both regressions (as implied in lines 347) or just the irrigated share regression (the one highlighted in the letter).

\textcolor{blue}{We modified the pertinent sentences to the following:}

\textcolor{blue}{Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations. Counties with solely irrigated or rainfed production observation data (Section 4.1) are used solely for estimating impacts of drought on per-area crop yields (Equation 1).}

Please also provide a short rationale, e.g. whether this is due to the model selected or substantive considerations. Finally, please clarify what this implies for the result interpretation. Fully irrigated county observations do not seem obviously irrelevant. The change in Fig. 3 results due to the correction is quantitatively minor but not irrelevant. It would, therefore, be helpful for readers to understand whether the exclusion of fully irrigated counties makes the results more accurate, or whether it creates a (necessary) bias.

\textcolor{blue}{Please see our replies about the rational. We simply cannot calculate irrigated area share for counties with only irrigated or ranifed observations.}

Relatedly, it seems that in Fig. 7, some of the previously displayed counties in both the irrigated (color shaded) and rainfed (grey) categories are now missing. Did the exclusion of fully irrigated counties also affect which rainfed areas are represented? If so, please also clarify this link for the readers.

\textcolor{blue}{Thank you so much for noticing this change. This has nothing to do with the exclusion of counties with only irrigation acres reported. The maps were supposed to present saturated thickness for "all" the counties used in the analysis. So, even the counties that were dropped for the irrigation share analyis should have been present. After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}

1. \textcolor{blue}{In the 1st round, we made a correction to the lower bound of saturated thickness according to reviewer 1's comment. We stated that 9m was used as the lower bound as below.}

\textcolor{blue}{"Aquifer thickness values were only estimated for counties with irrigated yield observations. Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [Fenichel et al., 2016, Haacker et al., 2016, Deines et al., 2020]."}

\textcolor{blue}{However, we were actually using 12m as the lower bound, which we realized after seeing reviewer's comment on this in the first round:}

\textcolor{blue}{"Figure 2 - Why do the plots stop at a thickness of about 12 m when you had earlier defined the minimum thickness as 9 m and mention 10 m in the text when discussing the relationships?"}

\textcolor{blue}{Therefore, we change the code so that 9m is used instead of 12m. We, however, did not update the map according to this change in the data. This is why the map in this version has several more counties in Kansas and New Mexico that did not appear in the map of the first round revision. They used to be dropped prior to chaging the threshold from 12m to 9m.}

2. \textcolor{blue}{The significant change in the map in the second round from the 1st round is a mistake on using the map we created when we were experimenting with codes. We then did not rerun the code to create a map with the whole dataset to overwrite the wrong map with the correct map. We made sure that we streamlined all the codes, rerun all of them, and cofirm that the map indeed reflect what we used for our analysis in this round. As stated in the manuscript, all the codes and data are available on a publicly accessible Github repository. Thank you again for your keen eyes on details. We would not have noticed this ourselves.}

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.tex
edits made

--- 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
@@
----
-output:
-  pdf_document: 
-    keep_tex: true
-  html_document: default
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{lmodern}
+\usepackage{amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
-# Reply to Reviewer 1
+\author{}
+\date{\vspace{-2.5em}}
 
-I have read the second revision of this manuscript and the authors' responses to my last review.
-I think the authors have done an excellent job in responding to the points I raised in that review.
-Thus, I recommend acceptance after the authors address the very minor points that I noted while reading this version of the manuscript.
-Those points are as follows:
+\begin{document}
+
+\hypertarget{reply-to-reviewer-1}{%
+\section{Reply to Reviewer 1}\label{reply-to-reviewer-1}}
+
+I have read the second revision of this manuscript and the authors'
+responses to my last review. I think the authors have done an excellent
+job in responding to the points I raised in that review. Thus, I
+recommend acceptance after the authors address the very minor points
+that I noted while reading this version of the manuscript. Those points
+are as follows:
 
@@
 
-+ Footnote 1 – change "saturated thickness" to "aquifer thickness" to be consistent with usage elsewhere. Same for x-axis label on Figure C.2.
+\begin{itemize}
+\tightlist
+\item
+  Footnote 1 -- change ``saturated thickness'' to ``aquifer thickness''
+  to be consistent with usage elsewhere. Same for x-axis label on Figure
+  C.2.
+\end{itemize}
 
@@
 
-+ Line 110 – can you briefly describe how this "accounting" is done?
+\begin{itemize}
+\tightlist
+\item
+  Line 110 -- can you briefly describe how this ``accounting'' is done?
+\end{itemize}
 
@@
 
-+ Line 147 – "leading in" should be "leading to".
+\begin{itemize}
+\tightlist
+\item
+  Line 147 -- ``leading in'' should be ``leading to''.
+\end{itemize}
 
@@
 
-+ Lines 282-283 – delete "of the aquifer" - redundant.
+\begin{itemize}
+\tightlist
+\item
+  Lines 282-283 -- delete ``of the aquifer'' - redundant.
+\end{itemize}
 
@@
 
-+ Line 289 – "thickness" should be "thicknesses".
+\begin{itemize}
+\tightlist
+\item
+  Line 289 -- ``thickness'' should be ``thicknesses''.
+\end{itemize}
 
@@
 
-+ Figure 7 – need a 200 m label for scale bar on soybean map.
-Are the blank areas in the aquifer in Texas and Oklahoma an indication that the aquifer is thin or that there is a lack of data?
-Please clarify.
-
-+ Line 308 – "y" should be "Y".
+\begin{itemize}
+\item
+  Figure 7 -- need a 200 m label for scale bar on soybean map. Are the
+  blank areas in the aquifer in Texas and Oklahoma an indication that
+  the aquifer is thin or that there is a lack of data? Please clarify.
+\item
+  Line 308 -- ``y'' should be ``Y''.
+\end{itemize}
 
@@
 
-+ Lines 357-359 – delete sentence as it is repeated on lines 363-364 where it fits better.
+\begin{itemize}
+\tightlist
+\item
+  Lines 357-359 -- delete sentence as it is repeated on lines 363-364
+  where it fits better.
+\end{itemize}
 
@@
 
-+ Lines 401-402 – a recent paper in Agricultural Water Management addresses this issue in more detail 
-(https://www.sciencedirect.com/science/article/pii/S0378377423002731).
+\begin{itemize}
+\tightlist
+\item
+  Lines 401-402 -- a recent paper in Agricultural Water Management
+  addresses this issue in more detail
+  (\url{https://www.sciencedirect.com/science/article/pii/S0378377423002731}).
+\end{itemize}
 
@@
 
-+ Lines 410-413 – very awkward sentence – please rewrite.
+\begin{itemize}
+\tightlist
+\item
+  Lines 410-413 -- very awkward sentence -- please rewrite.
+\end{itemize}
 
@@
 
-+ Line 446 – delete "across".
+\begin{itemize}
+\tightlist
+\item
+  Line 446 -- delete ``across''.
+\end{itemize}
 
@@
 
-# Reply to Reviewer 3
+\hypertarget{reply-to-reviewer-3}{%
+\section{Reply to Reviewer 3}\label{reply-to-reviewer-3}}
 
-The second round comments have been addressed convincingly.
-The discussion of the role of groundwater management policies in Appendix A is particularly useful addition, as it clarifies how the results presented here relate to the findings of analyses which focused more on policy factors.
-Ruling out regulations as a substantial causal factor based on the limited temporal overlap with the study period is coherent.
-The other edits implemented in response to the other comments have further improved the strength and accuracy of the presentation.
+The second round comments have been addressed convincingly. The
+discussion of the role of groundwater management policies in Appendix A
+is particularly useful addition, as it clarifies how the results
+presented here relate to the findings of analyses which focused more on
+policy factors. Ruling out regulations as a substantial causal factor
+based on the limited temporal overlap with the study period is coherent.
+The other edits implemented in response to the other comments have
+further improved the strength and accuracy of the presentation.
 
-Two relatively minor points remain to be addressed, based on one of the responses and the data and code correction you mentioned:
+Two relatively minor points remain to be addressed, based on one of the
+responses and the data and code correction you mentioned:
 
@@
 
-+ (1) I would recommend a minor addition regarding the response to my second round comment 5.b.
-You have pointed out that the claim of a significant yield decline in lines 77-78 is supported by Figure A.1 (and also A.2 for soybeans).
-Please add a reference to both figures in this sentence to clarify for readers where they can find evidence for the statistical significance.
+\begin{itemize}
+\item
+  \begin{enumerate}
+  \def\labelenumi{(\arabic{enumi})}
+  \tightlist
+  \item
+    I would recommend a minor addition regarding the response to my
+    second round comment 5.b. You have pointed out that the claim of a
+    significant yield decline in lines 77-78 is supported by Figure A.1
+    (and also A.2 for soybeans). Please add a reference to both figures
+    in this sentence to clarify for readers where they can find evidence
+    for the statistical significance.
+  \end{enumerate}
+\end{itemize}
 
@@
 
-+ (2) The beginning of the response letter describes a code and data correction unrelated to the previous reviewer comments.
-"[Counties] with only irrigated acres" were removed from the irrigation share regression.
-The immediate acknowledgement of this correction is appreciated.
-I trust that it was based on sensible considerations, though the rationale does not immediately become clear from the short statement, especially given that this does not seem to have been relevant previously.
-As you point out, the result changes due to this correction are quantitatively relatively minor (most visible in Figs.
-3 and 4), and do not affect the qualitative findings or conclusions.
+\begin{itemize}
+\item
+  \begin{enumerate}
+  \def\labelenumi{(\arabic{enumi})}
+  \setcounter{enumi}{1}
+  \tightlist
+  \item
+    The beginning of the response letter describes a code and data
+    correction unrelated to the previous reviewer comments.
+    ``{[}Counties{]} with only irrigated acres'' were removed from the
+    irrigation share regression. The immediate acknowledgement of this
+    correction is appreciated. I trust that it was based on sensible
+    considerations, though the rationale does not immediately become
+    clear from the short statement, especially given that this does not
+    seem to have been relevant previously. As you point out, the result
+    changes due to this correction are quantitatively relatively minor
+    (most visible in Figs. 3 and 4), and do not affect the qualitative
+    findings or conclusions.
+  \end{enumerate}
+\end{itemize}
 
@@
 
-I recommend, however, making sure that the correction, its rationale, and potential implications for the result interpretation are made clear to readers.
-Lines 347-348 state that "irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations".
-This statement covers the removal of fully irrigated counties, but could be clearer: Lines 348-350 mention the removal of "solely rainfed" counties explicitly - mentioning this but not the removal of fully irrigated counties could be confusing to readers.
-Please include an explicit statement about the removal of fully irrigated counties to avoid confusion.
-Please also make clear whether this removal pertains to both regressions (as implied in lines 347) or just the irrigated share regression (the one highlighted in the letter).
+I recommend, however, making sure that the correction, its rationale,
+and potential implications for the result interpretation are made clear
+to readers. Lines 347-348 state that ``irrigated area share and average
+crop yield estimates use data only from counties that have both
+irrigated and rainfed production observations''. This statement covers
+the removal of fully irrigated counties, but could be clearer: Lines
+348-350 mention the removal of ``solely rainfed'' counties explicitly -
+mentioning this but not the removal of fully irrigated counties could be
+confusing to readers. Please include an explicit statement about the
+removal of fully irrigated counties to avoid confusion. Please also make
+clear whether this removal pertains to both regressions (as implied in
+lines 347) or just the irrigated share regression (the one highlighted
+in the letter).
 
@@
 
-Please also provide a short rationale, e.g.
-whether this is due to the model selected or substantive considerations.
-Finally, please clarify what this implies for the result interpretation.
-Fully irrigated county observations do not seem obviously irrelevant.
-The change in Fig.
-3 results due to the correction is quantitatively minor but not irrelevant.
-It would, therefore, be helpful for readers to understand whether the exclusion of fully irrigated counties makes the results more accurate, or whether it creates a (necessary) bias.
+Please also provide a short rationale, e.g.~whether this is due to the
+model selected or substantive considerations. Finally, please clarify
+what this implies for the result interpretation. Fully irrigated county
+observations do not seem obviously irrelevant. The change in Fig. 3
+results due to the correction is quantitatively minor but not
+irrelevant. It would, therefore, be helpful for readers to understand
+whether the exclusion of fully irrigated counties makes the results more
+accurate, or whether it creates a (necessary) bias.
 
@@
 
-Relatedly, it seems that in Fig.
-7, some of the previously displayed counties in both the irrigated (color shaded) and rainfed (grey) categories are now missing.
-Did the exclusion of fully irrigated counties also affect which rainfed areas are represented?
-If so, please also clarify this link for the readers.
+Relatedly, it seems that in Fig. 7, some of the previously displayed
+counties in both the irrigated (color shaded) and rainfed (grey)
+categories are now missing. Did the exclusion of fully irrigated
+counties also affect which rainfed areas are represented? If so, please
+also clarify this link for the readers.
 
@@
 
-1. \textcolor{blue}{In the 1st round, we made a correction to the lower bound of saturated thickness according to reviewer 1's comment. We stated that 9m was used as the lower bound as below.}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textcolor{blue}{In the 1st round, we made a correction to the lower bound of saturated thickness according to reviewer 1's comment. We stated that 9m was used as the lower bound as below.}
+\end{enumerate}
 
@@
 
-2.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
   \textcolor{blue}{The significant change in the map in the second round from the 1st round is a mistake on using the map we created when we were experimenting with codes.
@@
 We would not have noticed this ourselves.}
+\end{enumerate}
+
+\end{document}

### 2023-10-09 Codes/test/map_of_removed_counties.rmd
edits made

```{r}
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    #--- saturated thickness has to be at least 9 meters ---#
    sat_thld_m = 9,
    #--- at least 75% of total county area has to be overlapped with HPA ---#
    ir_share_thld = 0.75,
    #--- drop obs with extreme balance ---#
    balance_thld = c(-1200, 200),
    #--- number of saturated thickness categories ---#
    sat_cat_num = 3,
    #--- select states of focus ---#
    # no observations are dropped with the current set of states
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop")
```

```{r}
sc_list <-
  reg_data$data[[1]] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
  .[comp %in% c(1, 2), ] %>%
  .[sat_cat != "dryland", .(sc_code, geometry, comp)] %>%
  unique(by = "sc_code") %>%
  st_as_sf()

reg_data$data[[1]] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
  .[comp == 1, ] %>%
  .[sat_cat != "dryland", ] %>%
  .[, .N, by = sc_code]

ggplot(data = sc_list) +
  geom_sf(aes(fill = factor(comp)))

reg_data$data[[1]] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
  .[comp == 2, ] %>%
  .[sat_cat != "dryland", ] %>%
  .[, unique(sc_code)]
```

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/manuscript.tex
edits made

--- 2023-09-13 Writing/NatureWater/2ndRoundRevision/manuscript.tex
@@
 
-Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}\footnote{Another pathway is through groundwater use regulation in response to low or declining saturated thickness.
+Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}\footnote{Another pathway is through groundwater use regulation in response to low or declining aquifer thickness.
 However, it is unlikely that this effects is captured in our analysis.
@@
 In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean.
-As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases.
+As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases (see Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} for the 95\% confidence intervals).
 This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling, li2020quantifying}.
@@
 Standard errors of this analysis are estimated using bootstrap methods clustered by county.
-Note that irrigated area share and average crop yield estimates use  data only from counties that have both irrigated and rainfed production observations.
-Counties with solely rainfed production observation data (Section \ref{data-sets}) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \ref{eq:eqn1}).
+Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations.
+Counties with solely irrigated or rainfed production observation data (Section \ref{data-sets}) are used solely for estimating impacts of drought on per-area crop yields (Equation \ref{eq:eqn1}).
 
@@
 not drought years).
-Furthermore, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining}.
+Furthermore, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining, whittemore2023108408}.
 These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
@@
 Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigators in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
-In contrast, as discussed in the \ref{main} section of our manuscript, there is extensive evidence highlighting the reductions in well yields that have occurred in these parts of the HPA along with anecdotal, theoretical and simulation based evidence for why these restrictions would impose meaningful restrictions on farmers ability to adequately meet crop water needs during droughts.
+As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
+This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
+Collectively, they suggest that these restrictions significantly constrain farmers' ability to meet crop water demands during droughts.
 

### 2023-10-09 Codes/0_7_merge_datasets.rmd
forgot to commit

--- 2023-08-30 Codes/0_6_merge_datasets.rmd
@@
   .[, num_obs := .N, by = sc_code] %>%
-  # === keep only if more than 2 observations are available ===#
+  # === keep only if more than 10 observations are available ===#
   .[num_obs >= 10, ] %>%

### 2023-10-09 Codes/2_figures_tables.rmd
forgot to commit

@@
 all_results <- readRDS("Shared/Results/all_results.rds")
+
+data_corn <- all_results[crop == "corn", data][[1]]
+data_soy <- all_results[crop == "soy", data][[1]]
 ```
@@
 ```{r}
-data_corn <- all_results[crop == "corn", data][[1]]
-data_soy <- all_results[crop == "soy", data][[1]]
-
 wd_data <-
@@
 
-# === all counties ===#
 data(fips_codes, package = "tidycensus")
 
-all_counties <-
-  tigris::counties() %>%
-  st_simplify(dTolerance = 100) %>%
-  rename(
-    state_code = STATEFP,
-    county_code = COUNTYFP,
-    county_name = NAME
-  ) %>%
-  left_join(., fips_codes, by = c("state_code", "county_code")) %>%
-  dplyr::select(state_code, state, county_code, county_name) %>%
-  mutate(sc_code = paste0(state_code, county_code)) %>%
-  data.table()
+# all_counties <-
+#   tigris::counties() %>%
+#   st_simplify(dTolerance = 100) %>%
+#   rename(
+#     state_code = STATEFP,
+#     county_code = COUNTYFP,
+#     county_name = NAME
+#   ) %>%
+#   left_join(., fips_codes, by = c("state_code", "county_code")) %>%
+#   dplyr::select(state_code, state, county_code, county_name) %>%
+#   mutate(sc_code = paste0(state_code, county_code)) %>%
+#   data.table()
+# saveRDS(all_counties, "Shared/Data/counties.rds")
 
-all_states <-
-  tigris::states() %>%
-  rename(state_code = STATEFP) %>%
-  left_join(
-    .,
-    unique(data.table(fips_codes), by = "state"),
-    by = "state_code"
-  )
+all_counties <- readRDS("Shared/Data/counties.rds")
+
+# all_states <-
+#   tigris::states() %>%
+#   rename(state_code = STATEFP) %>%
+#   left_join(
+#     .,
+#     unique(data.table(fips_codes), by = "state"),
+#     by = "state_code"
+#   )
+
+# saveRDS(all_states, "Shared/Data/states.rds")
+
+all_states <- readRDS("Shared/Data/states.rds")
 

### 2023-10-09 Writing/NatureWater/2ndRoundRevision/manuscript.tex
forgot to commit

@@
 
-Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
+Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean.
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
@@
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading in 1.57 tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.57 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
-These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935.
+These data provide annual estimates of the aquifer thickness at 250m resolution since 1935.
 Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
@@
 Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
+Furthermore, during the aggregation process, any aquifer cells within a county that had an aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thicknesses lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
 Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
@@
 
-where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(D\) is seasonal water deficit.
+where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(Y\) is crop yield, and \(D\) is seasonal water deficit.
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
@@
 Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same across when using different possible sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
 Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.

### 2023-10-09 Codes/Functions/functions.R

@@ -26,7 +26,6 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
         .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
         .[, state_year := paste0(state_name, "_", year)] %>%
         .[, num_obs := .N, by = sc_dry] %>%
-        .[num_obs > 5, ] %>%
         .[, num_obs := NULL]
     )) %>%
     # === sat breaks for grouping ===#

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/cover_letter.tex
cover letter modified

--- 2023-08-31 Writing/NatureWater/2ndRoundRevision/replies/cover_letter.tex
@@
 
-The vast majority of the changes we made are minor. Some notable changes are the following:
-
-\begin{itemize}
-    \item Fix a coding error that was creating invalid observations for our regressions analysis of the share of irrigated acres.
-Consequently, our results changed slightly.
-Our main conclusions and qualitative results are robust and did not change.
-    \item We added histograms of aquifer thickness in Figure 3.
-\end{itemize}
-
 We believe that these changes have substantially strengthen both the methodological approaches in our manuscript, as well the implications of our findings for groundwater conservation and management.

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
final edits by TM before Tim reviews

@@
 Please clarify.
+
+\textcolor{blue}{200m lable was added to the legend of the soybean map.
+The blank area are counties that were not used at all in either of the yield and irrigation share regressions because neither of them are reported by USDA.
+}
 
@@
 Our sentnce on this issue was not clear in the last round of revision.
-We meants that we excluded counties where only the irrigated acres are "observed" because only irrigated acres were reported by USDA, but not rainfed acres.
-So, it is very much possible that those counties had rainfed production, but rainfed acres were simply not reported by USDA for the counties we dropped.
+We meant that we excluded counties where only the irrigated acres are "observed" because only irrigated acres were reported by USDA, but not rainfed acres.
+It is very much possible that those counties had rainfed production, but rainfed acres were simply not reported by USDA for the counties we dropped.
+This means that we cannot even calculate the share of irrigation for those counties (This is also the case for counties with only rainfed production observations).
 So, we were not trying to drop the fully irrigated counties.
 Of course, they may be fully irrigated and actually had zero rainfed production though it is very unlikely.
-If we keep them, they would all have irrigated share of 1 (100\%) just because drayland acres were not repoted by USDA, which can be a significant over-estimation of their actual share of irrigated acres.
+Previously before fixing the coding error, those counties were assigned the value of 1 (100\%) as their share of irrigated acres just because drayland acres were not repoted by USDA, which can be a significant over-estimation of their actual share of irrigated acres.
 Therefore, we should drop them to avoid the problem lowering the quality of the data and consequently analysis.
@@
 We then did not rerun the code to create a map with the whole dataset to overwrite the wrong map with the correct map.
-We made sure that we streamlined all the codes, rerun all of them, and cofirm that the map indeed reflect what we used for our analysis in this round.
-As stated in the manuscript, all the codes and data are available on a publicly accessible Github repository.
-Thank you again for your keen eyes on details.
-We would not have noticed this ourselves.}
+In this round, we made sure that we streamlined all the codes, rerun all of them, and cofirmed that the map indeed reflect what we used for our analysis in this round.
+Thank you again for your keen eyes on details.}

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.tex
final edits by TM before Tim reviews

@@
 \begin{itemize}
+\tightlist
 \item
@@
   the aquifer is thin or that there is a lack of data? Please clarify.
+\end{itemize}
+
+\textcolor{blue}{200m lable was added to the legend of the soybean map.
+The blank area are counties that were not used at all in either of the yield and irrigation share regressions because neither of them are reported by USDA.
+}
+
+\begin{itemize}
+\tightlist
 \item
@@
 We then did not rerun the code to create a map with the whole dataset to overwrite the wrong map with the correct map.
-We made sure that we streamlined all the codes, rerun all of them, and cofirm that the map indeed reflect what we used for our analysis in this round.
-As stated in the manuscript, all the codes and data are available on a publicly accessible Github repository.
-Thank you again for your keen eyes on details.
-We would not have noticed this ourselves.}
+In this round, we made sure that we streamlined all the codes, rerun all of them, and cofirmed that the map indeed reflect what we used for our analysis in this round.
+Thank you again for your keen eyes on details.}
 \end{enumerate}

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/manuscript.tex
final edits by TM before Tim reviews

@@
 Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
+Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thicknesses lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
 Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
@@
 Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
-Aquifer thickness data were obtained from \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/} \citep{haacker2023}.
 The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data.

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
minor edit
edit

@@
 
-\textcolor{blue}{Please see our replies about the rational. We simply cannot calculate irrigated area share for counties with only irrigated or ranifed observations.}
+\textcolor{blue}{Please see our replies about the rational.}
 
@@
 This has nothing to do with the exclusion of counties with only irrigation acres reported.
-The maps were supposed to present saturated thickness for "all" the counties used in the analysis.
+The maps were supposed to present "all" the counties used in the analysis.
 So, even the counties that were dropped for the irrigation share analyis should have been present.

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
edit

@@
 The maps were supposed to present "all" the counties used in the analysis.
-So, even the counties that were dropped for the irrigation share analyis should have been present.
-After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}
+So, even the counties that were dropped for the irrigation share analyis should have been present.}
+
+\textcolor{blue}{After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}
 

### 2023-10-09 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
minor edit

@@
 
-\textcolor{blue}{"Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [Fenichel et al., 2016, Haacker et al., 2016, Deines et al., 2020]."}
+\textcolor{blue}{"Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than \textcolor{red}{9} m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [Fenichel et al., 2016, Haacker et al., 2016, Deines et al., 2020]."}
 

### 2023-10-19 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
minor edits

@@
 We, however, did not update the map according to this change in the data.
-This is why the map in this version has several more counties in Kansas and New Mexico that did not appear in the map of the first round revision.
+This is why the map in the current version has several more counties in Kansas and New Mexico that did not appear in the map of the first round revision.
 They used to be dropped prior to chaging the threshold from 12m to 9m.}
@@
 2.
-\textcolor{blue}{The significant change in the map in the second round from the 1st round is a mistake on using the map we created when we were experimenting with codes.
+\textcolor{blue}{The significant change in the map in the second round from the 1st round is simply an embarassing mistake of using the map we created when we were experimenting with the R codes.
 We then did not rerun the code to create a map with the whole dataset to overwrite the wrong map with the correct map.

### 2023-10-19 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.tex

@@
 
-\textcolor{blue}{200m lable was added to the legend of the soybean map.
-The blank area are counties that were not used at all in either of the yield and irrigation share regressions because neither of them are reported by USDA.
+\textcolor{blue}{The 200m label has been added to the legend of the soybean map.
+The blank area are counties that were not used in either of the yield and irrigation share regressions.
+This because neither of them have reported information on corn and/or soybean production from USDA, due to the small areas of land dvoted to production of these crops in these locations.
 }
@@
 
-\textcolor{blue}{Thank you for catching this. The suggested changes has been made.}
+\textcolor{blue}{Thank you for catching this. The suggested change has been made.}
 
@@
 
-\textcolor{blue}{We have rewritten the sentence to the following:}
+\textcolor{blue}{We have rewritten the sentence to now state the following, which we feel improves the clarity of the text:}
 
@@
 
-\textcolor{blue}{Thank you for this suggestion. We modify the sentence as follows:}
+\textcolor{blue}{Thank you for this suggestion. We have modified the sentence as follows:}
 
@@
 \textcolor{blue}{We apologize for the confusion we caused.
-Our sentnce on this issue was not clear in the last round of revision.
-We meant that we excluded counties where only the irrigated acres are "observed" because only irrigated acres were reported by USDA, but not rainfed acres.
-It is very much possible that those counties had rainfed production, but rainfed acres were simply not reported by USDA for the counties we dropped.
-This means that we cannot even calculate the share of irrigation for those counties (This is also the case for counties with only rainfed production observations).
-So, we were not trying to drop the fully irrigated counties.
-Of course, they may be fully irrigated and actually had zero rainfed production though it is very unlikely.
-Previously before fixing the coding error, those counties were assigned the value of 1 (100\%) as their share of irrigated acres just because drayland acres were not repoted by USDA, which can be a significant over-estimation of their actual share of irrigated acres.
-Therefore, we should drop them to avoid the problem lowering the quality of the data and consequently analysis.
-This is more of a data limitation than anything.}
+Our description of this issue was not clear in the last round of revision.
+We meant to highlight that we excluded counties where only the irrigated acres are "observed" because only irrigated production data were reported by USDA - typically this is because the total number of rainfed acres is small or was not captured by USDA surveys in the county.
+In these cases, we cannot  calculate the share of irrigated acres for those specific counties.
+Previously before fixing the coding error, those counties were assigned the value of 1 (100\%) for their share of irrigated acres because no rainfed production data was repoted by USDA, which may lead to an over-estimation of their actual share of irrigated acres.
+Therefore, in our revisions we removed these counties to avoid this issue and the consequent reduction of data quality.
+However, as previously, this did not alter the findings of our analysis, and therefore this is largely a cosmetic change to ensure we base our analysis on the highest quality data available.}
 
@@
 
-\textcolor{blue}{We modified the pertinent sentences to the following:}
+\textcolor{blue}{We modified the pertinent sentences to now state the following:}
 
 \textcolor{blue}{Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations.
-Counties with solely irrigated or rainfed production observation data (Section 4.1) are used solely for estimating impacts of drought on per-area crop yields (Equation 1).}
+Counties with solely irrigated or rainfed production observation data (Section 4.1) are used solely for estimating impacts of drought on per-area crop irrigated or rainfed yields respectively (Equation 1).}
 
@@
 
-\textcolor{blue}{Please see our replies about the rational. We simply cannot calculate irrigated area share for counties with only irrigated or ranifed observations.}
+\textcolor{blue}{Please see our replies above that provide detail on the rational behind this change.
+In summary, the change was made to avoid introducing assumptions about the irrigated area share for counties where rainfed production is not reported.
+It is not possible to calculate irrigated area share for counties with only irrigated or rainfed observations.
+We believe that ommitting counties lacking either of these data is preferable to prior assumptions made (e.g.
+that there is no irrigated or rainfed production if no data is reported in a county), as these assumptions may introduce noise into our analysis - for example if rainfed production is not reported due to sampling decisions made in the generation of estimates by USDA.}
 
@@
 
-\textcolor{blue}{Thank you so much for noticing this change.
+\textcolor{blue}{Thank you for noticing this change.
 This has nothing to do with the exclusion of counties with only irrigation acres reported.
-The maps were supposed to present saturated thickness for "all" the counties used in the analysis.
-So, even the counties that were dropped for the irrigation share analyis should have been present.
-After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}
+The maps were intended to present saturated thickness for "all" the counties used in the analysis.
+So, even the counties that were dropped for the irrigation share analsyis should have been present.
+The figure has been corrected to reflect this, and information added to the caption to explain this distinction.}
 

### 2023-10-19 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.rmd
final edits of replies to reviewers

@@
 
-\textcolor{blue}{200m lable was added to the legend of the soybean map.
-The blank area are counties that were not used at all in either of the yield and irrigation share regressions because neither of them are reported by USDA.
+\textcolor{blue}{The 200m label has been added to the legend of the soybean map.
+The blank area are counties that were not used in either of the yield and irrigation share regressions.
+This is because neither of them have reported information on corn and/or soybean production from USDA, due to the small areas of land dvoted to production of these crops in these locations.
 }
@@
 \textcolor{blue}{We apologize for the confusion we caused.
-Our sentnce on this issue was not clear in the last round of revision.
-We meant that we excluded counties where only the irrigated acres are "observed" because only irrigated acres were reported by USDA, but not rainfed acres.
-It is very much possible that those counties had rainfed production, but rainfed acres were simply not reported by USDA for the counties we dropped.
-This means that we cannot even calculate the share of irrigation for those counties (This is also the case for counties with only rainfed production observations).
-So, we were not trying to drop the fully irrigated counties.
-Of course, they may be fully irrigated and actually had zero rainfed production though it is very unlikely.
-Previously before fixing the coding error, those counties were assigned the value of 1 (100\%) as their share of irrigated acres just because drayland acres were not repoted by USDA, which can be a significant over-estimation of their actual share of irrigated acres.
-Therefore, we should drop them to avoid the problem lowering the quality of the data and consequently analysis.
-This is more of a data limitation than anything.}
+Our description of this issue was not clear in the last round of revision.
+We meant to highlight that we excluded counties where only the irrigated acres are "observed" because only irrigated production data were reported by USDA - typically this is because the total number of rainfed acres is small or was not captured by USDA surveys in the county.
+In these cases, we cannot calculate the share of irrigated acres for those specific counties.
+Previously before fixing the coding error, those counties were assigned the value of 1 (100\%) for their share of irrigated acres because no rainfed production data was repoted by USDA, which may lead to an over-estimation of their actual share of irrigated acres.
+Therefore, in our revisions we removed these counties to avoid this issue and the consequent reduction of data quality.
+However, as previously noted, this did not alter the findings of our analysis, and therefore this is largely a cosmetic change to ensure we base our analysis on the highest quality data available.}
 
@@
 
-\textcolor{blue}{Please see our replies about the rational.}
+\textcolor{blue}{Please see our replies above that provide detail on the rational behind this change.
+In summary, the change was made to avoid introducing assumptions about the irrigated area share for counties where rainfed production is not reported.
+It is not possible to calculate irrigated area share for counties with only irrigated or rainfed observations.
+We believe that ommitting counties lacking either of these data is preferable to prior assumptions made (e.g.
+that there is no irrigated or rainfed production if no data is reported in a county), as these assumptions may introduce noise into our analysis - for example if rainfed production is not reported due to sampling decisions made in the generation of estimates by USDA.}
 
@@
 
-\textcolor{blue}{Thank you so much for noticing this change.
+\textcolor{blue}{Thank you for noticing this change.
 This has nothing to do with the exclusion of counties with only irrigation acres reported.
-The maps were supposed to present "all" the counties used in the analysis.
-So, even the counties that were dropped for the irrigation share analyis should have been present.}
+The maps were intended to present aquifer thickness for "all" the counties used in the analysis.
+So, even the counties that were dropped for the irrigation share analsyis should have been present.}
 
-\textcolor{blue}{After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}
+\textcolor{blue}{We made an error of inserting a figure that originated during experimentation with our codes in the last round of revision.
+The figure has now been corrected.
+Additionally, please note that the current version includes several more counties in Kansas and New Mexico that did not appear in the map from the initial revision.
+In the first round of revision, we adjusted aquifer thickness threshold from 12m to 9m in response to a comment by reviewer 1.
+Due to this, we were using those counties in our analysis.
+However, we did not modify the map accordingly at that time.
+However, this change in the data has now been accurately reflected in the map.}
+
+<!-- \textcolor{blue}{After some digging, we found that we actually made mistakes twice on map creation during the whole review process so far.}
 
@@
 In this round, we made sure that we streamlined all the codes, rerun all of them, and cofirmed that the map indeed reflect what we used for our analysis in this round.
-Thank you again for your keen eyes on details.}
+Thank you again for your keen eyes on details.} -->

### 2023-10-19 Writing/NatureWater/3rdRoundRevision/replies/reply_to_reviewers_R3.tex
final edits of replies to reviewers

@@
 ]{article}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
-  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 The blank area are counties that were not used in either of the yield and irrigation share regressions.
-This because neither of them have reported information on corn and/or soybean production from USDA, due to the small areas of land dvoted to production of these crops in these locations.
+This is because neither of them have reported information on corn and/or soybean production from USDA, due to the small areas of land dvoted to production of these crops in these locations.
 }
@@
 Therefore, in our revisions we removed these counties to avoid this issue and the consequent reduction of data quality.
-However, as previously, this did not alter the findings of our analysis, and therefore this is largely a cosmetic change to ensure we base our analysis on the highest quality data available.}
+However, as previously noted, this did not alter the findings of our analysis, and therefore this is largely a cosmetic change to ensure we base our analysis on the highest quality data available.}
 
@@
 This has nothing to do with the exclusion of counties with only irrigation acres reported.
-The maps were intended to present saturated thickness for "all" the counties used in the analysis.
-So, even the counties that were dropped for the irrigation share analsyis should have been present.
-The figure has been corrected to reflect this, and information added to the caption to explain this distinction.}
+The maps were intended to present aquifer thickness for "all" the counties used in the analysis.
+So, even the counties that were dropped for the irrigation share analsyis should have been present.}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  \textcolor{blue}{In the 1st round, we made a correction to the lower bound of saturated thickness according to reviewer 1's comment. We stated that 9m was used as the lower bound as below.}
-\end{enumerate}
-
-\textcolor{blue}{"Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thickness lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production [Fenichel et al., 2016, Haacker et al., 2016, Deines et al., 2020]."}
-
-\textcolor{blue}{However, we were actually using 12m as the lower bound, which we realized after seeing reviewer's comment on this in the first round:}
-
-\textcolor{blue}{"Figure 2 - Why do the plots stop at a thickness of about 12 m when you had earlier defined the minimum thickness as 9 m and mention 10 m in the text when discussing the relationships?"}
-
-\textcolor{blue}{Therefore, we change the code so that 9m is used instead of 12m.
-We, however, did not update the map according to this change in the data.
-This is why the map in this version has several more counties in Kansas and New Mexico that did not appear in the map of the first round revision.
-They used to be dropped prior to chaging the threshold from 12m to 9m.}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  \textcolor{blue}{The significant change in the map in the second round from the 1st round is a mistake on using the map we created when we were experimenting with codes.
-We then did not rerun the code to create a map with the whole dataset to overwrite the wrong map with the correct map.
-In this round, we made sure that we streamlined all the codes, rerun all of them, and cofirmed that the map indeed reflect what we used for our analysis in this round.
-Thank you again for your keen eyes on details.}
-\end{enumerate}
+\textcolor{blue}{We made an error of inserting a figure that originated during experimentation with our codes in the last round of revision.
+The figure has now been corrected.
+Additionally, please note that the current version includes several more counties in Kansas and New Mexico that did not appear in the map from the initial revision.
+In the first round of revision, we adjusted aquifer thickness threshold from 12m to 9m in response to a comment by reviewer 1.
+Due to this, we were using those counties in our analysis.
+However, we did not modify the map accordingly at that time.
+However, this change in the data has now been accurately reflected in the map.}
 

### 2023-10-19 Writing/NatureWater/3rdRoundRevision/manuscript.tex
final edits of replies to reviewers

@@
 Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations.
-Counties with solely irrigated or rainfed production observation data (Section \ref{data-sets}) are used solely for estimating impacts of drought on per-area crop yields (Equation \ref{eq:eqn1}).
+Counties with solely irrigated or rainfed production observation data (Section \ref{data-sets}) are used solely for estimating impacts of drought on per-area crop irrigated or rainfed yields respectively (Equation \ref{eq:eqn1}).
 

### 2023-10-31 Codes/0_2_quickstat.Rmd
clean up the codes

@@
 ---
-title: "Get crop yield and acreage data: Corn and Soybeans"
+title: "Get crop yield and acreage data from USDA-NASS: Corn and Soybeans"
 ---
@@
 # Preparation
-Read the counties data
 
 ```{r }
-base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
-```
+#--- load the counties ---#
+base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))
 
-Get the list of states
-
-```{r }
+#--- get the list of states ---#
 all_states <-
@@
   toupper()
-```
 
-Define a function to get QuickStat data
-
-```{r }
-get_quickstat_data <- function(year, data_item) {
-  desc <-
-    tidyUSDA::allDataItem %>%
-    grep(pattern = data_item, ., value = TRUE)
-
-  temp_data <-
-    tidyUSDA::getQuickstat(
-      key = key_get("usda_nass_qs_api"),
-      program = "SURVEY",
-      data_item = desc,
-      geographic_level = "COUNTY",
-      state = all_states,
-      year = year,
-      geometry = FALSE
-    ) %>%
-    #--- keep only some of the variables ---#
-    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
-    rename("state_code" = "state_fips_code")
-
-  return(temp_data)
-}
-```
-
-List of items to download
-
-```{r }
+#--- list of items to download ---#
 item_list <-
@@
 
-# tidyUSDA::allDataItem %>%
-#     grep(pattern = "PLANTED", ., value = TRUE) %>%
-#     grep(pattern = "SOYBEANS", ., value = TRUE) %>%
-#     grep(pattern = "IRRIGATED", ., value = TRUE)
-
-# for (i in 1:length(item_list)) {
-#   desc <-
-#     tidyUSDA::allDataItem %>%
-#     grep(pattern = data_item, ., value = TRUE)
-#   print(desc)
-# }
-```
-
-List of year-item combinations
-
-```{r }
+#--- list of year-item combinations ---#
 par_list <-
@@
 
-# === save the data ===#
-saveRDS(nass_data, here("Shared/Data/RawData/nass_data_raw.rds"))
+#--- save the data ---#
+saveRDS(nass_data, here("Data/data-raw/nass_data_raw.rds"))
 ```
@@
   nass_data_wide <-
-    here("Shared/Data/RawData/nass_data_raw.rds") %>%
+    here("Data/data-raw/nass_data_raw.rds") %>%
     readRDS() %>%
@@
 
-Quick Summary
-
-```{r }
-library(modelsummary)
-datasummary(
-  crop * ir * (Mean + SD)
-  ~ acres + yield,
-  data = nass_data_wide
-)
-```
-
 # Save the data
@@
 ```{r eval = F}
-saveRDS(nass_data_wide, here("Shared/Data/ProcessedData/nass_data.rds"))
+saveRDS(nass_data_wide, here("Data/data-processed/nass_data.rds"))
 ```

### 2023-10-31 Codes/0_3_get_sat.Rmd
clean up the codes

@@
 ---
-title: "Assign saturated thickness value "
+title: "Assign saturated thickness values"
 author: "Taro Mieno"
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = FALSE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi=400,
-  fig.width=7.5,
-  fig.height=5,
-  out.width="750px",
-  out.height="500px"
-)
-```
-
 # Preparation
@@
 ```{r }
-sat_temp <- raster("/Volumes/Elements/Database/SaturatedThickness/SatThick_19851.tif")
+sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
 ```
@@
 hpa_counties <-
-  here("Shared/Data/ProcessedData/base_counties.rds") %>% 
+  here("Data/data-processed/base_counties.rds") %>%
   readRDS() %>%
@@
   sat_temp <-
-    paste0("/Volumes/Elements/Database/SaturatedThickness/SatThick_", y, "1.tif") %>% 
+    paste0("Data/data-raw/saturated_thickness/SatThick_", y, "1.tif") %>%
     raster()
@@
 ```{r extract-sat, eval = F}
-sat_all <- mclapply(1985:2016, get_sat, mc.cores = 12) %>%
+sat_all <-
+  mclapply(1985:2016, get_sat, mc.cores = 12) %>%
   rbindlist() %>%
-  .[sat < 0, sat := 0] %>% 
-  .[, sat := sat * 3.28084] # meter to feet
+  .[sat < 0, sat := 0]
 ```
@@
 ```{r save, eval = F}
-saveRDS(sat_all, here("Shared/Data/ProcessedData/sat_all.rds"))
+saveRDS(sat_all, here("Data/data-processed/sat_all.rds"))
 ```
-
-# Quick visualization
-
-```{r }
-sat_all <- 
-  here("Shared/Data/ProcessedData/sat_all.rds") %>% 
-  readRDS()
-```
-
-Saturate thickness history by county
-
-```{r }
-ggplot(sat_all) +
-  geom_line(aes(y = sat, x = year, group = sc_code))
-```
-
-Histogram
-
-```{r }
-ggplot(sat_all) +
-  geom_histogram(aes(x = sat))
-```

### 2023-10-31 Codes/0_4_get_weather.Rmd
clean up the codes

---
title: "Get Daymet weather data and calculate ET0"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

# Preparation

```{r}
# === counties ===#
counties <-
  here("Data/data-processed/base_counties.rds") %>%
  readRDS() %>%
  st_as_sf()

# === elevation ===#
elevation <-
  here("Data/data-raw/metdata_elevationdata.nc") %>%
  rast()
```

# Download gridMET data

## Set up parameters

```{r}
# === list of weather variables ===#
var_ls <-
  c(
    "pr", # precipitation
    "rmin", # min relative humidity (%)
    "rmax", # max relative humidity (%)
    "srad", # radiation (W/M^2)
    "tmmn", # tmin (Kelvin)
    "tmmx", # tmax (Kelvin)
    "vs", # wind speed,
    "pet" # et
  )

# === list of years ===#
year_ls <- 1985:2018

# === var-year data ===#
par_data <-
  expand.grid(
    var_name = var_ls,
    year = year_ls
  ) %>%
  data.table() %>%
  .[, var_name := as.character(var_name)]
```

## Download gridMET data

The total size of the gridMET netcdf files is about 35 GB.

```{r}
# mclapply(
#   seq_len(nrow(par_data)),
#   function(x) get_grid_MET(par_data[x, var_name], par_data[x, year]),
#   mc.cores = 12
# )
lapply(
  seq_len(nrow(par_data)),
  function(x) get_grid_MET(par_data[x, var_name], par_data[x, year])
)
```

# Calculate ET0 and summarize by county-date

## Setup
```{r}
#--- counties ---#
counties <- filter(counties, !(state %in% c("MO", "IL", "IN", "IA")))

#--- get the CRS of a gridMET data ---#
sample_gmet <- rast(here("Data/data-raw/gridMET/pr_1985.nc"))
gmet_crs <- crs(sample_gmet)
```

## Calculate ET0

For each variable-year combination, area-weighted average is calculated for each county, where area refers to the area of prism grids that are intersecting with the county polygon. This operations can take a while.

```{r}
if (!file.exists(here("data/data-processed", "gridMET.rds"))) {
  all_data <-
    mclapply(
      seq_len(nrow(counties)),
      function(x) {
        print(x)

        lapply(
          year_ls,
          function(y) get_et0_county_year(counties[x, ], y)
        ) %>%
          rbindlist() %>%
          return()
      },
      mc.cores = parallel::detectCores() - 4
    ) %>%
    rbindlist()

  saveRDS(all_data, here("Data/data-processed", "gridMET.rds"))
}
```

# Summarize the data for regression

```{r}
#--- read the raw gridMET data ---#
gm_data <- readRDS(here("Data/data-processed", "gridMET.rds"))

#--- summarize the data ---#
sum_gm_data <-
  gm_data %>%
  .[month(date) >= 5 & month(date) <= 9, ] %>%
  .[, t_min_max_2 := pmin(pmax((tmin + tmax) / 2 - 10, 0), 20)] %>%
  .[, .(
    # === total balance ===#
    balance = sum(et0 - prcp),
    # === total et0 ===#
    et0 = sum(et0),
    # === total precip ===#
    prcp = sum(prcp),
    # === GDD ===#,
    gdd = sum(t_min_max_2),
    # === EDD ===#
    edd = sum(pmax(0, tmax - 32)),
    # === number of days above x ===#
    days_ab_30 = sum(tmax >= 30),
    # === mean temp ===#
    mean_temp = mean(tmin + tmax) / 2
  ),
  by = .(sc_code, year(date))
  ]

saveRDS(sum_gm_data, here("Data/data-processed", "summrized_gridMET.rds"))
```

### 2023-10-31 Codes/0_5_ssurgo.Rmd
clean up the codes

---
title: "Prepare SSURGO Data"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

# Prepare

## Read the counties sf
```{r}
all_counties <-
  readRDS(here("Data/data-processed/base_counties.rds")) %>%
  st_as_sf() %>%
  dplyr::mutate(geometry = st_make_valid(geometry)) %>%
  st_transform(4326)
```

# Download and summarize

```{r download, cache = TRUE}
vars <- c("sandtotal_r", "silttotal_r", "awc_r")

lapply(
  1:nrow(all_counties),
  \(x) {
    print(x)
    file_name <- paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds")
    if (!file.exists(file_name)) {
      temp <- get_ssurgo_props(field = all_counties[x, ], vars)
      saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds"))
    } else {
      temp <- readRDS(file_name)
      if (ncol(temp) < 4) {
        temp <- get_ssurgo_props(field = all_counties[x, ], vars)
        saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds"))
      }
    }
  }
)

# parallel::mclapply(
#   1:nrow(all_counties),
#   \(x) {
#     print(x)
#     file_name <- here::here(paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds"))
#     if (!file.exists(file_name)) {
#       print(paste0("working on ", x))
#       temp <- get_ssurgo_props(all_counties[x, ], vars)
#       saveRDS(temp, here::here(paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds")))
#     } else {
#       temp <- readRDS(file_name)
#       if (ncol(temp) < 7) {
#         temp <- get_ssurgo_props(field = all_counties[x, ], vars)
#         saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds"))
#       } else {
#         print(paste0("skipping ", x))
#       }
#     }
#   },
#   mc.cores = 6,
#   mc.preschedule = FALSE
# )

```

# save the SSURGO data

```{r }
ssurgo_soil <-
  fs::dir_ls("Data/data-raw/SSURGO") %>%
  purrr::map(readRDS) %>%
  rbindlist()

saveRDS(ssurgo_soil, here("Data/data-processed/ssurgo_soil.rds"))
```

### 2023-10-31 Codes/0_6_merge_datasets.rmd
clean up the codes

@@
 
+## Read the datasets
+
 ```{r}
-base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
-
-#' NDMC Drought index
-drought_county <- readRDS(here("Shared/Data/ProcessedData/DI_NDMC.rds"))
-
-#' SPEI
-# spei_data <- readRDS(here("Shared/Data/ProcessedData/spei_data.rds"))
+base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))
 
 #' Weather
-weather_data <- readRDS(here("Shared/Data/ProcessedData", "summrized_gridMET.rds"))
+weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))
 
 #' Saturated thickness
-sat_all <- readRDS(here("Shared/Data/ProcessedData/sat_all.rds"))
+sat_all <- readRDS(here("Data/data-processed/sat_all.rds"))
 
 #' Yield, acreage data
-nass_data <- readRDS(here("Shared/Data/ProcessedData/nass_data.rds"))
+nass_data <- readRDS(here("Data/data-processed/nass_data.rds"))
 
 #* SSURGO
-
-ssurgo_soil <- readRDS(here("Shared/Data/ProcessedData/ssurgo_soil.rds"))
+ssurgo_soil <- readRDS(here("Data/data-processed/ssurgo_soil.rds"))
 ```
@@
   sat_all[., on = c("sc_code", "year")] %>%
-  drought_county[., on = c("sc_code", "year")] %>%
   .[order(sc_code, year), ] %>%
   ssurgo_soil[, on = "sc_code"]
-
-# merged_data[ir == "nir", state] %>% unique()
-# merged_data[ir == "ir", state] %>% unique()
 ```
@@
   merged_data %>%
+  #--- yield unit conversion (bu/acre to ton/ha)---#
+  .[, yield := yield * 12.553 / 200] %>%
   # === if not in_hpa and irrigated, drop them ===#
@@
   .[!is.na(sat), ] %>%
+  #--- indicator of whether dryland (rainfed) production or not ---#
+  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
+  #--- sc_code and dry_or_not (used for FE regression) ---#
+  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
+  #--- state-year combination ---#
+  .[, state_year := paste0(state_name, "_", year)] %>%
   # === number of observations by state-county code ===#
-  .[, num_obs := .N, by = sc_code] %>%
-  # === keep only if more than 2 observations are available ===#
+  .[, num_obs := .N, by = sc_dry] %>%
+  # .[, num_obs := .N, by = sc_code] %>%
+  # === keep only if more than 10 observations are available ===#
   .[num_obs >= 10, ] %>%
@@
 ```{r}
-saveRDS(final_data, here("Shared/Data/ProcessedData/final_data.rds"))
+saveRDS(final_data, here("Data/data-processed/final_data.rds"))
 ```

### 2023-10-31 Codes/1_Analysis.rmd
clean up the codes

@@
 final_data <-
-  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  here("Data/data-processed/final_data.rds") %>%
   # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
@@
     #--- drop obs with extreme balance ---#
-    balance_thld = c(-1200, 200),
+    balance_thld = c(-200, 1200),
     #--- number of saturated thickness categories ---#
@@
 
-saveRDS(main_analysis, "Shared/Results/main_analysis.rds")
+saveRDS(main_analysis, "Results/main_analysis.rds")
 ```
@@
 
-saveRDS(all_results, "Shared/Results/all_results.rds")
+saveRDS(all_results, "Results/all_results.rds")
 ```

### 2023-10-31 Codes/master.rmd
clean up the codes

# Data Preparation

```{r}
knitr::purl(here("Codes/0_1_hpa_counties.Rmd"), here("Codes/0_1_hpa_counties.R"))
knitr::purl(here("Codes/0_2_quickstat.Rmd"), here("Codes/0_2_quickstat.R"))
knitr::purl(here("Codes/0_3_get_sat.Rmd"), here("Codes/0_3_get_sat.R"))
knitr::purl(here("Codes/0_4_get_weather.Rmd"), here("Codes/0_4_get_weather.R"))
knitr::purl(here("Codes/0_5_ssurgo.Rmd"), here("Codes/0_5_ssurgo.R"))
knitr::purl(here("Codes/0_6_merge_datasets.Rmd"), here("Codes/0_6_merge_datasets.R"))
```

# Analysis

```{r}
knitr::purl(here("Codes/1_analysis.Rmd"), here("Codes/1_analysis.R"))
source(here("Codes/1_analysis.R"))
```

# Create figures

```{r}
knitr::purl(here("Codes/2_figures_tables.Rmd"), here("Codes/2_figures_tables.R"))
source(here("Codes/2_figures_tables.R"))
```

### 2023-10-31 Writing/manuscript.tex
clean up the codes

@@
 \usepackage{makecell}
+\usepackage{authblk}
 \usepackage{xcolor}
@@
 \title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
-\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of WWW, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
+
+\author[1]{Taro Mieno}
+\author[2]{Timothy Foster}
+\author[3]{Shunkei Kakimoto}
+\author[4]{Nicholas Brozovi\'{c}}
+\affil[1]{Department of Agricultural Economics, University of Nebraska Lincoln}
+\affil[2]{School of Mechanical, Aerospace and Civil Engineering, University of Manchester}
+\affil[3]{Department of Applied Economics, University of Minnesota}
+\affil[4]{Daugherty Water for Food Global Institute at the University of Nebraska}
+\affil[4]{Department of Agricultural Economics, University of Nebraska Lincoln}
+
 \date{}
@@
 
-Two pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
+Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}\footnote{Another pathway is through groundwater use regulation in response to low or declining aquifer thickness.
+However, it is unlikely that this effects is captured in our analysis.
+Please see Appendix \ref{pathway-reg}}.
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
@@
 This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
-If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well, rouhi2020downside}.
+If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
 The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
@@
 
-At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response} experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
+% At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
 In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_yield_intensive}
 
@@
 Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-level water deficit values included in regressions for each crop.}\label{fig:yield-response}
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
+Total number of observations are $8,773$ and $5,977$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_ir_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_ir_yield_dif}
 
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.89 (89\%) for an average county-level aquifer thickness of 130m to 0.72 (72\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings).
-Reductions in the share of production area under irrigation is also evident for soybean, although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100 m to 0.72 (72\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+Reductions in the share of production area under irrigation is also evident for soybean.
 
@@
 
-{\centering \includegraphics[width=5in,]{Figures/g_share}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share}
 
@@
 
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95 percent confidence interval)}\label{fig:ir-share}
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval).
+The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
+Total number of observations are $2,523$ and $1,558$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
 \end{figure}
@@
 
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 50 m, 90 m, and 130 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 350mm.
+Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 40 m, 70 m, and 100 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
+For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 400 mm.
 However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
-This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas and years where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600mm) and are smaller in overall magnitude.
-For example, for a seasonal water deficit of 950mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 130m and 10m aquifer thicknesses.
+This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas where aquifer thickness is lower.
+Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600 mm) and are smaller in overall magnitude.
+For example, for a seasonal water deficit of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 100 m and 10 m aquifer thicknesses.
 
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950mm, for example, reducing from 130m aquifer thickness to 90m (decline of 40m) would result in only 0.14 tonne/ha of production loss for corn.
+At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.09 tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 90m aquifer thickness to 50m (decline of 40m) would result in 0.94 tonne/ha of production loss for corn, with a further drop from 50m aquifer thickness to 10m (decline of 40m) leading in 1.44 tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.57 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
-This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers ability to effectively and reliably buffer crops against water deficits during the growing season.
+This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers' ability to effectively and reliably buffer crops against water deficits during the growing season.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_total_impact}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_total_impact}
 
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix A for estimated average yield response curves with 95 percent confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_avg_yield_dif}
 
@@
 
-\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 50 m, 90 m, and 130 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95 percent confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5 percent level of significance.}\label{fig:dif-tot-impact}
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}\label{fig:dif-tot-impact}
 \end{figure}
@@
 
+Soil characteristics data (sand percentage, silt percentage, and water holding capacity) were obtained from the Soil Survey Geographic
+Dataset (SSURGO).
+Observation units of the SSURGO data are polygons.
+For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables were calculated.
+
 \begin{figure}[H]
 
-{\centering \includegraphics[height=6in,]{Figures/g_y_wd_all}
+{\centering \includegraphics[width=750px,height=6in,]{../../Figures/g_y_wd_all}
 
@@
 
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water, haacker2023}.
-These data provide annual estimates of the aquifer thickness of the aquifer at 250m resolution since 1935.
+As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
+These data provide annual estimates of the aquifer thickness at 250m resolution since 1935.
 Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_map}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_map}
 
@@
 \begin{equation}
-y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k}^K \beta_{j,k}\phi_{j,k}(WB_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
+Y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(D_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
 \end{equation}
 
-where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit.
+where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(Y\) is crop yield, and \(D\) is seasonal water deficit.
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0-33.3%)
-\), \([33.3-66.7%)
-\), and \([66.7% - 100%]
-\) quantile ranges of aquifer thickness.
+The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0\%-33.3\%)\), \([33.3\%-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness.
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) represents the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+$K$ is the total number of spline basis functions.
+$K = 3$ was chosen for this regression as it minimizes Bayesian Information Criterion (BIC).
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
@@
 \begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k}^K \alpha_{k}\phi_{k}(WB_{i}) + \sum_{l}^L \beta_{l}\tau_{l}(AT_{i,t}) + X_i + \theta_{s,t} + \mu_t + v_{i,t} \label{eq:eqn2}
+    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \eta_{k}\phi_{k}(D_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(A_{i,t}) + X_i + \theta_{s,t} + v_{i,t} \label{eq:eqn2}
 \end{equation}
@@
 where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water balance is captured by \(\sum_{k}^K \alpha_{k}\phi_{k}(WB_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\alpha_{k}\) is its coefficient.
+The impact of average water balance is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(D_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
+$K$ and $L$ are the numbers of basis functions for $D$ and $A$, respectively.
+$K = 4$ and $L = 4$ were chosen as they minimize BIC.
 County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by \(X_i\).
@@
 The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
+In this combined model, aquifer thickness is a continuous variable for the irrigated area share regression component and a categorical variable for the irrigated yield regression following the format of Equations \eqref{eq:eqn2} and \eqref{eq:eqn1}, respectively.
 Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
 Standard errors of this analysis are estimated using bootstrap methods clustered by county.
-Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations.
-Counties with solely rainfed production observation data (Section \ref{data-sets}) are excluded from these analyses, and are used solely for estimating impacts of drought on per-area crop yields (Equation \eqref{eq:eqn1}).
+Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations.
+Counties with solely irrigated or rainfed production observation data (Section \ref{data-sets}) are used solely for estimating impacts of drought on per-area crop irrigated or rainfed yields respectively (Equation \ref{eq:eqn1}).
 
 \hypertarget{data-and-code-availability}{%
-\subsection{Data and Code Availability}\label{data-and-code-availability}}
+\subsection{Data Availability}\label{data-and-code-availability}}
 
@@
 
+\subsection{Code Availability}
+
+\clearpage
+
+\section*{Acknowledgements}
+
+Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.
+Timothy Foster was supported by Innovate UK award no.
+10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.'' The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
+
+\begin{itemize}
+\item OCE 58-0111-20-007 (TM, SK, NB)
+\item OCE 58-0111-21-007 (TM, SK, NB)
+\item Innovate UK award no. 10044695 (TF)
+\end{itemize}
+
+\subsection*{Author Contributions}
+
+\section*{Author Contribution Statement}
+
+\section*{Competing Interests Statement}
+
 \clearpage
@@
 \appendix}
+
+\hypertarget{pathway-reg}{%
+\section{Groundwater management policies as a potential causal pathway}\label{pathway-reg}}
+
+Our analysis evaluates the links between aquifer thickness and drought resilience, proposing a causal link between declining aquifer thickness and irrigated production during droughts.
+In the first section of our manuscript, we propose that this connection is driven by impacts of aquifer thickness on well yields and, hence, farmers ability to use groundwater to fully buffer crops against water deficits during drought.
+
+It is important to assess the extent to which this causal relationship may be affected or confounded by other factors.
+One key alternative hypothesis is that regulations limiting groundwater withdrawals could have been introduced as a response to declining aquifer thickness, with resulting reductions in irrigated areas and irrigated crop yields driven by regulatory rather than hydrologic (i.e.
+well yield) constraints.
+Below, we discuss the reasons that regulatory constraints are highly unlikely to be the cause of the relationships between aquifer thickness and drought impacts observed in our analysis for the HPA.
+
+First, it is important the majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period.
+It is also common for farmers to have been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}, further increasing their regulatory quotas above defined caps.
+Together, these factors mean that farmers in parts of the HPA where groundwater withdrawals are regulated or restricted are still able to increase water withdrawals substantially in drought years when demand is higher.
+Prior research has shown that as a result farmers water use decisions in HPA are not constrained by such regulations even in extreme drought years such as 2012 \citep{foster2019assessing}.
+
+Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to only 2016.
+The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in Northwest Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
+While this overlaps with the final 4 years of our study period, the Sheridan 6 LEMA area covers only parts of two counties in Kansas and hence represents a very small proportion of our overall study sample.
+Additionally, the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
+not drought years).
+Furthermore, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining, whittemore2023108408}.
+These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
+
+Finally, we also highlight that areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction.
+This is due to laws that typically provide farmers with absolute rights to use groundwater under their land, coupled with low levels of metering.
+Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigators in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
+As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
+This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
+Collectively, they suggest that these restrictions significantly constrain farmers' ability to meet crop water demands during droughts.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_corn}
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_soy}
 
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_avg_yield}
 
@@
 \hypertarget{test-dif-share}{%
-\section{Test of the Impact of Aquiefr Thickness on the Share of Irrigated Acres}\label{test-dif-share}}
+\section{Comparison of the impacts of different aquifer thickness levels on irrigated area share}\label{test-dif-share}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{B.\arabic{figure}}
+
+Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100 m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
+Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
+Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 100 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_share_dif}
+
+}
+
+\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 100 m (black line).
+The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
+\end{figure}
+
+\clearpage
+
+\hypertarget{county-fe}{%
+\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
 
@@
 
-Figure \ref{fig:test-sat-share} presents the difference in the share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100m.
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
+For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
 
@@
 
-{\centering \includegraphics[width=6.5in,]{Figures/g_share_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_variation}
 
@@
 
-\caption{Decline in the share of irrigated acres as a function of aquifer thickness relative to aquifer thickness level of 130m (The shaded areas represent the 95 percent confidence interval)}\label{fig:test-sat-share}
+\caption{Comparison of variations in aquifer thickness after demeaning by yearly average, state-year average, and county average.}\label{fig:variation-left}
 \end{figure}
 
-\hypertarget{county-fe}{%
-\section{County Fixed Effects in Irrigation Share Regression}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning them by its overall average (A), state average (B), and county average (C) to gain insights into how much variation for identification is left under no county or state fixed effects (A), state fixed effects (B), and county fixed effects (C).
-As you can see in the figure, while including state fixed effects do not reduce the total variation in aquifer thickness, including county fixed effects dramatically reduces its total variation.
-Consequently, we found that including county fixed effects results in extremely inaccurate estimates (high standard error).
-For this reason, we just included state fixed effects.
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
+Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_variation}
-
-}
-
-\caption{Comparison of variations in aquifer thikcness after demeaning by overall average, county average, and state average.}\label{fig:variation-left}
-\end{figure}
-
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of reduction in aquifer thickness on the share of irrigated acres are virtually the same across the sets of fixed effects.
-Importantly, adding state fixed effects, which control for unobservables at the state level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation is unlikely to cause significant bias to our estimation and findings.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,]{Figures/g_share_comp}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share_comp}
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_basis}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_basis}
 
@@
 Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
-Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate \(\beta_k\) to fit the data best.
+Unlike non-parametric regression, a GAM is a linear (in-parameter) model and estimates $\beta_k$ to best fit the observation data.
 
@@
 
-{\centering \includegraphics[width=6in,]{Figures/g_il_y}
+{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_il_y}
 

### 2023-10-31 Codes/0_1_hpa_counties.R

@@ -1,34 +1,13 @@
----
-title: "Prepare base-county data"
-author: "Taro Mieno"
-output:
-  html_document:
-    number_sections: yes
-    theme: flatly
-    highlight: zenburn
-    toc_float: yes
-    toc: yes
-    toc_depth: 3
-geometry: margin=1in
----
-
-# Identify counties the overlaps with the HPA boundary
-## Read the HPA border shapefile
-
-```{r Read-HPA, cache = TRUE}
+## ----Read-HPA, cache = TRUE-------------------------------------
 hpa <- st_read(here("Data/data-raw/hp_bound2010.shp"))
-```
 
-Map-of-HPA
 
-```{r, eval = FALSE}
-tm_shape(st_simplify(hpa)) +
-  tm_polygons()
-```
+## ---- eval = FALSE----------------------------------------------
+## tm_shape(st_simplify(hpa)) +
+##   tm_polygons()
 
-## Get all the US counties as an sf
 
-```{r Get-US-counties}
+## ----Get-US-counties--------------------------------------------
 (
   US_county <- tigris::counties(cb = TRUE) %>%
     st_as_sf() %>%
@@ -36,39 +15,30 @@ tm_shape(st_simplify(hpa)) +
     rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
     dplyr::select(state_code, county_code)
 )
-```
 
-## Identify counties that overlaps with `hpa`
 
-```{r Identify-overlap, cache = TRUE}
+## ----Identify-overlap, cache = TRUE-----------------------------
 hpa_counties <- US_county[hpa, ] %>%
   mutate(whole_area = st_area(.) %>% as.numeric())
-```
 
-Map-of-HPA and the intersecting counties
 
-```{r eval = FALSE}
-tm_shape(st_simplify(hpa)) +
-  tm_polygons() +
-  tm_shape(hpa_counties) +
-  tm_polygons(col = "blue", alpha = 0.4)
-```
+## ----eval = FALSE-----------------------------------------------
+## tm_shape(st_simplify(hpa)) +
+##   tm_polygons() +
+##   tm_shape(hpa_counties) +
+##   tm_polygons(col = "blue", alpha = 0.4)
 
-## Find the area of the portion that is intersecting with hpa
-Intersect HPA and counties in HPA 
 
-```{r intersect-HPA-county}
+## ----intersect-HPA-county---------------------------------------
 intersected_portion <-
[…]

### 2023-10-31 Codes/0_4_ndmc_di.R

@@ -1,44 +1,0 @@
-######################################
-# Drought Index
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#' According to Kuwayama (2018), the six key physical indicators consider percentiles
-#' from a soil moisture model, daily streamflow percentiles, the percent of normal precipitation,
-#' a standardized precipitation index, and a remotely-sensed vegetation
-#' health index (Svoboda et al. 2002). The authors also rely on supplementary indicators
-#' such as humidity and temperature departure from normal, reservoir, and lake levels,
-#' surface water supply indices, snowpack, and groundwater levels. The drought Index data is saved
-#' below:
-
-DI_1 <- fread(here("Data/RawData/NDMC_CO_IL_IN_DI_IA_KS_MS.csv"))
-DI_2 <- fread(here("Data/RawData/NDMC_DI_NM_OK_SD_TX_WY_NE.csv"))
-
-DI <- rbind(DI_1, DI_2)
-
-#' The default drought index is weekly persentage data. We changed them to monthly data by using the codes below.
-#' However, we are examing for different "seasons", from April to September, from May to October,
-#' from April to October, and from May to September. So we need to filter each season's drought index.
-
-drought_county <- DI %>%
-  .[, start_date := as.Date(ValidStart)] %>%
-  .[, end_date := as.Date(ValidEnd)] %>%
-  .[, FIPS := as.character(FIPS)] %>%
-  .[str_length(FIPS) == 4, FIPS := paste0("0", FIPS)] %>%
-  .[, state_code := str_sub(FIPS, 1, 2)] %>%
-  .[, county_code := str_sub(FIPS, 3, 5)] %>%
-  .[month(end_date) > 4 & month(start_date) < 10, ] %>%
-  # === state-county code ===#
-  .[, sc_code := paste0(state_code, county_code)] %>%
-  .[, .(
-    d0_5_9 = sum(D0) / 100,
-    d1_5_9 = sum(D1) / 100,
-    d2_5_9 = sum(D2) / 100,
-    d3_5_9 = sum(D3) / 100,
-    d4_5_9 = sum(D4) / 100
-  ), by = .(sc_code, year(end_date))] %>%
-  data.table()
-
-saveRDS(drought_county, here("Data/ProcessedData/DI_NDMC.rds"))

### 2023-10-31 Codes/1_analysis.R

@@ -1,24 +1,24 @@
----
-title: "Regression Analysis and Prediction"
-author: "Taro Mieno"
----
-
-# Prepare data
-
-```{r}
+## ----------------------------------------------------------
 #--- read data ---#
 final_data <-
-  here("Shared/Data/ProcessedData/final_data.rds") %>%
+  here("Data/data-processed/final_data.rds") %>%
+  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
   readRDS()
 
 #--- Data used for regression analysis ---#
 reg_data <-
   prepare_reg_data(
     data = final_data,
-    sat_thld_m = 9, # saturated thickness has to be at least 9 meters
-    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
-    balance_thld = c(-1200, 200),
+    #--- saturated thickness has to be at least 9 meters ---#
+    sat_thld_m = 9,
+    #--- at least 75% of total county area has to be overlapped with HPA ---#
+    ir_share_thld = 0.75,
+    #--- drop obs with extreme balance ---#
+    balance_thld = c(-200, 1200),
+    #--- number of saturated thickness categories ---#
     sat_cat_num = 3,
+    #--- select states of focus ---#
+    # no observations are dropped with the current set of states
     state_ls =
       list(
         corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
@@ -33,13 +33,18 @@ reg_data <-
 # ggplot(reg_data$data[[1]]) +
 #   geom_point(aes(y = acres_ratio, x = sat)) +
 #   geom_smooth(aes(y = acres_ratio, x = sat))
-```
 
-# Regression and prediciton
 
-## Yield
+## ----------------------------------------------------------
+# ir_share_data <- main_analysis$share_reg_data[[1]]
+# sat_seq <- main_analysis$sat_seq[[1]]
+# sandtotal_e <- main_analysis$sandtotal_med[[1]]
+# silttotal_e <- main_analysis$silttotal_med[[1]]
+# awc_e <- main_analysis$awc_med[[1]]
 
-```{r}
+# main_analysis$share_reg_data[[1]]
+# main_analysis$share_reg_data[[2]]
+
 main_analysis <-
   reg_data %>%
   rowwise() %>%
[…]

### 2023-10-31 Codes/2_figures_tables.R

@@ -1,21 +1,11 @@
----
-title: "Create Figures and Tables"
-author: "Taro Mieno"
----
-
-# Preparation
-
-## Load the data and results 
-
-```{r}
+## ---------------------------------------------------------------
 all_results <- readRDS("Results/all_results.rds")
 
 data_corn <- all_results[crop == "corn", data][[1]]
 data_soy <- all_results[crop == "soy", data][[1]]
-```
 
-## Set themes
-```{r}
+
+## ---------------------------------------------------------------
 theme_fig <-
   theme_bw() +
   theme(
@@ -61,13 +51,9 @@ theme_fig <-
     panel.background = element_blank(),
     panel.border = element_rect(fill = NA)
   )
-```
 
-# Summary statistics
 
-## Boxplot of water deficit by year
-
-```{r}
+## ---------------------------------------------------------------
 wd_data <-
   rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
   .[, .(wd = mean(balance)), by = .(year, crop)] %>%
@@ -201,11 +187,9 @@ ggsave(
   height = 6,
   width = 6
 )
-```
 
-## Saturated Thickness Map
 
-```{r echo = F}
+## ----echo = F---------------------------------------------------
 # === hpa ===#
 hpa_simplified <-
   # here("../Shared/Data/ProcessedData/hp_simplified.shp") %>%
@@ -331,11 +315,9 @@ g_soy <-
 g_map <- g_corn | g_soy
 
 ggsave(file = "Results/Figures/g_map.pdf", g_map, height = 7, width = 6)
-```
 
-# Yield response to water deficit by saturated thickness level (intensive margin) 
 
-```{r}
+## ---------------------------------------------------------------
 # all_results <-  data.table(all_results)
 
 plot_yield_data <-
@@ -521,11 +503,9 @@ g_yield_with_conf <-
 
 ggsave("Results/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
 ggsave("Results/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
-```
 
-# The impact of saturated thickness on the share of irrigated production
 
-```{r}
+## ---------------------------------------------------------------
 #++++++++++++++++++++++++++++++++++++
[…]

### 2023-10-31 Codes/ExploratoryAnalysis/1_exploratory_analysis.R

@@ -1,343 +1,0 @@
-######################################
-# Regression Analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(ggplot2)
-library(sf)
-library(raster)
-library(mgcv)
-library(data.table)
-library(lfe)
-library(stargazer)
-
-#+ read data
-
-final_data <- readRDS("Data/ProcessedData/final_data.rds") %>% 
-	.[sat == 0 & ir == "ir", sat := sat + 0.1] %>% 
-	.[, year_from_base := year - min(year)] %>% 
-	.[!is.na(sat), ] %>% 
-	.[, balance := balance / 1000]
-
-#+ define saturated thickness break
-
-sat_breaks <- 
-quantile(
-	final_data[ir == "ir", sat], 
-	prob = c(25, 50, 75, 100) / 100, 
-	na.rm = TRUE
-)
-
-#+ Create saturated thickness dummies
-
-reg_data <- 
-final_data[, 
-	sat_cat := cut(
-		sat, 
-	  breaks = c(0, 0.01, sat_breaks),
-	  include.lowest = TRUE
-	)  
-] %>% 
-.[, sat_cat_txt := paste0("sat_", as.numeric(sat_cat))] %>% 
-cbind(
-	., 
-	model.matrix(~ sat_cat_txt, data = .) %>% 
-	.[, -1] %>% 
-	data.table() %>% 
-	setnames(names(.), paste0("sat_d_", ((1:ncol(.)) + 1)))
-) %>% 
-.[, `:=`(
-		w_balance = balance - mean(balance),
-		w_yield = yield - mean(yield)
-	), 
-	by = .(sc_code, crop)
-] %>%  
-.[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)]
-
-# corn_data <- reg_data[crop == "corn", ]
-# soy_data <- reg_data[crop == "soy", ]
-
-# corn_data[state == "TX" & ir == "ir", ]
-# soy_data[state == "TX" & ir == "ir", ]
-
-
-#/*=================================================*/
-#' # Exploratory analysis
[…]

### 2023-10-31 Codes/ExploratoryAnalysis/analysis_explore.R

@@ -1,379 +1,0 @@
-######################################
-# Regression Analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(ggplot2)
-library(sf)
-library(raster)
-library(mgcv)
-library(data.table)
-library(lfe)
-library(stargazer)
-
-#+ read data
-
-final_data <- readRDS("Data/ProcessedData/final_data.rds") %>% 
-	.[sat == 0 & ir == "ir", sat := sat + 0.1] %>% 
-	.[, year_from_base := year - min(year)] %>% 
-	.[!is.na(sat), ] %>% 
-	.[, balance := balance / 1000] %>% 
-	.[, num_obs := .N, by = sc_code] %>% 
-	.[num_obs >= 10, ]
-
-#+ define saturated thickness break
-
-sat_breaks <- 
-quantile(
-	final_data[ir == "ir", sat], 
-	prob = c(25, 50, 75, 100) / 100, 
-	na.rm = TRUE
-)
-
-#+ Create saturated thickness dummies
-
-reg_data <- 
-final_data[, 
-	sat_cat := cut(
-		sat, 
-	  breaks = c(0, 0.01, sat_breaks),
-	  include.lowest = TRUE
-	)  
-] %>% 
-.[, sat_cat_txt := paste0("sat_", as.numeric(sat_cat))] %>% 
-cbind(
-	., 
-	model.matrix(~ sat_cat_txt, data = .) %>% 
-	.[, -1] %>% 
-	data.table() %>% 
-	setnames(names(.), paste0("sat_d_", ((1:ncol(.)) + 1)))
-) %>% 
-.[, 
-	`:=`(
-	w_balance = balance - mean(balance),
-		w_yield = yield - mean(yield)
-	), 
-	by = .(sc_code, crop)
-] %>%  
-.[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)]
-
-corn_states <- c("IL", "IA", "CO", "NE", "IN", "KS", "MS")
-corn_data <- reg_data[crop == "corn" & state %in% corn_states, ]
-soy_data <- reg_data[crop == "soy", ]
-
-#/*=================================================*/
-#' # Regression Analysis (GAM)
[…]

### 2023-10-31 Codes/Functions/functions.R

@@ -1,864 +1,0 @@
-prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
-  reg_data <-
-    data.table(
-      crop_type = c("corn", "soy"),
-      states = state_ls
-    ) %>%
-    rowwise() %>%
-    mutate(data = list(
-      data %>%
-        .[crop == crop_type, ] %>%
-        .[state %in% states, ] %>%
-        #--- yield ---#
-        .[, yield := yield * 12.553 / 200] %>%
-        #--- balance ---#
-        .[balance >= balance_thld[1], ] %>%
-        .[balance <= balance_thld[2], ] %>%
-        .[, balance := -balance] %>%
-        #--- unit conversion ---#
-        .[, sat := measurements::conv_unit(sat, "ft", "m")] %>%
-        #--- filter out counties with too low saturated thickness ---#
-        .[sat >= sat_thld_m | sat == 0, ] %>%
-        #--- share of irrigated/dryland acres ---#
-        # .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
-        .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ] %>%
-        .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
-        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
-        .[, state_year := paste0(state_name, "_", year)] %>%
-        .[, num_obs := .N, by = sc_dry] %>%
-        .[, num_obs := NULL]
-    )) %>%
-    # === sat breaks for grouping ===#
-    #' dryland production will have NA
-    mutate(sat_breaks = list(
-      quantile(
-        data[ir == "ir", sat],
-        prob = seq(0, 1, length = sat_cat_num + 1),
-        na.rm = TRUE
-      )
-    ))
-
-  #--- sat-breaks almost identical so, make them the same ---#
-  reg_data[2, ]$sat_breaks <- reg_data[1, ]$sat_breaks
-
-  return_data <-
-    reg_data %>%
-    mutate(data = list(
-      data %>%
-        # === define saturated thickness category variable ===#
-        .[
-          ,
-          sat_cat := cut(
-            sat,
-            breaks = sat_breaks,
-            include.lowest = TRUE
[…]

### 2023-10-31 Codes/Functions/retired_functions.R

@@ -1,686 +1,0 @@
-prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
-  # === saturated thickness threshold in meter ===#
-  # data <- reg_data$data[[1]]
-  # sat_breaks <- reg_data$sat_breaks[[1]]
-
-  # reg_data_y <- reg_data$reg_data_y[[1]]
-
-  reg_data <-
-    data.table(
-      crop_type = c("corn", "soy"),
-      states = list(state_ls[["corn"]], state_ls[["soy"]])
-    ) %>%
-    rowwise() %>%
-    mutate(base_data = list(
-      data %>%
-        # === lower bound ===#
-        .[balance >= balance_thld[1], ] %>%
-        # === upper bound ===#
-        .[balance <= balance_thld[2], ] %>%
-        #--- unit conversion ---#
-        .[, sat := measurements::conv_unit(sat, "ft", "m")]
-    )) %>%
-    # === select observations for each crop type ===#
-    mutate(data = list(
-      base_data[crop == crop_type & state %in% states, ]
-    )) %>%
-    mutate(data = list(
-      # === share of irrigated/dryland acres ===#
-      data[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)] %>%
-        # === filter out counties with too low saturated thickness ===#
-        # 0: dryland
-        # If irrigated, sat is set to be at least 0.01.
-        # See 0_6_merge_datasets.R
-        .[sat >= sat_thld_m | sat == 0, ]
-    )) %>%
-    # === sat breaks for grouping ===#
-    #' dryland production will have NA
-    mutate(sat_breaks = list(
-      quantile(
-        base_data[ir == "ir" & sat >= sat_thld_m, sat],
-        prob = seq(0, 1, length = sat_cat_num + 1),
-        na.rm = TRUE
-      )
-    )) %>%
-    # === prepare data for regression by crop type ===#
-    mutate(data = list(
-      data %>%
-        # === define saturated thickness category variable ===#
-        .[
-          ,
-          sat_cat := cut(
-            sat,
-            breaks = sat_breaks,
-            include.lowest = TRUE
-          )
-        ] %>%
[…]

### 2023-10-31 Codes/Functions/temp_1_0_functions_test.R

@@ -1,550 +1,0 @@
-
-prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, state_ls){
-
-  #=== saturated thickness threshold in meter ===#
-  sat_thld <- measurements::conv_unit(sat_thld_m, "m", "ft")
-
-  reg_data <-
-    data.table(
-      crop_type = c("corn", "soy"),
-      states = list(state_ls[["corn"]], state_ls[["soy"]])
-    ) %>%
-    rowwise() %>%
-    mutate(base_data = list(
-      data %>%
-      #=== lower bound ===#
-      .[balance >= balance_thld[1], ] %>%
-      #=== upper bound ===#
-      .[balance <= balance_thld[2], ]
-    )) %>%
-    #=== select observations for each crop type ===#
-    mutate(data = list(
-      base_data[crop == crop_type & state %in% states, ]
-    )) %>%
-    mutate(data = list(
-      #=== share of irrigated/dryland acres ===#
-      data[, acres_ratio := acres / sum(acres), by = .(sc_code, year, crop)] %>%
-      #=== filter out counties with too low saturated thickness ===#
-      # 0: dryland
-      # If irrigated, sat is set to be at least 0.01.
-      # See 0_6_merge_datasets.R
-      .[sat >= sat_thld | sat == 0, ]
-    )) %>%
-    #=== sat breaks for grouping ===#
-    #' dryland production will have NA
-    mutate(sat_breaks = list(
-      quantile(
-        base_data[ir == "ir" & sat >= sat_thld, sat],
-        # prob = c(0, 25, 50, 75, 100) / 100,
-        prob = c(0, 20, 40, 60, 80, 100) / 100,
-        na.rm = TRUE
-      )
-    )) %>%
-    #=== prepare data for regression by crop type ===#
-    mutate(reg_data_y = list(
-      data %>%
-      #=== define saturated thickness category variable ===#
-      .[,
-        sat_cat := cut(
-          sat,
-          breaks = sat_breaks,
-          include.lowest = TRUE
-        )
-      ] %>%
-      .[is.na(sat_cat), sat_cat := "dryland"] %>%
-      .[, sat_cat_i := factor(paste0("sc_", as.numeric(factor(sat_cat))))]
-    )) %>%
-    #=== prepare irrigation share data ===#
-    mutate(base_data_is = list(
[…]

### 2023-10-31 Codes/Retired/2_1_get_climate_scenarios.R

@@ -1,238 +1,0 @@
-######################################
-# Prepare daily weather data under climate change scenarios (2021 - 2060)
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-#=== packages ===#
-library(downloader)
-library(terra)
-library(lubridate)
-library(tidyverse)
-library(data.table)
-library(ncdf4)
-library(sf)
-library(raster)
-library(parallel)
-library(here)
-library(geosphere)
-library(exactextractr)
-
-#=== source functions ===#
-source(here("GitControlled/Codes", "functions.R"))
-
-#/*=================================================*/
-#' # Download climate scenario data
-#/*=================================================*/
-var_ls <- 
-  c(
-    "pr", # precipitation
-    "rhsmin", # min relative humidity (%)
-    "rhsmax", # max relative humidity (%)
-    "rsds", # radiation (W/M^2)
-    "tasmin", # tmin (Kelvin)
-    "tasmax", # tmax (Kelvin)
-    "uas", # wind speed,
-    "vas" # et
-  )
-
-s_year_ls <- seq(2021, 2096, by = 5)
-
-model_ls <- 
-  c(
-    "GFDL-ESM2M", 
-    "HadGEM2-ES365", 
-    "IPSL-CM5A-LR", 
-    "MIROC5"
-  )
-
-par_data_svmr <- 
-  expand.grid(
-    s_year = s_year_ls,
-    var_name = var_ls,
-    model = model_ls,
-    rcp = c(45, 85)
-  ) %>% 
-  data.table() %>% 
-  .[, var_name := as.character(var_name)] %>% 
-  .[order(model, rcp, s_year, var_name), ]
-
-# par_data_svmr <- arrange(par_data_svmr, s_year, var_name, rcp)
-
-#/*----------------------------------*/
-#' ## Download
-#/*----------------------------------*/
-mclapply(
-  seq_len(nrow(par_data_svmr)),
-  function(x) dl_cl_scenarios(par_data_svmr, x),
-  mc.cores = 12
-)
-
-# dl_cl_scenarios(par_data_svmr, 231)
-
-# par_data_svmr %>% 
-#   .[, id := 1:.N] %>% 
-#   .[var_name == "uas" & model == "HadGEM2-ES365" & rcp == 85 & s_year == 2041, ]
-
[…]

### 2023-10-31 Codes/Retired/2_2_simulation_analysis.R

@@ -1,281 +1,0 @@
-######################################
-# Simulation based on the statistical analysis
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-#=== load packages ===#
-library(tidyverse)
-library(RpackageTM)
-library(parallel)
-library(here)
-library(ggplot2)
-library(sf)
-library(raster)
-library(mgcv)
-library(gratia)
-library(data.table)
-library(lfe)
-library(stargazer)
-
-#=== Regression results ===#
-# created in 1_regression_analysis.R
-reg_results <- 
-  here("Shared/Results/reg_results.rds") %>% 
-  readRDS() %>% 
-  filter(model_name == "balance_d35") %>% 
-  dplyr::select(crop_type, states, data, gam_results, ir_share_gam, share_data, sat_breaks) %>% 
-  mutate(sat_ls = list(
-    ((sat_breaks + c(0, sat_breaks[-4])) / 2 / 10) %>% 
-    round(., digits = 0) %>% 
-    "*"(10) %>% 
-    c(0, .)
-  ))
-
-#=== source functions ===#
-source(here("GitControlled/Codes/functions.R"))
-
-#/*----------------------------------*/
-#' ## Set up parameters
-#/*----------------------------------*/
-cl_model_ls <- 
-  c(
-    "GFDL-ESM2M", 
-    "HadGEM2-ES365", 
-    "IPSL-CM5A-LR", 
-    "MIROC5"
-  )
-
-rcp_ls <- c(45, 85)
-
-par_data_base <- 
-  expand.grid(
-    crop_type = reg_results$crop_type %>% unique(),
-    cl_model = cl_model_ls,
-    rcp = rcp_ls
-  ) %>% 
-  data.table() 
-
-
-#/*=================================================*/
-#' # Truly hypothetical: everything modeled 
-#/*=================================================*/
-#' 
-#' Share of Irrigated Acres and Saturated Thickness modeled.
-#' What does it get: average representative farm experiencing different climate scenario.
-
-pred_yield_h <- function(i) {
-
-  print(i)
-
-  par_data <- 
-    par_data_base[i, ] %>% 
-    left_join(., reg_results, by = "crop_type") %>% 
[…]

### 2023-10-31 Codes/Retired/2_2_understand_weather_data.R

@@ -1,242 +1,0 @@
-######################################
-# Prepare daily weather data under climate change scenarios (2021 - 2060)
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-#=== packages ===#
-library(downloader)
-library(terra)
-library(lubridate)
-library(tidyverse)
-library(data.table)
-library(ncdf4)
-library(sf)
-library(raster)
-library(parallel)
-library(tmap)
-library(here)
-library(geosphere)
-library(exactextractr)
-
-#=== source functions ===#
-source(here("GitControlled/Codes", "functions.R"))
-
-#/*=================================================*/
-#' # Prepare data
-#/*=================================================*/
-model_ls <- 
-  c(
-    "GFDL-ESM2M", 
-    "HadGEM2-ES365", 
-    "IPSL-CM5A-LR", 
-    "MIROC5"
-  )
-
-rcp_ls <- c(45, 85)
-
-model_rcp <- 
-  expand.grid(
-    model = model_ls,
-    rcp = rcp_ls
-  ) %>% 
-  data.table()
-
-#/*----------------------------------*/
-#' ## Past
-#/*----------------------------------*/
-#=== data for regression ===#
-reg_data <- readRDS(here("Data/ProcessedData/final_data.rds"))
-
-sc_sf <- reg_data[, .(sc_code, geometry)] %>% unique(by = "sc_code")
-
-#=== weather data in the data fore regression ===#
-cl_data_current <- 
-  reg_data[, .(sc_code, year, balance, et0, prcp, days_ab_35)] %>% 
-  expand_grid_df(., model_rcp) 
-
-
-#/*----------------------------------*/
-#' ## Future
-#/*----------------------------------*/
-
-cl_data_future <- 
-  model_rcp %>% 
-  rowwise() %>% 
-  #=== define maca data file names ===#
-  mutate(file_name = list(
-    paste0(
-      "Data/ProcessedData/MACA/sum_", 
-      model,
-      "_",
-      rcp,
-      "_county.rds"
-    )  
-  )) %>% 
-  mutate(maca_data = list(
[…]

### 2023-10-31 Codes/Retired/3_1_make_figures_results.R

@@ -1,482 +1,0 @@
-#' ---
-#' title: "Make Figures"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include=FALSE
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = FALSE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi=400,
-  fig.width=7.5,
-  fig.height=5,
-  out.width="750px",
-  out.height="500px"
-)
-
-#/*=================================================*/
-#' # Prep
-#/*=================================================*/
-#+ package, include = F
-library(tmap)
-library(tidyverse)
-library(data.table)
-library(tidycensus)
-
-#=== regression results ===#
-reg_results <- readRDS(here("Shared/Results/reg_results.rds"))
-
-#=== hypothetical sim for regression results illustration ===#
-sim_results <- readRDS(here("Shared/Results", "hypothetical_sim_results.rds"))
-
-#=== simulation results ===#
-data_sim_yield <- readRDS(here("Shared/Results", "simulated_yield_data.rds"))  
-
-#=== county sf ===#
-data(fips_codes)
-
-all_counties <-
-  tigris::counties() %>% 
-  rename(
-    state_code = STATEFP, 
-    county_code = COUNTYFP,
-    county_name = NAME
-  ) %>% 
-  left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
-  dplyr::select(state_code, state, county_code, county_name) %>% 
-  mutate(sc_code = paste0(state_code, county_code)) %>% 
-  data.table()
-
-#/*=================================================*/
-#' # Yield response to drought and heat
-#/*=================================================*/
-
-#/*----------------------------------*/
-#' ## Water Deficit (Balance) on Yield by Saturated Thickness Level
-#/*----------------------------------*/
[…]

### 2023-10-31 Codes/Retired/3_2_make_figure_misc.R

@@ -1,454 +1,0 @@
-#' ---
-#' title: "Generate figures (climate, saturated thickness)"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include = FALSE
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-
-# /*=================================================*/
-#' # Preparation
-# /*=================================================*/
-# === packages ===#
-library(downloader)
-library(terra)
-library(lubridate)
-library(tidyverse)
-library(tidycensus)
-library(data.table)
-library(ncdf4)
-library(sf)
-library(raster)
-library(parallel)
-library(tmap)
-library(here)
-library(geosphere)
-library(exactextractr)
-
-# === source functions ===#
-source(here("GitControlled/Codes/functions.R"))
-
-# === Regression results ===#
-# created in 1_regression_analysis.R
-reg_results <-
-  here("Shared/Results", "reg_results.rds") %>%
-  readRDS() %>%
-  dplyr::select(crop_type, states, data, share_data, model_name) %>%
-  unique(by = "crop_type")
-
-# /*=================================================*/
-#' # Prepare Data
-# /*=================================================*/
-#+ echo = F
-
-# === sf for yield regression ===#
-sc_y_sf <-
-  reg_results %>%
-  dplyr::select(crop_type, data) %>%
-  unique(by = "crop_type") %>%
-  unnest() %>%
-  data.table() %>%
-  .[, .(sc_code, state, geometry, crop_type)] %>%
-  unique(by = c("sc_code", "crop_type"))
-
-# === sf for ir_share regression ===#
-sc_is_sf <-
-  reg_results %>%
-  dplyr::select(crop_type, share_data) %>%
-  unnest() %>%
-  data.table() %>%
[…]

### 2023-10-31 Codes/Retired/Outdated_0_2_quickstat_acres.R

@@ -1,261 +1,0 @@
-######################################
-# Get crop yield and acreage data: Corn and Soybeans
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-
-library(tidyverse)
-library(data.table)
-library(terra)
-library(raster)
-library(stars)
-library(here)
-library(sf)
-library(tidyUSDA)
-
-
-#/*=================================================*/
-#' # Prep
-#/*=================================================*/
-base_counties <- readRDS("./Data/ProcessedData/base_counties.rds")
-
-#' get the list of states
-all_states <- base_counties$state %>%
-  unique() %>% 
-  toupper() 
-
-#' get a list of data items
-tidyUSDA::allDataItem %>%
-  grep(pattern = "WHEAT", ., value = TRUE) %>%
-  grep(pattern = "YIELD", ., value = TRUE) %>%
-  grep(pattern = "IRRIGATED", ., value = TRUE)
-
-tidyUSDA::allDataItem %>%
-  grep(pattern = "COTTON", ., value = TRUE) %>%
-  grep(pattern = "YIELD", ., value = TRUE) %>%
-  grep(pattern = "IRRIGATED", ., value = TRUE)
-
-#' define a function to get QuickStat data
-get_quickstat_data <- function(year, data_item) {
-  ir_corn_yield <- getQuickstat(
-    key = "8FBF8552-57BD-38CB-9C16-E7008C5B8D7D",
-    program = "SURVEY",
-    data_item = data_item,
-    geographic_level = "COUNTY",
-    state = all_states,
-    year = year,
-    geometry = TRUE
-  ) %>%
-    #--- keep only some of the variables ---#
-    dplyr::select(year, NAME, county_code, state_fips_code, short_desc, Value) %>%
-    rename("state_code" = "state_fips_code")
-
-  return(ir_corn_yield)
-}
-
-#' Step 1: Major Crops (acre harvested) in the base states
- item_list0 <- c(
[…]

### 2023-10-31 Codes/Retired/Outdated_0_5_1_get_elevation.R

@@ -1,85 +1,0 @@
-######################################
-# Get elevation data
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-library(tidyverse)
-library(data.table)
-library(terra)
-library(raster)
-library(stars)
-library(here)
-library(sf)
-
-# /*=================================================*/
-#' # Elevation data
-# /*=================================================*/
-
-centroid_to_county <- readRDS("./Data/ProcessedData/centroid_to_county.rds")
-
-# ggplot() +
-# geom_sf(data = hpa_counties) +
-# geom_sf(data = hpa_grids, size = 0.2, color = "red")
-
-# === transforming the data into sp ===#
-centroid_to_county_sp <- as(centroid_to_county, "Spatial")
-
-#/*----------------------------------*/
-#' ## Download the elevation data and save it
-#/*----------------------------------*/
-
-if (!file.exists(here("Data/DEM/dem.tif"))) {
-  # === getting the extent of the data ===#
-  extent_counties <- ext(centroid_to_county_sp)
-
-  #=== create a table of lat and long ===#
-  extent_dt <- expand.grid(
-    lon  = seq(extent_counties[1], extent_counties[2], by = 5),
-    lat  = seq(extent_counties[3], extent_counties[4], by = 5)
-  ) %>% 
-  data.table()
-
-  #=== get the DEM data for all the lat/long combinations ===#
-  dem_data <- lapply(
-    seq_len(nrow(extent_dt)),
-    function(x) 
-    raster::getData(
-      "SRTM", 
-      lon = extent_dt[x, lon], 
-      lat = extent_dt[x, lat],
-      path = here("Data/DEM")
-    )
-  ) 
-
-  #=== combine them into a single raster ===#
-  dem_combined <- eval(parse(text=
-    paste0(
-      "mosaic(",
-      paste0("dem_data[[", seq_len(length(dem_data)), "]]", collapse = ","),
-      ", fun = mean)"
-    )
-  ))
-
-  #=== save it ===#
-  terra::writeRaster(dem_combined, here("Data/DEM/dem.tif"), format = "GTiff", overwrite = TRUE)
-
-  #=== remove all the individual dem files ===#
-  list.files(here("Data/DEM"), full.names = TRUE) %>% 
-    .[str_detect(., "srtm")] %>% 
-    unlink(recursive = TRUE)
-
-}
-
-#/*----------------------------------*/
[…]

### 2023-10-31 Codes/Retired/Outdated_0_5_2_get_daymet_et0.R

@@ -1,175 +1,0 @@
-######################################
-# Get Daymet weather data and calculate ET0
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-library(daymetr)
-
-#=== centroid-county data ===#
-centroid_to_county <- readRDS("./Data/ProcessedData/centroid_to_county.rds")
-
-#=== load the elevation data ===#
-elevation_data <- readRDS(here("Data/DEM/elevation.rds")) %>% 
-  .[, .(centroid_id, elev)]
-
-#=== merge ===#
-centroids <- left_join(
-  centroid_to_county, 
-  elevation_data,
-  by = "centroid_id"
-)
-
-# === ASCE_PenmanMonteith parameters ===#
-# define constant
-albedo <- 0.23 # Albedo for grass reference crop (Allen et al. 1998)
-G <- 0 # Assume soil heat flux is negligible
-sbc <- 4.903 * 10^(-9) # Stefan-Boltzmann constant (MJ/K^4/m2/day)
-Gs <- 0.0820 # Solar constant (MJ/m2/min)
-Cn <- 900 # Default for daily calculations for short-reference grass crop
-Cd <- 0.34 # Default for daily calculations for short-reference grass crop
-Wspd <- 3 # wind speed
-
-# /*----------------------------------*/
-#' ## define a function that download daymet and calculate et0
-# /*----------------------------------*/
-
-# i <- 1
-# get_weather(1)
-
-get_weather <- function(i) {
-
-  print(i)
-
-  temp_lat <- centroids[i, ] %>% pull(Y)
-  temp_lon <- centroids[i, ] %>% pull(X)
-  temp_site <- centroids[i, ] %>% pull(centroid_id)
-  temp_elev <- centroids[i, ] %>% pull(elev)
-
-  temp_daymet <-  
-  tryCatch(
-    {
-      download_daymet(
-        lat = temp_lat,
-        lon = temp_lon,
-        start = 1990,
-        end = 2018,
-        internal = TRUE
-      )  
-    }, 
-    error = function(cond){
-      return(NULL)
-    }
-  )
-
-  if (!is.null(temp_daymet)) {
-
[…]

### 2023-10-31 Codes/Retired/Outdated_0_5_3_summarize_weather.R

@@ -1,133 +1,0 @@
-######################################
-# Summarize weather data
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-# /*=================================================*/
-#' # Preparation
-# /*=================================================*/
-# === uploading the packages ===#
-library(SPEI)
-library(tidyverse)
-library(data.table)
-library(here)
-library(lubridate)
-
-#=== centroid-county ===#
-centroid_to_county <- readRDS(here("Data/ProcessedData/centroid_to_county.rds")) %>% 
-  data.table() %>% 
-  .[, .(centroid_id, sc_code)]
-
-#=== daily weather data ===#
-daily_weather <- readRDS(here("Data/ProcessedData/weather.rds")) %>% 
-  .[, year := year(date)]
-
-#/*=================================================*/
-#' # weekly balance data
-#/*=================================================*/
-
-#=== weekly balance ===#
-balance_data <- daily_weather %>% 
-  .[, .SD[!any(is.na(et0)), ], by = centroid_id] %>% 
-  .[, .SD[!any(is.na(prcp)), ], by = centroid_id] %>% 
-  .[, year := year(date)] %>% 
-  .[, balance := et0 - prcp] %>% 
-  .[, base_date := min(date), by = .(centroid_id, year)] %>% 
-  .[, date_dif := date - base_date] %>% 
-  .[, week := as.numeric(date_dif) %/% 7] %>% 
-  .[, .(balance = sum(balance), num_obs = .N), by = .(centroid_id, year, week)] %>% 
-  centroid_to_county[., on = "centroid_id"] %>% 
-  .[num_obs == 7, ] %>% 
-  .[, .(balance = mean(balance)), by = .(sc_code, year, week)]
-
-#=== break balance values into 10 groups ===#
-balance_data[, balance_cat := cut(
-  balance,
-  breaks = quantile(balance, prob = 0:10/10),
-  include.lowest = TRUE 
-)] %>% 
-.[, bal_txt := paste0("b_", as.numeric(balance_cat))] 
-
-balance_data[, .(bal_txt, balance_cat)] %>% 
-  unique(by = "bal_txt") %>% 
-  saveRDS(here("Data/ProcessedData/balance_txt_meaning.rds"))
-
-#=== weekly count of balance bins ===#
-balance_w <- balance_data[, .(num_week = .N), by = .(bal_txt, sc_code, year)] %>% 
[…]

### 2023-10-31 Codes/Retired/functions.R

@@ -1,602 +1,0 @@
-#/*=================================================*/
-#' # Download and process gridMET data
-#/*=================================================*/
-#' Used in 0_5_get_weather.R
-#' Objective: download gridMET data
-
-get_grid_MET <- function(var_name, year) {
-
-  target_url <-
-    paste0(
-      "http://www.northwestknowledge.net/metdata/data/",
-      var_name, "_", year,
-      ".nc"
-    )
-
-  file_name <-
-    paste0(
-      "/Volumes/Elements/Database/gridMET-historical/",
-      var_name, "_", year,
-      ".nc"
-    )
-
-  if (!file.exists(file_name)) {
-    downloader::download(
-      url = target_url,
-      destfile = file_name,
-      mode = 'wb'
-    )
-  }
-
-}
-
-#/*=================================================*/
-#' # Calculate ET0 from gridMET data
-#/*=================================================*/
-#' Used in 0_5_get_weather.R
-
-get_et0_county_year <- function(county, year) {
-
-  file_name <-
-    paste0(
-      "Shared/Data/ProcessedData/gMET-county/gMET-",
-      county$sc_code, "-", year,
-      ".rds"
-    )
-
-  if (!file.exists(file_name)) {
-    #--------------------------
-    # collect weather variables in a single data.table
-    #--------------------------
-    temp <-
-      lapply(
-        var_ls,
-        function(x) proess_gm_county(x, year, county)
-      ) %>%
-      rbindlist() %>%
-      dcast(
-        rowid + coverage_fraction + cell + x + y + date ~ var,
-        value.var = "value"
-      ) %>%
-      .[, Jday := as.numeric(date - min(date) + 1)] %>%
-      .[month(date) >= 4, ] %>%
-      .[month(date) <= 9, ] %>%
-      #=== day length in second ===#
-      .[, dayl := daylength(y, date) * 3600] %>%
-      setnames(
-        c("tmmx", "tmmn"),
-        c("tmax", "tmin"),
-      ) %>%
-      #=== Kelvin to Celsius ===#
-      .[, `:=`(
-        tmin = tmin - 273.15,
-        tmax = tmax - 273.15
-      )]
-
-    #--------------------------
-    # get elevation data
[…]

### 2023-10-31 Codes/Retired/get_spei.R

@@ -1,48 +1,0 @@
-#/*=================================================*/
-#' # Calculate SPEI
-#/*=================================================*/
-ET_data <- readRDS(here("Data", "weather.rds"))
-
-
-
-#/*----------------------------------*/
-#' ## Calculate SPEI
-#/*----------------------------------*/
-monthly_balance <- ET_data %>%
-.[, 
-  .(m_et = sum(ET0), m_prcp = sum(prcp)), 
-  by = .(centroid_id, state_code, county_code, year, month)
-] %>% 
-.[, m_balance := m_prcp - m_et] %>% 
-.[order(centroid_id, year, month), ] %>% 
-nest_by(centroid_id) %>% 
-mutate(data = list(
-  mutate(data, spei_6 = spei(data[, "m_balance"], 6)$fitted %>% as.vector())  
-)) %>% 
-mutate(data = list(
-  mutate(data, spei_3 = spei(data[, "m_balance"], 3)$fitted %>% as.vector())  
-)) %>% 
-unnest() %>% 
-data.table()
-
-#/*----------------------------------*/
-#' ## Combine SPEI 3 and 6
-#/*----------------------------------*/
-#=== Sep covers 4 ~ 9 ===#
-spei_6_Sep <- monthly_balance[month == 9, .(centroid_id, state_code, county_code, year, month, spei_6)] %>% 
-  .[, month := NULL]
-
-#=== June and Sep cover 4 ~ 6 and 7 ~ 9, respectively ===#
-spei_3_June_Sep <- monthly_balance[month %in% c(6, 9), .(centroid_id, state_code, county_code, year, month, spei_3)] %>% 
-  dcast(centroid_id + state_code + county_code + year ~ paste0("spei_3_", month), value.var = "spei_3")
-
-#=== Summarize by state and county codes ===#
-spei_data <- spei_3_June_Sep[spei_6_Sep, on = c("centroid_id", "year")] %>% 
-  .[, 
-    .(
-      spei_3_6 = mean(spei_3_6, na.rm = TRUE), 
-      spei_3_9 = mean(spei_3_9, na.rm = TRUE), 
-      spei_6 = mean(spei_6, na.rm = TRUE)
-    ), 
-    by = .(state_code, county_code, year)
-  ]  

### 2023-10-31 Codes/Retired/import_modules.py

@@ -1,24 +1,0 @@
-import econml
-
-# Main imports
-from econml.orf import DMLOrthoForest, DROrthoForest
-from econml.dml import CausalForestDML
-from econml.sklearn_extensions.linear_model import (
-    WeightedLassoCVWrapper,
-    WeightedLasso,
-    WeightedLassoCV,
-)
-
-from sklearn.ensemble import GradientBoostingRegressor
-from sklearn.multioutput import MultiOutputRegressor
-
-# Helper imports
-import numpy as np
-from itertools import product
-from sklearn.linear_model import (
-    Lasso,
-    LassoCV,
-    LogisticRegression,
-    LogisticRegressionCV,
-)
-import matplotlib.pyplot as plt

### 2023-10-31 Codes/Retired/make_figure_misc_explore.R

@@ -1,270 +1,0 @@
-#' ---
-#' title: "Generate figures (climate, saturated thickness)"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include = FALSE
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-
-#/*=================================================*/
-#' # Preparation
-#/*=================================================*/
-#=== packages ===#
-library(downloader)
-library(terra)
-library(lubridate)
-library(tidyverse)
-library(tidycensus)
-library(data.table)
-library(ncdf4)
-library(sf)
-library(raster)
-library(parallel)
-library(tmap)
-library(here)
-library(geosphere)
-library(exactextractr)
-
-#=== source functions ===#
-source(here("GitControlled/Codes/functions.R"))
-
-#=== Regression results ===#
-# created in 1_regression_analysis.R
-reg_results <- 
-  here("Shared/Results/all_res.rds") %>% 
-  readRDS() %>% 
-  dplyr::select(crop, data_int, ir_share_data) %>% 
-  unique(by = "crop")
-
-#/*=================================================*/
-#' # Prepare Data
-#/*=================================================*/
-#+ echo = F
-
-#/*----------------------------------*/
-#' ## Regression data
-#/*----------------------------------*/
-#=== sf for yield regression ===#
-sc_y_sf <-  
-  reg_results %>% 
-  dplyr::select(crop, data_int) %>% 
-  unique(by = "crop") %>% 
-  unnest() %>% 
-  data.table() %>% 
-  .[, .(sc_code, state, year, geometry, crop, sat_cat)] %>% 
-  unique(by = c("sc_code", "crop"))
-
-#=== sf for ir_share regression ===#
-sc_is_sf <-  
-  reg_results %>% 
[…]

### 2023-10-31 Codes/Retired/run_DML_OF_c.py

@@ -1,35 +1,0 @@
-def run_DML_OF_c(
-    Y,
-    T,
-    X,
-    W,
-    X_test,
-    n_trees=1000,
-    min_leaf_size=10,
-    max_depth=30,
-    subsample_ratio=0.7,
-):
-
-    est = DMLOrthoForest(
-        model_Y=WeightedLassoCVWrapper(cv=3),
-        model_T=WeightedLassoCVWrapper(cv=3),
-        model_Y_final=GradientBoostingRegressor(
-            n_estimators=300, max_depth=20, min_samples_leaf=10
-        ),
-        model_T_final=MultiOutputRegressor(
-            GradientBoostingRegressor(
-                n_estimators=300, max_depth=20, min_samples_leaf=10
-            )
-        ),
-        n_trees=n_trees,
-        min_leaf_size=min_leaf_size,
-        max_depth=max_depth,
-        subsample_ratio=subsample_ratio,
-    )
-
-    est.fit(Y, T, X=X, W=W)
-
-    treatment_effects = est.const_marginal_effect(X_test)
-    # treatment_effects = est.effect(X_test)
-
-    return treatment_effects

### 2023-10-31 Codes/Retired/spin.R

@@ -1,24 +1,0 @@
-library(knitr)
-library(tidyverse)
-library(here)
-library(prettydoc)
-
-here("GitControlled/Codes/0_1_hpa_counties.R") %>% 
-  knitr::spin(. , knit = FALSE) %>% 
-  rmarkdown::render()
-
-here("GitControlled/Codes/0_2_quickstat.R") %>% 
-  knitr::spin(. , knit = FALSE) %>% 
-  rmarkdown::render()
-
-here("GitControlled/Codes/0_3_get_sat.R") %>% 
-  knitr::spin(. , knit = FALSE) %>% 
-  rmarkdown::render()
-
-here("GitControlled/Codes/3_1_make_figures_results.R") %>% 
-  knitr::spin(. , knit = FALSE) %>% 
-  rmarkdown::render()
-
-here("GitControlled/Codes/3_2_make_figure_misc.R") %>% 
-  knitr::spin(. , knit = FALSE) %>% 
-  rmarkdown::render()

### 2023-10-31 Codes/functions.R

@@ -9,25 +9,13 @@ prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_
       data %>%
         .[crop == crop_type, ] %>%
         .[state %in% states, ] %>%
-        #--- yield ---#
-        .[, yield := yield * 12.553 / 200] %>%
         #--- balance ---#
         .[balance >= balance_thld[1], ] %>%
         .[balance <= balance_thld[2], ] %>%
-        .[, balance := -balance] %>%
-        #--- unit conversion ---#
-        .[, sat := measurements::conv_unit(sat, "ft", "m")] %>%
         #--- filter out counties with too low saturated thickness ---#
         .[sat >= sat_thld_m | sat == 0, ] %>%
-        #--- share of irrigated/dryland acres ---#
-        .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
-        .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ] %>%
-        .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
-        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
-        .[, state_year := paste0(state_name, "_", year)] %>%
-        .[, num_obs := .N, by = sc_dry] %>%
-        .[num_obs > 5, ] %>%
-        .[, num_obs := NULL]
+        #--- keep only if the ratio of irrigable acres is above the threshold ---#
+        .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ]
     )) %>%
     # === sat breaks for grouping ===#
     #' dryland production will have NA
@@ -276,10 +264,17 @@ share_analysis <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e, awc
   #++++++++++++++++++++++++++++++++++++
   #+ Debug
   #++++++++++++++++++++++++++++++++++++
+  # ir_share_data <- main_analysis$share_reg_data[[1]]
+  # sat_seq <- main_analysis$sat_seq[[1]]
+  # sandtotal_e <- main_analysis$sandtotal_med[[1]]
+  # silttotal_e <- main_analysis$silttotal_med[[1]]
+  # awc_e <- main_analysis$awc_med[[1]]
 
   #++++++++++++++++++++++++++++++++++++
   #+ Main
[…]

### 2023-10-31 Codes/spinned/0_1_hpa_counties.R

@@ -1,175 +1,0 @@
-#' ---
-#' title: "Prepare base-county data"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include=FALSE
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-# /*=================================================*/
-#' # Identify counties the overlaps with the HPA boundary
-# /*=================================================*/
-
-# /*----------------------------------*/
-#' ## Read the HPA border shapefile
-# /*----------------------------------*/
-#+ Read-HPA, cache = TRUE
-hpa <- st_read(here("Shared/Data/RawData/hp_bound2010.shp"))
-
-#' Simplify the HPA boundary for faster mapping
-hpa_simplified <-
-  here("Shared/Data/RawData/hp_bound2010.shp") %>%
-  st_read() %>%
-  st_simplify(dTolerance = 1000)
-
-saveRDS(hpa_simplified, here("Shared/Data/ProcessedData/hp_simplified.shp"))
-
-#' Map-of-HPA
-tm_shape(hpa_simplified) +
-  tm_polygons()
-
-# /*----------------------------------*/
-#' ## Get all the US counties as an sf
-# /*----------------------------------*/
-#+ Get-US-counties (entire US)
-(
-  US_county <-
-    tigris::counties(cb = TRUE) %>%
-    st_as_sf() %>%
-    st_transform(st_crs(hpa)) %>%
-    rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
-    dplyr::select(state_code, county_code)
-)
-
-# /*----------------------------------*/
-#' ## Identify counties that overlaps with `hpa`
-# /*----------------------------------*/
-#+ Identify-overlap, cache = TRUE
-hpa_counties <- US_county[hpa, ] %>%
[…]

### 2023-10-31 Codes/spinned/0_2_quickstat.R

@@ -1,173 +1,0 @@
-#' ---
-#' title: "Get crop yield and acreage data: Corn and Soybeans"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include=FALSE
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-# /*=================================================*/
-#' # Preparation
-# /*=================================================*/
-#' Read the counties data
-base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
-
-#' Get the list of states
-all_states <- base_counties$state_name %>%
-  unique() %>%
-  toupper()
-
-#' Define a function to get QuickStat data
-get_quickstat_data <- function(year, data_item) {
-  desc <-
-    tidyUSDA::allDataItem %>%
-    grep(pattern = data_item, ., value = TRUE) %>%
-    .[!grepl(., pattern = "NET")]
-
-  temp_data <- getQuickstat(
-    key = key_get("usda_nass_qs_api"),
-    program = "SURVEY",
-    data_item = desc,
-    geographic_level = "COUNTY",
-    # state = all_states,
-    year = year,
-    geometry = FALSE
-  ) %>%
-    #--- keep only some of the variables ---#
-    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
-    rename("state_code" = "state_fips_code")
-
-  return(temp_data)
-}
-
-#' List of items to download
-item_list <-
-  c(
-    "CORN, GRAIN, IRRIGATED - YIELD",
-    "CORN, GRAIN, NON-IRRIGATED - YIELD",
-    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
-    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
-    "SOYBEANS, IRRIGATED - YIELD",
-    "SOYBEANS, NON-IRRIGATED - YIELD",
[…]

### 2023-10-31 Codes/spinned/0_4_ndmc_di.R

@@ -1,46 +1,0 @@
-######################################
-# Drought Index
-######################################
-# written by Name on Date
-# objectives:
-# 1.
-
-#' According to Kuwayama (2018), the six key physical indicators consider percentiles
-#' from a soil moisture model, daily streamflow percentiles, the percent of normal precipitation,
-#' a standardized precipitation index, and a remotely-sensed vegetation
-#' health index (Svoboda et al. 2002). The authors also rely on supplementary indicators
-#' such as humidity and temperature departure from normal, reservoir, and lake levels,
-#' surface water supply indices, snowpack, and groundwater levels. The drought Index data is saved
-#' below:
-
-DI_1 <- fread(here("Data/RawData/NDMC_CO_IL_IN_DI_IA_KS_MS.csv"))
-DI_2 <- fread(here("Data/RawData/NDMC_DI_NM_OK_SD_TX_WY_NE.csv"))
-
-DI <- rbind(DI_1, DI_2)
-
-#' The default drought index is weekly persentage data. We changed them to monthly data by using the codes below.
-#' However, we are examing for different "seasons", from April to September, from May to October,
-#' from April to October, and from May to September. So we need to filter each season's drought index.
-
-drought_county <- DI %>%
-  .[, start_date := as.Date(ValidStart)] %>%
-  .[, end_date := as.Date(ValidEnd)] %>%
-  .[, FIPS := as.character(FIPS)] %>%
-  .[str_length(FIPS) == 4, FIPS := paste0("0", FIPS)] %>%
-  .[, state_code := str_sub(FIPS, 1, 2)] %>%
-  .[, county_code := str_sub(FIPS, 3, 5)] %>%
-  .[month(end_date) > 4 & month(start_date) < 10, ] %>%
-  #=== state-county code ===#
-  .[, sc_code := paste0(state_code, county_code)] %>% 
-  .[, .(
-    d0_5_9 = sum(D0) / 100,
-    d1_5_9 = sum(D1) / 100,
-    d2_5_9 = sum(D2) / 100,
-    d3_5_9 = sum(D3) / 100,
-    d4_5_9 = sum(D4) / 100
-  ), by = .(sc_code, year(end_date))] %>%
-  data.table()
-
-saveRDS(drought_county, here("Data/ProcessedData/DI_NDMC.rds"))
-
-

### 2023-10-31 Codes/spinned/2_figures_tables.R

@@ -1,298 +1,0 @@
-## ----r setup, include=FALSE---------------------------------------------------
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-
-## ----r------------------------------------------------------------------------
-#--- load pacakges ---#
-library(tidyverse)
-library(here)
-library(ggplot2)
-library(sf)
-library(broom)
-library(raster)
-library(mgcv)
-library(fixest)
-library(gratia)
-library(modelsummary)
-library(data.table)
-library(scales)
-
-#--- source functions ---#
-source(here("Codes/Functions/1_0_functions.R"))
-
-
-## ----r------------------------------------------------------------------------
-raw_data_both <-
-  rbind(data_corn, data_soy) %>%
-  .[, dry_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")]
-
-
-## ----r------------------------------------------------------------------------
-(
-  sum_table <-
-    modelsummary::datasummary(
-      type * (yield + balance) ~ dry_or_not * (Mean + SD),
-      data = raw_data_both,
-      output = "kableExtra"
-      # output = "latex"
-    )
-)
-
-
-## ----r------------------------------------------------------------------------
-(
-  g_wb_box <-
-    rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
-    ggplot(data = .) +
-    geom_boxplot(aes(y = -balance, x = factor(year), fill = type)) +
-    xlab("Year") +
-    ylab("Water Deficit (mm)") +
-    scale_fill_discrete(name = "") +
-    theme_bw() +
-    theme(
-      legend.position = "bottom",
-      axis.text.x = element_text(angle = 90)
-    )
-)
-
-
-## ----r------------------------------------------------------------------------
-g_yield_hist <-
[…]

### 2023-10-31 Codes/spinned/3_appendix.R

@@ -1,210 +1,0 @@
-## ----r------------------------------------------------------------------------
-#--- define the semi-parametric portion of the model ---#
-semi_formula <- formula(yield ~ s(spei, by = sat_cat, k = 3, m = 2))
-
-#--- semi-parametric FE wih fixest ---#
-semi_res <-
-  semi_ols(
-    semi_formula = semi_formula,
-    c_formula = "i(sat_cat, ref = 'dryland') + i(year) + i(sc_dry)",
-    # fe = "",
-    cluster = "sc_dry",
-    data = data_corn
-  )
-
-#--- main regression results ---#
-y_res_int <- semi_res$fe_res
-
-r2_corn_spei <- r2(y_res_int, type = "r2")
-
-
-## ----r, eval = FALSE----------------------------------------------------------
-modelsummary::msummary(y_res_int, stars = TRUE)
-
-
-## ----r------------------------------------------------------------------------
-gam_y_res <- semi_res$gam_res
-
-#--- raw data for prediction ---#
-data_for_pred_corn <-
-  CJ(
-    spei = seq(min(data_corn$spei), max(data_corn$spei), length = 50),
-    sat_cat = unique(data_corn$sat_cat)
-  ) %>%
-  .[, sat := parse_number(gsub("\\,.*", "", sat_cat))] %>%
-  .[is.na(sat), sat := 0] %>%
-  # === can be any sc_code (need to shift yield for "average" county) ===#
-  .[, sc_code := data_corn$sc_code[20]] %>%
-  .[, sc_dry := data_corn$sc_dry[20]] %>%
-  # === can be any year (need to shift yield for "average" year) ===#
-  .[, year := 2009] %>%
-  # === fake yield data (necessary to take advantage of predict.gam) ===#
-  .[, yield := 0] %>%
-  .[, type := "pred"] %>%
-  .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]
-
-#--- create bases  ---#
-yhat_data_corn <-
-  gen_smooth_data(
-    data = data_for_pred_corn,
-    gam_res = gam_y_res
-  ) %>%
-  .$data
-
-y_hat_with_se <- predict(y_res_int, newdata = yhat_data_corn, se.fit = TRUE)
-
-yhat_data_corn[, `:=`(
-  y_hat = y_hat_with_se$fit,
-  y_hat_se = y_hat_with_se$se.fit
[…]

### 2023-10-31 Writing/NatureWater/1stRoundRevision/knit.R

@@ -1,3 +1,0 @@
-setwd(here::here("GitControlled/Writing/NatureWater/1stRoundRevision"))
-
-

### 2023-10-31 Codes/master.rmd
functions.R organized

@@
 
-```{r}
+```{r, eval = FALSE}
 knitr::purl(here("Codes/0_1_hpa_counties.Rmd"), here("Codes/0_1_hpa_counties.R"))
@@
 knitr::purl(here("Codes/0_6_merge_datasets.Rmd"), here("Codes/0_6_merge_datasets.R"))
+```
+
+## Find counties in HPA
+```{r, eval = FALSE}
+source(here("Codes/0_1_hpa_counties.R"))
+```
+
+## Downlod yield and acreage data from USDA-NASS Quickstat
+
+```{r}
+source(here("Codes/0_2_quickstat.R"))
+```
+
+## Assign saturated thickness values to each county-year
+
+```{r, eval = FALSE}
+source(here("Codes/0_3_get_sat.R"))
+```
+
+## Download gridMET data and summarize by county-year
+
++ Downloading all the gridMET data and processing them take a while.
++ The total size of the gridMET files is about 35GB.
+
+```{r, eval = FALSE}
+source(here("Codes/0_4_get_weather.R"))
+```
+
+## Download SSURGO data and summarize by county
+
++ Running this takes a while.
++ SSURGO can be erratic and you may face downloading errors at times
+
+```{r, eval = FALSE}
+source(here("Codes/0_5_ssurgo.R"))
+```
+
+## Merge datasets, defined new variables, and drop observations
+
+```{r, eval = FALSE}
+source(here("Codes/0_6_merge_datasets.R"))
 ```

### 2023-10-31 Codes/0_2_quickstat.R

@@ -1,0 +1,94 @@
+## ---------------------------------------------------------------
+#--- load the counties ---#
+base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))
+
+#--- get the list of states ---#
+all_states <-
+  base_counties$state_name %>%
+  unique() %>%
+  toupper()
+
+#--- list of items to download ---#
+item_list <-
+  c(
+    "CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
+    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
+    "SOYBEANS, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
+    "SOYBEANS, IRRIGATED - ACRES HARVESTED",
+    "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED"
+  )
+
+#--- list of year-item combinations ---#
+par_list <-
+  data.table::CJ(
+    data_item = item_list,
+    year = as.character(1985:2018)
+  )
+
+
+## ----download-nass-data, eval = F-------------------------------
+## nass_data <-
+##   lapply(
+##     seq_len(nrow(par_list)),
+##     function(x) {
+##       print(x)
+##       data <-
+##         get_quickstat_data(
+##           year = par_list[x, year],
+##           data_item = par_list[x, data_item]
+##         ) %>%
+##         data.table()
+##       return(data)
+##     }
+##   )
+## 
+## #--- save the data ---#
+## saveRDS(nass_data, here("Data/data-raw/nass_data_raw.rds"))
+
+
+## ----process-nass-data------------------------------------------
+(
+  nass_data_wide <-
+    here("Data/data-raw/nass_data_raw.rds") %>%
+    readRDS() %>%
+    # === drop observations with county_code == 998 ===#
+    .[county_code != 998, ] %>%
+    # === define state-county code ===#
+    .[, sc_code := paste0(state_code, county_code)] %>%
+    .[, .(sc_code, year, short_desc, Value)] %>%
[…]

### 2023-10-31 Codes/0_4_get_weather.R

@@ -1,0 +1,114 @@
+## ---------------------------------------------------------------
+# === counties ===#
+counties <-
+  here("Data/data-processed/base_counties.rds") %>%
+  readRDS() %>%
+  st_as_sf()
+
+# === elevation ===#
+elevation <-
+  here("Data/data-raw/metdata_elevationdata.nc") %>%
+  rast()
+
+
+## ---------------------------------------------------------------
+# === list of weather variables ===#
+var_ls <-
+  c(
+    "pr", # precipitation
+    "rmin", # min relative humidity (%)
+    "rmax", # max relative humidity (%)
+    "srad", # radiation (W/M^2)
+    "tmmn", # tmin (Kelvin)
+    "tmmx", # tmax (Kelvin)
+    "vs", # wind speed,
+    "pet" # et
+  )
+
+# === list of years ===#
+year_ls <- 1985:2018
+
+# === var-year data ===#
+par_data <-
+  expand.grid(
+    var_name = var_ls,
+    year = year_ls
+  ) %>%
+  data.table() %>%
+  .[, var_name := as.character(var_name)]
+
+
+## ---------------------------------------------------------------
+# mclapply(
+#   seq_len(nrow(par_data)),
+#   function(x) get_grid_MET(par_data[x, var_name], par_data[x, year]),
+#   mc.cores = 12
+# )
+lapply(
+  seq_len(nrow(par_data)),
+  function(x) get_grid_MET(par_data[x, var_name], par_data[x, year])
+)
+
+
+## ---------------------------------------------------------------
+#--- counties ---#
+counties <- filter(counties, !(state %in% c("MO", "IL", "IN", "IA")))
+
+#--- get the CRS of a gridMET data ---#
+sample_gmet <- rast(here("Data/data-raw/gridMET/pr_1985.nc"))
+gmet_crs <- crs(sample_gmet)
+
+
+## ---------------------------------------------------------------
+if (!file.exists(here("data/data-processed", "gridMET.rds"))) {
+  all_data <-
+    mclapply(
+      seq_len(nrow(counties)),
+      function(x) {
+        print(x)
+
+        lapply(
+          year_ls,
+          function(y) get_et0_county_year(counties[x, ], y)
+        ) %>%
+          rbindlist() %>%
+          return()
+      },
+      mc.cores = parallel::detectCores() - 4
[…]

### 2023-10-31 Codes/0_5_ssurgo.R

@@ -1,0 +1,63 @@
+## ---------------------------------------------------------------
+all_counties <-
+  readRDS(here("Data/data-processed/base_counties.rds")) %>%
+  st_as_sf() %>%
+  dplyr::mutate(geometry = st_make_valid(geometry)) %>%
+  st_transform(4326)
+
+
+## ----download, cache = TRUE-------------------------------------
+vars <- c("sandtotal_r", "silttotal_r", "awc_r")
+
+
+lapply(
+  1:nrow(all_counties),
+  \(x) {
+    print(x)
+    file_name <- paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds")
+    if (!file.exists(file_name)) {
+      temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+      saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds"))
+    } else {
+      temp <- readRDS(file_name)
+      if (ncol(temp) < 4) {
+        temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+        saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, ".rds"))
+      }
+    }
+  }
+)
+
+# parallel::mclapply(
+#   1:nrow(all_counties),
+#   \(x) {
+#     print(x)
+#     file_name <- here::here(paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds"))
+#     if (!file.exists(file_name)) {
+#       print(paste0("working on ", x))
+#       temp <- get_ssurgo_props(all_counties[x, ], vars)
+#       saveRDS(temp, here::here(paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds")))
+#     } else {
+#       temp <- readRDS(file_name)
+#       if (ncol(temp) < 7) {
+#         temp <- get_ssurgo_props(field = all_counties[x, ], vars)
+#         saveRDS(temp, paste0("Data/data-raw/SSURGO/ssurgo_", x, "rds"))
+#       } else {
+#         print(paste0("skipping ", x))
+#       }
+#     }
+#   },
+#   mc.cores = 6,
+#   mc.preschedule = FALSE
+# )
+
+
+
+## ---------------------------------------------------------------
+ssurgo_soil <-
+  fs::dir_ls("Data/data-raw/SSURGO") %>%
+  purrr::map(readRDS) %>%
[…]

### 2023-10-31 Codes/0_6_merge_datasets.R

@@ -1,0 +1,64 @@
+## ---------------------------------------------------------------
+base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))
+
+#' Weather
+weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))
+
+#' Saturated thickness
+sat_all <- readRDS(here("Data/data-processed/sat_all.rds"))
+
+#' Yield, acreage data
+nass_data <- readRDS(here("Data/data-processed/nass_data.rds"))
+
+#* SSURGO
+ssurgo_soil <- readRDS(here("Data/data-processed/ssurgo_soil.rds"))
+
+
+## ---------------------------------------------------------------
+merged_data <-
+  base_counties[nass_data, on = c("sc_code")] %>%
+  weather_data[., on = c("sc_code", "year")] %>%
+  sat_all[., on = c("sc_code", "year")] %>%
+  .[order(sc_code, year), ] %>%
+  ssurgo_soil[, on = "sc_code"]
+
+
+## ---------------------------------------------------------------
+final_data <-
+  merged_data %>%
+  #--- yield unit conversion (bu/acre to ton/ha)---#
+  .[, yield := yield * 12.553 / 200] %>%
+  # === if not in_hpa and irrigated, drop them ===#
+  # irrigated, but saturated thickness not observed. These
+  # observations cannot be used.
+  .[!(in_hpa == 0 & ir == "ir"), ] %>%
+  # === if not irrigated, set sat = 0 ===#
+  .[ir == "nir", sat := 0] %>%
+  # === if sat == 0, but indicated irrigated ===#
+  # add 0.1 to distinguish dryland production, where
+  # sat is set at 0 above
+  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
+  # === drop if yield or acres are not observed ===#
+  .[!is.na(yield) & !is.na(acres), ] %>%
+  # === if sat is missing, drop ===#
+  #' just 2017 and 2018 data because saturated thickness data
+  #' is observed only until 2016
+  .[!is.na(sat), ] %>%
+  #--- indicator of whether dryland (rainfed) production or not ---#
+  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
+  #--- sc_code and dry_or_not (used for FE regression) ---#
+  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
+  #--- state-year combination ---#
+  .[, state_year := paste0(state_name, "_", year)] %>%
[…]

### 2023-10-31 Codes/1_analysis.R

@@ -165,7 +165,7 @@ all_results <-
           .[, balance_avg := mean(balance), by = sc_code]
 
         share_pred_data <-
-          share_analysis(
+          share_analysis_boot(
             ir_share_data = share_boot_data,
             sat_seq,
             sandtotal_med,

### 2023-10-31 Codes/functions.R

@@ -1,860 +1,145 @@
-prepare_reg_data <- function(data, sat_thld_m, ir_share_thld, balance_thld, sat_cat_num, state_ls) {
-  reg_data <-
-    data.table(
-      crop_type = c("corn", "soy"),
-      states = state_ls
+#!===========================================================
+# ! Quickstat (0_2_quickstat.rmd)
+# !===========================================================
+#++++++++++++++++++++++++++++++++++++
+#+ Download quick stat data
+#++++++++++++++++++++++++++++++++++++
+get_quickstat_data <- function(year, data_item) {
+  desc <-
+    tidyUSDA::allDataItem %>%
+    grep(pattern = data_item, ., value = TRUE)
+
+  temp_data <-
+    tidyUSDA::getQuickstat(
+      #--- replace API key with yours ---#
+      key = keyring::key_get("usda_nass_qs_api"),
+      program = "SURVEY",
+      data_item = desc,
+      geographic_level = "COUNTY",
+      state = all_states,
+      year = year,
+      geometry = FALSE
     ) %>%
-    rowwise() %>%
-    mutate(data = list(
-      data %>%
-        .[crop == crop_type, ] %>%
-        .[state %in% states, ] %>%
-        #--- balance ---#
-        .[balance >= balance_thld[1], ] %>%
-        .[balance <= balance_thld[2], ] %>%
-        #--- filter out counties with too low saturated thickness ---#
-        .[sat >= sat_thld_m | sat == 0, ] %>%
-        #--- keep only if the ratio of irrigable acres is above the threshold ---#
-        .[sat == 0 | (ir == "ir" & ir_area_ratio >= ir_share_thld), ]
-    )) %>%
-    # === sat breaks for grouping ===#
-    #' dryland production will have NA
-    mutate(sat_breaks = list(
-      quantile(
-        data[ir == "ir", sat],
-        prob = seq(0, 1, length = sat_cat_num + 1),
-        na.rm = TRUE
-      )
-    ))
+    #--- keep only some of the variables ---#
+    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
+    rename("state_code" = "state_fips_code")
 
-  #--- sat-breaks almost identical so, make them the same ---#
[…]

### 2023-11-01 Codes/0_1_hpa_counties.Rmd
manuscript.rnw created

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-```
-
 # Identify counties the overlaps with the HPA boundary
@@
 ```{r Read-HPA, cache = TRUE}
-hpa <- st_read(here("Shared/Data/RawData/hp_bound2010.shp"))
+hpa <- st_read(here("Data/data-raw/hp_bound2010.shp"))
 ```
@@
 
-```{r }
+```{r, eval = FALSE}
 tm_shape(st_simplify(hpa)) +
@@
 
-```{r }
+```{r eval = FALSE}
 tm_shape(st_simplify(hpa)) +
@@
 
-## Find the area of the portion that is intersecting with hpa
+## Find the area of the portion that is intersecting with `hpa`
 Intersect HPA and counties in HPA
@@
 ```{r intersect-HPA-county}
-intersected_portion <- st_intersection(hpa, hpa_counties) %>%
-  #=== get the intersecting area ===#
+intersected_portion <-
+  st_intersection(hpa, hpa_counties) %>%
+  # === get the intersecting area ===#
   mutate(area = st_area(.) %>% as.numeric()) %>%
@@
   ) %>%
-  #=== find the proportion of the county area that overlaps with HPA ===#
+  # === find the proportion of the county area that overlaps with HPA ===#
   mutate(ir_area_ratio = interesected_area / whole_area) %>%
@@
 
-```{r }
-hpa_counties$ir_area_ratio %>% hist
+```{r eval = FALSE}
+hpa_counties$ir_area_ratio %>% hist()
 ```
@@
 
-additional_states <- fips_codes %>%
+additional_states <-
+  fips_codes %>%
   data.table() %>%
@@
 ```{r }
-fips_codes <- fips_codes %>%
+fips_codes <-
+  fips_codes %>%
   data.table() %>%
@@
 (
-all_counties <- filter(US_county, state_code %in% all_states) %>%
+  all_counties <-
+    dplyr::filter(US_county, state_code %in% all_states) %>%
     data.table() %>%
-  #=== merge with the data on counties in HPA ===#
+    # === merge with the data on counties in HPA ===#
     hpa_counties[., on = c("state_code", "county_code")] %>%
-  #=== merge with FIPS codes data ===#
+    # === merge with FIPS codes data ===#
     fips_codes[., on = c("state_code", "county_code")] %>%
-  #=== indicator of whether intersecting with HPA or not ===#
+    # === indicator of whether intersecting with HPA or not ===#
     .[, in_hpa := 0] %>%
     .[!is.na(ir_area_ratio), in_hpa := 1] %>%
-  #=== generate state-code variable ===#
+    # === generate state-code variable ===#
     .[, sc_code := paste0(state_code, county_code)] %>%
-  #=== keep only the relevant variables ===#
+    # === keep only the relevant variables ===#
     .[, .(state, state_name, sc_code, ir_area_ratio, geometry, in_hpa)]
@@
 
-```{r map-all-counties, cache = TRUE}
+```{r map-all-counties, eval = FALSE}
 tm_shape(st_as_sf(all_counties)) +
@@
 ```{r }
-saveRDS(all_counties, here("Shared/Data/ProcessedData/base_counties.rds"))
-
-# all_counties[, state_name] %>% unique()
+saveRDS(all_counties, here("Data/data-processed/base_counties.rds"))
 ```

### 2023-11-01 Codes/1_Analysis.rmd
manuscript.rnw created

@@
   here("Data/data-processed/final_data.rds") %>%
-  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
   readRDS()
@@
   setnames("crop_type", "crop")
-
-# saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))
-
-# ggplot(reg_data$data[[1]]) +
-#   geom_point(aes(y = acres_ratio, x = sat)) +
-#   geom_smooth(aes(y = acres_ratio, x = sat))
 ```
 
-# Regression and prediciton
-
-## Yield
+# Regression and prediciton (Main)
 
 ```{r}
-# ir_share_data <- main_analysis$share_reg_data[[1]]
-# sat_seq <- main_analysis$sat_seq[[1]]
-# sandtotal_e <- main_analysis$sandtotal_med[[1]]
-# silttotal_e <- main_analysis$silttotal_med[[1]]
-# awc_e <- main_analysis$awc_med[[1]]
-
-# main_analysis$share_reg_data[[1]]
-# main_analysis$share_reg_data[[2]]
-
 main_analysis <-
@@
       .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-      .[, balance_avg := mean(balance), by = sc_code]
+      .[, balance_avg := mean(balance), by = sc_code] %>%
+      .[sat < 180, ]
   )) %>%
@@
 
+# Bootstrap
+
 ```{r}
-#++++++++++++++++++++++++++++++++++++
-#+ Bootstrap
-#++++++++++++++++++++++++++++++++++++
 set.seed(383954)
@@
           .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-          .[, balance_avg := mean(balance), by = sc_code]
+          .[, balance_avg := mean(balance), by = sc_code] %>%
+          .[sat < 180, ]
 
         share_pred_data <-
-          share_analysis(
+          share_analysis_boot(
             ir_share_data = share_boot_data,
@@
     yield_dif_test,
+    share_reg_data,
     share_pred_data,

### 2023-11-01 Codes/2_figures_tables.rmd
manuscript.rnw created

--- 2023-09-02 Codes/2_figures_tables.rmd
@@
 ```{r}
-all_results <- readRDS("Shared/Results/all_results.rds")
+all_results <- readRDS("Results/all_results.rds")
+
+data_corn <- all_results[crop == "corn", data][[1]]
+data_soy <- all_results[crop == "soy", data][[1]]
 ```
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_y_wd_all.pdf",
+  file = "Results/Figures/g_y_wd_all.pdf",
   g_y_wd,
@@
 
-```{r echo = F, eval = F}
-# === hpa ===#
-hpa_simplified <-
-  # here("../Shared/Data/ProcessedData/hp_simplified.shp") %>%
-  here("Shared/Data/ProcessedData/hp_simplified.shp") %>%
-  readRDS()
+Though not necessary, save the counties and states data so that you do not have to waste time downloading them every time you run this code.
 
-# === all counties ===#
-data(fips_codes, package = "tidycensus")
-
+```{r, eval = FALSE}
 all_counties <-
@@
 
+saveRDS(all_counties, "Data/data-processed/counties.rds")
+
 all_states <-
@@
   )
+
+saveRDS(all_states, "Data/data-processed/states.rds")
+```
+
+Simplify HPA boundary to make the figure light-weight for mapping:
+
+```{r}
+hpa_simplified <-
+  here("Data/data-raw/hp_bound2010.shp") %>%
+  st_read() %>%
+  st_simplify(dTolerance = 1000)
+
+saveRDS(hpa_simplified, here("Data/data-processed/hpa_simplified.shp"))
+```
+
+```{r}
+#--- hpa ---#
+hpa_simplified <-
+  here("Data/data-processed/hpa_simplified.shp") %>%
+  readRDS()
+
+#--- fips code ---#
+data(fips_codes, package = "tidycensus")
+
+#--- all relevant counties ---#
+all_counties <- readRDS("Shared/Data/counties.rds")
+
+#--- all relevant states ---#
+all_states <- readRDS("Shared/Data/states.rds")
 
@@
 
-ggsave(file = "GitControlled/Writing/Figures/g_map.pdf", g_map, height = 7, width = 6)
+ggsave(file = "Results/Figures/g_map.pdf", g_map, height = 7, width = 6)
 ```
@@
 
-# g_yield_response <-
-#   cowplot::plot_grid(
-#     g_wd_corn, g_ir_yield_corn,
-#     ncol = 1,
-#     align = "v",
-#     rel_heights = c(0.15, 0.72)
-#   )
-
-# g_yield_response_soy <-
-#   cowplot::plot_grid(
-#     g_wd_soy, g_ir_yield_soy,
-#     ncol = 1,
-#     align = "v",
-#     rel_heights = c(0.15, 0.72)
-#   )
-
-cowplot::ggsave2("GitControlled/Writing/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)
+cowplot::ggsave2("Results/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)
 
@@
 
-ggsave("GitControlled/Writing/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
-ggsave("GitControlled/Writing/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
+ggsave("Results/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
+ggsave("Results/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
 ```
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_share.pdf",
+  file = "Results/Figures/g_share.pdf",
   g_ir_share,
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_share_dif.pdf",
+  file = "Results/Figures/g_share_dif.pdf",
   g_share_dif,
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_total_impact.pdf",
+  file = "Results/Figures/g_total_impact.pdf",
   g_total_impact,
@@
 
-cowplot::ggsave2("GitControlled/Writing/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
+cowplot::ggsave2("Results/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
 ```
@@
 
-ggsave("GitControlled/Writing/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
+ggsave("Results/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
 ```
@@
 
-ggsave("GitControlled/Writing/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
+ggsave("Results/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
 ```

### 2023-11-01 Codes/A_1_state_fe.rmd
manuscript.rnw created

--- 2023-08-30 Codes/A2_state_fe.rmd
@@
 ```{r}
-all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
+all_results <- readRDS(here::here("Results/all_results.rds"))
 ```
@@
   )
-
-# ggplot(results$ir_share_data[[1]]) +
-#   geom_point(aes(y = acres_ratio, x = sat))
 
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_share_comp.pdf",
+  file = "Results/Figures/g_share_comp.pdf",
   g_share_comp,
@@
 ```{r}
-ir_share_data_corn[, sat_m := measurements::conv_unit(sat, "ft", "m")]
+ir_corn_share_data <- all_results[crop == "corn", share_reg_data][[1]]
 
 sy_demeaan_data <-
-  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = .(state_name, year)] %>%
+  copy(ir_corn_share_data)[, .(sat = sat - mean(sat)), by = .(state_name, year)] %>%
   .[, type := "B. Demeaned by State-Year Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
 c_y_demeaan_data <-
-  copy(ir_share_data_corn)[, sat_m := sat_m - mean(sat_m), by = sc_code] %>%
-  .[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  copy(ir_corn_share_data)[, sat := sat - mean(sat), by = sc_code] %>%
+  .[, .(sat = sat - mean(sat)), by = year] %>%
   .[, type := "C. Demeaned by County and Yearly Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
 y_demeaan_data <-
-  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  copy(ir_corn_share_data)[, .(sat = sat - mean(sat)), by = year] %>%
   .[, type := "A. Demeaned by Yearly Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
@@
   ggplot(plot_data) +
-  geom_histogram(aes(x = sat_m)) +
+  geom_histogram(aes(x = sat)) +
   facet_wrap(type ~ ., ncol = 1) +
@@
 ggsave(
-  file = "GitControlled/Writing/Figures/g_variation.pdf",
+  file = "Results/Figures/g_variation.pdf",
   g_variation,

### 2023-11-01 Writing/manuscript.Rnw
manuscript.rnw created

--- 2023-10-31 Writing/manuscript.tex
@@
 
+<<init, echo = FALSE>>=
+knitr::opts_chunk$set(
+  echo = FALSE
+)
+library(tidyverse)
+library(data.table)
+@
+
 \title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
@@
 
+<<load_results, echo = FALSE>>=
+setwd(here::here("Writing"))
+all_results <- readRDS("../Results/all_results.rds")
+@
+
 \hypertarget{impact-intensive}{%
@@
 
+<<echo = FALSE>>=
+delta_yield_lowest_highest <- 
+  all_results[crop == "corn", yield_dif_test][[1]] %>%
+  .[comp_q == 3, ] %>%  
+  .[which.min(abs(balance -950)), dif_hat] %>%
+  round(digits = 2)
+@
+
 Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95\% confidence interval for corn and soybean.
 For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is 1.15 tonne/ha.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is \Sexpr{delta_yield_lowest_highest} tonne/ha.
 Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
@@
 
-% At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
-In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
-These counter-intuitive results reflect a combination of two factors.
-First, in years with small water deficits there is less need for irrigation and hence irrigated and rainfed yield differences are reduced.
-Second, years with small or even negative water deficits are associated with wetter conditions during the growing season, which can negatively impact yield if waterlogging occurs leading to greater risks of physical crop damage (e.g., lodging of grains), delays to planting/harvesting, nutrient leaching, and restricted root growth in waterlogged soils \citep{li2019excessive}.
-Declines in irrigated yields could be indicative of over-irrigation by farmers, which has been shown to occur in the HPA region in wet years \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}, exacerbating these risks.
-However, the small declines in irrigated yields at low or negative levels of water deficit may also reflect the fact that wetter years are typically cooler, reducing positive heat effects on crop growth and yield formation.
-
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}
 
 }
+
+<<num_obs_yield_reg, echo = FALSE>>=
+nobs_corn <- 
+  all_results[crop == "corn", data][[1]] %>% 
+  nrow() %>%
+  format(big.mark=",")
+
+nobs_soy <- all_results[crop == "soy", data][[1]] %>% 
+  nrow() %>%
+  format(big.mark=",")
+@
 
@@
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $8,773$ and $5,977$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
+Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
 \end{figure}
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_ir_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_ir_yield_dif}
 
@@
 
+<<ir_share, echo = FALSE>>=
+corn_share <- all_results[crop == "corn", share_pred_data][[1]]
+
+share_high <-
+  corn_share[which.min(abs(sat - 130)), ir_share_hat] %>%
+  round(digits = 2)
+
+share_low <-
+  corn_share[which.min(abs(sat - 10)), ir_share_hat] %>%
+  round(digits = 2) 
+@
+
 Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100 m to 0.72 (72\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}
 
@@
 
+<<num_obs_share_reg, echo = FALSE>>=
+nobs_corn_share <- 
+  all_results[crop == "corn", share_reg_data][[1]] %>% 
+  nrow() %>%
+  format(big.mark=",")
+
+nobs_soy_share <- 
+  all_results[crop == "soy", share_reg_data][[1]] %>% 
+  nrow() %>%
+  format(big.mark=",")
+@
+
 \caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval).
 The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $2,523$ and $1,558$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
+Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
 \end{figure}
@@
 \subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
+
+<<res_at_950, echo = FALSE>>=
+res_at_950 <-
+  all_results[, .(crop, avg_yield_pred_data)] %>%
+  unnest(cols = c(avg_yield_pred_data)) %>%
+  data.table() %>%
+  .[crop == "corn", ] %>%
+  .[(abs(balance - 950)) < 5, ]
+@
 
@@
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.09 tonne/ha of production loss for corn.
+At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only \Sexpr{round(res_at_950[sat == 100 & crop == "corn", avg_yield] - res_at_950[sat == 70 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.57 tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in \Sexpr{round(res_at_950[sat == 70 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to \Sexpr{round(res_at_950[sat == 40 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_total_impact}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_avg_yield_dif}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}
 
@@
 
-{\centering \includegraphics[width=750px,height=6in,]{../../Figures/g_y_wd_all}
+{\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}
 
@@
 Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any aquifer cells within a county that had a aquifer thickness value lower than 9 m were removed before aggregation to reflect the fact that aquifer thicknesses lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
+Furthermore, during the aggregation process, any county that had an aquifer thickness value lower than 9 m was removed before aggregation to reflect the fact that aquifer thicknesses lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
 Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
@@
 
-These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($8,773$ observations), (2) per-acre soybean yield ($5,977$ observations), (3) share of irrigated corn ($2,523$ observations), and (4) share of irrigated soybean ($1,558$ observations).
+These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($\Sexpr{nobs_corn}$ observations), (2) per-acre soybean yield ($\Sexpr{nobs_soy}$ observations), (3) share of irrigated corn ($\Sexpr{nobs_corn_share}$ observations), and (4) share of irrigated soybean ($\Sexpr{nobs_soy_share}$ observations).
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_map}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\subsection{Data Availability}\label{data-and-code-availability}}
+\hypertarget{data-availability}{%
+\subsection{Data Availability}\label{data-availability}}
 
-All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All analyses were performed in R \citep{R} and all the raw data-sets used in this study are publicly accessible.
+All the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
@@
 \subsection{Code Availability}
+All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
+All analyses were performed in R \citep{R}.
 
@@
 
-\section*{Author Contribution Statement}
-
-\section*{Competing Interests Statement}
+\subsection*{Competing Interests Statement}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_corn}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_soy}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_avg_yield}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}
 
@@
 
-Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100 m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
+<<sat_ref, echo = FALSE>>=
+sat_ref <-
+  readRDS(here::here("Shared/Results/all_results.rds")) %>%
+  .[1, sat_seq] %>%
+  .[[1]] %>%
+  .[length(.)]
+@
+
+Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of \Sexpr{sat_ref} m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
 Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
-Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 100 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to \Sexpr{sat_ref} m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_share_dif}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}
 
@@
 
-\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 100 m (black line).
+\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
 The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_variation}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share_comp}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_basis}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_il_y}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}
 

### 2023-11-01 Codes/1_analysis.R

@@ -1,8 +1,7 @@
-## ----------------------------------------------------------
+## ---------------------------------------------------------------
 #--- read data ---#
 final_data <-
   here("Data/data-processed/final_data.rds") %>%
-  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
   readRDS()
 
 #--- Data used for regression analysis ---#
@@ -28,23 +27,8 @@ reg_data <-
   data.table() %>%
   setnames("crop_type", "crop")
 
-# saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))
 
-# ggplot(reg_data$data[[1]]) +
-#   geom_point(aes(y = acres_ratio, x = sat)) +
-#   geom_smooth(aes(y = acres_ratio, x = sat))
-
-
-## ----------------------------------------------------------
-# ir_share_data <- main_analysis$share_reg_data[[1]]
-# sat_seq <- main_analysis$sat_seq[[1]]
-# sandtotal_e <- main_analysis$sandtotal_med[[1]]
-# silttotal_e <- main_analysis$silttotal_med[[1]]
-# awc_e <- main_analysis$awc_med[[1]]
-
-# main_analysis$share_reg_data[[1]]
-# main_analysis$share_reg_data[[2]]
-
+## ---------------------------------------------------------------
 main_analysis <-
   reg_data %>%
   rowwise() %>%
@@ -66,7 +50,8 @@ main_analysis <-
       .[comp == 2, ] %>%
       .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
       .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
-      .[, balance_avg := mean(balance), by = sc_code]
+      .[, balance_avg := mean(balance), by = sc_code] %>%
+      .[sat < 180, ]
   )) %>%
   #--- the same sequence used for all the bootstrap iterations ---#
   dplyr::mutate(sat_seq = list(
@@ -125,10 +110,7 @@ main_analysis <-
 saveRDS(main_analysis, "Results/main_analysis.rds")
 
 
-## ----------------------------------------------------------
-#++++++++++++++++++++++++++++++++++++
-#+ Bootstrap
-#++++++++++++++++++++++++++++++++++++
+## ---------------------------------------------------------------
 set.seed(383954)
 
 all_results <-
@@ -162,7 +144,8 @@ all_results <-
           .[comp == 2, ] %>%
[…]

### 2023-11-01 Codes/A_1_state_fe.R

@@ -1,13 +1,8 @@
-In the main regresion analysis in the manuscript, state FEs are included. County FEs are not included as it eliminates too much of the variation in aquifer thickness. This appendix compares the impact of aquifer thickness on irrigated share with and without State FEs.
-
-# Regression with and without state fixed effects
-
-```{r}
-all_results <- readRDS(here::here("Shared/Results/all_results.rds"))
-```
+## ---------------------------------------------------------------
+all_results <- readRDS(here::here("Results/all_results.rds"))
 
 
-```{r}
+## ---------------------------------------------------------------
 #++++++++++++++++++++++++++++++++++++
 #+ Formula
 #++++++++++++++++++++++++++++++++++++
@@ -33,9 +28,6 @@ formula_data <-
     formula = c(share_formula_yfe, share_formula_sandyfe, share_formula_syfe),
     type = c("Year FE", "State and Year FE", "State-Year FE")
   )
-
-# ggplot(results$ir_share_data[[1]]) +
-#   geom_point(aes(y = acres_ratio, x = sat))
 
 results <-
   all_results[, .(crop, data)] %>%
@@ -91,9 +83,9 @@ results <-
   dplyr::select(crop, type, share_pred_data) %>%
   unnest(cols = c(share_pred_data)) %>%
   data.table()
-```
 
-```{r}
+
+## ---------------------------------------------------------------
 g_share_comp <-
   results[sat <= 150, ] %>%
   .[, crop := case_when(
@@ -123,16 +115,14 @@ g_share_comp <-
   )
 
 ggsave(
-  file = "GitControlled/Writing/Figures/g_share_comp.pdf",
+  file = "Results/Figures/g_share_comp.pdf",
   g_share_comp,
   height = 5,
   width = 6
 )
-```
 
-# Loss of variation in aquifer thickness 
 
-```{r}
+## ---------------------------------------------------------------
 ir_share_data_corn[, sat_m := measurements::conv_unit(sat, "ft", "m")]
 
 sy_demeaan_data <-
@@ -166,9 +156,9 @@ g_variation <-
   ylab("Count")
 
 ggsave(
-  file = "GitControlled/Writing/Figures/g_variation.pdf",
+  file = "Results/Figures/g_variation.pdf",
   g_variation,
   height = 5,
   width = 6
 )
-```
+

### 2023-11-01 Codes/A_2_spline.R

@@ -1,0 +1,70 @@
+## ---------------------------------------------------------------
+library(mgcv)
+set.seed(83944)
+
+# === generate data ===#
+N <- 400 # number of observations
+x <- seq(1, 250, length = N)
+y_det <- 240 * (1 - 0.4 * exp(-0.03 * x))
+e <- 25 * rnorm(N) # error
+data <- data.table(x = x, y = y_det + e, y_det = y_det)
+
+basis_data <-
+  gam(y_det ~ s(x, k = 5, bs = "cr"), data = data) %>%
+  predict(., type = "lpmatrix") %>%
+  data.table() %>%
+  .[, x := data[, x]] %>%
+  melt(id.var = "x") %>%
+  .[, variable := case_when(
+    variable == "(Intercept)" ~ "Intercept",
+    variable == "s(x).1" ~ "1st basis",
+    variable == "s(x).2" ~ "2nd basis",
+    variable == "s(x).3" ~ "3rd basis",
+    variable == "s(x).4" ~ "4th basis"
+  )] %>%
+  .[, variable := factor(
+    variable,
+    levels = c("Intercept", "1st basis", "2nd basis", "3rd basis", "4th basis")
+  )]
+
+g_basis <-
+  ggplot(data = basis_data) +
+  geom_line(aes(y = value, x = x)) +
+  facet_grid(variable ~ .) +
+  theme_fig +
+  ylab("")
+
+ggsave("Results/Figures/g_basis.pdf", g_basis, width = 6, height = 5)
+
+
+## ---------------------------------------------------------------
+g_il_1 <-
+  data.table(
+    variable = unique(basis_data$variable),
+    coef = rep(60, 5)
+  ) %>%
+  .[basis_data, on = "variable"] %>%
+  .[, sum(coef * value), by = x] %>%
+  ggplot(data = .) +
+  geom_line(aes(y = V1, x = x)) +
+  theme_fig +
+  ylab("y") +
+  ggtitle("Case 1")
+
+g_il_2 <-
+  data.table(
+    variable = unique(basis_data$variable),
+    coef = c(20, -20, -40, 40, 60)
+  ) %>%
+  .[basis_data, on = "variable"] %>%
+  .[, sum(coef * value), by = x] %>%
+  ggplot(data = .) +
+  geom_line(aes(y = V1, x = x)) +
+  theme_fig +
+  ylab("y") +
+  ggtitle("Case 2")
+
[…]

### 2023-11-01 Codes/functions.R

@@ -717,7 +717,7 @@ yield_analysis_boot <- function(yield_data, balance_seq, sat_seq_eval, sat_break
 
 #++++++++++++++++++++++++++++++++++++
 #+ Run Irrigation Share Analysis (main)
-#++++++++++++++++++++++++++++++++++++ 
+#++++++++++++++++++++++++++++++++++++
 share_analysis_gam <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e, awc_e) {
   #++++++++++++++++++++++++++++++++++++
   #+ Debug
@@ -731,8 +731,6 @@ share_analysis_gam <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e,
   #++++++++++++++++++++++++++++++++++++
   #+ Main
   #++++++++++++++++++++++++++++++++++++
-  ir_share_data <- ir_share_data[sat < 180, ]
-
   #---------------------
   #- define the formula for the flexible part
   #---------------------
@@ -807,8 +805,6 @@ share_analysis_boot <- function(ir_share_data, sat_seq, sandtotal_e, silttotal_e
   #++++++++++++++++++++++++++++++++++++
   #+ Main
   #++++++++++++++++++++++++++++++++++++
-  ir_share_data <- ir_share_data[sat < 180, ]
-
   #---------------------
   #- define the formula for the flexible part
   #---------------------

### 2023-11-01 Writing/manuscript.Rnw
updates on manuscript writing (mostly on data)

@@
 Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset \citep{Abatzoglou2013}.
-gridMET is a gridded daily weather dataset with a spatial resolution of \textasciitilde4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
+gridMET is a gridded daily weather dataset with a spatial resolution of \textasciitilde 4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
 Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample.
@@
 Seasonal water balance estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell.
+Any county-year combinations that had an extreme water balance value (lower than -200 or greater than 1200) were dropped.
 Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
@@
 $K$ is the total number of spline basis functions.
-$K = 3$ was chosen for this regression as it minimizes Bayesian Information Criterion (BIC).
+$K = 3$ was chosen for this regression to prevent our estimated response from becoming unrealistically oscillatory while maintaining a good fit based on Bayesian Information Criterion (BIC).
 To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
@@
 $K$ and $L$ are the numbers of basis functions for $D$ and $A$, respectively.
-$K = 4$ and $L = 4$ were chosen as they minimize BIC.
+$K = 4$ and $L = 4$ were chosen to balance preventing the estimated responses from being unrealistically oscillatory and ensuring a good fit as indicated by BIC.
+For this regression, counties with aquifer thickness levels greater than 180 m were excluded due to the sparsity of observations in that range because our flexible functional form approach cannot reliably estimate its impact.
 County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by \(X_i\).
@@
 Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}.
-This data is used for Figure \ref{fig:sat-map}.
+County and State boundaries were obtained using the R tigris package \citep{Rtigris}.
+These datasets were used for Figure \ref{fig:sat-map}.
 Corn and soybean yield data were obtained from the U.S.
@@
 This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-Weather data (precipitation, temperature, and evapotranspiration) was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}.
-This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
+Weather data was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}.
+Elevation data was obtained from the gridMET website at \url{https://www.climatologylab.org/gridmet.html}.
+These datasets were used to calculate water balance, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
 The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}.
 This data is used for Figure \ref{fig:sat-map} and the regression analysis.
-Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which is used for regression analysis.
+Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which was used for regression analysis.
 The R soilDB package \citep{Rsoildb} was used to download the data.
@@
 Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
-The data used for the regression analysis can be generated using our codes by following the instructions available at the linked Github repository.
 
 \subsection{Code Availability}
-All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All analyses were performed in R \citep{R}.
+All data processing and analyses were carried out using R \citep{R}.
+The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}.
 

### 2023-11-01 Writing/manuscript.tex
updates on manuscript writing (mostly on data)

@@
 
+\usepackage{Sweave}
 \begin{document}
@@
 For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is 1.15 tonne/ha.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is 1.31 tonne/ha.
 Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
@@
 
-% At low water deficits (below 200mm), Figure \ref{fig:yield-dif} illustrates that counties with the highest levels of aquifer thickness (i.e., 3rd quantile in Figure \ref{fig:yield-response}) experience statistically lower yields for corn and soybean than counties with the lowest level of aquifer thickness (1st quantile).
-In case of soybean, irrigated yields for the lowest aquifer thickness category converge to levels comparable to those obtained from rainfed production.
-These counter-intuitive results reflect a combination of two factors.
-First, in years with small water deficits there is less need for irrigation and hence irrigated and rainfed yield differences are reduced.
-Second, years with small or even negative water deficits are associated with wetter conditions during the growing season, which can negatively impact yield if waterlogging occurs leading to greater risks of physical crop damage (e.g., lodging of grains), delays to planting/harvesting, nutrient leaching, and restricted root growth in waterlogged soils \citep{li2019excessive}.
-Declines in irrigated yields could be indicative of over-irrigation by farmers, which has been shown to occur in the HPA region in wet years \citep{foster2019assessing, gibson2017case, gibson2019benchmarking}, exacerbating these risks.
-However, the small declines in irrigated yields at low or negative levels of water deficit may also reflect the fact that wetter years are typically cooler, reducing positive heat effects on crop growth and yield formation.
-
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_yield_intensive}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}
 
@@
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $8,773$ and $5,977$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
+Total number of observations are $8,814$ and $5,997$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
 \end{figure}
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100 m to 0.72 (72\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.89 (89\%) for an average county-level aquifer thickness of 100 m to 0.7 (70\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
@@
 The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $2,523$ and $1,558$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
+Total number of observations are $2,464$ and $1,553$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
 \end{figure}
@@
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.09 tonne/ha of production loss for corn.
+At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.15 tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.57 tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.45 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 
-These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($8,773$ observations), (2) per-acre soybean yield ($5,977$ observations), (3) share of irrigated corn ($2,523$ observations), and (4) share of irrigated soybean ($1,558$ observations).
+These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($8,814$ observations), (2) per-acre soybean yield ($5,997$ observations), (3) share of irrigated corn ($2,464$ observations), and (4) share of irrigated soybean ($1,553$ observations).
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\subsection{Data Availability}\label{data-and-code-availability}}
+\hypertarget{data-availability}{%
+\subsection{Data Availability}\label{data-availability}}
 
-All the codes used for this study are available at the following Github repository (\href{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}{link}).
-All analyses were performed in R \citep{R} and all the raw data-sets used in this study are publicly accessible.
+All the raw data-sets used in this study are publicly accessible.
 The boundary shape file of the High Plains aquifer was obtained from the U.S.
 Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}.
-This data is used for Figure \ref{fig:sat-map}.
+County and State boundaries were obtained using the R tigris package \citep{Rtigris}.
+These datasets were used for Figure \ref{fig:sat-map}.
 Corn and soybean yield data were obtained from the U.S.
@@
 Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
-The data used for the regression analysis can be generated using our codes by following the instructions available at the linked Github repository.
 
 \subsection{Code Availability}
+All data processing and analyses were carried out using R \citep{R}.
+The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}.
 

### 2023-11-01 Codes/2_figures_tables.rmd
separate rnw file created for NW submission. Figures have been modified.
1. font: Helvetica
2. title and legend separately

@@
 data_soy <- all_results[crop == "soy", data][[1]]
-```
-
-## Set themes
-```{r}
-theme_fig <-
-  theme_bw() +
-  theme(
-    axis.title.x =
-      element_text(
-        size = 9, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
-      ),
-    axis.title.y =
-      element_text(
-        size = 9, angle = 90, hjust = .5, vjust = .9, family = "Times"
-      ),
-    axis.text.x =
-      element_text(
-        size = 8, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
-      ),
-    axis.text.y =
-      element_text(
-        size = 8, angle = 0, hjust = 1, vjust = 0, family = "Times"
-      ),
-    axis.ticks =
-      element_line(
-        linewidth = 0.3, linetype = "solid"
-      ),
-    axis.ticks.length = unit(.15, "cm"),
-    #--- legend ---#
-    legend.text =
-      element_text(
-        size = 9, angle = 0, hjust = 0, vjust = 0.5, family = "Times"
-      ),
-    legend.title =
-      element_text(
-        size = 9, angle = 0, hjust = 0, vjust = 0, family = "Times"
-      ),
-    legend.key.size = unit(0.5, "cm"),
-    #--- strip (for faceting) ---#
-    strip.text = element_text(size = 9, family = "Times"),
-    #--- plot title ---#
-    plot.title = element_text(family = "Times", face = "bold", size = 9),
-    #--- margin ---#
-    # plot.margin = margin(0, 0, 0, 0, "cm"),
-    #--- panel ---#
-    panel.grid.minor = element_blank(),
-    panel.background = element_blank(),
-    panel.border = element_rect(fill = NA)
-  )
 ```
@@
     legend.key.width = unit(1, "cm"),
-    plot.title = element_text(size = 9, family = "Times"),
-    legend.title = element_text(family = "Times"),
-    legend.text = element_text(family = "Times")
+    plot.title = element_text(size = 7, family = "Helvetica"),
+    legend.title = element_text(size = 7, family = "Helvetica"),
+    legend.text = element_text(size = 7, family = "Helvetica")
   ) +
@@
         ylab("") +
-        # scale_x_continuous(breaks = (-2:11)*100) +
         coord_cartesian(xlim = c(-200, 1150))
-      # +
-      # # xlim(NA, max_balance) +
-      # theme_void()
     }
@@
   theme(
-    plot.title = element_text(size = 9, family = "Times", face = "bold"),
     axis.text.x = element_blank(),
@@
         theme(
-          plot.title = element_text(size = 9, family = "Times", face = "bold"),
           axis.text.x = element_blank(),
@@
     legend.position = "bottom",
-    strip.background = element_blank(),
-    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = 9)
+    strip.background = element_blank()
   )
@@
         strip.background = element_blank(),
-        strip.text = element_text(size = 9, family = "Times"),
         panel.spacing.y = unit(1, "cm")

### 2023-11-01 NatureWaterSubmission/nw_manuscript.Rnw
separate rnw file created for NW submission. Figures have been modified.
1. font: Helvetica
2. title and legend separately

--- 2023-11-01 Writing/manuscript.Rnw
@@
 \usepackage{lmodern}
+\usepackage[labelfont=bf]{caption} % \caption*{} will create legend (explanatory text)
 \usepackage{iftex}
@@
 
-<<init, echo = FALSE>>=
+<<init, include = FALSE>>=
 knitr::opts_chunk$set(
@@
 
-Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}\footnote{Another pathway is through groundwater use regulation in response to low or declining aquifer thickness.
-However, it is unlikely that this effects is captured in our analysis.
-Please see Appendix \ref{pathway-reg}}.
+Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
 First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
@@
 Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}.
+Another potential pathway is through groundwater use regulation in response to low or declining aquifer thickness.
+However, it is unlikely that this effects is captured in our analysis.
+Please see Appendix \ref{pathway-reg}.
 
@@
 \hypertarget{impact-intensive}{%
-\subsection{Aquifer depletion increases sensitivity of crop yields to water deficits}\label{impact-intensive}}
+\subsection{Sensitivity of per-area crop yields to water deficits}\label{impact-intensive}}
 
-Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B)\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.}.
+Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B).
+Confidence intervals were not presented to avoid over-crowding the figures.
+Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.
 The histogram of annual water deficit is presented at the top for each crop.
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}
-
-}
-
 <<num_obs_yield_reg, echo = FALSE>>=
@@
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.
-Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}\label{fig:yield-response}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_ir_yield_dif}
-
-}
-
-\caption{The difference in estimated county-level irrigated corn (top panels) and soybean (bottom panels) yields for counties in the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.}\label{fig:yield-dif}
-\end{figure}
-
 \hypertarget{impact-share}{%
-\subsection{Aquifer depletion substantially reduces the share of agricultural land under irrigation}\label{impact-share}}
+\subsection{Share of agricultural land under irrigation}\label{impact-share}}
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}
-
-}
-
 <<num_obs_share_reg, echo = FALSE>>=
@@
 
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains (Note: shaded area represents 95\% confidence interval).
-The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}\label{fig:ir-share}
-\end{figure}
-
 \hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}{%
-\subsection{\texorpdfstring{Aquifer depletion increases overall drought-related production losses by up to \(25\%\)}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
+\subsection{\texorpdfstring{Overall drought-related production losses}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
 
@@
 This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers' ability to effectively and reliably buffer crops against water deficits during the growing season.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}
-
-}
-
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}
-
-}
-
-\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}\label{fig:dif-tot-impact}
-\end{figure}
 
@@
 From these data, we computed the seasonal water deficit (\(D_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
-Seasonal water balance estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell.
-Any county-year combinations that had an extreme water balance value (lower than -200 or greater than 1200) were dropped.
+Seasonal water deficit estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell.
+Any county-year combinations that had an extreme water deficit value (lower than -200 or greater than 1200) were dropped.
 Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
@@
 For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables were calculated.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}
-
-}
-
-\caption{Historical rainfed yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
-\end{figure}
 
@@
 
-These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($\Sexpr{nobs_corn}$ observations), (2) per-acre soybean yield ($\Sexpr{nobs_soy}$ observations), (3) share of irrigated corn ($\Sexpr{nobs_corn_share}$ observations), and (4) share of irrigated soybean ($\Sexpr{nobs_soy_share}$ observations).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}
-
-}
-
-\caption{Map showing county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
-Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
-\end{figure}
+These data were combined to form four sets of data for regression analysis: (1) per-area corn yield ($\Sexpr{nobs_corn}$ observations), (2) per-area soybean yield ($\Sexpr{nobs_soy}$ observations), (3) share of irrigated corn ($\Sexpr{nobs_corn_share}$ observations), and (4) share of irrigated soybean ($\Sexpr{nobs_soy_share}$ observations).
 
@@
 where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water balance is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(D_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
+The impact of average water deficit is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(D_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
 The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
@@
 Elevation data was obtained from the gridMET website at \url{https://www.climatologylab.org/gridmet.html}.
-These datasets were used to calculate water balance, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
+These datasets were used to calculate water deficit, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
 The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}.
@@
 
-Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007.
-Timothy Foster was supported by Innovate UK award no.
+Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
+This research was also supported by Innovate UK award no.
 10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.'' The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
@@
 \begin{itemize}
-\item OCE 58-0111-20-007 (TM, SK, NB)
-\item OCE 58-0111-21-007 (TM, SK, NB)
+\item OCE 58-0111-20-007 (TM, SK)
+\item OCE 58-0111-21-007 (TM, SK)
 \item Innovate UK award no. 10044695 (TF)
+\item Daugherty Water for Food Global Institute (SK)
 \end{itemize}
@@
 \subsection*{Competing Interests Statement}
+
+\section*{Tables}
+
+\section*{Figures}
+
+\begin{figure}[H]
+
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.}
+
+% {\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}}
+
+\caption*{The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
+Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}
+\label{fig:yield-response}
+\end{figure}
+
+%------------------------
+%- Figure 2
+%------------------------
+
+\begin{figure}[H]
+
+\caption{The difference in estimated county-level irrigated corn and soybean yields}
+
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_ir_yield_dif}}
+
+\caption*{The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
+Shaded area represents 95\% confidence interval.}
+\label{fig:yield-dif}
+\end{figure}
+
+%------------------------
+%- Figure 3
+%------------------------
+
+\begin{figure}[H]
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains}
+
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}}
+
+\label{fig:ir-share}
+\caption*{The solid black line represents the estimated share of irrigated acres.
+Shaded area represents 95\% confidence interval.
+The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
+Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}
+\end{figure}
+
+%------------------------
+%- Figure 4
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Average productivity of corn and soybean in the US High Plains}
+
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
+\caption*{The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}
+\label{fig:tot-impact}
+\end{figure}
+
+%------------------------
+%- Figure 5
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels}
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}}
+
+\caption*{The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}
+\label{fig:dif-tot-impact}
+\end{figure}
+
+%------------------------
+%- Figure 6
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Historical rainfed yield, irrigated yield, and water deficit}
+
+% {\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}}
+
+\caption*{The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
+\end{figure}
+
+%------------------------
+%- Figure 7
+%------------------------
+
+\begin{figure}[H]
+
+\caption{County-average aquifer thickness in 2016}
+
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}}
+
+\caption{The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
+\end{figure}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{A.\arabic{figure}}
+\renewcommand{\thefigure}{B.\arabic{figure}}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
+\renewcommand{\thefigure}{C.\arabic{figure}}
 
@@
 
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}\label{fig:state-fe-with-without}
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects.
+Shaded area represents 95\% confidence interval.}\label{fig:state-fe-with-without}
 \end{figure}
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
+\renewcommand{\thefigure}{E.\arabic{figure}}
 

### 2023-11-01 NatureWaterSubmission/nw_manuscript.tex
separate rnw file created for NW submission. Figures have been modified.
1. font: Helvetica
2. title and legend separately

--- 2023-11-01 NatureWaterSubmission/nw_manuscript.Rnw
@@
 \documentclass[
-]{article}
+]{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{amsmath,amssymb}
@@
 
-<<init, include = FALSE>>=
-knitr::opts_chunk$set(
-  echo = FALSE
-)
-library(tidyverse)
-library(data.table)
-@
-
 \title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
@@
 \date{}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<load_results, echo = FALSE>>=
-setwd(here::here("Writing"))
-all_results <- readRDS("../Results/all_results.rds")
-@
-
 \hypertarget{impact-intensive}{%
-\subsection{Sensitivity of per-area crop yields to water deficits}\label{impact-intensive}}
+\subsection{Sensitivity of per-acre crop yields to water deficits}\label{impact-intensive}}
 
@@
 
-<<echo = FALSE>>=
-delta_yield_lowest_highest <- 
-  all_results[crop == "corn", yield_dif_test][[1]] %>%
-  .[comp_q == 3, ] %>%  
-  .[which.min(abs(balance -950)), dif_hat] %>%
-  round(digits = 2)
-@
-
 Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95\% confidence interval for corn and soybean.
 For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is \Sexpr{delta_yield_lowest_highest} tonne/ha.
+At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is 1.31 tonne/ha.
 Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
 This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower aquifer thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
-
-<<num_obs_yield_reg, echo = FALSE>>=
-nobs_corn <- 
-  all_results[crop == "corn", data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-
-nobs_soy <- all_results[crop == "soy", data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-@
 
@@
 
-<<ir_share, echo = FALSE>>=
-corn_share <- all_results[crop == "corn", share_pred_data][[1]]
-
-share_high <-
-  corn_share[which.min(abs(sat - 130)), ir_share_hat] %>%
-  round(digits = 2)
-
-share_low <-
-  corn_share[which.min(abs(sat - 10)), ir_share_hat] %>%
-  round(digits = 2) 
-@
-
 Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.89 (89\%) for an average county-level aquifer thickness of 100 m to 0.7 (70\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
-
-<<num_obs_share_reg, echo = FALSE>>=
-nobs_corn_share <- 
-  all_results[crop == "corn", share_reg_data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-
-nobs_soy_share <- 
-  all_results[crop == "soy", share_reg_data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-@
 
@@
 \subsection{\texorpdfstring{Overall drought-related production losses}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
-
-<<res_at_950, echo = FALSE>>=
-res_at_950 <-
-  all_results[, .(crop, avg_yield_pred_data)] %>%
-  unnest(cols = c(avg_yield_pred_data)) %>%
-  data.table() %>%
-  .[crop == "corn", ] %>%
-  .[(abs(balance - 950)) < 5, ]
-@
 
@@
 Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only \Sexpr{round(res_at_950[sat == 100 & crop == "corn", avg_yield] - res_at_950[sat == 70 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
+At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only 0.15 tonne/ha of production loss for corn.
 This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in \Sexpr{round(res_at_950[sat == 70 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to \Sexpr{round(res_at_950[sat == 40 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
+Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in 0.95 tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to 1.45 tonne/ha of production loss for corn.
 Similar but less pronounced effects are also observed for soybean.
@@
 
-These data were combined to form four sets of data for regression analysis: (1) per-area corn yield ($\Sexpr{nobs_corn}$ observations), (2) per-area soybean yield ($\Sexpr{nobs_soy}$ observations), (3) share of irrigated corn ($\Sexpr{nobs_corn_share}$ observations), and (4) share of irrigated soybean ($\Sexpr{nobs_soy_share}$ observations).
+These data were combined to form four sets of data for regression analysis: (1) per-acre corn yield ($8,814$ observations), (2) per-acre soybean yield ($5,997$ observations), (3) share of irrigated corn ($2,464$ observations), and (4) share of irrigated soybean ($1,553$ observations).
 
@@
 
-\caption*{The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+\caption*{Irrigated yield responses to water deficits are dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}
+Total number of observations are $8,814$ and $5,997$ for corn and soybean regressions, respectively.}
 \label{fig:yield-response}
@@
 
-\caption*{The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
+\caption*{The solid red lines represent the difference in estimated per-acre irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
 Shaded area represents 95\% confidence interval.}
@@
 \begin{figure}[H]
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains}
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains.}
 
@@
 The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}
+Total number of observations are $2,464$ and $1,553$ for corn and soybean regressions, respectively.}
 \end{figure}
-
-%------------------------
-%- Figure 4
-%------------------------
 
@@
 
-\caption{Average productivity of corn and soybean in the US High Plains}
+% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
 
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
-\caption*{The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}
-\label{fig:tot-impact}
+\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.}\label{fig:tot-impact}
 \end{figure}
-
-%------------------------
-%- Figure 5
-%------------------------
 
@@
 
-\caption{Difference in average yield of corn and soybean for aquifer thickness levels}
 % {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}}
 
-\caption*{The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}
-\label{fig:dif-tot-impact}
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}\label{fig:dif-tot-impact}
 \end{figure}
-
-%------------------------
-%- Figure 6
-%------------------------
 
@@
 
-\caption{Historical rainfed yield, irrigated yield, and water deficit}
-
 % {\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}}
 
-\caption*{The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
+\caption{Historical rainfed yield, irrigated yield, and water deficit.
+The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
 \end{figure}
-
-%------------------------
-%- Figure 7
-%------------------------
 
@@
 
-\caption{County-average aquifer thickness in 2016}
-
 % {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}}
 
-\caption{The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+\caption{Map showing county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
 Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
@@
 
-<<sat_ref, echo = FALSE>>=
-sat_ref <-
-  readRDS(here::here("Shared/Results/all_results.rds")) %>%
-  .[1, sat_seq] %>%
-  .[[1]] %>%
-  .[length(.)]
-@
-
-Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of \Sexpr{sat_ref} m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
+Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of 100 m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
 Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
-Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to \Sexpr{sat_ref} m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to 100 m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
 
@@
 
-\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
+\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of 100 m (black line).
 The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}

### 2023-11-01 Writing/manuscript.Rnw
separate rnw file created for NW submission. Figures have been modified.
1. font: Helvetica
2. title and legend separately

@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 \maketitle

### 2023-11-01 Codes/2_figures_tables.R

@@ -1,29 +1,29 @@
-## ---------------------------------------------------------------
+## --------------------------------------------------------
 all_results <- readRDS("Results/all_results.rds")
 
 data_corn <- all_results[crop == "corn", data][[1]]
 data_soy <- all_results[crop == "soy", data][[1]]
 
 
-## ---------------------------------------------------------------
+## --------------------------------------------------------
 theme_fig <-
   theme_bw() +
   theme(
     axis.title.x =
       element_text(
-        size = 9, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
+        size = 7, angle = 0, hjust = .5, vjust = -0.3, family = "Helvetica"
       ),
     axis.title.y =
       element_text(
-        size = 9, angle = 90, hjust = .5, vjust = .9, family = "Times"
+        size = 7, angle = 90, hjust = .5, vjust = .9, family = "Helvetica"
       ),
     axis.text.x =
       element_text(
-        size = 8, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
+        size = 7, angle = 0, hjust = .5, vjust = 1.5, family = "Helvetica"
       ),
     axis.text.y =
       element_text(
-        size = 8, angle = 0, hjust = 1, vjust = 0, family = "Times"
+        size = 7, angle = 0, hjust = 1, vjust = 0, family = "Helvetica"
       ),
     axis.ticks =
       element_line(
@@ -33,17 +33,17 @@ theme_fig <-
     #--- legend ---#
     legend.text =
       element_text(
-        size = 9, angle = 0, hjust = 0, vjust = 0.5, family = "Times"
+        size = 7, angle = 0, hjust = 0, vjust = 0.5, family = "Helvetica"
       ),
     legend.title =
       element_text(
-        size = 9, angle = 0, hjust = 0, vjust = 0, family = "Times"
+        size = 7, angle = 0, hjust = 0, vjust = 0, family = "Helvetica"
       ),
     legend.key.size = unit(0.5, "cm"),
     #--- strip (for faceting) ---#
-    strip.text = element_text(size = 9, family = "Times"),
+    strip.text = element_text(size = 7, family = "Helvetica"),
     #--- plot title ---#
-    plot.title = element_text(family = "Times", face = "bold", size = 9),
+    plot.title = element_text(family = "Helvetica", face = "bold", size = 7),
[…]

### 2023-11-01 Codes/A_1_state_fe.R

@@ -1,8 +1,8 @@
-## ---------------------------------------------------------------
+## --------------------------------------------------------
 all_results <- readRDS(here::here("Results/all_results.rds"))
 
 
-## ---------------------------------------------------------------
+## --------------------------------------------------------
 #++++++++++++++++++++++++++++++++++++
 #+ Formula
 #++++++++++++++++++++++++++++++++++++
@@ -85,7 +85,7 @@ results <-
   data.table()
 
 
-## ---------------------------------------------------------------
+## --------------------------------------------------------
 g_share_comp <-
   results[sat <= 150, ] %>%
   .[, crop := case_when(
@@ -122,30 +122,30 @@ ggsave(
 )
 
 
-## ---------------------------------------------------------------
-ir_share_data_corn[, sat_m := measurements::conv_unit(sat, "ft", "m")]
+## --------------------------------------------------------
+ir_corn_share_data <- all_results[crop == "corn", share_reg_data][[1]]
 
 sy_demeaan_data <-
-  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = .(state_name, year)] %>%
+  copy(ir_corn_share_data)[, .(sat = sat - mean(sat)), by = .(state_name, year)] %>%
   .[, type := "B. Demeaned by State-Year Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
 c_y_demeaan_data <-
-  copy(ir_share_data_corn)[, sat_m := sat_m - mean(sat_m), by = sc_code] %>%
-  .[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  copy(ir_corn_share_data)[, sat := sat - mean(sat), by = sc_code] %>%
+  .[, .(sat = sat - mean(sat)), by = year] %>%
   .[, type := "C. Demeaned by County and Yearly Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
 y_demeaan_data <-
-  copy(ir_share_data_corn)[, .(sat_m = sat_m - mean(sat_m)), by = year] %>%
+  copy(ir_corn_share_data)[, .(sat = sat - mean(sat)), by = year] %>%
   .[, type := "A. Demeaned by Yearly Average"] %>%
-  .[, .(sat_m, type)]
+  .[, .(sat, type)]
 
[…]

### 2023-11-01 Codes/functions.R

@@ -1141,6 +1141,55 @@ test_dif_in_yield <- function(balance_ls, base_q, comp_q, yield_semi_res, sat_te
     rbindlist()
 }
 
+#!===========================================================
+#! Figure Theme
+#!===========================================================
+theme_fig <-
+  theme_bw() +
+  theme(
+    axis.title.x =
+      element_text(
+        size = 7, angle = 0, hjust = .5, vjust = -0.3, family = "Helvetica"
+      ),
+    axis.title.y =
+      element_text(
+        size = 7, angle = 90, hjust = .5, vjust = .9, family = "Helvetica"
+      ),
+    axis.text.x =
+      element_text(
+        size = 7, angle = 0, hjust = .5, vjust = 1.5, family = "Helvetica"
+      ),
+    axis.text.y =
+      element_text(
+        size = 7, angle = 0, hjust = 1, vjust = 0, family = "Helvetica"
+      ),
+    axis.ticks =
+      element_line(
+        linewidth = 0.3, linetype = "solid"
+      ),
+    axis.ticks.length = unit(.15, "cm"),
+    #--- legend ---#
+    legend.text =
+      element_text(
+        size = 7, angle = 0, hjust = 0, vjust = 0.5, family = "Helvetica"
+      ),
+    legend.title =
+      element_text(
+        size = 7, angle = 0, hjust = 0, vjust = 0, family = "Helvetica"
+      ),
+    legend.key.size = unit(0.5, "cm"),
+    #--- strip (for faceting) ---#
+    strip.text = element_text(size = 7, family = "Helvetica"),
+    #--- plot title ---#
+    plot.title = element_text(family = "Helvetica", face = "bold", size = 7),
+    #--- margin ---#
+    # plot.margin = margin(0, 0, 0, 0, "cm"),
+    #--- panel ---#
+    panel.grid.minor = element_blank(),
+    panel.background = element_blank(),
+    panel.border = element_rect(fill = NA)
+  )
+
 # !===========================================================
 # ! Generic
 # !===========================================================

### 2023-11-02 Codes/0_1_hpa_counties.Rmd
metadata completed

@@
 (
-  US_county <- tigris::counties(cb = TRUE) %>%
+  US_county <-
+    tigris::counties(cb = TRUE) %>%
     st_as_sf() %>%
@@
 Additional state codes:
-these supplement the dryland production data for understanding the impact of drought severity
+these may supplement the dryland production data for understanding the impact of drought severity
 on dryland crop production

### 2023-11-02 Codes/0_2_quickstat.Rmd
metadata completed

@@
     )] %>%
-    # === irrigated or not? ===#
+    # === irrigated or not ===#
     .[, ir := ifelse(

### 2023-11-02 Codes/0_3_get_sat.Rmd
metadata completed

@@
 # Preparation
-Import saturated thickness data
+Import saturated thickness data to get the CRS.
 
 ```{r }
-sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
+sat_crs <-
+  raster("Data/data-raw/saturated_thickness/SatThick_19851.tif") %>%
+  raster::projection()
 ```
@@
   st_as_sf() %>%
-  st_transform(projection(sat_temp))
+  st_transform(sat_crs)
 ```

### 2023-11-02 Codes/0_4_get_weather.Rmd
metadata completed

@@
     # === total precip ===#
-    prcp = sum(prcp),
-    # === GDD ===#,
-    gdd = sum(t_min_max_2),
-    # === EDD ===#
-    edd = sum(pmax(0, tmax - 32)),
-    # === number of days above x ===#
-    days_ab_30 = sum(tmax >= 30),
-    # === mean temp ===#
-    mean_temp = mean(tmin + tmax) / 2
+    prcp = sum(prcp)
   ),
@@
 
-saveRDS(sum_gm_data, here("Data/data-processed", "summrized_gridMET.rds"))
+saveRDS(sum_gm_data, here("Data/data-processed", "summarized_gridMET.rds"))
 ```

### 2023-11-02 Codes/0_6_merge_datasets.rmd
metadata completed

@@
 #' Weather
-weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))
+weather_data <- readRDS(here("Data/data-processed/summarized_gridMET.rds"))
 
@@
   .[!(in_hpa == 0 & ir == "ir"), ] %>%
-  # === if not irrigated, set sat = 0 ===#
-  .[ir == "nir", sat := 0] %>%
-  # === if sat == 0, but indicated irrigated ===#
-  # add 0.1 to distinguish dryland production, where
-  # sat is set at 0 above
-  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
-  # === drop if yield or acres are not observed ===#
+  #--- drop if irrigated and year > 2016 ---#
+  # drops irrigated produciton observations that are missing sat (2017, 2018 fow which sat is not observed).
+This will keep all the dryland production observations beyond 2016 because sat does not have to be observed for the dryland production.
+  .[!(ir == "ir" & year > 2016), ] %>%
+  # === drop if both yield and acres are not observed ===#
   .[!is.na(yield) & !is.na(acres), ] %>%
-  # === if sat is missing, drop ===#
-  #' just 2017 and 2018 data because saturated thickness data
-  #' is observed only until 2016
-  .[!is.na(sat), ] %>%
-  #--- indicator of whether dryland (rainfed) production or not ---#
-  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
-  #--- sc_code and dry_or_not (used for FE regression) ---#
-  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
+  #--- set sat = NA if not irrigated ---#
+  .[ir == "nir", sat := NA] %>%
+  #--- sc_code and irrigation status (used for FE regression) ---#
+  .[, sc_dry := paste0(sc_code, "_", ir)] %>%
   #--- state-year combination ---#
@@
   .[, num_obs := .N, by = sc_dry] %>%
-  # .[, num_obs := .N, by = sc_code] %>%
   # === keep only if more than 10 observations are available ===#
   .[num_obs >= 10, ] %>%
+  .[, num_obs := NULL] %>%
   # === order (not necessary) ===#
-  .[order(sc_code, year), ]
+  .[order(sc_code, year), ] %>%
+  dplyr::relocate(sc_code, year)
 ```

### 2023-11-02 Writing/nw_manuscript.Rnw
metadata completed

--- 2023-11-01 NatureWaterSubmission/nw_manuscript.Rnw
@@
 \usepackage{lmodern}
-\usepackage[labelfont=bf]{caption} % \caption*{} will create legend (explanatory text)
+\usepackage[labelfont=bf, textfont=footnotesize]{caption} % \caption*{} will create legend (explanatory text)
 \usepackage{iftex}
@@
 However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks \citep{grafton2018paradox,perez2021agricultural}.
-Future research should explore potential strategies that farmers could adopt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
+Future research should explore potential strategies that farmers could adapt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
 
@@
 As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
-These data provide annual estimates of the aquifer thickness at 250m resolution since 1935.
+These data provide annual estimates of the aquifer thickness at 250 m resolution since 1935.
 Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
@@
 \begin{equation}
-Y_{i,t} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(D_{i,t})\cdot C_j  + \alpha_i + \lambda_t + v_{i,t} \label{eq:eqn1}
+Y_{i,t,ir} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(D_{i,t})\cdot C_j  + \alpha_{i,ir} + \lambda_t + v_{i,t,ir} \label{eq:eqn1}
 \end{equation}
 
-where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(Y\) is crop yield, and \(D\) is seasonal water deficit.
+where county and year are indicated by subscripts \(i\) and \(t\), respectively, whether rainfed or irrigated is indicated by $ir$,  \(Y\) is crop yield, and \(D\) is seasonal water deficit.
 \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
@@
 $K = 3$ was chosen for this regression to prevent our estimated response from becoming unrealistically oscillatory while maintaining a good fit based on Bayesian Information Criterion (BIC).
-To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
+To control for time-invariant heterogeneity across counties and yearly shocks, county-irrigation-status fixed effects $\alpha_{i, ir}$ and year fixed effects $\lambda_t$ are also included.
+County-irrigation-status fixed effects are used instead of county fixed effects because some counties have both rainfed and irrigated observations, and they need to be distinguished.
+Standard error estimation is clustered by county.
 
@@
 
+%------------------------
+%- Figure 1
+%------------------------
+
 \begin{figure}[H]
@@
 
-\caption*{The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+\caption*{Note: The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
@@
 
-\caption*{The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
+\caption*{Note: The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
 Shaded area represents 95\% confidence interval.}
@@
 \label{fig:ir-share}
-\caption*{The solid black line represents the estimated share of irrigated acres.
+\caption*{Note: The solid black line represents the estimated share of irrigated acres.
 Shaded area represents 95\% confidence interval.
@@
 % {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
-\caption*{The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+\caption*{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
 See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
@@
 
-\caption*{The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+\caption*{Note: The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
 Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}
@@
 
-\caption*{The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
+\caption*{Note: The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
 \end{figure}
@@
 
-\caption{The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
-Counties highlighted in grey are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
+\caption*{Note: The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+Counties highlighted in gray are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
 \end{figure}
 
+\clearpage
+
+%==============================================
+%= References
+%==============================================
+
+\bibliography{DRA.bib}
+
+%==============================================
+%= Supplementary Information
+%==============================================
 \clearpage
@@
 This is due to laws that typically provide farmers with absolute rights to use groundwater under their land, coupled with low levels of metering.
-Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigators in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
+Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigating farmers in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
 As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}
-
-}
-
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-corn}
-\end{figure}
+%------------------------
+%- Figure B.1
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}}
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-soy}
+\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of corn in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}
+\label{fig:irrigated-yield-ind-corn}
 \end{figure}
+
+%------------------------
+%- Figure B.2
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}}
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:avg-yield-ind}
+\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of soybean in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}
+\label{fig:irrigated-yield-ind-soy}
+\end{figure}
+
+%------------------------
+%- Figure B.3
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Average productivity of corn and soybean}
+
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}}
+
+\caption*{Note: The black solid lines represent average yield of corn (top panel) and soybean (bottom panel) in the US High Plains for different levels of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}
+\label{fig:avg-yield-ind}
 \end{figure}
@@
 Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of \Sexpr{sat_ref} m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
-Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
+Gray shaded areas represented the 95\% confidence intervals for the estimated relationships.
 Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to \Sexpr{sat_ref} m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+
+%------------------------
+%- Figure C.1
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}
+\caption{Decline in the share of acres that are irrigated}
 
-}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}}
 
-\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
-The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
+\caption*{Note: The solid black lines represent the share of acres that are irrigated as a function of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
+The gray shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
 \end{figure}
@@
 
+%------------------------
+%- Figure D.1
+%------------------------
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}
+\caption{Comparison of variations in aquifer thickness after demeaning}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
 
-\caption{Comparison of variations in aquifer thickness after demeaning by yearly average, state-year average, and county average.}\label{fig:variation-left}
+\caption*{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}
+\label{fig:variation-left}
+
 \end{figure}
@@
 For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
-Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
 We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
+
+%------------------------
+%- Figure D.2
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
 
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects.
-Shaded area represents 95\% confidence interval.}\label{fig:state-fe-with-without}
+\caption*{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
+Gray shaded area represents 95\% confidence interval.}
+\label{fig:state-fe-with-without}
 \end{figure}
@@
 
+%------------------------
+%- Figure E.1
+%------------------------
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}
+\caption{Example of a series of thin-plate spline basis functions}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}}
 
-\caption{Example of a series of thin-plate spline basis functions}\label{fig:sp-basis}
+\label{fig:sp-basis}
 \end{figure}
@@
 
+\caption{Example of a function represented by different coefficient values}
+
 {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}
@@
 
-\caption{Example of a function represented by different coefficient values}\label{fig:ex-fx}
+\label{fig:ex-fx}
 \end{figure}
 
-\clearpage
-
-  \bibliography{DRA.bib}
-
 \end{document}

### 2023-11-02 Codes/0_3_get_sat.R

@@ -1,17 +1,19 @@
-## ---------------------------------------------------------------
-sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
+## -------------------------------------------------------------------
+sat_crs <-
+  raster("Data/data-raw/saturated_thickness/SatThick_19851.tif") %>%
+  raster::projection()
 
 
-## ---------------------------------------------------------------
+## -------------------------------------------------------------------
 hpa_counties <-
   here("Data/data-processed/base_counties.rds") %>%
   readRDS() %>%
   .[in_hpa == 1, ] %>%
   st_as_sf() %>%
-  st_transform(projection(sat_temp))
+  st_transform(sat_crs)
 
 
-## ---------------------------------------------------------------
+## -------------------------------------------------------------------
 get_sat <- function(y) {
   # y <- 2000
   sat_temp <-
@@ -32,13 +34,13 @@ get_sat <- function(y) {
 }
 
 
-## ----extract-sat, eval = F--------------------------------------
+## ----extract-sat, eval = F------------------------------------------
 ## sat_all <-
 ##   mclapply(1985:2016, get_sat, mc.cores = 12) %>%
 ##   rbindlist() %>%
 ##   .[sat < 0, sat := 0]
 
 
-## ----save, eval = F---------------------------------------------
+## ----save, eval = F-------------------------------------------------
 ## saveRDS(sat_all, here("Data/data-processed/sat_all.rds"))
 

### 2023-11-02 Codes/0_data_prep.R

@@ -1,129 +1,0 @@
-#' ---
-#' title: "Data preparation"
-#' author: "Taro Mieno"
-#' output:
-#'   html_document:
-#'     number_sections: yes
-#'     theme: flatly
-#'     highlight: zenburn
-#'     toc_float: yes
-#'     toc: yes
-#'     toc_depth: 3
-#' geometry: margin=1in
-#' ---
-
-#+ setup, include=FALSE
-
-library(knitr)
-knitr::opts_chunk$set(
-  echo = TRUE,
-  cache = TRUE,
-  comment = NA,
-  message = FALSE,
-  warning = FALSE,
-  tidy = FALSE,
-  cache.lazy = FALSE,
-  #--- figure ---#
-  dpi = 400,
-  fig.width = 7.5,
-  fig.height = 5,
-  out.width = "750px",
-  out.height = "500px"
-)
-
-library(here)
-
-opts_knit$set(
-  root.dir = here()
-)
-
-# /*=================================================*/
-#' # Objectives
-# /*=================================================*/
-#' Prepare datasets for econometric analysis of the impacts of saturated thickness on the
-#' impacts of drought on crop yield
-#'
-
-#' + Irrigated corn and soybean yields at the county level from 2000 to 2018 within the High Plains Aquifer measured in bushels per acre, downloaded from the National Agricultural Statistic Service (NASS) website.
-#' + County-level depths of saturated thickness of High Plains Aquifer form the 2000 to 2018, measured in feet.
-#' + For different degrees of drought, we use the Drought Index used by the U.S. Drought Monitor (Kuwayama 2018). Along D0 to D4, the severity of drought increases, which represents abnormally dry, drought moderate, drought sever, drought extreme, drought exceptional, respectively.
-#'
-
-# /*=================================================*/
-#' # Preparation
-# /*=================================================*/
-
-#+ load packages
-
-library(tidyverse)
-library(ggplot2)
-library(sf)
-library(raster)
-library(rnassqs)
-library(data.table)
-library(lfe)
-library(stargazer)
-library(here)
-library(keyring)
-library(tidyUSDA)
-library(tigris)
-library(tidycensus)
-library(exactextractr)
-library(parallel)
-
-# /*=================================================*/
-#' # Identify counties the overlaps with the HPA boundary
-# /*=================================================*/
-#' in: "Data/RawData/hp_bound2010.shp"
[…]

### 2023-11-02 Codes/1_analysis.R

@@ -1,4 +1,4 @@
-## ---------------------------------------------------------------
+## ----------------------------------------
 #--- read data ---#
 final_data <-
   here("Data/data-processed/final_data.rds") %>%
@@ -28,7 +28,7 @@ reg_data <-
   setnames("crop_type", "crop")
 
 
-## ---------------------------------------------------------------
+## ----------------------------------------
 main_analysis <-
   reg_data %>%
   rowwise() %>%
@@ -110,7 +110,7 @@ main_analysis <-
 saveRDS(main_analysis, "Results/main_analysis.rds")
 
 
-## ---------------------------------------------------------------
+## ----------------------------------------
 set.seed(383954)
 
 all_results <-

### 2023-11-02 Codes/2_figures_tables.R

@@ -1,59 +1,11 @@
-## --------------------------------------------------------
+## ---------------------
 all_results <- readRDS("Results/all_results.rds")
 
 data_corn <- all_results[crop == "corn", data][[1]]
 data_soy <- all_results[crop == "soy", data][[1]]
 
 
-## --------------------------------------------------------
-theme_fig <-
-  theme_bw() +
-  theme(
-    axis.title.x =
-      element_text(
-        size = 7, angle = 0, hjust = .5, vjust = -0.3, family = "Helvetica"
-      ),
-    axis.title.y =
-      element_text(
-        size = 7, angle = 90, hjust = .5, vjust = .9, family = "Helvetica"
-      ),
-    axis.text.x =
-      element_text(
-        size = 7, angle = 0, hjust = .5, vjust = 1.5, family = "Helvetica"
-      ),
-    axis.text.y =
-      element_text(
-        size = 7, angle = 0, hjust = 1, vjust = 0, family = "Helvetica"
-      ),
-    axis.ticks =
-      element_line(
-        linewidth = 0.3, linetype = "solid"
-      ),
-    axis.ticks.length = unit(.15, "cm"),
-    #--- legend ---#
-    legend.text =
-      element_text(
-        size = 7, angle = 0, hjust = 0, vjust = 0.5, family = "Helvetica"
-      ),
-    legend.title =
-      element_text(
-        size = 7, angle = 0, hjust = 0, vjust = 0, family = "Helvetica"
-      ),
-    legend.key.size = unit(0.5, "cm"),
-    #--- strip (for faceting) ---#
-    strip.text = element_text(size = 7, family = "Helvetica"),
-    #--- plot title ---#
-    plot.title = element_text(family = "Helvetica", face = "bold", size = 7),
-    #--- margin ---#
-    # plot.margin = margin(0, 0, 0, 0, "cm"),
-    #--- panel ---#
-    panel.grid.minor = element_blank(),
-    panel.background = element_blank(),
-    panel.border = element_rect(fill = NA)
-  )
-
-
-## --------------------------------------------------------
+## ---------------------
 wd_data <-
   rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
   .[, .(wd = mean(balance)), by = .(year, crop)] %>%
@@ -189,7 +141,7 @@ ggsave(
 )
 
 
[…]

### 2023-11-02 Codes/functions.R

@@ -170,64 +170,51 @@ get_grid_MET <- function(var_name, year) {
 #++++++++++++++++++++++++++++++++++++
 
 get_et0_county_year <- function(county, year) {
-  file_name <-
-    paste0(
-      "Data/data-processed/gMET-county/gMET-",
-      county$sc_code, "-", year,
-      ".rds"
-    )
+  #--------------------------
+  # collect weather variables in a single data.table
+  #--------------------------
+  temp <-
+    lapply(
+      var_ls,
+      function(x) proess_gm_county(x, year, county)
+    ) %>%
+    rbindlist() %>%
+    dcast(
+      rowid + coverage_fraction + cell + x + y + date ~ var,
+      value.var = "value"
+    ) %>%
+    .[, Jday := as.numeric(date - min(date) + 1)] %>%
+    .[month(date) >= 4, ] %>%
+    .[month(date) <= 9, ] %>%
+    # === day length in second ===#
+    .[, dayl := geosphere::daylength(y, date) * 3600] %>%
+    setnames(
+      c("tmmx", "tmmn"),
+      c("tmax", "tmin"),
+    ) %>%
+    # === Kelvin to Celsius ===#
+    .[, `:=`(
+      tmin = tmin - 273.15,
+      tmax = tmax - 273.15
+    )]
 
-  if (!file.exists(file_name)) {
-    #--------------------------
-    # collect weather variables in a single data.table
-    #--------------------------
-    temp <-
-      lapply(
-        var_ls,
-        function(x) proess_gm_county(x, year, county)
-      ) %>%
-      rbindlist() %>%
-      dcast(
-        rowid + coverage_fraction + cell + x + y + date ~ var,
-        value.var = "value"
-      ) %>%
-      .[, Jday := as.numeric(date - min(date) + 1)] %>%
-      .[month(date) >= 4, ] %>%
-      .[month(date) <= 9, ] %>%
-      # === day length in second ===#
-      .[, dayl := geosphere::daylength(y, date) * 3600] %>%
-      setnames(
-        c("tmmx", "tmmn"),
-        c("tmax", "tmin"),
-      ) %>%
-      # === Kelvin to Celsius ===#
-      .[, `:=`(
-        tmin = tmin - 273.15,
-        tmax = tmax - 273.15
-      )]
+  #--------------------------
[…]

### 2023-11-02 Codes/2_figures_tables.rmd
modify readme

@@
 
-saveRDS(hpa_simplified, here("Data/data-processed/hpa_simplified.shp"))
+saveRDS(hpa_simplified, here("Data/data-processed/hpa_simplified.rds"))
 ```
@@
 hpa_simplified <-
-  here("Data/data-processed/hpa_simplified.shp") %>%
+  here("Data/data-processed/hpa_simplified.rds") %>%
   readRDS()
@@
 #--- all relevant counties ---#
-all_counties <- readRDS("Shared/Data/counties.rds")
+all_counties <- readRDS("Data/data-processed/counties.rds")
 
 #--- all relevant states ---#
-all_states <- readRDS("Shared/Data/states.rds")
+all_states <- readRDS("Data/data-processed/states.rds")
 

### 2023-11-02 Codes/2_figures_tables.R

@@ -175,23 +175,23 @@ hpa_simplified <-
   st_read() %>%
   st_simplify(dTolerance = 1000)
 
-saveRDS(hpa_simplified, here("Data/data-processed/hpa_simplified.shp"))
+saveRDS(hpa_simplified, here("Data/data-processed/hpa_simplified.rds"))
 
 
 ## ---------------------
 #--- hpa ---#
 hpa_simplified <-
-  here("Data/data-processed/hpa_simplified.shp") %>%
+  here("Data/data-processed/hpa_simplified.rds") %>%
   readRDS()
 
 #--- fips code ---#
 data(fips_codes, package = "tidycensus")
 
 #--- all relevant counties ---#
-all_counties <- readRDS("Shared/Data/counties.rds")
+all_counties <- readRDS("Data/data-processed/counties.rds")
 
 #--- all relevant states ---#
-all_states <- readRDS("Shared/Data/states.rds")
+all_states <- readRDS("Data/data-processed/states.rds")
 
 states_of_interest <-
   dplyr::filter(all_states, state %in% data_corn[, unique(state)])

### 2023-11-03 Data/metadata.Rmd
minimum set of datasets added

---
title: Metadata
author: Taro Mieno
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    number_sections: true
---

# Data/data-raw

+ **hp_bound2010.shp**: HPA aquifer boundary with no attribute (shape file)
+ **metadata_elevationdata.nc**: ~ 4km elevation data from the gridMET website (netcdf file)
+ **nass_data_raw.rds** (`data.table`): downloaded from USDA NASS using the R `tidyusda` package (generated in **Codes/0_2_quickstat.Rmd**)
  + `year`: year
  + `county_code`: county fips code
  + `state_code`: state fips code
  + `short_desc`: description of what `Value` variable represents
  + `Value`: value

## data-raw -> girdMET

Files in this folder stores the raw gridMET data saved as netcdf files. They follow the following naming convention:

"weather_variable"_"year.nc"

Single files records daily values for the corresponding variable.

Here are the list of the variables:

+ "pr": precipitation
+ "rmin": min relative humidity (%)
+ "rmax": max relative humidity (%)
+ "srad": radiation (W/M^2)
+ "tmmn": tmin (Kelvin)
+ "tmmx": tmax (Kelvin)
+ "vs": wind speed,
+ "pet": evapotranspiration

See **Codes/04_get_weather.Rmd** for the codes to download them.

## data-raw -> saturated_thickness

Files in this folder stores the aquifer thickness data saved as geotif files. They follow the following naming convention:

"SatThick"_"year.tif"

# Data/data-processed

+ **base_counties.rds** (`data.table`): boundary and attributes of counties (generated in **0_1_hpa_counties.Rmd**)
  - `state`: state (abbreviated)
  - `state_name`: state (full)
  - `sc_code`: State and County FIPS codes combined into one
  - `ir_area_ratio`: the ratio of the county area that is overlapping with the HPA
  - `geometry`: simple feature column reprsenting county boundary
  - `in_hpa`: whether intersecting with the HPA or not (1 if intersected, 0 otherwise)
+ **gridMET.rds** (`data.table`): daily weather data from gridMET (generated in **0_4_get_weather.Rmd**)
  + `date`: date 
  + `tmin`: minimum temperature
  + `tmax`: maximum temperature
  + `prcp`: precipitation
  + `et0`: reference evapotranspiration
  + `sc_code`: State and County FIPS codes combined into one
+ **nass_data.rds** (`data.table`): **nass_data_raw.rds** reshaped to a long format (generated in **0_2_quickstat.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `crop`: crop type (corn or soybean)
  + `ir`: irrigated or rainfed ("ir" if irrigated, "nir" if not.)
  + `acres`: harvested acres
  + `yield`: crop yield (bu/acre)
+ **sat_all.rds** (`data.table`): aquifer (saturated) thickness  (generated in **0_3_get_sat.Rmd**)
  + `sat`: aquifer (saturated) thickness level
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
+ **ssurgo_soil.rds** (`data.table`): SSURGO soil data combined into one (generated in **0_5_ssurgo.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `sandtotal_r`: sand percentage
  + `silttotal_r`: silt percentage
  + `awc_r`: water holding capacity
+ **summarized_gridMET.rds** (`data.table`): gridMET data summerized by production season (generated in **0_4_get_weather.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `balance`: water balance (deficit) = `et0` - `prcp`
  + `et0`: total reference evapotranspiration (mm) during the production season defined as May through September
  + `prcp`: total precipitation (mm) during the production season defined as May through September
+ **final_data.rds** (`data.table`): dataset for analysis (generated in **0_6_merge_datasets.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `sat`: aquifer (saturated) thickness (m)
  + `balance`: water balance (defict) = `et0` - `prcp`
  + `et0`: total reference evapotranspiration (mm) during the production season defined as May through September
  + `prcp`: total precipitation (mm) during the production season defined as May through September
  + `state`: abbreviated state name
  + `state_name`: full state name
  + `ir_area_ratio`: the ratio of the county area that is overlapping with the HPA
  + `geometry`: simple feature column reprsenting county boundary
  + `in_hpa`: whether intersecting with the HPA or not (1 if intersected, 0 otherwise)
  + `ir`: irrigated or rainfed ("ir" if irrigated, "nir" if not.)
  + `acres`: harvested acres
  + `yield`: crop yield (ton/ha)
  + `sandtotal_r`: sand percentage
  + `silttotal_r`: silt percentage
  + `awc_r`: water holding capacity
  + `sc_dry`: `sc_code`_`ir` (concatenation of `sc_code` and `ir`)
  + `state_year`: `state`_`year` (concatenation of `state` and `year`)
+ **hpa_simplified.rds** (`sf`): boundary of HPA simplified from **hp_bound2010.shp**, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)
+ **states.rds** (`sf`): state boundary, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)
+ **counties.rds** (`sf): county boundary, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)

```{r, eval = FALSE, echo = FALSE}
readRDS("Data/data-processed/final_data.rds")
```

## Data/data-processed/SSURGO

This folder houses a series of SSURGO soil data files. Each file pertains to a single county and contains the same variables as found in **ssurgo_soil.rds**:

+ `sc_code`: State and County FIPS codes combined into one
+ `sandtotal_r`: sand percentage
+ `silttotal_r`: silt percentage
+ `awc_r`: water holding capacity

Indeed, `ssurgo_soil.rds` is an aggregation of the files contained in this folder.

Files for individual counties are saved separately due to the instability of the SSURGO website.

### 2023-11-03 Writing/nw_manuscript.Rnw
modify readme
minimum set of datasets added

@@
 Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
+A minimum set of data files required to reproduce the results and figures presented in this article can be found on the Figshare website (DOI: 10.6084/m9.figshare.24492196).
+Due to file size constraints on Figshare, not every data file is hosted there.
+However, a complete set of data files can be accessed via the provided Dropbox link: \url{https://www.dropbox.com/scl/fo/bghhwlidmi7wx1ok0az5n/h?rlkey=tgbix1hp7g9np9etlo1z3biyr&dl=0}.
 
@@
 The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}.
+Additionally, instructions for reproducing all the results and figures presented in this article can be found at the bottom of the repository page in the README section.
 
@@
 sat_ref <-
-  readRDS(here::here("Shared/Results/all_results.rds")) %>%
+  readRDS(here::here("Results/all_results.rds")) %>%
   .[1, sat_seq] %>%

### 2023-11-03 Codes/2_figures_tables.rmd
minor error fix

@@
 ```{r, eval = FALSE}
+#--- fips code ---#
+data(fips_codes, package = "tidycensus")
+
 all_counties <-

### 2023-11-03 Writing/nw_manuscript.Rnw
update readme

@@
 This research was also supported by Innovate UK award no.
-10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.'' The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
+10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.''
+
+The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
 
@@
 
-\subsection*{Author Contributions}
+\section*{Author Contributions}
 
-\subsection*{Competing Interests Statement}
+T.M.
+conceived the paper, developed the computer programs for data collection, processing, analysis, and summarization, and wrote the paper.; T.F.
+conceived and wrote the paper; S.K.
+wrote the computer programs for data collection and processing; and N.B.
+conceived and helped improve the framing of the paper.
+
+\section*{Competing Interests Statement}
+
+The authors declare no competing interests.
 

### 2023-11-04 Codes/3_split_manuscript.Rmd
a code to split the all-in-one manuscript to the main and SI parts.

```{r}
master <- readLines(here::here("Writing/nw_manuscript.Rnw"))
```

# Main

```{r}
main_end_line <- which(str_detect(master, "%= Supplementary Information")) - 2

which(str_detect(main, "pathway-reg"))

main <-
  master[1:main_end_line] %>%
  #--- this assumes the section ordering is correct and do not change any more ---#
  gsub("\\\\ref\\{pathway-reg\\}", "A", .) %>%
  gsub("\\\\ref\\{reg-conf\\}", "B", .) %>%
  gsub("\\\\ref\\{test-dif-share\\}", "C", .) %>%
  gsub("\\\\ref\\{spline-basis\\}", "D", .) %>%
  gsub("\\\\ref\\{county-fe\\}", "E", .)

writeLines(main, here::here("Writing/nw_main.Rnw"))
```

# SI

### 2023-11-04 Writing/nw_main.Rnw
a code to split the all-in-one manuscript to the main and SI parts.

--- 2023-11-03 Writing/nw_manuscript.Rnw
@@
 \usepackage{lineno}
-\linenumbers
+% \linenumbers
 \usepackage{float}
@@
 \urlstyle{same} % disable monospaced font for URLs
+\thispagestyle{empty}
 \hypersetup{
@@
 However, it is unlikely that this effects is captured in our analysis.
-Please see Appendix \ref{pathway-reg}.
+For more details on this issue, please refer to section \ref{pathway-reg} of the Supplementary Information.
 
@@
 Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.
+Please see section \ref{reg-conf} of the Supplementary Information for estimated yield response curves with 95 percent confidence interval.
 The histogram of annual water deficit is presented at the top for each crop.
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (Please see section \ref{test-dif-share} of the Supplementary Information for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (For further explanations on how spline-based regression functions, please see section \ref{spline-basis} of the Supplementary Information.).
 $K$ is the total number of spline basis functions.
@@
 Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
-A formal discussion of the trade-offs associated with the choice of fixed effects is provided in Appendix \ref{county-fe}.
+A formal discussion of the trade-offs associated with the choice of fixed effects is provided in section \ref{county-fe} of the Supplementary Information.
 
@@
 \caption*{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
+Please see section \ref{reg-conf} of the Supplementary Information for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}
@@
 
+% \renewcommand{\thesection}{S.I. \arabic{section}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S.\arabic{figure}}
+
 \hypertarget{pathway-reg}{%
-\section{Groundwater management policies as a potential causal pathway}\label{pathway-reg}}
+\section{}\label{pathway-reg}}
 
@@
 
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-
 In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
@@
 %------------------------
-%- Figure B.1
+%- Figure S.1
 %------------------------
@@
 %------------------------
-%- Figure B.2
+%- Figure S.2
 %------------------------
@@
 %------------------------
-%- Figure B.3
+%- Figure S.3
 %------------------------
@@
 
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
-
 <<sat_ref, echo = FALSE>>=
@@
 %------------------------
-%- Figure C.1
+%- Figure S.4
 %------------------------
@@
 
-\hypertarget{county-fe}{%
-\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
-This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
-As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
-However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
-For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
-
-%------------------------
-%- Figure D.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of variations in aquifer thickness after demeaning}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
-
-\caption*{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}
-\label{fig:variation-left}
-
-\end{figure}
-
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
-Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
-We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
-
-%------------------------
-%- Figure D.2
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
-
-\caption*{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
-Gray shaded area represents 95\% confidence interval.}
-\label{fig:state-fe-with-without}
-\end{figure}
-
-\clearpage
-
 \hypertarget{spline-basis}{%
 \section{Spline basis functions and GAM estimation}\label{spline-basis}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{E.\arabic{figure}}
 
@@
 %------------------------
-%- Figure E.1
+%- Figure S.5
 %------------------------
@@
 
+%------------------------
+%- Figure S.6
+%------------------------
+
 \begin{figure}[H]
@@
 \end{figure}
+\clearpage
+
+\hypertarget{county-fe}{%
+\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
+
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
+For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
+
+%------------------------
+%- Figure S.7
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of variations in aquifer thickness after demeaning}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
+
+\caption*{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}
+\label{fig:variation-left}
+
+\end{figure}
+
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
+Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
+
+%------------------------
+%- Figure S.8
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
+
+\caption*{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
+Gray shaded area represents 95\% confidence interval.}
+\label{fig:state-fe-with-without}
+\end{figure}
 

### 2023-11-07 Codes/3_gen_manuscript.Rmd
manuscript split into two: main and SI

```{r}
master <- readLines(here::here("Writing/nw_manuscript.Rnw"))
```

# Main

```{r}
main_end_line <- which(str_detect(master, "%= Supplementary Information")) - 2

end_doc <- which(str_detect(master, "end\\{document\\}"))

main <-
  master[c(1:main_end_line, end_doc)] %>%
  #--- this assumes the section ordering is correct and do not change any more ---#
  gsub("\\\\ref\\{pathway-reg\\}", "A", .) %>%
  gsub("\\\\ref\\{reg-conf\\}", "B", .) %>%
  gsub("\\\\ref\\{test-dif-share\\}", "C", .) %>%
  gsub("\\\\ref\\{spline-basis\\}", "D", .) %>%
  gsub("\\\\ref\\{county-fe\\}", "E", .) %>%
  gsub("\\\\ref\\{fig:irrigated-yield-ind-corn\\}", "S.1", .) %>%
  gsub("\\\\ref\\{fig:irrigated-yield-ind-soy\\}", "S.2", .)

writeLines(main, here::here("Writing/nw_main.Rnw"))
```

# SI

```{r}
main_start_line <- which(str_detect(master, "maketitle"))

main_end_line <- which(str_detect(master, "appendix-appendix"))

bib_line <- which(str_detect(master, "DRA.bib"))

si <-
  master[-c(main_start_line:main_end_line)] %>%
  gsub("\\\\ref\\{main\\}", "Main", .)

si_end_doc <- which(str_detect(si, "end\\{document\\}"))

si <- c(si[-si_end_doc], master[bib_line], si[si_end_doc])

writeLines(si, here::here("Writing/nw_si.Rnw"))
```

### 2023-11-07 Writing/nw_main.Rnw
manuscript split into two: main and SI

--- 2023-11-03 Writing/nw_manuscript.Rnw
@@
 \usepackage{lmodern}
-\usepackage[labelfont=bf, textfont=footnotesize]{caption} % \caption*{} will create legend (explanatory text)
+\usepackage[labelfont=bf]{caption} 
 \usepackage{iftex}
@@
 \author[4]{Nicholas Brozovi\'{c}}
-\affil[1]{Department of Agricultural Economics, University of Nebraska Lincoln}
+\affil[1]{Department of Agricultural Economics, University of Nebraska-Lincoln}
 \affil[2]{School of Mechanical, Aerospace and Civil Engineering, University of Manchester}
@@
 \affil[4]{Daugherty Water for Food Global Institute at the University of Nebraska}
-\affil[4]{Department of Agricultural Economics, University of Nebraska Lincoln}
+\affil[4]{Department of Agricultural Economics, University of Nebraska-Lincoln}
 
@@
 However, it is unlikely that this effects is captured in our analysis.
-Please see Appendix \ref{pathway-reg}.
+For more details on this issue, please refer to section A of the Supplementary Information.
 
@@
 Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.
+Please see section B of the Supplementary Information for estimated yield response curves with 95 percent confidence interval.
 The histogram of annual water deficit is presented at the top for each crop.
 In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean.
-As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases (see Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} for the 95\% confidence intervals).
+As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases (see Figures S.1 and S.2 for the 95\% confidence intervals).
 This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling, li2020quantifying}.
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (Please see section C of the Supplementary Information for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (For further explanations on how spline-based regression functions, please see section D of the Supplementary Information.).
 $K$ is the total number of spline basis functions.
@@
 Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
-A formal discussion of the trade-offs associated with the choice of fixed effects is provided in Appendix \ref{county-fe}.
+A formal discussion of the trade-offs associated with the choice of fixed effects is provided in section E of the Supplementary Information.
 
@@
 
-Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
+Funding for the research in this article was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
 This research was also supported by Innovate UK award no.
@@
 
-The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
+The funders had no role in study design, data collection and analysis, decision to publish or preparation of the article.
 
@@
 
-\caption*{Note: The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+\caption*{\footnotesize{Note: The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
 The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}
+Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}}
 \label{fig:yield-response}
@@
 
-\caption*{Note: The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
-Shaded area represents 95\% confidence interval.}
+\caption*{\footnotesize{Note: The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
+Shaded area represents 95\% confidence interval.}}
 \label{fig:yield-dif}
@@
 
-\label{fig:ir-share}
-\caption*{Note: The solid black line represents the estimated share of irrigated acres.
+\caption*{\footnotesize{Note: The solid black line represents the estimated share of irrigated acres.
 Shaded area represents 95\% confidence interval.
 The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}
+Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}}
+\label{fig:ir-share}
 \end{figure}
@@
 % {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
-\caption*{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}
+\caption*{\footnotesize{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+Please see section B of the Supplementary Information for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}}
 \label{fig:tot-impact}
@@
 
-\caption*{Note: The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}
+\caption*{\footnotesize{Note: The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}}
 \label{fig:dif-tot-impact}
@@
 
-\caption*{Note: The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}\label{fig:deficit-yield-hist}
+\caption*{\footnotesize{Note: The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}}
+\label{fig:deficit-yield-hist}
 \end{figure}
@@
 
-\caption*{Note: The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
-Counties highlighted in gray are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system}\label{fig:sat-map}
+\caption*{\footnotesize{Note: The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+Counties highlighted in gray are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system.}}
+\label{fig:sat-map}
 \end{figure}
@@
 
-%==============================================
-%= Supplementary Information
-%==============================================
-\clearpage
-
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\hypertarget{pathway-reg}{%
-\section{Groundwater management policies as a potential causal pathway}\label{pathway-reg}}
-
-Our analysis evaluates the links between aquifer thickness and drought resilience, proposing a causal link between declining aquifer thickness and irrigated production during droughts.
-In the first section of our manuscript, we propose that this connection is driven by impacts of aquifer thickness on well yields and, hence, farmers ability to use groundwater to fully buffer crops against water deficits during drought.
-
-It is important to assess the extent to which this causal relationship may be affected or confounded by other factors.
-One key alternative hypothesis is that regulations limiting groundwater withdrawals could have been introduced as a response to declining aquifer thickness, with resulting reductions in irrigated areas and irrigated crop yields driven by regulatory rather than hydrologic (i.e.
-well yield) constraints.
-Below, we discuss the reasons that regulatory constraints are highly unlikely to be the cause of the relationships between aquifer thickness and drought impacts observed in our analysis for the HPA.
-
-First, it is important the majority of groundwater withdrawal regulations that do exist in the HPA limit pumping to a specified volume over a multiyear period.
-It is also common for farmers to have been allowed to bank and carryover significant volumes of water deficits from historic periods when allocation levels were higher \citep{rimvsaite2023groundwater}, further increasing their regulatory quotas above defined caps.
-Together, these factors mean that farmers in parts of the HPA where groundwater withdrawals are regulated or restricted are still able to increase water withdrawals substantially in drought years when demand is higher.
-Prior research has shown that as a result farmers water use decisions in HPA are not constrained by such regulations even in extreme drought years such as 2012 \citep{foster2019assessing}.
-
-Where more restrictive groundwater withdrawal limits have been introduced, for example in some Groundwater Management Districts in Kansas \citep{marston2022, rimvsaite2023groundwater}, these have universally been implemented in the last 5-10 years and therefore have no or only minimal overlap with our study period which extends up to only 2016.
-The earliest of these groundwater managements in Kansas interventions was the Sheridan 6 Local Enhanced Management Area (LEMA) in Northwest Kansas, which introduced restrictions to farmer groundwater pumping of 11 inches per year on average per 5-year period starting in 2013.
-While this overlaps with the final 4 years of our study period, the Sheridan 6 LEMA area covers only parts of two counties in Kansas and hence represents a very small proportion of our overall study sample.
-Additionally, the years 2013-2016 were all characterized by average or above average growing season rainfall (i.e.
-not drought years).
-Furthermore, evidence shows that the effect of restrictions introduced by the LEMA has primarily been a reduction in per-area water use, with irrigated crop yields largely unaffected and only small reductions in irrigated areas compared to pre-policy conditions \citep{drysdale2018adaptation, deines2021combining, whittemore2023108408}.
-These policies therefore would not be expected to confound our results on the impacts of aquifer thickness.
-
-Finally, we also highlight that areas of the southern HPA, in particular in northern Texas, where the lowest aquifer thickness values are observed have few if any regulatory restrictions on groundwater extraction.
-This is due to laws that typically provide farmers with absolute rights to use groundwater under their land, coupled with low levels of metering.
-Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigating farmers in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
-As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
-This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
-Collectively, they suggest that these restrictions significantly constrain farmers' ability to meet crop water demands during droughts.
-
-\hypertarget{reg-conf}{%
-\section{Uncertainties in estimated impacts of water deficit on irrigated and average yield}\label{reg-conf}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-
-In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
-This section provides visualization of yield response with the 95 \(\%\) confidence interval to convey the degree of uncertainty in our estimation.
-Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} present the impact of water deficit on crop yield by aquifer thickness category with the 95\(\%\) confidence interval for irrigated corn and soybean, respectively.
-Figure \ref{fig:avg-yield-ind} presents the impact of water deficit on average crop yield (irrigated and rainfed combined) by aquifer thickness category for corn and soybean with the 95\(\%\) confidence interval.
-
-%------------------------
-%- Figure B.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}}
-
-\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of corn in US High Plains at different values of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
-\label{fig:irrigated-yield-ind-corn}
-\end{figure}
-
-%------------------------
-%- Figure B.2
-%------------------------
-
-\begin{figure}[H]
-
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}}
-
-\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of soybean in US High Plains at different values of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
-\label{fig:irrigated-yield-ind-soy}
-\end{figure}
-
-%------------------------
-%- Figure B.3
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Average productivity of corn and soybean}
-
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}}
-
-\caption*{Note: The black solid lines represent average yield of corn (top panel) and soybean (bottom panel) in the US High Plains for different levels of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
-\label{fig:avg-yield-ind}
-\end{figure}
-
-\clearpage
-
-\hypertarget{test-dif-share}{%
-\section{Comparison of the impacts of different aquifer thickness levels on irrigated area share}\label{test-dif-share}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
-
-<<sat_ref, echo = FALSE>>=
-sat_ref <-
-  readRDS(here::here("Results/all_results.rds")) %>%
-  .[1, sat_seq] %>%
-  .[[1]] %>%
-  .[length(.)]
-@
-
-Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of \Sexpr{sat_ref} m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
-Gray shaded areas represented the 95\% confidence intervals for the estimated relationships.
-Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to \Sexpr{sat_ref} m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
-
-%------------------------
-%- Figure C.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Decline in the share of acres that are irrigated}
-
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}}
-
-\caption*{Note: The solid black lines represent the share of acres that are irrigated as a function of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
-The gray shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
-\end{figure}
-
-\clearpage
-
-\hypertarget{county-fe}{%
-\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
-This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
-As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
-However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
-For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
-
-%------------------------
-%- Figure D.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of variations in aquifer thickness after demeaning}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
-
-\caption*{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}
-\label{fig:variation-left}
-
-\end{figure}
-
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
-Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
-We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
-
-%------------------------
-%- Figure D.2
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
-
-\caption*{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
-Gray shaded area represents 95\% confidence interval.}
-\label{fig:state-fe-with-without}
-\end{figure}
-
-\clearpage
-
-\hypertarget{spline-basis}{%
-\section{Spline basis functions and GAM estimation}\label{spline-basis}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{E.\arabic{figure}}
-
-We used generalized additive models (GAM) for our regression analysis.
-In GAM, the impact of a variable on the dependent variable can be expressed as a series of spline basis functions and their coefficients.
-Suppose you are interested in modeling the following model.
-
-\begin{equation}
-y = f(x) + \mu
-\end{equation}
-
-Where \(f(x)\) represents the impact of variable \(x\) on \(y\).
-The shape of \(f(x)\) is not known to the researcher.
-GAM attempts to represent \(f(x)\) with a series of spline basis functions and their coefficients like below.
-
-\begin{equation}
-y = \sum_{k=1}^K \beta_k \phi_k(x) + \mu
-\end{equation}
-
-Where \(\phi_k(\cdot)\) is the \(k\)th spline basis function and \(\beta_k\) is its coefficient.
-Figure \ref{fig:sp-basis} provides a visual illustration of a series of thin-plate spline basis functions.
-
-%------------------------
-%- Figure E.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Example of a series of thin-plate spline basis functions}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}}
-
-\label{fig:sp-basis}
-\end{figure}
-
-Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the impact of \(x\) on \(y\).
-Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting representation of \(f(x)\).
-In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\).
-In the second case, \(\{\beta_1, \dots, \beta_5\} = \{20, -20, -40, 40, 60\}\).
-Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best.
-Unlike non-parametric regression, a GAM is a linear (in-parameter) model and estimates $\beta_k$ to best fit the observation data.
-
-\begin{figure}[H]
-
-\caption{Example of a function represented by different coefficient values}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}
-
-}
-
-\label{fig:ex-fx}
-\end{figure}
-
 \end{document}

### 2023-11-07 Writing/nw_manuscript.Rnw
manuscript split into two: main and SI

@@
 % {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
-\caption*{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-See Appendix \ref{reg-conf} for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}
+\caption*{\footnotesize{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+Please see section \ref{reg-conf} of the Supplementary Information for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}}
 \label{fig:tot-impact}
@@
 \clearpage
+% appendix-appendix
 
-\hypertarget{appendix-appendix}{%
-\appendix}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S.\arabic{figure}}
+\renewcommand{\thesection}{\Alph{section}}
 
@@
 Our analysis evaluates the links between aquifer thickness and drought resilience, proposing a causal link between declining aquifer thickness and irrigated production during droughts.
-In the first section of our manuscript, we propose that this connection is driven by impacts of aquifer thickness on well yields and, hence, farmers ability to use groundwater to fully buffer crops against water deficits during drought.
+In the first section of our article, we propose that this connection is driven by impacts of aquifer thickness on well yields and, hence, farmers ability to use groundwater to fully buffer crops against water deficits during drought.
 
@@
 Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigating farmers in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
-As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
+As discussed in section \ref{main} of our article, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
 This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
@@
 \section{Uncertainties in estimated impacts of water deficit on irrigated and average yield}\label{reg-conf}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
 
@@
 
-\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of corn in US High Plains at different values of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
+\caption*{\footnotesize{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of corn in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
 \label{fig:irrigated-yield-ind-corn}
@@
 
-\caption*{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of soybean in US High Plains at different values of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
+\caption*{\footnotesize{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of soybean in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
 \label{fig:irrigated-yield-ind-soy}
@@
 
-\caption*{Note: The black solid lines represent average yield of corn (top panel) and soybean (bottom panel) in the US High Plains for different levels of water deficit and aquifer thickness.
-The gray shaded area represents the 95\% confidence interval.}
+\caption*{\footnotesize{Note: The black solid lines represent average yield of corn (top panel) and soybean (bottom panel) in the US High Plains for different levels of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
 \label{fig:avg-yield-ind}
@@
 \section{Comparison of the impacts of different aquifer thickness levels on irrigated area share}\label{test-dif-share}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
 
@@
 
-\caption*{Note: The solid black lines represent the share of acres that are irrigated as a function of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
-The gray shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
-\end{figure}
-
-\clearpage
-
-\hypertarget{county-fe}{%
-\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
-This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
-As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
-However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
-For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
-
-%------------------------
-%- Figure D.1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of variations in aquifer thickness after demeaning}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
-
-\caption*{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}
-\label{fig:variation-left}
-
-\end{figure}
-
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
-Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
-We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
-
-%------------------------
-%- Figure D.2
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
-
-\caption*{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
-Gray shaded area represents 95\% confidence interval.}
-\label{fig:state-fe-with-without}
+\caption*{\footnotesize{Note: The solid black lines represent the share of acres that are irrigated as a function of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
+The gray shaded areas represent the 95\% confidence interval for this estimated relationship.}}
+\label{fig:test-sat-share}
 \end{figure}
@@
 \section{Spline basis functions and GAM estimation}\label{spline-basis}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{E.\arabic{figure}}
 
@@
 
+\caption*{\footnotesize{Note: This figure presents a series of thin-plate spline basis functions evaluated at a range of $x$ values.}}
 \label{fig:sp-basis}
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}}
+\caption*{\footnotesize{Note: This figure presents two example functions derived by applying different coefficients on the thin-plate spline basis functions.}}
+\label{fig:ex-fx}
+\end{figure}
+\clearpage
 
-}
+\hypertarget{county-fe}{%
+\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
 
-\label{fig:ex-fx}
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
+For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
+
+%------------------------
+%- Figure S.7
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of variations in aquifer thickness after demeaning}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
+
+\caption*{\footnotesize{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}}
+\label{fig:variation-left}
+\end{figure}
+
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
+Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
+
+%------------------------
+%- Figure S.8
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
+
+\caption*{\footnotesize{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
+Gray shaded area represents 95\% confidence interval.}}
+\label{fig:state-fe-with-without}
 \end{figure}

### 2023-11-07 Writing/nw_si.Rnw
manuscript split into two: main and SI

--- 2023-11-07 Writing/nw_manuscript.Rnw
@@
 \begin{document}
-\maketitle
-\begin{abstract}
-Aquifer depletion poses a major threat to the ability of farmers, food supply chains, and rural economies globally to use groundwater as a means of adapting to climate variability and change.
-Empirical research has demonstrated the large differences in drought risk exposure that exist between rainfed and irrigated croplands, but previous work commonly assumes water supply for the latter is unconstrained.
-In this paper, we evaluate how aquifer depletion affects the resilience of irrigated crop production to drought risk using over 30 years of data on historical corn and soybean yields, production areas, and aquifer conditions for the High Plains region in the United States.
-We show that aquifer depletion reduces the ability of farmers to sustain irrigated crop yields and production areas in years and locations with large growing season water deficits.
-Our findings demonstrate that drought-related production losses on irrigated croplands increase non-linearly with aquifer depletion, highlighting the need for proactive aquifer conservation interventions to support adaptation and resilience to future increases in rainfall variability under climate change.
-\end{abstract}
-
-\hypertarget{main}{%
-\section{Main}\label{main}}
-
-Groundwater is an essential input for agricultural production, providing a critical buffer against limited or variable surface water supplies for farmers in many parts of the world \citep{scanlon2023global}.
-Globally, agriculture's dependence on groundwater for irrigation is expected to increase in the future because of higher crop water requirements, more erratic rainfall, and more frequent and extreme drought events caused by climate change \citep{zhou2010impact, wada2013multimodel, wada2014sustainability, kreins2015quantification, florke2018water}.
-However, many of the world's most important aquifer systems have experienced large reductions in storage over recent decades \citep{wada2010global, famiglietti2011satellites, scanlon2012groundwater, konikow2015long, bierkens2019non}.
-If not addressed, these reductions will negatively affect the ability of farmers, food supply chains, and rural economies to use groundwater as a means of adapting to climate change.
-
-Despite widespread alarm about the risks that aquifer depletion poses to crop productivity and food security, empirical evidence is limited about how reductions in groundwater availability will alter farmers' capacity to adapt successfully to drought and rainfall variability.
-Most empirical studies that have assessed impacts of drought on agricultural productivity have focused primarily on rainfed production areas due to the greater exposure of rainfed agriculture to drought-related shocks \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact}.
-Existing studies also demonstrate the importance of existing irrigated areas or future irrigation intensification and expansion to mitigate negative impacts of drought \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes}.
-In contrast, most research assessing impacts of aquifer depletion on resilience of irrigated farmland to drought has relied on simulation modeling \citep{foster2015well, cotterman2018groundwater, kahil2015modeling, yoon2021coupled, rad2020mod}, with more limited work that attempts to empirically evaluate change in drought-related production losses and risks on irrigated lands as a function of changing aquifer storage \citep{jain2021groundwater, suter2021depletion}.
-
-Two major pathways exist by which aquifer depletion will impact farmers' ability to access and use groundwater as a buffer against drought and rainfall variability \citep{foster2015analysis}.
-First, depletion increases the energy requirements and costs of pumping water, potentially making it less economically viable for farmers to satisfy crop water needs \citep{mieno2017price, bhattarai2021impact}.
-Second, depletion also reduces the thickness of an aquifer, lowering the aquifer's transmissivity (i.e., the potential rate of water supply to wells).
-This latter effect reduces the rate at which water can be pumped and applied to a field as irrigation \citep{konikow2005groundwater, foster2014modeling, hrozencik2017heterogeneous}.
-If the reduction in well yields is substantial enough - for example as would occur when a large proportion of aquifer thickness is depleted \citep{hecox2002calculation, korus2020depletion} - then a farmer may be unable to extract sufficient water to meet crop water demands, in particular during periods of limited rainfall and in the peak of the growing season when crops are most sensitive to water deficits \citep{foster2015well,rouhi2020downside}.
-The combined effect of declining well yields and increasing pumping costs could force farmers to reduce irrigation depths, resulting in greater risk of drought-related crop stress.
-Alternatively, or in addition, a farmer may be forced to limit irrigated area to ensure crop water demands can be met, leading to reductions in total crop production \citep{foster2014modeling, rad2020effects}.
-Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}.
-Another potential pathway is through groundwater use regulation in response to low or declining aquifer thickness.
-However, it is unlikely that this effects is captured in our analysis.
-For more details on this issue, please refer to section \ref{pathway-reg} of the Supplementary Information.
-
-In this paper, we present empirical evidence about how aquifer depletion affects farmers' ability to effectively and reliably buffer crops against drought  using data on historical crop yields, production areas, and aquifer conditions for the High Plains region in the United States.
-Our analysis extends previous research on the impacts of drought on agricultural production by evaluating how drought-related production losses are influenced by severity of groundwater scarcity, moving beyond prior binary comparisons of rainfed and irrigated production \citep{schlenker2009nonlinear, lobell2014greater, lu2018crop}.
-We also provide new empirical insights about the relative roles of extensive (i.e., irrigated area) and intensive (i.e., per-area irrigation rates) margin water use adjustments in determining drought-related production losses, which previously had only been demonstrated through theoretical modeling \citep{foster2014modeling, foster2017effects, rad2020effects}.
-Our findings provide important evidence of the value of preserving aquifer storage as a buffer against drought events.
-In doing so, our analysis adds weight to the need to improve groundwater resource conservation and sustainability to enable adaptation to climate change in groundwater-dependent agricultural systems globally \citep{jain2021groundwater, scanlon2023global}.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-<<load_results, echo = FALSE>>=
-setwd(here::here("Writing"))
-all_results <- readRDS("../Results/all_results.rds")
-@
-
-\hypertarget{impact-intensive}{%
-\subsection{Sensitivity of per-area crop yields to water deficits}\label{impact-intensive}}
-
-Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B).
-Confidence intervals were not presented to avoid over-crowding the figures.
-Please see section \ref{reg-conf} of the Supplementary Information for estimated yield response curves with 95 percent confidence interval.
-The histogram of annual water deficit is presented at the top for each crop.
-In wetter years with low crop irrigation requirements (i.e., negative values of water deficit), rainfed and irrigated yields are very similar for both corn and soybean.
-As drought severity increases (i.e., as water deficit increases), rainfed yields of both crops decline significantly whereas irrigated yields are increasing or stable as deficit increases (see Figures \ref{fig:irrigated-yield-ind-corn} and \ref{fig:irrigated-yield-ind-soy} for the 95\% confidence intervals).
-This result reflects the positive benefits irrigation provides both as a buffer against precipitation variability, and as a means of raising overall yields through increased evapotranspiration and associated biomass accumulation, reductions of heat-related stressors, and other factors \citep{zhu2022untangling, li2020quantifying}.
-
-We now focus on comparing yield response to water deficit among the three aquifer thickness categories (all irrigated).
-First, for water deficits above around 400mm, yield changes become negative for counties with the lowest levels of aquifer thickness (i.e., 1st quantile in Figure \ref{fig:yield-response}).
-This likely reflects the restrictions that lower aquifer thickness and associated reductions in well yields place on farmers' ability to fully meet crop water requirements during periods of more severe or extended water deficits \citetext{\citealp[\citet{foster2014modeling}]{rad2020effects}; \citealp{hrozencik2017heterogeneous}}.
-In contrast, yield continues to increase gradually as water deficit increases beyond 400 mm for counties with the highest level of aquifer thickness.
-
-<<echo = FALSE>>=
-delta_yield_lowest_highest <- 
-  all_results[crop == "corn", yield_dif_test][[1]] %>%
-  .[comp_q == 3, ] %>%  
-  .[which.min(abs(balance -950)), dif_hat] %>%
-  round(digits = 2)
-@
-
-Figure \ref{fig:yield-dif} shows the difference in yield of the second and third (highest) aquifer quantiles relative to the first (lowest) quantile with the 95\% confidence interval for corn and soybean.
-For corn, in years with water deficit greater than about 700 mm, the highest aquifer thickness group has statistically significantly higher yields than the lowest aquifer thickness group.
-At an extreme water deficit level of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), the difference in yields between the smallest and largest aquifer thickness groups is \Sexpr{delta_yield_lowest_highest} tonne/ha.
-Unlike corn, irrigated soybean yields are not statistically distinguishable among the three aquifer thickness groups.
-This is likely because soybean is not as water demanding a crop as corn, meaning that even counties with lower aquifer thickness can meet the majority of crop water demands as well yield is not a significantly binding constraint on irrigation management decisions.
-
-<<num_obs_yield_reg, echo = FALSE>>=
-nobs_corn <- 
-  all_results[crop == "corn", data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-
-nobs_soy <- all_results[crop == "soy", data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-@
-
-\hypertarget{impact-share}{%
-\subsection{Share of agricultural land under irrigation}\label{impact-share}}
-
-While aquifer thickness has a statistically significant impact on per-area irrigated crop yields in drought conditions (water deficit greater than 700 mm), the magnitude of the yield impact is small when compared to differences between irrigated and rainfed production for the same weather conditions.
-One explanation for this result is that farmers may choose to retain a smaller proportion of land in irrigated production, so that limited groundwater pumping capacity can be used to adequately buffer crops against drought on land that remains under irrigation.
-Reductions in irrigated areas as an aquifer is depleted have been demonstrated theoretically \citep{rad2020effects, foster2014modeling, hrozencik2017heterogeneous, deines2020transitions}, but there has been little empirical research on farmers' irrigated area choices or how these are affected by combinations and magnitudes of weather risks and groundwater scarcity.
-
-<<ir_share, echo = FALSE>>=
-corn_share <- all_results[crop == "corn", share_pred_data][[1]]
-
-share_high <-
-  corn_share[which.min(abs(sat - 130)), ir_share_hat] %>%
-  round(digits = 2)
-
-share_low <-
-  corn_share[which.min(abs(sat - 10)), ir_share_hat] %>%
-  round(digits = 2) 
-@
-
-Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer.
-Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (Please see section \ref{test-dif-share} of the Supplementary Information for formal statistical testings).
-Reductions in the share of production area under irrigation is also evident for soybean.
-
-<<num_obs_share_reg, echo = FALSE>>=
-nobs_corn_share <- 
-  all_results[crop == "corn", share_reg_data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-
-nobs_soy_share <- 
-  all_results[crop == "soy", share_reg_data][[1]] %>% 
-  nrow() %>%
-  format(big.mark=",")
-@
-
-\hypertarget{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}{%
-\subsection{\texorpdfstring{Overall drought-related production losses}{Aquifer depletion increases overall drought-related production losses by up to 25\textbackslash\%}}\label{aquifer-depletion-increases-overall-drought-related-production-losses-by-up-to-25}}
-
-<<res_at_950, echo = FALSE>>=
-res_at_950 <-
-  all_results[, .(crop, avg_yield_pred_data)] %>%
-  unnest(cols = c(avg_yield_pred_data)) %>%
-  data.table() %>%
-  .[crop == "corn", ] %>%
-  .[(abs(balance - 950)) < 5, ]
-@
-
-The overall influence of aquifer depletion on drought-related production losses is a function of both per-area irrigated yield and irrigated area changes highlighted in Sections \ref{impact-intensive} and \ref{impact-share}.
-Combining both responses, Figure \ref{fig:tot-impact} shows the relationship between water deficit and average per-area production (weighted-average of irrigated and rainfed yields, where the weights are production area shares) for corn and soybean at four values of aquifer thickness: 10, 40, 70, and 100 meters.
-While similar trends are observed as in Figure \ref{fig:yield-response}, the difference in production loss caused by high water deficits among the aquifer thickness levels are much more pronounced due to lower estimated irrigated area share for smaller aquifer thickness values.
-For example, for an aquifer thickness of 10 m, reductions in average yields occur for smaller water deficits than are required to trigger reductions in per-area irrigated yields alone.
-Similarly, the overall magnitude of yield reduction is greater for average yields than for per-area irrigated yields.
-This is because low aquifer thickness values have a greater estimated share of crop area under rainfed production, which magnifies average yield declines with increasing water deficits.
-
-Figure \ref{fig:dif-tot-impact} shows the difference in average per-area productivity for 40 m, 70 m, and 100 m aquifer thickness levels relative to an aquifer thickness of 10 m for corn and soybean.
-For corn, aquifer thickness does not have a significant impact on productivity when water deficits are below about 400 mm.
-However, beyond this threshold, higher levels of water deficit result in noticeable and statistically significant reductions to average corn production.
-This reflects the combination of lower irrigated yields under drought conditions, and, more importantly, reduced overall productivity due to smaller share of land under irrigation in areas where aquifer thickness is lower.
-Similar effects are also observed for soybean, although reductions in average production only occur at a higher deficit threshold (greater than around 600 mm) and are smaller in overall magnitude.
-For example, for a seasonal water deficit of 950 mm (average water deficit in 2012, the most extreme drought in our historical record), average per-area corn and soybean productivity are about 25\% and 15\% lower, respectively, when differentiating between areas with the 100 m and 10 m aquifer thicknesses.
-
-Figure \ref{fig:tot-impact} and Figure \ref{fig:dif-tot-impact} also demonstrate that impact of aquifer thickness on average per-area production is highly non-linear at high seasonal water deficits.
-At a water deficit of 950 mm, for example, reducing from 100 m aquifer thickness to 70 m (decline of 30 m) would result in only \Sexpr{round(res_at_950[sat == 100 & crop == "corn", avg_yield] - res_at_950[sat == 70 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
-This negligible difference in average yield from the decline is due to the insensitivity of the share of irrigated acres to a decline in aquifer thickness when the initial aquifer thickness level is high as can be seen in Figure \ref{fig:ir-share}.
-Reducing from 70 m aquifer thickness to 40 m (decline of 30 m) would result in \Sexpr{round(res_at_950[sat == 70 & crop == "corn", avg_yield] - res_at_950[sat == 40 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn, with a further drop from 40 m aquifer thickness to 10 m (decline of 30 m) leading to \Sexpr{round(res_at_950[sat == 40 & crop == "corn", avg_yield] - res_at_950[sat == 10 & crop == "corn", avg_yield], digits = 2)} tonne/ha of production loss for corn.
-Similar but less pronounced effects are also observed for soybean.
-This non-linear impact of aquifer thickness is due to the influence of aquifer thickness on the share of irrigated corn acres (Figure \ref{fig:ir-share}), which implicitly reflects how aquifer thickness alters well yields and, in turn, farmers' ability to effectively and reliably buffer crops against water deficits during the growing season.
-
-\clearpage
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-We demonstrate the important role of aquifer thickness in determining farmers' ability to successfully buffer crops against drought risks in groundwater irrigated agricultural systems such as the High Plains of the United States.
-Previous research focused on impacts of drought on agricultural systems has largely focused either on risks for rainfed production \citep{schlenker2009nonlinear, lobell2014greater, schlenker2010robust, zhou2020connections, borgomeo2020impact} or for irrigated lands where water supply constraints are assumed to be minimal \citep{kuwayama2019estimating, zipper2016drought, zhu2022untangling, zhu2022warming, lu2020mapping, davis2019sensitivity, li2018changes, luan2021combined}.
-Our analysis adds new empirical insights to this literature by quantifying how reductions in aquifer thickness increase the vulnerability of irrigated agricultural production areas to drought events.
-Critically, our findings suggest that commonly adopted binary definitions of agricultural land areas into rainfed and irrigated categories are likely to underestimate exposure of crop production to drought risks, in particular in areas where water supplies for irrigation are insufficient to fully meet crop water needs in all years due to constraints imposed by hydrology, policies, and/or economics.
-
-An important implication of our findings is that depletion of groundwater resources is likely to increase significantly the vulnerability of agricultural production to current and future drought risks.
-Climate change is projected to increase the frequency and severity of extreme drought events in many parts of the world \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}, including in parts of Kansas and Texas \citep{bradford2020robust, cook2022projected, mullens2019quantitative} where the High Plains Aquifer continues to experience prolonged and significant declines in aquifer thickness \citep{scanlon2012groundwater, haacker2016water, cotterman2018groundwater}.
-Similar trends of increasing drought risks coupled with declining aquifer storage have been reported in other major groundwater irrigated farming systems worldwide \citep{wada2010global, doll2014global, famiglietti2014global, feng2018groundwater, bierkens2019non}.
-Our findings suggest that these simultaneous changes in aquifer conditions will amplify impacts of future climate change on crop production, with associated ramifications for resilience of supply chains, commodity markets, rural economies, and food security.
-
-Our findings also provide useful guidance for water managers and policymakers seeking to address unsustainable groundwater use in the US High Plains and globally.
-Consistent with prior theoretical modeling, we demonstrate that the relationship between aquifer thickness and climatic water deficits during the growing season is highly non-linear when accounting for farmers' decisions about irrigated production areas.
-This indicates that sustainable groundwater conservation efforts \citep{macewan2017hydroecon, butler2020charting, elshall2020groundwater} should be targeted in space and time to avoid depletion exceeding critical thresholds which impair farmers' ability to effectively buffer production against drought.
-Analysis such as ours can provide valuable evidence about these critical thresholds and tipping points to guide proactive management of groundwater stocks, and, in doing so, limit economically damaging transitions from irrigated to rainfed production \citep{foster2017effects, deines2020transitions}.
-For example, several groundwater management agencies in the Texas portion of the HPA have in recent years set out targets to ensure 50\% of current aquifer storage and thickness remains in 50 years time (so called 50/50 rule) \citep{closas2018chronicle}.
-The effectiveness of such targets could likely be improved substantially by setting conservation targets based on empirical evidence about the level of aquifer thickness that is required to ensure adequate levels of drought protection, both now and with future climate change.
-
-Several opportunities exist for future research to build upon the methods and analyses presented here.
-First, due to the limited resolution of our observational data, we were only able to estimate the impacts of different aquifer thickness categories on drought risk.
-Availability of sub-county crop yield data, for example from satellite remote sensing and/or in-situ monitoring \citep{edreira2020combining, deines2021million}, would allow models of the continuous effect of aquifer thickness on drought risk to be developed.
-Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk.
-For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness \citep{butler2013interpretation, korus2020depletion}.
-These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias \citep{bound1991extent, hyslop2001bias}, suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk.
-Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquifer thickness.
-Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA.
-This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies.
-
-Second, our analysis considered the effects of aquifer thickness on farmers' decisions to switch from irrigated to rainfed production, but did not assess other potential responses to aquifer depletion such as changing crop types or irrigation management practices.
-Switching to more drought resistant crops or varieties may enable farmers to maintain irrigated production at lower levels of aquifer thickness, and has been shown to be an important response to increasing agricultural water scarcity and abstraction policies in other studies \citep{bhattarai2021impact, deines2019quantifying, manning2017producer}.
-Improvements to irrigation scheduling practices and technologies, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas \citep{deines2019quantifying, glose2022quantifying} and in parts of Texas \citep{mrad2020peak} and which are proposed as part of California's Sustainable Groundwater Management Act (SGMA) \citep{berbel2019droughts, lubell2020sustainable}, could also help to improve the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities.
-Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials \citep{foster2019assessing} and the importance of conservation \citep{marston2022}.
-However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks \citep{grafton2018paradox,perez2021agricultural}.
-Future research should explore potential strategies that farmers could adapt to lessen the combined impacts of drought and aquifer depletion, alongside efforts to curb unsustainable groundwater extraction and climate change.
-
-Finally, we focus on the impacts of aquifer thickness on agricultural resilience to seasonal water deficits.
-However, where water deficits are distributed unevenly during the season, negative impacts of aquifer depletion may be even larger.
-This is because reductions in well yields caused by aquifer depletion will constrain farmers' ability to meet peak crop water demands, increasing yield losses compared with if water deficits had been distributed evenly throughout the season \citep{ortiz2019unpacking}.
-Furthermore, irrigation also has broader benefits for crops such as reducing heat-related stress through cooling of land surface temperatures \citep{adegoke2003impact, bonfils2007empirical, lobell2008effect, zhu2022untangling}.
-Reductions in aquifer thickness therefore may have other negative impacts on irrigated crop productivity, if constraints to water supply due to aquifer depletion are significant enough to limit cooling benefits of irrigation during the growing season.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{data-sets}{%
-\subsection{Data-sets}\label{data-sets}}
-
-Our study focuses on the High Plains Aquifer (HPA) in the Central United States.
-Irrigated agriculture is integral to the landscape and economy of the HPA, with the region accounting for the majority of grain production in the United States valued in the billions of dollars annually \citep{smidt2016complex,fenichel2016measuring}.
-To evaluate the impacts of aquifer depletion on agricultural production overlying the HPA, we rely on data from several sources, which are described below.
-
-County-level annual irrigated and rainfed yield data for corn and soybean for 1985 through 2016 were obtained from the U.S.
-Department of Agriculture's National Agricultural Statistics Service (USDA-NASS).
-USDA-NASS does not report irrigated yields disaggregated by the source of water used for irrigation.
-However, it is known that groundwater is the main source of water for irrigation for counties that overlie the HPA due to the limited availability of surface water resources in this area.
-Therefore, we retain both rainfed and irrigated yield observations for counties where greater than 75\% of the county area overlies the boundary of the HPA.
-For counties that do not meet this condition, only rainfed yield observations are obtained to avoid the possibility that irrigated production may be only partly dependent on use of groundwater from the HPA.
-We also include additional rainfed yield observations for counties that are outside the HPA, but within the main states covering the aquifer (i.e., South Dakota, Wyoming, Colorado, New Mexico, Nebraska, Kansas, and Texas), to enable more accurate estimation of the impact of drought for rainfed production as a point of comparison to our main irrigated crop yield analyses.
-
-Meteorological data needed to estimate spatial and temporal variations in drought conditions across the HPA were obtained from the gridMET dataset \citep{Abatzoglou2013}.
-gridMET is a gridded daily weather dataset with a spatial resolution of \textasciitilde 4km, and has been used in a range of climate impacts studies for agriculture and other sectors in the United States \citep{abatzoglou2016impact, pereira2015crop, crane2018machine, venkatappa2021impacts, zhu2019dissecting}.
-Daily precipitation values were extracted for each gridMET cell overlapping counties in our sample.
-Daily values of maximum and minimum temperature, surface radiation, wind speed, and humidity were also obtained from gridMET in order to calculate daily reference crop evapotranspiration for each grid cell.
-From these data, we computed the seasonal water deficit (\(D_{i,t}\)) for each grid cell by subtracting total precipitation from total reference evapotranspiration for the period May through September, which represents the typical growing season for corn and soybean in the HPA region.
-Seasonal water deficit estimates were subsequently aggregated to the county-level to match our agricultural production data using an area-weighted average based on the proportion of the county covered by each overlapping gridMET cell.
-Any county-year combinations that had an extreme water deficit value (lower than -200 or greater than 1200) were dropped.
-Figure \ref{fig:deficit-yield-hist} presents annual water deficit by year for the corn and soybean data, illustrating major drought events in the region in years such as 2000, 2002, and 2012.
-The importance of irrigation is clearly evident in years with large water deficits, with rainfed yields dipping substantially whereas irrigated yields are, on average, largely unaffected (Figure \ref{fig:deficit-yield-hist}).
-
-Soil characteristics data (sand percentage, silt percentage, and water holding capacity) were obtained from the Soil Survey Geographic
-Dataset (SSURGO).
-Observation units of the SSURGO data are polygons.
-For each county, the SSURGO polygons were overlaid with the county boundary, and the area-weighted average of the soil variables were calculated.
-
-As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
-These data provide annual estimates of the aquifer thickness at 250 m resolution since 1935.
-Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
-Aquifer thickness estimates for each year were aggregated to the county level using an area-weighted average to match the resolution of agricultural production data.
-Aquifer thickness values were only estimated for counties with irrigated yield observations.
-Furthermore, during the aggregation process, any county that had an aquifer thickness value lower than 9 m was removed before aggregation to reflect the fact that aquifer thicknesses lower than this threshold are not considered to be viable for extraction and therefore it is unlikely these areas contribute groundwater for irrigation production \citep{fenichel2016measuring, haacker2016water, deines2020transitions}.
-Figure \ref{fig:sat-map} presents the distribution of aquifer thickness across our study area for counties included in corn and soybean analyses.
-Aquifer thickness levels vary substantially over space and time, which we will exploit in our analysis to quantify how aquifer thickness influences farmers' exposure and responses to drought.
-
-These data were combined to form four sets of data for regression analysis: (1) per-area corn yield ($\Sexpr{nobs_corn}$ observations), (2) per-area soybean yield ($\Sexpr{nobs_soy}$ observations), (3) share of irrigated corn ($\Sexpr{nobs_corn_share}$ observations), and (4) share of irrigated soybean ($\Sexpr{nobs_soy_share}$ observations).
-
-\hypertarget{regression-models}{%
-\subsection{Regression Models}\label{regression-models}}
-
-Aquifer depletion may impact resilience of irrigated crop production to drought through two main channels: (1) a reduction in per-area irrigated yields due to the economic or physical inability to meet crop water requirements; and (2) a reduction in the area of land irrigated, resulting in reduced production due to lower yields obtained through rainfed agriculture.
-
-To assess the first of these impacts, we quantify the impact of seasonal water deficits (\(D\)) during the growing season (May through September) on county-level rainfed and irrigated yields (\(Y\)) using a hierarchical generalized additive model, where irrigated yield response to seasonal water deficits is conditioned on the level of aquifer thickness (\(A\)) in the county in a given year.
-Our regression model is given in Equation \eqref{eq:eqn1}, with the same model formulation used for corn and soybean:
-
-\begin{equation}
-Y_{i,t,ir} = \sum_{j=1}^4 \sigma_j C_j + \sum_{j=1}^4 \sum_{k=1}^K \beta_{j,k}\phi_{j,k}(D_{i,t})\cdot C_j  + \alpha_{i,ir} + \lambda_t + v_{i,t,ir} \label{eq:eqn1}
-\end{equation}
-
-where county and year are indicated by subscripts \(i\) and \(t\), respectively, whether rainfed or irrigated is indicated by $ir$,  \(Y\) is crop yield, and \(D\) is seasonal water deficit.
-\(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise.
-The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0\%-33.3\%)\), \([33.3\%-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness.
-The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (For further explanations on how spline-based regression functions, please see section \ref{spline-basis} of the Supplementary Information.).
-$K$ is the total number of spline basis functions.
-$K = 3$ was chosen for this regression to prevent our estimated response from becoming unrealistically oscillatory while maintaining a good fit based on Bayesian Information Criterion (BIC).
-To control for time-invariant heterogeneity across counties and yearly shocks, county-irrigation-status fixed effects $\alpha_{i, ir}$ and year fixed effects $\lambda_t$ are also included.
-County-irrigation-status fixed effects are used instead of county fixed effects because some counties have both rainfed and irrigated observations, and they need to be distinguished.
-Standard error estimation is clustered by county.
-
-To estimate the additional impact of aquifer thickness on the share of total acres that are irrigated in a county, a fractional logit model under the generalized additive model framework is used in which the impact of aquifer thickness, average water deficit, and their interactions are estimated without the prior requirement to assume any functional forms.
-This allows for flexible representation of the non-linear relationships between drought risks, aquifer conditions, and agricultural land use, including potential thresholds in farmers' irrigated area choices under water supply constraints that have been noted in prior work \citep{foster2015well, foster2015analysis, hrozencik2017heterogeneous}.
-The model is given in Equation \eqref{eq:eqn2} below:
-
-\begin{equation}
-    log(\frac{s_{i,t}}{1-s_{i,t}}) = \beta_0 + \sum_{k=1}^K \eta_{k}\phi_{k}(D_{i}) + \sum_{l=1}^L \beta_{l}\tau_{l}(A_{i,t}) + X_i + \theta_{s,t} + v_{i,t} \label{eq:eqn2}
-\end{equation}
-
-where \(s_{i,t}\) is the share of irrigated production.
-The impact of average water deficit is captured by \(\sum_{k}^K \eta_{k}\phi_{k}(D_{i})\), where \(\phi_{k}(\cdot)\) is the \(k\)th spline basis function and \(\eta_{k}\) is its coefficient.
-The impact of aquifer thickness is captured in a similar way with \(\tau_{l}(\cdot)\) and \(\beta_{l}\) being the \(l\)th spline basis function and its coefficient, respectively.
-$K$ and $L$ are the numbers of basis functions for $D$ and $A$, respectively.
-$K = 4$ and $L = 4$ were chosen to balance preventing the estimated responses from being unrealistically oscillatory and ensuring a good fit as indicated by BIC.
-For this regression, counties with aquifer thickness levels greater than 180 m were excluded due to the sparsity of observations in that range because our flexible functional form approach cannot reliably estimate its impact.
-County-level soil characteristics including sand percentage, silt percentage, and water holding capacity, are represented by \(X_i\).
-State-year fixed effects and error term are represented by \(\theta_{s,t}\) and \(v_{i,t}\), respectively.
-Unlike for our regression models of per-area yields, we do not include county fixed effects to avoid eliminating the majority of variation in the aquifer thickness variable.
-The vast majority of variation in aquifer thickness is cross-sectional (across counties), rather than over time.
-Including county fixed effects would effectively eliminate the cross-sectional variation in aquifer thickness.
-A formal discussion of the trade-offs associated with the choice of fixed effects is provided in section \ref{county-fe} of the Supplementary Information.
-
-In the final step of our analysis, we then estimate the average crop yield as a function of aquifer thickness and seasonal water deficit.
-The average crop yield is estimated as the average of rainfed and irrigated yields estimated on the basis of Equation \eqref{eq:eqn1}, weighted by the estimated share of total acres that are irrigated in the county estimated using Equation \eqref{eq:eqn2}.
-In this combined model, aquifer thickness is a continuous variable for the irrigated area share regression component and a categorical variable for the irrigated yield regression following the format of Equations \eqref{eq:eqn2} and \eqref{eq:eqn1}, respectively.
-Comparing the estimated average crop yield across different levels of aquifer thickness for a given seasonal water deficit provides an indication of the effect of aquifer conditions on drought impacts, considering impacts of drought through changes in both per-area irrigated yields and the area of land irrigated.
-Standard errors of this analysis are estimated using bootstrap methods clustered by county.
-Note that irrigated area share and average crop yield estimates use data only from counties that have both irrigated and rainfed production observations because irrigated area share cannot be calculated when missing either one of the irrigated and rainfed observations.
-Counties with solely irrigated or rainfed production observation data (Section \ref{data-sets}) are used solely for estimating impacts of drought on per-area crop irrigated or rainfed yields respectively (Equation \ref{eq:eqn1}).
-
-\hypertarget{data-availability}{%
-\subsection{Data Availability}\label{data-availability}}
-
-All the raw data-sets used in this study are publicly accessible.
-The boundary shape file of the High Plains aquifer was obtained from the U.S.
-Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}.
-County and State boundaries were obtained using the R tigris package \citep{Rtigris}.
-These datasets were used for Figure \ref{fig:sat-map}.
-Corn and soybean yield data were obtained from the U.S.
-Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}.
-The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data.
-This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-Weather data was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}.
-Elevation data was obtained from the gridMET website at \url{https://www.climatologylab.org/gridmet.html}.
-These datasets were used to calculate water deficit, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis.
-The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}.
-This data is used for Figure \ref{fig:sat-map} and the regression analysis.
-Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which was used for regression analysis.
-The R soilDB package \citep{Rsoildb} was used to download the data.
-Regression analyses are implemented using the R fixest package \citep{Rfixest}.
-Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.
-A minimum set of data files required to reproduce the results and figures presented in this article can be found on the Figshare website (DOI: 10.6084/m9.figshare.24492196).
-Due to file size constraints on Figshare, not every data file is hosted there.
-However, a complete set of data files can be accessed via the provided Dropbox link: \url{https://www.dropbox.com/scl/fo/bghhwlidmi7wx1ok0az5n/h?rlkey=tgbix1hp7g9np9etlo1z3biyr&dl=0}.
-
-\subsection{Code Availability}
-All data processing and analyses were carried out using R \citep{R}.
-The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}.
-Additionally, instructions for reproducing all the results and figures presented in this article can be found at the bottom of the repository page in the README section.
-
-\clearpage
-
-\section*{Acknowledgements}
-
-Funding for the research in this article was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
-This research was also supported by Innovate UK award no.
-10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.''
-
-The funders had no role in study design, data collection and analysis, decision to publish or preparation of the article.
-
-\begin{itemize}
-\item OCE 58-0111-20-007 (TM, SK)
-\item OCE 58-0111-21-007 (TM, SK)
-\item Innovate UK award no. 10044695 (TF)
-\item Daugherty Water for Food Global Institute (SK)
-\end{itemize}
-
-\section*{Author Contributions}
-
-T.M.
-conceived the paper, developed the computer programs for data collection, processing, analysis, and summarization, and wrote the paper.; T.F.
-conceived and wrote the paper; S.K.
-wrote the computer programs for data collection and processing; and N.B.
-conceived and helped improve the framing of the paper.
-
-\section*{Competing Interests Statement}
-
-The authors declare no competing interests.
-
-\section*{Tables}
-
-\section*{Figures}
-
-%------------------------
-%- Figure 1
-%------------------------
-
-\begin{figure}[H]
-
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.}
-
-% {\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}}
-
-\caption*{\footnotesize{Note: The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
-Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}}
-\label{fig:yield-response}
-\end{figure}
-
-%------------------------
-%- Figure 2
-%------------------------
-
-\begin{figure}[H]
-
-\caption{The difference in estimated county-level irrigated corn and soybean yields}
-
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_ir_yield_dif}}
-
-\caption*{\footnotesize{Note: The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
-Shaded area represents 95\% confidence interval.}}
-\label{fig:yield-dif}
-\end{figure}
-
-%------------------------
-%- Figure 3
-%------------------------
-
-\begin{figure}[H]
-\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains}
-
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}}
-
-\caption*{\footnotesize{Note: The solid black line represents the estimated share of irrigated acres.
-Shaded area represents 95\% confidence interval.
-The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
-Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}}
-\label{fig:ir-share}
-\end{figure}
-
-%------------------------
-%- Figure 4
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Average productivity of corn and soybean in the US High Plains}
-
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
-\caption*{\footnotesize{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
-Please see section \ref{reg-conf} of the Supplementary Information for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
-The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}}
-\label{fig:tot-impact}
-\end{figure}
-
-%------------------------
-%- Figure 5
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Difference in average yield of corn and soybean for aquifer thickness levels}
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}}
-
-\caption*{\footnotesize{Note: The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
-Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}}
-\label{fig:dif-tot-impact}
-\end{figure}
-
-%------------------------
-%- Figure 6
-%------------------------
-
-\begin{figure}[H]
-
-\caption{Historical rainfed yield, irrigated yield, and water deficit}
-
-% {\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}}
-
-\caption*{\footnotesize{Note: The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}}
-\label{fig:deficit-yield-hist}
-\end{figure}
-
-%------------------------
-%- Figure 7
-%------------------------
-
-\begin{figure}[H]
-
-\caption{County-average aquifer thickness in 2016}
-
-% {\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}}
-
-\caption*{\footnotesize{Note: The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
-Counties highlighted in gray are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system.}}
-\label{fig:sat-map}
-\end{figure}
-
-\clearpage
-
-%==============================================
-%= References
-%==============================================
-
-\bibliography{DRA.bib}
-
-%==============================================
-%= Supplementary Information
-%==============================================
-\clearpage
-% appendix-appendix
 
@@
 Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigating farmers in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
-As discussed in section \ref{main} of our article, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
+As discussed in section Main of our article, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
 This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
@@
 
+\bibliography{DRA.bib}
 \end{document}

### 2023-11-08 Codes/2_figures_tables.rmd
generic manuscript.rnw created include both main and appendix

@@
 
-# Summary statistics
+# Figure 1: Yield response to water deficit by saturated thickness level (intensive margin)
 
-## Boxplot of water deficit by year
+```{r}
+# all_results <-  data.table(all_results)
+
+plot_yield_data <-
+  all_results[, .(crop, yield_pred_data)] %>%
+  rowwise() %>%
+  dplyr::mutate(plot_data = list(
+    yield_pred_data[, crop_type := crop] %>%
+      .[, crop_type := case_when(
+        crop_type == "corn" ~ "A. Corn",
+        crop_type == "soy" ~ "B. Sybean"
+      )]
+  )) %>%
+  dplyr::mutate(sat_cat_text_ls = list(
+    plot_data[, .(sat_rank, sat_cat_text)] %>%
+      .[!is.na(sat_rank), ] %>%
+      unique(by = "sat_rank") %>%
+      .[order(sat_rank), sat_cat_text]
+  )) %>%
+  dplyr::mutate(plot_data = list(
+    plot_data %>%
+      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
+      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))]
+  ))
+
+g_ir_yield_both <-
+  plot_yield_data %>%
+  dplyr::mutate(g_yield = list(
+    ggplot(plot_data) +
+      geom_line(aes(
+        y = y_hat,
+        x = balance,
+        color = factor(sat_cat_text)
+      )) +
+      scale_x_continuous(
+        name = "Water Deficit (mm)",
+        breaks = (-2:11) * 100
+      ) +
+      scale_y_continuous(
+        name = "Yield (tonne/ha)",
+        breaks = 0:(ceiling(max(plot_data$y_hat)) + 0.5),
+        limits = c(NA, (ceiling(max(plot_data$y_hat)) + 0.5))
+      ) +
+      scale_color_discrete(
+        name = "Aquifer Thickness",
+        guide = guide_legend(
+          title.position = "top",
+          nrow = 2
+        )
+      ) +
+      scale_fill_discrete(guide = "none") +
+      theme_fig +
+      coord_cartesian(xlim = c(-200, 1150)) +
+      theme(
+        legend.position = "bottom",
+        plot.margin = margin(0, 0, 0.0, 0, "cm"),
+        plot.title = element_blank()
+      )
+  ))
+
+g_ir_yield_corn <- g_ir_yield_both$g_yield[[1]] + theme(legend.position = "none")
+g_ir_yield_soy <- g_ir_yield_both$g_yield[[2]]
+
+g_wds <-
+  all_results[, .(crop, data)] %>%
+  dplyr::mutate(g_fig = purrr::map(
+    data,
+    function(data) {
+      ggplot(data) +
+        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
+        theme_fig +
+        xlab("") +
+        ylab("") +
+        coord_cartesian(xlim = c(-200, 1150))
+    }
+  ))
+
+g_wd_corn <-
+  g_wds$g_fig[[1]] +
+  ggtitle("A. Corn") +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.border = element_blank(),
+    plot.margin = margin(0, 0, -0.5, 0, "cm")
+  )
+
+g_wd_soy <-
+  g_wds$g_fig[[2]] +
+  ggtitle("B. Soybean") +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.border = element_blank(),
+    plot.margin = margin(0, 0, -0.5, 0, "cm")
+  )
+
+g_yield_response <-
+  cowplot::plot_grid(
+    g_wd_corn,
+    g_ir_yield_corn + theme(plot.margin = margin(0, 0, 0.2, 0, "cm")), g_wd_soy,
+    g_ir_yield_soy,
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.18 / 2, 0.69 / 2, 0.18 / 2, 0.95 / 2)
+  )
+
+cowplot::ggsave2("Results/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)
+```
+
+# Figure 2: Test the significance of the difference in irrigated yield between the aquifer thickness categories
+
+```{r}
+g_dif <-
+  all_results[, .(crop, yield_dif_test, sat_text_data)] %>%
+  rowwise() %>%
+  dplyr::mutate(plot_data = list(
+    yield_dif_test[, crop_type := crop] %>%
+      .[, crop_type := case_when(
+        crop_type == "corn" ~ "A. Corn",
+        crop_type == "soy" ~ "B. Sybean"
+      )] %>%
+      setnames("comp_q", "sat_rank") %>%
+      sat_text_data[., on = "sat_rank"] %>%
+      .[, dif_upper := dif_hat + 1.96 * se] %>%
+      .[, dif_lower := dif_hat - 1.96 * se]
+  )) %>%
+  dplyr::mutate(sat_cat_text_ls = list(
+    plot_data[, .(sat_rank, sat_cat_text)] %>%
+      .[!is.na(sat_rank), ] %>%
+      unique(by = "sat_rank") %>%
+      .[order(sat_rank), sat_cat_text]
+  )) %>%
+  dplyr::mutate(plot_data = list(
+    plot_data[, sat_cat_text := factor(sat_cat_text, levels = sat_cat_text_ls)]
+  )) %>%
+  dplyr::mutate(g_fig = list(
+    ggplot(plot_data) +
+      facet_grid(. ~ sat_cat_text) +
+      geom_ribbon(
+        aes(
+          x = balance,
+          ymin = dif_lower,
+          ymax = dif_upper
+        ),
+        fill = "blue",
+        alpha = 0.3
+      ) +
+      geom_line(
+        aes(
+          x = balance,
+          y = dif_hat
+        ),
+        color = "red"
+      ) +
+      geom_hline(yintercept = 0, col = "black") +
+      scale_color_discrete(name = "") +
+      scale_x_continuous(breaks = (-1:6) * 200) +
+      ylab("Difference in Yield (tonne/ha)") +
+      xlab("Water Deficit (mm)") +
+      theme_fig +
+      theme(
+        strip.background = element_blank(),
+        legend.position = "bottom",
+        panel.spacing = unit(1, "lines")
+      )
+  ))
+
+g_ir_yield_dif <-
+  cowplot::plot_grid(
+    g_dif$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
+    g_dif$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.5, 0.5)
+  )
+
+cowplot::ggsave2("Results/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
+```
+
+# Figure 3: The impact of saturated thickness on the share of irrigated production
+
+```{r}
+#++++++++++++++++++++++++++++++++++++
+#+ Share response
+#++++++++++++++++++++++++++++++++++++
+plot_data_ir_share <-
+  all_results[, .(crop, share_pred_data)] %>%
+  unnest(cols = share_pred_data) %>%
+  data.table() %>%
+  .[, crop_type := case_when(
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
+  )]
+
+g_share <-
+  plot_data_ir_share %>%
+  nest_by(crop) %>%
+  dplyr::mutate(g_fig = list(
+    ggplot(dplyr::filter(data, sat <= 150)) +
+      geom_line(aes(y = ir_share_hat, x = sat)) +
+      geom_ribbon(
+        aes(
+          ymin = ir_share_hat - 1.96 * ir_share_hat_se,
+          ymax = ir_share_hat + 1.96 * ir_share_hat_se,
+          x = sat
+        ),
+        alpha = 0.4
+      ) +
+      scale_y_continuous(breaks = seq(0, 1, by = 0.1), limits = c(0.4, 1)) +
+      scale_x_continuous(breaks = seq(10, 150, by = 10)) +
+      coord_cartesian(xlim = c(9, 150)) +
+      xlab("Aquifer Thickness (meter)") +
+      ylab("Share of Irrigated Acres") +
+      theme_fig +
+      theme(
+        legend.position = "bottom"
+      )
+  ))
+
+#++++++++++++++++++++++++++++++++++++
+#+ Saturated thickness histogram
+#++++++++++++++++++++++++++++++++++++
+g_sat_hist <-
+  all_results[, .(crop, data)] %>%
+  dplyr::mutate(g_fig = purrr::map(
+    data,
+    function(data) {
+      ggplot(data[sat < 150 & ir == "ir", ]) +
+        geom_histogram(aes(x = sat), bins = 30, col = "darkgreen", fill = NA) +
+        theme_fig +
+        xlab("") +
+        ylab("") +
+        coord_cartesian(xlim = c(9, 150)) +
+        theme(
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.ticks.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank(),
+          plot.margin = margin(0, 0, -0.5, 0, "cm")
+        )
+    }
+  ))
+
+g_ir_share <-
+  cowplot::plot_grid(
+    g_sat_hist$g_fig[[1]] + ggtitle("A. Corn"),
+    g_share$g_fig[[1]] + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
+    g_sat_hist$g_fig[[2]] + ggtitle("B. Soybean"),
+    g_share$g_fig[[2]],
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.18 / 2, 0.64 / 2, 0.18 / 2, 0.64 / 2)
+  )
+
+ggsave(
+  file = "Results/Figures/g_share.pdf",
+  g_ir_share,
+  height = 6,
+  width = 6
+)
+```
+
+# Figure 4: Total impact of a decline in saturated thickness on average yield
+
+```{r}
+g_total_both <-
+  all_results[, .(crop, avg_yield_sum)] %>%
+  unnest() %>%
+  data.table() %>%
+  .[, crop_type := case_when(
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
+  )] %>%
+  nest_by(crop_type) %>%
+  data.table() %>%
+  dplyr::mutate(g_fig = purrr::map(data, \(data) {
+    ggplot(data) +
+      geom_line(aes(y = avg_yield, x = balance, color = factor(sat))) +
+      theme_fig +
+      ylim(min(data$avg_yield) - 0.5, max(data$avg_yield) + 0.5) +
+      ylab("Yield (tonne/ha)") +
+      scale_color_discrete(
+        "Aquifer Thickness (meter)",
+        guide = guide_legend(
+          title.position = "top",
+          nrow = 1
+        )
+      ) +
+      scale_x_continuous(
+        name = "Water Deficit (mm)",
+        breaks = (-2:12) * 100
+      ) +
+      coord_cartesian(xlim = c(-200, 1150)) +
+      theme(legend.position = "bottom")
+  }))
+
+g_total_corn <-
+  g_total_both$g_fig[[1]] +
+  theme(legend.position = "none") +
+  ylim(6, 13)
+
+g_total_soy <-
+  g_total_both$g_fig[[2]] +
+  ylim(1.5, 4)
+
+g_wds <-
+  all_results[, .(crop, data)] %>%
+  dplyr::mutate(g_fig = purrr::map(
+    data,
+    function(data) {
+      ggplot(data) +
+        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
+        theme_fig +
+        xlab("") +
+        ylab("") +
+        coord_cartesian(xlim = c(-200, 1150)) +
+        theme(
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.ticks.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank(),
+          plot.margin = margin(0, 0, -0.5, 0, "cm")
+        )
+    }
+  ))
+
+g_total_impact <-
+  cowplot::plot_grid(
+    g_wds$g_fig[[1]] + ggtitle("A. Corn"),
+    g_total_corn + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
+    g_wds$g_fig[[2]] + ggtitle("B. Soybean"),
+    g_total_soy,
+    ncol = 1,
+    align = "v",
+    rel_heights = c(0.18 / 2, 0.65 / 2, 0.18 / 2, 0.8 / 2)
+  )
+
+ggsave(
+  file = "Results/Figures/g_total_impact.pdf",
+  g_total_impact,
+  height = 6,
+  width = 6
+)
+```
+
+# Figure 5: Testing of the significance of yield difference (average total impact)
+
+```{r}
+g_avg_yield_dif <-
+  all_results[, .(crop, avg_yield_sum)] %>%
+  unnest(cols = c(avg_yield_sum)) %>%
+  data.table() %>%
+  .[, sat_text := paste0(
+    "Aquifer Thickness = ",
+    format(sat, digits = 3),
+    "m"
+  )] %>%
+  .[sat != min(sat), ] %>%
+  nest_by(crop) %>%
+  data.table() %>%
+  dplyr::mutate(g_fig = purrr::map(data, \(data){
+    ggplot(data) +
+      geom_ribbon(aes(x = balance, ymin = dif_lower, ymax = dif_upper), alpha = 0.3, fill = "blue") +
+      geom_line(aes(x = balance, y = dif_avg_yield), color = "red") +
+      geom_hline(yintercept = 0) +
+      facet_grid(. ~ sat_text, scales = "free_y") +
+      xlab("Water Deficit (mm)") +
+      ylab("Difference in Yield (tonne/ha)") +
+      xlim(-200, 1200) +
+      theme_fig +
+      theme(
+        strip.background = element_blank(),
+        panel.spacing.y = unit(1, "cm")
+      )
+  }))
+
+g_ay_corn <-
+  g_avg_yield_dif$g_fig[[1]] +
+  theme(
+    legend.position = "none",
+    axis.title.x = element_blank()
+  ) +
+  scale_y_continuous(breaks = -2:5, limits = c(-2, 6)) +
+  ggtitle("A. Corn")
+
+g_ay_soy <-
+  g_avg_yield_dif$g_fig[[2]] +
+  scale_y_continuous(breaks = -2:2, limits = c(-2, 2)) +
+  ggtitle("B. Soybean")
+
+g_ay <- g_ay_corn / g_ay_soy
+
+ggsave("Results/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
+```
+
+# Water deficit by year (Figure 6)
 
@@
 
-## Saturated Thickness Map
+# Saturated Thickness Map (Figure 7)
 
@@
 
-# Yield response to water deficit by saturated thickness level (intensive margin)
+# Supplementary Information
 
+## Figure S.1 and S.2 
 ```{r}
-# all_results <-  data.table(all_results)
-
-plot_yield_data <-
-  all_results[, .(crop, yield_pred_data)] %>%
-  rowwise() %>%
-  dplyr::mutate(plot_data = list(
-    yield_pred_data[, crop_type := crop] %>%
-      .[, crop_type := case_when(
-        crop_type == "corn" ~ "A. Corn",
-        crop_type == "soy" ~ "B. Sybean"
-      )]
-  )) %>%
-  dplyr::mutate(sat_cat_text_ls = list(
-    plot_data[, .(sat_rank, sat_cat_text)] %>%
-      .[!is.na(sat_rank), ] %>%
-      unique(by = "sat_rank") %>%
-      .[order(sat_rank), sat_cat_text]
-  )) %>%
-  dplyr::mutate(plot_data = list(
-    plot_data %>%
-      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
-      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))]
-  ))
-
-g_ir_yield_both <-
-  plot_yield_data %>%
-  dplyr::mutate(g_yield = list(
-    ggplot(plot_data) +
-      geom_line(aes(
-        y = y_hat,
-        x = balance,
-        color = factor(sat_cat_text)
-      )) +
-      scale_x_continuous(
-        name = "Water Deficit (mm)",
-        breaks = (-2:11) * 100
-      ) +
-      scale_y_continuous(
-        name = "Yield (tonne/ha)",
-        breaks = 0:(ceiling(max(plot_data$y_hat)) + 0.5),
-        limits = c(NA, (ceiling(max(plot_data$y_hat)) + 0.5))
-      ) +
-      scale_color_discrete(
-        name = "Aquifer Thickness",
-        guide = guide_legend(
-          title.position = "top",
-          nrow = 2
-        )
-      ) +
-      scale_fill_discrete(guide = "none") +
-      theme_fig +
-      coord_cartesian(xlim = c(-200, 1150)) +
-      theme(
-        legend.position = "bottom",
-        plot.margin = margin(0, 0, 0.0, 0, "cm"),
-        plot.title = element_blank()
-      )
-  ))
-
-g_ir_yield_corn <- g_ir_yield_both$g_yield[[1]] + theme(legend.position = "none")
-g_ir_yield_soy <- g_ir_yield_both$g_yield[[2]]
-
-g_wds <-
-  all_results[, .(crop, data)] %>%
-  dplyr::mutate(g_fig = purrr::map(
-    data,
-    function(data) {
-      ggplot(data) +
-        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
-        theme_fig +
-        xlab("") +
-        ylab("") +
-        coord_cartesian(xlim = c(-200, 1150))
-    }
-  ))
-
-g_wd_corn <-
-  g_wds$g_fig[[1]] +
-  ggtitle("A. Corn") +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.ticks.y = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    panel.border = element_blank(),
-    plot.margin = margin(0, 0, -0.5, 0, "cm")
-  )
-
-g_wd_soy <-
-  g_wds$g_fig[[2]] +
-  ggtitle("B. Soybean") +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.ticks.y = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    panel.border = element_blank(),
-    plot.margin = margin(0, 0, -0.5, 0, "cm")
-  )
-
-g_yield_response <-
-  cowplot::plot_grid(
-    g_wd_corn,
-    g_ir_yield_corn + theme(plot.margin = margin(0, 0, 0.2, 0, "cm")), g_wd_soy,
-    g_ir_yield_soy,
-    ncol = 1,
-    align = "v",
-    rel_heights = c(0.18 / 2, 0.69 / 2, 0.18 / 2, 0.95 / 2)
-  )
-
-cowplot::ggsave2("Results/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)
-
-#++++++++++++++++++++++++++++++++++++
-#+ With confidence interval (seprately)
-#++++++++++++++++++++++++++++++++++++
 g_yield_with_conf <-
@@
 
-ggsave("Results/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
-ggsave("Results/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
+ggsave("Results/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 5, height = 3.5)
+ggsave("Results/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 5, height = 3.5)
 ```
 
-# The impact of saturated thickness on the share of irrigated production
-
-```{r}
-#++++++++++++++++++++++++++++++++++++
-#+ Share response
-#++++++++++++++++++++++++++++++++++++
-plot_data_ir_share <-
-  all_results[, .(crop, share_pred_data)] %>%
-  unnest(cols = share_pred_data) %>%
-  data.table() %>%
-  .[, crop_type := case_when(
-    crop == "corn" ~ "A. Corn",
-    crop == "soy" ~ "B. Soybean"
-  )]
-
-g_share <-
-  plot_data_ir_share %>%
-  nest_by(crop) %>%
-  dplyr::mutate(g_fig = list(
-    ggplot(dplyr::filter(data, sat <= 150)) +
-      geom_line(aes(y = ir_share_hat, x = sat)) +
-      geom_ribbon(
-        aes(
-          ymin = ir_share_hat - 1.96 * ir_share_hat_se,
-          ymax = ir_share_hat + 1.96 * ir_share_hat_se,
-          x = sat
-        ),
-        alpha = 0.4
-      ) +
-      scale_y_continuous(breaks = seq(0, 1, by = 0.1), limits = c(0.4, 1)) +
-      scale_x_continuous(breaks = seq(10, 150, by = 10)) +
-      coord_cartesian(xlim = c(9, 150)) +
-      xlab("Aquifer Thickness (meter)") +
-      ylab("Share of Irrigated Acres") +
-      theme_fig +
-      theme(
-        legend.position = "bottom"
-      )
-  ))
-
-#++++++++++++++++++++++++++++++++++++
-#+ Saturated thickness histogram
-#++++++++++++++++++++++++++++++++++++
-g_sat_hist <-
-  all_results[, .(crop, data)] %>%
-  dplyr::mutate(g_fig = purrr::map(
-    data,
-    function(data) {
-      ggplot(data[sat < 150 & ir == "ir", ]) +
-        geom_histogram(aes(x = sat), bins = 30, col = "darkgreen", fill = NA) +
-        theme_fig +
-        xlab("") +
-        ylab("") +
-        coord_cartesian(xlim = c(9, 150)) +
-        theme(
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks.x = element_blank(),
-          axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(),
-          panel.grid.minor = element_blank(),
-          panel.border = element_blank(),
-          plot.margin = margin(0, 0, -0.5, 0, "cm")
-        )
-    }
-  ))
-
-g_ir_share <-
-  cowplot::plot_grid(
-    g_sat_hist$g_fig[[1]] + ggtitle("A. Corn"),
-    g_share$g_fig[[1]] + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
-    g_sat_hist$g_fig[[2]] + ggtitle("B. Soybean"),
-    g_share$g_fig[[2]],
-    ncol = 1,
-    align = "v",
-    rel_heights = c(0.18 / 2, 0.64 / 2, 0.18 / 2, 0.64 / 2)
-  )
-
-ggsave(
-  file = "Results/Figures/g_share.pdf",
-  g_ir_share,
-  height = 6,
-  width = 6
-)
-```
-
-# Difference in the share of irrigated production (from 100)
-
-```{r}
-plot_data_ir_share <-
-  all_results[, .(crop, share_sum)] %>%
-  unnest(cols = share_sum) %>%
-  data.table() %>%
-  .[, crop_type := case_when(
-    crop == "corn" ~ "A. Corn",
-    crop == "soy" ~ "B. Soybean"
-  )]
-
-g_share_dif <-
-  ggplot(plot_data_ir_share[sat <= 100, ]) +
-  geom_line(aes(y = dif_ir_share_hat, x = sat)) +
-  geom_ribbon(
-    aes(
-      ymin = dif_lower,
-      ymax = dif_upper,
-      x = sat
-    ),
-    alpha = 0.4
-  ) +
-  geom_hline(yintercept = 0, color = "red") +
-  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
-  scale_x_continuous(breaks = seq(10, 150, by = 10)) +
-  xlab("Aquifer Thickness (meter)") +
-  ylab("Decline in Estimated Share of Irrigated Acres") +
-  facet_wrap(crop_type ~ ., ncol = 1) +
-  theme_fig +
-  theme(
-    legend.position = "bottom",
-    strip.background = element_blank()
-  )
-
-ggsave(
-  file = "Results/Figures/g_share_dif.pdf",
-  g_share_dif,
-  height = 4,
-  width = 6
-)
-```
-
-# Total impact of a decline in saturated thickness on average yield
-
-```{r}
-g_total_both <-
-  all_results[, .(crop, avg_yield_sum)] %>%
-  unnest() %>%
-  data.table() %>%
-  .[, crop_type := case_when(
-    crop == "corn" ~ "A. Corn",
-    crop == "soy" ~ "B. Soybean"
-  )] %>%
-  nest_by(crop_type) %>%
-  data.table() %>%
-  dplyr::mutate(g_fig = purrr::map(data, \(data) {
-    ggplot(data) +
-      geom_line(aes(y = avg_yield, x = balance, color = factor(sat))) +
-      theme_fig +
-      ylim(min(data$avg_yield) - 0.5, max(data$avg_yield) + 0.5) +
-      ylab("Yield (tonne/ha)") +
-      scale_color_discrete(
-        "Aquifer Thickness (meter)",
-        guide = guide_legend(
-          title.position = "top",
-          nrow = 1
-        )
-      ) +
-      scale_x_continuous(
-        name = "Water Deficit (mm)",
-        breaks = (-2:12) * 100
-      ) +
-      coord_cartesian(xlim = c(-200, 1150)) +
-      theme(legend.position = "bottom")
-  }))
-
-g_total_corn <-
-  g_total_both$g_fig[[1]] +
-  theme(legend.position = "none") +
-  ylim(6, 13)
-
-g_total_soy <-
-  g_total_both$g_fig[[2]] +
-  ylim(1.5, 4)
-
-g_wds <-
-  all_results[, .(crop, data)] %>%
-  dplyr::mutate(g_fig = purrr::map(
-    data,
-    function(data) {
-      ggplot(data) +
-        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
-        theme_fig +
-        xlab("") +
-        ylab("") +
-        coord_cartesian(xlim = c(-200, 1150)) +
-        theme(
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks.x = element_blank(),
-          axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(),
-          panel.grid.minor = element_blank(),
-          panel.border = element_blank(),
-          plot.margin = margin(0, 0, -0.5, 0, "cm")
-        )
-    }
-  ))
-
-g_total_impact <-
-  cowplot::plot_grid(
-    g_wds$g_fig[[1]] + ggtitle("A. Corn"),
-    g_total_corn + theme(plot.margin = margin(0, 0, 0.5, 0, "cm")),
-    g_wds$g_fig[[2]] + ggtitle("B. Soybean"),
-    g_total_soy,
-    ncol = 1,
-    align = "v",
-    rel_heights = c(0.18 / 2, 0.65 / 2, 0.18 / 2, 0.8 / 2)
-  )
-
-ggsave(
-  file = "Results/Figures/g_total_impact.pdf",
-  g_total_impact,
-  height = 6,
-  width = 6
-)
-```
-
-# Test the significance of the difference in irrigated yield between the aquifer thickness categories
-
-```{r}
-g_dif <-
-  all_results[, .(crop, yield_dif_test, sat_text_data)] %>%
-  rowwise() %>%
-  dplyr::mutate(plot_data = list(
-    yield_dif_test[, crop_type := crop] %>%
-      .[, crop_type := case_when(
-        crop_type == "corn" ~ "A. Corn",
-        crop_type == "soy" ~ "B. Sybean"
-      )] %>%
-      setnames("comp_q", "sat_rank") %>%
-      sat_text_data[., on = "sat_rank"] %>%
-      .[, dif_upper := dif_hat + 1.96 * se] %>%
-      .[, dif_lower := dif_hat - 1.96 * se]
-  )) %>%
-  dplyr::mutate(sat_cat_text_ls = list(
-    plot_data[, .(sat_rank, sat_cat_text)] %>%
-      .[!is.na(sat_rank), ] %>%
-      unique(by = "sat_rank") %>%
-      .[order(sat_rank), sat_cat_text]
-  )) %>%
-  dplyr::mutate(plot_data = list(
-    plot_data[, sat_cat_text := factor(sat_cat_text, levels = sat_cat_text_ls)]
-  )) %>%
-  dplyr::mutate(g_fig = list(
-    ggplot(plot_data) +
-      facet_grid(. ~ sat_cat_text) +
-      geom_ribbon(
-        aes(
-          x = balance,
-          ymin = dif_lower,
-          ymax = dif_upper
-        ),
-        fill = "blue",
-        alpha = 0.3
-      ) +
-      geom_line(
-        aes(
-          x = balance,
-          y = dif_hat
-        ),
-        color = "red"
-      ) +
-      geom_hline(yintercept = 0, col = "black") +
-      scale_color_discrete(name = "") +
-      scale_x_continuous(breaks = (-1:6) * 200) +
-      ylab("Difference in Yield (tonne/ha)") +
-      xlab("Water Deficit (mm)") +
-      theme_fig +
-      theme(
-        strip.background = element_blank(),
-        legend.position = "bottom",
-        panel.spacing = unit(1, "lines")
-      )
-  ))
-
-g_ir_yield_dif <-
-  cowplot::plot_grid(
-    g_dif$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
-    g_dif$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
-    ncol = 1,
-    align = "v",
-    rel_heights = c(0.5, 0.5)
-  )
-
-cowplot::ggsave2("Results/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
-```
-
-# Testing of the significance of yield difference (average total impact)
-
-```{r}
-g_avg_yield_dif <-
-  all_results[, .(crop, avg_yield_sum)] %>%
-  unnest(cols = c(avg_yield_sum)) %>%
-  data.table() %>%
-  .[, sat_text := paste0(
-    "Aquifer Thickness = ",
-    format(sat, digits = 3),
-    "m"
-  )] %>%
-  .[sat != min(sat), ] %>%
-  nest_by(crop) %>%
-  data.table() %>%
-  dplyr::mutate(g_fig = purrr::map(data, \(data){
-    ggplot(data) +
-      geom_ribbon(aes(x = balance, ymin = dif_lower, ymax = dif_upper), alpha = 0.3, fill = "blue") +
-      geom_line(aes(x = balance, y = dif_avg_yield), color = "red") +
-      geom_hline(yintercept = 0) +
-      facet_grid(. ~ sat_text, scales = "free_y") +
-      xlab("Water Deficit (mm)") +
-      ylab("Difference in Yield (tonne/ha)") +
-      xlim(-200, 1200) +
-      theme_fig +
-      theme(
-        strip.background = element_blank(),
-        panel.spacing.y = unit(1, "cm")
-      )
-  }))
-
-g_ay_corn <-
-  g_avg_yield_dif$g_fig[[1]] +
-  theme(
-    legend.position = "none",
-    axis.title.x = element_blank()
-  ) +
-  scale_y_continuous(breaks = -2:5, limits = c(-2, 6)) +
-  ggtitle("A. Corn")
-
-g_ay_soy <-
-  g_avg_yield_dif$g_fig[[2]] +
-  scale_y_continuous(breaks = -2:2, limits = c(-2, 2)) +
-  ggtitle("B. Soybean")
-
-g_ay <- g_ay_corn / g_ay_soy
-
-ggsave("Results/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
-```
-
-# The impact of water deficit on average yield by aquifer thickness category with confidence interval
+## Figure S.3: The impact of water deficit on average yield by aquifer thickness category with confidence interval
 
@@
 
-ggsave("Results/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
+ggsave("Results/Figures/g_avg_yield.pdf", g_ay_response, width = 5, height = 4)
 ```
+
+# Figure S.4: Difference in the share of irrigated production (from 100)
+
+```{r}
+plot_data_ir_share <-
+  all_results[, .(crop, share_sum)] %>%
+  unnest(cols = share_sum) %>%
+  data.table() %>%
+  .[, crop_type := case_when(
+    crop == "corn" ~ "A. Corn",
+    crop == "soy" ~ "B. Soybean"
+  )]
+
+g_share_dif <-
+  ggplot(plot_data_ir_share[sat <= 100, ]) +
+  geom_line(aes(y = dif_ir_share_hat, x = sat)) +
+  geom_ribbon(
+    aes(
+      ymin = dif_lower,
+      ymax = dif_upper,
+      x = sat
+    ),
+    alpha = 0.4
+  ) +
+  geom_hline(yintercept = 0, color = "red") +
+  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
+  scale_x_continuous(breaks = seq(10, 150, by = 10)) +
+  xlab("Aquifer Thickness (meter)") +
+  ylab("Decline in Estimated Share of Irrigated Acres") +
+  facet_wrap(crop_type ~ ., ncol = 1) +
+  theme_fig +
+  theme(
+    legend.position = "bottom",
+    strip.background = element_blank()
+  )
+
+ggsave(
+  file = "Results/Figures/g_share_dif.pdf",
+  g_share_dif,
+  height = 4,
+  width = 5
+)
+```

### 2023-11-08 Writing/manuscript.Rnw
generic manuscript.rnw created include both main and appendix

@@
 %
-\documentclass[
-]{article}
+% \documentclass[sn-nature]{sn-jnl}
+\documentclass{article}
 \usepackage{amsmath,amssymb}
 \usepackage{lmodern}
+\usepackage[labelfont=bf]{caption} 
 \usepackage{iftex}
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+%\urlstyle{same} % disable monospaced font for URLs
+\thispagestyle{empty}
 \hypersetup{
@@
   hidelinks,
-  pdfcreator={LaTeX via pandoc}}
+  pdfcreator={LaTeX via pandoc}
+}
 
-<<init, echo = FALSE>>=
+\begin{document}
+
+<<init, include = FALSE>>=
 knitr::opts_chunk$set(
@@
 \author[4]{Nicholas Brozovi\'{c}}
-\affil[1]{Department of Agricultural Economics, University of Nebraska Lincoln}
-\affil[2]{School of Mechanical, Aerospace and Civil Engineering, University of Manchester}
+\affil[1]{Department of Agricultural Economics, University of Nebraska-Lincoln}
+\affil[2]{Department of Mechanical, Aerospace and Civil Engineering, University of Manchester}
 \affil[3]{Department of Applied Economics, University of Minnesota}
 \affil[4]{Daugherty Water for Food Global Institute at the University of Nebraska}
-\affil[4]{Department of Agricultural Economics, University of Nebraska Lincoln}
+\affil[4]{Department of Agricultural Economics, University of Nebraska-Lincoln}
 
-\date{}
+\maketitle
 
-\begin{document}
-\SweaveOpts{concordance=TRUE}
-\maketitle
-\begin{abstract}
+\abstract{
 Aquifer depletion poses a major threat to the ability of farmers, food supply chains, and rural economies globally to use groundwater as a means of adapting to climate variability and change.
@@
 Our findings demonstrate that drought-related production losses on irrigated croplands increase non-linearly with aquifer depletion, highlighting the need for proactive aquifer conservation interventions to support adaptation and resilience to future increases in rainfall variability under climate change.
-\end{abstract}
+}
 
@@
 Understanding how these two responses vary as a function of aquifer conditions is crucial to guide groundwater management planning decisions, in particular given projected increases in frequency and intensity of droughts under climate change \citep{ukkola2020robust, chiang2021evidence, cook2020twenty}.
+Another potential pathway is through groundwater use regulation in response to low or declining aquifer thickness.
+However, it is unlikely that this effects is captured in our analysis.
+For more details on this issue, please refer to section \ref{pathway-reg} of the Supplementary Information.
 
@@
 \hypertarget{impact-intensive}{%
-\subsection{Aquifer depletion increases sensitivity of crop yields to water deficits}\label{impact-intensive}}
+\subsection{Sensitivity of per-area crop yields to water deficits}\label{impact-intensive}}
 
-Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B)\footnote{Confidence intervals were not presented to avoid over-crowding the figures.
-Please see Appendix \ref{reg-conf} for estimated yield response curves with 95 percent confidence interval.}.
+Figure \ref{fig:yield-response} shows the estimated crop yield response to water deficit for rainfed production and groundwater-irrigated production at different levels of aquifer thickness categories for corn (panel A) and soybean (panel B).
+Confidence intervals were not presented to avoid over-crowding the figures.
+Please see section \ref{reg-conf} of the Supplementary Information for estimated yield response curves with 95 percent confidence interval.
 The histogram of annual water deficit is presented at the top for each crop.
@@
 Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean.
-As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (See Appendix \ref{test-dif-share} for formal statistical testings).
+As aquifer thickness declines, the share of irrigated acres for corn decreases from around \Sexpr{share_high} (\Sexpr{share_high * 100}\%) for an average county-level aquifer thickness of 100 m to \Sexpr{share_low} (\Sexpr{share_low * 100}\%) for an average county-level aquifer thickness of 10 m (Please see section \ref{test-dif-share} of the Supplementary Information for formal statistical testings).
 Reductions in the share of production area under irrigation is also evident for soybean.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}
-
-}
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}
-
-}
-
-\caption{Historical rainfed yield, irrigated yield, and water deficit}\label{fig:deficit-yield-hist}
-\end{figure}
-
 As a proxy for the extent of aquifer depletion we use annual estimates of aquifer thickness generated by existing research \citep{haacker2016water,haacker2023}.
-These data provide annual estimates of the aquifer thickness at 250m resolution since 1935.
+These data provide annual estimates of the aquifer thickness at 250 m resolution since 1935.
 Annual aquifer thickness maps were derived based on interpolation of 14,000 to 20,000 individual observations of water levels per year, along with detailed maps of both the land surface elevation and bedrock elevation of the HPA.
@@
 The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category.
-Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works).
+Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category-specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (For further explanations on how spline-based regression functions, please see section \ref{spline-basis} of the Supplementary Information.).
 $K$ is the total number of spline basis functions.
 $K = 3$ was chosen for this regression to prevent our estimated response from becoming unrealistically oscillatory while maintaining a good fit based on Bayesian Information Criterion (BIC).
-To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.
+To control for time-invariant heterogeneity across counties and yearly shocks, county-irrigation-status fixed effects $\alpha_{i, ir}$ and year fixed effects $\lambda_t$ are also included.
+County-irrigation-status fixed effects are used instead of county fixed effects because some counties have both rainfed and irrigated observations, and they need to be distinguished.
+Standard error estimation is clustered by county.
 
@@
 
-Funding for the research in this manuscript was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
+Funding for the research in this article was provided by the United States Department of Agriculture under contract numbers OCE 58-0111-20-007 and OCE 58-0111-21-007 and the Daugherty Water for Food Global Institute at the University of Nebraska.
 This research was also supported by Innovate UK award no.
-10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.'' The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
+10044695, as part of the UK Research and Innovation and European Commission funded project ``TRANSCEND: Transformational and robust adaptation to water scarcity and climate change under deep uncertainty.''
+
+The funders had no role in study design, data collection and analysis, decision to publish or preparation of the article.
 
@@
 
-\subsection*{Author Contributions}
+\section*{Author Contributions}
 
-\subsection*{Competing Interests Statement}
+T.M.
+conceived the paper, developed the computer programs for data collection, processing, analysis, and summarization, and wrote the paper.; T.F.
+conceived and wrote the paper; S.K.
+wrote the computer programs for data collection and processing; and N.B.
+conceived and helped improve the framing of the paper.
 
+\section*{Competing Interests Statement}
+
+The authors declare no competing interests.
+
+\section*{Tables}
+
+\section*{Figures}
+
+%------------------------
+%- Figure 1
+%------------------------
+
+\begin{figure}[H]
+
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area yields of corn and soybean in US High Plains.}
+
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_yield_intensive}}
+
+\caption*{\footnotesize{Note: The solid lines represent per-area irrigated yield responses to water deficits, dissagregated by aquifer thickness level, considering three quantile groups that capture the range of aquifer thickness values in our sample: (i) 9 m to 30.1 m; (ii) 30.1 m to 67.2 m; and (iii) 67.2 m to 215.8 m.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.
+Total number of observations are $\Sexpr{nobs_corn}$ and $\Sexpr{nobs_soy}$ for corn and soybean regressions, respectively.}}
+\label{fig:yield-response}
+\end{figure}
+
+%------------------------
+%- Figure 2
+%------------------------
+
+\begin{figure}[H]
+
+\caption{The difference in estimated county-level irrigated corn and soybean yields}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_ir_yield_dif}}
+
+\caption*{\footnotesize{Note: The solid red lines represent the difference in estimated per-area irrigated corn (top panels) and soybean (bottom panels) yields for the second and third (largest) aquifer thickness quantiles relative to the first (smallest) aquifer thickness quantile.
+Shaded area represents 95\% confidence interval.}}
+\label{fig:yield-dif}
+\end{figure}
+
+%------------------------
+%- Figure 3
+%------------------------
+
+\begin{figure}[H]
+\caption{The estimated impact of aquifer thickness on the share of agricultural production under irrigation for corn and soybean producing counties in the US High Plains}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share}}
+
+\caption*{\footnotesize{Note: The solid black line represents the estimated share of irrigated acres.
+Shaded area represents 95\% confidence interval.
+The histograms on the top of each panel show the distribution of county-level aquifer thickness values included for each crop.
+Total number of observations are $\Sexpr{nobs_corn_share}$ and $\Sexpr{nobs_soy_share}$ for corn and soybean regressions, respectively.}}
+\label{fig:ir-share}
+\end{figure}
+
+%------------------------
+%- Figure 4
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Average productivity of corn and soybean in the US High Plains}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_total_impact}}
+
+\caption*{\footnotesize{Note: The solid lines represent average yield at different levels of water deficit and aquifer thickness, considering both per-area yield responses and farmers' choices about irrigated area shares in response to different water deficit and aquifer conditions.
+Please see section \ref{reg-conf} of the Supplementary Information for estimated average yield response curves with 95\% confidence intervals displayed, which are omitted here for visual clarity.
+The histograms on the top of each panel show the distribution of county-level annual water deficit values included in regressions for each crop.}}
+\label{fig:tot-impact}
+\end{figure}
+
+%------------------------
+%- Figure 5
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Difference in average yield of corn and soybean for aquifer thickness levels}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_avg_yield_dif}}
+
+\caption*{\footnotesize{Note: The solid red lines represent difference in average yield of corn and soybean for aquifer thickness levels of 40 m, 70 m, and 100 m, relative to a baseline aquifer thickness of 10 m.
+Shaded areas denote 95\% confidence interval, which when greater than zero indicate average yields for the given aquifer thickness level are statistically significantly greater than for an aquifer thickness of 10 m at the 5\% level of significance.}}
+\label{fig:dif-tot-impact}
+\end{figure}
+
+%------------------------
+%- Figure 6
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Historical rainfed yield, irrigated yield, and water deficit}
+
+{\centering \includegraphics[width=750px,height=6in,]{../Results/Figures/g_y_wd_all}}
+
+\caption*{\footnotesize{Note: The blue solid, red solid, and black dotted lines represent irrigated yield, rainfed yield, and water deficit, respectively.}}
+\label{fig:deficit-yield-hist}
+\end{figure}
+
+%------------------------
+%- Figure 7
+%------------------------
+
+\begin{figure}[H]
+
+\caption{County-average aquifer thickness in 2016}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_map}}
+
+\caption*{\footnotesize{Note: The map shows county-average aquifer thickness in 2016 for each county included in our corn and soybean regression analyses.
+Counties highlighted in gray are rainfed only, and the red polygon denotes the boundary of the High Plains Aquifer system.}}
+\label{fig:sat-map}
+\end{figure}
+
+%==============================================
+%= References
+%==============================================
 \clearpage
+\bibliography{DRA.bib}
 
-\hypertarget{appendix-appendix}{%
-\appendix}
+%==============================================
+%= Supplementary Information
+%==============================================
+\clearpage
+% appendix-appendix
+
+\setcounter{figure}{0}
+\setcounter{section}{0}
+\renewcommand{\thefigure}{S.\arabic{figure}}
+\renewcommand{\thesection}{\Alph{section}}
 
@@
 This is due to laws that typically provide farmers with absolute rights to use groundwater under their land, coupled with low levels of metering.
-Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigators in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
-As discussed in section \ref{main} of our manuscript, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
+Research has shown that these conditions have historically limited or constrained potential for implementing extensive and binding pumping restrictions on irrigating farmers in these areas \citep{wheeler2016lessons, closas2018chronicle}, again indicating that regulations are not the binding driver of water use constraints in these areas.
+As discussed in section \ref{main} of our article, there is extensive evidence pointing to reduced well yields in these areas of the HPA.
 This evidence ranges from anecdotal and theoretical insights to simulation-based studies.
@@
 
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{A.\arabic{figure}}
-
 In the main article, the confidence intervals of the estimated yield responses to water deficit were not presented to enable easier comparison between response curves.
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}
-
-}
-
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-corn}
-\end{figure}
+%------------------------
+%- Figure S.1
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of corn}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_corn}}
 
-\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean in US High Plains (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:irrigated-yield-ind-soy}
+\caption*{\footnotesize{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of corn in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
+\label{fig:irrigated-yield-ind-corn}
 \end{figure}
+
+%------------------------
+%- Figure S.2
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}
+\caption{The impact of water deficit and aquifer thickness on rainfed and irrigated per-area irrigated yields of soybean}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_yield_with_conf_soy}}
 
-\caption{Average productivity of corn and soybean in the US High Plains for different levels of water deficit and aquifer thickness (solid black line) with 95\% confidence interval (grey shaded area)}\label{fig:avg-yield-ind}
+\caption*{\footnotesize{Note: The solid black lines represent the estimated rainfed and irrigated per-area irrigated yields of soybean in US High Plains at different values of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
+\label{fig:irrigated-yield-ind-soy}
+\end{figure}
+
+%------------------------
+%- Figure S.3
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Average productivity of corn and soybean}
+
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_avg_yield}}
+
+\caption*{\footnotesize{Note: The black solid lines represent average yield of corn (top panel) and soybean (bottom panel) in the US High Plains for different levels of water deficit and aquifer thickness.
+The gray shaded area represents the 95\% confidence interval.}}
+\label{fig:avg-yield-ind}
 \end{figure}
@@
 
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{B.\arabic{figure}}
-
 <<sat_ref, echo = FALSE>>=
 sat_ref <-
-  readRDS(here::here("Shared/Results/all_results.rds")) %>%
+  readRDS(here::here("Results/all_results.rds")) %>%
   .[1, sat_seq] %>%
@@
 Figure \ref{fig:test-sat-share} presents the difference in the estimated share of irrigated acres at different levels of aquifer thickness relative to the aquifer thickness level of \Sexpr{sat_ref} m, which represents a high level of aquifer thickness for our study area that exerts no constraints on water supply for irrigation.
-Grey shaded areas represented the 95\% confidence intervals for the estimated relationships.
+Gray shaded areas represented the 95\% confidence intervals for the estimated relationships.
 Where this range exceeds the red zero line, this indicates that the difference in aquifer thickness (relative to \Sexpr{sat_ref} m baseline) leads to a statistically significant (at 5\% level) reduction in irrigated production area share.
+
+%------------------------
+%- Figure S.4
+%------------------------
 
@@
 
-{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}
+\caption{Decline in the share of acres that are irrigated}
 
-}
+{\centering \includegraphics[width=6.5in,height=500px,]{../Results/Figures/g_share_dif}}
 
-\caption{Decline in the share of acres that are irrigated as a function of different levels of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
-The grey shaded areas represent the 95\% confidence interval for this estimated relationship.}\label{fig:test-sat-share}
-\end{figure}
-
-\clearpage
-
-\hypertarget{county-fe}{%
-\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{C.\arabic{figure}}
-
-Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
-This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
-As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
-However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
-For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}
-
-}
-
-\caption{Comparison of variations in aquifer thickness after demeaning by yearly average, state-year average, and county average.}\label{fig:variation-left}
-\end{figure}
-
-Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
-For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
-Importantly, adding state-year fixed effects, which control for unobservables at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
-This indicates that, even though there may be unobservables that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
-We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}
-
-}
-
-\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}\label{fig:state-fe-with-without}
+\caption*{\footnotesize{Note: The solid black lines represent the share of acres that are irrigated as a function of aquifer thickness relative to a baseline aquifer thickness level of \Sexpr{sat_ref} m (black line).
+The gray shaded areas represent the 95\% confidence interval for this estimated relationship.}}
+\label{fig:test-sat-share}
 \end{figure}
@@
 \section{Spline basis functions and GAM estimation}\label{spline-basis}}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{D.\arabic{figure}}
 
@@
 
+%------------------------
+%- Figure S.5
+%------------------------
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}
+\caption{Example of a series of thin-plate spline basis functions}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_basis}}
 
-\caption{Example of a series of thin-plate spline basis functions}\label{fig:sp-basis}
+\caption*{\footnotesize{Note: This figure presents a series of thin-plate spline basis functions evaluated at a range of $x$ values.}}
+\label{fig:sp-basis}
 \end{figure}
@@
 
+%------------------------
+%- Figure S.6
+%------------------------
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}
+\caption{Example of a function represented by different coefficient values}
 
-}
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_il_y}}
+\caption*{\footnotesize{Note: This figure presents two example functions derived by applying different coefficients on the thin-plate spline basis functions.}}
+\label{fig:ex-fx}
+\end{figure}
+\clearpage
 
-\caption{Example of a function represented by different coefficient values}\label{fig:ex-fx}
+\hypertarget{county-fe}{%
+\section{Comparison of the choice of spatial fixed effects in irrigation area share regressions}\label{county-fe}}
+
+Figure \ref{fig:variation-left} presents the histogram of aquifer thickness after demeaning these values by removing either their: (A) yearly average, (B) state-year average, or (C) county average values.
+This analysis is used to gain insights into how much variation in aquifer thickness for identification is left when implementing either: (A) only year fixed effects, (B) state-year fixed effects, or (C) county and year fixed effects.
+As shown in Figure \ref{fig:variation-left}, including state-year fixed effects does not substantially reduce the total variation in aquifer thickness compared to including only year fixed effects.
+However, including county fixed effects dramatically reduces the total variation in aquifer thickness, resulting in extremely inaccurate estimates of impacts of aquifer conditions on drought risk (high standard error).
+For this reason, we included state-year fixed effects in our irrigated production share regression analysis to strike a balance between preserving variability in aquifer thickness data and ensuring controls for other unobserved spatial factors that may influence estimates of the impacts of aquifer conditions on production shares.
+
+%------------------------
+%- Figure S.7
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of variations in aquifer thickness after demeaning}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_variation}}
+
+\caption*{\footnotesize{Note: This figure compares the variations in aquifer thickness after demeaning by yearly average (top), state-year average (middle), and county-year average (bottom).}}
+\label{fig:variation-left}
 \end{figure}
 
-\clearpage
+Figure \ref{fig:state-fe-with-without} compares the impact of aquifer thickness on the share of irrigated acres for corn and soybean with different sets of fixed effects.
+For both corn and soybean, the negative impact of a reduction in aquifer thickness on the share of irrigated acres is virtually the same when using different possible sets of fixed effects.
+Importantly, adding state-year fixed effects, which control for unobservable factors at the state-year level, does not alter the estimated impacts of aquifer thickness on share of acres irrigated.
+This indicates that, even though there may be unobservable factors that are spatially correlated with aquifer thickness beyond soil characteristics, their correlation may not cause significant bias to our estimation and findings.
+We note that more detailed assessment of the impacts of including county-fixed effects is beyond the scope of our available datasets, and we defer this matter to future studies where greater temporal variations in aquifer thickness levels may be available to allow incorporation of county-level fixed effects.
 
-  \bibliography{DRA.bib}
+%------------------------
+%- Figure S.8
+%------------------------
+
+\begin{figure}[H]
+
+\caption{Comparison of estimated impact of aquifer thickness on the share of irrigated acres with and without state fixed effects}
+
+{\centering \includegraphics[width=6in,height=500px,]{../Results/Figures/g_share_comp}}
+
+\caption*{\footnotesize{Note: The solid black lines represent the estimated share of irrigated acres for corn and soybean under different types of fixed effects.
+Gray shaded area represents 95\% confidence interval.}}
+\label{fig:state-fe-with-without}
+\end{figure}
 

### 2023-11-08 Codes/2_figures_tables.R

@@ -1,11 +1,413 @@
-## ---------------------
+## ---------------------------------------------------------------------
 all_results <- readRDS("Results/all_results.rds")
 
 data_corn <- all_results[crop == "corn", data][[1]]
 data_soy <- all_results[crop == "soy", data][[1]]
 
 
-## ---------------------
+## ---------------------------------------------------------------------
+# all_results <-  data.table(all_results)
+
+plot_yield_data <-
+  all_results[, .(crop, yield_pred_data)] %>%
+  rowwise() %>%
+  dplyr::mutate(plot_data = list(
+    yield_pred_data[, crop_type := crop] %>%
+      .[, crop_type := case_when(
+        crop_type == "corn" ~ "A. Corn",
+        crop_type == "soy" ~ "B. Sybean"
+      )]
+  )) %>%
+  dplyr::mutate(sat_cat_text_ls = list(
+    plot_data[, .(sat_rank, sat_cat_text)] %>%
+      .[!is.na(sat_rank), ] %>%
+      unique(by = "sat_rank") %>%
+      .[order(sat_rank), sat_cat_text]
+  )) %>%
+  dplyr::mutate(plot_data = list(
+    plot_data %>%
+      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
+      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))]
+  ))
+
+g_ir_yield_both <-
+  plot_yield_data %>%
+  dplyr::mutate(g_yield = list(
+    ggplot(plot_data) +
+      geom_line(aes(
+        y = y_hat,
+        x = balance,
+        color = factor(sat_cat_text)
+      )) +
+      scale_x_continuous(
+        name = "Water Deficit (mm)",
+        breaks = (-2:11) * 100
+      ) +
+      scale_y_continuous(
+        name = "Yield (tonne/ha)",
+        breaks = 0:(ceiling(max(plot_data$y_hat)) + 0.5),
+        limits = c(NA, (ceiling(max(plot_data$y_hat)) + 0.5))
+      ) +
+      scale_color_discrete(
+        name = "Aquifer Thickness",
+        guide = guide_legend(
+          title.position = "top",
+          nrow = 2
+        )
+      ) +
[…]

### 2023-11-08 Codes/A_1_state_fe.R

@@ -1,8 +1,8 @@
-## --------------------------------------------------------
+## ---------------------------------------------------------------------
 all_results <- readRDS(here::here("Results/all_results.rds"))
 
 
-## --------------------------------------------------------
+## ---------------------------------------------------------------------
 #++++++++++++++++++++++++++++++++++++
 #+ Formula
 #++++++++++++++++++++++++++++++++++++
@@ -85,7 +85,7 @@ results <-
   data.table()
 
 
-## --------------------------------------------------------
+## ---------------------------------------------------------------------
 g_share_comp <-
   results[sat <= 150, ] %>%
   .[, crop := case_when(
@@ -122,7 +122,7 @@ ggsave(
 )
 
 
-## --------------------------------------------------------
+## ---------------------------------------------------------------------
 ir_corn_share_data <- all_results[crop == "corn", share_reg_data][[1]]
 
 sy_demeaan_data <-
